---
title: "BioClim_Growth_Season"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# BioClim Variables (https://www.worldclim.org/data/bioclim.html#google_vignette)
-   annual mean temperature (BIO1)
-   mean diurnal range (BIO2) - (Mean of monthly (max temp - min temp))
-   temperature seasonality (BIO4) (standard deviation *100)
-   temperature annual range (BIO7) (Max Temperature of Warmest Month - Min Temperature of Coldest Month)
-   mean temp of wettest quarter (BIO8)
-   mean temp of driest quarter (BIO9)
-   annual precipitation (BIO12) - sum of ppt for the entire year (not the avg)
-   precipitation seasonality (BIO15)  (Coefficient of Variation)
-   precip of warmest quarter (BIO18)
-   precip of coldest quarter (BIO19)

Can't caluclate with the function in QBMS for calculating it because I only want it for the growth season, which for most populations is less than 12 months. 
https://search.r-project.org/CRAN/refmans/QBMS/html/calc_biovars.html  

A quarter is a period of three months in the OG bioclim. Can't do that here because I don't have contiguous sets of 3 months (for example some low elev pops start their growht in December). So I will modify the "quarter" variables to be "period" variables. 

-   mean temp of wettest quarter (BIO8) - CHANGED TO mean temp of wettest month 
-   mean temp of driest quarter (BIO9) - CHANGED TO mean temp of driest month 
-   precip of warmest quarter (BIO18) - CHANGED TO precip of warmest month
-   precip of coldest quarter (BIO19) - CHANGED TO precip of coldest month 

## Relevant Libraries and Functions

```{r}
library(raster)
library(tidyverse)
library(conflicted)
conflicts_prefer(dplyr::select())
conflicts_prefer(dplyr::filter)
library(ggrepel)
library(cowplot)
library(gridExtra)
library(corrplot) #plotting correlations 
library(rstatix) #performing cor_test

sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
} #legend function for grid_arrange

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")
```

## Load the Flint growth season data (from "Flint_Growth_Season.Rmd")
```{r}
flint_recent_grwseason <- read_csv("../output/Climate/flint_climate_growthseason_recent.csv") %>%
  dplyr::select(month, parent.pop:tmx)
head(flint_recent_grwseason, 12)
flint_recent_grwseason %>% filter(parent.pop=="DPR")

flint_historical_grwseason <- read_csv("../output/Climate/flint_climate_growthseason_historical.csv") %>%
  dplyr::select(month, parent.pop:tmx)
head(flint_historical_grwseason, 12)
```

## Manually Calculate BioClim Variables

Referenced this r code: https://github.com/rspatial/dismo/blob/master/R/biovars.R 

### Recent Climate

#### Prep

Calculating average temp and diurnal range for each month 
```{r}
bioclim_recent_meantemp_prep <- flint_recent_grwseason %>% 
  mutate(tavg = (tmn + tmx)/2, t_diurnal = (tmx-tmn))
```

Calculating wettest, driest, warmest, and coldest months 

```{r}
recent_wettest_month <- bioclim_recent_meantemp_prep %>%  #there's one wettest month per pop/year
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  slice_max(ppt)

recent_driest_month <- bioclim_recent_meantemp_prep %>%  #there's more than one driest month for some pops and years 
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  slice_min(ppt)

recent_warmest_month <- bioclim_recent_meantemp_prep %>% #there's more than one warmest month for 1 pop/year
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  slice_max(tavg)

recent_coldest_month <- bioclim_recent_meantemp_prep %>% #there's one coldest month per pop/year
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  slice_min(tavg)
```


#### Calculations 

Bio 1, 2, 4, 7, 12, 15
```{r}
bioclim_recent_calc <- bioclim_recent_meantemp_prep %>% 
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  summarise(ann_tmean=mean(tavg),  #Bio1 - Annual Mean Temperature
            mean_diurnal_range=mean(t_diurnal), #Bio2 - Mean Diurnal Range
            temp_seasonality=sd(tavg), #Bio4 - Temperature Seasonality
            temp_ann_range=(max(tmx))-(min(tmn)), #bio7 - temp annual range
            ann_ppt=sum(ppt), #bio12 - annual precip
            ppt_seasonality=cv(ppt+1)) #bio15 - Precipitation Seasonality (+1 to avoid strange CVs for areas where mean rainfaill is < 1)
bioclim_recent_calc
```

Bio 8(Q), 9(Q), 18(Q), 19(Q)
```{r}
#bio8 = tmean_wettest_month
bio8_recent <- recent_wettest_month %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, year, tmean_wettest_month=tavg)

#bio9 = tmean_driest_month
bio9_recent <- recent_driest_month %>% 
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  summarise(tmean_driest_month=mean(tavg)) #taking the average b/c some years where there are multiple driest months 

bio8_9_recent <- full_join(bio8_recent, bio9_recent)

#bio18 = ppt_warmest_month
bio18_recent <- recent_warmest_month %>% 
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  summarise(ppt_warmest_month=mean(ppt)) #taking the avg b/c more than one warmest month for one pop/year

#bio19 = ppt_coldest_month
bio19_recent <- recent_wettest_month %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, year, ppt_coldest_month=ppt)

bio18_19_recent <- full_join(bio18_recent, bio19_recent)

all_periods_recent <- full_join(bio8_9_recent, bio18_19_recent)
```

Merge all bioclims
```{r}
bioclim_recent <- full_join(bioclim_recent_calc, all_periods_recent)
summary(bioclim_recent)
write_csv(bioclim_recent, "../output/Climate/BioClim_growthseason_Recent.csv")
```

### Historical Climate
#### Prep

Calculating average temp and diurnal range for each month 
```{r}
bioclim_historical_meantemp_prep <- flint_historical_grwseason %>% 
  mutate(tavg = (tmn + tmx)/2, t_diurnal = (tmx-tmn))
```

Calculating wettest, driest, warmest, and coldest months 

```{r}
historical_wettest_month <- bioclim_historical_meantemp_prep %>%  
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  slice_max(ppt)

historical_driest_month <- bioclim_historical_meantemp_prep %>%  #there's more than one driest month for some pops and years 
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  slice_min(ppt)

historical_warmest_month <- bioclim_historical_meantemp_prep %>% #there's one warmest month per pop/year
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  slice_max(tavg)

historical_coldest_month <- bioclim_historical_meantemp_prep %>% #there's one coldest month per pop/year
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  slice_min(tavg)
```


#### Calculations 

Bio 1, 2, 4, 7, 12, 15
```{r}
bioclim_historical_calc <- bioclim_historical_meantemp_prep %>% 
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  summarise(ann_tmean=mean(tavg),  #Bio1 - Annual Mean Temperature
            mean_diurnal_range=mean(t_diurnal), #Bio2 - Mean Diurnal Range
            temp_seasonality=sd(tavg), #Bio4 - Temperature Seasonality
            temp_ann_range=(max(tmx))-(min(tmn)), #bio7 - temp annual range
            ann_ppt=sum(ppt), #bio12 - annual precip
            ppt_seasonality=cv(ppt+1)) #bio15 - Precipitation Seasonality (+1 to avoid strange CVs for areas where mean rainfaill is < 1)
bioclim_historical_calc
```

Bio 8(Q), 9(Q), 18(Q), 19(Q)
```{r}
#bio8 = tmean_wettest_month
bio8_historical <- historical_wettest_month %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, year, tmean_wettest_month=tavg)

#bio9 = tmean_driest_month
bio9_historical <- historical_driest_month %>% 
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  summarise(tmean_driest_month=mean(tavg)) #taking the average b/c some years where there are multiple driest months 

bio8_9_historical <- full_join(bio8_historical, bio9_historical)

#bio18 = ppt_warmest_month
bio18_historical <- historical_warmest_month %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, year, ppt_warmest_month=ppt)

#bio19 = ppt_coldest_month
bio19_historical <- historical_wettest_month %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, year, ppt_coldest_month=ppt)

bio18_19_historical <- full_join(bio18_historical, bio19_historical)

all_periods_historical <- full_join(bio8_9_historical, bio18_19_historical)
```

Merge all bioclims
```{r}
bioclim_historical <- full_join(bioclim_historical_calc, all_periods_historical)
summary(bioclim_historical)
write_csv(bioclim_historical, "../output/Climate/BioClim_growthseason_Historical.csv")
```


## Bioclim trends
```{r}
bioclim_recent %>% ggplot(aes(x=parent.pop, y=ann_tmean)) + geom_boxplot()
bioclim_historical %>% ggplot(aes(x=parent.pop, y=ann_tmean)) + geom_boxplot()

bioclim_recent %>% ggplot(aes(x=parent.pop, y=ann_ppt)) + geom_boxplot() #seemingly more variation in ppt across years than temperature 
bioclim_historical %>% ggplot(aes(x=parent.pop, y=ann_ppt)) + geom_boxplot()

bioclim_recent %>% ggplot(aes(x=year, y=temp_seasonality, group=parent.pop, color=parent.pop)) + geom_line()
bioclim_historical %>% ggplot(aes(x=year, y=temp_seasonality, group=parent.pop, color=parent.pop)) + geom_line()

bioclim_recent %>% ggplot(aes(x=year, y=ppt_seasonality, group=parent.pop, color=parent.pop)) + geom_line()
bioclim_historical %>% ggplot(aes(x=year, y=ppt_seasonality, group=parent.pop, color=parent.pop)) + geom_line()
```

## Averages

### Recent Years - BioClim
```{r}
bioclim_recent_avgs <- bioclim_recent %>% group_by(parent.pop, elevation.group, elev_m) %>%
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range", "tmean_wettest_month", "tmean_driest_month", "ann_ppt", "ppt_seasonality","ppt_warmest_month", "ppt_coldest_month"), c(mean, sem), na.rm = TRUE) 
names(bioclim_recent_avgs) <- gsub("fn2", "sem", colnames(bioclim_recent_avgs))
names(bioclim_recent_avgs) <-gsub("fn1", "avg", colnames(bioclim_recent_avgs))
bioclim_recent_avgs
#write_csv(bioclim_recent_avgs, "../output/Climate/growthseason_BioClimAvgs_Recent.csv")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_avg), y=ann_tmean_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ann_tmean_avg-ann_tmean_sem,ymax=ann_tmean_avg+ann_tmean_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Annual Mean Temp", x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_AnnMeanTmp_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_avg), y=mean_diurnal_range_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=mean_diurnal_range_avg-mean_diurnal_range_sem,ymax=mean_diurnal_range_avg+mean_diurnal_range_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Diurnal Range", x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_DiurnalRange_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_avg), y=temp_seasonality_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=temp_seasonality_avg-temp_seasonality_sem,ymax=temp_seasonality_avg+temp_seasonality_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_TempSeasonality_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_avg), y=temp_ann_range_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=temp_ann_range_avg-temp_ann_range_sem,ymax=temp_ann_range_avg+temp_ann_range_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_AnnTmpRange_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_month_avg), y=tmean_wettest_month_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=tmean_wettest_month_avg-tmean_wettest_month_sem,ymax=tmean_wettest_month_avg+tmean_wettest_month_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_TmpMeanWet_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_month_avg), y=tmean_driest_month_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=tmean_driest_month_avg-tmean_driest_month_sem,ymax=tmean_driest_month_avg+tmean_driest_month_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_TmpMeanDry_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_avg), y=ann_ppt_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ann_ppt_avg-ann_ppt_sem,ymax=ann_ppt_avg+ann_ppt_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_AnnPPT_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_avg), y=ppt_seasonality_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ppt_seasonality_avg-ppt_seasonality_sem,ymax=ppt_seasonality_avg+ppt_seasonality_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_PPTSeasonality_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_month_avg), y=ppt_warmest_month_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ppt_warmest_month_avg-ppt_warmest_month_sem,ymax=ppt_warmest_month_avg+ppt_warmest_month_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_PPTWarm_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_month_avg), y=ppt_coldest_month_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ppt_coldest_month_avg-ppt_coldest_month_sem,ymax=ppt_coldest_month_avg+ppt_coldest_month_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_PPTCold_RecentClim.png", width = 12, height = 6, units = "in")
```

### Historical Years - BioClim
```{r}
names(bioclim_historical)
bioclim_historical_avgs <- bioclim_historical %>% group_by(parent.pop, elevation.group, elev_m) %>%
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range", "tmean_wettest_month", "tmean_driest_month", "ann_ppt", "ppt_seasonality","ppt_warmest_month", "ppt_coldest_month"), c(mean, sem), na.rm = TRUE) 
names(bioclim_historical_avgs) <- gsub("fn2", "sem", colnames(bioclim_historical_avgs))
names(bioclim_historical_avgs) <-gsub("fn1", "avg", colnames(bioclim_historical_avgs))
bioclim_historical_avgs
#write_csv(bioclim_historical_avgs, "../output/Climate/growthseason_BioClimAvgs_Historical.csv")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_avg), y=ann_tmean_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ann_tmean_avg-ann_tmean_sem,ymax=ann_tmean_avg+ann_tmean_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Annual Mean Temp", x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_AnnMeanTmp_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_avg), y=mean_diurnal_range_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=mean_diurnal_range_avg-mean_diurnal_range_sem,ymax=mean_diurnal_range_avg+mean_diurnal_range_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Diurnal Range", x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_DiurnalRange_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_avg), y=temp_seasonality_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=temp_seasonality_avg-temp_seasonality_sem,ymax=temp_seasonality_avg+temp_seasonality_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_TempSeasonality_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_avg), y=temp_ann_range_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=temp_ann_range_avg-temp_ann_range_sem,ymax=temp_ann_range_avg+temp_ann_range_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_AnnTmpRange_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_month_avg), y=tmean_wettest_month_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=tmean_wettest_month_avg-tmean_wettest_month_sem,ymax=tmean_wettest_month_avg+tmean_wettest_month_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_TmpMeanWet_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_month_avg), y=tmean_driest_month_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=tmean_driest_month_avg-tmean_driest_month_sem,ymax=tmean_driest_month_avg+tmean_driest_month_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_TmpMeanDry_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_avg), y=ann_ppt_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ann_ppt_avg-ann_ppt_sem,ymax=ann_ppt_avg+ann_ppt_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_AnnPPT_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_avg), y=ppt_seasonality_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ppt_seasonality_avg-ppt_seasonality_sem,ymax=ppt_seasonality_avg+ppt_seasonality_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_PPTSeasonality_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_month_avg), y=ppt_warmest_month_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ppt_warmest_month_avg-ppt_warmest_month_sem,ymax=ppt_warmest_month_avg+ppt_warmest_month_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_PPTWarm_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_month_avg), y=ppt_coldest_month_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ppt_coldest_month_avg-ppt_coldest_month_sem,ymax=ppt_coldest_month_avg+ppt_coldest_month_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_PPTCold_HistoricalClim.png", width = 12, height = 6, units = "in")
```


## Climate trait correlations
```{r}
names(bioclim_recent)
bioclim_recent <- tibble(bioclim_recent)
bioclim_recent %>% cor_test(ann_tmean:ppt_coldest_month, method = "pearson")
recent_cor_bioclim_mat <- bioclim_recent %>% dplyr::select(ann_tmean:ppt_coldest_month) %>% cor_mat()
recent_cor_bioclim_mat

recent_cor_bioclim = bioclim_recent %>% dplyr::select(ann_tmean:ppt_coldest_month) %>% cor()
file_path= "../output/Climate/BioClim_RecentClim_Cors.png"
png(width = 12, height = 6, res= 300, units = "in", file=file_path, type = "cairo")
corrplot(recent_cor_bioclim)
dev.off()

bioclim_historical <- tibble(bioclim_historical)
bioclim_historical %>% cor_test(ann_tmean:ppt_coldest_month, method = "pearson")
historical_cor_bioclim_mat <- bioclim_historical %>% dplyr::select(ann_tmean:ppt_coldest_month) %>% cor_mat()
historical_cor_bioclim_mat

historical_cor_bioclim = bioclim_historical %>% dplyr::select(ann_tmean:ppt_coldest_month) %>% cor()
file_path= "../output/Climate/BioClim_HistoricalClim_Cors.png"
png(width = 12, height = 6, res= 300, units = "in", file=file_path, type = "cairo")
corrplot(historical_cor_bioclim)
dev.off()
```
Correlations vary slightly between recent and historical climate 