---
title: "Climatevars_all sites_all years_all timeframes"
author: "JR Gremer"
date: "2025-07-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#clear console if needed
#rm(list = ls())
```

# This code compiles and calculates all flint and bioclim variables for each site, year, and timeframe

Note: Due to changes to how bioclim is calculated for water year and how the growing season months are calculated in this script, we will have to update:

-   Climate data methods in the manuscript

-   Climate variable correlations (Figure S1)

-   Climate PCAs currently in the manuscript (Figure 1, S2; Table S2, S3)

-   Climate distance calculations (Figure S3; Table S4)

-   Fitness \~ climate distance regressions and figures (Figure 2-4; S4-S7; Table 1-2)

## Load libraries

```{r}
library(raster) #for the CV function 
library(tidyverse)

sem <- function(x, na.rm=FALSE) { #standard error function 
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 

true_round <- function(number, digits) { #rounding function that bypasses R's round to even default
  posneg <- sign(number)
  number <- abs(number) * 10^digits
  number <- number + 0.5 + sqrt(.Machine$double.eps)
  number <- trunc(number)
  number <- number / 10 ^ digits
  number * posneg
} 
```

## Load the pop and location data

```{r}
#pop info
pops_common_garden <- read_csv("../input/WL2_Data/Pops_for_2023_WL2.csv") #pops included in common garden 
pops_common_garden_nonotes <- pops_common_garden %>% select(parent.pop:phylogroup, elevation.group) #subset columns

#Need to add YO1 for JGI:
yose1 <- tibble(parent.pop="YO1", phylogroup=NA) 
pops_common_garden_jgi <- bind_rows(pops_common_garden_nonotes, yose1)

#extra location info 
pop_loc <- read_csv("../input/Strep_tort_locs.csv")

#need to change YOSE to YO
pop_loc_yo <- pop_loc %>% mutate(parent.pop = str_replace(`Site code`, "YOSE(\\d+)", "YO\\1")) %>% select(Lat, Long, elev_m=`Elevation (m)`, parent.pop)

#merge in location info
pop_elev <- left_join(pops_common_garden_jgi, pop_loc_yo)

#add in common garden sites
sites_loc <- tibble(parent.pop=c("UCD_Garden", "WL2_Garden"), phylogroup=NA, Lat=c("38.53250", "38.82599"), Long=c("-121.78299", "-120.25090"), elev_m=c(16,2020))
pops_gardens_locs <- bind_rows(pop_elev, sites_loc)

#add row names
pops_gardens_locs_rowids <- pops_gardens_locs %>% rowid_to_column() %>% mutate(rowid = as.character(rowid))
```

## Load raw flint data

```{r}
flint_data1 <- read_csv("../input/Updated_Flint_Climate_Oct2024.csv") #This only goes through Aug 2024
flint_data2 <- read_csv("../input/Updated_Flint_Sept-Dec2024.csv") #Sept-Dec 2024

#merge:
flint_data <- bind_rows(flint_data1, flint_data2)
```

## Convert raw flint into a workable format (original code in "Climate_Shapefile.Rmd")

```{r}
cwd <- flint_data %>% filter(str_detect(DataFileName, "cwd"))
cwd_long <- cwd %>% 
  pivot_longer(!DataFileName, names_to = "rowid", values_to = "cwd") %>% 
  extract(DataFileName, c("year", "month"), "(\\d+)(.*)", convert = TRUE) %>% 
  separate(month, into = c("month", "file_type")) %>% select(rowid, year, month, cwd)

pet <- flint_data %>% filter(str_detect(DataFileName, "pet"))
pet_long <- pet %>% 
  pivot_longer(!DataFileName, names_to = "rowid", values_to = "pet") %>% 
  extract(DataFileName, c("year", "month"), "(\\d+)(.*)", convert = TRUE) %>% 
  separate(month, into = c("month", "file_type")) %>% select(rowid, year, month, pet)

aet <- flint_data %>% 
  filter(str_detect(DataFileName, "aet"))
aet_long <- aet %>% 
  pivot_longer(!DataFileName, names_to = "rowid", values_to = "aet") %>% 
  extract(DataFileName, c("year", "month"), "(\\d+)(.*)", convert = TRUE) %>% 
  separate(month, into = c("month", "file_type")) %>% select(rowid, year, month, aet)

pck <- flint_data %>% filter(str_detect(DataFileName, "pck"))
pck_long <- pck %>% 
  pivot_longer(!DataFileName, names_to = "rowid", values_to = "pck") %>% 
  extract(DataFileName, c("year", "month"), "(\\d+)(.*)", convert = TRUE) %>% 
  separate(month, into = c("month", "file_type")) %>% select(rowid, year, month, pck)

ppt <- flint_data %>% filter(str_detect(DataFileName, "ppt"))
ppt_long <- ppt %>% 
  pivot_longer(!DataFileName, names_to = "rowid", values_to = "ppt") %>% 
  extract(DataFileName, c("year", "month"), "(\\d+)(.*)", convert = TRUE) %>% 
  separate(month, into = c("month", "file_type")) %>% select(rowid, year, month, ppt)

tmn <- flint_data %>% filter(str_detect(DataFileName, "tmn"))
tmn_long <- tmn %>% 
  pivot_longer(!DataFileName, names_to = "rowid", values_to = "tmn") %>% 
  extract(DataFileName, c("year", "month"), "(\\d+)(.*)", convert = TRUE) %>% 
  separate(month, into = c("month", "file_type")) %>% select(rowid, year, month, tmn)

tmx <- flint_data %>% filter(str_detect(DataFileName, "tmx"))
tmx_long <- tmx %>% 
  pivot_longer(!DataFileName, names_to = "rowid", values_to = "tmx") %>% 
  extract(DataFileName, c("year", "month"), "(\\d+)(.*)", convert = TRUE) %>% 
  separate(month, into = c("month", "file_type")) %>% select(rowid, year, month, tmx)

#merge:
flint_long <- full_join(aet_long, cwd_long)
flint_long <- full_join(flint_long, pck_long)
flint_long <- full_join(flint_long, pet_long)
flint_long <- full_join(flint_long, ppt_long)
flint_long <- full_join(flint_long, tmn_long)
flint_long <- full_join(flint_long, tmx_long)

flint_pops <- full_join(pops_gardens_locs_rowids, flint_long) 
#write_csv(flint_pops, "../output/Climate/Flint_AllSites_Dec2024.csv") 
```

Note: flint_pops includes the 2 common garden sites

### Subset to only the 23 pops plus the garden sites

```{r}
pop_elev_climate <- flint_pops %>% 
  select(parent.pop, elevation.group, elev_m, Lat:Long, year:tmx) %>% #organize the columns 
  mutate(elevation.group=if_else(parent.pop=="UCD_Garden", "Low",  #add elevation group for gardens
                                 if_else(parent.pop=="WL2_Garden", "High", elevation.group))) %>% 
  filter(parent.pop!="YO1") #remove YO1 from the dataset 
#write_csv(pop_elev_climate, "../output/Climate/flint_climate_UCDpops.csv") #this is the file used in "Flint_Growth_Season.Rmd" and "Climate_PCAs_AllYear.Rmd"
```

## Flint Water Year

```{r}
month_nums <- tibble(month=c("jan","feb","mar","apr","may","jun","jul","aug","sep","oct","nov","dec"),
                     month_nums=c(1:12)) #give calendar months a number 

#merge:
flint_all_year_mosnums <- left_join(pop_elev_climate, month_nums) %>% arrange(parent.pop, year, month_nums) 

#define the water year as November-October:
flint_all_year_wtr_yr <- flint_all_year_mosnums %>% 
  mutate(wtr_yr = year + month_nums %in% c(11:12)) %>% 
  mutate(timeframe = case_when(
                                wtr_yr < 1964 ~ "beforehistoric",
                                wtr_yr >= 1964 & wtr_yr < 1994 ~ "historic",
                                wtr_yr >= 1994 & wtr_yr < 2024 ~ "recent")) %>%
                         mutate(timeframe = as.factor(timeframe), month = as.factor(month)) %>%
                         filter(wtr_yr != 1895, wtr_yr < 2024) %>% #remove years with less than 12 months of data 
                         mutate(months_wtryr = case_when(
                           month == "nov" ~ 1,
                           month == "dec" ~ 2,
                           month == "jan" ~ 3,
                           month == "feb" ~ 4,
                           month == "mar" ~ 5,
                           month == "apr" ~ 6,
                           month == "may" ~ 7, 
                           month == "jun" ~ 8,
                           month == "jul" ~ 9, 
                           month == "aug" ~ 10,
                           month == "sep" ~ 11,
                           month == "oct" ~ 12 )) %>% 
  select(parent.pop:month, timeframe, wtr_yr, months_wtryr, aet:tmx)  #organize the columns 
summary(flint_all_year_wtr_yr)
#write_csv(flint_all_year_wtr_yr, "../output/Climate/flint_all_year_wtr_yr.csv")

#water year is calculated Nov - October.  Could have issues with this, but move forward for now. Can be changed later. 
```

## Bioclim water year calculations

Note: Changed this from using the calc_biovars function in the QBMS package to calculating by hand to match the growing season calculations. Referenced this r code: <https://github.com/rspatial/dismo/blob/master/R/biovars.R>

```{r}
# Calculating average temp and diurnal range for each month
bioclim_wtr_yr_prep <- flint_all_year_wtr_yr %>% 
  mutate(tavg = (tmn + tmx)/2, t_diurnal = (tmx-tmn))

#Bio 1, 2, 4, 7, 12, 15
bioclim_wtr_yr_calc <- bioclim_wtr_yr_prep %>% 
  group_by(parent.pop, elevation.group, elev_m, timeframe, wtr_yr) %>% 
  summarise(ann_tmean=mean(tavg),  #Bio1 - Annual Mean Temperature
            mean_diurnal_range=mean(t_diurnal), #Bio2 - Mean Diurnal Range
            temp_seasonality=sd(tavg), #Bio4 - Temperature Seasonality
            temp_ann_range=(max(tmx))-(min(tmn)), #bio7 - temp annual range
            ann_ppt=sum(ppt), #bio12 - annual precip
            ppt_seasonality=cv(ppt+1)) #bio15 - Precipitation Seasonality (+1 to avoid strange CVs for areas where mean rainfaill is < 1)

# Calculating wettest, driest, warmest, and coldest months 
wettest_month <- bioclim_wtr_yr_prep %>%  #there's one wettest month per pop/year
  group_by(parent.pop, elevation.group, elev_m, timeframe, wtr_yr) %>% 
  slice_max(ppt)
driest_month <- bioclim_wtr_yr_prep %>%  #there's more than one driest month for some pops and years 
  group_by(parent.pop, elevation.group, elev_m, timeframe, wtr_yr) %>% 
  slice_min(ppt)
warmest_month <- bioclim_wtr_yr_prep %>% #there's more than one warmest month for 1 pop/year
  group_by(parent.pop, elevation.group, elev_m, timeframe, wtr_yr) %>% 
  slice_max(tavg)
coldest_month <- bioclim_wtr_yr_prep %>% #there's more than one driest month for some pops and years 
  group_by(parent.pop, elevation.group, elev_m, timeframe, wtr_yr) %>% 
  slice_min(tavg)

# Bio 8, 9, 18, 19
bio8_wtryr <- wettest_month %>% #bio8 = tmean_wettest_month
  dplyr::select(parent.pop, elevation.group, elev_m, timeframe, wtr_yr, tmean_wettest_month=tavg)
bio9_wtryr <- driest_month %>% #bio9 = tmean_driest_month
  group_by(parent.pop, elevation.group, elev_m, timeframe, wtr_yr) %>% 
  summarise(tmean_driest_month=mean(tavg)) #taking the average b/c some years where there are multiple driest months 
bio18_wtryr <- warmest_month %>% #bio18 = ppt_warmest_month
  group_by(parent.pop, elevation.group, elev_m, timeframe, wtr_yr) %>% 
  summarise(ppt_warmest_month=mean(ppt)) #taking the avg b/c more than one warmest month for one pop/year
bio19_wtryr <- coldest_month %>% #bio19 = ppt_coldest_month
  group_by(parent.pop, elevation.group, elev_m, timeframe, wtr_yr) %>% 
  summarise(ppt_coldest_month=mean(ppt)) #taking the avg b/c more than one coldest month for one pop/year
#merge:
bio8_9_wtryr <- full_join(bio8_wtryr, bio9_wtryr) 
bio18_19_wtryr <- full_join(bio18_wtryr, bio19_wtryr) 
all_periods_wtryr <- full_join(bio8_9_wtryr, bio18_19_wtryr)

#Merge all bioclims
bioclim_wtryr <- full_join(bioclim_wtr_yr_calc, all_periods_wtryr)
summary(bioclim_wtryr)
#write_csv(bioclim_wtryr, "../output/Climate/bioclim_wtr_yr.csv")
```

Note: bioclim_wtryr includes the 2 common garden sites

## Calculate Growing Season Months

Note: Originally calculated this separately for recent and historical time periods and then averaged the grow months. Here, use 60 years summary stats for the calcs, but keep individual yearly growth seasons (don't average grow months)

-   Take the max end date and min start date for the years without one

-   Get a yearly value for the length of the growth season

Adapted from Brandie's Flint_Growth_Season.Rmd

### Prelim Calcs

```{r}
growyear_months <- tibble(month=c(1:12), growmonth=c(5, 6, 7, 8, 9, 10, 11, 12, 1, 2, 3, 4)) #start in Sept for low elev though due to after-ripening requirements 
pop_elev_climate_60 <- pop_elev_climate %>% 
  filter(parent.pop != "UCD_Garden") %>% #remove Davis garden 
  mutate(month = recode(month, #convert months to numbers 
  jan = 1,
  feb = 2,
  mar = 3,
  apr = 4,
  may = 5,
  jun = 6,
  jul = 7,
  aug = 8,
  sep = 9,
  oct = 10,
  nov = 11,
  dec = 12
)) %>% 
  left_join(growyear_months) %>% 
  mutate(grow_yr = year + month %in% c(9:12)) %>% #match grow year to grow months 
  filter(grow_yr >= 1964, grow_yr < 2024) %>%  #use the most recent 60 years
  select(parent.pop:Long, year_cal=year, grow_yr, month, growmonth, cwd, pck, ppt, tmn, tmx) %>% 
  arrange(parent.pop, year_cal, month)

pop_elev_climate_60_avgs <- pop_elev_climate_60 %>% 
  group_by(parent.pop, elevation.group, elev_m, month) %>% 
  summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) #60 year averages per month for each pop
names(pop_elev_climate_60_avgs) <- gsub("fn2", "sem", colnames(pop_elev_climate_60_avgs))
names(pop_elev_climate_60_avgs) <-gsub("fn1", "mean", colnames(pop_elev_climate_60_avgs))
pop_elev_climate_60_avgs <- pop_elev_climate_60_avgs %>% mutate(PckSum=sum(pck_mean)) #estimate of average total snowpack in a year 
```

### Populations that get less than 70 mm (b/c that's the avg height in Oct at WL2 garden) of snow pack in a year (on average)

-   First month = ppt \>= 25

    -   Remember this month is not included in the final "growth season"

    -   Jenny has used 25 mm as germinating inducing rain (esp b/c it's a sum of ppt for the whole month), 10 mm would probably be fine for sustaining growth

-   Last month = cwd \> 3rd quartile of CWD for those groups of pops

    -   Remember this month is included in the final "growth season"

```{r}
nosnow_pops_60 <- pop_elev_climate_60_avgs %>% filter(PckSum < 70)
unique(nosnow_pops_60$parent.pop) #BH, CC, TM2, SC, IH 
summary(nosnow_pops_60) #3rd quartile of CWD = 85.15

nosnow_pops_60_tojoin <- nosnow_pops_60 %>% select(parent.pop:elev_m, PckSum) %>% distinct() #pull out PckSum
nosnow_pops_60_years <- left_join(nosnow_pops_60_tojoin, pop_elev_climate_60) #add PckSum to full 60 year data frame 

#first month
nosnow_60_first_month <- nosnow_pops_60_years %>%
  group_by(parent.pop, elevation.group, elev_m, grow_yr) %>% 
  filter(ppt>=25) %>% 
  arrange(growmonth) %>% 
  filter(row_number()==1) #get first month for each pop and year with germinating inducing rain 
nosnow_60_first_month_tomerge <- nosnow_60_first_month %>% 
  select(parent.pop:elev_m, grow_yr, firstmonth=growmonth) #first month is in grow month not calendar month format
#merge:
nosnow_60_first_month_col <- full_join(nosnow_pops_60_years, nosnow_60_first_month_tomerge)

#last month
nosnow_60_last_month <- nosnow_60_first_month_col %>%
  group_by(parent.pop, elevation.group, elev_m, grow_yr) %>% 
  filter(growmonth>firstmonth) %>% #start the search at the first month of the growth season 
  filter(cwd>85.15) %>% #last month is when cwd gets too high 
  arrange(growmonth) %>% 
  filter(row_number()==1)
nosnow_60_last_month_tomerge <- nosnow_60_last_month %>% 
  select(parent.pop:elev_m, grow_yr, lastmonth=growmonth) #last month is in grow month not calendar month format
nosnow_60_last_month_col <- full_join(nosnow_60_first_month_col, nosnow_60_last_month_tomerge) %>% 
  select(parent.pop:growmonth, firstmonth, lastmonth, cwd:tmx)
```

#### Checking for weird cases

```{r}
nosnow_60_last_month_col %>% filter(is.na(lastmonth)) %>% arrange(grow_yr) #16 cases in IH with no last month 
nosnow_60_last_month_col %>% filter(parent.pop=="IH") #most of above = cases where the next Sept is the month with cwd past the threshold --> figure out how to add this month into the final climate 
##except 2005 & 2019 (for those 2 cwd didn't get above 85.15 until after the next growth season started) --> use max end date for these 
pop_elev_climate %>% filter(parent.pop=="IH") %>% filter(month=="sep", year==2023) #high cwd in sept 2023

nosnow_60_last_month_col %>% filter(is.na(firstmonth)) %>% arrange(grow_yr) #0 cases 
nosnow_60_last_month_col %>% filter(firstmonth>4) #1 case in BH (2014) where first month is after December, this is fine, right?

nosnow_60_last_month_col %>% filter(lastmonth<4) %>% arrange(parent.pop, grow_yr) #1 case in BH 1979 where last month is less than 4 (earlier than December), it's actually only 1 month post start
nosnow_60_last_month_col %>% filter(lastmonth==firstmonth) %>% arrange(parent.pop, grow_yr) #0 cases
nosnow_60_last_month_col %>% filter(lastmonth==1) %>% arrange(parent.pop, grow_yr) #0 cases where Sept is last month 
nosnow_60_last_month_col %>% filter(lastmonth<6) #1 case where Oct-Dec is the last month (same BH as above)
nosnow_60_last_month_col %>% filter(growmonth==firstmonth+1, cwd>85.15) %>% arrange(parent.pop, grow_yr) #1 case where cwd is high in the second growth month (same BH as above - 1979)
nosnow_60_last_month_col %>% filter(parent.pop=="BH", grow_yr==1979) #next growmonth with high enough CWD = 10, use that instead of 2? 
```

#### Deal with weird cases

```{r}
summary(nosnow_60_last_month_col) #first month in data set = 1, last month in data set = 12 (in grow months)
nosnow_60_last_month_col %>% filter(parent.pop=="IH") %>% summarise(maxEnd=max(lastmonth, na.rm=TRUE)) #max end = 12 in grow months 

nosnow_60first_last_month <- nosnow_60_last_month_col %>% 
  mutate(lastmonth=if_else(parent.pop=="IH" & grow_yr==2005, 12, #use the max end for the cases were CWD never > 85.15
         if_else(parent.pop=="IH" & grow_yr==2019, 12, #use the max end for the cases were CWD never > 85.15
          if_else(parent.pop=="BH" & grow_yr==1979, 10, #use the next month with high CWD
                 lastmonth))))

#deal with cases for IH where sept is the last month 
IH_sept_last_month <- nosnow_60first_last_month %>% 
  filter(is.na(lastmonth)) %>% 
  mutate(lastmonth=13) %>% #last month will be the 13th month 
  arrange(grow_yr) #14 years left for IH with no last month (last month = next year's Sept)

IH_sept_last_month_only <- IH_sept_last_month %>% 
  ungroup() %>% 
  select(parent.pop, year_cal=grow_yr, firstmonth) %>% 
  distinct() %>% #just have 1 row per year where we need to add sept
  left_join(pop_elev_climate_60) %>% 
  filter(month==9) %>% 
  mutate(grow_yr=year_cal, growmonth=13, lastmonth=13) #change grow_yr to include sept 
#still need 2023 sept:
IH_2023_sept <- pop_elev_climate %>% filter(parent.pop=="IH") %>% filter(month=="sep", year==2023) %>% 
  mutate(grow_yr=year, month=9, growmonth=13, lastmonth=13, firstmonth=1) %>% 
  select(parent.pop, year_cal=year, elevation.group:Long, grow_yr, month,growmonth,firstmonth, lastmonth, cwd, pck, ppt:tmx)
```

#### Filter to only months b/t the first and last for the full growth season

```{r}
IH_sept_last_month_fixed <- IH_sept_last_month %>% 
  bind_rows(IH_sept_last_month_only) %>% #add rows for sept in 
  bind_rows(IH_2023_sept) %>% 
  arrange(year_cal, month) %>% 
  select(-PckSum) %>% #remove this since we lost it on the new rows 
  group_by(parent.pop, elevation.group, elev_m, grow_yr) %>% 
  filter(growmonth>firstmonth) %>% #first and last month are in grow month format not calendar year 
  filter(growmonth<=lastmonth) %>% 
  arrange(parent.pop, grow_yr, growmonth)

nosnow_grwseason_60 <- nosnow_60first_last_month %>% 
  group_by(parent.pop, elevation.group, elev_m, grow_yr) %>% 
  filter(growmonth>firstmonth) %>% #first and last month are in grow month format not calendar year 
  filter(growmonth<=lastmonth) %>% 
  bind_rows(IH_sept_last_month_fixed) %>% 
  arrange(parent.pop, grow_yr, growmonth) 

xtabs(~parent.pop+month, data=nosnow_grwseason_60)
nosnow_grwseason_60 %>% ggplot(aes(x=growmonth)) + geom_histogram() +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5,6,7,8,9,10, 11, 12, 13)) +
  facet_wrap(~parent.pop)

nosnow_grwseason_60 %>% filter(parent.pop=="IH", grow_yr==1965)
```

### Populations that get more than 70 mm of snow pack in a year (on average)

-   First month = snowpack = 0 and min temp \> 0

    -   Remember this month is not included in the final "growth season"

-   Last month = snowpack. \> 70 mm OR pck \> 0 and min temp \< 0 OR min temp \< -5 (moderate freeze)

Note: “The month name of the snowpack file (that is, pckmar.asc) relates to the first day of the next month (that is, station observations on April 1st correlate with the snowpack file for March)." But this makes sense, a survey on April 1 gives a sense for the total snowpack in March.

```{r}
snow_pops_60 <- pop_elev_climate_60_avgs %>% filter(PckSum >= 70)
unique(snow_pops_60$parent.pop) #18 pops + garden site get some significant snowpack per year 

#add rest of 2023 calendar year 
snow_pop_ids <- snow_pops_60 %>% select(parent.pop) %>% distinct()
snow_pops_2023_cal_year <- pop_elev_climate %>% 
  filter(parent.pop != "UCD_Garden") %>%  
  mutate(month = recode(month, 
  jan = 1,
  feb = 2,
  mar = 3,
  apr = 4,
  may = 5,
  jun = 6,
  jul = 7,
  aug = 8,
  sep = 9,
  oct = 10,
  nov = 11,
  dec = 12
)) %>% 
  left_join(growyear_months) %>% 
  mutate(grow_yr = year + month %in% c(9:12)) %>%
  select(parent.pop:Long, year_cal=year, grow_yr, month, growmonth, cwd, pck, ppt, tmn, tmx) %>% 
  filter(grow_yr == 2024, year_cal==2023) %>%  
  right_join(snow_pop_ids) 

snow_pops_60_tojoin <- snow_pops_60 %>% select(parent.pop:elev_m, PckSum) %>% distinct()
snow_pops_60_years <- left_join(snow_pops_60_tojoin, pop_elev_climate_60) %>% 
  filter(year_cal >1963) %>%  #remove 1963 calendar year 
  bind_rows(snow_pops_2023_cal_year) %>% #add rest of 2023 calendar year 
  arrange(parent.pop, year_cal, month)

#first month
snow_60_first_month <- snow_pops_60_years %>%
  group_by(parent.pop, elevation.group, elev_m, year_cal) %>% #find snow pops growth season within the normal calendar year since they can end as late as December 
  filter(pck==0) %>% 
  filter(tmn > 0) %>% 
  arrange(month) %>% 
  filter(row_number()==1) #get first month for each pop and year with no snowpack for germ
snow_60_first_month_tomerge <- snow_60_first_month %>% 
  select(parent.pop:elev_m, year_cal, firstmonth=month)
snow_60_first_month_col <- full_join(snow_pops_60_years, snow_60_first_month_tomerge)

#last month 
snow_60_last_month <- snow_60_first_month_col %>%
  group_by(parent.pop, elevation.group, elev_m, year_cal) %>% 
  filter(month>firstmonth) %>% 
  filter(month>4) %>% #before adding this, the min last month was 10 (except for cases with Jan germ and Feb snow)
  filter(if_else(pck>70, pck>70, 
                 if_else(pck>0, tmn < 0, tmn <= -5))) %>%  #-5 is moderate freeze (Pardee et al. 2017)
  arrange(month) %>% 
  filter(row_number()==1) #get first month after growstart for each pop and year with pck >70 OR pck>0 & tmn <0 OR tmn <-5
snow_60_last_month_tomerge <- snow_60_last_month %>% 
  select(parent.pop:elev_m, year_cal, firstmonth,lastmonth=month)
snow_60_last_month_col <- full_join(snow_60_first_month_col, snow_60_last_month_tomerge) %>% 
  select(firstmonth, lastmonth, year_cal:tmx)
```

#### Checking for weird cases

```{r}
snow_60_last_month_col %>% filter(is.na(firstmonth)) #no cases where there isn't a firstmonth
snow_60_last_month_col %>% filter(lastmonth==firstmonth) %>% arrange(parent.pop, year_cal) #no cases where last month is first month 

snow_60_last_month_col %>% filter(is.na(lastmonth)) #238 cases where there isn't a lastmonth #*no snow_pack, or min temp > -5 
#***WHAT TO DO FOR THESE CASES? check the next calendar year to see when pck > 70 and make the last month 12 (Dec) for the ones without that 

snow_60_last_month_col %>% filter(lastmonth==firstmonth+1) %>% arrange(parent.pop, year_cal) #0 cases 
snow_60_last_month_col %>% filter(lastmonth==firstmonth+2) #6 cases where there was sig snowpack in the second month after the first month of 0 snowpack 
heavysnow_60 <- snow_60_last_month_col %>% filter(lastmonth<=firstmonth+2) %>% filter(month==firstmonth)
summary(heavysnow_60) 
unique(heavysnow_60$parent.pop) #3 pops included (all Lassen)
#high snowpack all year except Aug-Nov, these cases are fine to leave 

snow_60_last_month_col %>% filter(firstmonth<4) %>% filter(pck>70) %>% filter(month>firstmonth, month<10) #13 cases where Jan or Feb = first month, but high snow 1 or 2 months later 
#look at the distribution of start and end months 
snow_60_last_month_col %>% 
  ggplot(aes(x=parent.pop, y=firstmonth)) +
  geom_boxplot() +
  scale_y_continuous(breaks=c(1, 2, 3, 4, 5,6,7,8,9,10, 11, 12)) 
#Jan and Feb are clearly outliers for some higher elev pops 
#SQ1, SQ2, WL1, WL2 (MAR OUTLIER TOO?), WV, YO4
snow_60_last_month_col %>% filter(parent.pop=="SQ1", firstmonth==1) %>% 
  filter(pck>70, month>firstmonth, month<10) #no cases of high snowpack soon after germ
snow_60_last_month_col %>% filter(parent.pop=="SQ2", firstmonth==2) %>% 
  filter(pck>70, month>firstmonth, month<10) #no cases of high snowpack soon after germ
snow_60_last_month_col %>% filter(parent.pop=="WV", firstmonth==1) %>% 
  filter(pck>70, month>firstmonth, month<10) #no cases of high snowpack soon after germ
snow_60_last_month_col %>% filter(parent.pop=="YO4", firstmonth<3) %>% 
  filter(pck>70, month>firstmonth, month<10) #no cases of high snowpack soon after germ
snow_60_last_month_col %>% filter(parent.pop=="WL1", firstmonth<3) %>% 
  filter(pck>70, month>firstmonth, month<10) #3 years with high snow pack 1 or 2 months post germ (1991, 2009, 2018)
snow_60_last_month_col %>% filter(parent.pop=="WL2", firstmonth<=3) %>% 
  filter(pck>70, month>firstmonth, month<10) #1 year with high snow pack 1 month post germ (2014) 
#Check with J&J about the above WL1 and WL2 cases, for now just leave them in 
```

#### Deal with weird cases
```{r}
#search for the last month in the grow year rather than calendar year for the cases without a last month 
snow_60_last_month_grwyr <- snow_60_last_month_col %>% 
  filter(month<9) %>% 
  group_by(parent.pop, elevation.group, elev_m, grow_yr) %>% 
  filter(if_else(pck>70, pck>70, 
                 if_else(pck>0, tmn < 0, tmn <= -5))) %>%  #-5 is moderate freeze (Pardee et al. 2017)
  arrange(growmonth) %>%  
  filter(row_number()==1)
snow_60_last_month_grwyr_tomerge <- snow_60_last_month_grwyr %>% 
  ungroup() %>% 
  mutate(year_cal=grow_yr-1) %>% #will apply these last months to the previous cal year
  select(parent.pop:elev_m, year_cal,lastmonth_grwyr=growmonth) %>% 
  arrange(parent.pop, year_cal)
snow_60_last_grw_month_col <- left_join(snow_60_last_month_col, snow_60_last_month_grwyr_tomerge) %>% 
  select(firstmonth, lastmonth, lastmonth_grwyr, year_cal:tmx) 


snow_60_last_grw_month_col %>% filter(is.na(lastmonth), is.na(lastmonth_grwyr)) #still some with no last month - use the max last month for these 
snow_60_last_grw_month_col %>% group_by(parent.pop) %>% summarise(maxEnd=max(lastmonth, na.rm = TRUE)) #max end for every pop = 12 
snow_60first_last_month <- snow_60_last_grw_month_col %>% 
  mutate(lastmonth=if_else(is.na(lastmonth) & is.na(lastmonth_grwyr), 12,#assign the max end 
                           lastmonth))

snow_60_last_grw_month_col %>% filter(is.na(lastmonth), !is.na(lastmonth_grwyr)) #use the grow year last month for these
snow_no_end <- snow_60first_last_month %>% 
  filter(is.na(lastmonth)) %>% 
  select(-lastmonth)
snow_no_end_nextyr <- snow_no_end %>%
  filter(grow_yr>year_cal) %>% #want the next calendar year
  ungroup() %>% 
  select(parent.pop, year_cal=grow_yr, firstmonth, lastmonth_grwyr) %>% 
  distinct() %>% 
  left_join(pop_elev_climate_60) %>% 
  filter(growmonth>4) %>% #want only Jan + 
  filter(growmonth<=lastmonth_grwyr) %>% #only keep the rows for the months between Jan and whatever the end month is in growthmonth format 
  mutate(year_cal=year_cal-1, #change the calendar year to the previous 
         month=month+12, #change the month number appropriately 
          lastmonth=(lastmonth_grwyr-4)+12)

snow_no_end_nextyr_lastmonth <- snow_no_end_nextyr %>% 
  select(parent.pop, year_cal, lastmonth) %>% 
  distinct() %>% 
  left_join(snow_no_end) #add last month info to snow_no_end
  
#snow_no_end_nextyr
#snow_60_last_grw_month_col %>% filter(parent.pop=="CP2", year_cal>1975) #the above code worked!
```

#### Fill in all the months b/t the first and last

```{r}
snow_no_end_last_month_fixed <- snow_no_end_nextyr_lastmonth %>% 
  bind_rows(snow_no_end_nextyr) %>% #add rows for additional months in 
  arrange(year_cal, month) %>% 
  select(-Lat, -Long, -growmonth) %>% #remove this since we lost it on the new rows 
  group_by(parent.pop, elevation.group, elev_m, year_cal) %>% 
  filter(month>firstmonth) %>%  
  filter(month<=lastmonth) %>% 
  arrange(parent.pop, year_cal, month)

snow_grwseason_60 <- snow_60first_last_month %>% 
  group_by(parent.pop, elevation.group, elev_m, year_cal) %>% 
  filter(month>firstmonth) %>% #first and last month are in grow month format not calendar year 
  filter(month<=lastmonth) %>% 
  bind_rows(snow_no_end_last_month_fixed) %>% 
  arrange(parent.pop, year_cal, month) %>% 
  select(parent.pop:elev_m, firstmonth, lastmonth, year_cal:month, cwd:tmx)

xtabs(~parent.pop+month, data=snow_grwseason_60)
snow_grwseason_60 %>% ggplot(aes(x=month)) + geom_histogram() +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5,6,7,8,9,10, 11, 12,13)) +
  facet_wrap(~parent.pop)
```

## Flint for growing season months only

Bind snow and no_snow pops together

```{r, eval=FALSE}
flint_60_grwseason <- rbind(nosnow_grwseason_60, snow_grwseason_60) %>% 
  mutate(timeframe = case_when(
                                year >= 1964 & year < 1994 ~ "historic",
                                year >= 1994 & year < 2024 ~ "recent")) %>% 
  select(parent.pop:elev_m, Lat:Long, timeframe, year, month, PckSum, cwd:tmx) #organize the columns 
unique(flint_60_grwseason$parent.pop)

#write_csv(flint_60_grwseason, "../output/Climate/flint_climate_growthseason_60.csv")
```

## Bioclim growing season calculations

Can't calculate with the function in QBMS for calculating it because I only want it for the growth season, which for most populations is less than 12 months. <https://search.r-project.org/CRAN/refmans/QBMS/html/calc_biovars.html>

Adapted from Brandie's BioClim_Growth_Season.Rmd

```{r, eval=FALSE}
# Calculating average temp and diurnal range for each month
bioclim_60_meantemp_prep <- flint_60_grwseason %>% 
  mutate(tavg = (tmn + tmx)/2, t_diurnal = (tmx-tmn))

# Bio 1, 2, 4, 7, 12, 15
bioclim_60_calc <- bioclim_60_meantemp_prep %>% 
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  summarise(ann_tmean=mean(tavg),  #Bio1 - Annual Mean Temperature
            mean_diurnal_range=mean(t_diurnal), #Bio2 - Mean Diurnal Range
            temp_seasonality=sd(tavg), #Bio4 - Temperature Seasonality
            temp_ann_range=(max(tmx))-(min(tmn)), #bio7 - temp annual range
            ann_ppt=sum(ppt), #bio12 - annual precip
            ppt_seasonality=cv(ppt+1)) #bio15 - Precipitation Seasonality (+1 to avoid strange CVs for areas where mean rainfaill is < 1)

# Calculating wettest, driest, warmest, and coldest months 
wettest_month <- bioclim_60_meantemp_prep %>%  #there's one wettest month per pop/year
  group_by(parent.pop, elevation.group, elev_m, timeframe, year) %>% 
  slice_max(ppt)
driest_month <- bioclim_60_meantemp_prep %>%  #there's more than one driest month for some pops and years 
  group_by(parent.pop, elevation.group, elev_m, timeframe, year) %>% 
  slice_min(ppt)
warmest_month <- bioclim_60_meantemp_prep %>% #there's one warmest month per pop/year
  group_by(parent.pop, elevation.group, elev_m, timeframe, year) %>% 
  slice_max(tavg)
coldest_month <- bioclim_60_meantemp_prep %>% #there's one coldest month per pop/year
  group_by(parent.pop, elevation.group, elev_m, timeframe, year) %>% 
  slice_min(tavg)

# Bio 8, 9, 18, 19
bio8_60 <- wettest_month %>% #bio8 = tmean_wettest_month
  dplyr::select(parent.pop, elevation.group, elev_m, year, tmean_wettest_month=tavg)
bio9_60 <- driest_month %>% #bio9 = tmean_driest_month
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  summarise(tmean_driest_month=mean(tavg)) #taking the average b/c some years where there are multiple driest months 
bio18_60 <- warmest_month %>% #bio18 = ppt_warmest_month
  dplyr::select(parent.pop, elevation.group, elev_m, year, ppt_warmest_month=ppt)
bio19_60 <- coldest_month %>% #bio19 = ppt_coldest_month
  dplyr::select(parent.pop, elevation.group, elev_m, year, ppt_coldest_month=ppt)
# Merge:
bio8_9_60 <- full_join(bio8_60, bio9_60)
bio18_19_60 <- full_join(bio18_60, bio19_60)
all_periods_60 <- full_join(bio8_9_60, bio18_19_60)

# Merge all bioclims:
bioclim_60_grwseason <- full_join(bioclim_60_calc, all_periods_60)
#write_csv(bioclim_60_grwseason, "../output/Climate/BioClim_growthseason_60.csv")
```

## Merge water year and growing season

```{r, eval=FALSE}
#flint_all_year_wtr_yr  #flint water year climate data frame with all years and months + both garden sites
#bioclim_wtryr          #bioclim water year climate data frame with all years + both garden sites
#flint_60_grwseason     #flint growth season climate data frame with most recent 60 years and grow months + WL2 garden 
#bioclim_60_grwseason   #bioclim water year climate data frame with most recent 60 years + WL2 garden 

#summarize flint by year to merge it with the bioclim data
flint_yrly_avgs_wtr_yr <- flint_all_year_wtr_yr %>% 
  filter(timeframe!="beforehistoric", parent.pop!="UCD_Garden") %>% 
  rename(year=wtr_yr, year_cal=year) %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, timeframe, year) %>% 
  summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE)

flint_yrly_avgs_grwssn <- flint_60_grwseason %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, timeframe, year) %>% 
  summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE)

#subset bioclim water year
bioclim_wtryr_to_merge <- bioclim_wtryr %>% 
  filter(timeframe!="beforehistoric", parent.pop!="UCD_Garden") %>% 
  rename(year=wtr_yr) 

#merge
wtr_yr_clim <- left_join(flint_yrly_avgs_wtr_yr, bioclim_wtryr_to_merge) %>% mutate(Season="Water Year")
grwssn_clim <- left_join(flint_yrly_avgs_grwssn, bioclim_60_grwseason) %>% mutate(Season="Growth Season")
all_clim <- bind_rows(wtr_yr_clim, grwssn_clim)
#write_csv(all_clim, "../output/Climate/All_Clim.csv")
```
