---
title: "Climate Data Prep"
author: "Brandie Quarles"
date: "2023-10-18"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  markdown: 
    wrap: 72
---

# Climate Data Preparation 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Relevant Libraries and Functions

```{r}
library(tidyverse)
library(ggpubr) ##allows for background images
library(ggrepel)
library(cowplot)
library(gridExtra)
library(naniar) #replaces values with NA
library(jpeg)
#making topo maps:
library(elevatr)
library(terra)
library(sf)
library(giscoR)
library(marmap)

sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
} #legend function for grid_arrange

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")
```

## Load the pop and location data

```{r}
#pop info
pops_common_garden <- read_csv("../input/UCD_Data/Pops_for_2022_UCD.csv") #pops included in common garden 
summary(pops_common_garden)
names(pops_common_garden)
pops_common_garden_nonotes <- pops_common_garden %>% select(parent.pop:elevation.group, UCD.seed.year)
pops_common_garden_nonotes
pops_common_garden_nonotes$elevation.group <- str_to_title(pops_common_garden$elevation.group)

#extra location info 
pop_loc <- read_csv("../input/Strep_tort_locs.csv")
names(pop_loc)
head(pop_loc)

unique(pop_loc$`Site code`)
unique(pops_common_garden_nonotes$parent.pop)
#need to change YOSE to YO
pop_loc_yo <- pop_loc %>% mutate(parent.pop = str_replace(`Site code`, "YOSE(\\d+)", "YO\\1")) %>% select(Lat, Long, elev_m=`Elevation (m)`, parent.pop)
unique(pop_loc_yo$parent.pop)

#merge in location info
pop_elev <- left_join(pops_common_garden_nonotes, pop_loc_yo)
head(pop_elev)
```

## Map populations

```{r}
names(pop_elev)
sapply(pop_elev, class) #lat and long are characters, need to be numeric 
pop_elev <- pop_elev %>% mutate_at(c("Lat", "Long"), as.double)
states <- map_data("state") %>% filter(region == "california")


#without labels 
pops_locs_no_labels <- ggplot() +
  geom_polygon(data = states, aes(x = long, y = lat, group = group), fill = "gray") +
  coord_quickmap(xlim = c(-125, -114), ylim = c(35, 43))+
  geom_point(data = pop_elev,
             aes(x = Long, y = Lat, color=elev_m),
             size = 4) +
  labs(x="Long", y="Lat", color="Elevation (m)") +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic() +
  theme(text=element_text(size=25))
#ggsave("../output/Pop_Map_NoLabels.png", width = 6, height = 4, units = "in")

#with labels 
ggplot() +
  geom_polygon(data = states, aes(x = long, y = lat, group = group), fill = "gray") +
  coord_quickmap(xlim = c(-125, -114), ylim = c(35, 43))+
  geom_point(data = pop_elev,
             aes(x = Long, y = Lat, color=elev_m),
             size = 4) +
  geom_text_repel(data = pop_elev,
         aes(x = Long, y = Lat,
             label = `parent.pop`),
         min.segment.length = 0,
         max.overlaps = 100,
        # label.padding = 1,
        # point.padding = 0.5,
         size = 4) +
  labs(x="Long", y="Lat", color="Elevation (m)") +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic() +
  theme(text=element_text(size=25))
#ggsave("../output/Pop_Map_Labels.png", width = 12, height = 8, units = "in")
```

### Map with elevation
Overlay topographical image on graph
```{r}
imgfile <- "../input/CA_TOPO_9.jpg"
img <- readJPEG(imgfile)

pops_locs_no_labels_topo <- pops_locs_no_labels +
  background_image(img)

ggplot() +
  background_image(img) +
  geom_polygon(data = states, aes(x = long, y = lat, group = group), fill = "gray")
#ggsave("../output/California.png", width = 6.69, height=7.5, units="in")

pops_locs_no_labels_topo <- ggplot() +
  #geom_polygon(data = states, aes(x = long, y = lat, group = group), fill = "gray") +
  background_image(img) + #not perfectly aligned with the polygon 
  coord_quickmap(xlim = c(-125, -114), ylim = c(35, 43))+
  geom_point(data = pop_elev,
             aes(x = Long, y = Lat, color=elev_m),
             size = 4) +
  labs(x="Long", y="Lat", color="Elevation (m)") +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic() +
  theme(text=element_text(size=25))
ggsave("../output/Pop_Map_Topo.png", width = 6, height=4, units="in")
```


Get elevation data 
```{r}
cali <- states %>% rename(lon = long)
summary(cali)

##Get Country data 
crsLONGLAT <- "+proj=longlat +datum=WGS84 +no_defs"
get_sf <- function(country_sf, country_transformed) {
	
	country_sf <- giscoR::gisco_get_countries(
    	year = "2020",
    	epsg = "4326",
    	resolution = "10",
    	country = "USA") %>% 
	  sf::st_crop(c(xmin = -124.4, xmax = -114.1, ymin = 32.5, ymax = 42.02))
	
	country_transformed <- st_transform(country_sf, crs = crsLONGLAT)

	return(country_transformed)
}

country_transformed <- get_sf() 

##Get elevation data 
get_elevation_data <- function(country_elevation, country_elevation_df) {

	country_elevation <- get_elev_raster(
		locations = country_transformed, 
		z = 14, # Please decrease the z value if you experience R crashing
		#values ranging from 1 (high-level detail) to 14 (low-level detail).
		clip = "locations") 

	country_elevation_df <- as.data.frame(country_elevation, xy = T) %>%
		na.omit()
	
	colnames(country_elevation_df)[3] <- "elevation"

	return(country_elevation_df)
}

country_elevation_df <- get_elevation_data()
head(country_elevation_df)
```

```{r}
get_elevation_map <- function(country_map) {

	country_map <- ggplot() +
  		geom_tile(data = country_elevation_df, 
  			aes(x = x, y = y, fill = elevation)) +
  		scale_fill_etopo() +
  		coord_sf(crs = crsLONGLAT)+
  		theme_minimal() +
  		theme(text = element_text(family = "georg", color = "#22211d"),
    		axis.line = element_blank(),
    		axis.text.x = element_blank(),
    		axis.text.y = element_blank(),
    		axis.ticks = element_blank(),
    		axis.title.x = element_blank(),
    		axis.title.y = element_blank(),
    		legend.position = "none",
   		  	panel.grid.major = element_line(color = "white", linewidth = 0.2),
    		panel.grid.minor = element_blank(),
    		plot.margin = unit(c(t=0, r=0, b=0, l=0),"lines"), #added these narrower margins to enlarge maps
    		plot.background = element_rect(fill = "white", color = NA), 
    		panel.background = element_rect(fill = "white", color = NA),
    		panel.border = element_blank()) 

	return(country_map)
}

country_map <- get_elevation_map()
```


## Elevation group changes

```{r, eval=FALSE}
names(pop_elev)
unique(pop_elev$parent.pop)
xtabs(~parent.pop+elevation.group, data = pop_elev)

pop_elev$elevation.group <- factor(pop_elev$elevation.group, levels=elev_order)
pop_elev %>% ggplot(aes(x=fct_reorder(parent.pop, elev_m), y=elev_m, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(title="New Elevation Classifications", x="Parent Populations", y= "Elevation (m)") + 
  theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Elevation_Class_New.png", width = 12, height = 6, units = "in")
```

# Flint Climate Data (1895-2022)

-   aet (mm) = actual evapotranspiration (water transpired by plants, if
    water not limited)
-   cwd (mm) = climate water deficit (estimate of drought) = pet-aet
-   pck (mm) = snowpack (prior month pck + snow - lost snow)
-   pet (mm) = potential evapotranspiration (total amount of water that
    can evaporate from the ground or be transpired by plants)
-   ppt (mm) = precipitation
-   rch (mm) = recharge (water that penetrates below the root zone)
-   run (mm) = runoff (water that becomes streamflow)
-   str (mm) = soil water storage (avg amount of water stored in the
    soil annually)
-   all above are totals or sums per month
-   tmn (deg C) = min air temp (for a given month)
-   tmx (deg C) = max air temp (for a given month) 127 years

Notes

-   Did some lit review to see what variables other people have focused
    on or found to be significant (CWD, temp, precip, AET)

-   Temperature likely highly correlated within pops w/in years

    -   Take all variables (including avg temp) for the 3 months when we
        think germination is occurring (Low: Oct-Jan, Mid: Nov-Jan,
        High: April-June)?

    -   Decided to use growth season rather than germ season: Low:
        Oct-July, Mid: Nov-Aug, High: April-Sept

## Load the climate data

```{r}
climate_data <- read_csv("../input/Dimensions_All_1895-2022.csv")
head(climate_data)
names(climate_data)
unique(climate_data$pop)
climate_data$year = as.character(climate_data$year)
head(climate_data)

climate_data_yo <- climate_data %>% mutate(pop = str_replace(pop, "YOSE(\\d+)", "YO\\1")) #changing YOSE to YO 
#`(\\d+)`: This part contains a regular expression within parentheses: 
#`\\d`: Matches a digit (equivalent to `[0-9]`). 
#The double backslash `\\` is used to escape the `\` #character in R strings, so `\\d` is interpreted as `\d` in the regex, denoting a digit.
#`+`: This qualifier means "one or more of the preceding element," so `\\d+` will match one or more digits.
#`()`: Parentheses are used to define a capturing group, which means that the part of the regex within the parentheses `(\\d+)` is captured to be used in back-references or for extracting matched parts.
#\\1 references the first capturing group
unique(climate_data_yo$pop)
```

## Combine pop info with climate data
```{r}
names(pop_elev)
names(climate_data_yo)
pop_elev_climate <- left_join(pop_elev, climate_data_yo, by=c("parent.pop"="pop")) %>% select(parent.pop, elevation.group, elev_m, year:tmx)
#write_csv(pop_elev_climate, "../output/Climate/flint_climate_UCDpops.csv") #this is the file used in "Flint_Growth_Season.Rmd"
unique(pop_elev_climate$parent.pop)
#unique(pop_elev$parent.pop)
head(pop_elev_climate, 30)
names(pop_elev_climate)
```

## Calculation of recent (last 30 years) and historical climate (prior 30 years)

Took out 2022 b/c only 9 months for that year Flint variables

```{r}
pop_elev_climate_recent <- pop_elev_climate %>% filter(year>1991 & year<2022) %>% select(parent.pop:month, cwd, pck, ppt, tmn, tmx)
head(pop_elev_climate_recent)
tail(pop_elev_climate_recent)
summary(pop_elev_climate_recent)
xtabs(~parent.pop+month, data=pop_elev_climate_recent)

pop_elev_climate_historical <- pop_elev_climate %>% filter(year<=1991 & year>1961) %>% select(parent.pop:month, cwd, pck, ppt, tmn, tmx)
head(pop_elev_climate_historical, 13)
xtabs(~parent.pop+month, data=pop_elev_climate_historical)
```

## Snow Cover - Average across years (all months included)

```{r}
names(pop_elev_climate_recent)
pop_elev_climate_recent %>% filter(year==2006) %>% ggplot(aes(x=month, y=pck, group=parent.pop, color=parent.pop)) + geom_point() + geom_line()
pop_elev_climate_recent %>% filter(year==2006) %>% ggplot(aes(x=month, y=pck, group=parent.pop, color=elevation.group)) + geom_point() + geom_line()

#average snowpack across entire year for 30-year period 
recent_climate_avgs <- pop_elev_climate_recent %>% group_by(parent.pop,elevation.group ,elev_m) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(recent_climate_avgs) <- gsub("fn2", "sem", colnames(recent_climate_avgs))
names(recent_climate_avgs) <-gsub("fn1", "mean", colnames(recent_climate_avgs))
recent_climate_avgs #30 year averages

recent_climate_avgs$elevation.group <- factor(recent_climate_avgs$elevation.group, levels=elev_order)
recent_snwpck <- recent_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, pck_mean), y=pck_mean, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=pck_mean-pck_sem,ymax=pck_mean+pck_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  #scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(x="Population", y="Avg SnwPck (mm)" ,title = "Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
recent_snwpck
#ggsave("../output/Climate/All_Year_Avg_PCK_RecentClim.png", width = 12, height = 6, units = "in")

historical_climate_avgs <- pop_elev_climate_historical %>% group_by(parent.pop, elevation.group ,elev_m) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(historical_climate_avgs) <- gsub("fn2", "sem", colnames(historical_climate_avgs))
names(historical_climate_avgs) <-gsub("fn1", "mean", colnames(historical_climate_avgs))
historical_climate_avgs #30 year averages

historical_climate_avgs$elevation.group <- factor(historical_climate_avgs$elevation.group, levels=elev_order)
hist_snwpck <- historical_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, pck_mean), y=pck_mean, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=pck_mean-pck_sem,ymax=pck_mean+pck_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  #scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(x="Population", y="Avg SnwPck (mm)" ,title = "Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
hist_snwpck
#ggsave("../output/Climate/All_Year_Avg_PCK_HistoricalClim.png", width = 8, height = 4, units = "in")

legend <- get_legend(hist_snwpck)
hist_snwpck <- hist_snwpck + theme(legend.position="none")
recent_snwpck <- recent_snwpck + theme(legend.position="none")
grid.arrange(hist_snwpck, recent_snwpck, legend, ncol=3, widths=c(3.12, 3.12, 1.09))
```

## Consistency of winter snow cover

```{r}
#monthly_pck <- pop_elev_climate_recent %>% filter(parent.pop==c("DPR","WR","WL1", "SQ1", "SQ2", "WL2", "YO4")) %>% group_by(parent.pop, elev_m, month) %>%
 # summarise(pck_mean=mean(pck), pck_sem=sem(pck))
#monthly_pck$parent.pop <- factor(monthly_pck$parent.pop, levels=c("DPR","WR","WL1", "SQ1", "SQ2", "WL2", "YO4"))

monthly_pck <- pop_elev_climate_recent %>% 
  filter(parent.pop==c("WL1", "SQ1", "SQ2", "WL2", "YO4")) %>% 
  group_by(parent.pop, elev_m, month) %>%
  summarise(pck_mean=mean(pck), pck_sem=sem(pck))
monthly_pck$parent.pop <- factor(monthly_pck$parent.pop, levels=c("WL1", "SQ1", "SQ2", "WL2", "YO4"))
monthly_pck

monthly_pck %>% ggplot(aes(x=month, y=pck_mean, group=parent.pop, fill=parent.pop)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=pck_mean-pck_sem,ymax=pck_mean+pck_sem),width=.2, position=position_dodge(0.75)) +
  labs(title="Recent Climate", y="Avg SnwPck (mm)")  + 
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5,6,7,8,9,10, 11, 12)) +
  theme_classic() + 
  theme(text=element_text(size=30))
#ggsave("../output/Climate/Monthly_Avg_PCK_RecnetClim.png", width = 14, height = 6, units = "in")

#check even higher elevation
monthly_pck_high_elev <- pop_elev_climate_recent %>% 
  filter(elev_m>2200) %>%
  group_by(parent.pop, elev_m, month) %>%
  summarise(pck_mean=mean(pck), pck_sem=sem(pck))
monthly_pck_high_elev$parent.pop <- factor(monthly_pck_high_elev$parent.pop, levels=c("CP2", "CP3", "LV3", "SQ3", "YO7","YO8", "LV1", "LVTR1", "YO11"))

monthly_pck_high_elev %>% ggplot(aes(x=month, y=pck_mean, group=parent.pop, fill=parent.pop)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=pck_mean-pck_sem,ymax=pck_mean+pck_sem),width=.2, position=position_dodge(0.75)) +
  labs(title="Recent Climate", y="Avg SnwPck (mm)")  + 
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5,6,7,8,9,10, 11, 12)) +
  theme_classic() + 
  theme(text=element_text(size=30))
#ggsave("../output/Climate/Monthly_Avg_PCK_HIGHELEV_RecnetClim.png", width = 14, height = 6, units = "in")
```

## CWD Across the Year

```{r}
names(pop_elev_climate_recent)

monthly_cwd <- pop_elev_climate_recent %>% 
  group_by(parent.pop, elev_m, month) %>%
  summarise(cwd_mean=mean(cwd), pck_sem=sem(cwd))
monthly_cwd

monthly_cwd %>% #low elev
  filter(elev_m<520) %>% 
  ggplot(aes(x=month, y=cwd_mean)) +
  geom_line() +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5,6,7,8,9,10, 11, 12)) +
  theme_classic() +
  facet_wrap(~parent.pop)

monthly_cwd %>% #mid elev
  filter(elev_m>520, elev_m<1940) %>% 
  ggplot(aes(x=month, y=cwd_mean)) +
  geom_line() +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5,6,7,8,9,10, 11, 12)) +
  theme_classic() +
  facet_wrap(~parent.pop)

monthly_cwd %>% #high elev
  filter(elev_m>1940) %>% 
  ggplot(aes(x=month, y=cwd_mean)) +
  geom_line() +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5,6,7,8,9,10, 11, 12)) +
  theme_classic() +
  facet_wrap(~parent.pop)
```

### Averages

#### Across last 30 years (all months included)

```{r}
pop_elev_climate_avgs <- pop_elev_climate %>% filter(year>1992) %>% group_by(parent.pop, elevation.group) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), mean, na.rm = TRUE)
pop_elev_climate_avgs #30 year averages of all climate variables 
pop_elev_climate_avgs$elevation.group <- factor(pop_elev_climate_avgs$elevation.group, levels=elev_order)
```

