---
title: "Climate Data Prep"
author: "Brandie Quarles"
date: "2023-10-18"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

To Do:

-   Double check elevation classifications

-   Create PCAs for recent and historical climate

    -   Color monthly PCA by month relative to growth season?

-   Edit consistency of snow cover code

-   Calculate historical climate avgs

Growth Seasons:

-   Low: Oct-July

-   Mid: Nov-Aug

-   High: April-Sept

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Relevant Libraries and Functions

```{r}
library(tidyverse)
library(ggrepel)
library(cowplot)
library(gridExtra)
library(corrplot) #plotting correlations 
library(rstatix) #performing cor_test
library(lmerTest)
library(naniar) #replaces values with NA
sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
} #legend function for grid_arrange

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")

#For scree plots 
#library("devtools") #The package devtools is required for the installation as factoextra is hosted on github.
#1install_github("kassambara/factoextra")
library("factoextra")
```

## Load the pop and location data

```{r}
#pop info
pops_common_garden <- read_csv("../input/UCD_Data/Pops_for_2022_UCD.csv") #pops included in common garden 
summary(pops_common_garden)
names(pops_common_garden)
pops_common_garden_nonotes <- pops_common_garden %>% select(parent.pop:elevation.group, UCD.seed.year)
pops_common_garden_nonotes
pops_common_garden_nonotes$elevation.group <- str_to_title(pops_common_garden$elevation.group)

#extra location info 
pop_loc <- read_csv("../input/Strep_tort_locs.csv")
names(pop_loc)
head(pop_loc)

unique(pop_loc$`Site code`)
unique(pops_common_garden_nonotes$parent.pop)
#need to change YOSE to YO
pop_loc_yo <- pop_loc %>% mutate(parent.pop = str_replace(`Site code`, "YOSE(\\d+)", "YO\\1")) %>% select(Lat, Long, elev_m=`Elevation (m)`, parent.pop)
unique(pop_loc_yo$parent.pop)

#merge in location info
pop_elev <- left_join(pops_common_garden_nonotes, pop_loc_yo)
head(pop_elev)
```

## Map populations

```{r}
names(pop_elev)
sapply(pop_elev, class) #lat and long are characters, need to be numeric 
pop_elev <- pop_elev %>% mutate_at(c("Lat", "Long"), as.double)
states <- map_data("state") %>% filter(region == "california")


#without labels 
ggplot() +
  geom_polygon(data = states, aes(x = long, y = lat, group = group), fill = "gray") +
  coord_quickmap(xlim = c(-125, -114), ylim = c(35, 43))+
  geom_point(data = pop_elev,
             aes(x = Long, y = Lat, color=elev_m),
             size = 4) +
  labs(x="Long", y="Lat", color="Elevation (m)") +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic() +
  theme(text=element_text(size=25))
ggsave("../output/Pop_Map_NoLabels.png", width = 6, height = 4, units = "in")

#with labels 
ggplot() +
  geom_polygon(data = states, aes(x = long, y = lat, group = group), fill = "gray") +
  coord_quickmap(xlim = c(-125, -114), ylim = c(35, 43))+
  geom_point(data = pop_elev,
             aes(x = Long, y = Lat, color=elev_m),
             size = 4) +
  geom_text_repel(data = pop_elev,
         aes(x = Long, y = Lat,
             label = `parent.pop`),
         min.segment.length = 0,
         max.overlaps = 100,
        # label.padding = 1,
        # point.padding = 0.5,
         size = 4) +
  labs(x="Long", y="Lat", color="Elevation (m)") +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic() +
  theme(text=element_text(size=25))
ggsave("../output/Pop_Map_Labels.png", width = 6, height = 4, units = "in")
```

## Elevation group changes

```{r}
names(pop_elev)
unique(pop_elev$parent.pop)
xtabs(~parent.pop+elevation.group, data = pop_elev)

pop_elev$elevation.group <- factor(pop_elev$elevation.group, levels=elev_order)
pop_elev %>% ggplot(aes(x=fct_reorder(parent.pop, elev_m), y=elev_m, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(title="New Elevation Classifications", x="Parent Populations", y= "Elevation (m)") + 
  theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Elevation_Class_New.png", width = 12, height = 6, units = "in")
```

# Climate Data (1895-2022)

-   aet (mm) = actual evapotranspiration (water transpired by plants, if
    water not limited)
-   cwd (mm) = climate water deficit (estimate of drought) = pet-aet
-   pck (mm) = snowpack (prior month pck + snow - lost snow)
-   pet (mm) = potential evapotranspiration (total amount of water that
    can evaporate from the ground or be transpired by plants)
-   ppt (mm) = precipitation
-   rch (mm) = recharge (water that penetrates below the root zone)
-   run (mm) = runoff (water that becomes streamflow)
-   str (mm) = soil water storage (avg amount of water stored in the
    soil annually)
-   all above are totals or sums per month
-   tmn (deg C) = min air temp (for a given month)
-   tmx (deg C) = max air temp (for a given month) 127 years

Notes

-   Did some lit review to see what variables other people have focused
    on or found to be significant (CWD, temp, precip, AET)

-   Temperature likely highly correlated within pops w/in years

    -   Take all variables (including avg temp) for the 3 months when we
        think germination is occurring (Low: Oct-Jan, Mid: Nov-Jan,
        High: April-June)?

    -   Decided to use growth season rather than germ season: Low:
        Oct-July, Mid: Nov-Aug, High: April-Sept

## Load the climate data

```{r}
climate_data <- read_csv("../input/Dimensions_All_1895-2022.csv")
head(climate_data)
names(climate_data)
unique(climate_data$pop)
climate_data$year = as.character(climate_data$year)
head(climate_data)

climate_data_yo <- climate_data %>% mutate(pop = str_replace(pop, "YOSE(\\d+)", "YO\\1")) #changing YOSE to YO 
#`(\\d+)`: This part contains a regular expression within parentheses: 
#`\\d`: Matches a digit (equivalent to `[0-9]`). 
#The double backslash `\\` is used to escape the `\` #character in R strings, so `\\d` is interpreted as `\d` in the regex, denoting a digit.
#`+`: This qualifier means "one or more of the preceding element," so `\\d+` will match one or more digits.
#`()`: Parentheses are used to define a capturing group, which means that the part of the regex within the parentheses `(\\d+)` is captured to be used in back-references or for extracting matched parts.
#\\1 references the first capturing group
unique(climate_data_yo$pop)

#combine pop info with climate data
names(pop_elev)
names(climate_data_yo)
pop_elev_climate <- left_join(pop_elev, climate_data_yo, by=c("parent.pop"="pop")) %>% select(parent.pop, elevation.group, elev_m, year:tmx)
unique(pop_elev_climate$parent.pop)
#unique(pop_elev$parent.pop)
head(pop_elev_climate, 30)
names(pop_elev_climate)
```

## Calculation of recent (last 30 years) and historical climate (prior 30 years)

```{r}
pop_elev_climate_recent <- pop_elev_climate %>% filter(year>1992) %>% select(parent.pop:month, cwd, pck, ppt, tmn, tmx)
head(pop_elev_climate_recent)
summary(pop_elev_climate_recent)

pop_elev_climate_historical <- pop_elev_climate %>% filter(year<=1992 & year>1962) %>% select(parent.pop:month, cwd, pck, ppt, tmn, tmx)
head(pop_elev_climate_historical, 13)
```

## Snow Cover - Average across years (all months included)

```{r}
names(pop_elev_climate_recent)
pop_elev_climate_recent %>% filter(year==2006) %>% ggplot(aes(x=month, y=pck, group=parent.pop, color=parent.pop)) + geom_point() + geom_line()
pop_elev_climate_recent %>% filter(year==2006) %>% ggplot(aes(x=month, y=pck, group=parent.pop, color=elevation.group)) + geom_point() + geom_line()

#average snowpack across entire year for 30-year period 
recent_climate_avgs <- pop_elev_climate_recent %>% group_by(parent.pop,elevation.group ,elev_m) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(recent_climate_avgs) <- gsub("fn2", "sem", colnames(recent_climate_avgs))
names(recent_climate_avgs) <-gsub("fn1", "mean", colnames(recent_climate_avgs))
recent_climate_avgs #30 year averages

recent_climate_avgs$elevation.group <- factor(recent_climate_avgs$elevation.group, levels=elev_order)
recent_snwpck <- recent_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, pck_mean), y=pck_mean, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=pck_mean-pck_sem,ymax=pck_mean+pck_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  #scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(x="Population", y="Avg SnwPck (mm)" ,title = "Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
recent_snwpck
ggsave("../output/All_Year_Avg_PCK_RecentClim.png", width = 12, height = 6, units = "in")

historical_climate_avgs <- pop_elev_climate_historical %>% group_by(parent.pop, elevation.group ,elev_m) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(historical_climate_avgs) <- gsub("fn2", "sem", colnames(historical_climate_avgs))
names(historical_climate_avgs) <-gsub("fn1", "mean", colnames(historical_climate_avgs))
historical_climate_avgs #30 year averages

historical_climate_avgs$elevation.group <- factor(historical_climate_avgs$elevation.group, levels=elev_order)
hist_snwpck <- historical_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, pck_mean), y=pck_mean, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=pck_mean-pck_sem,ymax=pck_mean+pck_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  #scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(x="Population", y="Avg SnwPck (mm)" ,title = "Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
hist_snwpck
#ggsave("../output/All_Year_Avg_PCK_HistoricalClim.png", width = 8, height = 4, units = "in")

legend <- get_legend(hist_snwpck)
hist_snwpck <- hist_snwpck + theme(legend.position="none")
recent_snwpck <- recent_snwpck + theme(legend.position="none")
grid.arrange(hist_snwpck, recent_snwpck, legend, ncol=3, widths=c(3.12, 3.12, 1.09))
```

## Consistency of winter snow cover - come back to this !

```{r}
#monthly_pck <- pop_elev_climate_recent %>% filter(parent.pop==c("DPR","WR","WL1", "SQ1", "SQ2", "WL2", "YO4")) %>% group_by(parent.pop, elev_m, month) %>%
  summarise(pck_mean=mean(pck), pck_sem=sem(pck))
#monthly_pck$parent.pop <- factor(monthly_pck$parent.pop, levels=c("DPR","WR","WL1", "SQ1", "SQ2", "WL2", "YO4"))

monthly_pck <- pop_elev_climate_recent %>% filter(parent.pop==c("WL1", "SQ1", "SQ2", "WL2", "YO4")) %>% group_by(parent.pop, elev_m, month) %>%
  summarise(pck_mean=mean(pck), pck_sem=sem(pck))
monthly_pck$parent.pop <- factor(monthly_pck$parent.pop, levels=c("WL1", "SQ1", "SQ2", "WL2", "YO4"))

monthly_pck %>% ggplot(aes(x=month, y=pck_mean, group=parent.pop, fill=parent.pop)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=pck_mean-pck_sem,ymax=pck_mean+pck_sem),width=.2, position=position_dodge(0.75)) +
  labs(title="Recent Climate", y="Avg SnwPck (mm)")  + 
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5,6,7,8,9,10, 11, 12)) +
  theme_classic() + 
  theme(text=element_text(size=30))
ggsave("../output/Monthly_Avg_PCK_RecnetClim.png", width = 14, height = 6, units = "in")

#NEED TO EDIT THE BELOW
monthly_pck <- pop_elev_climate_recent %>% filter(parent.pop==c("CP2", "CP3", "LV3")) %>% group_by(parent.pop, elev_m, month) %>%
  summarise(pck_mean=mean(pck), pck_sem=sem(pck))
monthly_pck$parent.pop <- factor(monthly_pck$parent.pop, levels=c("CP2", "CP3", "LV3", "SQ3", "YO7","YO8", "LV1", "LVTR1", "YO11"))

monthly_pck %>% ggplot(aes(x=month, y=pck_mean, group=parent.pop, fill=parent.pop)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=pck_mean-pck_sem,ymax=pck_mean+pck_sem),width=.2, position=position_dodge(0.75)) +
  labs(title="Recent Climate", y="Avg SnwPck (mm)")  + 
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5,6,7,8,9,10, 11, 12)) +
  theme_classic() + 
  theme(text=element_text(size=30))
ggsave("../output/Monthly_Avg_PCK_HIGHELEV_RecnetClim.png", width = 14, height = 6, units = "in")
```

## Climate during growth season (Low: Oct-July, Mid: Nov-Aug, High: April-Sept)

```{r}
recent_climate_low <- pop_elev_climate_recent %>% filter(elevation.group=="Low") %>% filter(month<=7 | month>9)
head(recent_climate_low, 30)
tail(recent_climate_low, 30)
summary(recent_climate_low)

recent_climate_mid <- pop_elev_climate_recent %>% filter(elevation.group=="Mid") %>% filter(month<=8 | month>10)
head(recent_climate_mid, 30)
tail(recent_climate_mid, 30)
summary(recent_climate_mid)

recent_climate_high <- pop_elev_climate_recent %>% filter(elevation.group=="High") %>% filter(month>3 & month<=9)
head(recent_climate_high, 30)
tail(recent_climate_high, 30)
summary(recent_climate_high)

recent_climate_growthseason <- rbind(recent_climate_low, recent_climate_mid, recent_climate_high)
summary(recent_climate_growthseason)

historical_climate_low <- pop_elev_climate_historical %>% filter(elevation.group=="Low") %>% filter(month<=7 | month>9)
head(historical_climate_low, 30)
tail(historical_climate_low, 30)
summary(historical_climate_low)
unique(historical_climate_low$parent.pop) #low elevation pops = BH, CC, IH, SC, TM2

historical_climate_mid <- pop_elev_climate_historical %>% filter(elevation.group=="Mid") %>% filter(month<=8 | month>10)
head(historical_climate_mid, 30)
tail(historical_climate_mid, 30)
summary(historical_climate_mid)
unique(historical_climate_mid$parent.pop) #mid elevation pops = DPR, FR, SQ1, SQ2, WL1, WR, WV

historical_climate_high <- pop_elev_climate_historical %>% filter(elevation.group=="High") %>% filter(month>3 & month<=9)
head(historical_climate_high, 30)
tail(historical_climate_high, 30)
summary(historical_climate_high) 
unique(historical_climate_high$parent.pop) #high elevation pops =  CP2, CP3, LV1, LV3, LVTR1, SQ3, WL2, Y011, Y04, Y07, Y08

historical_climate_growthseason <- rbind(historical_climate_low, historical_climate_mid, historical_climate_high)
summary(historical_climate_growthseason)
```

## Climate traits across pops

### Totals

```{r}
names(recent_climate_growthseason)
recent_climate_growthseason_yearlytot <- recent_climate_growthseason %>%  group_by(parent.pop, year, elevation.group) %>% summarise_at(c("pck", "ppt"), sum, na.rm = TRUE)
recent_climate_growthseason_yearlytot$elevation.group <- factor(recent_climate_growthseason_yearlytot$elevation.group, levels = elev_order)

recent_pck_total <- recent_climate_growthseason_yearlytot %>% ggplot(aes(x=year, y=pck, group=parent.pop, color=elevation.group)) + 
  geom_point() + geom_line() + 
  scale_colour_manual(values=elev_three_palette) + 
  ggtitle("Recent Climate") + 
  theme_classic() + 
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45, hjust = 1))
#ggsave("../output/growthseasonTot_SnwPck_RecentClim.png", width = 12, height = 6, units = "in")

recent_ppt_total <- recent_climate_growthseason_yearlytot %>% ggplot(aes(x=year, y=ppt, group=parent.pop, color=elevation.group)) + 
  geom_point() + geom_line() + 
  scale_colour_manual(values=elev_three_palette) + 
  ggtitle("Recent Climate")  + 
  theme_classic() + 
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45, hjust = 1))
#ggsave("../output/growthseasonTot_Precip_RecentClim.png", width = 12, height = 6, units = "in")

historical_climate_growthseason_yearlytot <- historical_climate_growthseason %>% group_by(parent.pop, year, elevation.group) %>% summarise_at(c("pck", "ppt"), sum, na.rm = TRUE)
historical_climate_growthseason_yearlytot$elevation.group <- factor(historical_climate_growthseason_yearlytot$elevation.group, levels = elev_order)

hist_pck_total <- historical_climate_growthseason_yearlytot %>% ggplot(aes(x=year, y=pck, group=parent.pop, color=elevation.group)) + 
  geom_point() + geom_line() + 
  scale_colour_manual(values=elev_three_palette) + 
  ggtitle("Historical Climate") + 
  theme_classic() + 
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45, hjust = 1))
#ggsave("../output/growthseasonTot_SnwPck_HistoricalClim.png", width = 12, height = 6, units = "in")

hist_ppt_total <- historical_climate_growthseason_yearlytot %>% ggplot(aes(x=year, y=ppt, group=parent.pop, color=elevation.group)) + 
  geom_point() + geom_line() + 
  scale_colour_manual(values=elev_three_palette) + 
  ggtitle("Historical Climate") + 
  theme_classic() + 
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45, hjust = 1))
#ggsave("../output/growthseasonTot_Precip_HistoricalClim.png", width = 12, height = 6, units = "in")


#should combine these into one figure and save that instead
legend <- get_legend(hist_pck_total)
hist_pck_total <- hist_pck_total + theme(legend.position="none")
recent_pck_total <- recent_pck_total + theme(legend.position="none")
grid.arrange(hist_pck_total, recent_pck_total, legend, ncol=3, widths=c(3.12, 3.12, 1.09))
 #2400 x 800

legend <- get_legend(hist_ppt_total)
hist_ppt_total <- hist_ppt_total + theme(legend.position="none")
recent_ppt_total <- recent_ppt_total + theme(legend.position="none")
grid.arrange(hist_ppt_total, recent_ppt_total, legend, ncol=3, widths=c(3.12, 3.12, 1.09))
 #2000 x 850
```

### Averages

#### Across last 30 years (all months included)

```{r}
pop_elev_climate_avgs <- pop_elev_climate %>% filter(year>1992) %>% group_by(parent.pop, elevation.group) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), mean, na.rm = TRUE)
pop_elev_climate_avgs #30 year averages of all climate variables 
pop_elev_climate_avgs$elevation.group <- factor(pop_elev_climate_avgs$elevation.group, levels=elev_order)
```

#### Recent Years - Growth Season

```{r}
recent_climate_growthseason_avgs <- recent_climate_growthseason %>% group_by(parent.pop, elevation.group) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(recent_climate_growthseason_avgs) <- gsub("fn2", "sem", colnames(recent_climate_growthseason_avgs))
names(recent_climate_growthseason_avgs) <-gsub("fn1", "mean", colnames(recent_climate_growthseason_avgs))
recent_climate_growthseason_avgs #30 year averages during growth season months 
recent_climate_growthseason_avgs$elevation.group <- factor(recent_climate_growthseason_avgs$elevation.group, levels=elev_order)       

recent_climate_growthseason_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, cwd_mean), y=cwd_mean, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=cwd_mean-cwd_sem,ymax=cwd_mean+cwd_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation",x="Population", y="Avg CWD" ,title = "Average CWD during Growth Season - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/growthseasonAvg_CWD_RecentClim.png", width = 12, height = 6, units = "in")

recent_climate_growthseason_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_mean), y=ppt_mean, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=ppt_mean-ppt_sem,ymax=ppt_mean+ppt_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation",x="Population", y="Avg Precip" ,title = "Average Precip during Growth Season - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/growthseasonAvg_Precip_RecentClim.png", width = 12, height = 6, units = "in")

recent_climate_growthseason_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, pck_mean), y=pck_mean, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=pck_mean-pck_sem,ymax=pck_mean+pck_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation",x="Population", y="Avg SnwPck" ,title = "Average SnwPck during Growth Season - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/growthseasonAvg_SnwPck_RecentClim.png", width = 12, height = 6, units = "in")

recent_climate_growthseason_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmn_mean), y=tmn_mean, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmn_mean-tmn_sem,ymax=tmn_mean+tmn_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation",x="Population", y="Avg MinTemp" ,title = "Average MinTemp during Growth Season - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/growthseasonAvg_MinTemp_RecentClim.png", width = 12, height = 6, units = "in")

recent_climate_growthseason_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmx_mean), y=tmx_mean, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmx_mean-tmx_sem,ymax=tmx_mean+tmx_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation",x="Population", y="Avg MaxTemp" ,title = "Average MaxTemp during Growth Season - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/growthseasonAvg_MaxTemp_RecentClim.png", width = 12, height = 6, units = "in")
```

#### Historical Years - Growth Season

Need to edit this

```{r}
#Historical climate
#historical_climate_growthseason_avgs
```

## Climate trait correlations

```{r}
recent_climate_growthseason %>% cor_test(cwd, pck, ppt, tmn, tmx, method = "pearson")
recent_cor_mat <- recent_climate_growthseason %>% select(cwd, pck, ppt, tmn, tmx) %>% cor_mat()
recent_cor_mat

recent_cor = recent_climate_growthseason %>% select(cwd, pck, ppt, tmn, tmx) %>% cor()
file_path= "../output/GrowthSeason_RecentClim_Cors.png"
png(width = 12, height = 6, res= 300, units = "in", file=file_path, type = "cairo")
corrplot(recent_cor)
dev.off()

historical_climate_growthseason %>% cor_test(cwd, pck, ppt, tmn, tmx, method = "pearson")
historical_cor_mat <- historical_climate_growthseason %>% select(cwd, pck, ppt, tmn, tmx) %>% cor_mat()
historical_cor_mat

historical_cor = historical_climate_growthseason %>% select(cwd, pck, ppt, tmn, tmx) %>% cor()
file_path= "../output/GrowthSeason_HistoricalClim_Cors.png"
png(width = 12, height = 6, res= 300, units = "in", file=file_path, type = "cairo")
corrplot(historical_cor)
dev.off()
```

Trait correlations are the same across recent and historical time
periods (during the germination season)

## Climate PCAs

-   Should use this code to check the significance of the PCA:
    <https://github.com/StatQuest/pca_demo/blob/master/pca_demo.R>
-   Remember this paper: Björklund, M. 2019. Be careful with your
    principal components. Evolution 73: 2151--2158.

### All years and months included

Produce basic PCA plots

```{r}
#normalize the data
climate_normalized <- recent_climate_growthseason %>% select(cwd, pck, ppt, tmn, tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized)
cor.norm = cor(climate_normalized) #test correlations among the traits
corrplot(cor.norm)

climate.pca <- princomp(cor.norm) 
summary(climate.pca) #PC1 explains 90% of the variation, PC2 explains 8.1%, together they explain 98%
climate.pca$loadings[, 1:2] #PC1 has high positive loading values for tmn, tmx, cwd
fviz_eig(climate.pca) #scree plot 
fviz_pca_var(climate.pca, col.var = "black") #biplot

#The goal of the third visualization is to determine how much each variable is represented in a given component. Such a quality of representation is called the Cos2 and corresponds to the square cosine, and it is computed using the fviz_cos2 function. (https://www.datacamp.com/tutorial/pca-analysis-r)
#A low value means that the variable is not perfectly represented by that component. 
#A high value, on the other hand, means a good representation of the variable on that component.
fviz_cos2(climate.pca, choice = "var", axes = 1:2) #axes says which PCs to use 
#biplot combined wiht cos2
fviz_pca_var(climate.pca, col.var = "cos2", 
            gradient.cols = c("black", "orange", "green"),
            repel = TRUE)
```

Overlay populations on PCA plot

```{r}
pc = prcomp(climate_normalized, scale = TRUE)

pc_data = data.frame(pc$x)

locs_pc = cbind(recent_climate_growthseason, pc_data)

loadings = data.frame(varnames=rownames(pc$rotation), pc$rotation)

ggplot() +
  geom_point(data = filter(locs_pc, elevation.group == "High"), aes(x = PC1, y = PC2), color = "#0043F0", alpha = 0.6) +
  geom_point(data = filter(locs_pc, elevation.group == "Mid"), aes(x = PC1, y = PC2), color = "#C9727F", alpha = 0.6) +
  geom_point(data = filter(locs_pc, elevation.group == "Low"), aes(x = PC1, y = PC2), color = "#F5A540", alpha = 0.6) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
ggsave("../output/RecentClim_PCA_Elev.png", width = 6, height = 4, units = "in")
```

### Average across al years and months (1 datapoint per pop)

Produce basic PCA plots

```{r}
#normalize the data
climate_normalized_avgs <- recent_climate_growthseason_avgs %>% ungroup() %>% select(cwd_mean, pck_mean, ppt_mean, tmn_mean, tmx_mean) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_avgs)
cor.norm_avgs = cor(climate_normalized_avgs) #test correlations among the traits
corrplot(cor.norm_avgs)

climate.pca_avgs <- princomp(cor.norm_avgs) 
summary(climate.pca_avgs) #PC1 explains 90% of the variation, PC2 explains 8.1%, together they explain 98%
climate.pca_avgs$loadings[, 1:2] #PC1 has high positive loading values for tmn, tmx, cwd
fviz_eig(climate.pca_avgs) #scree plot 
fviz_pca_var(climate.pca_avgs, col.var = "black") #biplot

#The goal of the third visualization is to determine how much each variable is represented in a given component. Such a quality of representation is called the Cos2 and corresponds to the square cosine, and it is computed using the fviz_cos2 function. (https://www.datacamp.com/tutorial/pca-analysis-r)
#A low value means that the variable is not perfectly represented by that component. 
#A high value, on the other hand, means a good representation of the variable on that component.
fviz_cos2(climate.pca_avgs, choice = "var", axes = 1:2) #axes says which PCs to use 
#biplot combined wiht cos2
fviz_pca_var(climate.pca_avgs, col.var = "cos2", 
            gradient.cols = c("black", "orange", "green"),
            repel = TRUE)

#RecentClim_GrowthSeason_PCA_Avgs 5.5 x 5
```

Overlay populations on PCA plot

```{r}
pc_avgs = prcomp(climate_normalized_avgs, scale = TRUE)

pc_data = data.frame(pc_avgs$x)

locs_pc = cbind(recent_climate_growthseason_avgs, pc_data)
names(locs_pc)

loadings = data.frame(varnames=rownames(pc_avgs$rotation), pc_avgs$rotation)

ggplot() +
  geom_point(data = filter(locs_pc, elevation.group == "High"), size=2, aes(x = PC1, y = PC2), color = "#0043F0") +
  geom_point(data = filter(locs_pc, elevation.group == "Mid"), size=2, aes(x = PC1, y = PC2), color = "#C9727F") +
  geom_point(data = filter(locs_pc, elevation.group == "Low"), size=2, aes(x = PC1, y = PC2), color = "#F5A540") + 
  geom_label_repel(data = locs_pc, aes(x = PC1, y = PC2, label = parent.pop),
                  min.segment.length = 0, box.padding = 0.5) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
ggsave("../output/RecentClim_PCA_Elev_Avgs.png", width = 6, height = 4, units = "in")
```

### Average across years per month

Calculate monthly averages

```{r}
names(recent_climate_growthseason)
recent_climate_growthseason_mosavgs <- recent_climate_growthseason %>% group_by(parent.pop, elevation.group, month) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(recent_climate_growthseason_mosavgs) <- gsub("fn2", "sem", colnames(recent_climate_growthseason_mosavgs))
names(recent_climate_growthseason_mosavgs) <-gsub("fn1", "mean", colnames(recent_climate_growthseason_mosavgs))
recent_climate_growthseason_mosavgs #30 year averages per growth season month 
recent_climate_growthseason_mosavgs$elevation.group <- factor(recent_climate_growthseason_mosavgs$elevation.group, levels=elev_order) 
```

Produce basic PCA plots

*Color by growth month relative to germination season?*

```{r}
names(recent_climate_growthseason_mosavgs)
#normalize the data
climate_normalized_mosavgs <- recent_climate_growthseason_mosavgs %>% ungroup() %>% select(cwd_mean, pck_mean, ppt_mean, tmn_mean, tmx_mean) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_mosavgs)
cor.norm_mosavgs = cor(climate_normalized_mosavgs) #test correlations among the traits
corrplot(cor.norm_mosavgs)

climate.pca_mosavgs <- princomp(cor.norm_mosavgs) 
summary(climate.pca_mosavgs) #PC1 explains 90% of the variation, PC2 explains 8.1%, together they explain 98%
climate.pca_mosavgs$loadings[, 1:2] #PC1 has high positive loading values for tmn, tmx, cwd
fviz_eig(climate.pca_mosavgs) #scree plot 
fviz_pca_var(climate.pca_mosavgs, col.var = "black") #biplot

#The goal of the third visualization is to determine how much each variable is represented in a given component. Such a quality of representation is called the Cos2 and corresponds to the square cosine, and it is computed using the fviz_cos2 function. (https://www.datacamp.com/tutorial/pca-analysis-r)
#A low value means that the variable is not perfectly represented by that component. 
#A high value, on the other hand, means a good representation of the variable on that component.
fviz_cos2(climate.pca_mosavgs, choice = "var", axes = 1:2) #axes says which PCs to use 
#biplot combined wiht cos2
fviz_pca_var(climate.pca_mosavgs, col.var = "cos2", 
            gradient.cols = c("black", "orange", "green"),
            repel = TRUE)
#RecentClim_GrowthSeason_PCA_MonthlyAvgs 5.5 x 5
```

Overlay populations on PCA plot

```{r}
pc_mosavgs = prcomp(climate_normalized_mosavgs, scale = TRUE)

pc_data = data.frame(pc_mosavgs$x)

locs_pc = cbind(recent_climate_growthseason_mosavgs, pc_data)

loadings = data.frame(varnames=rownames(pc_mosavgs$rotation), pc_mosavgs$rotation)

ggplot() +
  geom_point(data = filter(locs_pc, elevation.group == "High"), size=2, aes(x = PC1, y = PC2), color = "#0043F0") +
  geom_point(data = filter(locs_pc, elevation.group == "Mid"), size=2, aes(x = PC1, y = PC2), color = "#C9727F") +
  geom_point(data = filter(locs_pc, elevation.group == "Low"), size=2, aes(x = PC1, y = PC2), color = "#F5A540") + 
  geom_text_repel(data = locs_pc, aes(x = PC1, y = PC2, label = parent.pop),
             min.segment.length = 0) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
ggsave("../output/RecentClim_PCA_Elev_MonthlyAvgs_Pops.png", width = 6, height = 4, units = "in")

ggplot() +
  geom_point(data = filter(locs_pc, elevation.group == "High"), size=2, aes(x = PC1, y = PC2), color = "#0043F0") +
  geom_point(data = filter(locs_pc, elevation.group == "Mid"), size=2, aes(x = PC1, y = PC2), color = "#C9727F") +
  geom_point(data = filter(locs_pc, elevation.group == "Low"), size=2, aes(x = PC1, y = PC2), color = "#F5A540") + 
  geom_text_repel(data = locs_pc, aes(x = PC1, y = PC2, label = month),
                  min.segment.length = 0) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
ggsave("../output/RecentClim_PCA_Elev_MonthlyAvgs_Months.png", width = 6, height = 4, units = "in")
```
