---
title: "Climate PCAs_Growth Season"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To Do:

-   Compare historical 60 years to recent 30 years as another climate change test

-   Check the significance of the PCA

    -   Remember this paper: Björklund, M. 2019. Be careful with your principal components. Evolution 73: 2151--2158.

# Climate PCAs for Growth Season

Note: No timeperiod effects when 2024 included in the PCA

## Load necessary libraries

```{r}
library(tidyverse)
library(ggrepel)
#library(cowplot)
library(gridExtra)
library(corrplot) #plotting correlations 
library(QBMS) #for function calc_biovars to calculate bioclim variables
library(ggfortify) #easier PCA figures
library(vegan) #for permanova 
sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
} #legend function for grid_arrange

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")
month_order <- c("jan","feb","mar","apr","may","jun","jul","aug","sep","oct","nov","dec")
```

## Load the Flint growth season data (from "Flint_Growth_Season.Rmd")

```{r}
flint_recent_grwseason <- read_csv("../output/Climate/flint_climate_growthseason_recent.csv") %>%
  dplyr::select(month, parent.pop:tmx)
head(flint_recent_grwseason, 12)

flint_historical_grwseason <- read_csv("../output/Climate/flint_climate_growthseason_historical.csv") %>%
  dplyr::select(month, parent.pop:tmx)
head(flint_historical_grwseason, 12)
```

### For 2024
```{r}
flint_recent_grwseason_2024 <- read_csv("../output/Climate/flint_climate_growthseason_recent_2024.csv") %>%
  dplyr::select(month, parent.pop:tmx)
head(flint_recent_grwseason_2024, 12)

flint_historical_grwseason_2024 <- read_csv("../output/Climate/flint_climate_growthseason_historical_2024.csv") %>%
  dplyr::select(month, parent.pop:tmx)
head(flint_historical_grwseason_2024, 12)
```


## Load the Bioclim growth season data (from "BioClim_Growth_Season.Rmd")

```{r}
bioclim_recent_grwseason <- read_csv("../output/Climate/BioClim_growthseason_Recent.csv")
bioclim_historical_grwseason <- read_csv("../output/Climate/BioClim_growthseason_Historical.csv")
```

### For 2024
```{r}
bioclim_recent_grwseason_2024 <- read_csv("../output/Climate/BioClim_growthseason_Recent_2024.csv")
bioclim_historical_grwseason_2024 <- read_csv("../output/Climate/BioClim_growthseason_Historical_2024.csv")
```

## Yearly averages for Flint (for climate dist calcs)
```{r}
flint_grwssn_recent_yravgs <- flint_recent_grwseason %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, year) %>% 
  summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE)
flint_grwssn_recent_yravgs

flint_grwssn_historical_yravgs <- flint_historical_grwseason %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, year) %>% 
  summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) 
flint_grwssn_historical_yravgs
```

### For 2024
```{r}
flint_grwssn_recent_yravgs_2024 <- flint_recent_grwseason_2024 %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, year) %>% 
  summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE)
flint_grwssn_recent_yravgs_2024

flint_grwssn_historical_yravgs_2024 <- flint_historical_grwseason_2024 %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, year) %>% 
  summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) 
flint_grwssn_historical_yravgs_2024
```

## Merge flint yrly avgs with bioclim
```{r}
bioclim_flint_yrly_avgs_recent <- full_join(flint_grwssn_recent_yravgs, bioclim_recent_grwseason) 
#write_csv(bioclim_flint_yrly_avgs_recent, "../output/Climate/growthseason_yrly_avgs_Recent.csv")
bioclim_flint_yrly_avgs_historical <- full_join(flint_grwssn_historical_yravgs, bioclim_historical_grwseason) 
#write_csv(bioclim_flint_yrly_avgs_historical, "../output/Climate/growthseason_yrly_avgs_Historical.csv")
```

### For 2024
```{r}
bioclim_flint_yrly_avgs_recent_2024 <- full_join(flint_grwssn_recent_yravgs_2024, bioclim_recent_grwseason_2024) 
#write_csv(bioclim_flint_yrly_avgs_recent_2024, "../output/Climate/growthseason_yrly_avgs_Recent_2024.csv")
bioclim_flint_yrly_avgs_historical_2024 <- full_join(flint_grwssn_historical_yravgs_2024, bioclim_historical_grwseason_2024) 
#write_csv(bioclim_flint_yrly_avgs_historical_2024, "../output/Climate/growthseason_yrly_avgs_Historical_2024.csv")
```

## WL2 Garden 

### 2023
```{r}
WL2Grdn_flint_bioclim_grwseason <- read_csv("../output/Climate/WL2_Garden_GrwSsnClim_2023.csv") %>% 
  rename_with(~str_remove(., "_WL2"), everything())
```

### 2024
```{r}
WL2Grdn_flint_bioclim_2024_grwseason <- read_csv("../output/Climate/WL2_Garden_GrwSsnClim_2024.csv") %>% 
  rename_with(~str_remove(., "_WL2"), everything())
```

## 2023 Avg across years and months (Flint + bioclim)

### Calculate avgs for flint

```{r}
flint_recent_grwseason_avgs <- flint_recent_grwseason %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Recent") 
flint_recent_grwseason_avgs

flint_historical_grwseason_avgs <- flint_historical_grwseason %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Historical")
flint_historical_grwseason_avgs

flint_grwseason_avgs <- bind_rows(flint_recent_grwseason_avgs, flint_historical_grwseason_avgs) #combine into 1 dataframe 
head(flint_grwseason_avgs)
tail(flint_grwseason_avgs)
#write_csv(flint_grwseason_avgs, "../output/Climate/growthseason_FlintAvgs.csv")
```

### Calculate avgs for bioclim

```{r}
bioclim_recent_grwseason_avgs <- bioclim_recent_grwseason %>% 
  group_by(parent.pop, elevation.group, elev_m) %>% 
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_month", "tmean_driest_month", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_month", "ppt_coldest_month"),
               c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Recent") 
bioclim_recent_grwseason_avgs

bioclim_historical_grwseason_avgs <- bioclim_historical_grwseason %>% 
  group_by(parent.pop, elevation.group, elev_m) %>% 
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_month", "tmean_driest_month", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_month", "ppt_coldest_month"),
               c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Historical") 
bioclim_historical_grwseason_avgs

bioclim_grwseason_avgs <- bind_rows(bioclim_recent_grwseason_avgs, bioclim_historical_grwseason_avgs) #combine into 1 dataframe 
head(bioclim_grwseason_avgs)
tail(bioclim_grwseason_avgs)
#write_csv(bioclim_grwseason_avgs, "../output/Climate/growthseason_BioClimAvgs.csv")
```

Merge with flint

```{r}
bioclim_flint_grwseason_avgs <- full_join(flint_grwseason_avgs, bioclim_grwseason_avgs) %>% 
  select(TimePd, parent.pop:ppt_coldest_month)
head(bioclim_flint_grwseason_avgs)
```

### Correlations - Recent + Historical

```{r}
#normalize the data
climate_normalized_bioclim_flint_grwseason_avgs <- bioclim_flint_grwseason_avgs %>% ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_bioclim_flint_grwseason_avgs)
cor.norm = cor(climate_normalized_bioclim_flint_grwseason_avgs) #test correlations among the traits
cor.norm
corrplot(cor.norm)

#tmn, tmx, tmean_driest_month and ann_tmean all highly correlated (90-98%) - only keep ann_tmean 
```

Recent
```{r}
climate_normalized_bioclim_flint_grwseason_avgs_recent <- bioclim_flint_grwseason_avgs %>% 
  ungroup() %>% 
  filter(TimePd=="Recent") %>% 
  select(tmn:tmean_driest_month, cwd, ppt, ann_ppt:ppt_coldest_month) %>% 
  scale() #normalize the data so they're all on the same scale

cor.norm_recent = cor(climate_normalized_bioclim_flint_grwseason_avgs_recent) #test correlations among the traits
cor.sig_recent <- cor.mtest(climate_normalized_bioclim_flint_grwseason_avgs_recent, method = "pearson")

corrplot(cor.norm_recent, type="upper",
         tl.srt = 45, p.mat = cor.sig_recent$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```

Historical
```{r}
climate_normalized_bioclim_flint_grwseason_avgs_historic <- bioclim_flint_grwseason_avgs %>% 
  ungroup() %>% 
  filter(TimePd=="Historical") %>% 
  select(tmn:tmean_driest_month, cwd, ppt, ann_ppt:ppt_coldest_month) %>% 
  scale() #normalize the data so they're all on the same scale

cor.norm_historic = cor(climate_normalized_bioclim_flint_grwseason_avgs_historic) #test correlations among the traits
cor.sig_historic <- cor.mtest(climate_normalized_bioclim_flint_grwseason_avgs_historic, method = "pearson")

corrplot(cor.norm_historic, type="upper",
         tl.srt = 45, p.mat = cor.sig_historic$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```

### PCA - Recent + Historical

```{r}
all_bioclim_flint_avgs.pc = prcomp(bioclim_flint_grwseason_avgs[c(7:8, 11:15, 17:20)], scale = TRUE, center = TRUE) 
str(all_bioclim_flint_avgs.pc)
```

plot % Variance Explained

```{r}
summary(all_bioclim_flint_avgs.pc)

tibble(PC=str_c("PC",str_pad(1:10,2,pad="0")),
       percent_var=all_bioclim_flint_avgs.pc$sdev[1:10]^2/sum(all_bioclim_flint_avgs.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
all_bioclim_flint_avgs.pc.dat = data.frame(all_bioclim_flint_avgs.pc$x)

all_bioclim_flint_avgs_locs.pc = cbind(bioclim_flint_grwseason_avgs, all_bioclim_flint_avgs.pc.dat)

all_bioclim_flint_avgs_loadings = data.frame(varnames=rownames(all_bioclim_flint_avgs.pc$rotation), all_bioclim_flint_avgs.pc$rotation)
all_bioclim_flint_avgs_loadings
```

```{r}
autoplot(all_bioclim_flint_avgs.pc, data = bioclim_flint_grwseason_avgs,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
autoplot(all_bioclim_flint_avgs.pc, data = bioclim_flint_grwseason_avgs,
         x=2, y=4,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
all_bioclim_flint_avgs_locs.pc_avg <- all_bioclim_flint_avgs_locs.pc %>%
  group_by(parent.pop, elev_m, TimePd) %>%
  summarise(across(.cols=starts_with("PC"), .fns = mean)) %>%
  ungroup()

#all_bioclim_flint_avgs_locs.pc_avg
```

```{r}
all_bioclim_flint_avgs_locs.pc_avg %>% 
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point() +
  facet_wrap(~TimePd) +
  coord_fixed(ratio = 1.5)
```

## 2023 - Permanova
See Jenny's github: https://github.com/jrgremer/Seedbanks_drought_deluge  
```{r}
all_bioclim_flint_avgs_locs.pc_dist <- all_bioclim_flint_avgs_locs.pc %>% ungroup() %>% select(PC1:PC7)
dist_matrix_grwssn <- dist(all_bioclim_flint_avgs_locs.pc_dist, method = "euclidian") #use a distance function to calculate euclidian distance in PCA space
permanova_results_grwssn <- adonis2(dist_matrix_grwssn ~ TimePd*elev_m*Lat, data = all_bioclim_flint_avgs_locs.pc) #use adonis2 to run the permanova
#rows in data must be in same order as rows in distance matrix 
#by = NULL will assess the overall significance of all terms together

#look at output 
summary(permanova_results_grwssn)
permanova_results_grwssn
permanova_results_grwssn$`Pr(>F)`

permanova_results_grwssn_terms <- adonis2(dist_matrix_grwssn ~ TimePd*elev_m*Lat, data = all_bioclim_flint_avgs_locs.pc, by = "terms") #by = "terms" will assess significance for each term (sequentially from first to last)
permanova_results_grwssn_terms
```

## 2023 - LM on PCs (follow up on permanova)

For each PC fit a model with timepd and parent.pop/elev to test for climate change

```{r}
lmer_results_grwssn <- all_bioclim_flint_avgs_locs.pc %>%
  ungroup() %>% 
  select(TimePd, parent.pop, elev_m, Lat, Long, PC1:PC10) %>% 
  pivot_longer(starts_with("PC", ignore.case = FALSE), 
               names_to = "PC", values_to = "value") %>% 
  group_by(PC) %>% 
  nest(data=c(TimePd, parent.pop, elev_m, Lat, Long, value)) %>% 
  mutate(glm=map(data, ~ glm(value ~ TimePd*elev_m*Lat,
                               data=.x)),
         anova = map(glm, ~ broom.mixed::tidy(anova(.x))))

#lmer_results_grwssn
```

```{r}
PC_anova <- lmer_results_grwssn %>% select(-data, -glm) %>% unnest(anova) %>%
  select(PC, term, p.value) %>%
  filter(p.value < 0.05) %>%
  arrange(term, p.value)

PC_anova

lmer_results_grwssn %>% select(-data, -glm) %>% unnest(anova) %>%
  filter(PC=="PC2"| PC=="PC4") #the two PCs with most sig p-values for timepd

#mod_test <- glm(PC4 ~ TimePd*elev_m*Lat, data=all_bioclim_flint_avgs_locs.pc)
#summary(mod_test)
#anova(mod_test)
```

```{r}
autoplot(all_bioclim_flint_avgs.pc, data = bioclim_flint_grwseason_avgs,
         x=2, y=4,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()

all_bioclim_flint_avgs_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC2, y=PC4, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)
#high elev climate seems to be shifting to be more similar to low elev 

home_sites_pca <- all_bioclim_flint_avgs_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC2, y=PC4, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  #geom_point(size=2, alpha=0.7) +
  labs(x="PC2 (30.76%)", y="PC4 (10.65%)", color="Elevation (m)") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  annotate("text", x = -0.06, y= -0.13, label = "WL2", colour = "purple") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8) +
  annotate("text", x = -4.2, y = -4.5, label = "Low Temp \n Seasonality") +
  annotate("text", x = 3.6, y = -4.5, label = "High Temp \n Seasonality") +
  annotate("text", x = -5.24, y = -3.3, label = "Low Diurnal \n Range") +
  annotate("text", x = -5.24, y = 2.5, label = "High Diurnal \n Range") +
  coord_cartesian(ylim = c(-3.5, 3), xlim = c(-4, 3.5), clip = "off") +
  theme_classic() +
  theme(text=element_text(size=28))
home_sites_pca
#ggsave("../output/Climate/GRWSSN_PC2-PC4.png", width = 7, height = 5.5, units = "in")

#all_bioclim_flint_avgs_locs.pc_avg %>% filter(parent.pop=="WL2")
```

## 2023 Add the Garden to the Final PCA 
```{r}
WL2Grdn_pc_prep <- WL2Grdn_flint_bioclim_grwseason %>% select(cwd, ppt, ann_tmean, mean_diurnal_range, temp_seasonality, temp_ann_range, tmean_wettest_month, ann_ppt, ppt_seasonality, ppt_warmest_month, ppt_coldest_month)
#scale(WL2Grdn_pc_prep, all_bioclim_flint_avgs.pc$center, all_bioclim_flint_avgs.pc$scale) %*% all_bioclim_flint_avgs.pc$rotation  #gives same result as below 
WL2Grdn_predicted <- predict(all_bioclim_flint_avgs.pc, newdata = WL2Grdn_pc_prep)

WL2Grdn_pc_prep_2024 <- WL2Grdn_flint_bioclim_2024_grwseason %>% select(cwd, ppt, ann_tmean, mean_diurnal_range, temp_seasonality, temp_ann_range, tmean_wettest_month, ann_ppt, ppt_seasonality, ppt_warmest_month, ppt_coldest_month)
WL2Grdn_2024_predicted <- predict(all_bioclim_flint_avgs.pc, newdata = WL2Grdn_pc_prep_2024)

#str(home_sites_pca) #can add predicted WL2 point to the existing plot's data 
home_sites_pca$data <- rbind(home_sites_pca$data, 
  data.frame(
    parent.pop = "WL2_Garden",
    elev_m = 2020,
    TimePd = c("2023", "2024"),
    PC1 = c(WL2Grdn_predicted[, "PC1"], WL2Grdn_2024_predicted[, "PC1"]),
    PC2 = c(WL2Grdn_predicted[, "PC2"], WL2Grdn_2024_predicted[, "PC2"]),
    PC3 = c(WL2Grdn_predicted[, "PC3"], WL2Grdn_2024_predicted[, "PC3"]),
    PC4 = c(WL2Grdn_predicted[, "PC4"], WL2Grdn_2024_predicted[, "PC4"]),
    PC5 = c(WL2Grdn_predicted[, "PC5"], WL2Grdn_2024_predicted[, "PC5"]),
    PC6 = c(WL2Grdn_predicted[, "PC6"], WL2Grdn_2024_predicted[, "PC6"]),
    PC7 = c(WL2Grdn_predicted[, "PC7"], WL2Grdn_2024_predicted[, "PC7"]),
    PC8 = c(WL2Grdn_predicted[, "PC8"], WL2Grdn_2024_predicted[, "PC8"]),
    PC9 = c(WL2Grdn_predicted[, "PC9"], WL2Grdn_2024_predicted[, "PC9"]),
    PC10 = c(WL2Grdn_predicted[, "PC10"], WL2Grdn_2024_predicted[, "PC10"]),
    PC11 = c(WL2Grdn_predicted[, "PC11"], WL2Grdn_2024_predicted[, "PC11"]),
    group = c("new", "new2")
  )
)

home_sites_pca + 
  geom_point(data=filter(home_sites_pca$data, parent.pop == "WL2_Garden"), size=3, shape = 8, show.legend = FALSE) +
  annotate("text", x = -.6, y= -2.7, label = "WL2 Garden \n 2023", colour = "purple") +
  annotate("text", x = 2.7, y= -2.9, label = "WL2 Garden \n 2024", colour = "purple") 
ggsave("../output/Climate/GRWSSN_PC2-PC4_PlusGarden.png", width = 7.4, height = 6, units = "in")
```

## Add the Garden to the PCA with PC1 & PC2
```{r}
home_sites_pca2 <- all_bioclim_flint_avgs_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  #geom_point(size=2, alpha=0.7) +
  labs(y="PC2 (30.76%)", x="PC1 (39.15%)", color="Elevation (m)") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  annotate("text", x = 0.3, y= -0.18, label = "WL2", colour = "purple") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8) +
  annotate("text", y = -3.4, x = -5.9, label = "Low Temp \n Seasonality") +
  annotate("text", y = 3, x = -5.9, label = "High Temp \n Seasonality") +
  annotate("text", y = -5, x = 3.3, label = "Low Ann \n PPT") +
  annotate("text", y = -5, x = -4.1, label = "High Ann \n PPT") +
  coord_cartesian(ylim = c(-3.9, 3), xlim = c(-4.5, 3.5), clip = "off") +
  theme_classic() +
  theme(text=element_text(size=28))
home_sites_pca2

#all_bioclim_flint_avgs_locs.pc_avg %>% filter(parent.pop=="WL2")


home_sites_pca2$data <- rbind(home_sites_pca2$data, 
  data.frame(
    parent.pop = "WL2_Garden",
    elev_m = 2020,
    TimePd = c("2023", "2024"),
    PC1 = c(WL2Grdn_predicted[, "PC1"], WL2Grdn_2024_predicted[, "PC1"]),
    PC2 = c(WL2Grdn_predicted[, "PC2"], WL2Grdn_2024_predicted[, "PC2"]),
    PC3 = c(WL2Grdn_predicted[, "PC3"], WL2Grdn_2024_predicted[, "PC3"]),
    PC4 = c(WL2Grdn_predicted[, "PC4"], WL2Grdn_2024_predicted[, "PC4"]),
    PC5 = c(WL2Grdn_predicted[, "PC5"], WL2Grdn_2024_predicted[, "PC5"]),
    PC6 = c(WL2Grdn_predicted[, "PC6"], WL2Grdn_2024_predicted[, "PC6"]),
    PC7 = c(WL2Grdn_predicted[, "PC7"], WL2Grdn_2024_predicted[, "PC7"]),
    PC8 = c(WL2Grdn_predicted[, "PC8"], WL2Grdn_2024_predicted[, "PC8"]),
    PC9 = c(WL2Grdn_predicted[, "PC9"], WL2Grdn_2024_predicted[, "PC9"]),
    PC10 = c(WL2Grdn_predicted[, "PC10"], WL2Grdn_2024_predicted[, "PC10"]),
    PC11 = c(WL2Grdn_predicted[, "PC11"], WL2Grdn_2024_predicted[, "PC11"]),
    group = c("new", "new2")
  )
)

home_sites_pca2 + 
  geom_point(data=filter(home_sites_pca2$data, parent.pop == "WL2_Garden"), size=3, shape = 8, show.legend = FALSE) +
  annotate("text", y = -1.1, x= 0.5, label = "WL2 Garden \n 2023", colour = "purple") +
  annotate("text", y = 2.2, x= 2.8, label = "WL2 Garden \n 2024", colour = "purple") 
ggsave("../output/Climate/GRWSSN_PC1-PC2_PlusGarden.png", width = 7.4, height = 6, units = "in")
```


## 2024 Avg across years and months

### Calculate avgs for flint
```{r}
flint_grwssn_recent_2024_avgs <- flint_recent_grwseason_2024 %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Recent") 
#flint_grwssn_recent_2024_avgs

flint_grwssn_historical_2024_avgs <- flint_historical_grwseason_2024 %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Historical")
#flint_grwssn_historical_2024_avgs

flint_grwssn_avgs_2024 <- bind_rows(flint_grwssn_recent_2024_avgs, flint_grwssn_historical_2024_avgs) #combine into 1 dataframe 
#head(flint_grwssn_avgs_2024)
#tail(flint_grwssn_avgs_2024)
#write_csv(flint_grwssn_avgs_2024, "../output/Climate/grwssn_FlintAvgs_2024.csv")
```

### Calculate avgs for bioclim
```{r}
bioclim_grwssn_recent_2024_avgs <- bioclim_recent_grwseason_2024 %>% 
  group_by(parent.pop, elevation.group, elev_m) %>% 
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_month", "tmean_driest_month", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_month", "ppt_coldest_month"),
               c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Recent") 
#bioclim_grwssn_recent_2024_avgs

bioclim_grwssn_historical_2024_avgs <- bioclim_recent_grwseason_2024 %>% 
  group_by(parent.pop, elevation.group, elev_m) %>% 
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_month", "tmean_driest_month", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_month", "ppt_coldest_month"),
               c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Historical") 
#bioclim_grwssn_historical_2024_avgs

bioclim_grwssn_avgs_2024 <- bind_rows(bioclim_grwssn_recent_2024_avgs, bioclim_grwssn_historical_2024_avgs) #combine into 1 dataframe 
#head(bioclim_grwssn_avgs_2024)
#tail(bioclim_grwssn_avgs_2024)
#write_csv(bioclim_grwssn_avgs_2024, "../output/Climate/grwssn_BioClimAvgs_2024.csv")
```

Merge the two 
```{r}
bioclim_flint_grwseason_avgs_2024 <- full_join(flint_grwssn_avgs_2024, bioclim_grwssn_avgs_2024) %>% 
  select(TimePd, parent.pop:ppt_coldest_month) %>% 
  select(-pck)
#head(bioclim_flint_grwseason_avgs_2024)
```

### Correlations - Recent + Historical

```{r}
#normalize the data
climate_normalized_bioclim_flint_grwseason_avgs_2024 <- bioclim_flint_grwseason_avgs_2024 %>% ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_bioclim_flint_grwseason_avgs_2024)
cor.norm_2024 = cor(climate_normalized_bioclim_flint_grwseason_avgs_2024) #test correlations among the traits
cor.norm_2024
corrplot(cor.norm_2024)

#tmn, tmx, tmean_driest_month and ann_tmean all highly correlated (90-98%) - only keep ann_tmean 
```

Recent
```{r}
climate_normalized_bioclim_flint_grwseason_avgs_2024_recent <- bioclim_flint_grwseason_avgs_2024 %>% 
  ungroup() %>% 
  filter(TimePd=="Recent") %>% 
  select(tmn:tmean_driest_month, cwd, ppt, ann_ppt:ppt_coldest_month) %>% 
  scale() #normalize the data so they're all on the same scale

cor.norm_2024_recent = cor(climate_normalized_bioclim_flint_grwseason_avgs_2024_recent) #test correlations among the traits
cor.sig_2024_recent <- cor.mtest(climate_normalized_bioclim_flint_grwseason_avgs_2024_recent, method = "pearson")

corrplot(cor.norm_2024_recent, type="upper",
         tl.srt = 45, p.mat = cor.sig_2024_recent$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```

Historical
```{r}
climate_normalized_bioclim_flint_grwseason_avgs_2024_historic <- bioclim_flint_grwseason_avgs_2024 %>% 
  ungroup() %>% 
  filter(TimePd=="Historical") %>% 
  select(tmn:tmean_driest_month, cwd, ppt, ann_ppt:ppt_coldest_month) %>% 
  scale() #normalize the data so they're all on the same scale

cor.norm_2024_historic = cor(climate_normalized_bioclim_flint_grwseason_avgs_2024_historic) #test correlations among the traits
cor.sig_2024_historic <- cor.mtest(climate_normalized_bioclim_flint_grwseason_avgs_2024_historic, method = "pearson")

corrplot(cor.norm_2024_historic, type="upper",
         tl.srt = 45, p.mat = cor.sig_2024_historic$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```

### PCA - Recent + Historical

```{r}
all_bioclim_flint_avgs_2024.pc = prcomp(bioclim_flint_grwseason_avgs_2024[c(7:8, 11:15, 17:20)], scale = TRUE, center = TRUE) 
str(all_bioclim_flint_avgs_2024.pc)
```

plot % Variance Explained

```{r}
summary(all_bioclim_flint_avgs_2024.pc)

tibble(PC=str_c("PC",str_pad(1:10,2,pad="0")),
       percent_var=all_bioclim_flint_avgs_2024.pc$sdev[1:10]^2/sum(all_bioclim_flint_avgs_2024.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
all_bioclim_flint_avgs_2024.pc.dat = data.frame(all_bioclim_flint_avgs_2024.pc$x)

all_bioclim_flint_avgs_2024_locs.pc = cbind(bioclim_flint_grwseason_avgs_2024, all_bioclim_flint_avgs_2024.pc.dat)

all_bioclim_flint_avgs_2024_loadings = data.frame(varnames=rownames(all_bioclim_flint_avgs_2024.pc$rotation), all_bioclim_flint_avgs_2024.pc$rotation)
all_bioclim_flint_avgs_2024_loadings #pretty similar to 2023, PC4 a bit different though
```

```{r}
autoplot(all_bioclim_flint_avgs_2024.pc, data = bioclim_flint_grwseason_avgs_2024,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
autoplot(all_bioclim_flint_avgs_2024.pc, data = bioclim_flint_grwseason_avgs_2024,
         x=2, y=4,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
all_bioclim_flint_avgs_2024_locs.pc_avg <- all_bioclim_flint_avgs_2024_locs.pc %>%
  group_by(parent.pop, elev_m, TimePd) %>%
  summarise(across(.cols=starts_with("PC"), .fns = mean)) %>%
  ungroup()

#all_bioclim_flint_avgs_2024_locs.pc_avg
```

```{r}
all_bioclim_flint_avgs_2024_locs.pc_avg %>% 
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point() +
  facet_wrap(~TimePd) +
  coord_fixed(ratio = 1.5)
```

## 2024 - Permanova
See Jenny's github: https://github.com/jrgremer/Seedbanks_drought_deluge  
```{r}
all_bioclim_flint_avgs_2024_locs.pc_dist <- all_bioclim_flint_avgs_2024_locs.pc %>% ungroup() %>% select(PC1:PC7)
dist_matrix_grwssn_2024 <- dist(all_bioclim_flint_avgs_2024_locs.pc_dist, method = "euclidian") #use a distance function to calculate euclidian distance in PCA space
permanova_results_grwssn_2024 <- adonis2(dist_matrix_grwssn_2024 ~ TimePd*elev_m*Lat, data = all_bioclim_flint_avgs_2024_locs.pc) #use adonis2 to run the permanova
#rows in data must be in same order as rows in distance matrix 
#by = NULL will assess the overall significance of all terms together

#look at output 
summary(permanova_results_grwssn_2024)
permanova_results_grwssn_2024
permanova_results_grwssn_2024$`Pr(>F)`

permanova_results_grwssn_2024_terms <- adonis2(dist_matrix_grwssn_2024 ~ TimePd*elev_m*Lat, data = all_bioclim_flint_avgs_2024_locs.pc, by = "terms") #by = "terms" will assess significance for each term (sequentially from first to last)
permanova_results_grwssn_2024_terms #no significant time period effect!! 
```

## 2024 - LM on PCs (follow up on permanova)

For each PC fit a model with timepd and parent.pop/elev to test for climate change

```{r}
lmer_results_grwssn_2024 <- all_bioclim_flint_avgs_2024_locs.pc %>%
  ungroup() %>% 
  select(TimePd, parent.pop, elev_m, Lat, Long, PC1:PC10) %>% 
  pivot_longer(starts_with("PC", ignore.case = FALSE), 
               names_to = "PC", values_to = "value") %>% 
  group_by(PC) %>% 
  nest(data=c(TimePd, parent.pop, elev_m, Lat, Long, value)) %>% 
  mutate(glm=map(data, ~ glm(value ~ TimePd*elev_m*Lat,
                               data=.x)),
         anova = map(glm, ~ broom.mixed::tidy(anova(.x))))

#lmer_results_grwssn_2024
```

```{r}
PC_anova_2024 <- lmer_results_grwssn_2024 %>% select(-data, -glm) %>% unnest(anova) %>%
  select(PC, term, p.value) %>%
  filter(p.value < 0.05) %>%
  arrange(term, p.value)

PC_anova_2024 #no main timeperiod effects
#PC7 has a timepd*lat and timepd*elev effect

lmer_results_grwssn_2024 %>% select(-data, -glm) %>% unnest(anova) %>%
  filter(PC=="PC2"| PC=="PC4") #the two PCs with most sig p-values for timepd

#mod_test <- glm(PC4 ~ TimePd*elev_m*Lat, data=all_bioclim_flint_avgs_2024_locs.pc)
#summary(mod_test)
#anova(mod_test)
```

```{r}
autoplot(all_bioclim_flint_avgs_2024.pc, data = bioclim_flint_grwseason_avgs_2024,
         x=2, y=4,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()

all_bioclim_flint_avgs_2024_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC2, y=PC4, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)
#high elev climate seems to be shifting to be more similar to low elev 

home_sites_pca_2024 <- all_bioclim_flint_avgs_2024_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC2, y=PC4, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  #geom_point(size=2, alpha=0.7) +
  labs(x="PC2 (27.64%)", y="PC4 (8.91%)", color="Elevation (m)") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8) +
  annotate("text", x = -4.2, y = -2.5, label = "Low Temp \n Seasonality") +
  annotate("text", x = 3.6, y = -2.5, label = "High Temp \n Seasonality") +
  annotate("text", x = -5.28, y = -1.5, label = "Low Diurnal \n Range") +
  annotate("text", x = -5.31, y = 2.5, label = "High Diurnal \n Range") +
  coord_cartesian(ylim = c(-3, 3), xlim = c(-4, 3.5), clip = "off") +
  theme_classic() +
  theme(text=element_text(size=28))
home_sites_pca_2024
#ggsave("../output/Climate/GRWSSN_PC2-PC4_2024.png", width = 7, height = 5.5, units = "in")
```

## 2024 Add the Garden to the Final PCA 
```{r}
WL2Grdn_pc_prep_2024 <- WL2Grdn_flint_bioclim_2024_grwseason %>% select(cwd, ppt, ann_tmean, mean_diurnal_range, temp_seasonality, temp_ann_range, tmean_wettest_month, ann_ppt, ppt_seasonality, ppt_warmest_month, ppt_coldest_month)
#scale(WL2Grdn_pc_prep_2024, all_bioclim_flint_avgs_2024.pc$center, all_bioclim_flint_avgs_2024.pc$scale) %*% all_bioclim_flint_avgs_2024.pc$rotation  #gives same result as below 
WL2Grdn_predicted_2024 <- predict(all_bioclim_flint_avgs_2024.pc, newdata = WL2Grdn_pc_prep_2024)
WL2Grdn_predicted_2024

#str(home_sites_pca_2024) #can add predicted WL2 point to the existing plot's data 
home_sites_pca_2024$data <- rbind(home_sites_pca_2024$data, 
  data.frame(
    parent.pop = "WL2_Garden",
    elev_m = 2020,
    TimePd = "2024",
    PC1 = WL2Grdn_predicted_2024[, "PC1"],
    PC2 = WL2Grdn_predicted_2024[, "PC2"],
    PC3 = WL2Grdn_predicted_2024[, "PC3"],
    PC4 = WL2Grdn_predicted_2024[, "PC4"],
    PC5 = WL2Grdn_predicted_2024[, "PC5"],
    PC6 = WL2Grdn_predicted_2024[, "PC6"],
    PC7 = WL2Grdn_predicted_2024[, "PC7"],
    PC8 = WL2Grdn_predicted_2024[, "PC8"],
    PC9 = WL2Grdn_predicted_2024[, "PC9"],
    PC10 = WL2Grdn_predicted_2024[, "PC10"],
    PC11 = WL2Grdn_predicted_2024[, "PC11"],
    group = "new"
  )
)

home_sites_pca_2024 + 
  geom_point(data=filter(home_sites_pca_2024$data, parent.pop == "WL2_Garden"), size=3, shape = 8, show.legend = FALSE)
#ggsave("../output/Climate/GRWSSN_PC2-PC4_PlusGarden_2024.png", width = 7.4, height = 6, units = "in")
```

---

## 2023 Old PCAs

### All years and months included (Flint)

#### Correlations - Flint Recent

```{r}
#normalize the data
climate_normalized_all_flint_recent <- flint_recent_grwseason %>% select(cwd, ppt, tmn, tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_flint_recent)
cor.norm = cor(climate_normalized_all_flint_recent) #test correlations among the traits
corrplot(cor.norm)
#tmn and tmx highly correlated (95%), consider removing one 
```

#### PCA - Flint Recent

```{r}
all_flint_recent.pc = prcomp(flint_recent_grwseason[c(9, 11:12)], scale = TRUE, center = TRUE)

str(all_flint_recent.pc)
```

plot % Variance Explained

```{r}
summary(all_flint_recent.pc)

tibble(PC=str_c("PC",str_pad(1:3,2,pad="0")),
       percent_var=all_flint_recent.pc$sdev[1:3]^2/sum(all_flint_recent.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
all_flint_recent.pc.dat = data.frame(all_flint_recent.pc$x)

all_flint_recent_locs.pc = cbind(flint_recent_grwseason, all_flint_recent.pc.dat)

all_flint_recent_loadings = data.frame(varnames=rownames(all_flint_recent.pc$rotation), all_flint_recent.pc$rotation)
all_flint_recent_loadings
```

```{r}
autoplot(all_flint_recent.pc, data = flint_recent_grwseason,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=8, loadings.label.colour="black", loadings.label.vjust = -0.2) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()

#for plot customizations see: ?ggbiplot
```

PCs 1 and 3

```{r}
autoplot(all_flint_recent.pc, data = flint_recent_grwseason,
         x=1, y=3,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=8, loadings.label.colour="black", loadings.label.vjust = -0.2) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

#### Correlations - Flint Historical

```{r}
#normalize the data
climate_normalized_all_flint_historical <- flint_historical_grwseason %>% select(cwd, ppt, tmn, tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_flint_historical)
cor.norm = cor(climate_normalized_all_flint_historical) #test correlations among the traits
corrplot(cor.norm)
#tmn and tmx highly correlated, consider removing one 
```

#### PCA - Flint Historical

```{r}
all_flint_historical.pc = prcomp(flint_historical_grwseason[c(9, 11:12)], scale = TRUE, center = TRUE)

str(all_flint_historical.pc)
```

plot % Variance Explained

```{r}
summary(all_flint_historical.pc)

tibble(PC=str_c("PC",str_pad(1:3,2,pad="0")),
       percent_var=all_flint_historical.pc$sdev[1:3]^2/sum(all_flint_historical.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
all_flint_historical.pc.dat = data.frame(all_flint_historical.pc$x)

all_flint_historical_locs.pc = cbind(flint_historical_grwseason, all_flint_historical.pc.dat)

all_flint_historical_loadings = data.frame(varnames=rownames(all_flint_historical.pc$rotation), all_flint_historical.pc$rotation)
all_flint_historical_loadings
```

```{r}
autoplot(all_flint_historical.pc, data = flint_historical_grwseason,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=8, loadings.label.colour="black", loadings.label.vjust = -0.2) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

PCs 1 and 3

```{r}
autoplot(all_flint_historical.pc, data = flint_historical_grwseason,
         x=1, y=3,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=8, loadings.label.colour="black", loadings.label.vjust = -0.2) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

### All years included (bioclim)

#### Correlations - bioclim Recent

```{r}
#normalize the data
climate_normalized_all_bioclim_recent <- bioclim_recent_grwseason %>% select(ann_tmean:ppt_coldest_month) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_bioclim_recent)
cor.norm = cor(climate_normalized_all_bioclim_recent) #test correlations among the traits
corrplot(cor.norm)
```

#### PCA - bioclim Recent

```{r}
all_bioclim_recent.pc = prcomp(bioclim_recent_grwseason[c(5:14)], scale = TRUE, center = TRUE) 

str(all_bioclim_recent.pc)
```

plot % Variance Explained

```{r}
summary(all_bioclim_recent.pc)

tibble(PC=str_c("PC",str_pad(1:10,2,pad="0")),
       percent_var=all_bioclim_recent.pc$sdev[1:10]^2/sum(all_bioclim_recent.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
all_bioclim_recent.pc.dat = data.frame(all_bioclim_recent.pc$x)

all_bioclim_recent_locs.pc = cbind(bioclim_recent_grwseason, all_bioclim_recent.pc.dat)

all_bioclim_recent_loadings = data.frame(varnames=rownames(all_bioclim_recent.pc$rotation), all_bioclim_recent.pc$rotation)
all_bioclim_recent_loadings
```

```{r}
autoplot(all_bioclim_recent.pc, data = bioclim_recent_grwseason,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

PCs 3 and 4

```{r}
autoplot(all_bioclim_recent.pc, data = bioclim_recent_grwseason,
         x=3, y=4,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

PCs 4 and 5

```{r}
autoplot(all_bioclim_recent.pc, data = bioclim_recent_grwseason,
         x=4, y=5,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

#### Correlations - bioclim Historical

```{r}
#normalize the data
climate_normalized_all_bioclim_historical <- bioclim_historical_grwseason %>% select(ann_tmean:ppt_coldest_month) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_bioclim_historical)
cor.norm = cor(climate_normalized_all_bioclim_historical) #test correlations among the traits
corrplot(cor.norm)
```

#### PCA - bioclim Historical

```{r}
all_bioclim_historical.pc = prcomp(bioclim_historical_grwseason[c(5:14)], scale = TRUE, center = TRUE) 

str(all_bioclim_historical.pc)
```

plot % Variance Explained

```{r}
summary(all_bioclim_historical.pc)

tibble(PC=str_c("PC",str_pad(1:10,2,pad="0")),
       percent_var=all_bioclim_historical.pc$sdev[1:10]^2/sum(all_bioclim_historical.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
all_bioclim_historical.pc.dat = data.frame(all_bioclim_historical.pc$x)

all_bioclim_historical_locs.pc = cbind(bioclim_historical_grwseason, all_bioclim_historical.pc.dat)

all_bioclim_historical_loadings = data.frame(varnames=rownames(all_bioclim_historical.pc$rotation), all_bioclim_historical.pc$rotation)
all_bioclim_historical_loadings
```

```{r}
autoplot(all_bioclim_historical.pc, data = bioclim_historical_grwseason,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

PCs 3 and 4

```{r}
autoplot(all_bioclim_historical.pc, data = bioclim_historical_grwseason,
         x=3, y=4,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

PCs 4 and 5

```{r}
autoplot(all_bioclim_historical.pc, data = bioclim_historical_grwseason,
         x=4, y=5,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

### Monthly Averages - Flint

#### Calculate avgs

```{r}
flint_recent_grwseason_mosavgs <- flint_recent_grwseason %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, month) %>% 
  summarise_at(c("cwd", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Recent") 
flint_recent_grwseason_mosavgs

flint_historical_grwseason_mosavgs <- flint_historical_grwseason %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, month) %>% 
  summarise_at(c("cwd", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Historical")
flint_historical_grwseason_mosavgs

flint_grwseason_mosavgs <- bind_rows(flint_recent_grwseason_mosavgs, flint_historical_grwseason_mosavgs) #combine into 1 dataframe 
head(flint_grwseason_mosavgs)
tail(flint_grwseason_mosavgs)
```

#### Correlations - Recent + Historical

```{r}
#normalize the data
climate_normalized_flint_grwseason_mosavgs <- flint_grwseason_mosavgs %>% ungroup() %>% 
  select(cwd:tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_flint_grwseason_mosavgs)
cor.norm = cor(climate_normalized_flint_grwseason_mosavgs) #test correlations among the traits
corrplot(cor.norm)

#tmn and tmx highly correlated, consider removing one (96%)
#tmx and ppt highly neg correlated (-82%)
```

#### PCA - Recent + Historical

```{r}
#flint_grwseason_mosavgs[c(8:12)]
mos_flint.pc = prcomp(flint_grwseason_mosavgs[c(7:9)], scale = TRUE, center = TRUE)

str(mos_flint.pc)
```

plot % Variance Explained

```{r}
summary(mos_flint.pc)

tibble(PC=str_c("PC",str_pad(1:3,2,pad="0")),
       percent_var=mos_flint.pc$sdev[1:3]^2/sum(mos_flint.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
mos_flint.pc.dat = data.frame(mos_flint.pc$x)

mos_flint_locs.pc = cbind(flint_grwseason_mosavgs, mos_flint.pc.dat)

mos_flint_loadings = data.frame(varnames=rownames(mos_flint.pc$rotation), mos_flint.pc$rotation)
mos_flint_loadings
```

```{r}
autoplot(mos_flint.pc, data = flint_grwseason_mosavgs,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="month",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
#high elev seems most similar to low elev in summer months 
```

```{r}
mos_flint_locs.pc_avg <- mos_flint_locs.pc %>%
  group_by(parent.pop, elev_m, TimePd,month) %>%
  summarise(across(.cols=starts_with("PC"), .fns = mean)) %>%
  ungroup()

mos_flint_locs.pc_avg
```

```{r}
mos_flint_locs.pc_avg %>% 
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point() +
  facet_wrap(~month) +
  coord_fixed(ratio = 1.5)
#Remember: this is growth season so high elev pops don't have winter months 

```

```{r}
mos_flint_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m, month))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point() +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")))
#high elev climate seems to be shifting to be more similar to low elev 
#hard to see which months though 
```

### Avg across years and months (Flint)

#### Correlations - Recent + Historical

```{r}
#normalize the data
climate_normalized_flint_grwseason_avgs <- flint_grwseason_avgs %>% ungroup() %>% 
  select(cwd:tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_flint_grwseason_avgs)
cor.norm = cor(climate_normalized_flint_grwseason_avgs) #test correlations among the traits
corrplot(cor.norm)

#tmn and tmx highly correlated, consider removing one (90%)
```

#### PCA - Recent + Historical

```{r}
avgs_flint.pc = prcomp(flint_grwseason_avgs[c(6:8)], scale = TRUE, center = TRUE)

str(avgs_flint.pc)
```

plot % Variance Explained

```{r}
summary(avgs_flint.pc)

tibble(PC=str_c("PC",str_pad(1:3,2,pad="0")),
       percent_var=avgs_flint.pc$sdev[1:3]^2/sum(avgs_flint.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
avgs_flint.pc.dat = data.frame(avgs_flint.pc$x)

avgs_flint_locs.pc = cbind(flint_grwseason_avgs, avgs_flint.pc.dat)

avgs_flint_loadings = data.frame(varnames=rownames(avgs_flint.pc$rotation), avgs_flint.pc$rotation)
avgs_flint_loadings
```

```{r}
autoplot(avgs_flint.pc, data = flint_grwseason_avgs,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
avgs_flint_locs.pc_avg <- avgs_flint_locs.pc %>%
  group_by(parent.pop, elev_m, TimePd) %>%
  summarise(across(.cols=starts_with("PC"), .fns = mean)) %>%
  ungroup()

avgs_flint_locs.pc_avg
```

```{r}
avgs_flint_locs.pc_avg %>% 
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point() +
  facet_wrap(~TimePd) +
  coord_fixed(ratio = 1.5)
```

```{r}
avgs_flint_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)

avgs_flint_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)
#growth season climate has shifted for high elev way more than for low elev & directions are varied 
```
