---
title: "Climate_PCAs_AllYear"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To Do:

-   Compare historical 60 years to recent 30 years as another climate change test

-   Check the significance of the PCA

    -   Remember this paper: Björklund, M. 2019. Be careful with your principal components. Evolution 73: 2151--2158.

# Climate PCAs for Full "Water" Year

## Load necessary libraries

```{r}
library(tidyverse)
library(lmerTest)
library(ggrepel)
#library(cowplot)
library(gridExtra)
library(corrplot) #plotting correlations 
library(QBMS) #for function calc_biovars to calculate bioclim variables
library(ggfortify) #easier PCA figures
library(vegan) #for permanova 
sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
} #legend function for grid_arrange

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")
month_order <- c("jan","feb","mar","apr","may","jun","jul","aug","sep","oct","nov","dec")
```

## Load Flint Data

```{r}
flint_all_year <- read_csv("../output/Climate/flint_climate_UCDpops.csv")
head(flint_all_year)
```

### Add month numbers

```{r}
month_nums <- tibble(month=c("jan","feb","mar","apr","may","jun","jul","aug","sep","oct","nov","dec"),
                     month_nums=c(1:12))
month_nums

flint_all_year_mosnums <- left_join(flint_all_year, month_nums) %>% arrange(parent.pop, year, month_nums)
tail(flint_all_year_mosnums, 12) 
```

### Water Year

```{r}
flint_all_year_wtr_yr <- flint_all_year_mosnums %>% mutate(wtr_yr = year + month_nums %in% c(11:12))
```

## Generate bioclim for all year - 2023

-   annual mean temperature (BIO1)
-   mean diurnal range (BIO2) - (Mean of monthly (max temp - min temp))
-   temperature seasonality (BIO4) (standard deviation \*100)
-   temperature annual range (BIO7) (Max Temperature of Warmest Month - Min Temperature of Coldest Month)
-   mean temp of wettest quarter (BIO8)
-   mean temp of driest quarter (BIO9)
-   annual precipitation (BIO12) - sum of ppt for the entire year (not the avg)
-   precipitation seasonality (BIO15) (Coefficient of Variation)
-   precip of warmest quarter (BIO18)
-   precip of coldest quarter (BIO19)

### Prep

```{r}
bioclim_allyear_prep <- flint_all_year_wtr_yr %>% 
  filter(parent.pop != "UCD_Garden", parent.pop != "WL2_Garden") %>%  #remove garden sites 
  rename(tmin=tmn, tmax=tmx, year_cal=year, year=wtr_yr) %>% #rename columns to match what calc_biovars expects, also make sure it uses water year 
  filter(year != "1895", year < "2024") %>%  #remove years with less than 12 months of data
  arrange(parent.pop, year, month)

bioclim_all_year <- tibble(bio1=NA, bio2=NA, bio4=NA, bio7=NA, bio8=NA, bio9=NA, bio12=NA, bio15=NA, bio18=NA, bio19=NA, year=2025) #blank tibble to bind calculations to
bioclim_all_year
 
popids <- unique(bioclim_allyear_prep$parent.pop) #list of pop ids for for loop 

pop_elev <- flint_all_year_wtr_yr %>% select(parent.pop:Long) %>% distinct()
```

### Calculation

```{r}
for(i in popids) {
  A <- bioclim_allyear_prep %>% filter(parent.pop==i) %>% calc_biovars() %>% mutate(parent.pop=i)
  #print(A)
  bioclim_all_year <- bind_rows(bioclim_all_year, A)
}
unique(bioclim_all_year$parent.pop) #has all the populations in there!

unique(bioclim_allyear_prep$year)
```

### SUBSET

```{r}
bioclim_all_year_final <- bioclim_all_year %>% 
  select(parent.pop, year, ann_tmean=bio1, mean_diurnal_range=bio2, 
         temp_seasonality=bio4, temp_ann_range=bio7, tmean_wettest_quarter=bio8,
         tmean_driest_quarter=bio9, ann_ppt=bio12, ppt_seasonality=bio15,
         ppt_warmest_quarter=bio18, ppt_coldest_quarter=bio19) %>%
  filter(year!=2025)
head(bioclim_all_year_final)
```

### Merge with pop info

```{r}
pop_elev_bioclim_all_year <- left_join(bioclim_all_year_final, pop_elev) %>% 
  select(parent.pop, elevation.group:Long, year:ppt_coldest_quarter)
head(pop_elev_bioclim_all_year)
#write_csv(pop_elev_bioclim_all_year, "../output/Climate/bioclim_all_year_UCD_pops_wtr_year.csv")
```

## Calculation of recent (last 30 years) and historical climate (prior 30 years) - 2023

```{r}
flint_all_year_recent <- flint_all_year_wtr_yr %>% 
  filter(parent.pop != "UCD_Garden", parent.pop != "WL2_Garden") %>%  #remove garden sites 
  filter(wtr_yr>1993 & wtr_yr<=2023) %>% 
  select(parent.pop:month, wtr_yr, cwd, pck, ppt, tmn, tmx)
head(flint_all_year_recent)
tail(flint_all_year_recent)

flint_all_year_historical <- flint_all_year_wtr_yr %>% 
  filter(parent.pop != "UCD_Garden", parent.pop != "WL2_Garden") %>%  #remove garden sites 
  filter(wtr_yr<=1993 & wtr_yr>1963) %>% 
  select(parent.pop:month, wtr_yr, cwd, pck, ppt, tmn, tmx)
head(flint_all_year_historical, 13)
tail(flint_all_year_historical, 13)

bioclim_all_year_recent <- pop_elev_bioclim_all_year %>% filter(year>1993 & year<=2023) #year here means water year (see above code where bioclim vars were calculated)
#head(bioclim_all_year_recent)
#tail(bioclim_all_year_recent)

bioclim_all_year_historical <- pop_elev_bioclim_all_year %>% filter(year<=1993 & year>1963)
#head(bioclim_all_year_historical, 13)
#tail(bioclim_all_year_historical, 13)
```

### Yearly averages for Flint (for climate distance calc)
```{r}
flint_all_year_recent_yravgs <- flint_all_year_recent %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, wtr_yr) %>% 
  summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  rename(year=wtr_yr)
flint_all_year_recent_yravgs

flint_all_year_historical_yravgs <- flint_all_year_historical %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, wtr_yr) %>% 
  summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  rename(year=wtr_yr)
flint_all_year_historical_yravgs
```

### Merge with bioclim
```{r}
bioclim_flint_yrly_avgs_recent <- full_join(flint_all_year_recent_yravgs, bioclim_all_year_recent) 
#write_csv(bioclim_flint_yrly_avgs_recent, "../output/Climate/fullyear_wtr_year_avgs_Recent.csv")
bioclim_flint_yrly_avgs_historical <- full_join(flint_all_year_historical_yravgs, bioclim_all_year_historical) 
#write_csv(bioclim_flint_yrly_avgs_historical, "../output/Climate/fullyear_wtr_year_avgs_Historical.csv")
```

## Calculation of extra historical climate (prior 60 years) - 2023
```{r}
flint_all_year_historical_extra <- flint_all_year_wtr_yr %>% 
  filter(parent.pop != "UCD_Garden", parent.pop != "WL2_Garden") %>%  #remove garden sites 
  filter(wtr_yr<=1993 & wtr_yr>1933) %>% 
  select(parent.pop:month, wtr_yr, cwd, pck, ppt, tmn, tmx)
head(flint_all_year_historical_extra, 13)
tail(flint_all_year_historical_extra, 13)

bioclim_all_year_historical_extra <- pop_elev_bioclim_all_year %>% filter(year<=1993 & year>1933)
#head(bioclim_all_year_historical_extra, 13)
#tail(bioclim_all_year_historical_extra, 13)
```


## WL2 Garden 2023
```{r}
bioclim_allyear_prep_WL2Grdn <- flint_all_year_wtr_yr %>% 
  filter(parent.pop == "WL2_Garden") %>%  #only keep WL2 garden
  rename(tmin=tmn, tmax=tmx, year_cal=year, year=wtr_yr) %>% #rename columns to match what calc_biovars expects, also make sure it uses water year 
  filter(year=="2023") %>% #year of the experiment only 
  arrange(parent.pop, year, month)

bioclim_all_year_WL2Grdn <- bioclim_allyear_prep_WL2Grdn %>% 
  calc_biovars() %>% 
  mutate(parent.pop="WL2_Garden", year=="2023") 

bioclim_all_year_final_WL2Grdn <- bioclim_all_year_WL2Grdn %>% 
  select(parent.pop, year, ann_tmean=bio1, mean_diurnal_range=bio2, 
         temp_seasonality=bio4, temp_ann_range=bio7, tmean_wettest_quarter=bio8,
         tmean_driest_quarter=bio9, ann_ppt=bio12, ppt_seasonality=bio15,
         ppt_warmest_quarter=bio18, ppt_coldest_quarter=bio19) 

WL2Grdn_elev_bioclim_all_year <- left_join(bioclim_all_year_final_WL2Grdn, pop_elev) %>% 
  select(parent.pop, elevation.group:Long, year:ppt_coldest_quarter)
WL2Grdn_elev_bioclim_all_year

WL2Grdn_flint_all_year <- flint_all_year_wtr_yr %>% 
  filter(parent.pop == "WL2_Garden", wtr_yr=="2023") %>%  #only keep WL2 garden
  select(parent.pop:month, wtr_yr, cwd, pck, ppt, tmn, tmx)

WL2Grdn_flint_all_year_summary <- WL2Grdn_flint_all_year %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) 

WL2Grdn_flint_bioclim_all_year <- left_join(WL2Grdn_flint_all_year_summary, WL2Grdn_elev_bioclim_all_year) %>% 
  select(-year)
```

## Avg across years and months (Flint + bioclim) - 2023

### Calculate avgs for flint

```{r}
flint_all_year_recent_avgs <- flint_all_year_recent %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Recent") 
#flint_all_year_recent_avgs

flint_all_year_historical_avgs <- flint_all_year_historical %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Historical")
#flint_all_year_historical_avgs

flint_all_year_avgs <- bind_rows(flint_all_year_recent_avgs, flint_all_year_historical_avgs) #combine into 1 dataframe 
head(flint_all_year_avgs)
tail(flint_all_year_avgs)
#write_csv(flint_all_year_avgs, "../output/Climate/fullyear_FlintAvgs_wtr_year.csv")
```

### Calculate avgs for bioclim 

```{r}
bioclim_all_year_recent_avgs <- bioclim_all_year_recent %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_quarter", "tmean_driest_quarter", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_quarter", "ppt_coldest_quarter"),
               c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Recent") 
#bioclim_all_year_recent_avgs

bioclim_all_year_historical_avgs <- bioclim_all_year_historical %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_quarter", "tmean_driest_quarter", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_quarter", "ppt_coldest_quarter"),
               c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Historical") 
#bioclim_all_year_historical_avgs

bioclim_all_year_avgs <- bind_rows(bioclim_all_year_recent_avgs, bioclim_all_year_historical_avgs) #combine into 1 dataframe 
head(bioclim_all_year_avgs)
tail(bioclim_all_year_avgs)
#write_csv(bioclim_all_year_avgs, "../output/Climate/fullyear_BioClimAvgs_wtr_year.csv")
```

Merge with flint

```{r}
bioclim_flint_all_year_avgs <- full_join(flint_all_year_avgs, bioclim_all_year_avgs) %>% 
  select(TimePd, parent.pop:ppt_coldest_quarter)
head(bioclim_flint_all_year_avgs)
```

### Correlations - Recent + Historical

```{r}
#normalize the data
climate_normalized_bioclim_flint_all_year_avgs <- bioclim_flint_all_year_avgs %>% ungroup() %>% 
  select(cwd:ppt_coldest_quarter) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_bioclim_flint_all_year_avgs)
dim(climate_normalized_bioclim_flint_all_year_avgs)
cor.norm = cor(climate_normalized_bioclim_flint_all_year_avgs) #test correlations among the traits
cor.sig <- cor.mtest(climate_normalized_bioclim_flint_all_year_avgs, method = "pearson")

corrplot(cor.norm, type="upper",
         tl.srt = 45, p.mat = cor.sig$p, 
         sig.level = 0.05, insig="blank")

#tmn, tmx, tmean_wettest_quarter, tmean_driest_quarter and ann_tmean all highly correlated (97-99%) - only keep ann_tmean 
#temp_ann_range and mean_diurnal range 92% correlated - keep temp_ann_range 
#ann_ppt and ppt 100% correlated (ppt = avg across monts, ann_ppt = avg of the total ppt in a year) - only keep ann_ppt 
#ppt_warmest and coldest quarter highly correlated with ann_ppt and ppt - take both out 
```

Recent
```{r}
climate_normalized_bioclim_flint_all_year_avgs_recent <- bioclim_flint_all_year_avgs %>% 
  ungroup() %>% 
  filter(TimePd=="Recent") %>% 
  select(tmn:tmean_driest_quarter, cwd, pck, ppt, ann_ppt:ppt_coldest_quarter) %>% 
  scale() #normalize the data so they're all on the same scale

cor.norm_recent = cor(climate_normalized_bioclim_flint_all_year_avgs_recent) #test correlations among the traits
cor.sig_recent <- cor.mtest(climate_normalized_bioclim_flint_all_year_avgs_recent, method = "pearson")

corrplot(cor.norm_recent, type="upper",
         tl.srt = 45, p.mat = cor.sig_recent$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```

Historical
```{r}
climate_normalized_bioclim_flint_all_year_avgs_historic <- bioclim_flint_all_year_avgs %>% 
  ungroup() %>% 
  filter(TimePd=="Historical") %>% 
  select(tmn:tmean_driest_quarter, cwd, pck, ppt, ann_ppt:ppt_coldest_quarter) %>% 
  scale() #normalize the data so they're all on the same scale

cor.norm_historic = cor(climate_normalized_bioclim_flint_all_year_avgs_historic) #test correlations among the traits
cor.sig_historic <- cor.mtest(climate_normalized_bioclim_flint_all_year_avgs_historic, method = "pearson")

corrplot(cor.norm_historic, type="upper",
         tl.srt = 45, p.mat = cor.sig_historic$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```


### PCA - Recent + Historical

```{r}
all_bioclim_flint_avgs.pc = prcomp(bioclim_flint_all_year_avgs[c(7:8, 12, 14:15, 18:19)], scale = TRUE, center = TRUE)
str(all_bioclim_flint_avgs.pc)
```

plot % Variance Explained

```{r}
summary(all_bioclim_flint_avgs.pc)

tibble(PC=str_c("PC",str_pad(1:7,2,pad="0")),
       percent_var=all_bioclim_flint_avgs.pc$sdev[1:7]^2/sum(all_bioclim_flint_avgs.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
all_bioclim_flint_avgs.pc.dat = data.frame(all_bioclim_flint_avgs.pc$x)

all_bioclim_flint_avgs_locs.pc = cbind(bioclim_flint_all_year_avgs, all_bioclim_flint_avgs.pc.dat)

all_bioclim_flint_avgs_loadings = data.frame(varnames=rownames(all_bioclim_flint_avgs.pc$rotation), all_bioclim_flint_avgs.pc$rotation)
all_bioclim_flint_avgs_loadings
```

```{r}
autoplot(all_bioclim_flint_avgs.pc, data = bioclim_flint_all_year_avgs,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
autoplot(all_bioclim_flint_avgs.pc, data = bioclim_flint_all_year_avgs,
         x=1, y=4,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
all_bioclim_flint_avgs_locs.pc_avg <- all_bioclim_flint_avgs_locs.pc %>%
  group_by(parent.pop, elev_m, TimePd) %>%
  summarise(across(.cols=starts_with("PC", ignore.case = FALSE), .fns = mean)) %>%
  ungroup()
all_bioclim_flint_avgs.pc
```

```{r}
all_bioclim_flint_avgs_locs.pc_avg %>% 
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point() +
  facet_wrap(~TimePd) +
  coord_fixed(ratio = 1.5)
```

## Permanova - 2023
See Jenny's github: https://github.com/jrgremer/Seedbanks_drought_deluge  
```{r}
all_bioclim_flint_avgs_locs.pc_dist <- all_bioclim_flint_avgs_locs.pc %>% ungroup() %>% select(PC1:PC7)
dist_matrix_all_year <- dist(all_bioclim_flint_avgs_locs.pc_dist, method = "euclidian") #use a distance function to calculate euclidian distance in PCA space
permanova_results_all_year <- adonis2(dist_matrix_all_year ~ TimePd*elev_m*Lat, data = all_bioclim_flint_avgs_locs.pc) #use adonis2 to run the permanova
#rows in data must be in same order as rows in distance matrix 

#look at output 
summary(permanova_results_all_year)
permanova_results_all_year
permanova_results_all_year$`Pr(>F)`

permanova_results_all_year_terms <- adonis2(dist_matrix_all_year ~ TimePd*elev_m*Lat, data = all_bioclim_flint_avgs_locs.pc, by = "terms")
permanova_results_all_year_terms
```

## LM on PCs (follow up on permanova) - 2023

For each PC fit a model with timepd and parent.pop/elev to test for climate change

```{r}
lmer_results_all_year <- all_bioclim_flint_avgs_locs.pc %>%
  ungroup() %>% 
  select(TimePd, parent.pop, elev_m, Lat, Long, PC1:PC7) %>% 
  pivot_longer(starts_with("PC", ignore.case = FALSE), 
               names_to = "PC", values_to = "value") %>% 
  group_by(PC) %>% 
  nest(data=c(TimePd, parent.pop, elev_m, Lat, Long, value)) %>% 
  mutate(glm=map(data, ~ glm(value ~ TimePd*elev_m*Lat,
                               data=.x)),
         anova = map(glm, ~ broom.mixed::tidy(anova(.x))))

#lmer_results_all_year
```

```{r}
PC_anova <- lmer_results_all_year %>% select(-data, -glm) %>% unnest(anova) %>%
  select(PC, term, p.value) %>%
  filter(p.value < 0.05) %>%
  arrange(term, p.value)

PC_anova

lmer_results_all_year %>% select(-data, -glm) %>% unnest(anova) %>% 
  filter(PC=="PC1"| PC=="PC4") #the two PCs with most sig p-values for timepd

#mod_test <- glm(PC4 ~ TimePd*elev_m*Lat, data=all_bioclim_flint_avgs_locs.pc)
#summary(mod_test)
#anova(mod_test)
```

```{r}
autoplot(all_bioclim_flint_avgs.pc, data = bioclim_flint_all_year_avgs,
         x=1, y=4,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()

all_bioclim_flint_avgs_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC4, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)

all_bioclim_flint_avgs_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC4, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)
#high elev climate seems to be shifting to be more similar to low elev 
```

```{r}
home_sites_pca <- all_bioclim_flint_avgs_locs.pc_avg %>%  
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC4, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  #geom_point(size=2, alpha=0.7) +
  labs(x="PC1 (51.11%)", y="PC4 (8.07%)", color="Elevation (m)") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8) +
  annotate("text", x = -4, y = -5.1, label = "Warm \n No Snow") +
  annotate("text", x = 4, y = -5.1, label = "Cold \n Snow") +
  annotate("text", x = -5.4, y = -3.5, label = "High Temp \n Seasonality") +
  annotate("text", x = -5.4, y = 1.5, label = "Low Temp \n Seasonality") +
  coord_cartesian(ylim = c(-4.1, 2), xlim = c(-4,4), clip = "off") +
  theme_classic() +
  theme(text=element_text(size=28))
#ggsave("../output/Climate/Wtr_Year_PC1-PC4.png", width = 7.4, height = 6, units = "in")

## add WL2 garden 2023
WL2Grdn_pc_prep <- WL2Grdn_flint_bioclim_all_year %>% ungroup() %>% select(cwd:pck, ann_tmean, temp_seasonality, temp_ann_range, ann_ppt, ppt_seasonality)
#scale(WL2Grdn_pc_prep, all_bioclim_flint_avgs.pc$center, all_bioclim_flint_avgs.pc$scale) %*% all_bioclim_flint_avgs.pc$rotation  #gives same result as below 
WL2Grdn_predicted <- predict(all_bioclim_flint_avgs.pc, newdata = WL2Grdn_pc_prep)
WL2Grdn_predicted

#str(home_sites_pca) #can add predicted WL2 point to the existing plot's data 
home_sites_pca$data <- rbind(home_sites_pca$data, 
  data.frame(
    parent.pop = "WL2_Garden",
    elev_m = 2020,
    TimePd = "2023",
    PC1 = WL2Grdn_predicted[, "PC1"],
    PC2 = WL2Grdn_predicted[, "PC2"],
    PC3 = WL2Grdn_predicted[, "PC3"],
    PC4 = WL2Grdn_predicted[, "PC4"],
    PC5 = WL2Grdn_predicted[, "PC5"],
    PC6 = WL2Grdn_predicted[, "PC6"],
    PC7 = WL2Grdn_predicted[, "PC7"],
    group = "new"
  )
)

home_sites_pca + 
  geom_point(data=filter(home_sites_pca$data, parent.pop == "WL2_Garden"), size=3, shape = 8, show_guide = FALSE)
#ggsave("../output/Climate/Wtr_Year_PC1-PC4_PlusGarden.png", width = 7.4, height = 6, units = "in")
```

## Extra historical PCA (60 years) - 2023

### Extra Historical Averages
```{r}
flint_all_year_historical_avgs_extra <- flint_all_year_historical_extra %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Historical")
#flint_all_year_historical_avgs_extra

flint_all_year_avgs_extra <- bind_rows(flint_all_year_recent_avgs, flint_all_year_historical_avgs_extra) #combine into 1 dataframe 

bioclim_all_year_historical_avgs_extra <- bioclim_all_year_historical_extra %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_quarter", "tmean_driest_quarter", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_quarter", "ppt_coldest_quarter"),
               c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Historical") 
#bioclim_all_year_historical_avgs_extra

bioclim_all_year_avgs_extra <- bind_rows(bioclim_all_year_recent_avgs, bioclim_all_year_historical_avgs_extra) #combine into 1 dataframe 


bioclim_flint_all_year_avgs_extra <- full_join(flint_all_year_avgs_extra, bioclim_all_year_avgs_extra) %>% 
  select(TimePd, parent.pop:ppt_coldest_quarter)
head(bioclim_flint_all_year_avgs_extra)
```

### Correlations - Extra Historical 

```{r}
#normalize the data
climate_normalized_bioclim_flint_all_year_avgs_extra <- bioclim_flint_all_year_avgs_extra %>% ungroup() %>% 
  select(cwd:ppt_coldest_quarter) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_bioclim_flint_all_year_avgs_extra)
dim(climate_normalized_bioclim_flint_all_year_avgs_extra)
cor.norm_extra = cor(climate_normalized_bioclim_flint_all_year_avgs_extra) #test correlations among the traits
cor.sig_extra <- cor.mtest(climate_normalized_bioclim_flint_all_year_avgs_extra, method = "pearson")

corrplot(cor.norm_extra, type="upper",
         tl.srt = 45, p.mat = cor.sig_extra$p, 
         sig.level = 0.05, insig="blank")

#tmn, tmx, tmean_wettest_quarter, tmean_driest_quarter and ann_tmean all highly correlated - only keep ann_tmean 
#temp_ann_range and mean_diurnal range correlated - keep temp_ann_range 
#ann_ppt and ppt 100% correlated (ppt = avg across monts, ann_ppt = avg of the total ppt in a year) - only keep ann_ppt 
#ppt_warmest and coldest quarter highly correlated with ann_ppt and ppt - take both out 
```

Extra Historical
```{r}
climate_normalized_bioclim_flint_all_year_avgs_extra_historic <- bioclim_flint_all_year_avgs_extra %>% 
  ungroup() %>% 
  filter(TimePd=="Historical") %>% 
  select(tmn:tmean_driest_quarter, cwd, pck, ppt, ann_ppt:ppt_coldest_quarter) %>% 
  scale() #normalize the data so they're all on the same scale

cor.norm_historic_extra = cor(climate_normalized_bioclim_flint_all_year_avgs_extra_historic) #test correlations among the traits
cor.sig_historic_extra <- cor.mtest(climate_normalized_bioclim_flint_all_year_avgs_extra_historic, method = "pearson")

corrplot(cor.norm_historic_extra, type="upper",
         tl.srt = 45, p.mat = cor.sig_historic_extra$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```


### PCA - Recent + Extra Historical

```{r}
all_bioclim_flint_avgs.pc_extra = prcomp(bioclim_flint_all_year_avgs_extra[c(7:8, 12, 14:15, 18:19)], scale = TRUE, center = TRUE)
str(all_bioclim_flint_avgs.pc_extra)
```

plot % Variance Explained

```{r}
summary(all_bioclim_flint_avgs.pc_extra)

tibble(PC=str_c("PC",str_pad(1:7,2,pad="0")),
       percent_var=all_bioclim_flint_avgs.pc_extra$sdev[1:7]^2/sum(all_bioclim_flint_avgs.pc_extra$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
all_bioclim_flint_avgs.pc_extra.dat = data.frame(all_bioclim_flint_avgs.pc_extra$x)

all_bioclim_flint_avgs_locs.pc_extra = cbind(bioclim_flint_all_year_avgs_extra, all_bioclim_flint_avgs.pc_extra.dat)

all_bioclim_flint_avgs_loadings_extra = data.frame(varnames=rownames(all_bioclim_flint_avgs.pc_extra$rotation), all_bioclim_flint_avgs.pc_extra$rotation)
all_bioclim_flint_avgs_loadings_extra
```

```{r}
autoplot(all_bioclim_flint_avgs.pc_extra, data = bioclim_flint_all_year_avgs_extra,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
autoplot(all_bioclim_flint_avgs.pc_extra, data = bioclim_flint_all_year_avgs_extra,
         x=1, y=4,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
all_bioclim_flint_avgs_locs.pc_avg_extra <- all_bioclim_flint_avgs_locs.pc_extra %>%
  group_by(parent.pop, elev_m, TimePd) %>%
  summarise(across(.cols=starts_with("PC", ignore.case = FALSE), .fns = mean)) %>%
  ungroup()
```

```{r}
all_bioclim_flint_avgs_locs.pc_avg_extra %>% 
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point() +
  facet_wrap(~TimePd) +
  coord_fixed(ratio = 1.5)
```

## Permanova - Extra Historical 2023
See Jenny's github: https://github.com/jrgremer/Seedbanks_drought_deluge  
```{r}
all_bioclim_flint_avgs_locs.pc_dist_extra <- all_bioclim_flint_avgs_locs.pc_extra %>% ungroup() %>% select(PC1:PC7)
dist_matrix_all_year_extra <- dist(all_bioclim_flint_avgs_locs.pc_dist_extra, method = "euclidian") #use a distance function to calculate euclidian distance in PCA space
permanova_results_all_year_extra <- adonis2(dist_matrix_all_year_extra ~ TimePd*elev_m*Lat, data = all_bioclim_flint_avgs_locs.pc_extra) #use adonis2 to run the permanova
#rows in data must be in same order as rows in distance matrix 

#look at output 
summary(permanova_results_all_year_extra)
permanova_results_all_year_extra
permanova_results_all_year_extra$`Pr(>F)`

permanova_results_all_year_terms_extra <- adonis2(dist_matrix_all_year_extra ~ TimePd*elev_m*Lat, data = all_bioclim_flint_avgs_locs.pc_extra, by = "terms")
permanova_results_all_year_terms_extra
```

## LM on PCs (follow up on permanova) - Extra Historical 2023

For each PC fit a model with timepd and parent.pop/elev to test for climate change

```{r}
lmer_results_all_year_extra <- all_bioclim_flint_avgs_locs.pc_extra %>%
  ungroup() %>% 
  select(TimePd, parent.pop, elev_m, Lat, Long, PC1:PC7) %>% 
  pivot_longer(starts_with("PC", ignore.case = FALSE), 
               names_to = "PC", values_to = "value") %>% 
  group_by(PC) %>% 
  nest(data=c(TimePd, parent.pop, elev_m, Lat, Long, value)) %>% 
  mutate(glm=map(data, ~ glm(value ~ TimePd*elev_m*Lat,
                               data=.x)),
         anova = map(glm, ~ broom.mixed::tidy(anova(.x))))

#lmer_results_all_year_extra
```

```{r}
PC_anova_extra <- lmer_results_all_year_extra %>% select(-data, -glm) %>% unnest(anova) %>%
  select(PC, term, p.value) %>%
  filter(p.value < 0.05) %>%
  arrange(term, p.value)

PC_anova_extra

lmer_results_all_year_extra %>% select(-data, -glm) %>% unnest(anova) %>% 
  filter(PC=="PC1"| PC=="PC4") #the two PCs with most sig p-values for timepd

#mod_test <- glm(PC4 ~ TimePd*elev_m*Lat, data=all_bioclim_flint_avgs_locs.pc_extra)
#summary(mod_test)
#anova(mod_test)
```

```{r}
autoplot(all_bioclim_flint_avgs.pc_extra, data = bioclim_flint_all_year_avgs_extra,
         x=1, y=4,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()

all_bioclim_flint_avgs_locs.pc_avg_extra %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC4, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)

all_bioclim_flint_avgs_locs.pc_avg_extra %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC4, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)
#high elev climate seems to be shifting to be more similar to low elev 
```

```{r}
home_sites_pca_extra <- all_bioclim_flint_avgs_locs.pc_avg_extra %>%  
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC4, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  #geom_point(size=2, alpha=0.7) +
  labs(x="PC1 (50.61%)", y="PC4 (8.5%)", color="Elevation (m)") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8) +
  annotate("text", x = -4, y = -5.1, label = "Warm \n No Snow") +
  annotate("text", x = 4, y = -5.1, label = "Cold \n Snow") +
  annotate("text", x = -5.4, y = -3.5, label = "High Temp \n Seasonality") +
  annotate("text", x = -5.4, y = 1.5, label = "Low Temp \n Seasonality") +
  coord_cartesian(ylim = c(-4.1, 2), xlim = c(-4,4), clip = "off") +
  theme_classic() +
  theme(text=element_text(size=28))
#ggsave("../output/Climate/Wtr_Year_PC1-PC4_ExtraHist.png", width = 7.4, height = 6, units = "in")

#str(home_sites_pca) #can add predicted WL2 point to the existing plot's data 
home_sites_pca_extra$data <- rbind(home_sites_pca_extra$data, 
  data.frame(
    parent.pop = "WL2_Garden",
    elev_m = 2020,
    TimePd = "2023",
    PC1 = WL2Grdn_predicted[, "PC1"],
    PC2 = WL2Grdn_predicted[, "PC2"],
    PC3 = WL2Grdn_predicted[, "PC3"],
    PC4 = WL2Grdn_predicted[, "PC4"],
    PC5 = WL2Grdn_predicted[, "PC5"],
    PC6 = WL2Grdn_predicted[, "PC6"],
    PC7 = WL2Grdn_predicted[, "PC7"],
    group = "new"
  )
)

home_sites_pca_extra + 
  geom_point(data=filter(home_sites_pca_extra$data, parent.pop == "WL2_Garden"), size=3, shape = 8, show_guide = FALSE)
ggsave("../output/Climate/Wtr_Year_PC1-PC4_PlusGarden_ExtraHist.png", width = 7.4, height = 6, units = "in")
```

## For 2024 (Water Year Avgs)
### Prep

```{r}
bioclim_2024_allyear_prep <- flint_all_year_wtr_yr %>% 
  filter(parent.pop != "UCD_Garden", parent.pop != "WL2_Garden") %>%  #remove garden sites 
  rename(tmin=tmn, tmax=tmx, year_cal=year, year=wtr_yr) %>% #rename columns to match what calc_biovars expects, also make sure it uses water year 
  filter(year != "1895", year !="2025") %>%  #remove years with less than 12 months of data
  arrange(parent.pop, year, month)

bioclim_2024_all_year <- tibble(bio1=NA, bio2=NA, bio4=NA, bio7=NA, bio8=NA, bio9=NA, bio12=NA, bio15=NA, bio18=NA, bio19=NA, year=2025) #blank tibble to bind calculations to
 
popids <- unique(bioclim_2024_allyear_prep$parent.pop) #list of pop ids for for loop 

pop_elev <- flint_all_year_wtr_yr %>% select(parent.pop:Long) %>% distinct()
```

### Calculation

```{r}
for(i in popids) {
  A <- bioclim_2024_allyear_prep %>% filter(parent.pop==i) %>% calc_biovars() %>% mutate(parent.pop=i)
  #print(A)
  bioclim_2024_all_year <- bind_rows(bioclim_2024_all_year, A)
}
unique(bioclim_2024_all_year$parent.pop) #has all the populations in there!
```

### SUBSET

```{r}
bioclim_2024_all_year_final <- bioclim_2024_all_year %>% 
  select(parent.pop, year, ann_tmean=bio1, mean_diurnal_range=bio2, 
         temp_seasonality=bio4, temp_ann_range=bio7, tmean_wettest_quarter=bio8,
         tmean_driest_quarter=bio9, ann_ppt=bio12, ppt_seasonality=bio15,
         ppt_warmest_quarter=bio18, ppt_coldest_quarter=bio19) %>%
  filter(year!=2025)
head(bioclim_2024_all_year_final)
```

### Merge with pop info

```{r}
pop_elev_bioclim_2024_all_year <- left_join(bioclim_2024_all_year_final, pop_elev) %>% 
  select(parent.pop, elevation.group:Long, year:ppt_coldest_quarter)
head(pop_elev_bioclim_2024_all_year)
#write_csv(pop_elev_bioclim_2024_all_year, "../output/Climate/bioclim_all_year_UCD_pops_wtr_year_2024.csv")
```

### Calculation of recent (last 30 years) and historical climate (prior 30 years)

```{r}
flint_all_year_recent_2024 <- flint_all_year_wtr_yr %>% 
  filter(parent.pop != "UCD_Garden", parent.pop != "WL2_Garden") %>%  #remove garden sites 
  filter(wtr_yr>1994 & wtr_yr<=2024) %>% 
  select(parent.pop:month, wtr_yr, cwd, pck, ppt, tmn, tmx)
head(flint_all_year_recent_2024)
tail(flint_all_year_recent_2024)

flint_all_year_historical_2024 <- flint_all_year_wtr_yr %>% 
  filter(parent.pop != "UCD_Garden", parent.pop != "WL2_Garden") %>%  #remove garden sites 
  filter(wtr_yr<=1994 & wtr_yr>1964) %>% 
  select(parent.pop:month, wtr_yr, cwd, pck, ppt, tmn, tmx)
head(flint_all_year_historical_2024, 13)
tail(flint_all_year_historical_2024, 13)

bioclim_2024_all_year_recent <- pop_elev_bioclim_2024_all_year %>% filter(year>1994 & year<=2024) #year here means water year (see above code where bioclim vars were calculated)
#head(bioclim_2024_all_year_recent)
#tail(bioclim_2024_all_year_recent)

bioclim_2024_all_year_historical <- pop_elev_bioclim_2024_all_year %>% filter(year<=1994 & year>1964)
#head(bioclim_2024_all_year_historical, 13)
#tail(bioclim_2024_all_year_historical, 13)
```

### Yearly averages for Flint
```{r}
flint_all_year_recent_yravgs_2024 <- flint_all_year_recent_2024 %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, wtr_yr) %>% 
  summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  rename(year=wtr_yr)
#flint_all_year_recent_yravgs_2024

flint_all_year_historical_yravgs_2024 <- flint_all_year_historical_2024 %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, wtr_yr) %>% 
  summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  rename(year=wtr_yr)
#flint_all_year_historical_yravgs_2024
```

### Merge with bioclim
```{r}
bioclim_2024_flint_yrly_avgs_recent <- full_join(flint_all_year_recent_yravgs_2024, bioclim_2024_all_year_recent) 
#write_csv(bioclim_2024_flint_yrly_avgs_recent, "../output/Climate/fullyear_wtr_year_avgs_Recent_2024.csv")
bioclim_2024_flint_yrly_avgs_historical <- full_join(flint_all_year_historical_yravgs_2024, bioclim_2024_all_year_historical) 
#write_csv(bioclim_2024_flint_yrly_avgs_historical, "../output/Climate/fullyear_wtr_year_avgs_Historical_2024.csv")
```

### Calculation of extra historical climate (prior 60 years)
```{r}
flint_all_year_historical_extra_2024 <- flint_all_year_wtr_yr %>% 
  filter(parent.pop != "UCD_Garden", parent.pop != "WL2_Garden") %>%  #remove garden sites 
  filter(wtr_yr<=1994 & wtr_yr>1934) %>% 
  select(parent.pop:month, wtr_yr, cwd, pck, ppt, tmn, tmx)
head(flint_all_year_historical_extra_2024, 13)
tail(flint_all_year_historical_extra_2024, 13)

bioclim_all_year_historical_extra_2024 <- pop_elev_bioclim_2024_all_year %>% filter(year<=1994 & year>1934)
#head(bioclim_all_year_historical_extra_2024, 13)
#tail(bioclim_all_year_historical_extra_2024, 13)
```

## WL2 Garden 2024
```{r}
bioclim_2024_allyear_prep_WL2Grdn <- flint_all_year_wtr_yr %>% 
  filter(parent.pop == "WL2_Garden") %>%  #only keep WL2 garden
  rename(tmin=tmn, tmax=tmx, year_cal=year, year=wtr_yr) %>% #rename columns to match what calc_biovars expects, also make sure it uses water year 
  filter(year=="2024") %>% #year of the experiment only 
  arrange(parent.pop, year, month)

bioclim_2024_all_year_WL2Grdn <- bioclim_2024_allyear_prep_WL2Grdn %>% 
  calc_biovars() %>% 
  mutate(parent.pop="WL2_Garden", year=="2024") 

bioclim_2024_all_year_final_WL2Grdn <- bioclim_2024_all_year_WL2Grdn %>% 
  select(parent.pop, year, ann_tmean=bio1, mean_diurnal_range=bio2, 
         temp_seasonality=bio4, temp_ann_range=bio7, tmean_wettest_quarter=bio8,
         tmean_driest_quarter=bio9, ann_ppt=bio12, ppt_seasonality=bio15,
         ppt_warmest_quarter=bio18, ppt_coldest_quarter=bio19) 

WL2Grdn_elev_bioclim_2024_all_year <- left_join(bioclim_2024_all_year_final_WL2Grdn, pop_elev) %>% 
  select(parent.pop, elevation.group:Long, year:ppt_coldest_quarter)
WL2Grdn_elev_bioclim_2024_all_year

WL2Grdn_flint_2024_all_year <- flint_all_year_wtr_yr %>% 
  filter(parent.pop == "WL2_Garden", wtr_yr=="2024") %>%  #only keep WL2 garden
  select(parent.pop:month, wtr_yr, cwd, pck, ppt, tmn, tmx)

WL2Grdn_flint_2024_all_year_summary <- WL2Grdn_flint_2024_all_year %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) 

WL2Grdn_flint_bioclim_2024_all_year <- left_join(WL2Grdn_flint_2024_all_year_summary, WL2Grdn_elev_bioclim_2024_all_year) %>% 
  select(-year)
```

## Avg across years and months (Flint + bioclim) - 2024

### Calculate avgs for flint

```{r}
flint_all_year_recent_avgs_2024 <- flint_all_year_recent_2024 %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Recent") 
#flint_all_year_recent_avgs_2024

flint_all_year_historical_avgs_2024 <- flint_all_year_historical_2024 %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Historical")
#flint_all_year_historical_avgs_2024

flint_all_year_avgs_2024 <- bind_rows(flint_all_year_recent_avgs_2024, flint_all_year_historical_avgs_2024) #combine into 1 dataframe 
head(flint_all_year_avgs_2024)
tail(flint_all_year_avgs_2024)
#write_csv(flint_all_year_avgs_2024, "../output/Climate/fullyear_FlintAvgs_wtr_year_2024.csv")
```

### Calculate avgs for bioclim 

```{r}
bioclim_all_year_recent_avgs_2024 <- bioclim_2024_all_year_recent %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_quarter", "tmean_driest_quarter", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_quarter", "ppt_coldest_quarter"),
               c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Recent") 
#bioclim_all_year_recent_avgs_2024

bioclim_all_year_historical_avgs_2024 <- bioclim_2024_all_year_historical %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_quarter", "tmean_driest_quarter", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_quarter", "ppt_coldest_quarter"),
               c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Historical") 
#bioclim_all_year_historical_avgs_2024

bioclim_all_year_avgs_2024 <- bind_rows(bioclim_all_year_recent_avgs_2024, bioclim_all_year_historical_avgs_2024) #combine into 1 dataframe 
head(bioclim_all_year_avgs_2024)
tail(bioclim_all_year_avgs_2024)
#write_csv(bioclim_all_year_avgs_2024, "../output/Climate/fullyear_BioClimAvgs_wtr_year_2024.csv")
```

Merge with flint

```{r}
bioclim_flint_all_year_avgs_2024 <- full_join(flint_all_year_avgs_2024, bioclim_all_year_avgs_2024) %>% 
  select(TimePd, parent.pop:ppt_coldest_quarter)
head(bioclim_flint_all_year_avgs_2024)
```

### Correlations - Recent + Historical

```{r}
#normalize the data
climate_normalized_bioclim_flint_all_year_avgs_2024 <- bioclim_flint_all_year_avgs_2024 %>% ungroup() %>% 
  select(cwd:ppt_coldest_quarter) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_bioclim_flint_all_year_avgs_2024)
dim(climate_normalized_bioclim_flint_all_year_avgs_2024)
cor.norm_2024 = cor(climate_normalized_bioclim_flint_all_year_avgs_2024) #test correlations among the traits
cor.sig_2024 <- cor.mtest(climate_normalized_bioclim_flint_all_year_avgs_2024, method = "pearson")

corrplot(cor.norm_2024, type="upper",
         tl.srt = 45, p.mat = cor.sig_2024$p, 
         sig.level = 0.05, insig="blank")

#tmn, tmx, tmean_wettest_quarter, tmean_driest_quarter and ann_tmean all highly correlated (97-99%) - only keep ann_tmean 
#temp_ann_range and mean_diurnal range 92% correlated - keep temp_ann_range 
#ann_ppt and ppt 100% correlated (ppt = avg across monts, ann_ppt = avg of the total ppt in a year) - only keep ann_ppt 
#ppt_warmest and coldest quarter highly correlated with ann_ppt and ppt - take both out 
```

Recent
```{r}
climate_normalized_bioclim_flint_all_year_avgs_recent_2024 <- bioclim_flint_all_year_avgs_2024 %>% 
  ungroup() %>% 
  filter(TimePd=="Recent") %>% 
  select(tmn:tmean_driest_quarter, cwd, pck, ppt, ann_ppt:ppt_coldest_quarter) %>% 
  scale() #normalize the data so they're all on the same scale

cor.norm_recent_2024 = cor(climate_normalized_bioclim_flint_all_year_avgs_recent_2024) #test correlations among the traits
cor.sig_recent_2024 <- cor.mtest(climate_normalized_bioclim_flint_all_year_avgs_recent_2024, method = "pearson")

corrplot(cor.norm_recent_2024, type="upper",
         tl.srt = 45, p.mat = cor.sig_recent_2024$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```

Historical
```{r}
climate_normalized_bioclim_flint_all_year_avgs_historic_2024 <- bioclim_flint_all_year_avgs_2024 %>% 
  ungroup() %>% 
  filter(TimePd=="Historical") %>% 
  select(tmn:tmean_driest_quarter, cwd, pck, ppt, ann_ppt:ppt_coldest_quarter) %>% 
  scale() #normalize the data so they're all on the same scale

cor.norm_historic_2024 = cor(climate_normalized_bioclim_flint_all_year_avgs_historic_2024) #test correlations among the traits
cor.sig_historic_2024 <- cor.mtest(climate_normalized_bioclim_flint_all_year_avgs_historic_2024, method = "pearson")

corrplot(cor.norm_historic_2024, type="upper",
         tl.srt = 45, p.mat = cor.sig_historic_2024$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```

### PCA - Recent + Historical

```{r}
all_bioclim_flint_avgs.pc_2024 = prcomp(bioclim_flint_all_year_avgs_2024[c(7:8, 12, 14:15, 18:19)], scale = TRUE, center = TRUE)
str(all_bioclim_flint_avgs.pc_2024)
```

plot % Variance Explained

```{r}
summary(all_bioclim_flint_avgs.pc_2024)

tibble(PC=str_c("PC",str_pad(1:7,2,pad="0")),
       percent_var=all_bioclim_flint_avgs.pc_2024$sdev[1:7]^2/sum(all_bioclim_flint_avgs.pc_2024$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
all_bioclim_flint_avgs.pc.dat_2024 = data.frame(all_bioclim_flint_avgs.pc_2024$x)

all_bioclim_flint_avgs_locs.pc_2024 = cbind(bioclim_flint_all_year_avgs_2024, all_bioclim_flint_avgs.pc.dat_2024)

all_bioclim_flint_avgs_loadings_2024 = data.frame(varnames=rownames(all_bioclim_flint_avgs.pc_2024$rotation), all_bioclim_flint_avgs.pc_2024$rotation)
all_bioclim_flint_avgs_loadings_2024 #very similar to 2023 
```

```{r}
autoplot(all_bioclim_flint_avgs.pc_2024, data = bioclim_flint_all_year_avgs_2024,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
autoplot(all_bioclim_flint_avgs.pc_2024, data = bioclim_flint_all_year_avgs_2024,
         x=1, y=4,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
all_bioclim_flint_avgs_locs.pc_2024_avg <- all_bioclim_flint_avgs_locs.pc_2024 %>%
  group_by(parent.pop, elev_m, TimePd) %>%
  summarise(across(.cols=starts_with("PC", ignore.case = FALSE), .fns = mean)) %>%
  ungroup()
all_bioclim_flint_avgs.pc
```

```{r}
all_bioclim_flint_avgs_locs.pc_2024_avg %>% 
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point() +
  facet_wrap(~TimePd) +
  coord_fixed(ratio = 1.5)
```

## Permanova - 2024
See Jenny's github: https://github.com/jrgremer/Seedbanks_drought_deluge  
```{r}
all_bioclim_flint_avgs_locs.pc_dist_2024 <- all_bioclim_flint_avgs_locs.pc_2024 %>% ungroup() %>% select(PC1:PC7)
dist_matrix_all_year_2024 <- dist(all_bioclim_flint_avgs_locs.pc_dist_2024, method = "euclidian") #use a distance function to calculate euclidian distance in PCA space
permanova_results_all_year_2024 <- adonis2(dist_matrix_all_year_2024 ~ TimePd*elev_m*Lat, data = all_bioclim_flint_avgs_locs.pc_2024) #use adonis2 to run the permanova
#rows in data must be in same order as rows in distance matrix 

#look at output 
summary(permanova_results_all_year_2024)
permanova_results_all_year_2024
permanova_results_all_year_2024$`Pr(>F)`

permanova_results_all_year_terms_2024 <- adonis2(dist_matrix_all_year_2024 ~ TimePd*elev_m*Lat, data = all_bioclim_flint_avgs_locs.pc_2024, by = "terms")
permanova_results_all_year_terms_2024 #very similar to 2023
```

## LM on PCs (follow up on permanova) - 2024

For each PC fit a model with timepd and parent.pop/elev to test for climate change

```{r}
lmer_results_all_year_2024 <- all_bioclim_flint_avgs_locs.pc_2024 %>%
  ungroup() %>% 
  select(TimePd, parent.pop, elev_m, Lat, Long, PC1:PC7) %>% 
  pivot_longer(starts_with("PC", ignore.case = FALSE), 
               names_to = "PC", values_to = "value") %>% 
  group_by(PC) %>% 
  nest(data=c(TimePd, parent.pop, elev_m, Lat, Long, value)) %>% 
  mutate(glm=map(data, ~ glm(value ~ TimePd*elev_m*Lat,
                               data=.x)),
         anova = map(glm, ~ broom.mixed::tidy(anova(.x))))

#lmer_results_all_year_2024
```

```{r}
PC_anova_2024 <- lmer_results_all_year_2024 %>% select(-data, -glm) %>% unnest(anova) %>%
  select(PC, term, p.value) %>%
  filter(p.value < 0.05) %>%
  arrange(term, p.value)

PC_anova_2024

lmer_results_all_year_2024 %>% select(-data, -glm) %>% unnest(anova) %>% 
  filter(PC=="PC1"| PC=="PC4") #the two PCs with most sig p-values for timepd

#mod_test <- glm(PC4 ~ TimePd*elev_m*Lat, data=all_bioclim_flint_avgs_locs.pc_2024)
#summary(mod_test)
#anova(mod_test)
```

```{r}
autoplot(all_bioclim_flint_avgs.pc_2024, data = bioclim_flint_all_year_avgs_2024,
         x=1, y=4,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()

all_bioclim_flint_avgs_locs.pc_2024_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC4, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)

all_bioclim_flint_avgs_locs.pc_2024_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC4, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)
#high elev climate seems to be shifting to be more similar to low elev 
```

```{r}
home_sites_pca_2024 <- all_bioclim_flint_avgs_locs.pc_2024_avg %>%  
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC4, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  #geom_point(size=2, alpha=0.7) +
  labs(x="PC1 (50%)", y="PC4 (8.94%)", color="Elevation (m)") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8) +
  annotate("text", x = -4, y = -5.1, label = "Warm \n No Snow") +
  annotate("text", x = 4, y = -5.1, label = "Cold \n Snow") +
  annotate("text", x = -5.4, y = -3.5, label = "High Temp \n Seasonality") +
  annotate("text", x = -5.4, y = 1.5, label = "Low Temp \n Seasonality") +
  coord_cartesian(ylim = c(-4.1, 2), xlim = c(-4,4), clip = "off") +
  theme_classic() +
  theme(text=element_text(size=28))
#ggsave("../output/Climate/Wtr_Year_PC1-PC4_2024.png", width = 7.4, height = 6, units = "in")

## add WL2 garden 2024
WL2Grdn_pc_prep_2024 <- WL2Grdn_flint_bioclim_2024_all_year %>% ungroup() %>% select(cwd:pck, ann_tmean, temp_seasonality, temp_ann_range, ann_ppt, ppt_seasonality)
#scale(WL2Grdn_pc_prep_2024, all_bioclim_flint_avgs.pc$center, all_bioclim_flint_avgs.pc$scale) %*% all_bioclim_flint_avgs.pc$rotation  #gives same result as below 
WL2Grdn_predicted_2024 <- predict(all_bioclim_flint_avgs.pc_2024, newdata = WL2Grdn_pc_prep_2024)
WL2Grdn_predicted_2024

#str(home_sites_pca_2024) #can add predicted WL2 point to the existing plot's data 
home_sites_pca_2024$data <- rbind(home_sites_pca_2024$data, 
  data.frame(
    parent.pop = "WL2_Garden",
    elev_m = 2020,
    TimePd = "2024",
    PC1 = WL2Grdn_predicted_2024[, "PC1"],
    PC2 = WL2Grdn_predicted_2024[, "PC2"],
    PC3 = WL2Grdn_predicted_2024[, "PC3"],
    PC4 = WL2Grdn_predicted_2024[, "PC4"],
    PC5 = WL2Grdn_predicted_2024[, "PC5"],
    PC6 = WL2Grdn_predicted_2024[, "PC6"],
    PC7 = WL2Grdn_predicted_2024[, "PC7"],
    group = "new"
  )
)

home_sites_pca_2024 + 
  geom_point(data=filter(home_sites_pca_2024$data, parent.pop == "WL2_Garden"), size=3, shape = 8, show_guide = FALSE)
ggsave("../output/Climate/Wtr_Year_PC1-PC4_PlusGarden_2024.png", width = 7.4, height = 6, units = "in")
```

---

## 2023 Old PCAs

### All years and months included (Flint)

#### Correlations - Flint Recent

```{r}
#normalize the data
climate_normalized_all_flint_recent <- flint_all_year_recent %>% select(cwd, pck, ppt, tmn, tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_flint_recent)
cor.norm = cor(climate_normalized_all_flint_recent) #test correlations among the traits
corrplot(cor.norm, type = "upper")
#tmn and tmx highly correlated, consider removing one 
```

#### PCA - Flint Recent

```{r}
#flint_all_year_recent[c(8:12)]
all_flint_recent.pc = prcomp(flint_all_year_recent[c(8:12)], scale = TRUE, center = TRUE)

str(all_flint_recent.pc)
```

plot % Variance Explained

```{r}
summary(all_flint_recent.pc)

tibble(PC=str_c("PC",str_pad(1:5,2,pad="0")),
       percent_var=all_flint_recent.pc$sdev[1:5]^2/sum(all_flint_recent.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
all_flint_recent.pc.dat = data.frame(all_flint_recent.pc$x)

all_flint_recent_locs.pc = cbind(flint_all_year_recent, all_flint_recent.pc.dat)

all_flint_recent_loadings = data.frame(varnames=rownames(all_flint_recent.pc$rotation), all_flint_recent.pc$rotation)
all_flint_recent_loadings
```

```{r}
autoplot(all_flint_recent.pc, data = flint_all_year_recent,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=8, loadings.label.colour="black", loadings.label.vjust = -0.2) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()

#for plot customizations see: ?ggbiplot
```

PCs 3 and 4

```{r}
autoplot(all_flint_recent.pc, data = flint_all_year_recent,
         x=3, y=4,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=8, loadings.label.colour="black", loadings.label.vjust = -0.2) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

#### Correlations - Flint Historical

```{r}
#normalize the data
climate_normalized_all_flint_historical <- flint_all_year_historical %>% select(cwd, pck, ppt, tmn, tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_flint_historical)
cor.norm = cor(climate_normalized_all_flint_historical) #test correlations among the traits
corrplot(cor.norm)
#tmn and tmx highly correlated, consider removing one 
```

#### PCA - Flint Historical

```{r}
#flint_all_year_historical[c(8:12)]
all_flint_historical.pc = prcomp(flint_all_year_historical[c(8:12)], scale = TRUE, center = TRUE)

str(all_flint_historical.pc)
```

plot % Variance Explained

```{r}
summary(all_flint_historical.pc)

tibble(PC=str_c("PC",str_pad(1:5,2,pad="0")),
       percent_var=all_flint_historical.pc$sdev[1:5]^2/sum(all_flint_historical.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
all_flint_historical.pc.dat = data.frame(all_flint_historical.pc$x)

all_flint_historical_locs.pc = cbind(flint_all_year_historical, all_flint_historical.pc.dat)

all_flint_historical_loadings = data.frame(varnames=rownames(all_flint_historical.pc$rotation), all_flint_historical.pc$rotation)
all_flint_historical_loadings
```

```{r}
autoplot(all_flint_historical.pc, data = flint_all_year_historical,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=8, loadings.label.colour="black", loadings.label.vjust = -0.2) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

PCs 3 and 4

```{r}
autoplot(all_flint_historical.pc, data = flint_all_year_historical,
         x=3, y=4,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=8, loadings.label.colour="black", loadings.label.vjust = -0.2) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

### All years included (bioclim)

#### Correlations - bioclim Recent

```{r}
#normalize the data
climate_normalized_all_bioclim_recent <- bioclim_all_year_recent %>% select(ann_tmean:ppt_coldest_quarter) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_bioclim_recent)
cor.norm = cor(climate_normalized_all_bioclim_recent) #test correlations among the traits
corrplot(cor.norm)

#ann_ppt and ppt_coldest quarter highly correlated, consider removing one 
#ann_tmean and t_mean wettest and driest quarter are highly correlated, consider only keeping one 
```

#### PCA - bioclim Recent

```{r}
all_bioclim_recent.pc = prcomp(bioclim_all_year_recent[c(7:10, 13:15)], scale = TRUE, center = TRUE) #took out ppt_coldest quarter & t_mean wettest and driest quarter

str(all_bioclim_recent.pc)
```

plot % Variance Explained

```{r}
summary(all_bioclim_recent.pc)

tibble(PC=str_c("PC",str_pad(1:5,2,pad="0")),
       percent_var=all_bioclim_recent.pc$sdev[1:5]^2/sum(all_bioclim_recent.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
all_bioclim_recent.pc.dat = data.frame(all_bioclim_recent.pc$x)

all_bioclim_recent_locs.pc = cbind(bioclim_all_year_recent, all_bioclim_recent.pc.dat)

all_bioclim_recent_loadings = data.frame(varnames=rownames(all_bioclim_recent.pc$rotation), all_bioclim_recent.pc$rotation)
all_bioclim_recent_loadings
```

```{r}
autoplot(all_bioclim_recent.pc, data = bioclim_all_year_recent,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

PCs 3 and 4

```{r}
autoplot(all_bioclim_recent.pc, data = bioclim_all_year_recent,
         x=3, y=4,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

PCs 4 and 5

```{r}
autoplot(all_bioclim_recent.pc, data = bioclim_all_year_recent,
         x=4, y=5,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

#### Correlations - bioclim Historical

```{r}
#normalize the data
climate_normalized_all_bioclim_historical <- bioclim_all_year_historical %>% select(ann_tmean:ppt_coldest_quarter) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_bioclim_historical)
cor.norm = cor(climate_normalized_all_bioclim_historical) #test correlations among the traits
corrplot(cor.norm)

#ann_ppt and ppt_coldest quarter highly correlated, consider removing one 
#ann_tmean and t_mean wettest and driest quarter are highly correlated, consider only keeping one 
```

#### PCA - bioclim Historical

```{r}
all_bioclim_historical.pc = prcomp(bioclim_all_year_historical[c(7:10, 13:15)], scale = TRUE, center = TRUE) #took out ppt_coldest quarter & t_mean wettest and driest quarter

str(all_bioclim_historical.pc)
```

plot % Variance Explained

```{r}
summary(all_bioclim_historical.pc)

tibble(PC=str_c("PC",str_pad(1:5,2,pad="0")),
       percent_var=all_bioclim_historical.pc$sdev[1:5]^2/sum(all_bioclim_historical.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
all_bioclim_historical.pc.dat = data.frame(all_bioclim_historical.pc$x)

all_bioclim_historical_locs.pc = cbind(bioclim_all_year_historical, all_bioclim_historical.pc.dat)

all_bioclim_historical_loadings = data.frame(varnames=rownames(all_bioclim_historical.pc$rotation), all_bioclim_historical.pc$rotation)
all_bioclim_historical_loadings
```

```{r}
autoplot(all_bioclim_historical.pc, data = bioclim_all_year_historical,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

PCs 3 and 4

```{r}
autoplot(all_bioclim_historical.pc, data = bioclim_all_year_historical,
         x=3, y=4,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

PCs 4 and 5

```{r}
autoplot(all_bioclim_historical.pc, data = bioclim_all_year_historical,
         x=4, y=5,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

### Monthly Averages - Flint

#### Calculate avgs

```{r}
flint_all_year_recent_mosavgs <- flint_all_year_recent %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, month) %>% 
  summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Recent") 
flint_all_year_recent_mosavgs

flint_all_year_historical_mosavgs <- flint_all_year_historical %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, month) %>% 
  summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Historical")
flint_all_year_historical_mosavgs

flint_all_year_mosavgs <- bind_rows(flint_all_year_recent_mosavgs, flint_all_year_historical_mosavgs) #combine into 1 dataframe 
head(flint_all_year_mosavgs)
tail(flint_all_year_mosavgs)
```

#### Correlations - Recent + Historical

```{r}
#normalize the data
climate_normalized_flint_all_year_mosavgs <- flint_all_year_mosavgs %>% ungroup() %>% 
  select(cwd:tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_flint_all_year_mosavgs)
cor.norm = cor(climate_normalized_flint_all_year_mosavgs) #test correlations among the traits
corrplot(cor.norm)

#tmn and tmx highly correlated, consider removing one (98%)
#tmx and pck highly neg correlated (-85%)
#tmn and pck highly neg correlated (-86%)
```

#### PCA - Recent + Historical

```{r}
#flint_all_year_mosavgs[c(8:12)]
mos_flint.pc = prcomp(flint_all_year_mosavgs[c(7:11)], scale = TRUE, center = TRUE)

str(mos_flint.pc)
```

plot % Variance Explained

```{r}
summary(mos_flint.pc)

tibble(PC=str_c("PC",str_pad(1:5,2,pad="0")),
       percent_var=mos_flint.pc$sdev[1:5]^2/sum(mos_flint.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
mos_flint.pc.dat = data.frame(mos_flint.pc$x)

mos_flint_locs.pc = cbind(flint_all_year_mosavgs, mos_flint.pc.dat)

mos_flint_loadings = data.frame(varnames=rownames(mos_flint.pc$rotation), mos_flint.pc$rotation)
mos_flint_loadings
```

```{r}
autoplot(mos_flint.pc, data = flint_all_year_mosavgs,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="month",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
#high elev seems most similar to low elev in summer months 
```

```{r}
mos_flint_locs.pc_avg <- mos_flint_locs.pc %>%
  group_by(parent.pop, elev_m, TimePd,month) %>%
  summarise(across(.cols=starts_with("PC"), .fns = mean)) %>%
  ungroup()

mos_flint_locs.pc_avg
mos_flint_locs.pc_avg$month <- factor(mos_flint_locs.pc_avg$month, levels=month_order) 
```

```{r}
mos_flint_locs.pc_avg %>% 
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point() +
  facet_wrap(~month) +
  coord_fixed(ratio = 1.5)
#as I suspected from the raw PC figure, high elev pop cliamte is most similar to low elev climate in the summer months 
#hard to see differences between recent and historical with this fig though
```

```{r}
mos_flint_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m, month))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point() +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")))
#high elev climate seems to be shifting to be more similar to low elev 
#hard to see which months though 
```

### Avg across years and months (Flint)

#### Correlations - Recent + Historical

```{r}
#normalize the data
climate_normalized_flint_all_year_avgs <- flint_all_year_avgs %>% ungroup() %>% 
  select(cwd:tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_flint_all_year_avgs)
cor.norm = cor(climate_normalized_flint_all_year_avgs) #test correlations among the traits
corrplot(cor.norm)

#tmn and tmx highly correlated, consider removing one (98%)
#tmx and pck highly neg correlated (-85%)
#tmn and pck highly neg correlated (-86%)
```

#### PCA - Recent + Historical

```{r}
#flint_all_year_avgs[c(8:12)]
avgs_flint.pc = prcomp(flint_all_year_avgs[c(6:10)], scale = TRUE, center = TRUE)

str(avgs_flint.pc)
```

plot % Variance Explained

```{r}
summary(avgs_flint.pc)

tibble(PC=str_c("PC",str_pad(1:5,2,pad="0")),
       percent_var=avgs_flint.pc$sdev[1:5]^2/sum(avgs_flint.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
avgs_flint.pc.dat = data.frame(avgs_flint.pc$x)

avgs_flint_locs.pc = cbind(flint_all_year_avgs, avgs_flint.pc.dat)

avgs_flint_loadings = data.frame(varnames=rownames(avgs_flint.pc$rotation), avgs_flint.pc$rotation)
avgs_flint_loadings
```

```{r}
autoplot(avgs_flint.pc, data = flint_all_year_avgs,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
avgs_flint_locs.pc_avg <- avgs_flint_locs.pc %>%
  group_by(parent.pop, elev_m, TimePd) %>%
  summarise(across(.cols=starts_with("PC"), .fns = mean)) %>%
  ungroup()

avgs_flint_locs.pc_avg
```

```{r}
avgs_flint_locs.pc_avg %>% 
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point() +
  facet_wrap(~TimePd) +
  coord_fixed(ratio = 1.5)
```

```{r}
avgs_flint_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)

avgs_flint_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)
#high elev climate seems to be shifting to be more similar to low elev 
```
