---
title: "Climate_Figs_All_Year"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Climate Figures with Year-Round data (as opposed to growth season)

## Relevant Libraries and Functions

```{r}
library(tidyverse)
library(ggrepel)
library(cowplot)
library(gridExtra)

sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
} #legend function for grid_arrange

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")
```

## Load the data
```{r}
flint_all_year <- read_csv("../output/Climate/flint_climate_UCDpops.csv")
head(flint_all_year)

bioclim_all_year <- read_csv("../output/Climate/bioclim_all_year_UCD_pops.csv")
head(bioclim_all_year)
```

## Historical and Recent Climate
```{r}
flint_recent_all_year <- flint_all_year %>% 
  mutate(month = recode(month, #convert months to numbers 
  jan = 1,
  feb = 2,
  mar = 3,
  apr = 4,
  may = 5,
  jun = 6,
  jul = 7,
  aug = 8,
  sep = 9,
  oct = 10,
  nov = 11,
  dec = 12
)) %>% 
  filter(parent.pop != "UCD_Garden", parent.pop != "WL2_Garden") %>%  #remove garden sites 
  filter(year>1993 & year<=2023) %>% 
  select(parent.pop:month, cwd, pck, ppt, tmn, tmx)
  
flint_historical_all_year <- flint_all_year %>% 
  mutate(month = recode(month, #convert months to numbers 
  jan = 1,
  feb = 2,
  mar = 3,
  apr = 4,
  may = 5,
  jun = 6,
  jul = 7,
  aug = 8,
  sep = 9,
  oct = 10,
  nov = 11,
  dec = 12
)) %>% 
  filter(parent.pop != "UCD_Garden", parent.pop != "WL2_Garden") %>%  #remove garden sites 
  filter(year<=1993 & year>1963) %>% 
  select(parent.pop:month, cwd, pck, ppt, tmn, tmx)

bioclim_recent <- bioclim_all_year %>% 
  filter(year>1993 & year<=2023)
  
bioclim_historical <- bioclim_all_year %>% 
  filter(year<=1993 & year>1963)
```


## Flint 

### Monthly Averages

#### Recent Years - Flint Year Round
```{r}
flint_recent_all_year_mosavgs <- flint_recent_all_year %>% group_by(parent.pop, elevation.group, elev_m, month) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(flint_recent_all_year_mosavgs) <- gsub("fn2", "sem", colnames(flint_recent_all_year_mosavgs))
names(flint_recent_all_year_mosavgs) <-gsub("fn1", "mean", colnames(flint_recent_all_year_mosavgs))
flint_recent_all_year_mosavgs #30 year averages during growth season months 
flint_recent_all_year_mosavgs$elevation.group <- factor(flint_recent_all_year_mosavgs$elevation.group, levels=elev_order)       

flint_recent_all_year_mosavgs %>% ggplot(aes(x=month, y=cwd_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=cwd_mean-cwd_sem,ymax=cwd_mean+cwd_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0")+ 
  labs(fill="Elevation",x="Population", y="Avg CWD" ,title = "Average CWD Year-Round - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1)) + 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/all_year_Avg_Monthly_CWD_RecentClim.png", width = 16, height = 10, units = "in")

flint_recent_all_year_mosavgs %>% ggplot(aes(x=month, y=pck_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=pck_mean-pck_sem,ymax=pck_mean+pck_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0")+ 
  labs(fill="Elevation",x="Population", y="Avg PCK" ,title = "Average PCK Year-Round - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1)) + 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/all_year_Avg_Monthly_PCK_RecentClim.png", width = 16, height = 10, units = "in")

flint_recent_all_year_mosavgs %>% ggplot(aes(x=month, y=ppt_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=ppt_mean-ppt_sem,ymax=ppt_mean+ppt_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg Precip" ,title = "Average Precip Year-Round - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))+ 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/all_year_Avg_Monthly_Precip_RecentClim.png", width = 12, height = 6, units = "in")

flint_recent_all_year_mosavgs %>% ggplot(aes(x=month, y=tmn_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmn_mean-tmn_sem,ymax=tmn_mean+tmn_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MinTemp" ,title = "Average MinTemp Year-Round - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))+ 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/all_year_Avg_Monthly_MinTemp_RecentClim.png", width = 12, height = 6, units = "in")

flint_recent_all_year_mosavgs %>% ggplot(aes(x=month, y=tmx_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmx_mean-tmx_sem,ymax=tmx_mean+tmx_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MaxTemp" ,title = "Average MaxTemp Year-Round - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))+ 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/all_year_Avg_Monthly_MaxTemp_RecentClim.png", width = 12, height = 6, units = "in")
```

#### Historical Years - Flint Year Round
```{r}
flint_historical_all_year_mosavgs <- flint_historical_all_year %>% group_by(parent.pop, elevation.group, elev_m, month) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(flint_historical_all_year_mosavgs) <- gsub("fn2", "sem", colnames(flint_historical_all_year_mosavgs))
names(flint_historical_all_year_mosavgs) <-gsub("fn1", "mean", colnames(flint_historical_all_year_mosavgs))
flint_historical_all_year_mosavgs #30 year averages during growth season months 
flint_historical_all_year_mosavgs$elevation.group <- factor(flint_historical_all_year_mosavgs$elevation.group, levels=elev_order)       

flint_historical_all_year_mosavgs %>% ggplot(aes(x=month, y=cwd_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=cwd_mean-cwd_sem,ymax=cwd_mean+cwd_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0")+ 
  labs(fill="Elevation",x="Population", y="Avg CWD" ,title = "Average CWD Year-Round - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1)) + 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/all_year_Avg_Monthly_CWD_HistoricalClim.png", width = 16, height = 10, units = "in")

flint_historical_all_year_mosavgs %>% ggplot(aes(x=month, y=pck_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=pck_mean-pck_sem,ymax=pck_mean+pck_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0")+ 
  labs(fill="Elevation",x="Population", y="Avg PCK" ,title = "Average PCK Year-Round - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1)) + 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/all_year_Avg_Monthly_PCK_HistoricalClim.png", width = 16, height = 10, units = "in")

flint_historical_all_year_mosavgs %>% ggplot(aes(x=month, y=ppt_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=ppt_mean-ppt_sem,ymax=ppt_mean+ppt_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg Precip" ,title = "Average Precip Year-Round - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))+ 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/all_year_Avg_Monthly_Precip_HistoricalClim.png", width = 12, height = 6, units = "in")

flint_historical_all_year_mosavgs %>% ggplot(aes(x=month, y=tmn_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmn_mean-tmn_sem,ymax=tmn_mean+tmn_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MinTemp" ,title = "Average MinTemp Year-Round - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))+ 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/all_year_Avg_Monthly_MinTemp_HistoricalClim.png", width = 12, height = 6, units = "in")

flint_historical_all_year_mosavgs %>% ggplot(aes(x=month, y=tmx_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmx_mean-tmx_sem,ymax=tmx_mean+tmx_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MaxTemp" ,title = "Average MaxTemp Year-Round - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))+ 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/all_year_Avg_Monthly_MaxTemp_HistoricalClim.png", width = 12, height = 6, units = "in")
```


### 30 Year Averages

#### Recent Years - Flint Year Round

```{r}
flint_recent_all_year_avgs <- flint_recent_all_year %>% group_by(parent.pop, elevation.group, elev_m) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(flint_recent_all_year_avgs) <- gsub("fn2", "sem", colnames(flint_recent_all_year_avgs))
names(flint_recent_all_year_avgs) <-gsub("fn1", "mean", colnames(flint_recent_all_year_avgs))
flint_recent_all_year_avgs #30 year averages during growth season months 
flint_recent_all_year_avgs$elevation.group <- factor(flint_recent_all_year_avgs$elevation.group, levels=elev_order)    
#write_csv(flint_recent_all_year_avgs, "../output/Climate/all_year__FlintAvgs_Recent.csv")

flint_recent_all_year_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, cwd_mean), y=cwd_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=cwd_mean-cwd_sem,ymax=cwd_mean+cwd_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0")+ 
  labs(fill="Elevation",x="Population", y="Avg CWD" ,title = "Average CWD Year-Round - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_Avg_CWD_RecentClim.png", width = 12, height = 6, units = "in")

flint_recent_all_year_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, pck_mean), y=pck_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=pck_mean-pck_sem,ymax=pck_mean+pck_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0")+ 
  labs(fill="Elevation",x="Population", y="Avg PCK" ,title = "Average PCK Year-Round - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_Avg_PCK_RecentClim.png", width = 12, height = 6, units = "in")

flint_recent_all_year_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_mean), y=ppt_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=ppt_mean-ppt_sem,ymax=ppt_mean+ppt_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg Precip" ,title = "Average Precip Year-Round - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_Avg_Precip_RecentClim.png", width = 12, height = 6, units = "in")

flint_recent_all_year_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmn_mean), y=tmn_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmn_mean-tmn_sem,ymax=tmn_mean+tmn_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MinTemp" ,title = "Average MinTemp Year-Round - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_Avg_MinTemp_RecentClim.png", width = 12, height = 6, units = "in")

flint_recent_all_year_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmx_mean), y=tmx_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmx_mean-tmx_sem,ymax=tmx_mean+tmx_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MaxTemp" ,title = "Average MaxTemp Year-Round - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_Avg_MaxTemp_RecentClim.png", width = 12, height = 6, units = "in")
```

#### Historical Years - Flint Year Round

```{r}
flint_historical_all_year_avgs <- flint_historical_all_year %>% group_by(parent.pop, elevation.group, elev_m) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(flint_historical_all_year_avgs) <- gsub("fn2", "sem", colnames(flint_historical_all_year_avgs))
names(flint_historical_all_year_avgs) <-gsub("fn1", "mean", colnames(flint_historical_all_year_avgs))
flint_historical_all_year_avgs #30 year averages during growth season months 
flint_historical_all_year_avgs$elevation.group <- factor(flint_historical_all_year_avgs$elevation.group, levels=elev_order)    
#write_csv(flint_historical_all_year_avgs, "../output/Climate/all_year__FlintAvgs_Historic.csv")

flint_historical_all_year_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, cwd_mean), y=cwd_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=cwd_mean-cwd_sem,ymax=cwd_mean+cwd_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0")+ 
  labs(fill="Elevation",x="Population", y="Avg CWD" ,title = "Average CWD Year-Round - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_Avg_CWD_HistoricalClim.png", width = 12, height = 6, units = "in")

flint_historical_all_year_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, pck_mean), y=pck_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=pck_mean-pck_sem,ymax=pck_mean+pck_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0")+ 
  labs(fill="Elevation",x="Population", y="Avg PCK" ,title = "Average PCK Year-Round - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_Avg_PCK_HistoricalClim.png", width = 12, height = 6, units = "in")

flint_historical_all_year_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_mean), y=ppt_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=ppt_mean-ppt_sem,ymax=ppt_mean+ppt_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg Precip" ,title = "Average Precip Year-Round - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_Avg_Precip_HistoricalClim.png", width = 12, height = 6, units = "in")

flint_historical_all_year_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmn_mean), y=tmn_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmn_mean-tmn_sem,ymax=tmn_mean+tmn_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MinTemp" ,title = "Average MinTemp Year-Round - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_Avg_MinTemp_HistoricalClim.png", width = 12, height = 6, units = "in")

flint_historical_all_year_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmx_mean), y=tmx_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmx_mean-tmx_sem,ymax=tmx_mean+tmx_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MaxTemp" ,title = "Average MaxTemp Year-Round - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_Avg_MaxTemp_HistoricalClim.png", width = 12, height = 6, units = "in")
```

## Bioclim - 30 year averages 

### Recent Years - BioClim
```{r}
bioclim_recent_avgs <- bioclim_recent %>% group_by(parent.pop, elevation.group, elev_m) %>%
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range", "tmean_wettest_quarter", "tmean_driest_quarter", "ann_ppt", "ppt_seasonality","ppt_warmest_quarter", "ppt_coldest_quarter"), c(mean, sem), na.rm = TRUE) 
names(bioclim_recent_avgs) <- gsub("fn2", "sem", colnames(bioclim_recent_avgs))
names(bioclim_recent_avgs) <-gsub("fn1", "avg", colnames(bioclim_recent_avgs))
bioclim_recent_avgs
#write_csv(bioclim_recent_avgs, "../output/Climate/all_year_BioClimAvgs_Recent.csv")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_avg), y=ann_tmean_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ann_tmean_avg-ann_tmean_sem,ymax=ann_tmean_avg+ann_tmean_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Annual Mean Temp", x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_AnnMeanTmp_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_avg), y=mean_diurnal_range_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=mean_diurnal_range_avg-mean_diurnal_range_sem,ymax=mean_diurnal_range_avg+mean_diurnal_range_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Diurnal Range", x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_DiurnalRange_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_avg), y=temp_seasonality_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=temp_seasonality_avg-temp_seasonality_sem,ymax=temp_seasonality_avg+temp_seasonality_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_TempSeasonality_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_avg), y=temp_ann_range_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=temp_ann_range_avg-temp_ann_range_sem,ymax=temp_ann_range_avg+temp_ann_range_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_AnnTmpRange_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_quarter_avg), y=tmean_wettest_quarter_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=tmean_wettest_quarter_avg-tmean_wettest_quarter_sem,ymax=tmean_wettest_quarter_avg+tmean_wettest_quarter_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_TmpMeanWet_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_quarter_avg), y=tmean_driest_quarter_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=tmean_driest_quarter_avg-tmean_driest_quarter_sem,ymax=tmean_driest_quarter_avg+tmean_driest_quarter_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_TmpMeanDry_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_avg), y=ann_ppt_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ann_ppt_avg-ann_ppt_sem,ymax=ann_ppt_avg+ann_ppt_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_AnnPPT_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_avg), y=ppt_seasonality_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ppt_seasonality_avg-ppt_seasonality_sem,ymax=ppt_seasonality_avg+ppt_seasonality_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_PPTSeasonality_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_quarter_avg), y=ppt_warmest_quarter_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ppt_warmest_quarter_avg-ppt_warmest_quarter_sem,ymax=ppt_warmest_quarter_avg+ppt_warmest_quarter_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_PPTWarm_RecentClim.png", width = 12, height = 6, units = "in")

bioclim_recent_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_quarter_avg), y=ppt_coldest_quarter_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ppt_coldest_quarter_avg-ppt_coldest_quarter_sem,ymax=ppt_coldest_quarter_avg+ppt_coldest_quarter_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Recent Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_PPTCold_RecentClim.png", width = 12, height = 6, units = "in")
```

### Historical Years - BioClim
```{r}
names(bioclim_historical)
bioclim_historical_avgs <- bioclim_historical %>% group_by(parent.pop, elevation.group, elev_m) %>%
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range", "tmean_wettest_quarter", "tmean_driest_quarter", "ann_ppt", "ppt_seasonality","ppt_warmest_quarter", "ppt_coldest_quarter"), c(mean, sem), na.rm = TRUE) 
names(bioclim_historical_avgs) <- gsub("fn2", "sem", colnames(bioclim_historical_avgs))
names(bioclim_historical_avgs) <-gsub("fn1", "avg", colnames(bioclim_historical_avgs))
bioclim_historical_avgs
#write_csv(bioclim_historical_avgs, "../output/Climate/all_year_BioClimAvgs_Historical.csv")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_avg), y=ann_tmean_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ann_tmean_avg-ann_tmean_sem,ymax=ann_tmean_avg+ann_tmean_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Annual Mean Temp", x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_AnnMeanTmp_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_avg), y=mean_diurnal_range_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=mean_diurnal_range_avg-mean_diurnal_range_sem,ymax=mean_diurnal_range_avg+mean_diurnal_range_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Diurnal Range", x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_DiurnalRange_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_avg), y=temp_seasonality_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=temp_seasonality_avg-temp_seasonality_sem,ymax=temp_seasonality_avg+temp_seasonality_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_TempSeasonality_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_avg), y=temp_ann_range_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=temp_ann_range_avg-temp_ann_range_sem,ymax=temp_ann_range_avg+temp_ann_range_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_AnnTmpRange_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_quarter_avg), y=tmean_wettest_quarter_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=tmean_wettest_quarter_avg-tmean_wettest_quarter_sem,ymax=tmean_wettest_quarter_avg+tmean_wettest_quarter_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_TmpMeanWet_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_quarter_avg), y=tmean_driest_quarter_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=tmean_driest_quarter_avg-tmean_driest_quarter_sem,ymax=tmean_driest_quarter_avg+tmean_driest_quarter_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_TmpMeanDry_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_avg), y=ann_ppt_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ann_ppt_avg-ann_ppt_sem,ymax=ann_ppt_avg+ann_ppt_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_AnnPPT_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_avg), y=ppt_seasonality_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ppt_seasonality_avg-ppt_seasonality_sem,ymax=ppt_seasonality_avg+ppt_seasonality_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_PPTSeasonality_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_quarter_avg), y=ppt_warmest_quarter_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ppt_warmest_quarter_avg-ppt_warmest_quarter_sem,ymax=ppt_warmest_quarter_avg+ppt_warmest_quarter_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_PPTWarm_HistoricalClim.png", width = 12, height = 6, units = "in")

bioclim_historical_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_quarter_avg), y=ppt_coldest_quarter_avg, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=ppt_coldest_quarter_avg-ppt_coldest_quarter_sem,ymax=ppt_coldest_quarter_avg+ppt_coldest_quarter_sem),width=.2, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",  x="Population", title ="Historical Climate - Growth Season") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/all_year_PPTCold_HistoricalClim.png", width = 12, height = 6, units = "in")
```
