---
title: "Climate_PCAs_AllYear"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Climate PCAs

-   Should use this code to check the significance of the PCA:
    <https://github.com/StatQuest/pca_demo/blob/master/pca_demo.R>
-   Remember this paper: Björklund, M. 2019. Be careful with your
    principal components. Evolution 73: 2151--2158.

-   Notes

    -   Include snowpack? Not in the growth season PCAs since growth season is based off snowpack. 
    -   Take out UCD_garden? It is pretty different from the other sites so it may be skewing things... depends on if we want the PCA to describe home sites-only or all home sites and gardens

## Load necessary libraries 
```{r}
library(tidyverse)
library(ggrepel)
#library(cowplot)
library(gridExtra)
library(corrplot) #plotting correlations 
library(rstatix) #performing cor_test
library(QBMS) #for function calc_biovars to calculate bioclim variables
library(ggfortify) #easier PCA figures
sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
} #legend function for grid_arrange

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")
```

## Load Flint Data
```{r}
flint_all_year <- read_csv("../output/Climate/flint_climate_UCDpops.csv")
head(flint_all_year)
```

## Generate bioclim for all year
-   annual mean temperature (BIO1)
-   mean diurnal range (BIO2) - (Mean of monthly (max temp - min temp))
-   temperature seasonality (BIO4) (standard deviation *100)
-   temperature annual range (BIO7) (Max Temperature of Warmest Month - Min Temperature of Coldest Month)
-   mean temp of wettest quarter (BIO8)
-   mean temp of driest quarter (BIO9)
-   annual precipitation (BIO12) - sum of ppt for the entire year (not the avg)
-   precipitation seasonality (BIO15)  (Coefficient of Variation)
-   precip of warmest quarter (BIO18)
-   precip of coldest quarter (BIO19)

### Prep
```{r}
bioclim_allyear_prep <- flint_all_year %>% 
  rename(tmin=tmn, tmax=tmx) %>% #rename columns to match what calc_biovars expects
  filter(year != "1895", year !="2024") %>%  #remove years with less than 12 months of data
  arrange(parent.pop, year, month)

bioclim_all_year <- tibble(bio1=NA, bio2=NA, bio4=NA, bio7=NA, bio8=NA, bio9=NA, bio12=NA, bio15=NA, bio18=NA, bio19=NA, year=2025) #blank tibble to bind calculations to
bioclim_all_year
 
popids <- unique(bioclim_allyear_prep$parent.pop) #list of pop ids for for loop 

pop_elev <- flint_all_year %>% select(parent.pop:Long) %>% distinct()
```

### Calculation
```{r}
for(i in popids) {
  A <- bioclim_allyear_prep %>% filter(parent.pop==i) %>% calc_biovars() %>% mutate(parent.pop=i)
  #print(A)
  bioclim_all_year <- bind_rows(bioclim_all_year, A)
}
unique(bioclim_all_year$parent.pop) #has all the populations in there!
```

### SUBSET
```{r}
bioclim_all_year_final <- bioclim_all_year %>% 
  select(parent.pop, year, ann_tmean=bio1, mean_diurnal_range=bio2, 
         temp_seasonality=bio4, temp_ann_range=bio7, tmean_wettest_quarter=bio8,
         tmean_driest_quarter=bio9, ann_ppt=bio12, ppt_seasonality=bio15,
         ppt_warmest_quarter=bio18, ppt_coldest_quarter=bio19) %>%
  filter(year!=2025)
head(bioclim_all_year_final)
```

### Merge with pop info
```{r}
pop_elev_bioclim_all_year <- left_join(bioclim_all_year_final, pop_elev) %>% 
  select(parent.pop, elevation.group:Long, year:ppt_coldest_quarter)
head(pop_elev_bioclim_all_year)
```

## Calculation of recent (last 30 years) and historical climate (prior 30 years)

Note: Removed 2024 to facilitate use of QBMS package (also most of the plants did not experience 2024)

```{r}
flint_all_year_recent <- flint_all_year %>% filter(year>1993 & year<=2023) %>% select(parent.pop:month, cwd, pck, ppt, tmn, tmx)
head(flint_all_year_recent)
tail(flint_all_year_recent)

flint_all_year_historical <- flint_all_year %>% filter(year<=1993 & year>1963) %>% select(parent.pop:month, cwd, pck, ppt, tmn, tmx)
head(flint_all_year_historical, 13)
tail(flint_all_year_historical, 13)

bioclim_all_year_recent <- pop_elev_bioclim_all_year %>% filter(year>1993 & year<=2023)
head(bioclim_all_year_recent)
tail(bioclim_all_year_recent)

bioclim_all_year_historical <- pop_elev_bioclim_all_year %>% filter(year<=1993 & year>1963)
head(bioclim_all_year_historical, 13)
tail(bioclim_all_year_historical, 13)
```

## All years and months included (Flint)

Correlations - Flint Recent

```{r}
#normalize the data
climate_normalized_all_flint_recent <- flint_all_year_recent %>% select(cwd, ppt, tmn, tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_flint_recent)
cor.norm = cor(climate_normalized_all_flint_recent) #test correlations among the traits
corrplot(cor.norm)
```

Overlay populations on PCA plot - Flint Recent

```{r}
#flint_all_year_recent[c(8:12)]
all_flint_recent.pc = prcomp(flint_all_year_recent[c(8, 10:12)], scale = TRUE, center = TRUE)
summary(all_flint_recent.pc)

all_flint_recent.pc.dat = data.frame(all_flint_recent.pc$x)

all_flint_recent_locs.pc = cbind(flint_all_year_recent, all_flint_recent.pc.dat)

all_flint_recent_loadings = data.frame(varnames=rownames(all_flint_recent.pc$rotation), all_flint_recent.pc$rotation)

autoplot(all_flint_recent.pc, all_flint_recent_loadings=TRUE, loadings.label=TRUE, 
         loadings.colour="blue", loadings.label.colour="red",  loadings.label.size = 10)
#this method results in different amounts of variation explained by PCs compared to princomp

autoplot(all_flint_recent.pc, data = flint_all_year_recent,
         colour='elevation.group', loadings=TRUE,
         loadings.colour='black', loadings.label = TRUE)

all_flint_recent_locs.pc %>% 
  ggplot(aes(x = PC1, y = PC2, colour=elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") 

```

Correlations - Flint Historical

```{r}
#normalize the data
climate_normalized_all_flint_historical <- flint_all_year_historical %>% select(cwd, ppt, tmn, tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_flint_historical)
cor.norm = cor(climate_normalized_all_flint_historical) #test correlations among the traits
corrplot(cor.norm)
```

Overlay populations on PCA plot - Flint Historical

```{r}
#flint_all_year_historical[c(8:12)]
all_flint_historical.pc = prcomp(flint_all_year_historical[c(8, 10:12)], scale = TRUE, center = TRUE)
summary(all_flint_historical.pc)

all_flint_historical.pc.dat = data.frame(all_flint_historical.pc$x)

all_flint_historical_locs.pc = cbind(flint_all_year_historical, all_flint_historical.pc.dat)

all_flint_historical_loadings = data.frame(varnames=rownames(all_flint_historical.pc$rotation), all_flint_historical.pc$rotation)

autoplot(all_flint_historical.pc, all_flint_historical_loadings=TRUE, loadings.label=TRUE, 
         loadings.colour="blue", loadings.label.colour="red",  loadings.label.size = 10)
#this method results in different amounts of variation explained by PCs compared to princomp

autoplot(all_flint_historical.pc, data = flint_all_year_historical,
         colour='elevation.group', loadings=TRUE,
         loadings.colour='black', loadings.label = TRUE)

all_flint_historical_locs.pc %>% 
  ggplot(aes(x = PC1, y = PC2, colour=elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") 

```

## All years included (bioclim)

Correlations - bioclim Recent

```{r}
#normalize the data
climate_normalized_all_bioclim_recent <- bioclim_all_year_recent %>% select(ann_tmean:ppt_coldest_quarter) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_bioclim_recent)
cor.norm = cor(climate_normalized_all_bioclim_recent) #test correlations among the traits
corrplot(cor.norm)
```

Overlay populations on PCA plot - bioclim Recent

```{r}
all_bioclim_recent.pc = prcomp(bioclim_all_year_recent[c(7:16)], scale = TRUE, center = TRUE)
summary(all_bioclim_recent.pc)

all_bioclim_recent.pc.dat = data.frame(all_bioclim_recent.pc$x)

all_bioclim_recent_locs.pc = cbind(bioclim_all_year_recent, all_bioclim_recent.pc.dat)

all_bioclim_recent_loadings = data.frame(varnames=rownames(all_bioclim_recent.pc$rotation), all_bioclim_recent.pc$rotation)

autoplot(all_bioclim_recent.pc, all_bioclim_recent_loadings=TRUE, loadings.label=TRUE, 
         loadings.colour="blue", loadings.label.colour="red",  loadings.label.size = 5)
#this method results in different amounts of variation explained by PCs compared to princomp

autoplot(all_bioclim_recent.pc, data = bioclim_all_year_recent,
         colour='elevation.group', loadings=TRUE,
         loadings.colour='black', loadings.label = TRUE)

all_bioclim_recent_locs.pc %>% 
  ggplot(aes(x = PC1, y = PC2, colour=elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") 

```

Correlations - bioclim Historical

```{r}
#normalize the data
climate_normalized_all_bioclim_historical <- bioclim_all_year_historical %>% select(ann_tmean:ppt_coldest_quarter) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_bioclim_historical)
cor.norm = cor(climate_normalized_all_bioclim_historical) #test correlations among the traits
corrplot(cor.norm)
```

Overlay populations on PCA plot - bioclim Historical

```{r}
all_bioclim_historical.pc = prcomp(bioclim_all_year_historical[c(7:16)], scale = TRUE, center = TRUE)
summary(all_bioclim_historical.pc)

all_bioclim_historical.pc.dat = data.frame(all_bioclim_historical.pc$x)

all_bioclim_historical_locs.pc = cbind(bioclim_all_year_historical, all_bioclim_historical.pc.dat)

all_bioclim_historical_loadings = data.frame(varnames=rownames(all_bioclim_historical.pc$rotation), all_bioclim_historical.pc$rotation)

autoplot(all_bioclim_historical.pc, all_bioclim_historical_loadings=TRUE, loadings.label=TRUE, 
         loadings.colour="blue", loadings.label.colour="red",  loadings.label.size = 5)
#this method results in different amounts of variation explained by PCs compared to princomp

autoplot(all_bioclim_historical.pc, data = bioclim_all_year_historical,
         colour='elevation.group', loadings=TRUE,
         loadings.colour='black', loadings.label = TRUE)

all_bioclim_historical_locs.pc %>% 
  ggplot(aes(x = PC1, y = PC2, colour=elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") 
```

## Yearly Flint Avgs + bioclim 

Calculate avgs

```{r}
flint_all_year_recent_yrlyavgs <- flint_all_year_recent %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, year) %>% 
  summarise_at(c("cwd",  "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(flint_all_year_recent_yrlyavgs) <- gsub("fn2", "sem", colnames(flint_all_year_recent_yrlyavgs))
names(flint_all_year_recent_yrlyavgs) <-gsub("fn1", "mean", colnames(flint_all_year_recent_yrlyavgs))
flint_all_year_recent_yrlyavgs

flint_all_year_historical_yrlyavgs <- flint_all_year_historical %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, year) %>% 
  summarise_at(c("cwd",  "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(flint_all_year_historical_yrlyavgs) <- gsub("fn2", "sem", colnames(flint_all_year_historical_yrlyavgs))
names(flint_all_year_historical_yrlyavgs) <-gsub("fn1", "mean", colnames(flint_all_year_historical_yrlyavgs))
flint_all_year_historical_yrlyavgs

```

Merge with bioclim
```{r}
flint_all_year_recent_yrlyavgs_prep <- flint_all_year_recent_yrlyavgs %>% select(parent.pop:cwd_mean, tmn_mean:tmx_mean) #removed ppt_mean because that parameter is included in bioclim vars 
bioclim_flint_all_year_recent <- full_join(flint_all_year_recent_yrlyavgs_prep, bioclim_all_year_recent)
head(bioclim_flint_all_year_recent)

flint_all_year_historical_yrlyavgs_prep <- flint_all_year_historical_yrlyavgs %>% select(parent.pop:cwd_mean, tmn_mean:tmx_mean)
bioclim_flint_all_year_historical <- full_join(flint_all_year_historical_yrlyavgs_prep, bioclim_all_year_historical)
head(bioclim_flint_all_year_historical)
```

Correlations - Recent

```{r}
#normalize the data
climate_normalized_bioclim_flint_all_year_recent <- bioclim_flint_all_year_recent %>% ungroup() %>% 
  select(cwd_mean:ppt_coldest_quarter) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_bioclim_flint_all_year_recent)
cor.norm = cor(climate_normalized_bioclim_flint_all_year_recent) #test correlations among the traits
corrplot(cor.norm)
```

Overlay populations on PCA plot - Recent

```{r}
all_bioclim_flint_recent.pc = prcomp(bioclim_flint_all_year_recent[c(7:19)], scale = TRUE, center = TRUE)
summary(all_bioclim_flint_recent.pc)

all_bioclim_flint_recent.pc.dat = data.frame(all_bioclim_flint_recent.pc$x)

all_bioclim_flint_recent_locs.pc = cbind(bioclim_flint_all_year_recent, all_bioclim_flint_recent.pc.dat)

all_bioclim_flint_recent_loadings = data.frame(varnames=rownames(all_bioclim_flint_recent.pc$rotation), all_bioclim_flint_recent.pc$rotation)

autoplot(all_bioclim_flint_recent.pc, all_bioclim_flint_recent_loadings=TRUE, loadings.label=TRUE, 
         loadings.colour="blue", loadings.label.colour="red",  loadings.label.size = 5)
#this method results in different amounts of variation explained by PCs compared to princomp

autoplot(all_bioclim_flint_recent.pc, data = bioclim_flint_all_year_recent,
         colour='elevation.group', loadings=TRUE,
         loadings.colour='black', loadings.label = TRUE)

all_bioclim_flint_recent_locs.pc %>% 
  ggplot(aes(x = PC1, y = PC2, colour=elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") 

```

Correlations - historical

```{r}
#normalize the data
climate_normalized_bioclim_flint_all_year_historical <- bioclim_flint_all_year_historical %>% ungroup() %>% 
  select(cwd_mean:ppt_coldest_quarter) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_bioclim_flint_all_year_historical)
cor.norm = cor(climate_normalized_bioclim_flint_all_year_historical) #test correlations among the traits
corrplot(cor.norm)
```

Overlay populations on PCA plot - historical

```{r}
all_bioclim_flint_historical.pc = prcomp(bioclim_flint_all_year_historical[c(7:19)], scale = TRUE, center = TRUE)
summary(all_bioclim_flint_historical.pc)

all_bioclim_flint_historical.pc.dat = data.frame(all_bioclim_flint_historical.pc$x)

all_bioclim_flint_historical_locs.pc = cbind(bioclim_flint_all_year_historical, all_bioclim_flint_historical.pc.dat)

all_bioclim_flint_historical_loadings = data.frame(varnames=rownames(all_bioclim_flint_historical.pc$rotation), all_bioclim_flint_historical.pc$rotation)

autoplot(all_bioclim_flint_historical.pc, all_bioclim_flint_historical_loadings=TRUE, loadings.label=TRUE, 
         loadings.colour="blue", loadings.label.colour="red",  loadings.label.size = 5)
#this method results in different amounts of variation explained by PCs compared to princomp

autoplot(all_bioclim_flint_historical.pc, data = bioclim_flint_all_year_historical,
         colour='elevation.group', loadings=TRUE,
         loadings.colour='black', loadings.label = TRUE)

all_bioclim_flint_historical_locs.pc %>% 
  ggplot(aes(x = PC1, y = PC2, colour=elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +

  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") 

```

## Avg across years and months (Flint)
Calculate avgs

```{r}
flint_all_year_recent_avgs <- flint_all_year_recent %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd",  "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(flint_all_year_recent_avgs) <- gsub("fn2", "sem", colnames(flint_all_year_recent_avgs))
names(flint_all_year_recent_avgs) <-gsub("fn1", "mean", colnames(flint_all_year_recent_avgs))
flint_all_year_recent_avgs

flint_all_year_historical_avgs <- flint_all_year_historical %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd",  "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(flint_all_year_historical_avgs) <- gsub("fn2", "sem", colnames(flint_all_year_historical_avgs))
names(flint_all_year_historical_avgs) <-gsub("fn1", "mean", colnames(flint_all_year_historical_avgs))
flint_all_year_historical_avgs

```

Correlations - Recent

```{r}
#normalize the data
climate_normalized_flint_all_year_recent_avgs <- flint_all_year_recent_avgs %>% ungroup() %>% 
  select(cwd_mean:tmx_mean) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_flint_all_year_recent_avgs)
cor.norm = cor(climate_normalized_flint_all_year_recent_avgs) #test correlations among the traits
corrplot(cor.norm)
```

Overlay populations on PCA plot - Recent

```{r}
all_flint_avgs_recent.pc = prcomp(flint_all_year_recent_avgs[c(6:9)], scale = TRUE, center = TRUE)
summary(all_flint_avgs_recent.pc)

all_flint_avgs_recent.pc.dat = data.frame(all_flint_avgs_recent.pc$x)

all_flint_avgs_recent_locs.pc = cbind(flint_all_year_recent_avgs, all_flint_avgs_recent.pc.dat)

all_flint_avgs_recent_loadings = data.frame(varnames=rownames(all_flint_avgs_recent.pc$rotation), all_flint_avgs_recent.pc$rotation)

autoplot(all_flint_avgs_recent.pc, all_flint_avgs_recent_loadings=TRUE, loadings.label=TRUE, 
         loadings.colour="blue", loadings.label.colour="red",  loadings.label.size = 5)
#this method results in different amounts of variation explained by PCs compared to princomp

autoplot(all_flint_avgs_recent.pc, data = flint_all_year_recent_avgs,
         colour='elevation.group', loadings=TRUE,
         loadings.colour='black', loadings.label = TRUE)

all_flint_avgs_recent_locs.pc %>% 
  ggplot(aes(x = PC1, y = PC2, colour=elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_text_repel(aes(label = parent.pop),
                  min.segment.length = 0) +
  geom_segment(data = all_flint_avgs_recent_loadings, 
               aes(x = 0, y = 0, xend = (PC1*5), yend = (PC2*5)), 
               arrow = arrow(length = unit(1/2, "picas")), color = "black") +
  annotate("text", x = (all_flint_avgs_recent_loadings$PC1*5), y = (all_flint_avgs_recent_loadings$PC2*5),
     label = all_flint_avgs_recent_loadings$varnames) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") 

```

Correlations - historical

```{r}
#normalize the data
climate_normalized_flint_all_year_historical_avgs <- flint_all_year_historical_avgs %>% ungroup() %>% 
  select(cwd_mean:tmx_mean) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_flint_all_year_historical_avgs)
cor.norm = cor(climate_normalized_flint_all_year_historical_avgs) #test correlations among the traits
corrplot(cor.norm)
```

Overlay populations on PCA plot - historical

```{r}
all_flint_avgs_historical.pc = prcomp(flint_all_year_historical_avgs[c(6:9)], scale = TRUE, center = TRUE)
summary(all_flint_avgs_historical.pc)

all_flint_avgs_historical.pc.dat = data.frame(all_flint_avgs_historical.pc$x)

all_flint_avgs_historical_locs.pc = cbind(flint_all_year_historical_avgs, all_flint_avgs_historical.pc.dat)

all_flint_avgs_historical_loadings = data.frame(varnames=rownames(all_flint_avgs_historical.pc$rotation), all_flint_avgs_historical.pc$rotation)

autoplot(all_flint_avgs_historical.pc, all_flint_avgs_historical_loadings=TRUE, loadings.label=TRUE, 
         loadings.colour="blue", loadings.label.colour="red",  loadings.label.size = 5)
#this method results in different amounts of variation explained by PCs compared to princomp

autoplot(all_flint_avgs_historical.pc, data = flint_all_year_historical_avgs,
         colour='elevation.group', loadings=TRUE,
         loadings.colour='black', loadings.label = TRUE)

all_flint_avgs_historical_locs.pc %>% 
  ggplot(aes(x = PC1, y = PC2, colour=elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_text_repel(aes(label = parent.pop),
                  min.segment.length = 0) +
  geom_segment(data = all_flint_avgs_historical_loadings, 
               aes(x = 0, y = 0, xend = (PC1*5), yend = (PC2*5)), 
               arrow = arrow(length = unit(1/2, "picas")), color = "black") +
  annotate("text", x = (all_flint_avgs_historical_loadings$PC1*5), y = (all_flint_avgs_historical_loadings$PC2*5),
     label = all_flint_avgs_historical_loadings$varnames) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") 
```

## Avg across years and months (Flint + bioclim)

Calculate avgs

```{r}
bioclim_all_year_recent_avgs <- bioclim_all_year_recent %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_quarter", "tmean_driest_quarter", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_quarter", "ppt_coldest_quarter"),
               c(mean, sem), na.rm = TRUE) 
names(bioclim_all_year_recent_avgs) <- gsub("fn2", "sem", colnames(bioclim_all_year_recent_avgs))
names(bioclim_all_year_recent_avgs) <-gsub("fn1", "mean", colnames(bioclim_all_year_recent_avgs))
bioclim_all_year_recent_avgs

bioclim_all_year_historical_avgs <- bioclim_all_year_historical %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_quarter", "tmean_driest_quarter", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_quarter", "ppt_coldest_quarter"),
               c(mean, sem), na.rm = TRUE) 
names(bioclim_all_year_historical_avgs) <- gsub("fn2", "sem", colnames(bioclim_all_year_historical_avgs))
names(bioclim_all_year_historical_avgs) <-gsub("fn1", "mean", colnames(bioclim_all_year_historical_avgs))
bioclim_all_year_historical_avgs
```

Merge with flint
```{r}
flint_all_year_recent_avgs_prep <- flint_all_year_recent_avgs %>% select(parent.pop:cwd_mean, tmn_mean:tmx_mean) #removed ppt_mean because that parameter is included in bioclim vars 
bioclim_flint_all_year_recent_avgs <- full_join(flint_all_year_recent_avgs_prep, bioclim_all_year_recent_avgs)
head(bioclim_flint_all_year_recent_avgs)

flint_all_year_historical_avgs_prep <- flint_all_year_historical_avgs %>% select(parent.pop:cwd_mean, tmn_mean:tmx_mean) #removed ppt_mean because that parameter is included in bioclim vars 
bioclim_flint_all_year_historical_avgs <- full_join(flint_all_year_historical_avgs_prep, bioclim_all_year_historical_avgs)
head(bioclim_flint_all_year_historical_avgs)
```


Correlations - Recent

```{r}
#normalize the data
climate_normalized_bioclim_flint_all_year_recent_avgs <- bioclim_flint_all_year_recent_avgs %>% ungroup() %>% 
  select(cwd_mean:ppt_coldest_quarter_mean) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_bioclim_flint_all_year_recent_avgs)
cor.norm = cor(climate_normalized_bioclim_flint_all_year_recent_avgs) #test correlations among the traits
corrplot(cor.norm)
```

Overlay populations on PCA plot - Recent

```{r}
all_bioclim_flint_avgs_recent.pc = prcomp(bioclim_flint_all_year_recent_avgs[c(6:18)], scale = TRUE, center = TRUE)
summary(all_bioclim_flint_avgs_recent.pc)

all_bioclim_flint_avgs_recent.pc.dat = data.frame(all_bioclim_flint_avgs_recent.pc$x)

all_bioclim_flint_avgs_recent_locs.pc = cbind(bioclim_flint_all_year_recent_avgs, all_bioclim_flint_avgs_recent.pc.dat)

all_bioclim_flint_avgs_recent_loadings = data.frame(varnames=rownames(all_bioclim_flint_avgs_recent.pc$rotation), all_bioclim_flint_avgs_recent.pc$rotation)

autoplot(all_bioclim_flint_avgs_recent.pc, all_bioclim_flint_avgs_recent_loadings=TRUE, loadings.label=TRUE, 
         loadings.colour="blue", loadings.label.colour="red",  loadings.label.size = 5)
#this method results in different amounts of variation explained by PCs compared to princomp

autoplot(all_bioclim_flint_avgs_recent.pc, data = bioclim_flint_all_year_recent_avgs,
         colour='elevation.group', loadings=TRUE,
         loadings.colour='black', loadings.label = TRUE)

all_bioclim_flint_avgs_recent_locs.pc %>% 
  ggplot(aes(x = PC1, y = PC2, colour=elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_text_repel(aes(label = parent.pop),
                  min.segment.length = 0) +
  geom_segment(data = all_bioclim_flint_avgs_recent_loadings, 
               aes(x = 0, y = 0, xend = (PC1*5), yend = (PC2*5)), 
               arrow = arrow(length = unit(1/2, "picas")), color = "black") +
  annotate("text", x = (all_bioclim_flint_avgs_recent_loadings$PC1*5), y = (all_bioclim_flint_avgs_recent_loadings$PC2*5),
     label = all_bioclim_flint_avgs_recent_loadings$varnames) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") 

```

Correlations - historical

```{r}
#normalize the data
climate_normalized_bioclim_flint_all_year_historical_avgs <- bioclim_flint_all_year_historical_avgs %>% ungroup() %>% 
  select(cwd_mean:ppt_coldest_quarter_mean) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_bioclim_flint_all_year_historical_avgs)
cor.norm = cor(climate_normalized_bioclim_flint_all_year_historical_avgs) #test correlations among the traits
corrplot(cor.norm)
```

Overlay populations on PCA plot - historical

```{r}
all_bioclim_flint_avgs_historical.pc = prcomp(bioclim_flint_all_year_historical_avgs[c(6:18)], scale = TRUE, center = TRUE)
summary(all_bioclim_flint_avgs_historical.pc)

all_bioclim_flint_avgs_historical.pc.dat = data.frame(all_bioclim_flint_avgs_historical.pc$x)

all_bioclim_flint_avgs_historical_locs.pc = cbind(bioclim_flint_all_year_historical_avgs, all_bioclim_flint_avgs_historical.pc.dat)

all_bioclim_flint_avgs_historical_loadings = data.frame(varnames=rownames(all_bioclim_flint_avgs_historical.pc$rotation), all_bioclim_flint_avgs_historical.pc$rotation)

autoplot(all_bioclim_flint_avgs_historical.pc, all_bioclim_flint_avgs_historical_loadings=TRUE, loadings.label=TRUE, 
         loadings.colour="blue", loadings.label.colour="red",  loadings.label.size = 5)
#this method results in different amounts of variation explained by PCs compared to princomp

autoplot(all_bioclim_flint_avgs_historical.pc, data = bioclim_flint_all_year_historical_avgs,
         colour='elevation.group', loadings=TRUE,
         loadings.colour='black', loadings.label = TRUE)

all_bioclim_flint_avgs_historical_locs.pc %>% 
  ggplot(aes(x = PC1, y = PC2, colour=elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_text_repel(aes(label = parent.pop),
                  min.segment.length = 0) +
  geom_segment(data = all_bioclim_flint_avgs_historical_loadings, 
               aes(x = 0, y = 0, xend = (PC1*5), yend = (PC2*5)), 
               arrow = arrow(length = unit(1/2, "picas")), color = "black") +
  annotate("text", x = (all_bioclim_flint_avgs_historical_loadings$PC1*5), y = (all_bioclim_flint_avgs_historical_loadings$PC2*5),
     label = all_bioclim_flint_avgs_historical_loadings$varnames) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") 
```