---
title: "Climate Data Prep"
author: "Brandie Quarles"
date: "2023-10-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Relevant Libraries and Functions
```{r}
library(tidyverse)
library(corrplot) #plotting correlations 
library(rstatix) #performing cor_test
library(lmerTest)
library(naniar) #replaces values with NA
sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")

#For scree plots 
#library("devtools") #The package devtools is required for the installation as factoextra is hosted on github.
#1install_github("kassambara/factoextra")
library("factoextra")
```

#Climate Data (1895-2022)
- aet (mm) = actual evapotranspiration (water transpired by plants, if water not limited)
- cwd (mm) = climate water deficit (estimate of drought) = pet-aet
- pck (mm) = snowpack (prior month pck + snow - lost snow)
- pet (mm) = potential evapotranspiration (total amount of water that can evaporate from the ground or be trasnpired by plants)
- ppt (mm) = precipitation
- rch (mm) = recharge (water that penetrates below the root zone)
- run (mm) = runoff (water that becomes streamflow)
- str (mm) = soil water storage (avg amount of water stored in the soil annually)
all above are totals/sums per month 
- tmn (deg C) = min air temp (for a given month)
- tmx (deg C) = max air temp (for a given month)
127 years


Notes
- did some lit review to see what variables other people have focused on or found to be significant (CWD, temp, precip, AET)
-Try PCR (principal components regression)
-Temperature highly correlated within pops w/in years - take all variables (including avg temp) for the 3 months when we think germination is occurring (Low: Oct-Jan, Mid: Nov-Jan, High: April-June)

#Load the data 
```{r}
climate_data <- read_csv("../input/Dimensions_All_1895-2022.csv")
head(climate_data)
names(climate_data)
unique(climate_data$pop)
climate_data$year = as.character(climate_data$year)
head(climate_data)

climate_data_yo <- climate_data %>% mutate(pop = str_replace(pop, "YOSE(\\d+)", "YO\\1")) #changing YOSE to YO 
#`(\\d+)`: This part contains a regular expression within parentheses: 
#`\\d`: Matches a digit (equivalent to `[0-9]`). 
#The double backslash `\\` is used to escape the `\` #character in R strings, so `\\d` is interpreted as `\d` in the regex, denoting a digit.
#`+`: This qualifier means "one or more of the preceding element," so `\\d+` will match one or more digits.
#`()`: Parentheses are used to define a capturing group, which means that the part of the regex within the parentheses `(\\d+)` is captured to be used in back-references or for extracting matched parts.
#\\1 references the first capturing group
unique(climate_data_yo$pop)

#pop info
pop_elev <- read_csv("../input/UCD_Data/Pops_for_2022_UCD.csv") #elevation info for each population 
summary(pop_elev)
pop_elev_nonotes <- pop_elev %>% select(-notes, -JGI.DNA)
pop_elev_nonotes
pop_elev_nonotes$elevation.group <- str_to_title(pop_elev_nonotes$elevation.group)

#names(pop_elev_nonotes)
pop_elev_climate <- left_join(pop_elev_nonotes, climate_data_yo, by=c("parent.pop"="pop")) %>% select(parent.pop, elevation.group, year:tmx)
unique(pop_elev_climate$parent.pop)
#unique(pop_elev_nonotes$parent.pop)
head(pop_elev_climate, 30)
names(pop_elev_climate)
```

#Calculation of recent (last 30 years) and historical climate (prior 30 years)
```{r}
pop_elev_climate_recent <- pop_elev_climate %>% filter(year>1992)
head(pop_elev_climate_recent)
pop_elev_climate_historical <- pop_elev_climate %>% filter(year<=1992 & year>1962)
pop_elev_climate_historical
```


#Climate across pops (during germ season)!
```{r}

```


#Climate traits across pops 
```{r}
pop_elev_climate_yearlytot <- pop_elev_climate %>% group_by(parent.pop, year, elevation.group) %>% summarise_at(c("pck", "ppt"), sum, na.rm = TRUE)
names(pop_elev_climate_yearlytot)
pop_elev_climate_yearlytot %>% filter(year>1992) %>% ggplot(aes(x=year, y=pck, group=parent.pop, color=elevation.group)) + geom_point() + geom_line() + scale_colour_manual(values=elev_three_palette)
pop_elev_climate_yearlytot %>% filter(year>1992) %>% ggplot(aes(x=year, y=ppt, group=parent.pop, color=elevation.group)) + geom_point() + geom_line() + scale_colour_manual(values=elev_three_palette)

pop_elev_climate_avgs <- pop_elev_climate %>% filter(year>1992) %>% group_by(parent.pop, elevation.group) %>% summarise_at(c("aet", "cwd", "pck", "pet", "ppt", "rch", "run", "str", "tmn", "tmx"), mean, na.rm = TRUE)
pop_elev_climate_avgs #30 year averages of all climate variables 
pop_elev_climate_avgs$elevation.group <- factor(pop_elev_climate_avgs$elevation.group, levels=elev_order)

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, aet), y=aet, fill=elevation.group)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, cwd), y=cwd, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, pck), y=pck, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, pet), y=pet, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt), y=ppt, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, rch), y=rch, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, run), y=run, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, str), y=str, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1)) #this could be a better metric of available water (compared to ppt)
#TM2 may be higher b/c of clay in the area that was being mapped (off the outcrops)

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmn), y=tmn, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmx), y=tmx, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))
```

#Climate trait correlations 
```{r}
names(pop_elev_climate)
pop_elev_climate %>% cor_test(aet, cwd, pck, pet, ppt, rch, run, str, tmn, tmx, method="pearson")

cor.mat <- pop_elev_climate %>% select(aet, cwd, pck, pet, ppt, rch, run, str, tmn, tmx) %>% cor_mat()
cor.mat
	
mydata.cor = pop_elev_climate %>% select(aet, cwd, pck, pet, ppt, rch, run, str, tmn, tmx) %>% cor()
mydata.cor
corrplot(mydata.cor)

#PCA
#normalize the data
climate_normalized <- pop_elev_climate %>% select(aet, cwd, pck, pet, ppt, rch, run, str, tmn, tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized)
cor.norm = cor(climate_normalized) #test correlations among the traits
corrplot(cor.norm)

climate.pca <- princomp(cor.norm) 
summary(climate.pca) #PC1 explains 80% of the variation, PC2 explains 7.9%, together they explain 88%
climate.pca$loadings[, 1:2] #PC1 has high positive loading values for tmn, tmx, pet, cwd, aet
fviz_eig(climate.pca) #scree plot 
fviz_pca_var(climate.pca, col.var = "black") #biplot

#The goal of the third visualization is to determine how much each variable is represented in a given component. Such a quality of representation is called the Cos2 and corresponds to the square cosine, and it is computed using the fviz_cos2 function. (https://www.datacamp.com/tutorial/pca-analysis-r)
#A low value means that the variable is not perfectly represented by that component. 
#A high value, on the other hand, means a good representation of the variable on that component.
fviz_cos2(climate.pca, choice = "var", axes = 1:2) #axes says which PCs to use 
#biplot combined wiht cos2
fviz_pca_var(climate.pca, col.var = "cos2", 
            gradient.cols = c("black", "orange", "green"),
            repel = TRUE)

#Should use this code to check the significance of the PCA: https://github.com/StatQuest/pca_demo/blob/master/pca_demo.R
#Remember this paper: Björklund, M. 2019. Be careful with your principal components. Evolution 73: 2151–2158.

```