---
title: "Manuscript-Figures"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Figures for WL2 Climate Distance Manuscript

## Libraries
```{r}
library(tidyverse)
```

## Load Data
```{r}
wl2_establishment <- read_csv("../output/WL2_Traits/WL2_Establishment.csv")
wl2_y1surv <- read_csv("../output/WL2_Traits/WL2_Y1Surv.csv")
winter_surv <- read_csv("../output/WL2_Traits/WL2_WinterSurv_2324.csv")
wl2_surv_to_bud_y2 <- read_csv("../output/WL2_Traits/WL2_Surv_to_Rep_Y2_2324.csv")
wl2_prob_rep
```

## Figure 2: Geo distance

## Figure 3: Gower's recent water year climate distance

## Figure 4: Temp and PPT recent water climate distance

## Figure S5: Additional Gower's recent water year climate distance 

## Figure S6: Additonal Temp and PPT recent water yar climate distance

## Additional figures (come back to these)

-   Figure 1: Population map and PCAs

-   Figure S1: Pearson's correlations

-   Figure S2: PCAs with PC1 and PC2

-   Figure S3: Gower's climate distance across pops

-   Figure S4: Establishment relat with geo distance

-   Figure S7: Growth season climate distance effects
