---
title: "Climate_Shapefile"
output: html_document
date: "2024-10-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Create a shapefile for USGS to extract the climate data for all populations and common garden sites

## Libraries
```{r}
library(tidyverse)
```

## Load the pop and location data

```{r}
#pop info
pops_common_garden <- read_csv("../input/UCD_Data/Pops_for_2022_UCD.csv") #pops included in common garden 
names(pops_common_garden)
pops_common_garden_nonotes <- pops_common_garden %>% select(parent.pop:phylogroup)
head(pops_common_garden_nonotes)

yose1 <- tibble(parent.pop="YO1", phylogroup=NA) #Need to add YO1 for JGI

pops_common_garden_jgi <- bind_rows(pops_common_garden_nonotes, yose1)

#extra location info 
pop_loc <- read_csv("../input/Strep_tort_locs.csv")
names(pop_loc)
head(pop_loc)

unique(pop_loc$`Site code`)
unique(pops_common_garden_jgi$parent.pop)
#need to change YOSE to YO
pop_loc_yo <- pop_loc %>% mutate(parent.pop = str_replace(`Site code`, "YOSE(\\d+)", "YO\\1")) %>% select(Lat, Long, elev_m=`Elevation (m)`, parent.pop)
unique(pop_loc_yo$parent.pop)

#merge in location info
pop_elev <- left_join(pops_common_garden_jgi, pop_loc_yo)
tail(pop_elev)
```

## Common Garden Sites Lat/Long
```{r}
#38.532504, -121.782986 - Davis Veg Crops
#38.82599, -120.25090 - WL2 Garden

sites_loc <- tibble(parent.pop=c("UCD_Garden", "WL2_Garden"), phylogroup=NA, Lat=c("38.53250", "38.82599"), Long=c("-121.78299", "-120.25090"), elev_m=c(16,2020))
sites_loc
```

## Make coordinates file that will be converted into a shapefile 
```{r}
pops_gardens_locs <- bind_rows(pop_elev, sites_loc)
pops_gardens_locs

pops_gardens_coord <- pops_gardens_locs %>% select(Lat:Long)
#write_csv(pops_gardens_coord, "../output/Climate/Shapefile_coords.csv")
```

## Connecting pop ids to new Flint data

### add row names to pops_gardens
```{r}
pops_gardens_locs_rowids <- pops_gardens_locs %>% rowid_to_column() %>% mutate(rowid = as.character(rowid))
```

### import new data 
Note: This only goes through Aug 2024 - will need to reach back out to get Sept and Oct 2024 probably in December 2024?
```{r}
new_flint_data <- read_csv("../input/Updated_Flint_Climate_Oct2024.csv")
new_flint_data
```

```{r}
new_cwd <- new_flint_data %>% filter(str_detect(
DataFileName, "cwd"))
head(new_cwd)
tail(new_cwd)
new_cwd %>% filter(str_detect(DataFileName, "2024"))

new_cwd_long <- new_cwd %>% pivot_longer(!DataFileName, names_to = 
"rowid", values_to = "cwd") %>% extract(DataFileName, c("year", "month"), "(\\d+)(.*)", convert = TRUE) %>% separate(month, into = c("month", "file_type")) %>% select(rowid, year, month, cwd)
head(new_cwd_long, 30)
```

```{r}
new_pet <- new_flint_data %>% filter(str_detect(
DataFileName, "pet"))
head(new_pet)
tail(new_pet)

new_pet_long <- new_pet %>% pivot_longer(!DataFileName, names_to = 
"rowid", values_to = "pet") %>% extract(DataFileName, c("year", "month"), "(\\d+)(.*)", convert = TRUE) %>% separate(month, into = c("month", "file_type")) %>% select(rowid, year, month, pet)
head(new_pet_long, 30)
```

```{r}
new_aet <- new_flint_data %>% filter(str_detect(
DataFileName, "aet"))
head(new_aet)
tail(new_aet)

new_aet_long <- new_aet %>% pivot_longer(!DataFileName, names_to = 
"rowid", values_to = "aet") %>% extract(DataFileName, c("year", "month"), "(\\d+)(.*)", convert = TRUE) %>% separate(month, into = c("month", "file_type")) %>% select(rowid, year, month, aet)
head(new_aet_long, 30)
```

```{r}
new_pck <- new_flint_data %>% filter(str_detect(
DataFileName, "pck"))
head(new_pck)
tail(new_pck)

new_pck_long <- new_pck %>% pivot_longer(!DataFileName, names_to = 
"rowid", values_to = "pck") %>% extract(DataFileName, c("year", "month"), "(\\d+)(.*)", convert = TRUE) %>% separate(month, into = c("month", "file_type")) %>% select(rowid, year, month, pck)
head(new_pck_long, 30)
```

```{r}
new_ppt <- new_flint_data %>% filter(str_detect(
DataFileName, "ppt"))
head(new_ppt)
tail(new_ppt)

new_ppt_long <- new_ppt %>% pivot_longer(!DataFileName, names_to = 
"rowid", values_to = "ppt") %>% extract(DataFileName, c("year", "month"), "(\\d+)(.*)", convert = TRUE) %>% separate(month, into = c("month", "file_type")) %>% select(rowid, year, month, ppt)
head(new_ppt_long, 30)
```

```{r}
new_tmn <- new_flint_data %>% filter(str_detect(
DataFileName, "tmn"))
head(new_tmn)
tail(new_tmn)

new_tmn_long <- new_tmn %>% pivot_longer(!DataFileName, names_to = 
"rowid", values_to = "tmn") %>% extract(DataFileName, c("year", "month"), "(\\d+)(.*)", convert = TRUE) %>% separate(month, into = c("month", "file_type")) %>% select(rowid, year, month, tmn)
head(new_tmn_long, 30)
```

```{r}
new_tmx <- new_flint_data %>% filter(str_detect(
DataFileName, "tmx"))
head(new_tmx)
tail(new_tmx)

new_tmx_long <- new_tmx %>% pivot_longer(!DataFileName, names_to = 
"rowid", values_to = "tmx") %>% extract(DataFileName, c("year", "month"), "(\\d+)(.*)", convert = TRUE) %>% separate(month, into = c("month", "file_type")) %>% select(rowid, year, month, tmx)
head(new_tmx_long, 30)
```

Merge
```{r}
new_flint_long <- full_join(new_aet_long, new_cwd_long)
new_flint_long <- full_join(new_flint_long, new_pck_long)
new_flint_long <- full_join(new_flint_long, new_pet_long)
new_flint_long <- full_join(new_flint_long, new_ppt_long)
new_flint_long <- full_join(new_flint_long, new_tmn_long)
new_flint_long <- full_join(new_flint_long, new_tmx_long)
head(new_flint_long)


new_flint_pops <- full_join(pops_gardens_locs_rowids, new_flint_long)
head(new_flint_pops)
write_csv(new_flint_pops, "../output//Flint_AllSites_Aug2024.csv")
```

