---
title: "Climate PCAs_Growth Season"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To Do:
- Save PC figures of interest 

# Climate PCAs

-   Should use this code to check the significance of the PCA:
    <https://github.com/StatQuest/pca_demo/blob/master/pca_demo.R>
-   Remember this paper: Björklund, M. 2019. Be careful with your
    principal components. Evolution 73: 2151--2158.

-   PCAs for climate during the growth season

    -   Color monthly PCA by month relative to growth season?
    -   Bioclim vars only PCA okay? or combine with other Flint
        variables? Hard to combine b/c Flint data is for each month
        within a year and Bioclim variables are yearly summaries. Could
        do one with just Flint with all months in each year included and
        one with Flint growth season yearly averages combined with
        Bioclim variables
 
## Load necessary libraries 
```{r}
library(tidyverse)
library(ggrepel)
#library(cowplot)
library(gridExtra)
library(corrplot) #plotting correlations 
library(rstatix) #performing cor_test
library(QBMS) #for function calc_biovars to calculate bioclim variables
library(ggfortify) #easier PCA figures
sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
} #legend function for grid_arrange

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")
month_order <- c("jan","feb","mar","apr","may","jun","jul","aug","sep","oct","nov","dec")
```

## Load the Flint growth season data (from "Flint_Growth_Season.Rmd")
```{r}
flint_recent_grwseason <- read_csv("../output/Climate/flint_climate_growthseason_recent.csv") %>%
  dplyr::select(month, parent.pop:tmx)
head(flint_recent_grwseason, 12)
flint_recent_grwseason %>% filter(parent.pop=="DPR")

flint_historical_grwseason <- read_csv("../output/Climate/flint_climate_growthseason_historical.csv") %>%
  dplyr::select(month, parent.pop:tmx)
head(flint_historical_grwseason, 12)
```

## Load the Bioclim growth season data (from "BioClim_Growth_Season.Rmd")
```{r}
bioclim_recent_grwseason <- read_csv("../output/Climate/BioClim_growthseason_Recent.csv")
bioclim_historical_grwseason <- read_csv("../output/Climate/BioClim_growthseason_Historical.csv")
```


## All years and months included (Flint)

### Correlations - Flint Recent

```{r}
#normalize the data
climate_normalized_all_flint_recent <- flint_recent_grwseason %>% select(cwd, ppt, tmn, tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_flint_recent)
cor.norm = cor(climate_normalized_all_flint_recent) #test correlations among the traits
corrplot(cor.norm)
#tmn and tmx highly correlated (95%), consider removing one 
```

### PCA - Flint Recent

```{r}
all_flint_recent.pc = prcomp(flint_recent_grwseason[c(9, 11:12)], scale = TRUE, center = TRUE)

str(all_flint_recent.pc)
```

plot % Variance Explained
```{r}
summary(all_flint_recent.pc)

tibble(PC=str_c("PC",str_pad(1:3,2,pad="0")),
       percent_var=all_flint_recent.pc$sdev[1:3]^2/sum(all_flint_recent.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata
```{r}
all_flint_recent.pc.dat = data.frame(all_flint_recent.pc$x)

all_flint_recent_locs.pc = cbind(flint_recent_grwseason, all_flint_recent.pc.dat)

all_flint_recent_loadings = data.frame(varnames=rownames(all_flint_recent.pc$rotation), all_flint_recent.pc$rotation)
all_flint_recent_loadings
```

```{r}
autoplot(all_flint_recent.pc, data = flint_recent_grwseason,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=8, loadings.label.colour="black", loadings.label.vjust = -0.2) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()

#for plot customizations see: ?ggbiplot
```

PCs 1 and 3
```{r}
autoplot(all_flint_recent.pc, data = flint_recent_grwseason,
         x=1, y=3,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=8, loadings.label.colour="black", loadings.label.vjust = -0.2) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

### Correlations - Flint Historical

```{r}
#normalize the data
climate_normalized_all_flint_historical <- flint_historical_grwseason %>% select(cwd, ppt, tmn, tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_flint_historical)
cor.norm = cor(climate_normalized_all_flint_historical) #test correlations among the traits
corrplot(cor.norm)
#tmn and tmx highly correlated, consider removing one 
```

### PCA - Flint Historical

```{r}
all_flint_historical.pc = prcomp(flint_historical_grwseason[c(9, 11:12)], scale = TRUE, center = TRUE)

str(all_flint_historical.pc)
```

plot % Variance Explained
```{r}
summary(all_flint_historical.pc)

tibble(PC=str_c("PC",str_pad(1:3,2,pad="0")),
       percent_var=all_flint_historical.pc$sdev[1:3]^2/sum(all_flint_historical.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata
```{r}
all_flint_historical.pc.dat = data.frame(all_flint_historical.pc$x)

all_flint_historical_locs.pc = cbind(flint_historical_grwseason, all_flint_historical.pc.dat)

all_flint_historical_loadings = data.frame(varnames=rownames(all_flint_historical.pc$rotation), all_flint_historical.pc$rotation)
all_flint_historical_loadings
```

```{r}
autoplot(all_flint_historical.pc, data = flint_historical_grwseason,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=8, loadings.label.colour="black", loadings.label.vjust = -0.2) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

PCs 1 and 3
```{r}
autoplot(all_flint_historical.pc, data = flint_historical_grwseason,
         x=1, y=3,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=8, loadings.label.colour="black", loadings.label.vjust = -0.2) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

## All years included (bioclim)

### Correlations - bioclim Recent

```{r}
#normalize the data
climate_normalized_all_bioclim_recent <- bioclim_recent_grwseason %>% select(ann_tmean:ppt_coldest_month) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_bioclim_recent)
cor.norm = cor(climate_normalized_all_bioclim_recent) #test correlations among the traits
corrplot(cor.norm)
```

### PCA - bioclim Recent

```{r}
all_bioclim_recent.pc = prcomp(bioclim_recent_grwseason[c(5:14)], scale = TRUE, center = TRUE) 

str(all_bioclim_recent.pc)
```

plot % Variance Explained
```{r}
summary(all_bioclim_recent.pc)

tibble(PC=str_c("PC",str_pad(1:10,2,pad="0")),
       percent_var=all_bioclim_recent.pc$sdev[1:10]^2/sum(all_bioclim_recent.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata
```{r}
all_bioclim_recent.pc.dat = data.frame(all_bioclim_recent.pc$x)

all_bioclim_recent_locs.pc = cbind(bioclim_recent_grwseason, all_bioclim_recent.pc.dat)

all_bioclim_recent_loadings = data.frame(varnames=rownames(all_bioclim_recent.pc$rotation), all_bioclim_recent.pc$rotation)
all_bioclim_recent_loadings
```

```{r}
autoplot(all_bioclim_recent.pc, data = bioclim_recent_grwseason,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

PCs 3 and 4
```{r}
autoplot(all_bioclim_recent.pc, data = bioclim_recent_grwseason,
         x=3, y=4,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

PCs 4 and 5
```{r}
autoplot(all_bioclim_recent.pc, data = bioclim_recent_grwseason,
         x=4, y=5,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

### Correlations - bioclim Historical

```{r}
#normalize the data
climate_normalized_all_bioclim_historical <- bioclim_historical_grwseason %>% select(ann_tmean:ppt_coldest_month) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_all_bioclim_historical)
cor.norm = cor(climate_normalized_all_bioclim_historical) #test correlations among the traits
corrplot(cor.norm)
```

### PCA - bioclim Historical

```{r}
all_bioclim_historical.pc = prcomp(bioclim_historical_grwseason[c(5:14)], scale = TRUE, center = TRUE) 

str(all_bioclim_historical.pc)
```

plot % Variance Explained
```{r}
summary(all_bioclim_historical.pc)

tibble(PC=str_c("PC",str_pad(1:10,2,pad="0")),
       percent_var=all_bioclim_historical.pc$sdev[1:10]^2/sum(all_bioclim_historical.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata
```{r}
all_bioclim_historical.pc.dat = data.frame(all_bioclim_historical.pc$x)

all_bioclim_historical_locs.pc = cbind(bioclim_historical_grwseason, all_bioclim_historical.pc.dat)

all_bioclim_historical_loadings = data.frame(varnames=rownames(all_bioclim_historical.pc$rotation), all_bioclim_historical.pc$rotation)
all_bioclim_historical_loadings
```

```{r}
autoplot(all_bioclim_historical.pc, data = bioclim_historical_grwseason,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

PCs 3 and 4
```{r}
autoplot(all_bioclim_historical.pc, data = bioclim_historical_grwseason,
         x=3, y=4,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

PCs 4 and 5
```{r}
autoplot(all_bioclim_historical.pc, data = bioclim_historical_grwseason,
         x=4, y=5,
         colour='elev_m', alpha=0.5,
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

## Monthly Averages - Flint  

### Calculate avgs

```{r}
flint_recent_grwseason_mosavgs <- flint_recent_grwseason %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, month) %>% 
  summarise_at(c("cwd", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Recent") 
flint_recent_grwseason_mosavgs

flint_historical_grwseason_mosavgs <- flint_historical_grwseason %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, month) %>% 
  summarise_at(c("cwd", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Historical")
flint_historical_grwseason_mosavgs

flint_grwseason_mosavgs <- bind_rows(flint_recent_grwseason_mosavgs, flint_historical_grwseason_mosavgs) #combine into 1 dataframe 
head(flint_grwseason_mosavgs)
tail(flint_grwseason_mosavgs)
```

### Correlations - Recent + Historical

```{r}
#normalize the data
climate_normalized_flint_grwseason_mosavgs <- flint_grwseason_mosavgs %>% ungroup() %>% 
  select(cwd:tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_flint_grwseason_mosavgs)
cor.norm = cor(climate_normalized_flint_grwseason_mosavgs) #test correlations among the traits
corrplot(cor.norm)

#tmn and tmx highly correlated, consider removing one (96%)
#tmx and ppt highly neg correlated (-82%)
```

### PCA - Recent + Historical

```{r}
#flint_grwseason_mosavgs[c(8:12)]
mos_flint.pc = prcomp(flint_grwseason_mosavgs[c(7:9)], scale = TRUE, center = TRUE)

str(mos_flint.pc)
```

plot % Variance Explained
```{r}
summary(mos_flint.pc)

tibble(PC=str_c("PC",str_pad(1:3,2,pad="0")),
       percent_var=mos_flint.pc$sdev[1:3]^2/sum(mos_flint.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata
```{r}
mos_flint.pc.dat = data.frame(mos_flint.pc$x)

mos_flint_locs.pc = cbind(flint_grwseason_mosavgs, mos_flint.pc.dat)

mos_flint_loadings = data.frame(varnames=rownames(mos_flint.pc$rotation), mos_flint.pc$rotation)
mos_flint_loadings
```

```{r}
autoplot(mos_flint.pc, data = flint_grwseason_mosavgs,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="month",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
#high elev seems most similar to low elev in summer months 
```


```{r}
mos_flint_locs.pc_avg <- mos_flint_locs.pc %>%
  group_by(parent.pop, elev_m, TimePd,month) %>%
  summarise(across(.cols=starts_with("PC"), .fns = mean)) %>%
  ungroup()

mos_flint_locs.pc_avg
```

```{r}
mos_flint_locs.pc_avg %>% 
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point() +
  facet_wrap(~month) +
  coord_fixed(ratio = 1.5)
#Remember: this is growth season so high elev pops don't have winter months 

```

```{r}
mos_flint_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m, month))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point() +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")))
#high elev climate seems to be shifting to be more similar to low elev 
#hard to see which months though 
```

## Avg across years and months (Flint)

### Calculate avgs

```{r}
flint_recent_grwseason_avgs <- flint_recent_grwseason %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Recent") 
flint_recent_grwseason_avgs

flint_historical_grwseason_avgs <- flint_historical_grwseason %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Historical")
flint_historical_grwseason_avgs

flint_grwseason_avgs <- bind_rows(flint_recent_grwseason_avgs, flint_historical_grwseason_avgs) #combine into 1 dataframe 
head(flint_grwseason_avgs)
tail(flint_grwseason_avgs)
write_csv(flint_grwseason_avgs, "../output/Climate/growthseason_FlintAvgs.csv")
```

### Correlations - Recent + Historical

```{r}
#normalize the data
climate_normalized_flint_grwseason_avgs <- flint_grwseason_avgs %>% ungroup() %>% 
  select(cwd:tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_flint_grwseason_avgs)
cor.norm = cor(climate_normalized_flint_grwseason_avgs) #test correlations among the traits
corrplot(cor.norm)

#tmn and tmx highly correlated, consider removing one (90%)
```

### PCA - Recent + Historical

```{r}
avgs_flint.pc = prcomp(flint_grwseason_avgs[c(6:8)], scale = TRUE, center = TRUE)

str(avgs_flint.pc)
```

plot % Variance Explained
```{r}
summary(avgs_flint.pc)

tibble(PC=str_c("PC",str_pad(1:3,2,pad="0")),
       percent_var=avgs_flint.pc$sdev[1:3]^2/sum(avgs_flint.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata
```{r}
avgs_flint.pc.dat = data.frame(avgs_flint.pc$x)

avgs_flint_locs.pc = cbind(flint_grwseason_avgs, avgs_flint.pc.dat)

avgs_flint_loadings = data.frame(varnames=rownames(avgs_flint.pc$rotation), avgs_flint.pc$rotation)
avgs_flint_loadings
```

```{r}
autoplot(avgs_flint.pc, data = flint_grwseason_avgs,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
avgs_flint_locs.pc_avg <- avgs_flint_locs.pc %>%
  group_by(parent.pop, elev_m, TimePd) %>%
  summarise(across(.cols=starts_with("PC"), .fns = mean)) %>%
  ungroup()

avgs_flint_locs.pc_avg
```

```{r}
avgs_flint_locs.pc_avg %>% 
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point() +
  facet_wrap(~TimePd) +
  coord_fixed(ratio = 1.5)
```

```{r}
avgs_flint_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)

avgs_flint_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)
#growth season climate has shifted for high elev way more than for low elev & directions are varied 
```

## Avg across years and months (Flint + bioclim)

### Calculate avgs

```{r}
bioclim_recent_grwseason_avgs <- bioclim_recent_grwseason %>% 
  group_by(parent.pop, elevation.group, elev_m) %>% 
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_month", "tmean_driest_month", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_month", "ppt_coldest_month"),
               c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Recent") 
bioclim_recent_grwseason_avgs

bioclim_historical_grwseason_avgs <- bioclim_historical_grwseason %>% 
  group_by(parent.pop, elevation.group, elev_m) %>% 
  summarise_at(c("ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_month", "tmean_driest_month", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_month", "ppt_coldest_month"),
               c(mean), na.rm = TRUE) %>% 
  mutate(TimePd = "Historical") 
bioclim_historical_grwseason_avgs

bioclim_grwseason_avgs <- bind_rows(bioclim_recent_grwseason_avgs, bioclim_historical_grwseason_avgs) #combine into 1 dataframe 
head(bioclim_grwseason_avgs)
tail(bioclim_grwseason_avgs)
write_csv(bioclim_grwseason_avgs, "../output/Climate/growthseason_BioClimAvgs.csv")
```

Merge with flint
```{r}
bioclim_flint_grwseason_avgs <- full_join(flint_grwseason_avgs, bioclim_grwseason_avgs) %>% 
  select(TimePd, parent.pop:ppt_coldest_month)
head(bioclim_flint_grwseason_avgs)
```


### Correlations - Recent + Historical

```{r}
#normalize the data
climate_normalized_bioclim_flint_grwseason_avgs <- bioclim_flint_grwseason_avgs %>% ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized_bioclim_flint_grwseason_avgs)
cor.norm = cor(climate_normalized_bioclim_flint_grwseason_avgs) #test correlations among the traits
cor.norm
corrplot(cor.norm)

#tmn, tmx, tmean_driest_month and ann_tmean all highly correlated (90-98%) - only keep ann_tmean 
```

### PCA - Recent + Historical

```{r}
all_bioclim_flint_avgs.pc = prcomp(bioclim_flint_grwseason_avgs[c(7, 11:15, 17:20)], scale = TRUE, center = TRUE) 
str(all_bioclim_flint_avgs.pc)
```

plot % Variance Explained
```{r}
summary(all_bioclim_flint_avgs.pc)

tibble(PC=str_c("PC",str_pad(1:10,2,pad="0")),
       percent_var=all_bioclim_flint_avgs.pc$sdev[1:10]^2/sum(all_bioclim_flint_avgs.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata
```{r}
all_bioclim_flint_avgs.pc.dat = data.frame(all_bioclim_flint_avgs.pc$x)

all_bioclim_flint_avgs_locs.pc = cbind(bioclim_flint_grwseason_avgs, all_bioclim_flint_avgs.pc.dat)

all_bioclim_flint_avgs_loadings = data.frame(varnames=rownames(all_bioclim_flint_avgs.pc$rotation), all_bioclim_flint_avgs.pc$rotation)
all_bioclim_flint_avgs_loadings
```

```{r}
autoplot(all_bioclim_flint_avgs.pc, data = bioclim_flint_grwseason_avgs,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
autoplot(all_bioclim_flint_avgs.pc, data = bioclim_flint_grwseason_avgs,
         x=1, y=3,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
autoplot(all_bioclim_flint_avgs.pc, data = bioclim_flint_grwseason_avgs,
         x=4, y=5,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
all_bioclim_flint_avgs_locs.pc_avg <- all_bioclim_flint_avgs_locs.pc %>%
  group_by(parent.pop, elev_m, TimePd) %>%
  summarise(across(.cols=starts_with("PC"), .fns = mean)) %>%
  ungroup()

all_bioclim_flint_avgs_locs.pc_avg
```

```{r}
all_bioclim_flint_avgs_locs.pc_avg %>% 
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point() +
  facet_wrap(~TimePd) +
  coord_fixed(ratio = 1.5)
```

```{r}
all_bioclim_flint_avgs_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)

all_bioclim_flint_avgs_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)
#high elev climate seems to be shifting to be more similar to low elev 
```

## LM on PCs
For each PC fit a model with timepd and parent.pop/elev to test for climate change
```{r}
lmer_results <- all_bioclim_flint_avgs_locs.pc %>%
  ungroup() %>% 
  select(TimePd, parent.pop, elev_m, Lat, Long, PC1:PC10) %>% 
  pivot_longer(starts_with("PC", ignore.case = FALSE), 
               names_to = "PC", values_to = "value") %>% 
  group_by(PC) %>% 
  nest(data=c(TimePd, parent.pop, elev_m, Lat, Long, value)) %>% 
  mutate(glm=map(data, ~ glm(value ~ TimePd*elev_m*Lat,
                               data=.x)),
         anova = map(glm, ~ broom.mixed::tidy(anova(.x))))

lmer_results
```

```{r}
PC_anova <- lmer_results %>% select(-data, -glm) %>% unnest(anova) %>%
  select(PC, term, p.value) %>%
  filter(p.value < 0.05) %>%
  arrange(term, p.value)

PC_anova
```

```{r}
autoplot(all_bioclim_flint_avgs.pc, data = bioclim_flint_grwseason_avgs,
         x=2, y=3,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()

all_bioclim_flint_avgs_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC2, y=PC3, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)

all_bioclim_flint_avgs_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC2, y=PC3, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)
#high elev climate seems to be shifting to be more similar to low elev 
```

```{r}
lmer_results_flint <- avgs_flint_locs.pc %>%
  ungroup() %>% 
  select(TimePd, parent.pop, elev_m, Lat, Long, PC1:PC3) %>% 
  pivot_longer(starts_with("PC", ignore.case = FALSE), 
               names_to = "PC", values_to = "value") %>% 
  group_by(PC) %>% 
  nest(data=c(TimePd, parent.pop, elev_m, Lat, Long, value)) %>% 
  mutate(glm=map(data, ~ glm(value ~ TimePd*elev_m*Lat,
                               data=.x)),
         anova = map(glm, ~ broom.mixed::tidy(anova(.x))))

lmer_results_flint
```

```{r}
PC_anova_flint <- lmer_results_flint %>% select(-data, -glm) %>% unnest(anova) %>%
  select(PC, term, p.value) %>%
  filter(p.value < 0.05) %>%
  arrange(term, p.value)

PC_anova_flint
```
TimePd (evidence for climate change) is significant on PC1 with just Flint climate and Flint+Bioclim (dry/warm vs. cool/snowpack)

```{r}
autoplot(avgs_flint.pc, data = flint_grwseason_avgs,
         x=1, y=2,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()

avgs_flint_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)

avgs_flint_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=TimePd, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)
#high elev climate seems to be shifting to be more similar to low elev 
```