---
title: "Manuscript-Figures"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Figures for WL2 Climate Distance Manuscript

## Libraries
```{r}
library(tidyverse)
library(ggpubr) #ggarrange

sem <- function(x, na.rm=FALSE) {           #for caclulating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 

arcsin_transform <- function(x) { #for arcsine transforming probabilities 
  asin(sqrt(x))
}
```

## Load Data
```{r}
wl2_establishment <- read_csv("../output/WL2_Traits/WL2_Establishment.csv")
wl2_y1surv <- read_csv("../output/WL2_Traits/WL2_Y1Surv.csv")
wl2_surv_to_bud_y1 <- read_csv("../output/WL2_Traits/WL2_SurvtoRep_y1.csv")
wl2_fruits_y1 <- read_csv("../output/WL2_Traits/WL2_Fruits_Y1.csv")
winter_surv <- read_csv("../output/WL2_Traits/WL2_WinterSurv_2324.csv")
wl2_surv_to_bud_y2 <- read_csv("../output/WL2_Traits/WL2_Surv_to_Rep_Y2_2324.csv")
wl2_fruits_y2 <- read_csv("../output/WL2_Traits/WL2_Fruits_Y2_2324.csv")
wl2_prob_rep <- read_csv("../output/WL2_Traits/WL2_ProbFitness.csv")
```

### Subtraction Climate Distance
```{r}
wl2_wtr_year_sub_recent <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_Recent.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Recent=ann_tmean_dist, 
         Wtr_Year_PPTDist_Recent=ann_ppt_dist,
         Wtr_Year_CWDDist_Recent=cwd_dist)
wl2_wtr_year_sub_historic <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_Historical.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Historic=ann_tmean_dist, 
         Wtr_Year_PPTDist_Historic=ann_ppt_dist,
         Wtr_Year_CWDDist_Historic=cwd_dist)
wl2_sub_dist_2023 <- wl2_wtr_year_sub_recent %>% 
  left_join(wl2_wtr_year_sub_historic) %>% 
  rename(pop=parent.pop)

wl2_wtr_year_sub_recent_2324 <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_2324_Recent.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Recent=ann_tmean_dist, 
         Wtr_Year_PPTDist_Recent=ann_ppt_dist,
         Wtr_Year_CWDDist_Recent=cwd_dist)
wl2_wtr_year_sub_historic_2324 <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_2324_Historical.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Historic=ann_tmean_dist, 
         Wtr_Year_PPTDist_Historic=ann_ppt_dist,
         Wtr_Year_CWDDist_Historic=cwd_dist)

wl2_sub_dist_2324 <- wl2_wtr_year_sub_recent_2324 %>% 
  left_join(wl2_wtr_year_sub_historic_2324) %>% 
  rename(pop=parent.pop)

wl2_sub_dist_2023_prep <- wl2_sub_dist_2023 %>% 
  select(pop:Wtr_Year_CWDDist_Historic)
wl2_sub_dist_2324_prep <- wl2_sub_dist_2324 %>% 
  select(pop:Wtr_Year_CWDDist_Historic) 

wl2_sub_dist_avg <- bind_rows(wl2_sub_dist_2023_prep, wl2_sub_dist_2324_prep) %>% 
  group_by(pop) %>% 
  summarise_at(c("Wtr_Year_TempDist_Recent", "Wtr_Year_PPTDist_Recent", "Wtr_Year_CWDDist_Recent", 
                 "Wtr_Year_TempDist_Historic", "Wtr_Year_PPTDist_Historic", "Wtr_Year_CWDDist_Historic"),
               c(mean), na.rm = TRUE) 
```

## Data Summaries

### Maternal family sample sizes upon transplant
```{r}
wl2_establishment %>% 
  unite(pop.mf, pop:mf, remove = FALSE) %>% 
  group_by(pop.mf) %>% 
  summarise(n=n()) %>% 
  arrange(pop.mf) #number of plants per maternal family 

wl2_establishment %>% 
  select(pop, mf) %>% 
  distinct() %>% 
  group_by(pop) %>% 
  summarise(n=n()) %>% 
  arrange(n) #number of mfs per pop
```


### Year 2 survival to budding
```{r}
wl2_surv_to_bud_y2_summary <- wl2_surv_to_bud_y2 %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, Geographic_Dist, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurvtoBudY2=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) 

wl2_surv_to_bud_y2_summary_sub <- wl2_surv_to_bud_y2 %>% 
  left_join(wl2_sub_dist_2324) %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_PPTDist_Recent) %>% 
  summarise(meanSurvtoBudY2=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE))
```

### Prob Fitness
```{r}
prob_fitness_summary <- wl2_prob_rep %>%  
  group_by(pop, elev_m, Geographic_Dist, Wtr_Year_GD_Recent) %>% 
  summarise(meanProbFit=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>%
  mutate(RepYear=if_else(meanProbFit==0, "None",
                         if_else(pop=="TM2", "Both Years", "Year 2 Only")))
prob_fitness_summary$RepYear <- factor(prob_fitness_summary$RepYear,
                                               levels = c('None','Year 2 Only', 'Both Years'))


prob_fitness_summary_sub <- wl2_prob_rep %>% 
  left_join(wl2_sub_dist_avg) %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_PPTDist_Recent) %>% 
  summarise(meanProbFit=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>%
  mutate(RepYear=if_else(meanProbFit==0, "None",
                         if_else(pop=="TM2", "Both Years", "Year 2 Only")))
prob_fitness_summary_sub$RepYear <- factor(prob_fitness_summary_sub$RepYear,
                                               levels = c('None','Year 2 Only', 'Both Years'))
```

### Establishment, y1surv, winter surv
```{r}
est_summary <- wl2_establishment %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE))

y1surv_summary <- wl2_y1surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanY1Surv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) 

wintersurv_summary <- winter_surv %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent) %>% 
  summarise(meanWintSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE))

y1_fruits_summary <- wl2_fruits_y1 %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruitsY1=mean(fruits, na.rm = TRUE), semFruits=sem(fruits, na.rm=TRUE))

y2_fruits_summary <- wl2_fruits_y2 %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruitsY2=mean(fruits, na.rm = TRUE), semFruits=sem(fruits, na.rm=TRUE))

y1_surv_to_bud_summary <- wl2_surv_to_bud_y1 %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurvtoBudY1=mean(SurvtoRep_Y1, na.rm = TRUE), semSurv=sem(SurvtoRep_Y1, na.rm=TRUE))
```

## Figure 2: Gower's recent water year climate distance
```{r}
probfit_gower <- prob_fitness_summary %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanProbFit, group = pop, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(aes(shape=RepYear), size=6) + 
  geom_errorbar(aes(ymin=meanProbFit-semEst,ymax=meanProbFit+semEst),width=.02,linewidth = 2) +
  annotate("text", x = 0.3959884, y= 0.095, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of \n Successfully Reproducing", x="Recent Water Year Climate Dist", 
       color="Elevation (m)", shape="Rep Year(s)") +
  theme(text=element_text(size=30))

repsurvy2_gower <- wl2_surv_to_bud_y2_summary %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanSurvtoBudY2, group = pop, colour = elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurvtoBudY2-semEst,ymax=meanSurvtoBudY2+semEst),width=.02,linewidth = 2) +
  annotate("text", x = 0.25, y= 0.8333333, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0), limits = c(0.4,1.05)) +
  labs(y="Survival to Budding Y2", x="Recent Water Year Climate Dist", colour = "Elevation (m)") +
  theme(text=element_text(size=30))

fig3_gower <- ggarrange(probfit_gower, repsurvy2_gower, 
                        labels = c("A)", "B)"), 
                        font.label = list(size=30, face = "plain"), 
                        ncol=2, nrow=1) 
#ggsave("../output/WL2_Traits/Figure2_WtrYrGowers.png", width = 26, height = 9, units = "in")
```

## Figure 3: Temp and PPT recent water climate distance
```{r}
WYtemp_probfit <- prob_fitness_summary_sub %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanProbFit, group = pop, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(aes(shape=RepYear), size=6) + 
  geom_errorbar(aes(ymin=meanProbFit-semEst,ymax=meanProbFit+semEst),width=.3,linewidth = 2) +
  annotate("text", x = 0.6300972	, y= 0.095, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of \n Successfully Reproducing", x="Recent Water Year Temp Dist (°C)", 
       color="Elevation (m)", shape="Rep Year(s)") +
  theme(text=element_text(size=30))

WYppt_probfit <- prob_fitness_summary_sub %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Recent, y=meanProbFit, group = pop, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(aes(shape=RepYear), size=6) + 
  geom_errorbar(aes(ymin=meanProbFit-semEst,ymax=meanProbFit+semEst),width=.3,linewidth = 2) +
  annotate("text", x = -546.96536, y= 0.095, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of \n Successfully Reproducing", x="Recent Water Year PPT Dist (mm)", 
       color="Elevation (m)", shape="Rep Year(s)") +
  theme(text=element_text(size=30))

WYtemp_repsurvy2 <- wl2_surv_to_bud_y2_summary_sub %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanSurvtoBudY2, group = pop, colour = elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurvtoBudY2-semEst,ymax=meanSurvtoBudY2+semEst),width=.3,linewidth = 2) +
  annotate("text", x = 0.4, y= 0.8333333, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0), limits = c(0.4,1.05)) +
  labs(y="Survival to Budding Y2", x="Recent Water Year Temp Dist (°C)", colour = "Elevation (m)") +
  theme(text=element_text(size=30))

WYppt_repsurvy2 <- wl2_surv_to_bud_y2_summary_sub %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Recent, y=meanSurvtoBudY2, group = pop, colour = elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurvtoBudY2-semEst,ymax=meanSurvtoBudY2+semEst),width=.3,linewidth = 2) +
  annotate("text", x = 143, y= 0.8333333, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0), limits = c(0.4,1.05)) +
  labs(y="Survival to Budding Y2", x="Recent Water Year PPT Dist (mm)", colour = "Elevation (m)") +
  theme(text=element_text(size=30))

fig4_sub <- ggarrange(WYtemp_probfit, WYtemp_repsurvy2, 
                        WYppt_probfit, WYppt_repsurvy2, 
                        labels = c("A)", "B)", "C)", "D)"), 
                        font.label = list(size=30, face = "plain"), 
                        ncol=2, nrow=2) 
#ggsave("../output/WL2_Traits/Figure3_TmpPPT.png", width = 26, height = 18, units = "in")
```

## Figure S4: Geographic Distance
```{r}
est_geo <- wl2_establishment %>% 
  group_by(pop, elev_m, Geographic_Dist) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  annotate("text", x = 136.2622, y=0.50, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Geographic Distance (m)", 
       color="Elevation (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

probfit_geo <- prob_fitness_summary %>% 
  ggplot(aes(x=Geographic_Dist, y=meanProbFit, group = pop, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(aes(shape=RepYear), size=6) + 
  geom_errorbar(aes(ymin=meanProbFit-semEst,ymax=meanProbFit+semEst),width=.2, linewidth = 2) +
  annotate("text", x = 136.2622, y= 0.095, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of \n Successfully Reproducing", x="Geographic Distance (m)", 
       color="Elevation (m)", shape="Rep Year(s)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

repsurvy2_geo <- wl2_surv_to_bud_y2_summary %>% 
  ggplot(aes(x=Geographic_Dist, y=meanSurvtoBudY2, group = pop, colour = elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurvtoBudY2-semEst,ymax=meanSurvtoBudY2+semEst),width=.2, linewidth = 2) +
  annotate("text", x = 9500, y= 0.8333333, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0), limits = c(0.4,1.05)) +
  labs(y="Survival to Budding Y2", x="Geographic Distance (m)", colour = "Elevation (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

fig2_geo <- ggarrange(probfit_geo, est_geo, repsurvy2_geo, 
                        labels = c("A)", "B)", "C)"), 
                        font.label = list(size=30, face = "plain"), 
                        ncol=2, nrow=2) 
#ggsave("../output/WL2_Traits/FigureS4_Geo.png", width = 26, height = 18, units = "in")
```

## Figure S5: Additional Gower's recent water year climate distance 
```{r}
est_gower <- wl2_establishment %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>%
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanEst, group = pop, colour = elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  annotate("text", x = 0.3959884, y= 0.65, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Water Year Climate Dist", 
       color="Elevation (m)") +
  theme(text=element_text(size=30))

y1surv_gower <- wl2_y1surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanSurv, group = pop, colour = elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02,linewidth = 2) +
  annotate("text", x = 0.42, y= 0.7884615, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Recent Water Year Climate Dist", colour="Elevation (m)") +
  theme(text=element_text(size=30))

wintersurv_gower <- winter_surv %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanSurv, group = pop, colour = elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02,linewidth = 2) +
  annotate("text", x = 0.227, y= 0.24, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Survival", x="Recent Water Year Climate Dist", colour="Elevation (m)") +
  theme(text=element_text(size=30))

figs5_sub <- ggarrange(est_gower, y1surv_gower, 
                        wintersurv_gower, 
                        labels = c("A)", "B)", "C)"), 
                        font.label = list(size=30, face = "plain"), 
                        ncol=2, nrow=2) 
#ggsave("../output/WL2_Traits/FigureS5_WTRYRGowers.png", width = 26, height = 18, units = "in")
```

## Figure S6: Additonal Temp and PPT recent water yar climate distance
```{r}
WYtemp_est <- wl2_establishment %>% 
  left_join(wl2_sub_dist_2023) %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanEst, group = pop, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  annotate("text", x = 1.6352917, y= 0.65, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Water Year Temp Dist", 
       color="Elevation (m)") +
  theme(text=element_text(size=30))

WYppt_est <- wl2_establishment %>% 
  left_join(wl2_sub_dist_2023) %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Recent, y=meanEst, group = pop, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  annotate("text", x = -1120, y= 0.5714286, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Water Year PPT Dist", 
       color="Elevation (m)") +
  theme(text=element_text(size=30))

WYtemp_y1surv <- wl2_y1surv %>% 
  left_join(wl2_sub_dist_2023) %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanY1Surv=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanY1Surv, group = pop, colour = elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanY1Surv-semEst,ymax=meanY1Surv+semEst),width=.02,linewidth = 2) +
  annotate("text", x = 1.6352917, y= 0.88, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Recent Water Year Temp Dist", colour="Elevation (m)") +
  theme(text=element_text(size=30))

figs6_gower <- ggarrange(WYtemp_est, WYtemp_y1surv, 
                        WYppt_est, 
                        labels = c("A)", "B)", "C)"), 
                        font.label = list(size=30, face = "plain"), 
                        ncol=2, nrow=2) 
#ggsave("../output/WL2_Traits/FigureS6_TmpPPT.png", width = 26, height = 18, units = "in")
```

## Tradeoffs between fitness metrics

```{r}
all_fitness <- est_summary %>% 
  left_join(y1surv_summary) %>% 
  select(-semSurv) %>% 
  left_join(y1_surv_to_bud_summary) %>% 
  select(-semSurv) %>%
  left_join(y1_fruits_summary) %>% 
  ungroup() %>% 
  select(-Wtr_Year_GD_Recent, -Wtr_Year_GD_Historical, -GrwSsn_GD_Recent, -semEst, -semFruits) %>% 
  left_join(wintersurv_summary) %>% 
  select(-semSurv, -Wtr_Year_GD_Recent) %>%
  left_join(wl2_surv_to_bud_y2_summary) %>% 
  select(-semEst, -Wtr_Year_GD_Recent) %>%
  left_join(y2_fruits_summary) %>% 
  select(-semFruits, -Wtr_Year_GD_Recent, -Geographic_Dist)
all_fitness
```

```{r}
all_fitness %>% 
  ggplot(aes(meanEst, meanY1Surv, colour = elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic()

all_fitness %>% 
  ggplot(aes(meanEst, meanWintSurv, colour = elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic()

all_fitness %>% 
  ggplot(aes(meanEst, meanSurvtoBudY2, colour = elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic()

all_fitness %>% 
  ggplot(aes(meanEst, meanFruitsY2, colour = elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic()

all_fitness %>% 
  ggplot(aes(meanY1Surv, meanWintSurv, colour = elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic()

all_fitness %>% 
  ggplot(aes(meanY1Surv, meanSurvtoBudY2, colour = elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic()

all_fitness %>% 
  ggplot(aes(meanY1Surv, meanFruitsY2, colour = elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic()

all_fitness %>% 
  ggplot(aes(meanWintSurv, meanSurvtoBudY2, colour = elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic()

all_fitness %>% 
  ggplot(aes(meanWintSurv, meanFruitsY2, colour = elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic()

all_fitness %>% 
  ggplot(aes(meanSurvtoBudY1, meanSurvtoBudY2, colour = elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic()

all_fitness %>% 
  ggplot(aes(meanSurvtoBudY2, meanFruitsY2, colour = elev_m)) +
  geom_point() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_classic()
```

### Quick Stats
Arcsin transformed proportions
which is: asin(sqrt(p))
```{r}
allfitness_asin <- all_fitness %>% 
  #change NAs to neg so that they will come out as NAs (no built in na.rm to sart function):
  mutate(meanY1Surv = if_else(is.na(meanY1Surv), -1, meanY1Surv), 
         meanSurvtoBudY1 = if_else(is.na(meanSurvtoBudY1), -1, meanSurvtoBudY1),
         meanFruitsY1 = if_else(is.na(meanFruitsY1), -1, meanFruitsY1),
         meanWintSurv = if_else(is.na(meanWintSurv), -1, meanWintSurv),
         meanSurvtoBudY2 = if_else(is.na(meanSurvtoBudY2), -1, meanSurvtoBudY2),
         meanFruitsY2 = if_else(is.na(meanFruitsY2), -1, meanFruitsY2)) %>% 
  #arcsin transform mean proportions 
  mutate(across(starts_with("mean"), arcsin_transform, .names = "arctr_{.col}"))
head(allfitness_asin)
names(allfitness_asin)
```

```{r}
est_y1surv <- lm(arctr_meanY1Surv ~ arctr_meanEst, data=allfitness_asin)
summary(est_y1surv) #sig pos relat

est_wintsurv <- lm(arctr_meanWintSurv ~ arctr_meanEst, data=allfitness_asin)
summary(est_wintsurv) #sig pos relat

est_y2surv <- lm(arctr_meanSurvtoBudY2 ~ arctr_meanEst, data=allfitness_asin)
summary(est_y2surv) #not sig neg relat

est_y2fruits <- lm(meanFruitsY2 ~ arctr_meanEst, data=allfitness_asin)
summary(est_y2fruits) #sig pos relat - weird given figure..


y1surv_wintsurv <- lm(arctr_meanWintSurv ~ arctr_meanY1Surv, data=allfitness_asin)
summary(y1surv_wintsurv) #sig pos relat

y1surv_y2surv <- lm(arctr_meanSurvtoBudY2 ~ arctr_meanY1Surv, data=allfitness_asin)
summary(y1surv_y2surv) #not sig neg relat

y1surv_y2fruits <- lm(meanFruitsY2 ~ arctr_meanY1Surv, data=allfitness_asin)
summary(y1surv_y2fruits) #sig pos relat - weird given figure..


wintsurv_y2surv <- lm(arctr_meanSurvtoBudY2 ~ arctr_meanWintSurv, data=allfitness_asin)
summary(wintsurv_y2surv) #not sig neg relat

wintsurv_y2fruits <- lm(meanFruitsY2 ~ arctr_meanWintSurv, data=allfitness_asin)
summary(wintsurv_y2fruits) #sig pos relat - weird given figure..


y2surv_y2fruits <- lm(meanFruitsY2 ~ arctr_meanSurvtoBudY2, data=allfitness_asin)
summary(y2surv_y2fruits) #not sig neg relat 
```

## Additional figures (come back to these)

-   Figure 1: Population map and PCAs

-   Figure S1: Pearson's correlations

-   Figure S2: PCAs with PC1 and PC2

-   Figure S3: Gower's climate distance across pops

-   Figure S7: Growth season climate distance effects
