---
title: "Data Exploration"
author: "Brandie Quarles"
date: "2023-08-03"
output: html_document:
  keep_md: yes
---

To Do
- read through notes to distinguish between "transplants" and "germination"
- figure out what germination surveys are - see UCD notes 
- read protocol of annual census - what's the purpose?

#Note: Not all data has been added to the corrected csvs folder, some are still in raw csvs

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Relevant Libraries and Functions
```{r}
library(tidyverse)
library(naniar) #replaces values with NA
sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("high", "mid", "low")
```


##Figuring out what the different data are 
#Annual Censuses of Germination and Transplant Beds
Purpose: match native populations annual demog. censuses? 
```{r}
#Note: BQC edited the csvs prior to importing to R b/c of issues with the naming of columns and the first row of the transplants sheets not being the heading row (veg, repo info and not block, mf etc...) 
pop_elev <- read_csv("../input/Pops_for_2022_UCD.csv")
summary(pop_elev)
pop_elev_nonotes <- pop_elev %>% select(-notes)
pop_elev_nonotes

anncensusgerm <- read_csv("../input/Annual_census_germination_20230626_corrected.csv")
head(anncensusgerm)
summary(anncensusgerm)
#convert "n/a" to NA
anncensusgerm_nacorrected <- anncensusgerm %>% replace_with_na_all(condition = ~.x == "n/a")
head(anncensusgerm_nacorrected, 30)
summary(anncensusgerm_nacorrected) #only block:mf, pheno, herb_dam, wilt_status, and notes should be character
anncensusgerm_nanumcorrected <- anncensusgerm_nacorrected %>% mutate_at(c("diam_mm", "height_cm", "total_branch", "longest_leaf_cm", "flowers", "fruits", "longest_fruit_cm", "repro_branch"), as.double)
summary(anncensusgerm_nanumcorrected)

anncensustrans_0522 <- read_csv("../input/Annual_Census_Transplants_20230522_corrected.csv")
head(anncensustrans_0522)
summary(anncensustrans_0522)
anncensustrans_0522 %>% filter(diam_mm=="X") #747 rows 
anncensustrans_0522 %>% filter(is.na(diam_mm)) #53 rows 
anncensustrans_0522_numcorrected <- anncensustrans_0522 %>% mutate_at(c("diam_mm", "height_cm", "total_branch", "longest_leaf_cm_veg", "flowers_repro", "fruits_repro", "longest_fruit_cm_repro", "repro_branch"), as.double)
summary(anncensustrans_0522_numcorrected) #correct number of NAs

anncensustrans_0626 <- read_csv("../input/Annual_Census_transplants_20230626_corrected.csv")
head(anncensustrans_0626)
summary(anncensustrans_0626)

names(anncensusgerm_nanumcorrected)
names(anncensustrans_0522_numcorrected)
names(anncensustrans_0626)
names(pop_elev_nonotes)

#merge with elevation info 
anncensustrans_0522_elev <- left_join(anncensustrans_0522_numcorrected, pop_elev_nonotes, by=c("pop"="parent.pop"))
head(anncensustrans_0522_elev, 30)
summary(anncensustrans_0522_elev)
```


#Germination Surveys
Purpose: monitor germination of planted seeds in germ beds (plants that didn't germinate in the lab)
```{r}
germsurvey_0127 = read_csv("../input/Germination_survey_20230127_corrected.csv")
summary(germsurvey_0127)
head(germsurvey_0127)

germsurvey_0210 = read_csv("../input/Germination_survey_20230210_corrected.csv")
summary(germsurvey_0210)
head(germsurvey_0210)

germsurvey_0508 = read_csv("../input/Germination_survey_20230508_corrected.csv")
summary(germsurvey_0508)
head(germsurvey_0508)
```


#Mortality Survey?
```{r}
mortsurvey <- read_csv("../input/Mortality_survey_20230217_corrected.csv")
summary(mortsurvey) #859 rows so proably only transplant beds?
head(mortsurvey)
```


#Phenology Survey of Germination Bed?
```{r}
pheno_germs <- read_csv("../input/Phenology_survey_germination_20230515_corrected.csv")
summary(pheno_germs)
head(pheno_germs)
```


#Size Surveys of Germination and Transplant Beds?
```{r}
size_germs_0303
size_germs_0317
size_germs_0324
size_germs_0403
size_germs_0417
size_germs_0424
size_germs_0508
size_germs_0515
size_germs_0605

size_trans_1213
sie_trans_0303
sie_trans_0317
sie_trans_0324
sie_trans_0403
sie_trans_0417
sie_trans_0424
sie_trans_0401
sie_trans_0508
sie_trans_0605

ucd_garden_size_1128 <- read_csv("../input/UCD_garden_size_measurements_20221128_corrected.csv")
summary(ucd_garden_size_1128) #2264 rows... maybe germs and trans combined?
head (ucd_garden_size_1128)
```

##Annual Census Analysis 
Questions:
- Is there more variation b/t populations than b/t mfs?
- Are certain traits related to fitness?
- Did low elevation populations perform better than high elevation populations?
- Connection to climate of home site?
```{r}
names(anncensusgerm_nanumcorrected) #not a lot of actual observations, when was this collected?
names(anncensustrans_0522_numcorrected)
names(anncensustrans_0626)

#histograms
anncensustrans_0522_numcorrected %>% ggplot(aes(x=diam_mm)) + geom_histogram()
anncensustrans_0522_numcorrected %>% ggplot(aes(x=height_cm)) + geom_histogram()
anncensustrans_0522_numcorrected %>% ggplot(aes(x=total_branch)) + geom_histogram()
anncensustrans_0522_numcorrected %>% ggplot(aes(x=longest_leaf_cm_veg)) + geom_histogram()
anncensustrans_0522_numcorrected %>% ggplot(aes(x=flowers_repro)) + geom_histogram()
anncensustrans_0522_numcorrected %>% ggplot(aes(x=fruits_repro)) + geom_histogram()
anncensustrans_0522_numcorrected %>% ggplot(aes(x=longest_fruit_cm_repro)) + geom_histogram()

#plot traits across pops and mfs 
anncensustrans_0522_numcorrected %>% ggplot(aes(x=pop, y=diam_mm)) + geom_boxplot() #more pops with data on 0522 than 0626
anncensustrans_0626 %>% ggplot(aes(x=pop, y=diam_mm)) + geom_boxplot()

```

Checking Survival to May 22, 2023
```{r}
#table of phenology variable 
xtabs(~pheno, data = anncensustrans_0522_numcorrected) 
#747 dead (X)
#32 vegetative (V)
#15 post flowering (fruiting) (P)
#8 budding (B)
#3 flowering (F)

#convert everything not dead to alive
anncensustrans_0522_elev_surv <- anncensustrans_0522_elev %>% mutate(Surv=ifelse(pheno == "X", 0, 1)) %>% select(phylogroup, elevation.group, UCD.seed.year, date:repro_branch, Surv) #1=survived, 0=dead 
head(anncensustrans_0522_elev_surv, 30)
tail(anncensustrans_0522_elev_surv, 30)

xtabs(~Surv, data = anncensustrans_0522_elev_surv) #only 58 plants alive late May 
surv_popmeans <- anncensustrans_0522_elev_surv %>% group_by(pop, elevation.group) %>% summarise(meanSurv=mean(Surv, na.rm = TRUE), sem.mort=sem(Surv, na.rm=TRUE))
surv_popmeans <- surv_popmeans %>% subset(pop != "buffer")
surv_popmeans

surv_popmeans$elevation.group <- factor(surv_popmeans$elevation.group, levels=elev_order)
surv_popmeans$pop <- factor(surv_popmeans$pop)

surv_popmeans %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-sem.mort,ymax=meanSurv+sem.mort),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  #geom_hline(yintercept=0.5, linetype="dashed", color="red") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

xtabs(~elevation.group + Surv, data = anncensustrans_0522_elev_surv) #low elevation has the most alive and mid-elevation has the least, but this isn't standardized by the total amount planted

surv_popmeans %>% ggplot(aes(x=elevation.group, y=meanSurv)) + geom_boxplot()
```

