---
title: "Establishment_Figures"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Establishment (survival 3 weeks post transplant)

## Libraries
```{r}
library(tidyverse)
library(magrittr)
#conflicted::conflicts_prefer(dplyr::filter)
library(viridis) #for pretty colors
library(ggrepel) #for non-overlapping labels on plots
library(ggdist) #visualizations of distributions and uncertainty 
library(ggpubr)

#library(sp) #for calculating geographic distance 
library(geosphere) #for calculating geographic distance
library(corrplot) #plotting correlations  
library(rstatix) #performing cor_test

library(lmerTest) #mixed models
conflicted::conflicts_prefer(lmerTest::lmer)
library(broom.mixed)
library(tidymodels)
library(furrr) #Apply Mapping Functions in Parallel using Futures
tidymodels_prefer()
# install.packages("multilevelmod")
library(multilevelmod)

library(brms)
#conflicted::conflicts_prefer(brms::ar)
#conflicted::conflicts_prefer(dplyr::combine)
#conflicted::conflicts_prefer(brms::dstudent_t)
library(tidybayes) #for extracting and visiaulizing brms model output 
library(modelr) #for data grid

sem <- function(x, na.rm=FALSE) {           #for caclulating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 

cbbPalette2 <- c("#E69F00","#000000", "#56B4E9","#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
timepd_palette <- c("#56B4E9","#D55E00")

options(mc.cores = parallel::detectCores())
```

## Load data from UCD and WL2
```{r}
ucd_surv <- read_csv("../input/UCD_Data/CorrectedCSVs/UCD_transplants_pheno_mort_20231016_corrected.csv") %>% 
  rename(death.date=`Death Date`, bud.date=`Date First Bud`, flower.date=`Date First Flower`, 
         fruit.date=`Date First Fruit`, last.flower.date=`Date Last Flower`, last.fruit.date=`Date Last Fruit`) %>% 
  filter(!is.na(pop)) %>% 
  unite(Genotype, pop:rep, sep="_", remove = FALSE) %>% 
  filter(!str_detect(Genotype, "buffer")) 
#unique(ucd_surv$pop)

wl2_surv_1020 <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_mort_pheno_20231020_corrected.csv") #need to add in 10/27 mortality 

wl2_ann_cens <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_annual_census_20231027_corrected.csv") %>% 
  unite(Genotype, pop:rep, sep="_", remove = FALSE) %>% 
  unite(BedLoc, bed:`bed-col`, sep="_", remove = FALSE) %>% 
  filter(BedLoc!="K_5_C") %>% #get rid of duplicate locations
  filter(BedLoc!="B_32_A") %>% #get rid of duplicate locations
  filter(!is.na(pop), !str_detect(Genotype, ".*buff*")) #remove buffers 

wl2_surv_1027 <- wl2_ann_cens %>% 
  filter(pheno=="X") %>% 
  select(death.date_2=date, block:bed, bed.row=`bed-row`, bed.col=`bed-col`, pop:rep) #add in 10/27 death dates
#note: 10/27 pheno dates (like flowering), not updated with this code 

wl2_surv_y1 <- left_join(wl2_surv_1020, wl2_surv_1027) %>% 
  mutate(death.date=if_else(is.na(death.date), death.date_2, death.date)) %>% 
  mutate(pop= str_replace(pop, "Y08", "YO8")) %>% 
  mutate(pop= str_replace(pop, "Y04", "YO4")) %>% 
  unite(BedLoc, bed:bed.col, sep="_", remove = FALSE) %>% 
  filter(BedLoc!="K_5_C") %>% #get rid of duplicate locations
  filter(BedLoc!="B_32_A") %>% #get rid of duplicate locations
  unite(Genotype, pop:rep, sep="_", remove = FALSE) %>% 
  filter(!is.na(pop), !str_detect(Genotype, ".*buff*")) %>%  #remove buffers 
  select(block:survey.notes)
#wl2_surv_y1_to_export <- wl2_surv_y1 %>% select(block:rep, death.date, survey.notes)
#write_csv(wl2_surv_y1_to_export, "../output/WL2_Traits/WL2_Mortality_2023.csv")
#unique(wl2_surv_y1$pop)
```

## Gower's Distance
```{r}
garden_climate <- read_csv("../output/Climate/flint_climate_UCDpops.csv") %>% 
  filter(parent.pop=="WL2_Garden" | parent.pop=="UCD_Garden") %>% 
  select(parent.pop:Long) %>% 
  distinct()
garden_climate
#UCD LAT/LONG = 38.53250, -121.7830
#WL2 Lat/Long = 38.82599, -120.2509

ucd_gowers <- read_csv("../output/Climate/Gowers_UCD.csv") %>% 
  select(parent.pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  mutate(UCD_Lat=38.53250, UCD_Long=-121.7830, UCD_Elev=16) %>% 
  mutate(Geographic_Dist=distHaversine(cbind(UCD_Long, UCD_Lat), cbind(Long, Lat)),
         Elev_Dist=elev_m-UCD_Elev) %>% # Calculate the distance using the haversine formula (dist in meters)
  #mutate(Lat_Dist=UCD_Lat-Lat, Long_Dist=UCD_Long-Long) %>% #Garden-Home - lat and long per Gerst et al 2011 which kept them separate for some directionality
  rename(pop=parent.pop)

wl2_gowers_2023 <- read_csv("../output/Climate/Gowers_WL2.csv") %>% 
  select(parent.pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  mutate(WL2_Lat=38.82599, WL2_Long=-120.2509, WL2_Elev=2020) %>% 
  mutate(Geographic_Dist=distHaversine(cbind(WL2_Long, WL2_Lat), cbind(Long, Lat)),
         Elev_Dist=elev_m-WL2_Elev) %>% # Calculate the distance using the haversine formula
  #mutate(Lat_Dist=WL2_Lat-Lat, Long_Dist=WL2_Long-Long) %>% #Garden-Home - lat and long per Gerst et al 2011 which kept them separate for some directionality
  rename(pop=parent.pop)
```

## Climate Subtraction Distances
```{r}
ucd_wtr_year_sub_recent <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_Davis_Recent.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Recent=ann_tmean_dist, Wtr_Year_PPTDist_Recent=ann_ppt_dist)
ucd_wtr_year_sub_historic <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_Davis_Historical.csv") %>%
  select(parent.pop, Wtr_Year_TempDist_Historic=ann_tmean_dist, Wtr_Year_PPTDist_Historic=ann_ppt_dist)
ucd_grwssn_sub_recent <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_Home_Davis_Recent.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Recent=ann_tmean_dist, GrwSsn_PPTDist_Recent=ann_ppt_dist)
ucd_grwssn_sub_historic <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_Home_Davis_Historical.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Historic=ann_tmean_dist, GrwSsn_PPTDist_Historic=ann_ppt_dist)
ucd_sub_dist <- ucd_wtr_year_sub_recent %>% 
  left_join(ucd_wtr_year_sub_historic) %>% 
  left_join(ucd_grwssn_sub_recent) %>% 
  left_join(ucd_grwssn_sub_historic) %>% 
  rename(pop=parent.pop) %>% 
  left_join(ucd_gowers)

wl2_wtr_year_sub_recent <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_Recent.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Recent=ann_tmean_dist, Wtr_Year_PPTDist_Recent=ann_ppt_dist)
wl2_wtr_year_sub_historic <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_Historical.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Historic=ann_tmean_dist, Wtr_Year_PPTDist_Historic=ann_ppt_dist)
wl2_grwssn_sub_recent <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_Home_WL2_Recent.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Recent=ann_tmean_dist, GrwSsn_PPTDist_Recent=ann_ppt_dist)
wl2_grwssn_sub_historic <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_Home_WL2_Historical.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Historic=ann_tmean_dist, GrwSsn_PPTDist_Historic=ann_ppt_dist)
wl2_sub_dist_2023 <- wl2_wtr_year_sub_recent %>% 
  left_join(wl2_wtr_year_sub_historic) %>% 
  left_join(wl2_grwssn_sub_recent) %>% 
  left_join(wl2_grwssn_sub_historic) %>% 
  rename(pop=parent.pop) %>% 
  left_join(wl2_gowers_2023)
```


## Davis - used 3 weeks to match WL2
```{r}
unique(ucd_surv$pop)

ucd_surv %>% filter(rep==100)
ucd_surv %>% filter(Genotype=="WL2_4_10")
#there are 3 plants that died early and later had germ WL2_4_10(1/6/23), WL2_3_100(L1_5_A)-1/6/23 and SQ2_6_100(L2_37_C)-12/13/22

ucd_establishment <- ucd_surv %>% 
  left_join(ucd_gowers) %>% 
  select(block:rep, elevation.group:Wtr_Year_GD_Historical, Geographic_Dist, Elev_Dist, death.date) %>% 
  mutate(death.date=mdy(death.date)) %>% 
  mutate(Establishment=if_else(is.na(death.date), 1,
                               if_else(death.date=="2022-11-30" | death.date=="2022-12-13" | death.date=="2022-12-21", 0,
                                       1))) %>% 
  mutate(Establishment=if_else(Genotype=="SQ2_6_100", 0, Establishment))

#ucd_establishment %>% filter(Establishment==0)
  
#write_csv(ucd_establishment, "../output/UCD_Traits/UCD_Establishment.csv")
```

```{r}
ucd_establishment_sub_dist <- ucd_establishment %>% 
  select(block:rep, Establishment) %>% 
  left_join(ucd_sub_dist)

unique(ucd_establishment_sub_dist$block)
unique(ucd_establishment_sub_dist$pop)
unique(ucd_establishment_sub_dist$mf)
```


### Bar plots
```{r}
ucd_establishment %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanEst), y=meanEst, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/UCD_Traits/UCD_Establishment_GrwSsn_GD_Recent.png", width = 12, height = 8, units = "in")

ucd_establishment %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanEst), y=meanEst, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/UCD_Traits/UCD_Establishment_Wtr_Year_GD_Recent.png", width = 12, height = 8, units = "in")
```


## WL2 - 3 weeks based off survival curve
```{r}
wl2_establishment <- wl2_surv_y1 %>% 
  left_join(wl2_gowers_2023) %>% 
  select(block:rep, elevation.group:Wtr_Year_GD_Historical, Geographic_Dist, Elev_Dist, death.date) %>% 
  mutate(death.date=mdy(death.date)) %>% 
  mutate(Establishment=if_else(is.na(death.date), 1,
                               if_else(death.date=="2023-07-26" | death.date=="2023-08-02" | death.date=="2023-08-09", 0,
                                       1)))
#write_csv(wl2_establishment, "../output/WL2_Traits/WL2_Establishment.csv")
```

```{r}
wl2_establishment_sub_dist <- wl2_establishment %>% 
  select(block:rep, Establishment) %>% 
  left_join(wl2_sub_dist_2023)

unique(wl2_establishment_sub_dist$block)
unique(wl2_establishment_sub_dist$pop)
unique(wl2_establishment_sub_dist$mf)
```

### Bar plots
```{r}
wl2_establishment %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanEst), y=meanEst, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/WL2_Traits/WL2_Establishment_GrwSsn_GD_Recent.png", width = 12, height = 8, units = "in")

wl2_establishment %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanEst), y=meanEst, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/WL2_Traits/WL2_Establishment_Wtr_Year_GD_Recent.png", width = 12, height = 8, units = "in")
```

## Figures for Paper

### Gower's, Temp, and Ppt
```{r}
WYCD_recent <- wl2_establishment %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>%
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanEst, group = pop, colour = elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  annotate("text", x = 0.3959884, y= 0.65, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Water Year Climate Dist", 
       color="Elevation (m)") +
  theme(text=element_text(size=30))

WYCD_temp_recent <- wl2_establishment_sub_dist %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanEst, group = pop, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  annotate("text", x = 1.6352917, y= 0.65, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Water Year Temp Dist", 
       color="Elevation (m)") +
  theme(text=element_text(size=30))

WYCD_ppt_recent <- wl2_establishment_sub_dist %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Recent, y=meanEst, group = pop, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  annotate("text", x = -1120, y= 0.5714286, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Water Year PPT Dist", 
       color="Elevation (m)") +
  theme(text=element_text(size=30))

wl2_est_FIG_recent <- ggarrange(WYCD_recent, "", 
                                WYCD_temp_recent,
                                WYCD_ppt_recent,
                                labels = c("A)", "", "B)", "C)"), 
                                font.label = list(size=30, 
                                                  face = "plain"), 
                                ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_Estab_RecentDist.png", width = 26, height = 18, units = "in")
```

### Just Geo Distance
```{r}
wl2_establishment %>% 
  group_by(pop, elev_m, Geographic_Dist) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  annotate("text", x = 136.2622, y=0.50, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Geographic Distance (m)", 
       color="Elevation (m)") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/WL2_Traits/WL2_Establishment_GeoDist.png", width = 10, height = 5, units = "in")
```

## Scatterplots

### Davis
```{r}
#scatter plots
GSCD_recent <- ucd_establishment %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Growth Season CD") +
  theme(text=element_text(size=25))

WYCD_recent <- ucd_establishment %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>%
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Water Year CD") +
  theme(text=element_text(size=25))

GD <- ucd_establishment %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, Geographic_Dist) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Geographic Distance (m)") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))


ED <- ucd_establishment %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Elevation Distance (m)") +
  theme(text=element_text(size=25))

ucd_establishment_FIG <- ggarrange(GSCD_recent, WYCD_recent, GD, ED, ncol=2, nrow=2) 
#ggsave("../output/UCD_Traits/UCD_Establishment_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots
GSCD_historic <- ucd_establishment %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Historical, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Historic Growth Season CD") +
  theme(text=element_text(size=25))

WYCD_historic <- ucd_establishment %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>%
  ggplot(aes(x=Wtr_Year_GD_Historical, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Historic Water Year CD") +
  theme(text=element_text(size=25))

ucd_establishment_FIG <- ggarrange(GSCD_historic, WYCD_historic, GD, ED, ncol=2, nrow=2) 
#ggsave("../output/UCD_Traits/UCD_Establishment_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

### WL2
```{r}
#scatter plots
GSCD_recent <- wl2_establishment %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Growth Season CD") +
  theme(text=element_text(size=25))

WYCD_recent <- wl2_establishment %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>%
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Water Year CD") +
  theme(text=element_text(size=25))

GD <- wl2_establishment %>% 
  group_by(pop, elev_m, Geographic_Dist) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Geographic Distance (m)") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))


ED <- wl2_establishment %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Elevation Distance (m)") +
  theme(text=element_text(size=25))

wl2_establishment_FIG <- ggarrange(GSCD_recent, WYCD_recent, GD, ED, ncol=2, nrow=2) 
#ggsave("../output/WL2_Traits/WL2_Establishment_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots
GSCD_historic <- wl2_establishment %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Historical, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Historic Growth Season CD") +
  theme(text=element_text(size=25))

WYCD_historic <- wl2_establishment %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>%
  ggplot(aes(x=Wtr_Year_GD_Historical, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Historic Water Year CD") +
  theme(text=element_text(size=25))

wl2_establishment_FIG <- ggarrange(GSCD_historic, WYCD_historic, GD, ED, ncol=2, nrow=2) 
#ggsave("../output/WL2_Traits/WL2_Establishment_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

### Directional Distance
```{r}
#scatter plots - recent
GSCD_prob_recent <- wl2_establishment_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- wl2_establishment_sub_dist %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Water Year Temp Dist", color="Water Year \n Climate Distance") +
  theme(text=element_text(size=30))

GD_prob <- wl2_establishment_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, Wtr_Year_TempDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- wl2_establishment_sub_dist %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

WL2_establishment_sub_FIG_prob_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
#ggsave("../output/WL2_Traits/WL2_Establishment_TmpSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- wl2_establishment_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Historic Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- wl2_establishment_sub_dist %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Historic Water Year Temp Dist", color="Water Year \n Climate Distance") +
  theme(text=element_text(size=30))

WL2_establishment_sub_FIG_prob_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
#ggsave("../output/WL2_Traits/WL2_Establishment_TmpSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
GSCD_prob_recent <- ucd_establishment_sub_dist %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- ucd_establishment_sub_dist %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Water Year Temp Dist", color="Water Year \n Climate Distance") +
  theme(text=element_text(size=30))

GD_prob <- ucd_establishment_sub_dist %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, Wtr_Year_TempDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- ucd_establishment_sub_dist %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

UCD_establishment_sub_FIG_prob_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
#ggsave("../output/UCD_Traits/UCD_Establishment_TmpSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- ucd_establishment_sub_dist %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Historic Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- ucd_establishment_sub_dist %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Historic Water Year Temp Dist", color="Water Year \n Climate Distance") +
  theme(text=element_text(size=30))

UCD_establishment_sub_FIG_prob_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
#ggsave("../output/UCD_Traits/UCD_Establishment_TmpSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
GSCD_prob_recent <- wl2_establishment_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=0.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- wl2_establishment_sub_dist %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Water Year PPT Dist", color="Water Year \n Climate Distance") +
  theme(text=element_text(size=30))

GD_prob <- wl2_establishment_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, Wtr_Year_PPTDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- wl2_establishment_sub_dist %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

WL2_establishment_sub_FIG_prob_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
#ggsave("../output/WL2_Traits/WL2_Establishment_PPTSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- wl2_establishment_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Historic Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- wl2_establishment_sub_dist %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Historic Water Year PPT Dist", color="Water Year \n Climate Distance") +
  theme(text=element_text(size=30))

WL2_establishment_sub_FIG_prob_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
#ggsave("../output/WL2_Traits/WL2_Establishment_PPTSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
GSCD_prob_recent <- ucd_establishment_sub_dist %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- ucd_establishment_sub_dist %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Recent Water Year PPT Dist", color="Water Year \n Climate Distance") +
  theme(text=element_text(size=30))

GD_prob <- ucd_establishment_sub_dist %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, Wtr_Year_PPTDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- ucd_establishment_sub_dist %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

UCD_establishment_sub_FIG_prob_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
#ggsave("../output/UCD_Traits/UCD_Establishment_PPTSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- ucd_establishment_sub_dist %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Historic Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- ucd_establishment_sub_dist %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Establishment", x="Historic Water Year PPT Dist", color="Water Year \n Climate Distance") +
  theme(text=element_text(size=30))

UCD_establishment_sub_FIG_prob_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
#ggsave("../output/UCD_Traits/UCD_Establishment_PPTSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

## Stats

### Scaling 
```{r}
wl2_establishment %>% group_by(pop) %>% summarise(n=n()) %>% arrange(n) #CHECK SAMPLE SIZE
wl2_establishment_scaled <- wl2_establishment %>% mutate_at(c("GrwSsn_GD_Recent","Wtr_Year_GD_Recent",                                                           "GrwSsn_GD_Historical","Wtr_Year_GD_Historical","Geographic_Dist"),
                                                            scale) 

ucd_establishment %>% group_by(pop) %>% summarise(n=n()) %>% arrange(n) #CHECK SAMPLE SIZE
ucd_establishment_scaled <- ucd_establishment %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  mutate_at(c("GrwSsn_GD_Recent","Wtr_Year_GD_Recent",                                                           "GrwSsn_GD_Historical","Wtr_Year_GD_Historical","Geographic_Dist"),
                                                            scale) 

#CHECK MODEL PARAMS
unique(wl2_establishment_scaled$pop)
unique(wl2_establishment_scaled$mf)
unique(wl2_establishment_scaled$block)
summary(wl2_establishment_scaled)

unique(ucd_establishment_scaled$pop)
unique(ucd_establishment_scaled$mf)
unique(ucd_establishment_scaled$block)
summary(ucd_establishment_scaled)
```

```{r}
wl2_establishment_scaled_sub <- wl2_establishment_sub_dist %>% 
  mutate_at(c("Wtr_Year_TempDist_Recent",  "Wtr_Year_PPTDist_Recent", 
                 "Wtr_Year_TempDist_Historic", "Wtr_Year_PPTDist_Historic",
                 "GrwSsn_TempDist_Recent", "GrwSsn_PPTDist_Recent",
                 "GrwSsn_TempDist_Historic", "GrwSsn_PPTDist_Historic",
              "Geographic_Dist"), scale)

ucd_establishment_scaled_sub <- ucd_establishment_sub_dist %>% 
  filter(pop!="WV") %>% #Only 2 indivs for WV
  mutate_at(c("Wtr_Year_TempDist_Recent",  "Wtr_Year_PPTDist_Recent", 
                 "Wtr_Year_TempDist_Historic", "Wtr_Year_PPTDist_Historic",
                 "GrwSsn_TempDist_Recent", "GrwSsn_PPTDist_Recent",
                 "GrwSsn_TempDist_Historic", "GrwSsn_PPTDist_Historic",
              "Geographic_Dist"), scale)
```

### Basic Model Workflow 
```{r}
glmer.model_binomial <- 
  linear_reg() %>% 
  set_engine("glmer", family=binomial)

surv_wflow <- workflow() %>% 
  add_variables(outcomes = Establishment, predictors = c(pop, mf, block))

surv_fits <- tibble(wflow=list(
  pop = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ (1|pop))},
  
  pop.mf = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ (1|pop/mf))},
  
  pop.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ (1|pop) + (1|block))},
  
  pop.mf.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ (1|pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)

surv_fits_wl2 <- surv_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_establishment_scaled))
#mod_test <- glmer(Establishment ~ (1|pop) + (1|block), data=wl2_establishment_scaled, family=binomial)
#summary(mod_test)
#boundary (singular) fit: see help('isSingular') errors with mf in the model 

surv_fits_ucd <- surv_fits %>%
  mutate(fit = map(wflow, fit, data = ucd_establishment_scaled))
#mod_test <- glmer(Establishment ~ (1|pop/mf) + (1|block), data=ucd_establishment_scaled, family=binomial)
#summary(mod_test)
#full model: Warning: unable to evaluate scaled gradientWarning: Model failed to converge: degenerate  Hessian with 1 negative eigenvalues
#pop.mf: boundary (singular) fit: see help('isSingular')

surv_fits_wl2 %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#pop block model best by AIC and BIC

surv_fits_ucd %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#pop.block model best by AIC and BIC, issues with mf so exclude it 
```

#### Test climate and geographic distance 
```{r}
surv_GD_wflow <- workflow() %>%
  add_variables(outcomes = Establishment, predictors = c(pop, mf, block, contains("GD"), Geographic_Dist)) 

surv_GD_fits <- tibble(wflow=list(
  pop.block = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ (1|pop) + (1|block))},
  
  GS_Recent = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ GrwSsn_GD_Recent + Geographic_Dist + (1|pop) + (1|block))},
  
  GS_Historical = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ GrwSsn_GD_Historical + Geographic_Dist + (1|pop) + (1|block))},
  
  WY_Recent = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ Wtr_Year_GD_Recent + Geographic_Dist + (1|pop) + (1|block))},
  
  WY_Historical = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ Wtr_Year_GD_Historical + Geographic_Dist + (1|pop) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) 

surv_GD_fits_wl2 <- surv_GD_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_establishment_scaled))

surv_GD_fits_ucd <- surv_GD_fits %>%
  mutate(fit = map(wflow, fit, data = ucd_establishment_scaled))

surv_GD_fits_wl2 %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#water year models the best

surv_GD_fits_ucd %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#pop.block model best by AIC and BIC

surv_GD_fits_wl2 %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "GD") | term=="Geographic_Dist") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)

surv_GD_fits_ucd %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "GD") | term=="Geographic_Dist") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)
```

```{r}
surv_GD_wflow_sub <- workflow() %>%
  add_variables(outcomes = Establishment, predictors = c(pop, mf, block, contains("Dist"))) 

surv_GD_fits_sub <- tibble(wflow=list(
  pop.block = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ (1|pop) + (1|block))},
  
  GS_Recent = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ GrwSsn_TempDist_Recent + GrwSsn_PPTDist_Recent + Geographic_Dist + (1|pop) + (1|block))},
  
  GS_Historical = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ GrwSsn_TempDist_Historic + GrwSsn_PPTDist_Historic + Geographic_Dist + (1|pop) + (1|block))},
  
  WY_Recent = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ Wtr_Year_TempDist_Recent + Wtr_Year_PPTDist_Recent + Geographic_Dist + (1|pop) + (1|block))},
  
  WY_Historical = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ Wtr_Year_TempDist_Historic + Wtr_Year_PPTDist_Historic + Geographic_Dist + (1|pop) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) 

surv_GD_fits_wl2_sub <- surv_GD_fits_sub %>%
  mutate(fit = map(wflow, fit, data = wl2_establishment_scaled_sub))

surv_GD_fits_ucd_sub <- surv_GD_fits_sub %>%
  mutate(fit = map(wflow, fit, data = ucd_establishment_scaled_sub))

surv_GD_fits_wl2_sub %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#water year models the best

surv_GD_fits_ucd_sub %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#water year recent the best 

surv_GD_fits_wl2_sub %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "Dist")) %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)

surv_GD_fits_ucd_sub %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "Dist")) %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)
```

