---
title: "FirstYear_Survival"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

# First Year Survival (surv to ann census)

To Do:
- Look at relationship between size and survival
- Standard error correction on scatter plots?


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Relevant Libraries and Functions

```{r}
library(tidyverse)
library(magrittr)
#conflicted::conflicts_prefer(dplyr::filter)
library(viridis) #for pretty colors
library(ggrepel) #for non-overlapping labels on plots
library(ggdist) #visualizations of distributions and uncertainty 
library(ggpubr)

#library(sp) #for calculating geographic distance 
library(geosphere) #for calculating geographic distance
library(corrplot) #plotting correlations  
library(rstatix) #performing cor_test

library(lmerTest) #mixed models
conflicted::conflicts_prefer(lmerTest::lmer)
library(broom.mixed)
library(tidymodels)
library(furrr) #Apply Mapping Functions in Parallel using Futures
tidymodels_prefer()
# install.packages("multilevelmod")
library(multilevelmod)

library(brms)
#conflicted::conflicts_prefer(brms::ar)
#conflicted::conflicts_prefer(dplyr::combine)
#conflicted::conflicts_prefer(brms::dstudent_t)
library(tidybayes) #for extracting and visiaulizing brms model output 
library(modelr) #for data grid

sem <- function(x, na.rm=FALSE) {           #for caclulating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 

cbbPalette2 <- c("#E69F00","#000000", "#56B4E9","#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
timepd_palette <- c("#56B4E9","#D55E00")

options(mc.cores = parallel::detectCores())
```

## Gower's Distance
```{r}
garden_climate <- read_csv("../output/Climate/flint_climate_UCDpops.csv") %>% 
  rename(pop=parent.pop) %>% 
  filter(pop=="WL2_Garden" | pop=="UCD_Garden") %>% 
  select(pop:Long) %>% 
  distinct()
garden_climate
#UCD LAT/LONG = 38.53250, -121.7830
#WL2 Lat/Long = 38.82599, -120.2509

ucd_gowers <- read_csv("../output/Climate/Gowers_UCD.csv") %>% 
  rename(pop=parent.pop) %>% 
  select(pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  mutate(UCD_Lat=38.53250, UCD_Long=-121.7830, UCD_Elev=16) %>% 
  mutate(Geographic_Dist=distHaversine(cbind(UCD_Long, UCD_Lat), cbind(Long, Lat)),
         Elev_Dist=UCD_Elev-elev_m) #%>% # Calculate the distance using the haversine formula (dist in meters)
  #mutate(Lat_Dist=UCD_Lat-Lat, Long_Dist=UCD_Long-Long) %>% #Garden-Home - lat and long per Gerst et al 2011 which kept them separate for some directionality

wl2_gowers_2023 <- read_csv("../output/Climate/Gowers_WL2.csv") %>% 
  rename(pop=parent.pop) %>% 
  select(pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  mutate(WL2_Lat=38.82599, WL2_Long=-120.2509, WL2_Elev=2020) %>% 
  mutate(Geographic_Dist=distHaversine(cbind(WL2_Long, WL2_Lat), cbind(Long, Lat)),
         Elev_Dist=WL2_Elev-elev_m) #%>% # Calculate the distance using the haversine formula
  #mutate(Lat_Dist=WL2_Lat-Lat, Long_Dist=WL2_Long-Long) %>% #Garden-Home - lat and long per Gerst et al 2011 which kept them separate for some directionality
```

## Climate Subtraction Distances
```{r}
ucd_wtr_year_sub_recent <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Davis_Recent.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Recent=ann_tmean_dist, Wtr_Year_PPTDist_Recent=ann_ppt_dist)
ucd_wtr_year_sub_historic <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Davis_Historical.csv") %>%
  select(parent.pop, Wtr_Year_TempDist_Historic=ann_tmean_dist, Wtr_Year_PPTDist_Historic=ann_ppt_dist)
ucd_grwssn_sub_recent <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_Davis_Recent.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Recent=ann_tmean_dist, GrwSsn_PPTDist_Recent=ann_ppt_dist)
ucd_grwssn_sub_historic <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_Davis_Historical.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Historic=ann_tmean_dist, GrwSsn_PPTDist_Historic=ann_ppt_dist)
ucd_sub_dist <- ucd_wtr_year_sub_recent %>% 
  left_join(ucd_wtr_year_sub_historic) %>% 
  left_join(ucd_grwssn_sub_recent) %>% 
  left_join(ucd_grwssn_sub_historic) %>% 
  rename(pop=parent.pop) %>% 
  left_join(ucd_gowers)

wl2_wtr_year_sub_recent <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_WL2_Recent.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Recent=ann_tmean_dist, Wtr_Year_PPTDist_Recent=ann_ppt_dist)
wl2_wtr_year_sub_historic <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_WL2_Historical.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Historic=ann_tmean_dist, Wtr_Year_PPTDist_Historic=ann_ppt_dist)
wl2_grwssn_sub_recent <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_WL2_Recent.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Recent=ann_tmean_dist, GrwSsn_PPTDist_Recent=ann_ppt_dist)
wl2_grwssn_sub_historic <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_WL2_Historical.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Historic=ann_tmean_dist, GrwSsn_PPTDist_Historic=ann_ppt_dist)
wl2_sub_dist_2023 <- wl2_wtr_year_sub_recent %>% 
  left_join(wl2_wtr_year_sub_historic) %>% 
  left_join(wl2_grwssn_sub_recent) %>% 
  left_join(wl2_grwssn_sub_historic) %>% 
  rename(pop=parent.pop) %>% 
  left_join(wl2_gowers_2023)
```

## Load Survival data from both Gardens
```{r}
WL2_firstyear_mort <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_annual_census_20231027_corrected.csv",
                                 na = c("", "NA", "-", "N/A")) %>% 
  mutate(pop= str_replace(pop, ".*VTR.*", "LVTR1")) %>% 
  mutate(pop= str_replace(pop, "Y08", "YO8")) %>% 
  mutate(pop= str_replace(pop, "Y04", "YO4")) %>% 
  filter(!is.na(pop)) %>% 
  unite(BedLoc, bed:`bed-col`, sep="_", remove = FALSE) %>% 
  unite(Genotype, pop:rep, sep="_", remove = FALSE) %>% 
  filter(!str_detect(Genotype, ".*buff*")) %>% 
  unite(pop.mf, pop:mf, sep="_", remove = FALSE)
head(WL2_firstyear_mort)
unique(WL2_firstyear_mort$pop)
WL2_firstyear_mort %>% filter(Genotype=="CC_1_2") #there are 2 CC_1_2 plants (in different field locs), CC-1-2 was planted in 13-A and not 5C according to planting notes 
WL2_firstyear_mort %>% filter(Genotype=="IH_4_5") #there are 2 IH_4_5 plants (in different field locs), IH_4_5 was planted in 22B and not 32A according to planting notes 
WL2_firstyear_mort %>% rowwise() %>%  #checking if mf and rep can be converted to numeric 
  filter(!is.na(rep)) %>%  
  filter(is.na(as.numeric(rep)))

UCD_firstyear_mort <- read_csv("../input/UCD_Data/CorrectedCSVs/Annual_Census_Transplants_All_May_2023_corrected.csv",
                           na = c("", "NA", "-", "N/A")) %>% 
  mutate(pop= str_replace(pop, ".*VTR.*", "LVTR1")) %>% 
  unite(Genotype, pop:rep, sep="_", remove = FALSE) %>% 
  filter(!str_detect(Genotype, ".*buff*")) %>% 
  unite(pop.mf, pop:mf, sep="_", remove = FALSE)
head(UCD_firstyear_mort)
unique(UCD_firstyear_mort$pop)
UCD_firstyear_mort %>% rowwise() %>%  #checking if mf and rep can be converted to numeric 
  filter(!is.na(rep)) %>%  
  filter(is.na(as.numeric(rep)))

#UCD_firstyear_mort %>% filter(rep==100) #need to convert these to dead
```

```{r}
WL2_firstyear_mort %>% filter(pheno=="X") %>% filter(!is.na(num.fruit)) #no dead that had mature fruit. 
```

## Clean the data 
```{r}
WL2_firstyear_mort_prep <- WL2_firstyear_mort %>%
  filter(BedLoc!="K_5_C") %>% 
  filter(BedLoc!="B_32_A") %>% 
  select(block, Genotype:rep, pheno) %>% 
  mutate(Site="WL2") %>%
  mutate(mf=as.double(mf), rep=as.double(rep))
names(WL2_firstyear_mort_prep)


UCD_firstyear_mort_prep <- UCD_firstyear_mort %>% 
  select(block, Genotype:rep, pheno) %>%
  mutate(Site="UCD") %>%
  mutate(mf=as.double(mf), rep=as.double(rep))
names(UCD_firstyear_mort_prep)
```

## Establishment
```{r}
ucd_est <- read_csv("../output/UCD_Traits/UCD_Establishment.csv") %>% select(block:rep, Establishment) 

wl2_est <- read_csv("../output/WL2_Traits/WL2_Establishment.csv") %>% select(block:rep, Establishment) 
```

## Calculate First Year Survival 
```{r}
#WL2_firstyear_mort %>% filter(is.na(pheno)) %>% filter(!is.na(survey.notes)) #double checking that all NAs = dead 
#UCD_firstyear_mort %>% filter(is.na(pheno)) %>% filter(!is.na(notes))

wl2_y1_surv <- WL2_firstyear_mort_prep %>% 
  left_join(wl2_est) %>% 
  left_join(wl2_gowers_2023) %>% 
  filter(Establishment==1) %>% 
  mutate(Survival=if_else(pheno=="X" | is.na(pheno), 0, 1)) %>% 
  select(block:Genotype, pop:rep, elevation.group:Wtr_Year_GD_Historical, Geographic_Dist, Elev_Dist, pheno, Survival)
#wl2_y1_surv %>% group_by(pop) %>% summarise(n=n()) %>% arrange(n) #min sample size = 4 

ucd_y1_surv <- UCD_firstyear_mort_prep %>% 
  left_join(ucd_est) %>% 
  mutate(Establishment = if_else(Genotype=="WL2_4_100", 1, Establishment)) %>% #take into account the later germ plant that established and later died (coded as WL2_4_10 in establishment data sheet)
  left_join(ucd_gowers) %>% 
  filter(Establishment==1) %>% 
  mutate(Survival=if_else(pheno=="X" | is.na(pheno), 0, 1)) %>% 
  mutate(Survival=if_else(rep==100, 0, Survival)) %>% 
  select(block:Genotype, pop:rep, elevation.group:Wtr_Year_GD_Historical, Geographic_Dist, Elev_Dist, pheno, Survival)
#ucd_y1_surv %>% group_by(pop) %>% summarise(n=n()) %>% arrange(n) #min sample size = 2 

#ucd_y1_surv %>% filter(rep==100)
```

```{r}
wl2_y1_surv_sub_dist <- wl2_y1_surv %>% 
  select(block:rep, Survival) %>% 
  left_join(wl2_sub_dist_2023)

unique(wl2_y1_surv_sub_dist$block)
unique(wl2_y1_surv_sub_dist$pop)
unique(wl2_y1_surv_sub_dist$mf)

ucd_y1_surv_sub_dist <- ucd_y1_surv %>% 
  select(block:rep, Survival) %>% 
  left_join(ucd_sub_dist)

unique(ucd_y1_surv_sub_dist$block)
unique(ucd_y1_surv_sub_dist$pop)
unique(ucd_y1_surv_sub_dist$mf)
```

### Bar plots 
WL2
```{r}
wl2_y1_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/WL2_Traits/WL2_Y1Surv_GrwSsn_GD_Recent.png", width = 12, height = 8, units = "in")

wl2_y1_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/WL2_Traits/WL2_Y1Surv_Wtr_Year_GD_Recent.png", width = 12, height = 8, units = "in")
```

Davis
```{r}
ucd_y1_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/UCD_Traits/UCD_Y1Surv_GrwSsn_GD_Recent.png", width = 12, height = 8, units = "in")

ucd_y1_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/UCD_Traits/UCD_Y1Surv_Wtr_Year_GD_Recent.png", width = 12, height = 8, units = "in")
```

### Scatterplots
WL2
```{r}
#scatter plots
GSCD_recent <- wl2_y1_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Recent, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Recent Growth Season CD") +
  theme(text=element_text(size=25))

WYCD_recent <- wl2_y1_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Recent Water Year CD") +
  theme(text=element_text(size=25))

GD <- wl2_y1_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent, Geographic_Dist) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Geographic Distance (m)") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

ED <- wl2_y1_surv %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Elevation Distance (m)") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

wl2_y1_surv_FIG <- ggarrange(GSCD_recent, WYCD_recent, GD, ED, ncol=2, nrow=2) 
#ggsave("../output/WL2_Traits/WL2_Y1_Surv_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots
GSCD_historic <- wl2_y1_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Historical, Wtr_Year_GD_Historical) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Historical, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Historic Growth Season CD") +
  theme(text=element_text(size=25))

WYCD_historic <- wl2_y1_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Historical, Wtr_Year_GD_Historical) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Historical, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Historic Water Year CD") +
  theme(text=element_text(size=25))

wl2_y1_surv_FIG <- ggarrange(GSCD_historic, WYCD_historic, GD, ED, ncol=2, nrow=2) 
#ggsave("../output/WL2_Traits/WL2_Y1_Surv_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

Davis
```{r}
#scatter plots
GSCD_recent <- ucd_y1_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Recent, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Recent Growth Season CD") +
  theme(text=element_text(size=25))

WYCD_recent <- ucd_y1_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Recent Water Year CD") +
  theme(text=element_text(size=25))

GD <- ucd_y1_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent, Geographic_Dist) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Geographic Distance (m)") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

ED <- ucd_y1_surv %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Elevation Distance (m)") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

ucd_y1_surv_FIG <- ggarrange(GSCD_recent, WYCD_recent, GD, ED, ncol=2, nrow=2) 
#ggsave("../output/UCD_Traits/UCD_Y1_Surv_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots
GSCD_historic <- ucd_y1_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Historical, Wtr_Year_GD_Historical) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Historical, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Historic Growth Season CD") +
  theme(text=element_text(size=25))

WYCD_historic <- ucd_y1_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Historical, Wtr_Year_GD_Historical) %>% 
  summarise(meanSurv=mean(Survival, na.rm = TRUE), semSurv=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Historical, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Historic Water Year CD") +
  theme(text=element_text(size=25))

ucd_y1_surv_FIG <- ggarrange(GSCD_historic, WYCD_historic, GD, ED, ncol=2, nrow=2) 
#ggsave("../output/UCD_Traits/UCD_Y1_Surv_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

#### Directional Distance
```{r}
#scatter plots - recent
GSCD_prob_recent <- wl2_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Recent Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- wl2_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Recent Water Year Temp Dist") +
  theme(text=element_text(size=30))

GD_prob <- wl2_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, Wtr_Year_TempDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- wl2_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

WL2_y1_surv_sub_FIG_prob_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_Y1_Surv_TmpSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- wl2_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Historic Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- wl2_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Historic Water Year Temp Dist") +
  theme(text=element_text(size=30))

WL2_y1_surv_sub_FIG_prob_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_Y1_Surv_TmpSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
GSCD_prob_recent <- ucd_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Recent Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- ucd_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Recent Water Year Temp Dist") +
  theme(text=element_text(size=30))

GD_prob <- ucd_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, Wtr_Year_TempDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- ucd_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

UCD_y1_surv_sub_FIG_prob_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/UCD_Traits/UCD_Y1_Surv_TmpSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- ucd_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Historic Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- ucd_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Historic Water Year Temp Dist") +
  theme(text=element_text(size=30))

UCD_y1_surv_sub_FIG_prob_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/UCD_Traits/UCD_Y1_Surv_TmpSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
GSCD_prob_recent <- wl2_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=0.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Recent Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- wl2_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Recent Water Year PPT Dist") +
  theme(text=element_text(size=30))

GD_prob <- wl2_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, Wtr_Year_PPTDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- wl2_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

WL2_y1_surv_sub_FIG_prob_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_Y1_Surv_PPTSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- wl2_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Historic Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- wl2_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Historic Water Year PPT Dist") +
  theme(text=element_text(size=30))

WL2_y1_surv_sub_FIG_prob_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_Y1_Surv_PPTSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
GSCD_prob_recent <- ucd_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Recent Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- ucd_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Recent Water Year PPT Dist") +
  theme(text=element_text(size=30))

GD_prob <- ucd_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, Wtr_Year_PPTDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- ucd_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

UCD_y1_surv_sub_FIG_prob_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/UCD_Traits/UCD_Y1_Surv_PPTSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- ucd_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Historic Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- ucd_y1_surv_sub_dist %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Survival, na.rm = TRUE), semEst=sem(Survival, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Y1 Survival", x="Historic Water Year PPT Dist") +
  theme(text=element_text(size=30))

UCD_y1_surv_sub_FIG_prob_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/UCD_Traits/UCD_Y1_Surv_PPTSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

## Stats

### Scaling 
```{r}
wl2_y1_surv_scaled <- wl2_y1_surv %>% mutate_at(c("GrwSsn_GD_Recent","Wtr_Year_GD_Recent",                                                           "GrwSsn_GD_Historical","Wtr_Year_GD_Historical","Geographic_Dist"),
                                                            scale) 

ucd_y1_surv_scaled <- ucd_y1_surv %>% mutate_at(c("GrwSsn_GD_Recent","Wtr_Year_GD_Recent",                                                           "GrwSsn_GD_Historical","Wtr_Year_GD_Historical","Geographic_Dist"),
                                                            scale) 

#CHECK MODEL PARAMS
unique(wl2_y1_surv_scaled$pop)
unique(wl2_y1_surv_scaled$mf)
unique(wl2_y1_surv_scaled$block)
summary(wl2_y1_surv_scaled)

unique(ucd_y1_surv_scaled$pop)
unique(ucd_y1_surv_scaled$mf)
unique(ucd_y1_surv_scaled$block)
summary(ucd_y1_surv_scaled)
```

```{r}
wl2_y1_surv_scaled_sub <- wl2_y1_surv_sub_dist %>% 
  mutate_at(c("Wtr_Year_TempDist_Recent",  "Wtr_Year_PPTDist_Recent", 
                 "Wtr_Year_TempDist_Historic", "Wtr_Year_PPTDist_Historic",
                 "GrwSsn_TempDist_Recent", "GrwSsn_PPTDist_Recent",
                 "GrwSsn_TempDist_Historic", "GrwSsn_PPTDist_Historic",
              "Geographic_Dist"), scale)

ucd_y1_surv_scaled_sub <- ucd_y1_surv_sub_dist %>% 
  mutate_at(c("Wtr_Year_TempDist_Recent",  "Wtr_Year_PPTDist_Recent", 
                 "Wtr_Year_TempDist_Historic", "Wtr_Year_PPTDist_Historic",
                 "GrwSsn_TempDist_Recent", "GrwSsn_PPTDist_Recent",
                 "GrwSsn_TempDist_Historic", "GrwSsn_PPTDist_Historic",
              "Geographic_Dist"), scale)
```


### Basic Model Workflow 
```{r}
glmer.model_binomial <- 
  linear_reg() %>% 
  set_engine("glmer", family=binomial)

surv_wflow <- workflow() %>% 
  add_variables(outcomes = Survival, predictors = c(pop, mf, block))

surv_fits <- tibble(wflow=list(
  pop = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = Survival ~ (1|pop))},
  
  pop.mf = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = Survival ~ (1|pop/mf))},
  
  pop.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = Survival ~ (1|pop) + (1|block))},
  
  pop.mf.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = Survival ~ (1|pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)

surv_fits_wl2 <- surv_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_y1_surv_scaled))
#mod_test <- glmer(Survival ~ (1|pop/mf) + (1|block), data=wl2_y1_surv_scaled, family=binomial)
#summary(mod_test)
#no issues with mf 

surv_fits_ucd <- surv_fits %>%
  mutate(fit = map(wflow, fit, data = ucd_y1_surv_scaled))
#mod_test <- glmer(Survival ~ (1|pop/mf) + (1|block), data=ucd_y1_surv_scaled, family=binomial)

surv_fits_wl2 %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#pop block model best by AIC and BIC, full model close behind

surv_fits_ucd %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#pop.mf model best by AIC and BIC, but full model = close behind and has no issues so will use that 
```

#### Test climate and geographic distance 
```{r}
surv_GD_wflow <- workflow() %>%
  add_variables(outcomes = Survival, predictors = c(pop, mf, block, contains("GD"), Geographic_Dist)) 

surv_GD_fits <- tibble(wflow=list(
  pop.block = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Survival ~ (1|pop/mf) + (1|block))},
  
  GS_Recent = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Survival ~ GrwSsn_GD_Recent + Geographic_Dist + (1|pop/mf) + (1|block))},
  
  GS_Historical = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Survival ~ GrwSsn_GD_Historical + Geographic_Dist + (1|pop/mf) + (1|block))},
  
  WY_Recent = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Survival ~ Wtr_Year_GD_Recent + Geographic_Dist + (1|pop/mf) + (1|block))},
  
  WY_Historical = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Survival ~ Wtr_Year_GD_Historical + Geographic_Dist + (1|pop/mf) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) 

surv_GD_fits_wl2 <- surv_GD_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_y1_surv_scaled))

surv_GD_fits_ucd <- surv_GD_fits %>%
  mutate(fit = map(wflow, fit, data = ucd_y1_surv_scaled))

surv_GD_fits_wl2 %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#water year models the best

surv_GD_fits_ucd %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#water year models the best 

surv_GD_fits_wl2 %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "GD") | term=="Geographic_Dist") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)

surv_GD_fits_ucd %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "GD") | term=="Geographic_Dist") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)

#mod_test <- glmer(Survival ~ Wtr_Year_GD_Historical + Geographic_Dist + (1|pop/mf) + (1|block), data=ucd_y1_surv_scaled, family=binomial)
#summary(mod_test)
```

```{r}
surv_GD_wflow_sub <- workflow() %>%
  add_variables(outcomes = Survival, predictors = c(pop, mf, block, contains("Dist"))) 

surv_GD_fits_sub <- tibble(wflow=list(
  pop.block = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Survival ~ (1|pop/mf) + (1|block))},
  
  GS_Recent = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Survival ~ GrwSsn_TempDist_Recent + GrwSsn_PPTDist_Recent + Geographic_Dist + (1|pop/mf) + (1|block))},
  
  GS_Historical = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Survival ~ GrwSsn_TempDist_Historic + GrwSsn_PPTDist_Historic + Geographic_Dist + (1|pop/mf) + (1|block))},
  
  WY_Recent = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Survival ~ Wtr_Year_TempDist_Recent + Wtr_Year_PPTDist_Recent + Geographic_Dist + (1|pop/mf) + (1|block))},
  
  WY_Historical = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Survival ~ Wtr_Year_TempDist_Historic + Wtr_Year_PPTDist_Historic + Geographic_Dist + (1|pop/mf) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) 

surv_GD_fits_wl2_sub <- surv_GD_fits_sub %>%
  mutate(fit = map(wflow, fit, data = wl2_y1_surv_scaled_sub))

surv_GD_fits_ucd_sub <- surv_GD_fits_sub %>%
  mutate(fit = map(wflow, fit, data = ucd_y1_surv_scaled_sub))

surv_GD_fits_wl2_sub %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#water year models the best

surv_GD_fits_ucd_sub %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#water year models the best 

surv_GD_fits_wl2_sub %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "Dist")) %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)

surv_GD_fits_ucd_sub %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "Dist")) %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)
```


## Reaction Norms (old code)
```{r, eval=FALSE}
firstyear_mort_rxnnorms <- bind_rows(UCD_firstyear_mort_prep, WL2_firstyear_mort_prep) %>% 
  mutate(Survival=if_else(pheno=="X" | is.na(pheno), 0, 1)) %>% 
  arrange(Genotype, Site)
xtabs(~pop.mf+Site, data=firstyear_mort_rxnnorms) #some pop/mfs without data at one of the sites 
head(firstyear_mort_rxnnorms)

firstyear_mort_rxnnorms_loc <- left_join(firstyear_mort_rxnnorms, gowersdist_all)
head(firstyear_mort_rxnnorms_loc)
#write_csv(firstyear_mort_rxnnorms_loc, "../output/firstyear_mort_both_sites.csv")
```

### Means by maternal family
```{r, eval=FALSE}
firstyear_mort_rxnnorms_mfs <- firstyear_mort_rxnnorms_loc %>% 
  group_by(pop.mf, pop, Site, elev_m, Recent_Gowers_Dist_WL2) %>% 
  summarise(N_Surv = sum(!is.na(Survival)), mean_Surv = mean(Survival,na.rm=(TRUE)), 
            sem_surv=sem(Survival, na.rm=(TRUE)))
firstyear_mort_rxnnorms_mfs %>% arrange(N_Surv)

#restrict to mfs with plants at both sites
firstyear_mort_rxnnorms_summary_mfs_wide <- firstyear_mort_rxnnorms_mfs %>% 
  select(pop.mf, Site, N_Surv) %>% 
  spread(Site, N_Surv) %>% 
  mutate(Both.Sites=if_else(!is.na(UCD) & !is.na(WL2), TRUE, FALSE)) %>% 
 filter(Both.Sites != "FALSE")
firstyear_mort_rxnnorms_summary_mfs_wide %>% arrange(Both.Sites)

firstyear_mort_rxnnorms_summary_mfs_bothsites <- left_join(firstyear_mort_rxnnorms_summary_mfs_wide, firstyear_mort_rxnnorms_mfs)
head(firstyear_mort_rxnnorms_summary_mfs_bothsites)
unique(firstyear_mort_rxnnorms_summary_mfs_bothsites$pop) #still have all 23 pops 

firstyear_mort_rxnnorms_summary_mfs2 <- firstyear_mort_rxnnorms_summary_mfs_bothsites %>% 
  mutate(Site=str_replace_all(Site, "UCD", "Low Elev"), 
         Site=str_replace_all(Site, "WL2", "High Elev")) 
firstyear_mort_rxnnorms_summary_mfs2$Site <- factor(firstyear_mort_rxnnorms_summary_mfs2$Site,
                                               levels = c('Low Elev','High Elev'))

firstyear_mort_rxnnorms_summary_mfs2 %>% arrange(desc(mean_Surv))
firstyear_mort_rxnnorms_summary_mfs2 %>% 
  filter(N_Surv == 1) #26 maternal families with only 1 plant at one of the sites
```

#### Plot mf avgs 
```{r, eval=FALSE}
firstyear_mort_rxnnorms_summary_mfs2 %>% 
   ggplot(aes(x=Site, y=mean_Surv, group=pop.mf, color=pop.mf)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() +
   theme(text=element_text(size=25))
```

```{r, eval=FALSE}
#note that there are a lot of maternal families with only 1 plant - removed those from these figures 
firstyear_mort_rxnnorms_summary_mfs2 %>% 
  filter(N_Surv != 1) %>% 
  ggplot(aes(x=Site, y=mean_Surv, group=pop.mf, color=elev_m)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
 geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Survival", color="Elevation (m)") +
  theme(text=element_text(size=28))
#ggsave("../output/Surv_FirstYear_RxNorms_MFAvgs.png", width = 14, height = 8, units = "in")
```

```{r, eval=FALSE}
firstyear_mort_rxnnorms_summary_mfs2 %>% 
  filter(N_Surv != 1) %>% 
  filter(pop=="TM2"|pop=="WL2") %>%  
  ggplot(aes(x=Site, y=mean_Surv, group=pop.mf, color=pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() + 
  labs(y="Survival", color="Elevation (m)") +
  theme(text=element_text(size=25))
#ggsave("../output/Surv_FirstYear_RxNorms_TM2_WL2_MFAvgs.png", width = 12, height = 8, units = "in")

firstyear_mort_rxnnorms_summary_mfs2 %>% 
  filter(N_Surv != 1) %>% 
  filter(pop=="WL2") %>%  
  ggplot(aes(x=Site, y=mean_Surv, group=pop.mf, color=pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() + 
  labs(y="Survival", color="Elevation (m)") +
  theme(text=element_text(size=25))
```

### Means by Pop
```{r, eval=FALSE}
xtabs(~pop+Survival, data=firstyear_mort_rxnnorms_loc)
firstyear_mort_rxnnorms_summary <- firstyear_mort_rxnnorms_loc %>% 
  group_by(pop, Site, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(Survival)), mean_Surv = mean(Survival,na.rm=(TRUE)), 
            sem_surv=sem(Survival, na.rm=(TRUE)))
firstyear_mort_rxnnorms_summary

firstyear_mort_rxnnorms_summary %>% arrange(N_Surv)
firstyear_mort_rxnnorms_summary %>% arrange(desc(mean_Surv))

firstyear_mort_rxnnorms_summary2 <- firstyear_mort_rxnnorms_summary %>% 
  mutate(Site=str_replace_all(Site, "UCD", "Low Elev"), 
         Site=str_replace_all(Site, "WL2", "High Elev"))
firstyear_mort_rxnnorms_summary2$Site <- factor(firstyear_mort_rxnnorms_summary2$Site,
                                               levels = c('Low Elev','High Elev'))

```

#### Plot Pop Avgs
```{r, eval=FALSE}
firstyear_mort_rxnnorms_summary %>% 
   ggplot(aes(x=Site, y=mean_Surv, group=pop, color=pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() +
   theme(text=element_text(size=25))
```

```{r, eval=FALSE}
firstyear_mort_rxnnorms_summary2 %>% 
  filter(N_Surv != 1) %>% 
  ggplot(aes(x=Site, y=mean_Surv, group=pop, color=elev_m)) + 
  geom_point(size=1.5) + geom_line(linewidth=1.5) +
  geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Survival", color="Elevation (m)") + #ylim(0,1) +
  theme(text=element_text(size=28))
#ggsave("../output/Surv_FirstYear_RxNorms_ALL_PopAvgs.png", width = 14, height = 8, units = "in")
```

```{r, eval=FALSE}
firstyear_mort_rxnnorms_summary2 %>% 
  filter(N_Surv != 1) %>% 
  filter(Site=="High Elev") %>% 
  ggplot(aes(x=fct_reorder(pop, mean_Surv), y=mean_Surv, group=pop, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() + scale_fill_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Survival", fill="Elevation (m)", x="Population") + #ylim(0,1) +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Surv_FirstYear_WL2ONLY_ALL_PopAvgs.png", width = 14, height = 8, units = "in")
```

Combine pop and mf avgs 
```{r, eval=FALSE}
firstyear_mort_rxnnorms_summary2 %>% 
  mutate(pop.mf=pop) %>% 
  filter(N_Surv != 1) %>% 
  ggplot(aes(x=Site, y=mean_Surv, group=pop.mf, color=elev_m)) + 
  geom_point(size=1.5) + geom_line(linewidth=1.5, alpha=0.7) +
  geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Survival", color="Elevation (m)") + #ylim(0,1) +
  theme(text=element_text(size=28)) +
 geom_line(data = firstyear_mort_rxnnorms_summary_mfs2, size=0.2) +
  geom_point(data = firstyear_mort_rxnnorms_summary_mfs2, size=0.8) 
#ggsave("../output/Surv_FirstYear_RxNorms_ALL_PopMFAvgs.png", width = 14, height = 8, units = "in")
```

```{r, eval=FALSE}
firstyear_mort_rxnnorms_summary2 %>% 
  filter(!is.na(pop), pop!="YO8*") %>% 
  filter(N_Surv != 1) %>% 
  filter(pop=="TM2"|pop=="WL2") %>%  
  ggplot(aes(x=Site, y=mean_Surv, group=pop, color=pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() + 
 ylim(0,1) +
  labs(y="Survival", color="Elevation (m)") +
  theme(text=element_text(size=25))
#ggsave("../output/Surv_FirstYear_RxNorms_TM2_WL2_PopAvgs.png", width = 12, height = 8, units = "in")
```

### Stats

#### Combine size (two months) and survival (three months)
```{r, eval=FALSE}
size_survival <- left_join(twomonths_rxnnorms_loc, firstyear_mort_rxnnorms_loc) %>% 
  #twomonths_rxnnorms_loc is from "Single_Time_Size_RxnNorms.Rmd"
  mutate(mf=as.character(mf))
head(size_survival)
names(size_survival)
summary(size_survival)
```

#### Log Reg Survival 
```{r, eval=FALSE}
lmesurv1 <- glmer(Survival ~ elev_m*Site + (1|pop/mf), 
                  data = size_survival, 
                  family = binomial(link = "logit"), 
                  nAGQ=0, #uses the adaptive Gaussian quadrature instead the Laplace approximation. Ask Julin about this 
                  control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))) 
lmesurv2 <- glmer(Survival ~ elev_m*Site + (1|pop/mf), data = size_survival, family = binomial(link = "logit"), nAGQ=0) 
lmesurv3 <- glmer(Survival ~ elev_m*Site + (1|pop/mf), data = size_survival, family = binomial(link = "logit"), control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))) #this one results in a warnign 
summary(lmesurv1)
anova(lmesurv1)
glance(lmesurv1)
emtrends(lmesurv1, pairwise ~ Site, var = "elev_m")
emmip(lmesurv1, Site ~ elev_m, cov.reduce = range) #low elev pops have a steeper increase in survival from low to high elev garden than higher elev pops 
#ranova(lmesurv) doesn't work for lme4
#to test for significance of random effect: 
#the most common way to do this is to use a likelihood ratio test, i.e. fit the full and reduced models (the reduced model is the model with the focal variance(s) set to zero). 
m0 <- glmer(Survival ~ elev_m + (1|pop), data = size_survival, family = binomial("logit")) 
m00 <- glm(Survival~ elev_m, size_survival, family = binomial("logit"))
anova(lmesurv1,m0, m00) #model with both random effects has a higher likelihood (better fit)
```

#### Log Reg survival ~ size 
Height
```{r, eval=FALSE}
basiclogit_height <- glm(Survival ~ height.cm, data = size_survival, family = "binomial")
basiclogit_height
summary(basiclogit_height)

logit_height <- glm(Survival ~ height.cm + Site, data = size_survival, family = "binomial")
summary(logit_height)

logit_height2 <- glmer(Survival ~ height.cm*Site + (1|pop/mf), data = size_survival, family = binomial(link = "logit"), nAGQ=0) 
summary(logit_height2)
emtrends(logit_height2, pairwise ~ Site, var = "height.cm")
emmip(logit_height2, Site ~ height.cm, cov.reduce = range)
```

Length
```{r, eval=FALSE}
logit_length2 <- glmer(Survival ~ long.leaf.cm*Site + (1|pop/mf), data = size_survival, family = binomial(link = "logit"), nAGQ=0) 
summary(logit_length2)
emtrends(logit_length2, pairwise ~ Site, var = "long.leaf.cm")
emmip(logit_length2, Site ~ long.leaf.cm, cov.reduce = range)
```


#### Figures 
```{r, eval=FALSE}
size_survival %>% 
  ggplot(aes(x=height.cm, y=Survival)) +
  geom_point()
```

```{r, eval=FALSE}
size_surv_pop_avgs <- size_survival %>% 
  group_by(pop, Site, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(Survival)), mean_Surv = mean(Survival,na.rm=(TRUE)), 
            sem_surv=sem(Survival, na.rm=(TRUE)), 
            N_height = sum(!is.na(height.cm)), mean_height.cm = mean(height.cm,na.rm=(TRUE)), 
            sem_height.cm=sem(height.cm, na.rm=(TRUE)), N_length = sum(!is.na(long.leaf.cm)),
            mean_long.leaf.cm=mean(long.leaf.cm, na.rm=(TRUE)), 
            sem_long.leaf.cm=sem(long.leaf.cm, na.rm=TRUE))

size_surv_pop_avgs %>% arrange(desc(mean_height.cm))

size_surv_pop_avgs2 <- size_surv_pop_avgs %>% 
  mutate(Site=str_replace_all(Site, "UCD", "Low Elev"), 
         Site=str_replace_all(Site, "WL2", "High Elev"))
size_surv_pop_avgs2$Site <- factor(size_surv_pop_avgs2$Site,
                                               levels = c('Low Elev','High Elev'))
```

```{r, eval=FALSE}
size_surv_pop_avgs2 %>% 
  ggplot(aes(x=mean_height.cm, y=mean_Surv, group=pop, color=elev_m)) +
  geom_point(size=8) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Height (cm)" ,y="Survival", color="Elevation (m)") +
  theme(text=element_text(size=28)) +
  facet_wrap(vars(Site), scales="free")
#ggsave("../output/Surv_Height_BothGardens.png", width = 18, height = 8, units = "in")

size_surv_pop_avgs2 %>% 
  ggplot(aes(x=mean_long.leaf.cm, y=mean_Surv, group=pop, color=elev_m)) +
  geom_point(size=8) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Leaf Length (cm)" ,y="Survival", color="Elevation (m)") +
  theme(text=element_text(size=28)) +
  facet_wrap(vars(Site), scales="free")
#ggsave("../output/Surv_Length_BothGardens.png", width = 18, height = 8, units = "in")

size_surv_pop_avgs %>% 
  filter(Site=="UCD") %>% 
  ggplot(aes(x=mean_height.cm, y=mean_Surv, group=pop, color=elev_m)) +
  geom_point(size=1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Height (cm)" ,y="Survival", color="Elevation (m)", title="Low Elevation Garden") +
  theme(text=element_text(size=25))

size_surv_pop_avgs %>% 
  filter(Site=="UCD") %>% 
  ggplot(aes(x=mean_height.cm, y=mean_Surv)) +
  geom_point(size=1) +
  geom_smooth() +
  theme_classic()  +
  labs(x="Height (cm)" ,y="Survival", color="Elevation (m)", title="Low Elevation Garden") +
  theme(text=element_text(size=25))

size_surv_pop_avgs %>% 
  filter(Site=="WL2") %>% 
  ggplot(aes(x=mean_height.cm, y=mean_Surv, group=pop, color=elev_m)) +
  geom_point(size=1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Height (cm)" ,y="Survival", color="Elevation (m)", title="High Elevation Garden") +
  theme(text=element_text(size=25))

size_surv_pop_avgs %>% 
  filter(Site=="WL2") %>% 
  ggplot(aes(x=mean_height.cm, y=mean_Surv)) +
  geom_point(size=1) +
  geom_smooth() +
  theme_classic()  +
  labs(x="Height (cm)" ,y="Survival", color="Elevation (m)", title="High Elevation Garden") +
  theme(text=element_text(size=25))
```




