---
title: "FirstYear_Survival"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: html_document
---

To Do:
- Look at relationship between size and survival
- Do stats on size and survival separately 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Relevant Libraries and Functions

```{r}
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
library(lmerTest) #for mixed effect models
conflicted::conflicts_prefer(lmerTest::lmer)
library(broom.mixed) #tidy method for lmerTest
library(emmeans) #for post-hoc pairwise comparisons 
library(naniar) #replaces values with NA
library(corrplot) #plotting correlations 
library(rstatix) #performing cor_test
sem <- function(x, na.rm=FALSE) {           #for caclulating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low") #for proper arrangement in figures 

#install.packages("viridis") #for more better color palettes (if want to distinguish between all pops)
#library(viridis)
```

## Load the pop and location data and envtal dist data 

```{r}
gowersdist_UCD <- read_csv("../output/Climate/Pops_GowersEnvtalDist_UCD.csv") %>% 
  rename(Recent_Gowers_Dist_UCD = Recent_Gowers_Dist, Historic_Gowers_Dist_UCD = Historic_Gowers_Dist)
head(gowersdist_UCD)
gowersdist_WL2 <- read_csv("../output/Climate/Pops_GowersEnvtalDist_WL2.csv") %>% 
  rename(Recent_Gowers_Dist_WL2 = Recent_Gowers_Dist, Historic_Gowers_Dist_WL2 = Historic_Gowers_Dist)
head(gowersdist_WL2)

gowersdist_all <- full_join(gowersdist_UCD, gowersdist_WL2)
head(gowersdist_all)
```


## Load Survival data from both Gardens
```{r}
WL2_firstyear_mort <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_annual_census_20231027_corrected.csv",
                                 na = c("", "NA", "-", "N/A")) %>% 
  rename(parent.pop=pop) %>% 
  mutate(parent.pop= str_replace(parent.pop, ".*VTR.*", "LVTR1")) %>% 
  mutate(parent.pop= str_replace(parent.pop, "Y08", "YO8")) %>% 
  mutate(parent.pop= str_replace(parent.pop, "Y04", "YO4")) %>% 
  filter(!is.na(parent.pop)) %>% 
  unite(BedLoc, bed:`bed-col`, sep="_", remove = FALSE) %>% 
  unite(Genotype, parent.pop:rep, sep="_", remove = FALSE) %>% 
  filter(!str_detect(Genotype, ".*buff*")) %>% 
  unite(pop.mf, parent.pop:mf, sep="_", remove = FALSE)
head(WL2_firstyear_mort)
unique(WL2_firstyear_mort$parent.pop)
WL2_firstyear_mort %>% filter(Genotype=="CC_1_2") #there are 2 CC_1_2 plants (in different field locs), CC-1-2 was planted in 13-A and not 5C according to planting notes 
WL2_firstyear_mort %>% rowwise() %>%  #checking if mf and rep can be converted to numeric 
  filter(!is.na(rep)) %>%  
  filter(is.na(as.numeric(rep)))

UCD_firstyear_mort <- read_csv("../input/UCD_Data/CorrectedCSVs/Annual_Census_Transplants_All_May_2023_corrected.csv",
                           na = c("", "NA", "-", "N/A")) %>% 
  rename(parent.pop=pop) %>% 
  filter(rep != 100) %>%  #get rid of individuals that germinated in the field 
  mutate(parent.pop= str_replace(parent.pop, ".*VTR.*", "LVTR1")) %>% 
  unite(Genotype, parent.pop:rep, sep="_", remove = FALSE) %>% 
  filter(!str_detect(Genotype, ".*buff*")) %>% 
  unite(pop.mf, parent.pop:mf, sep="_", remove = FALSE)
head(UCD_firstyear_mort)
unique(UCD_firstyear_mort$parent.pop)
UCD_firstyear_mort %>% rowwise() %>%  #checking if mf and rep can be converted to numeric 
  filter(!is.na(rep)) %>%  
  filter(is.na(as.numeric(rep)))
```

## Merge the Data
```{r}
WL2_firstyear_mort_prep <- WL2_firstyear_mort %>%
  filter(BedLoc!="K_5_C") %>% 
  select(Genotype:rep, pheno) %>% 
  mutate(Site="WL2") %>%
  mutate(mf=as.double(mf), rep=as.double(rep))
names(WL2_firstyear_mort_prep)


UCD_firstyear_mort_prep <- UCD_firstyear_mort %>% 
  select(Genotype:rep, pheno) %>%
  mutate(Site="UCD") %>%
  mutate(mf=as.double(mf), rep=as.double(rep))
names(UCD_firstyear_mort_prep)

firstyear_mort_rxnnorms <- bind_rows(UCD_firstyear_mort_prep, WL2_firstyear_mort_prep) %>% 
  mutate(Survival=if_else(pheno=="X" | is.na(pheno), 0, 1)) %>% 
  arrange(Genotype, Site)
xtabs(~pop.mf+Site, data=firstyear_mort_rxnnorms) #some pop/mfs without data at one of the sites 
head(firstyear_mort_rxnnorms)
```

## Add in location info
```{r}
firstyear_mort_rxnnorms_loc <- left_join(firstyear_mort_rxnnorms, gowersdist_all)
head(firstyear_mort_rxnnorms_loc)
```

## Means by maternal family 
```{r}
firstyear_mort_rxnnorms_mfs <- firstyear_mort_rxnnorms_loc %>% 
  group_by(pop.mf, parent.pop, Site, elev_m, Recent_Gowers_Dist_WL2) %>% 
  summarise(N_Surv = sum(!is.na(Survival)), mean_Surv = mean(Survival,na.rm=(TRUE)), 
            sem_surv=sem(Survival, na.rm=(TRUE)))
firstyear_mort_rxnnorms_mfs %>% arrange(N_Surv)

#restrict to mfs with plants at both sites
firstyear_mort_rxnnorms_summary_mfs_wide <- firstyear_mort_rxnnorms_mfs %>% 
  select(pop.mf, Site, N_Surv) %>% 
  spread(Site, N_Surv) %>% 
  mutate(Both.Sites=if_else(!is.na(UCD) & !is.na(WL2), TRUE, FALSE)) %>% 
 filter(Both.Sites != "FALSE")
firstyear_mort_rxnnorms_summary_mfs_wide %>% arrange(Both.Sites)

firstyear_mort_rxnnorms_summary_mfs_bothsites <- left_join(firstyear_mort_rxnnorms_summary_mfs_wide, firstyear_mort_rxnnorms_mfs)
head(firstyear_mort_rxnnorms_summary_mfs_bothsites)
unique(firstyear_mort_rxnnorms_summary_mfs_bothsites$parent.pop) #still have all 23 pops 

firstyear_mort_rxnnorms_summary_mfs2 <- firstyear_mort_rxnnorms_summary_mfs_bothsites %>% 
  mutate(Site=str_replace_all(Site, "UCD", "Low Elev"), 
         Site=str_replace_all(Site, "WL2", "High Elev")) 
firstyear_mort_rxnnorms_summary_mfs2$Site <- factor(firstyear_mort_rxnnorms_summary_mfs2$Site,
                                               levels = c('Low Elev','High Elev'))

firstyear_mort_rxnnorms_summary_mfs2 %>% arrange(desc(mean_Surv))
firstyear_mort_rxnnorms_summary_mfs2 %>% 
  filter(N_Surv == 1) #26 maternal families with only 1 plant at one of the sites
```

### Plot mf avgs
```{r}
firstyear_mort_rxnnorms_summary_mfs2 %>% 
   ggplot(aes(x=Site, y=mean_Surv, group=pop.mf, color=pop.mf)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() +
   theme(text=element_text(size=25))
```

```{r}
#note that there are a lot of maternal families with only 1 plant - removed those from these figures 
firstyear_mort_rxnnorms_summary_mfs2 %>% 
  filter(N_Surv != 1) %>% 
  ggplot(aes(x=Site, y=mean_Surv, group=pop.mf, color=elev_m)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
 geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Survival", color="Elevation (m)") +
  theme(text=element_text(size=28))
ggsave("../output/Surv_FirstYear_RxNorms_MFAvgs.png", width = 14, height = 8, units = "in")
```

```{r}
firstyear_mort_rxnnorms_summary_mfs2 %>% 
  filter(N_Surv != 1) %>% 
  filter(parent.pop=="TM2"|parent.pop=="WL2") %>%  
  ggplot(aes(x=Site, y=mean_Surv, group=pop.mf, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() + 
  labs(y="Survival", color="Elevation (m)") +
  theme(text=element_text(size=25))
ggsave("../output/Surv_FirstYear_RxNorms_TM2_WL2_MFAvgs.png", width = 12, height = 8, units = "in")

firstyear_mort_rxnnorms_summary_mfs2 %>% 
  filter(N_Surv != 1) %>% 
  filter(parent.pop=="WL2") %>%  
  ggplot(aes(x=Site, y=mean_Surv, group=pop.mf, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() + 
  labs(y="Survival", color="Elevation (m)") +
  theme(text=element_text(size=25))
```


## Means by Pop
```{r}
firstyear_mort_rxnnorms_summary <- firstyear_mort_rxnnorms_loc %>% 
  group_by(parent.pop, Site, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(Survival)), mean_Surv = mean(Survival,na.rm=(TRUE)), 
            sem_surv=sem(Survival, na.rm=(TRUE)))
firstyear_mort_rxnnorms_summary

firstyear_mort_rxnnorms_summary %>% arrange(N_Surv)
firstyear_mort_rxnnorms_summary %>% arrange(desc(mean_Surv))

firstyear_mort_rxnnorms_summary2 <- firstyear_mort_rxnnorms_summary %>% 
  mutate(Site=str_replace_all(Site, "UCD", "Low Elev"), 
         Site=str_replace_all(Site, "WL2", "High Elev"))
firstyear_mort_rxnnorms_summary2$Site <- factor(firstyear_mort_rxnnorms_summary2$Site,
                                               levels = c('Low Elev','High Elev'))

```

### Plot Pop Avgs
```{r}
firstyear_mort_rxnnorms_summary %>% 
   ggplot(aes(x=Site, y=mean_Surv, group=parent.pop, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() +
   theme(text=element_text(size=25))
```

```{r}
firstyear_mort_rxnnorms_summary2 %>% 
  filter(N_Surv != 1) %>% 
  ggplot(aes(x=Site, y=mean_Surv, group=parent.pop, color=elev_m)) + 
  geom_point(size=1.5) + geom_line(linewidth=1.5) +
  geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Survival", color="Elevation (m)") + #ylim(0,1) +
  theme(text=element_text(size=28))
ggsave("../output/Surv_FirstYear_RxNorms_ALL_PopAvgs.png", width = 14, height = 8, units = "in")
```

Combine pop and mf avgs 
```{r}
firstyear_mort_rxnnorms_summary2 %>% 
  mutate(pop.mf=parent.pop) %>% 
  filter(N_Surv != 1) %>% 
  ggplot(aes(x=Site, y=mean_Surv, group=pop.mf, color=elev_m)) + 
  geom_point(size=1.5) + geom_line(linewidth=1.5, alpha=0.7) +
  geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Survival", color="Elevation (m)") + #ylim(0,1) +
  theme(text=element_text(size=28)) +
 geom_line(data = firstyear_mort_rxnnorms_summary_mfs2, size=0.2) +
  geom_point(data = firstyear_mort_rxnnorms_summary_mfs2, size=0.8) 
ggsave("../output/Surv_FirstYear_RxNorms_ALL_PopMFAvgs.png", width = 14, height = 8, units = "in")
```

```{r}
firstyear_mort_rxnnorms_summary2 %>% 
  filter(!is.na(parent.pop), parent.pop!="YO8*") %>% 
  filter(N_Surv != 1) %>% 
  filter(parent.pop=="TM2"|parent.pop=="WL2") %>%  
  ggplot(aes(x=Site, y=mean_Surv, group=parent.pop, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_Surv-sem_surv,ymax=mean_Surv+sem_surv),width=.1) +
  theme_classic() + 
 ylim(0,1) +
  labs(y="Survival", color="Elevation (m)") +
  theme(text=element_text(size=25))
ggsave("../output/Surv_FirstYear_RxNorms_TM2_WL2_PopAvgs.png", width = 12, height = 8, units = "in")
```

## Stats


### Combine size (two months) and survival (three months)
```{r}
size_survival <- left_join(twomonths_rxnnorms_loc, firstyear_mort_rxnnorms_loc) %>% 
  #twomonths_rxnnorms_loc is from "Single_Time_Size_RxnNorms.Rmd"
  mutate(mf=as.character(mf))
head(size_survival)
names(size_survival)
summary(size_survival)
```

### Log Reg Survival 
```{r}
lmesurv1 <- glmer(Survival ~ elev_m*Site + (1|parent.pop/mf), 
                  data = size_survival, 
                  family = binomial(link = "logit"), 
                  nAGQ=0, #uses the adaptive Gaussian quadrature instead the Laplace approximation. Ask Julin about this 
                  control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))) 
lmesurv2 <- glmer(Survival ~ elev_m*Site + (1|parent.pop/mf), data = size_survival, family = binomial(link = "logit"), nAGQ=0) 
lmesurv3 <- glmer(Survival ~ elev_m*Site + (1|parent.pop/mf), data = size_survival, family = binomial(link = "logit"), control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))) #this one results in a warnign 
summary(lmesurv1)
anova(lmesurv1)
glance(lmesurv1)
emtrends(lmesurv1, pairwise ~ Site, var = "elev_m")
emmip(lmesurv1, Site ~ elev_m, cov.reduce = range) #low elev pops have a steeper increase in survival from low to high elev garden than higher elev pops 
#ranova(lmesurv) doesn't work for lme4
#to test for significance of random effect: 
#the most common way to do this is to use a likelihood ratio test, i.e. fit the full and reduced models (the reduced model is the model with the focal variance(s) set to zero). 
m0 <- glmer(Survival ~ elev_m + (1|parent.pop), data = size_survival, family = binomial("logit")) 
m00 <- glm(Survival~ elev_m, size_survival, family = binomial("logit"))
anova(lmesurv1,m0, m00) #model with both random effects has a higher likelihood (better fit)
```


UCD only
```{r}
size_survival_ucd <- size_survival %>% filter(Site=="UCD")

##Recent models 
survmod_UCD <- glm(Survival ~ Recent_Gowers_Dist_UCD, data = size_survival_ucd, family = "binomial")
summary(survmod_UCD)
survmod_UCD2 <- glmer(Survival ~ elev_m + (1|parent.pop/mf), data = size_survival_ucd, family = binomial(link = "logit"))
summary(survmod_UCD2)

heightmod_UCD <- glm(height.cm ~ Recent_Gowers_Dist_UCD, data=size_survival_ucd)
summary(heightmod_UCD)

heightmod_UCD_mixed <- lmer(height.cm ~ Recent_Gowers_Dist_UCD + (1|parent.pop/mf), data=size_survival_ucd)
summary(heightmod_UCD_mixed)
ranova(heightmod_UCD_mixed)

lengthmod_UCD <- glm(long.leaf.cm ~ Recent_Gowers_Dist_UCD, data=size_survival_ucd)
summary(lengthmod_UCD)

lengthmod_UCD_mixed <- lmer(long.leaf.cm ~ Recent_Gowers_Dist_UCD + (1|parent.pop/mf), data=size_survival_ucd)
summary(lengthmod_UCD_mixed)
ranova(lengthmod_UCD_mixed)

##Historic models
survmod_UCD <- glm(Survival ~ Historic_Gowers_Dist_UCD, data = size_survival_ucd, family = "binomial")
summary(survmod_UCD)

heightmod_UCD <- glm(height.cm ~ Historic_Gowers_Dist_UCD, data=size_survival_ucd)
summary(heightmod_UCD)

heightmod_UCD_mixed <- lmer(height.cm ~ Historic_Gowers_Dist_UCD + (1|parent.pop/mf), data=size_survival_ucd)
summary(heightmod_UCD_mixed)
ranova(heightmod_UCD_mixed)

lengthmod_UCD <- glm(long.leaf.cm ~ Historic_Gowers_Dist_UCD, data=size_survival_ucd)
summary(lengthmod_UCD)

lengthmod_UCD_mixed <- lmer(long.leaf.cm ~ Historic_Gowers_Dist_UCD + (1|parent.pop/mf), data=size_survival_ucd)
summary(lengthmod_UCD_mixed)
ranova(lengthmod_UCD_mixed)
```

WL2 Only 
```{r}
size_survival_WL2 <- size_survival %>% filter(Site=="WL2")

##Recent models 
survmod_WL2 <- glm(Survival ~ Recent_Gowers_Dist_WL2, data = size_survival_WL2, family = "binomial")
summary(survmod_WL2)
survmod_WL22 <- glmer(Survival ~ elev_m + (1|parent.pop/mf), data = size_survival_WL2, family = binomial(link = "logit"))
summary(survmod_WL22)

heightmod_WL2 <- glm(height.cm ~ Recent_Gowers_Dist_WL2, data=size_survival_WL2)
summary(heightmod_WL2)

heightmod_WL2_mixed <- lmer(height.cm ~ Recent_Gowers_Dist_WL2 + (1|parent.pop/mf), data=size_survival_WL2)
summary(heightmod_WL2_mixed)
ranova(heightmod_WL2_mixed)

lengthmod_WL2 <- glm(long.leaf.cm ~ Recent_Gowers_Dist_WL2, data=size_survival_WL2)
summary(lengthmod_WL2)

lengthmod_WL2_mixed <- lmer(long.leaf.cm ~ Recent_Gowers_Dist_WL2 + (1|parent.pop/mf), data=size_survival_WL2)
summary(lengthmod_WL2_mixed)
ranova(lengthmod_WL2_mixed)

##Historic models
survmod_WL2 <- glm(Survival ~ Historic_Gowers_Dist_WL2, data = size_survival_WL2, family = "binomial")
summary(survmod_WL2)

heightmod_WL2 <- glm(height.cm ~ Historic_Gowers_Dist_WL2, data=size_survival_WL2)
summary(heightmod_WL2)

heightmod_WL2_mixed <- lmer(height.cm ~ Historic_Gowers_Dist_WL2 + (1|parent.pop/mf), data=size_survival_WL2)
summary(heightmod_WL2_mixed)
ranova(heightmod_WL2_mixed)

lengthmod_WL2 <- glm(long.leaf.cm ~ Historic_Gowers_Dist_WL2, data=size_survival_WL2)
summary(lengthmod_WL2)

lengthmod_WL2_mixed <- lmer(long.leaf.cm ~ Historic_Gowers_Dist_WL2 + (1|parent.pop/mf), data=size_survival_WL2)
summary(lengthmod_WL2_mixed)
ranova(lengthmod_WL2_mixed)
```


```{r}
survmod <- glm(Survival ~ Site + elev_m, data = size_survival, family = "binomial")
summary(survmod)
#go back and find how I incorporated random effects into log reg in R 
```

### Log Reg survival ~ size 
Height
```{r}
basiclogit_height <- glm(Survival ~ height.cm, data = size_survival, family = "binomial")
basiclogit_height
summary(basiclogit_height)

logit_height <- glm(Survival ~ height.cm + Site, data = size_survival, family = "binomial")
summary(logit_height)

logit_height2 <- glmer(Survival ~ height.cm*Site + (1|parent.pop/mf), data = size_survival, family = binomial(link = "logit"), nAGQ=0) 
summary(logit_height2)
emtrends(logit_height2, pairwise ~ Site, var = "height.cm")
emmip(logit_height2, Site ~ height.cm, cov.reduce = range)
```

Length
```{r}
logit_length2 <- glmer(Survival ~ long.leaf.cm*Site + (1|parent.pop/mf), data = size_survival, family = binomial(link = "logit"), nAGQ=0) 
summary(logit_length2)
emtrends(logit_length2, pairwise ~ Site, var = "long.leaf.cm")
emmip(logit_length2, Site ~ long.leaf.cm, cov.reduce = range)
```


### Figures 
```{r}
size_survival %>% 
  ggplot(aes(x=height.cm, y=Survival)) +
  geom_point()
```

```{r}
size_surv_pop_avgs <- size_survival %>% 
  group_by(parent.pop, Site, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(Survival)), mean_Surv = mean(Survival,na.rm=(TRUE)), 
            sem_surv=sem(Survival, na.rm=(TRUE)), 
            N_height = sum(!is.na(height.cm)), mean_height.cm = mean(height.cm,na.rm=(TRUE)), 
            sem_height.cm=sem(height.cm, na.rm=(TRUE)), N_length = sum(!is.na(long.leaf.cm)),
            mean_long.leaf.cm=mean(long.leaf.cm, na.rm=(TRUE)), 
            sem_long.leaf.cm=sem(long.leaf.cm, na.rm=TRUE))

size_surv_pop_avgs %>% arrange(desc(mean_height.cm))

size_surv_pop_avgs2 <- size_surv_pop_avgs %>% 
  mutate(Site=str_replace_all(Site, "UCD", "Low Elev"), 
         Site=str_replace_all(Site, "WL2", "High Elev"))
size_surv_pop_avgs2$Site <- factor(size_surv_pop_avgs2$Site,
                                               levels = c('Low Elev','High Elev'))
```

```{r}
size_surv_pop_avgs2 %>% 
  ggplot(aes(x=mean_height.cm, y=mean_Surv, group=parent.pop, color=elev_m)) +
  geom_point(size=8) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Height (cm)" ,y="Survival", color="Elevation (m)") +
  theme(text=element_text(size=28)) +
  facet_wrap(vars(Site), scales="free")
ggsave("../output/Surv_Height_BothGardens.png", width = 18, height = 8, units = "in")

size_surv_pop_avgs2 %>% 
  ggplot(aes(x=mean_long.leaf.cm, y=mean_Surv, group=parent.pop, color=elev_m)) +
  geom_point(size=8) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Leaf Length (cm)" ,y="Survival", color="Elevation (m)") +
  theme(text=element_text(size=28)) +
  facet_wrap(vars(Site), scales="free")
ggsave("../output/Surv_Length_BothGardens.png", width = 18, height = 8, units = "in")

size_surv_pop_avgs %>% 
  filter(Site=="UCD") %>% 
  ggplot(aes(x=mean_height.cm, y=mean_Surv, group=parent.pop, color=elev_m)) +
  geom_point(size=1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Height (cm)" ,y="Survival", color="Elevation (m)", title="Low Elevation Garden") +
  theme(text=element_text(size=25))

size_surv_pop_avgs %>% 
  filter(Site=="UCD") %>% 
  ggplot(aes(x=mean_height.cm, y=mean_Surv)) +
  geom_point(size=1) +
  geom_smooth() +
  theme_classic()  +
  labs(x="Height (cm)" ,y="Survival", color="Elevation (m)", title="Low Elevation Garden") +
  theme(text=element_text(size=25))

size_surv_pop_avgs %>% 
  filter(Site=="WL2") %>% 
  ggplot(aes(x=mean_height.cm, y=mean_Surv, group=parent.pop, color=elev_m)) +
  geom_point(size=1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Height (cm)" ,y="Survival", color="Elevation (m)", title="Low Elevation Garden") +
  theme(text=element_text(size=25))

size_surv_pop_avgs %>% 
  filter(Site=="WL2") %>% 
  ggplot(aes(x=mean_height.cm, y=mean_Surv)) +
  geom_point(size=1) +
  geom_smooth() +
  theme_classic()  +
  labs(x="Height (cm)" ,y="Survival", color="Elevation (m)", title="Low Elevation Garden") +
  theme(text=element_text(size=25))
```




