---
title: "Flint_Growth_Season"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  markdown: 
    wrap: 72
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Growth Season by snow/rain

-   First month with no snow pack or w/ significant rain as start of
    growth season

    -   This month isn't included in the final "growth season" - because
        we're saying how close are the conditions upon establishment

-   First month with snow pack or w/ significant CWD as end of growth
    season
    
    -   This month is included in the final "growth season" - because
        that month is likely a month when they're maturing seeds (esp at lower elevs)

Mean height at WL2 in Oct was 7 cm --\> 70 mm

## Notes: 
 - Need to come back and do this to include 2024 months for biennials 
 - Calculate grow season for gardens in the specific year of the garden (2022-2023)

## Relevant Libraries and Functions

```{r}
library(tidyverse)
library(ggrepel)
library(cowplot)
library(gridExtra)
library(corrplot) #plotting correlations 
library(rstatix) #performing cor_test
library(naniar) #replaces values with NA
sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
} #legend function for grid_arrange

true_round <- function(number, digits) {
  posneg <- sign(number)
  number <- abs(number) * 10^digits
  number <- number + 0.5 + sqrt(.Machine$double.eps)
  number <- trunc(number)
  number <- number / 10 ^ digits
  number * posneg
} #rounding function that bypasses R's round to even default

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")
```

## Load the data (from "Climate_Prep.Rmd")

```{r}
pop_elev_climate <- read_csv("../output/Climate/flint_climate_UCDpops.csv")
```

## Calculation of recent (last 30 years) and historical climate (prior 30 years)

For 2023 (annuals-only analysis)

```{r}
unique(pop_elev_climate$parent.pop) #double check all 23 populations and garden sites are in the data set 

pop_elev_climate_recent <- pop_elev_climate %>% 
  filter(parent.pop != "UCD_Garden", parent.pop != "WL2_Garden") %>% #remove garden sites since we don't need their historic growth seasons 
  filter(year>1993 & year<2024) %>% 
  select(parent.pop:month, cwd, pck, ppt, tmn, tmx) %>% 
  mutate(month = recode(month, #convert months to numbers 
  jan = 1,
  feb = 2,
  mar = 3,
  apr = 4,
  may = 5,
  jun = 6,
  jul = 7,
  aug = 8,
  sep = 9,
  oct = 10,
  nov = 11,
  dec = 12
))
head(pop_elev_climate_recent)
xtabs(~parent.pop+month, data=pop_elev_climate_recent) #check that there's 30 years for all pops and months 
unique(pop_elev_climate_recent$month)

pop_elev_climate_historical <- pop_elev_climate %>% 
  filter(parent.pop != "UCD_Garden", parent.pop != "WL2_Garden") %>% #remove garden sites since we don't need their historic growth seasons 
  filter(year<=1993 & year>1963) %>% 
  select(parent.pop:month, cwd, pck, ppt, tmn, tmx) %>% 
  mutate(month = recode(month, #convert months to numbers 
  jan = 1,
  feb = 2,
  mar = 3,
  apr = 4,
  may = 5,
  jun = 6,
  jul = 7,
  aug = 8,
  sep = 9,
  oct = 10,
  nov = 11,
  dec = 12
))
head(pop_elev_climate_historical, 13)
summary(pop_elev_climate_historical)
xtabs(~parent.pop+month, data=pop_elev_climate_historical) #check that there's 30 years for all pops and months 
```

## Averages and Preliminary Calcs

### Recent climate

```{r}
pop_elev_climate_recent %>% filter(ppt == 0) #months 7-10 usually have ppt = 0 for low elev 
pop_elev_climate_recent %>% filter(pck >0, ppt == 0) #in sequoia pops, 2022, ppt=0, but there was still snowpack in January 
pop_elev_climate_recent %>% filter(ppt > 0, tmx<1.5) %>% arrange(pck) #if the temp is less than 1.5 (Flint's criteria for snowfall) and there is precipitation then pck > 0

#first calculate 30 year monthly averages for pck and ppt 
pop_elev_climate_recent_avgs <- pop_elev_climate_recent %>% 
  group_by(parent.pop, elevation.group, elev_m, month) %>% 
  summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE)
names(pop_elev_climate_recent_avgs) <- gsub("fn2", "sem", colnames(pop_elev_climate_recent_avgs))
names(pop_elev_climate_recent_avgs) <-gsub("fn1", "mean", colnames(pop_elev_climate_recent_avgs))
pop_elev_climate_recent_avgs #30 year averages per month for each pop

pop_elev_climate_recent_avgs <- pop_elev_climate_recent_avgs %>% mutate(PckSum=sum(pck_mean)) #estimate of average total snowpack in a year 
pop_elev_climate_recent_avgs %>% arrange(PckSum)

#Exploratory filters
pop_elev_climate_recent %>% filter(parent.pop=="SC") %>% filter(year==2016 | year==2017) #snow pack in Jan 2017
pop_elev_climate_recent %>% filter(parent.pop=="IH") %>% filter(pck >0) #snow pack in Jan 2017 and Feb 2019
pop_elev_climate_recent %>% filter(parent.pop=="SQ3") %>% filter(pck==0, ppt < 10) #high elev pops get very little rain and high cwd during growth season ...
```

### Historical climate

```{r}
pop_elev_climate_historical_avgs <- pop_elev_climate_historical %>% 
  group_by(parent.pop, elevation.group, elev_m, month) %>% 
  summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE)
names(pop_elev_climate_historical_avgs) <- gsub("fn2", "sem", colnames(pop_elev_climate_historical_avgs))
names(pop_elev_climate_historical_avgs) <-gsub("fn1", "mean", colnames(pop_elev_climate_historical_avgs))
pop_elev_climate_historical_avgs #30 year averages per month for each pop

pop_elev_climate_historical_avgs <- pop_elev_climate_historical_avgs %>%
  mutate(PckSum=sum(pck_mean)) #estimate of average total snowpack in a year 
pop_elev_climate_historical_avgs %>% arrange(PckSum)

#Exploratory filters
pop_elev_climate_historical %>% filter(parent.pop=="BH") %>% filter(pck >0) #snow pack in Jan 1982
pop_elev_climate_historical %>% filter(parent.pop=="SC") %>% filter(pck >0) #snow pack in Jan 1973 and 1982
pop_elev_climate_historical %>% filter(parent.pop=="CC") %>% filter(pck >0) #snow pack in Jan 1969 and 1973
pop_elev_climate_historical %>% filter(parent.pop=="TM2") %>% filter(pck >0) #snow pack in Dec '68, Jan-Feb '69, Dec '71, Dec '72, Jan '73, Jan '82
pop_elev_climate_historical %>% filter(parent.pop=="IH") %>% filter(pck >0) #snow pack in Dec '68, Jan-Feb '69, Dec '70, Dec '71, Jan & Dec '72, Jan-Feb '73, Jan '79
```

### Simpler version (similar results to above calcs)
Take the sum of snowpack within each year and then average that 
In code above, take average snowpack for each month (across all years) and then sum that 
```{r}
##recent
pop_elev_climate_recent_sums <- pop_elev_climate_recent %>% 
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  summarise(PckSum=sum(pck))
pop_elev_climate_recent_sums_avgs <- pop_elev_climate_recent_sums %>% 
  group_by(parent.pop, elevation.group, elev_m) %>% 
  summarise(meanPckSum=mean(PckSum), semPckSum=sem(PckSum)) %>% 
  arrange(meanPckSum)
pop_elev_climate_recent_sums_avgs

##historical
pop_elev_climate_historical_sums <- pop_elev_climate_historical %>% 
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  summarise(PckSum=sum(pck))
pop_elev_climate_historical_sums_avgs <- pop_elev_climate_historical_sums %>% 
  group_by(parent.pop, elevation.group, elev_m) %>% 
  summarise(meanPckSum=mean(PckSum), semPckSum=sem(PckSum)) %>% 
  arrange(meanPckSum)
pop_elev_climate_historical_sums_avgs
```

## Populations that get less than 70 mm (b/c that's the avg height in Oct at WL2 garden) of snow pack in a year (on average)

-   First month = ppt \>= 25

    -   Remember this month is not included in the final "growth season"

    -   Jenny has used 25 mm as germinating inducing rain (esp b/c it's
        a sum of ppt for the whole month), 10 mm would probably be fine
        for sustaining growth

-   Last month = cwd \> 3rd quartile of CWD for those groups of pops

    -   Remember this month is included in the final "growth season"

### Recent Climate

Prep
```{r}
nosnow_pops_recent <- pop_elev_climate_recent_avgs %>% filter(PckSum < 70)
unique(nosnow_pops_recent$parent.pop) #BH, CC, TM2, SC, IH 
summary(nosnow_pops_recent) #3rd quartile of CWD = 89

nosnow_pops_recent_tojoin <- nosnow_pops_recent %>% select(parent.pop:elev_m, PckSum) %>% distinct()

nosnow_pops_recent_years <- left_join(nosnow_pops_recent_tojoin, pop_elev_climate_recent) 

nosnow_pops_recent_years %>% filter(month=="7", ppt>=25) #never get >=25mm of ppt in July 
nosnow_pops_recent_years %>% filter(month=="8", ppt>=25) #sometimes get >= 25 mm of ppt in Aug
#going to start the "grow year" in Sept though due to after ripening requirements 

growyear_months <- tibble(month=c(1:12), growmonth=c(5, 6, 7, 8, 9, 10, 11, 12, 1, 2, 3, 4))

nosnow_recent_growyear_months <- full_join(growyear_months, nosnow_pops_recent_years)
dim(nosnow_recent_growyear_months)

#nosnow_recent_growyear_months %>% filter(ppt==0, cwd<89)
```

First month 
```{r}
nosnow_recent_first_month <- nosnow_recent_growyear_months %>%
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  filter(ppt>=25) %>% 
  arrange(growmonth) %>% 
  filter(row_number()==1) #get first month for each pop and year with germinating inducing rain 

#nosnow_recent_first_month %>% filter(parent.pop=="IH") %>% arrange(year)

nosnow_recent_first_month_tomerge <- nosnow_recent_first_month %>% 
  select(parent.pop:elev_m, year, firstmonth=growmonth) #first month is in grow month not calendar month format

nosnow_recent_first_month_col <- full_join(nosnow_recent_growyear_months, nosnow_recent_first_month_tomerge)
dim(nosnow_recent_first_month_col)
```

Last month
```{r}
nosnow_recent_last_month <- nosnow_recent_growyear_months %>%
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  filter(cwd>89) %>% #last month is when cwd gets too high 
  arrange(month) %>% 
  filter(row_number()==1)

nosnow_recent_last_month_tomerge <- nosnow_recent_last_month %>% 
  select(parent.pop:elev_m, year, lastmonth=growmonth) #last month is in grow month not calendar month format

nosnow_recent_last_month_col <- full_join(nosnow_recent_first_month_col, nosnow_recent_last_month_tomerge) %>% 
  select(firstmonth, lastmonth, month:tmx)
dim(nosnow_recent_last_month_col)
```

Checking for weird cases 
```{r}
summary(nosnow_recent_last_month_col)

nosnow_recent_last_month_col %>% filter(is.na(lastmonth)) %>% arrange(year) #3 years (1998, 2005, 2019) when IH doesn't have a last month based on these criteria (cwd never goes above 89 in those years)
#****CAN EXCLUDE THESE CASES
nosnow_recent_last_month_col %>% filter(is.na(firstmonth)) %>% arrange(year) #1 year (2013) when BH doesn't have a first month based on these criteria (ppt never above 25)
#****CAN EXCLUDE THIS CASE 
nosnow_recent_last_month_col %>% filter(lastmonth<4) %>% arrange(parent.pop, year) #some cases where last month is less than 4 (earlier than December)
nosnow_recent_last_month_col %>% filter(lastmonth<firstmonth, lastmonth<4) %>% arrange(parent.pop, year) #all of the above are when the last month is before the first month (in growyear, not calendar year)
#****NEED TO CONVERT LAST MONTH TO CALENDAR YEAR FORMAT FOR AVG CALCS 
nosnow_recent_last_month_col %>% filter(lastmonth>firstmonth & lastmonth<4) %>% arrange(parent.pop, year) #0 cases where last month is after the first month 

nosnow_recent_last_month_col %>% filter(lastmonth==1) %>% arrange(parent.pop, year) #10 cases where Sept is last month 
nosnow_recent_last_month_col %>% filter(lastmonth==2) #0 cases where Oct-Dec is the last month
nosnow_recent_last_month_col %>% filter(firstmonth>4) #0 cases where first month is after December
nosnow_recent_last_month_col %>% filter(lastmonth==firstmonth) %>% arrange(parent.pop, year) #first month and last month are never the same 
nosnow_recent_last_month_col %>% filter(growmonth==firstmonth+1, cwd>89) %>% arrange(parent.pop, year) #1 case at UCD garden where cwd is high in the second growth month (2013) - see above note

#nosnow_recent_last_month_col %>% filter(growmonth==lastmonth) %>% arrange(cwd, ppt)
```

Average first and last month
```{r}
#remove IH: 1998, 2005, 2019
#remove BH: 2013
#convert last month to cal year for avg calc 

nosnow_recentfirst_last_month_for_calc <- nosnow_recent_last_month_col %>% 
  filter(!(parent.pop=="IH" & year==1998)) %>% 
  filter(!(parent.pop=="IH" & year==2005)) %>% 
  filter(!(parent.pop=="IH" & year==2019)) %>% 
  filter(!(parent.pop=="BH" & year==2013)) %>% 
  mutate(lastmonth_cal=if_else(lastmonth<=4, lastmonth+8, lastmonth-4)) %>% 
  select(firstmonth:lastmonth, lastmonth_cal, parent.pop, elevation.group, elev_m, year) %>% 
  distinct()

nosnow_recent_avg_first_last_month <- nosnow_recentfirst_last_month_for_calc %>% 
  group_by(parent.pop, elevation.group, elev_m) %>% 
  summarise(AvgFirstMonth=mean(firstmonth), First_sem=sem(firstmonth),
            AvgLastMonth=mean(lastmonth), Last_sem=sem(lastmonth),
            AvgLastMonth_cal=mean(lastmonth_cal), Last_cal_sem=sem(lastmonth_cal))
nosnow_recent_avg_first_last_month
#2 in growmonth = October 
#8 in growmonth = April - w/o converting last month to calendar year format, the cases where growmonth=1 is pulling the last month much earlier and with a much greater standard error --> wrong!
#8 in lastmonth_cal = August 

#Used Standard Rounding
nosnow_avg_first_last_month_recent <- nosnow_recent_avg_first_last_month %>% 
  mutate(AvgFirstMonth=true_round(AvgFirstMonth, 0), 
          AvgLastMonth=true_round(AvgLastMonth_cal, 0) + 4) %>% #avg in calendar year first to account for cases where CC and TM2 had a last month of Sept, then convert back to grow month format
  select(parent.pop:elev_m, AvgFirstMonth, AvgLastMonth)
nosnow_recent_avg_first_last_month
nosnow_avg_first_last_month_recent

nosnow_recent_avg_first_last_month_allmonths <- left_join(nosnow_recent_growyear_months, nosnow_avg_first_last_month_recent)
```
Apriori Assumption for growth season: Low: Oct-July

Fill in all the months b/t the first and last for the full growth season 
```{r}
nosnow_grwseason_recent <- nosnow_recent_avg_first_last_month_allmonths %>% 
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  filter(growmonth>AvgFirstMonth) %>% #first and last month are in grow month format not calendar year 
  filter(growmonth<=AvgLastMonth) %>% 
  arrange(year,parent.pop, growmonth)
summary(nosnow_grwseason_recent) 

xtabs(~parent.pop+month, data=nosnow_grwseason_recent)

nosnow_grwseason_recent %>% ggplot(aes(x=month)) + geom_histogram() +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5,6,7,8,9,10, 11, 12)) +
  facet_wrap(~parent.pop)
```


### Historical Climate

Prep
```{r}
nosnow_pops_historical <- pop_elev_climate_historical_avgs %>% filter(PckSum < 70)
unique(nosnow_pops_historical$parent.pop)
summary(nosnow_pops_historical) #3rd quartile of CWD = 83 - was 87 before we shifted historical to be 1964-1993 (from 1962-1991)

nosnow_pops_historical_tojoin <- nosnow_pops_historical %>% select(parent.pop:elev_m, PckSum) %>% distinct()

nosnow_pops_historical_years <- left_join(nosnow_pops_historical_tojoin, pop_elev_climate_historical) 

nosnow_pops_historical_years %>% filter(month=="7", ppt>=25) #in 1974, ppt>25 for 4 pops in July + BH in 1992... unlikely seeds would have after ripened enough 
nosnow_pops_historical_years %>% filter(month=="8", ppt>=25) %>% arrange(year, parent.pop) #11 cases of ppt>25 in Aug 
nosnow_pops_historical_years %>% filter(year==1965) %>% arrange(parent.pop)
nosnow_pops_historical_years %>% filter(year==1968) %>% arrange(parent.pop) 
nosnow_pops_historical_years %>% filter(year==1975) %>% arrange(parent.pop) 
nosnow_pops_historical_years %>% filter(year==1976) %>% arrange(parent.pop)
#some cases where ppt remains high in following month 

#going to start in Sept though due to afterripening requirements 
growyear_months <- tibble(month=c(1:12), growmonth=c(5, 6, 7, 8, 9, 10, 11, 12, 1, 2, 3, 4))

nosnow_historical_growyear_months <- full_join(growyear_months, nosnow_pops_historical_years)
dim(nosnow_historical_growyear_months)
```

First month 
```{r}
nosnow_historical_first_month <- nosnow_historical_growyear_months %>%
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  filter(ppt>=25) %>% 
  arrange(growmonth) %>% 
  filter(row_number()==1) #get first month for each pop and year with germinating inducing rain 

nosnow_historical_first_month_tomerge <- nosnow_historical_first_month %>% 
  select(parent.pop:elev_m, year, firstmonth=growmonth)

nosnow_historical_first_month_col <- full_join(nosnow_historical_growyear_months, nosnow_historical_first_month_tomerge)
dim(nosnow_historical_first_month_col)
```

Last month 
```{r}
nosnow_historical_last_month <- nosnow_historical_growyear_months %>%
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  filter(cwd>83) %>% #adjusted for the historical 3rd quartile of cwd
  arrange(month) %>% 
  filter(row_number()==1)

nosnow_historical_last_month_tomerge <- nosnow_historical_last_month %>% 
  select(parent.pop:elev_m, year, lastmonth=growmonth) #last month is in grow month not calendar month format

nosnow_historical_last_month_col <- full_join(nosnow_historical_first_month_col, nosnow_historical_last_month_tomerge)  %>% 
  select(firstmonth, lastmonth, month:tmx)
dim(nosnow_historical_last_month_col)
```

Checking for weird cases
```{r}
summary(nosnow_historical_last_month_col)

nosnow_historical_last_month_col %>% filter(is.na(lastmonth)) %>% arrange(year) #0 years where there isn't a last month
nosnow_historical_last_month_col %>% filter(is.na(firstmonth)) %>% arrange(year) #0 years where there isn't a first month
nosnow_historical_last_month_col %>% filter(lastmonth<4) %>% arrange(parent.pop, year) #some cases where last month is less than 4 (earlier than December)
nosnow_historical_last_month_col %>% filter(lastmonth<firstmonth, lastmonth<4) %>% arrange(parent.pop, year) #most of the above are when the last month is before the first month (in growyear, not calendar year)
#****NEED TO CONVERT LAST MONTH TO CALENDAR YEAR FORMAT FOR AVG CALCS 
nosnow_historical_last_month_col %>% filter(lastmonth>firstmonth & lastmonth<4) %>% arrange(parent.pop, year) #0 cases where the last month is after the first month before Dec
nosnow_historical_last_month_col %>% filter(lastmonth==firstmonth) %>% arrange(parent.pop, year) #first month and last month are the same for IH in 1 year ('89)
#****CAN DROP THIS CASE 

nosnow_historical_last_month_col %>% filter(lastmonth==1) %>% arrange(parent.pop, year) #6 cases where Sept is last month 
nosnow_historical_last_month_col %>% filter(lastmonth==2) #0 cases where Oct-Dec is the last month
nosnow_historical_last_month_col %>% filter(firstmonth>4) #0 cases where first month is after December

nosnow_historical_last_month_col %>% filter(growmonth==firstmonth+1, cwd>83) %>% arrange(parent.pop, year) #3 cases at BH where cwd is high in the second growth month 

#nosnow_historical_last_month_col %>% filter(growmonth==lastmonth) %>% arrange(cwd, ppt)
```

Average first and last month
```{r}
#remove IH: 1989
#convert last month to cal year for avg calc 

nosnow_historicalfirst_last_month_for_calc <- nosnow_historical_last_month_col %>% 
  filter(!(parent.pop=="IH" & year==1989)) %>% 
  mutate(lastmonth_cal=if_else(lastmonth<=4, lastmonth+8, lastmonth-4)) %>% 
  select(firstmonth:lastmonth, lastmonth_cal, parent.pop, elevation.group, elev_m, year) %>% 
  distinct()

nosnow_historical_avg_first_last_month <- nosnow_historicalfirst_last_month_for_calc %>% 
  group_by(parent.pop, elevation.group, elev_m) %>% 
  summarise(AvgFirstMonth=mean(firstmonth), First_sem=sem(firstmonth),
            AvgLastMonth=mean(lastmonth), Last_sem=sem(lastmonth),
            AvgLastMonth_cal=mean(lastmonth_cal), Last_cal_sem=sem(lastmonth_cal))
nosnow_historical_avg_first_last_month
#2 in growmonth = October 
#10 in growmonth = June - w/o converting last month to calendar year format, the cases where growmonth=1 is pulling the last month much earlier and with a much greater standard error --> wrong!
#8 in lastmonth_cal = August 

#Used Standard Rounding
nosnow_avg_first_last_month_historical <- nosnow_historical_avg_first_last_month %>% 
  mutate(AvgFirstMonth=true_round(AvgFirstMonth, 0), 
         AvgLastMonth=true_round(AvgLastMonth_cal, 0) + 4) %>% #avg in calendar year first to account for cases where CC and TM2 had a last month of Sept, then convert back to grow month format
  select(parent.pop:elev_m, AvgFirstMonth, AvgLastMonth)
nosnow_historical_avg_first_last_month
nosnow_avg_first_last_month_historical

nosnow_historical_avg_first_last_month_allmonths <- left_join(nosnow_historical_growyear_months, nosnow_avg_first_last_month_historical)
```

Fill in all the months b/t the first and last for the full growth season 
```{r}
nosnow_grwseason_historical <- nosnow_historical_avg_first_last_month_allmonths %>% 
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  filter(growmonth>AvgFirstMonth) %>% #first and last month are in grow month format not calendar year 
  filter(growmonth<=AvgLastMonth) %>% 
  arrange(year,parent.pop, growmonth)
summary(nosnow_grwseason_historical) 

xtabs(~parent.pop+month, data=nosnow_grwseason_historical)

nosnow_grwseason_historical %>% ggplot(aes(x=month)) + geom_histogram() +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5,6,7,8,9,10, 11, 12)) +
  facet_wrap(~parent.pop)
#note this includes a UCD_Garden growing season that uses the same season criteria as the other low elevation pops, but CWD is higher at UCD so the last month of the growing season is probably later than this indicates
```

## Populations that get more than 70 mm of snow pack in a year (on average)

-   First month = snowpack = 0 and min temp > 0

    -   Remember this month is not included in the final "growth season"

-   Last month = snowpack. \> 70 mm OR pck > 0 and min temp < 0 OR min temp < -5 (moderate freeze)

Note: “The month name of the snowpack file (that is, pckmar.asc) relates to the first day of the next month (that is, station observations on April 1st correlate with the snowpack file for March)." But this makes sense, a survey on April 1 gives a sense for the total snowpack in March. 

### Recent climate

Prep
```{r}
snow_pops_recent <- pop_elev_climate_recent_avgs %>% filter(PckSum >= 70)
unique(snow_pops_recent$parent.pop) #18 pops get some significant snowpack per year 
summary(snow_pops_recent)

snow_pops_recent_tojoin <- snow_pops_recent %>% select(parent.pop:elev_m, PckSum) %>% distinct()

snow_pops_recent_years <- left_join(snow_pops_recent_tojoin, pop_elev_climate_recent) 
summary(snow_pops_recent_years)
```

First month 
```{r}
snow_recent_first_month <- snow_pops_recent_years %>%
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  filter(pck==0) %>% 
  filter(tmn > 0) %>% 
  arrange(month) %>% 
  filter(row_number()==1) #get first month for each pop and year with no snowpack for germ

snow_recent_first_month_tomerge <- snow_recent_first_month %>% 
  select(parent.pop:elev_m, year, firstmonth=month)

snow_recent_first_month_col <- full_join(snow_pops_recent_years, snow_recent_first_month_tomerge)
dim(snow_recent_first_month_col)
```

Last month
```{r}
snow_recent_last_month <- snow_recent_first_month_col %>%
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  filter(month>firstmonth) %>% 
  filter(if_else(pck>70, pck>70, 
                 if_else(pck>0, tmn < 0, tmn <= -5))) %>%  #-5 is moderate freeze (Pardee et al. 2017)
  arrange(month) %>% 
  filter(row_number()==1) #get first month after growstart for each pop and year with pck >70 OR pck>0 & tmn <0 OR tmn <-5

#snow_recent_last_month %>% filter(tmn <=-5, pck <=0)
#snow_recent_last_month %>% filter(pck >0, tmn < 0)

snow_recent_last_month_tomerge <- snow_recent_last_month %>% 
  select(parent.pop:elev_m, year, firstmonth,lastmonth=month)

snow_recent_last_month_col <- full_join(snow_recent_first_month_col, snow_recent_last_month_tomerge) %>% 
  select(firstmonth, lastmonth, year:tmx)
dim(snow_recent_last_month_col)
```

Check weird cases
```{r}
summary(snow_recent_last_month_col)

snow_recent_last_month_col %>% filter(is.na(firstmonth)) #no cases where there isn't a firstmonth
snow_recent_last_month_col %>% filter(is.na(lastmonth)) #117 cases where there isn't a lastmonth 
#***WHAT TO DO FOR THESE CASES? Make the last month 12 (Dec)
#*no snow_pack, or min temp > -5 

snow_recent_last_month_col %>% filter(lastmonth==firstmonth) %>% arrange(parent.pop, year) #no cases where last month is first month 

snow_recent_last_month_col %>% filter(lastmonth==firstmonth+1) %>% arrange(parent.pop, year) 
#***I added a temperature requirement for temp >0 for the first month and still get 12 cases where there is heavy snow in the month after the first month 
snow_recent_last_month_col %>% filter(lastmonth==firstmonth+2) #1 case where there was sig snowpack in the second month after the first month of 0 snowpack (WL1 - 2018)

heavysnow_recent <- snow_recent_last_month_col %>% filter(lastmonth==firstmonth+1) %>% filter(month==firstmonth)
summary(heavysnow_recent)
unique(heavysnow_recent$parent.pop)
heavysnow_recent %>% filter(elev_m>1800)
#"DPR" "WL1" "WL2" "WR", "YO4" 
#*Temps range from just above freezing (0.080) to 2.9 (which seems like an acceptable temp given we vernalize at 4C)
#*OR START THE CHECK IN FEB FOR POPS > 6000 FT (ONLY 2 ABOVE 6000 FEET, both in 2014)
#*OR just take these cases out
```

Average first and last month
```{r}
#for cases with no last month - make last month 12
#remove cases where last month is first month +1 or 2 

snow_recentfirst_last_month_for_calc <- snow_recent_last_month_col %>% 
  mutate(lastmonth=if_else(is.na(lastmonth), 12, lastmonth)) %>% 
  filter(lastmonth > firstmonth+2) %>% 
  select(firstmonth:lastmonth, parent.pop, elevation.group, elev_m, year) %>% 
  distinct()
#snow_recentfirst_last_month_for_calc %>% filter(parent.pop=="CP2", year=="2011") #converting last month to 12 worked for the cases where it was NA before 

snow_recent_avg_first_last_month <- snow_recentfirst_last_month_for_calc %>% 
  group_by(parent.pop, elevation.group, elev_m) %>% 
  summarise(AvgFirstMonth=mean(firstmonth), First_sem=sem(firstmonth),
            AvgLastMonth=mean(lastmonth), Last_sem=sem(lastmonth)) %>% 
  arrange(elevation.group, elev_m)
snow_recent_avg_first_last_month
#Start = May-June for "high" elev, Feb-May for "mid" elevation 
#End = Nov-Dec

#Used Standard Rounding
snow_avg_first_last_month_recent <- snow_recent_avg_first_last_month %>% 
  mutate(AvgFirstMonth=true_round(AvgFirstMonth, 0), AvgLastMonth=true_round(AvgLastMonth, 0)) %>% 
  select(parent.pop:elev_m, AvgFirstMonth, AvgLastMonth)
snow_recent_avg_first_last_month
snow_avg_first_last_month_recent

snow_recent_avg_first_last_month_allmonths <- left_join(snow_pops_recent_years, snow_avg_first_last_month_recent)
```

Fill in months b/t start and stop 
```{r}
snow_grwseason_recent <- snow_recent_avg_first_last_month_allmonths %>% 
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  filter(month>AvgFirstMonth) %>% 
  filter(month<=AvgLastMonth)
summary(snow_grwseason_recent) 

xtabs(~parent.pop+month, data=snow_grwseason_recent)

snow_grwseason_recent %>% ggplot(aes(x=month)) + geom_histogram() +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5,6,7,8,9,10, 11, 12)) +
  facet_wrap(~parent.pop)
```

### Historical climate

Prep
```{r}
snow_pops_historical <- pop_elev_climate_historical_avgs %>% filter(PckSum >= 70)
unique(snow_pops_historical$parent.pop) #18 pops get some significant snowpack per year 
summary(snow_pops_historical)

snow_pops_historical_tojoin <- snow_pops_historical %>% select(parent.pop:elev_m, PckSum) %>% distinct()

snow_pops_historical_years <- left_join(snow_pops_historical_tojoin, pop_elev_climate_historical) 
summary(snow_pops_historical_years)
snow_pops_historical_years %>% filter(pck < 2, pck >0) %>% arrange(parent.pop, pck) #What about when snowpack is 1 mm? This mostly occurs Nov-Jan, 
#one case in Feb (SQ3), one case in March (FR) and one case in July (LV1)
```

First month 
```{r}
snow_historical_first_month <- snow_pops_historical_years %>%
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  filter(pck==0) %>% 
  filter(tmn > 0) %>% 
  arrange(month) %>% 
  filter(row_number()==1) #get first month for each pop and year with no snowpack for germ

snow_historical_first_month_tomerge <- snow_historical_first_month %>% 
  select(parent.pop:elev_m, year, firstmonth=month)

snow_historical_first_month_col <- full_join(snow_pops_historical_years, snow_historical_first_month_tomerge)
dim(snow_historical_first_month_col)
```

Last month
```{r}
snow_historical_last_month <- snow_historical_first_month_col %>%
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  filter(month>firstmonth) %>% 
  filter(if_else(pck>70, pck>70, 
                 if_else(pck>0, tmn < 0, tmn <= -5))) %>%  #-5 is moderate freeze (Pardee et al. 2017) 
  arrange(month) %>% 
  filter(row_number()==1) #get first month after growstart for each pop and year with pck >70 OR PCK > 0 & TMN <0 OR TMN <-5

#snow_historical_last_month %>% filter(tmn <= -5, pck <=0)

snow_historical_last_month_tomerge <- snow_historical_last_month %>% 
  select(parent.pop:elev_m, year, firstmonth,lastmonth=month)

snow_historical_last_month_col <- full_join(snow_historical_first_month_col, snow_historical_last_month_tomerge) %>% 
  select(firstmonth, lastmonth, year:tmx)
dim(snow_historical_last_month_col)
```

check weird cases
```{r}
snow_historical_last_month_col %>% filter(is.na(firstmonth)) #no cases where there isn't a firstmonth
snow_historical_last_month_col %>% filter(is.na(lastmonth)) %>% arrange(elev_m) #99 cases where there isn't a lastmonth 
#****range of elevations 

snow_historical_last_month_col %>% filter(lastmonth==firstmonth) #no cases where last month is first month 

snow_historical_last_month_col %>% filter(lastmonth==firstmonth+1) #4 cases where there was sig snowpack in the month after the first month of 0 snowpack
snow_historical_last_month_col %>% filter(lastmonth==firstmonth+2) #7 cases where there was sig snowpack in the second month after the first month of 0 snowpack


heavysnow_historical <- snow_historical_last_month_col %>% filter(lastmonth==firstmonth+1) %>% filter(month==firstmonth)
summary(heavysnow_historical)
unique(heavysnow_historical$parent.pop)
heavysnow_historical %>% filter(elev_m>1800)
#"FR" "WL1" "WR"  
#****CHECK TEMPS FOR THESE CASES 
#*Temps range from 1.640 to 3.3 (which seems like an acceptable temp given we vernalize at 4C)
#*OR START THE CHECK IN FEB FOR POPS > 6000 FT (ALL BELOW 6000 FT)
#*OR just take these cases out

#snow_historical_last_month_col %>% filter(year==1991)
#snow_historical_last_month_col %>% filter(parent.pop=="DPR") %>% filter(year==1962|year==1975)
#snow_historical_last_month_col %>% filter(parent.pop=="WR") %>% filter(year==1975|year==1986)
#snow_historical_last_month_col %>% filter(parent.pop=="YO7") %>% filter(year==1976)
#snow_historical_last_month_col %>% filter(parent.pop=="FR") %>% filter(year==1975)
#snow_historical_last_month_col %>% filter(parent.pop=="SQ3") %>% filter(year==1976)
```

Average first and last month 
```{r}
#for cases with no last month - make last month 12
#remove cases where last month is first month +1 or 2 

snow_historicalfirst_last_month_for_calc <- snow_historical_last_month_col %>% 
  mutate(lastmonth=if_else(is.na(lastmonth), 12, lastmonth)) %>% 
  filter(lastmonth > firstmonth+2) %>% 
  select(firstmonth:lastmonth, parent.pop, elevation.group, elev_m, year) %>% 
  distinct()

snow_historical_avg_first_last_month <- snow_historicalfirst_last_month_for_calc %>% 
  group_by(parent.pop, elevation.group, elev_m) %>% 
  summarise(AvgFirstMonth=mean(firstmonth), First_sem=sem(firstmonth),
            AvgLastMonth=mean(lastmonth), Last_sem=sem(lastmonth)) %>% 
  arrange(elevation.group, elev_m)
snow_historical_avg_first_last_month
#Start = May-July for "high" elev, March-May for "mid" elevation 
#End = Nov-Dec
#interestingly, the starts are later than recent climate and ends are earlier in some cases 

#Used Standard Rounding
snow_avg_first_last_month_historical <- snow_historical_avg_first_last_month %>% 
  mutate(AvgFirstMonth=true_round(AvgFirstMonth, 0), AvgLastMonth=true_round(AvgLastMonth, 0)) %>% 
  select(parent.pop:elev_m, AvgFirstMonth, AvgLastMonth)
snow_historical_avg_first_last_month
snow_avg_first_last_month_historical

snow_historical_avg_first_last_month_allmonths <- left_join(snow_pops_historical_years, snow_avg_first_last_month_historical)
```

Fill in months b/t start and stop
```{r}
snow_grwseason_historical <- snow_historical_avg_first_last_month_allmonths %>%
  group_by(parent.pop, elevation.group, elev_m, year) %>% 
  filter(month>AvgFirstMonth) %>% 
  filter(month<=AvgLastMonth)
summary(snow_grwseason_historical) 

xtabs(~parent.pop+month, data=snow_grwseason_historical)

snow_grwseason_historical %>% ggplot(aes(x=month)) + geom_histogram() +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5,6,7,8,9,10, 11, 12)) +
  facet_wrap(~parent.pop)
```

## Bind snow and no_snow pops together

```{r}
unique(nosnow_grwseason_recent$parent.pop)
unique(snow_grwseason_recent$parent.pop)
allpops_recent_grwseason <- rbind(nosnow_grwseason_recent, snow_grwseason_recent)
summary(allpops_recent_grwseason)
unique(allpops_recent_grwseason$parent.pop)
write_csv(allpops_recent_grwseason, "../output/Climate/flint_climate_growthseason_recent.csv")

unique(nosnow_grwseason_historical$parent.pop)
unique(snow_grwseason_historical$parent.pop)
allpops_historical_grwseason <- rbind(nosnow_grwseason_historical, snow_grwseason_historical)
summary(allpops_historical_grwseason)
unique(allpops_historical_grwseason$parent.pop)
write_csv(allpops_historical_grwseason, "../output/Climate/flint_climate_growthseason_historical.csv")
```


## Climate traits across pops

### Totals

```{r}
names(allpops_recent_grwseason)
allpops_recent_grwseason_yearlytot <- allpops_recent_grwseason %>%  group_by(parent.pop, year, elev_m) %>% summarise_at(c("pck", "ppt"), sum, na.rm = TRUE)

recent_ppt_total <- allpops_recent_grwseason_yearlytot %>% ggplot(aes(x=year, y=ppt, group=parent.pop, color=elev_m)) + 
  geom_point() + geom_line() + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  ggtitle("Recent Climate")  + 
  theme_classic() + 
  ylim(0, 2000) +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45, hjust = 1))
#ggsave("../output/Climate/growthseasonTot_Precip_RecentClim.png", width = 12, height = 6, units = "in")

allpops_historical_grwseason_yearlytot <- allpops_historical_grwseason %>% group_by(parent.pop, year, elev_m) %>% summarise_at(c("pck", "ppt"), sum, na.rm = TRUE)

hist_ppt_total <- allpops_historical_grwseason_yearlytot %>% ggplot(aes(x=year, y=ppt, group=parent.pop, color=elev_m)) + 
  geom_point() + geom_line() + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  ggtitle("Historical Climate") + 
  theme_classic() + 
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45, hjust = 1))
#ggsave("../output/Climate/growthseasonTot_Precip_HistoricalClim.png", width = 12, height = 6, units = "in")

#should combine these into one figure and save that instead
legend <- get_legend(hist_ppt_total)
hist_ppt_total <- hist_ppt_total + theme(legend.position="none")
recent_ppt_total <- recent_ppt_total + theme(legend.position="none")
grid.arrange(hist_ppt_total, recent_ppt_total, legend, ncol=3, widths=c(3.12, 3.12, 1.09))
 #2000 x 850
```

### Monthly Averages

#### Recent Years - Growth Season
```{r}
allpops_recent_grwseason_mosavgs <- allpops_recent_grwseason %>% group_by(parent.pop, elevation.group, elev_m, month) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(allpops_recent_grwseason_mosavgs) <- gsub("fn2", "sem", colnames(allpops_recent_grwseason_mosavgs))
names(allpops_recent_grwseason_mosavgs) <-gsub("fn1", "mean", colnames(allpops_recent_grwseason_mosavgs))
allpops_recent_grwseason_mosavgs #30 year averages during growth season months 
allpops_recent_grwseason_mosavgs$elevation.group <- factor(allpops_recent_grwseason_mosavgs$elevation.group, levels=elev_order)       

allpops_recent_grwseason_mosavgs %>% ggplot(aes(x=month, y=cwd_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=cwd_mean-cwd_sem,ymax=cwd_mean+cwd_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0")+ 
  labs(fill="Elevation",x="Population", y="Avg CWD" ,title = "Average CWD during Growth Season - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1)) + 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/growthseasonAvg_Monthly_CWD_RecentClim.png", width = 16, height = 10, units = "in")

allpops_recent_grwseason_mosavgs %>% ggplot(aes(x=month, y=ppt_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=ppt_mean-ppt_sem,ymax=ppt_mean+ppt_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg Precip" ,title = "Average Precip during Growth Season - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))+ 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/growthseasonAvg_Monthly_Precip_RecentClim.png", width = 12, height = 6, units = "in")

allpops_recent_grwseason_mosavgs %>% ggplot(aes(x=month, y=tmn_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmn_mean-tmn_sem,ymax=tmn_mean+tmn_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MinTemp" ,title = "Average MinTemp during Growth Season - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))+ 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/growthseasonAvg_Monthly_MinTemp_RecentClim.png", width = 12, height = 6, units = "in")

allpops_recent_grwseason_mosavgs %>% ggplot(aes(x=month, y=tmx_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmx_mean-tmx_sem,ymax=tmx_mean+tmx_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MaxTemp" ,title = "Average MaxTemp during Growth Season - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))+ 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/growthseasonAvg_Monthly_MaxTemp_RecentClim.png", width = 12, height = 6, units = "in")
```

#### Historical Years - Growth Season
```{r}
allpops_historical_grwseason_mosavgs <- allpops_historical_grwseason %>% group_by(parent.pop, elevation.group, elev_m, month) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(allpops_historical_grwseason_mosavgs) <- gsub("fn2", "sem", colnames(allpops_historical_grwseason_mosavgs))
names(allpops_historical_grwseason_mosavgs) <-gsub("fn1", "mean", colnames(allpops_historical_grwseason_mosavgs))
allpops_historical_grwseason_mosavgs #30 year averages during growth season months 
allpops_historical_grwseason_mosavgs$elevation.group <- factor(allpops_historical_grwseason_mosavgs$elevation.group, levels=elev_order)       

allpops_historical_grwseason_mosavgs %>% ggplot(aes(x=month, y=cwd_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=cwd_mean-cwd_sem,ymax=cwd_mean+cwd_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(breaks=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0")+ 
  labs(fill="Elevation",x="Population", y="Avg CWD" ,title = "Average CWD during Growth Season - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1)) + 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/growthseasonAvg_Monthly_CWD_HistoricalClim.png", width = 16, height = 10, units = "in")

allpops_historical_grwseason_mosavgs %>% ggplot(aes(x=month, y=ppt_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=ppt_mean-ppt_sem,ymax=ppt_mean+ppt_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg Precip" ,title = "Average Precip during Growth Season - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))+ 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/growthseasonAvg_Monthly_Precip_HistoricalClim.png", width = 12, height = 6, units = "in")

allpops_historical_grwseason_mosavgs %>% ggplot(aes(x=month, y=tmn_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmn_mean-tmn_sem,ymax=tmn_mean+tmn_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MinTemp" ,title = "Average MinTemp during Growth Season - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))+ 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/growthseasonAvg_Monthly_MinTemp_HistoricalClim.png", width = 12, height = 6, units = "in")

allpops_historical_grwseason_mosavgs %>% ggplot(aes(x=month, y=tmx_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmx_mean-tmx_sem,ymax=tmx_mean+tmx_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MaxTemp" ,title = "Average MaxTemp during Growth Season - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))+ 
  facet_wrap(~parent.pop)
ggsave("../output/Climate/growthseasonAvg_Monthly_MaxTemp_HistoricalClim.png", width = 12, height = 6, units = "in")
```


### 30 Year Averages

#### Recent Years - Growth Season

```{r}
allpops_recent_grwseason_avgs <- allpops_recent_grwseason %>% group_by(parent.pop, elevation.group, elev_m) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(allpops_recent_grwseason_avgs) <- gsub("fn2", "sem", colnames(allpops_recent_grwseason_avgs))
names(allpops_recent_grwseason_avgs) <-gsub("fn1", "mean", colnames(allpops_recent_grwseason_avgs))
allpops_recent_grwseason_avgs #30 year averages during growth season months 
allpops_recent_grwseason_avgs$elevation.group <- factor(allpops_recent_grwseason_avgs$elevation.group, levels=elev_order)    

allpops_recent_grwseason_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, cwd_mean), y=cwd_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=cwd_mean-cwd_sem,ymax=cwd_mean+cwd_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0")+ 
  labs(fill="Elevation",x="Population", y="Avg CWD" ,title = "Average CWD during Growth Season - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseasonAvg_CWD_RecentClim.png", width = 12, height = 6, units = "in")

allpops_recent_grwseason_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_mean), y=ppt_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=ppt_mean-ppt_sem,ymax=ppt_mean+ppt_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg Precip" ,title = "Average Precip during Growth Season - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseasonAvg_Precip_RecentClim.png", width = 12, height = 6, units = "in")

allpops_recent_grwseason_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmn_mean), y=tmn_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmn_mean-tmn_sem,ymax=tmn_mean+tmn_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MinTemp" ,title = "Average MinTemp during Growth Season - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseasonAvg_MinTemp_RecentClim.png", width = 12, height = 6, units = "in")

allpops_recent_grwseason_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmx_mean), y=tmx_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmx_mean-tmx_sem,ymax=tmx_mean+tmx_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MaxTemp" ,title = "Average MaxTemp during Growth Season - Recent Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseasonAvg_MaxTemp_RecentClim.png", width = 12, height = 6, units = "in")
```

#### Historical Years - Growth Season

```{r}
allpops_historical_grwseason_avgs <- allpops_historical_grwseason %>% group_by(parent.pop, elevation.group, elev_m) %>% summarise_at(c("cwd", "pck", "ppt", "tmn", "tmx"), c(mean, sem), na.rm = TRUE) 
names(allpops_historical_grwseason_avgs) <- gsub("fn2", "sem", colnames(allpops_historical_grwseason_avgs))
names(allpops_historical_grwseason_avgs) <-gsub("fn1", "mean", colnames(allpops_historical_grwseason_avgs))
allpops_historical_grwseason_avgs #30 year averages during growth season months 
allpops_historical_grwseason_avgs$elevation.group <- factor(allpops_historical_grwseason_avgs$elevation.group, levels=elev_order)    

allpops_historical_grwseason_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, cwd_mean), y=cwd_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=cwd_mean-cwd_sem,ymax=cwd_mean+cwd_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0")+ 
  labs(fill="Elevation",x="Population", y="Avg CWD" ,title = "Average CWD during Growth Season - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseasonAvg_CWD_HistoricalClim.png", width = 12, height = 6, units = "in")

allpops_historical_grwseason_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt_mean), y=ppt_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=ppt_mean-ppt_sem,ymax=ppt_mean+ppt_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg Precip" ,title = "Average Precip during Growth Season - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseasonAvg_Precip_HistoricalClim.png", width = 12, height = 6, units = "in")

allpops_historical_grwseason_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmn_mean), y=tmn_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmn_mean-tmn_sem,ymax=tmn_mean+tmn_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MinTemp" ,title = "Average MinTemp during Growth Season - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseasonAvg_MinTemp_HistoricalClim.png", width = 12, height = 6, units = "in")

allpops_historical_grwseason_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmx_mean), y=tmx_mean, fill=elev_m)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=tmx_mean-tmx_sem,ymax=tmx_mean+tmx_sem),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") + 
  labs(fill="Elevation",x="Population", y="Avg MaxTemp" ,title = "Average MaxTemp during Growth Season - Historical Climate") +
   theme_classic() +
  theme(text=element_text(size=20), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseasonAvg_MaxTemp_HistoricalClim.png", width = 12, height = 6, units = "in")
```

## Climate trait correlations
```{r}
head(allpops_recent_grwseason)
allpops_recent_grwseason <- tibble(allpops_recent_grwseason)
allpops_recent_grwseason %>% cor_test(cwd, ppt, tmn, tmx, method = "pearson")
recent_cor_mat <- allpops_recent_grwseason %>% select(cwd, ppt, tmn, tmx) %>% cor_mat()
recent_cor_mat

recent_cor = allpops_recent_grwseason %>% select(cwd, ppt, tmn, tmx) %>% cor()
file_path= "../output/Climate/GrowthSeason_RecentClim_Cors.png"
png(width = 12, height = 6, res= 300, units = "in", file=file_path, type = "cairo")
corrplot(recent_cor)
dev.off()

allpops_historical_grwseason <- tibble(allpops_historical_grwseason)
allpops_historical_grwseason %>% cor_test(cwd, ppt, tmn, tmx, method = "pearson")
historical_cor_mat <- allpops_historical_grwseason %>% select(cwd, ppt, tmn, tmx) %>% cor_mat()
historical_cor_mat

historical_cor = allpops_historical_grwseason %>% select(cwd, ppt, tmn, tmx) %>% cor()
file_path= "../output/Climate/GrowthSeason_HistoricalClim_Cors.png"
png(width = 12, height = 6, res= 300, units = "in", file=file_path, type = "cairo")
corrplot(historical_cor)
dev.off()
```
Trait correlations are the same across recent and historical time
periods (during the growth season)