---
title: "PCAs_Combined"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Additional PCAs for paper 

## Libraries 
```{r}
library(tidyverse)
library(ggalt) #for geom_encircle
library(corrplot) #plotting correlations 
library(ggfortify) #easier PCA figures
library(QBMS) #for function calc_biovars to calculate bioclim variables
library(viridis) #for color scale 
```

## Combine Water Year and Growing Season PCAs

### Water Year Averages
```{r}
flint_all_year_avgs <- read_csv("../output/Climate/fullyear_FlintAvgs_wtr_year.csv")
bioclim_all_year_avgs <- read_csv("../output/Climate/fullyear_BioClimAvgs_wtr_year.csv")

#Merge
bioclim_flint_all_year_avgs <- full_join(flint_all_year_avgs, bioclim_all_year_avgs) %>% 
  select(TimePd, parent.pop:ppt_coldest_quarter) %>% 
  mutate(Season="Water Year")
names(bioclim_flint_all_year_avgs)
```

### Growth Season Averages
```{r}
flint_grwseason_avgs <- read_csv("../output/Climate/growthseason_FlintAvgs.csv")
bioclim_grwseason_avgs <- read_csv("../output/Climate/growthseason_BioClimAvgs.csv")

#Merge
bioclim_flint_grwseason_avgs <- full_join(flint_grwseason_avgs, bioclim_grwseason_avgs) %>% 
  select(TimePd, parent.pop:ppt_coldest_month) %>% 
  rename(tmean_wettest_quarter=tmean_wettest_month, 
         tmean_driest_quarter=tmean_driest_month,
         ppt_warmest_quarter=ppt_warmest_month,
         ppt_coldest_quarter=ppt_coldest_month) %>% 
  mutate(Season="Growth Season")
names(bioclim_flint_grwseason_avgs) #change month to quarter
```

### Dataframe with a column for seasonal summary 
```{r}
wtryr_grwssn_avgs <- bind_rows(bioclim_flint_all_year_avgs, bioclim_flint_grwseason_avgs)
```

### Correlations - Recent + Historical

```{r}
#normalize the data
climate_normalized_wtryr_grwssn_avgs <- wtryr_grwssn_avgs %>% ungroup() %>% 
  select(cwd:ppt_coldest_quarter) %>% scale() #normalize the data so they're all on the same scale

cor.norm = cor(climate_normalized_wtryr_grwssn_avgs) #test correlations among the traits
cor.sig <- cor.mtest(climate_normalized_wtryr_grwssn_avgs, method = "pearson")

corrplot(cor.norm, type="upper",
         tl.srt = 45, p.mat = cor.sig$p, 
         sig.level = 0.05, insig="blank")

#tmn, tmx, tmean_wettest_quarter, tmean_driest_quarter and ann_tmean all highly correlated (92-99%) - only keep ann_tmean 
#ann_ppt and ppt_coldest_quarter highly correlated (96%) - only keep ann_ppt 
#ppt_coldest_quarter and ppt_warmest_quarter (92%) - only keek ppt_warmest_quarter
```

### PCA - Recent + Historical
```{r}
wtryr_grwssn_avgs.pc = prcomp(wtryr_grwssn_avgs[c(7:9, 12:15, 18:20)], scale = TRUE, center = TRUE)
str(wtryr_grwssn_avgs.pc)
```

plot % Variance Explained

```{r}
summary(wtryr_grwssn_avgs.pc)

tibble(PC=str_c("PC",str_pad(1:10,2,pad="0")),
       percent_var=wtryr_grwssn_avgs.pc$sdev[1:10]^2/sum(wtryr_grwssn_avgs.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
wtryr_grwssn_avgs.pc.dat = data.frame(wtryr_grwssn_avgs.pc$x)

wtryr_grwssn_avgs_locs.pc = cbind(wtryr_grwssn_avgs, wtryr_grwssn_avgs.pc.dat)

wtryr_grwssn_avgs_loadings = data.frame(varnames=rownames(wtryr_grwssn_avgs.pc$rotation), wtryr_grwssn_avgs.pc$rotation)
wtryr_grwssn_avgs_loadings
```

```{r}
autoplot(wtryr_grwssn_avgs.pc, data = wtryr_grwssn_avgs,
         colour='elev_m', alpha=0.5,
         #label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
autoplot(wtryr_grwssn_avgs.pc, data = wtryr_grwssn_avgs,
         colour='Season', alpha=0.5,
         #label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic() +
  geom_encircle(aes(group=Season, colour = Season))
```

```{r}
autoplot(wtryr_grwssn_avgs.pc, data = wtryr_grwssn_avgs,
          x=1, y=3,
         colour='Season', alpha=0.5,
         #label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic() +
  geom_encircle(aes(group=Season, colour = Season))
```


## WL2-only PCA to show climate variation across years

### Load Data for WL2

#### Water Year
```{r}
yrly_avgs_recent_wtryr <- read_csv("../output/Climate/fullyear_wtr_year_avgs_Recent.csv") %>% 
  filter(parent.pop=="WL2")
yrly_avgs_hist_wtryr <- read_csv("../output/Climate/fullyear_wtr_year_avgs_Historical.csv") %>% 
  filter(parent.pop=="WL2")
```

#### Growth Season 
```{r}
yrly_avgs_recent_grwseason <- read_csv("../output/Climate/growthseason_yrly_avgs_Recent.csv") %>% 
  filter(parent.pop=="WL2")
yrly_avgs_hist_grwseason <- read_csv("../output/Climate/growthseason_yrly_avgs_Historical.csv") %>% 
  filter(parent.pop=="WL2")
```

#### WL2 Garden 2023 - Water Year 
```{r}
flint_all_year_wtr_yr <- read_csv("../output/Climate/flint_all_year_wtr_yr.csv")
pop_elev <- flint_all_year_wtr_yr %>% select(parent.pop:Long) %>% distinct()

bioclim_allyear_prep_WL2Grdn <- flint_all_year_wtr_yr %>% 
  filter(parent.pop == "WL2_Garden") %>%  #only keep WL2 garden
  rename(tmin=tmn, tmax=tmx, year_cal=year, year=wtr_yr) %>% #rename columns to match what calc_biovars expects, also make sure it uses water year 
  filter(year=="2023") %>% #year of the experiment only 
  arrange(parent.pop, year, month)

bioclim_all_year_WL2Grdn <- bioclim_allyear_prep_WL2Grdn %>% 
  calc_biovars() %>% 
  mutate(parent.pop="WL2_Garden", year=="2023") 

bioclim_all_year_final_WL2Grdn <- bioclim_all_year_WL2Grdn %>% 
  select(parent.pop, year, ann_tmean=bio1, mean_diurnal_range=bio2, 
         temp_seasonality=bio4, temp_ann_range=bio7, tmean_wettest_quarter=bio8,
         tmean_driest_quarter=bio9, ann_ppt=bio12, ppt_seasonality=bio15,
         ppt_warmest_quarter=bio18, ppt_coldest_quarter=bio19) 

WL2Grdn_elev_bioclim_all_year <- left_join(bioclim_all_year_final_WL2Grdn, pop_elev) %>% 
  select(parent.pop, elevation.group:Long, year:ppt_coldest_quarter)
WL2Grdn_elev_bioclim_all_year

WL2Grdn_flint_all_year <- flint_all_year_wtr_yr %>% 
  filter(parent.pop == "WL2_Garden", wtr_yr=="2023") %>%  #only keep WL2 garden
  select(parent.pop:month, wtr_yr, cwd, pck, ppt, tmn, tmx)

WL2Grdn_flint_all_year_summary <- WL2Grdn_flint_all_year %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) 

WL2Grdn_flint_bioclim_all_year <- left_join(WL2Grdn_flint_all_year_summary, WL2Grdn_elev_bioclim_all_year) %>% 
  select(-year)
```

#### WL2 Garden 2024 - Water Year (need to add flint_all_year_wtr_yr to this script)
```{r}
bioclim_2024_allyear_prep_WL2Grdn <- flint_all_year_wtr_yr %>% 
  filter(parent.pop == "WL2_Garden") %>%  #only keep WL2 garden
  rename(tmin=tmn, tmax=tmx, year_cal=year, year=wtr_yr) %>% #rename columns to match what calc_biovars expects, also make sure it uses water year 
  filter(year=="2024") %>% #year of the experiment only 
  arrange(parent.pop, year, month)

bioclim_2024_all_year_WL2Grdn <- bioclim_2024_allyear_prep_WL2Grdn %>% 
  calc_biovars() %>% 
  mutate(parent.pop="WL2_Garden", year=="2024") 

bioclim_2024_all_year_final_WL2Grdn <- bioclim_2024_all_year_WL2Grdn %>% 
  select(parent.pop, year, ann_tmean=bio1, mean_diurnal_range=bio2, 
         temp_seasonality=bio4, temp_ann_range=bio7, tmean_wettest_quarter=bio8,
         tmean_driest_quarter=bio9, ann_ppt=bio12, ppt_seasonality=bio15,
         ppt_warmest_quarter=bio18, ppt_coldest_quarter=bio19) 

WL2Grdn_elev_bioclim_2024_all_year <- left_join(bioclim_2024_all_year_final_WL2Grdn, pop_elev) %>% 
  select(parent.pop, elevation.group:Long, year:ppt_coldest_quarter)
WL2Grdn_elev_bioclim_2024_all_year

WL2Grdn_flint_2024_all_year <- flint_all_year_wtr_yr %>% 
  filter(parent.pop == "WL2_Garden", wtr_yr=="2024") %>%  #only keep WL2 garden
  select(parent.pop:month, wtr_yr, cwd, pck, ppt, tmn, tmx)

WL2Grdn_flint_2024_all_year_summary <- WL2Grdn_flint_2024_all_year %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx"), c(mean), na.rm = TRUE) 

WL2Grdn_flint_bioclim_2024_all_year <- left_join(WL2Grdn_flint_2024_all_year_summary, WL2Grdn_elev_bioclim_2024_all_year) %>% 
  select(-year)
```

#### WL2 Garden 2023 - Growth Season
```{r}
WL2Grdn_flint_bioclim_grwseason <- read_csv("../output/Climate/WL2_Garden_GrwSsnClim_2023.csv") %>% 
  rename_with(~str_remove(., "_WL2"), everything())
```

#### WL2 Garden2024 Growth Season 
```{r}
WL2Grdn_flint_bioclim_2024_grwseason <- read_csv("../output/Climate/WL2_Garden_GrwSsnClim_2024.csv") %>% 
  rename_with(~str_remove(., "_WL2"), everything())
```

### Correlations - Recent

#### Water Year 
```{r}
#normalize the data
clim_norm_wtryr_yrly_avgs_recent <- yrly_avgs_recent_wtryr %>% ungroup() %>% 
  select(cwd:ppt_coldest_quarter) %>% scale() #normalize the data so they're all on the same scale

cor.norm = cor(clim_norm_wtryr_yrly_avgs_recent) #test correlations among the traits
cor.sig <- cor.mtest(clim_norm_wtryr_yrly_avgs_recent, method = "pearson")

corrplot(cor.norm, type="upper",
         tl.srt = 45, p.mat = cor.sig$p, 
         sig.level = 0.05, insig="blank")
cor.norm
#tmn, tmx, and ann_tmean highly correlated - only keep ann_tmean
#ann_ppt and ppt 100% correlated - only keep ann_ppt 
```

#### Growth Season 
```{r}
#normalize the data
clim_norm_grwssn_yrly_avgs_recent <- yrly_avgs_recent_grwseason %>% ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% scale() #normalize the data so they're all on the same scale

cor.norm = cor(clim_norm_grwssn_yrly_avgs_recent) #test correlations among the traits
cor.sig <- cor.mtest(clim_norm_grwssn_yrly_avgs_recent, method = "pearson")

corrplot(cor.norm, type="upper",
         tl.srt = 45, p.mat = cor.sig$p, 
         sig.level = 0.05, insig="blank")
cor.norm
#tmn, tmx, and ann_tmean highly correlated - only keep ann_tmean
#ann_ppt, and ppt_coldest_month highly correlated - only keep ann_ppt 
```

### PCA - Recent

#### Water Year
```{r}
yrly_avgs_recent_wtryr.pc = prcomp(yrly_avgs_recent_wtryr[c(7:8, 12:21)], scale = TRUE, center = TRUE)
str(yrly_avgs_recent_wtryr.pc)
```

plot % Variance Explained

```{r}
summary(yrly_avgs_recent_wtryr.pc)

tibble(PC=str_c("PC",str_pad(1:12,2,pad="0")),
       percent_var=yrly_avgs_recent_wtryr.pc$sdev[1:12]^2/sum(yrly_avgs_recent_wtryr.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
yrly_avgs_recent_wtryr.pc.dat = data.frame(yrly_avgs_recent_wtryr.pc$x)

yrly_avgs_recent_wtryr_locs.pc = cbind(yrly_avgs_recent_wtryr, yrly_avgs_recent_wtryr.pc.dat) %>% 
  select(parent.pop:year, PC1:PC12)

yrly_avgs_recent_wtryr_loadings = data.frame(varnames=rownames(yrly_avgs_recent_wtryr.pc$rotation), yrly_avgs_recent_wtryr.pc$rotation)
yrly_avgs_recent_wtryr_loadings
```

```{r}
autoplot(yrly_avgs_recent_wtryr.pc, data = yrly_avgs_recent_wtryr,
         colour='year', alpha=0.6,
         #label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_color_viridis() +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
home_sites_pca <- yrly_avgs_recent_wtryr_locs.pc %>%  
  mutate(group=str_c(year))  %>%
  ggplot(aes(x=PC1, y=PC2, color=year)) +
  scale_color_viridis(breaks=c(1995, 2007, 2020), direction = -1) +
  geom_point(size=4, alpha=0.7) +
  labs(x="PC1 (40.1%)", y="PC2 (18.64%)", color="Year") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  annotate("text", x = -4, y = -7.6, label = "Dry \n No Snow", size=6) +
  annotate("text", x = 6.5, y = -7.6, label = "Wet \n Snow", size=6) +
  annotate("text", x = -5.1, y = -5, label = "Low Temp \n Range", size=6) +
  annotate("text", x = -5.1, y = 3, label = "High Temp \n Range", size=6) +
  coord_cartesian(ylim = c(-6, 4), xlim = c(-4,7.3), clip = "off") +
  theme_classic() +
  theme(text=element_text(size=28))

## add WL2 garden 2023 and 2024
WL2Grdn_pc_prep <- WL2Grdn_flint_bioclim_all_year %>% ungroup() %>% select(cwd:pck, ann_tmean, mean_diurnal_range, temp_seasonality, temp_ann_range, tmean_wettest_quarter, tmean_driest_quarter, ann_ppt, ppt_seasonality, ppt_warmest_quarter, ppt_coldest_quarter)
WL2Grdn_predicted <- predict(yrly_avgs_recent_wtryr.pc, newdata = WL2Grdn_pc_prep)

WL2Grdn_pc_prep_2024 <- WL2Grdn_flint_bioclim_2024_all_year %>% ungroup() %>% select(cwd:pck, ann_tmean, mean_diurnal_range, temp_seasonality, temp_ann_range, tmean_wettest_quarter, tmean_driest_quarter, ann_ppt, ppt_seasonality, ppt_warmest_quarter, ppt_coldest_quarter)
WL2Grdn_2024_predicted <- predict(yrly_avgs_recent_wtryr.pc, newdata = WL2Grdn_pc_prep_2024)

#str(home_sites_pca) #can add predicted WL2 point to the existing plot's data 
home_sites_pca$data <- rbind(home_sites_pca$data, 
  data.frame(
    parent.pop = "WL2_Garden",
    elevation.group="High",
    elev_m = 2020,
    Lat = 38.8263,
    Long=-120.2524,
    year = c(2023, 2024),
    PC1 = c(WL2Grdn_predicted[, "PC1"], WL2Grdn_2024_predicted[, "PC1"]),
    PC2 = c(WL2Grdn_predicted[, "PC2"], WL2Grdn_2024_predicted[, "PC2"]),
    PC3 = c(WL2Grdn_predicted[, "PC3"], WL2Grdn_2024_predicted[, "PC3"]),
    PC4 = c(WL2Grdn_predicted[, "PC4"], WL2Grdn_2024_predicted[, "PC4"]),
    PC5 = c(WL2Grdn_predicted[, "PC5"], WL2Grdn_2024_predicted[, "PC5"]),
    PC6 = c(WL2Grdn_predicted[, "PC6"], WL2Grdn_2024_predicted[, "PC6"]),
    PC7 = c(WL2Grdn_predicted[, "PC7"], WL2Grdn_2024_predicted[, "PC7"]),
    PC8 = c(WL2Grdn_predicted[, "PC8"], WL2Grdn_2024_predicted[, "PC8"]),
    PC9 = c(WL2Grdn_predicted[, "PC9"], WL2Grdn_2024_predicted[, "PC9"]),
    PC10 = c(WL2Grdn_predicted[, "PC10"], WL2Grdn_2024_predicted[, "PC10"]),
    PC11 = c(WL2Grdn_predicted[, "PC11"], WL2Grdn_2024_predicted[, "PC11"]),
    PC12 = c(WL2Grdn_predicted[, "PC12"], WL2Grdn_2024_predicted[, "PC12"]),
    group = c("new", "new2")
  )
)

home_sites_pca + 
  geom_point(data=filter(home_sites_pca$data, parent.pop == "WL2_Garden"), size=8, shape = 18, show_guide = FALSE) +
  annotate("text", x = 7, y= -1.6, label = "WL2 \n Garden \n 2023", colour = "purple", size=7) +
  annotate("text", x = -.14, y= -1.1, label = "WL2 \n Garden \n 2024", colour = "purple", size=7) 
ggsave("../output/Climate/Wtr_Year_WL2ONLYRecent_PC1-PC2.png", width = 7.4, height = 6, units = "in")
```

#### Growth Season 
```{r}
yrly_avgs_recent_grwseason.pc = prcomp(yrly_avgs_recent_grwseason[c(7:8, 12:20)], scale = TRUE, center = TRUE)
str(yrly_avgs_recent_grwseason.pc)
```

plot % Variance Explained

```{r}
summary(yrly_avgs_recent_grwseason.pc)

tibble(PC=str_c("PC",str_pad(1:11,2,pad="0")),
       percent_var=yrly_avgs_recent_grwseason.pc$sdev[1:11]^2/sum(yrly_avgs_recent_grwseason.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
yrly_avgs_recent_grwseason.pc.dat = data.frame(yrly_avgs_recent_grwseason.pc$x)

yrly_avgs_recent_grwseason_locs.pc = cbind(yrly_avgs_recent_grwseason, yrly_avgs_recent_grwseason.pc.dat) %>% 
  select(parent.pop:year, PC1:PC11)

yrly_avgs_recent_grwseason_loadings = data.frame(varnames=rownames(yrly_avgs_recent_grwseason.pc$rotation), yrly_avgs_recent_grwseason.pc$rotation)
yrly_avgs_recent_grwseason_loadings
```

```{r}
autoplot(yrly_avgs_recent_grwseason.pc, data = yrly_avgs_recent_grwseason,
         colour='year', alpha=0.6,
         #label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_color_viridis() +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```


```{r}
home_sites_pca <- yrly_avgs_recent_grwseason_locs.pc %>%  
  mutate(group=str_c(year))  %>%
  ggplot(aes(x=PC1, y=PC2, color=year)) +
  scale_color_viridis(breaks=c(1995, 2007, 2020), direction = -1) +
  geom_point(size=4, alpha=0.7) +
  labs(x="PC1 (31.74%)", y="PC2 (19.02%)", color="Year") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  annotate("text", x = -3.5, y = -5.3, label = "High Temp \n Seasonality", size=6) +
  annotate("text", x = 3, y = -5.3, label = "Low Temp \n Seasonality", size=6) +
  annotate("text", x = -4.9, y = -3, label = "High Ann \n Temp", size=6) +
  annotate("text", x = -4.9, y = 3, label = "Low Ann \n Temp", size=6) +
  coord_cartesian(ylim = c(-4, 4), xlim = c(-4,4), clip = "off") +
  theme_classic() +
  theme(text=element_text(size=28))

## add WL2 garden 2023 and 2024
WL2Grdn_pc_prep <- WL2Grdn_flint_bioclim_grwseason %>% ungroup() %>% select(cwd, pck, ann_tmean, mean_diurnal_range, temp_seasonality, temp_ann_range, tmean_wettest_month, tmean_driest_month, ann_ppt, ppt_seasonality, ppt_warmest_month)
WL2Grdn_predicted <- predict(yrly_avgs_recent_grwseason.pc, newdata = WL2Grdn_pc_prep)

WL2Grdn_pc_prep_2024 <- WL2Grdn_flint_bioclim_2024_grwseason %>% ungroup() %>% select(cwd, pck, ann_tmean, mean_diurnal_range, temp_seasonality, temp_ann_range, tmean_wettest_month, tmean_driest_month, ann_ppt, ppt_seasonality, ppt_warmest_month)
WL2Grdn_2024_predicted <- predict(yrly_avgs_recent_grwseason.pc, newdata = WL2Grdn_pc_prep_2024)

#str(home_sites_pca) #can add predicted WL2 point to the existing plot's data 
home_sites_pca$data <- rbind(home_sites_pca$data, 
  data.frame(
    parent.pop = "WL2_Garden",
    elevation.group="High",
    elev_m = 2020,
    Lat = 38.8263,
    Long=-120.2524,
    year = c(2023, 2024),
    PC1 = c(WL2Grdn_predicted[, "PC1"], WL2Grdn_2024_predicted[, "PC1"]),
    PC2 = c(WL2Grdn_predicted[, "PC2"], WL2Grdn_2024_predicted[, "PC2"]),
    PC3 = c(WL2Grdn_predicted[, "PC3"], WL2Grdn_2024_predicted[, "PC3"]),
    PC4 = c(WL2Grdn_predicted[, "PC4"], WL2Grdn_2024_predicted[, "PC4"]),
    PC5 = c(WL2Grdn_predicted[, "PC5"], WL2Grdn_2024_predicted[, "PC5"]),
    PC6 = c(WL2Grdn_predicted[, "PC6"], WL2Grdn_2024_predicted[, "PC6"]),
    PC7 = c(WL2Grdn_predicted[, "PC7"], WL2Grdn_2024_predicted[, "PC7"]),
    PC8 = c(WL2Grdn_predicted[, "PC8"], WL2Grdn_2024_predicted[, "PC8"]),
    PC9 = c(WL2Grdn_predicted[, "PC9"], WL2Grdn_2024_predicted[, "PC9"]),
    PC10 = c(WL2Grdn_predicted[, "PC10"], WL2Grdn_2024_predicted[, "PC10"]),
    PC11 = c(WL2Grdn_predicted[, "PC11"], WL2Grdn_2024_predicted[, "PC11"]),
    group = c("new", "new2")
  )
)

home_sites_pca + 
  geom_point(data=filter(home_sites_pca$data, parent.pop == "WL2_Garden"), size=8, shape = 18, show_guide = FALSE) +
  annotate("text", x = 2, y= 2, label = "WL2 Garden 2023", colour = "purple", size=7) +
  annotate("text", x = -1.3, y= -3.8, label = "WL2 Garden 2024", colour = "purple", size=7) 
ggsave("../output/Climate/GRWSSN_WL2ONLYRecent_PC1-PC2.png", width = 7.4, height = 6, units = "in")
```

### Corrleations - Historical 

#### Water Year
```{r}
#normalize the data
clim_norm_wtryr_yrly_avgs_hist <- yrly_avgs_hist_wtryr %>% ungroup() %>% 
  select(cwd:ppt_coldest_quarter) %>% scale() #normalize the data so they're all on the same scale

cor.norm = cor(clim_norm_wtryr_yrly_avgs_hist) #test correlations among the traits
cor.sig <- cor.mtest(clim_norm_wtryr_yrly_avgs_hist, method = "pearson")

corrplot(cor.norm, type="upper",
         tl.srt = 45, p.mat = cor.sig$p, 
         sig.level = 0.05, insig="blank")
cor.norm
#tmn, tmx and ann_tmean highly correlated - only keep ann_tmean
#ann_ppt and ppt 100% correlated - only keep ann_ppt 
```

#### Growth Season 
```{r}
#normalize the data
clim_norm_grwssn_yrly_avgs_hist <- yrly_avgs_hist_grwseason %>% ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% scale() #normalize the data so they're all on the same scale

cor.norm = cor(clim_norm_grwssn_yrly_avgs_hist) #test correlations among the traits
cor.sig <- cor.mtest(clim_norm_grwssn_yrly_avgs_hist, method = "pearson")

corrplot(cor.norm, type="upper",
         tl.srt = 45, p.mat = cor.sig$p, 
         sig.level = 0.05, insig="blank")
cor.norm
#tmx, and ann_tmean highly correlated - only keep ann_tmean
#ppt, ppt_coldest_month, pck, and ann_ppt highly corr - only keep ann_ppt
```

### PCA - Historical 

#### Water Year
```{r}
yrly_avgs_hist_wtryr.pc = prcomp(yrly_avgs_hist_wtryr[c(7:8, 12:21)], scale = TRUE, center = TRUE)
str(yrly_avgs_hist_wtryr.pc)
```

plot % Variance Explained

```{r}
summary(yrly_avgs_hist_wtryr.pc)

tibble(PC=str_c("PC",str_pad(1:12,2,pad="0")),
       percent_var=yrly_avgs_hist_wtryr.pc$sdev[1:12]^2/sum(yrly_avgs_hist_wtryr.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
yrly_avgs_hist_wtryr.pc.dat = data.frame(yrly_avgs_hist_wtryr.pc$x)

yrly_avgs_hist_wtryr_locs.pc = cbind(yrly_avgs_hist_wtryr, yrly_avgs_hist_wtryr.pc.dat) %>% 
  select(parent.pop:year, PC1:PC12)

yrly_avgs_hist_wtryr_loadings = data.frame(varnames=rownames(yrly_avgs_hist_wtryr.pc$rotation), yrly_avgs_hist_wtryr.pc$rotation)
yrly_avgs_hist_wtryr_loadings
```

```{r}
autoplot(yrly_avgs_hist_wtryr.pc, data = yrly_avgs_hist_wtryr,
         colour='year', alpha=0.6,
         #label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_color_viridis() +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
home_sites_pca <- yrly_avgs_hist_wtryr_locs.pc %>%  
  mutate(group=str_c(year))  %>%
  ggplot(aes(x=PC1, y=PC2, color=year)) +
  scale_color_viridis(breaks=c(1965, 1992, 2020), direction = -1) +
  geom_point(size=4, alpha=0.7) +
  labs(x="PC1 (43.18%)", y="PC2 (16.24)", color="Year") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  annotate("text", x = -6.2, y = -5.1, label = "Wet \n Snow", size=6) +
  annotate("text", x = 3.6, y = -5.1, label = "Dry \n No Snow", size=6) +
  annotate("text", x = -7.5, y = -3, label = "High \n Temp \n Seasonality", size=6) +
  annotate("text", x = -7.5, y = 2.8, label = "Low \n Temp \n Seasonality", size=6) +
  coord_cartesian(ylim = c(-4, 3), xlim = c(-6.5,4), clip = "off") +
  theme_classic() +
  theme(text=element_text(size=28))

## add WL2 garden 2023 and 2024
WL2Grdn_pc_prep <- WL2Grdn_flint_bioclim_all_year %>% ungroup() %>% select(cwd, pck, ann_tmean, mean_diurnal_range, temp_seasonality, temp_ann_range, tmean_wettest_quarter, tmean_driest_quarter, ann_ppt, ppt_seasonality, ppt_warmest_quarter, ppt_coldest_quarter)
WL2Grdn_predicted <- predict(yrly_avgs_hist_wtryr.pc, newdata = WL2Grdn_pc_prep)

WL2Grdn_pc_prep_2024 <- WL2Grdn_flint_bioclim_2024_all_year %>% ungroup() %>% select(cwd, pck, ann_tmean, mean_diurnal_range, temp_seasonality, temp_ann_range, tmean_wettest_quarter, tmean_driest_quarter, ann_ppt, ppt_seasonality, ppt_warmest_quarter, ppt_coldest_quarter)
WL2Grdn_2024_predicted <- predict(yrly_avgs_hist_wtryr.pc, newdata = WL2Grdn_pc_prep_2024)

#str(home_sites_pca) #can add predicted WL2 point to the existing plot's data 
home_sites_pca$data <- rbind(home_sites_pca$data, 
  data.frame(
    parent.pop = "WL2_Garden",
    elevation.group="High",
    elev_m = 2020,
    Lat = 38.8263,
    Long=-120.2524,
    year = c(2023, 2024),
    PC1 = c(WL2Grdn_predicted[, "PC1"], WL2Grdn_2024_predicted[, "PC1"]),
    PC2 = c(WL2Grdn_predicted[, "PC2"], WL2Grdn_2024_predicted[, "PC2"]),
    PC3 = c(WL2Grdn_predicted[, "PC3"], WL2Grdn_2024_predicted[, "PC3"]),
    PC4 = c(WL2Grdn_predicted[, "PC4"], WL2Grdn_2024_predicted[, "PC4"]),
    PC5 = c(WL2Grdn_predicted[, "PC5"], WL2Grdn_2024_predicted[, "PC5"]),
    PC6 = c(WL2Grdn_predicted[, "PC6"], WL2Grdn_2024_predicted[, "PC6"]),
    PC7 = c(WL2Grdn_predicted[, "PC7"], WL2Grdn_2024_predicted[, "PC7"]),
    PC8 = c(WL2Grdn_predicted[, "PC8"], WL2Grdn_2024_predicted[, "PC8"]),
    PC9 = c(WL2Grdn_predicted[, "PC9"], WL2Grdn_2024_predicted[, "PC9"]),
    PC10 = c(WL2Grdn_predicted[, "PC10"], WL2Grdn_2024_predicted[, "PC10"]),
    PC11 = c(WL2Grdn_predicted[, "PC11"], WL2Grdn_2024_predicted[, "PC11"]),
    PC12 = c(WL2Grdn_predicted[, "PC12"], WL2Grdn_2024_predicted[, "PC12"]),
    group = c("new", "new2")
  )
)

home_sites_pca + 
  geom_point(data=filter(home_sites_pca$data, parent.pop == "WL2_Garden"), size=8, shape = 18, show_guide = FALSE) +
  annotate("text", x = -6, y= -1.8, label = "WL2 \n Garden \n 2023", colour = "purple", size=7) +
  annotate("text", x = 1.4, y= -0.9, label = "WL2 \n Garden \n 2024", colour = "purple", size=7) 
ggsave("../output/Climate/Wtr_Year_WL2ONLYHist_PC1-PC2.png", width = 7.4, height = 6, units = "in")
```

#### Growth Season 
```{r}
yrly_avgs_hist_grwseason.pc = prcomp(yrly_avgs_hist_grwseason[c(7, 12:20)], scale = TRUE, center = TRUE)
str(yrly_avgs_hist_grwseason.pc)
```

plot % Variance Explained

```{r}
summary(yrly_avgs_hist_grwseason.pc)

tibble(PC=str_c("PC",str_pad(1:10,2,pad="0")),
       percent_var=yrly_avgs_hist_grwseason.pc$sdev[1:10]^2/sum(yrly_avgs_hist_grwseason.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")
```

Combine PCs with metadata

```{r}
yrly_avgs_hist_grwseason.pc.dat = data.frame(yrly_avgs_hist_grwseason.pc$x)

yrly_avgs_hist_grwseason_locs.pc = cbind(yrly_avgs_hist_grwseason, yrly_avgs_hist_grwseason.pc.dat) %>% 
  select(parent.pop:year, PC1:PC10)

yrly_avgs_hist_grwseason_loadings = data.frame(varnames=rownames(yrly_avgs_hist_grwseason.pc$rotation), yrly_avgs_hist_grwseason.pc$rotation)
yrly_avgs_hist_grwseason_loadings
```

```{r}
autoplot(yrly_avgs_hist_grwseason.pc, data = yrly_avgs_hist_grwseason,
         colour='year', alpha=0.6,
         #label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_color_viridis() +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
home_sites_pca <- yrly_avgs_hist_grwseason_locs.pc %>%  
  mutate(group=str_c(year))  %>%
  ggplot(aes(x=PC1, y=PC2, color=year)) +
  scale_color_viridis(breaks = c(1965, 1992, 2020), direction = -1) +
  geom_point(size=4, alpha=0.7) +
  labs(x="PC1 (30.58%)", y="PC2 (20.91%)", color="Year") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  annotate("text", x = -3.3, y = -6.8, label = "High Temp \n Seasonality", size=6) +
  annotate("text", x = 5, y = -6.8, label = "Low Temp \n Seasonality", size=6) +
  annotate("text", x = -5.3, y = -4, label = "Wet \n Cold", size=6) +
  annotate("text", x = -5.3, y = 5, label = "Dry \n Hot", size=6) +
  coord_cartesian(ylim = c(-5, 6), xlim = c(-4,6), clip = "off") +
  theme_classic() +
  theme(text=element_text(size=28))

## add WL2 garden 2023 and 2024
WL2Grdn_pc_prep <- WL2Grdn_flint_bioclim_grwseason %>% ungroup() %>% select(cwd, ann_tmean, mean_diurnal_range, temp_seasonality, temp_ann_range, tmean_wettest_month, tmean_driest_month, ann_ppt, ppt_seasonality, ppt_warmest_month)
WL2Grdn_predicted <- predict(yrly_avgs_hist_grwseason.pc, newdata = WL2Grdn_pc_prep)

WL2Grdn_pc_prep_2024 <- WL2Grdn_flint_bioclim_2024_grwseason %>% ungroup() %>% select(cwd, ann_tmean, mean_diurnal_range, temp_seasonality, temp_ann_range, tmean_wettest_month, tmean_driest_month, ann_ppt, ppt_seasonality, ppt_warmest_month)
WL2Grdn_2024_predicted <- predict(yrly_avgs_hist_grwseason.pc, newdata = WL2Grdn_pc_prep_2024)

#str(home_sites_pca) #can add predicted WL2 point to the existing plot's data 
home_sites_pca$data <- rbind(home_sites_pca$data, 
  data.frame(
    parent.pop = "WL2_Garden",
    elevation.group="High",
    elev_m = 2020,
    Lat = 38.8263,
    Long=-120.2524,
    year = c(2023, 2024),
    PC1 = c(WL2Grdn_predicted[, "PC1"], WL2Grdn_2024_predicted[, "PC1"]),
    PC2 = c(WL2Grdn_predicted[, "PC2"], WL2Grdn_2024_predicted[, "PC2"]),
    PC3 = c(WL2Grdn_predicted[, "PC3"], WL2Grdn_2024_predicted[, "PC3"]),
    PC4 = c(WL2Grdn_predicted[, "PC4"], WL2Grdn_2024_predicted[, "PC4"]),
    PC5 = c(WL2Grdn_predicted[, "PC5"], WL2Grdn_2024_predicted[, "PC5"]),
    PC6 = c(WL2Grdn_predicted[, "PC6"], WL2Grdn_2024_predicted[, "PC6"]),
    PC7 = c(WL2Grdn_predicted[, "PC7"], WL2Grdn_2024_predicted[, "PC7"]),
    PC8 = c(WL2Grdn_predicted[, "PC8"], WL2Grdn_2024_predicted[, "PC8"]),
    PC9 = c(WL2Grdn_predicted[, "PC9"], WL2Grdn_2024_predicted[, "PC9"]),
    PC10 = c(WL2Grdn_predicted[, "PC10"], WL2Grdn_2024_predicted[, "PC10"]),
    group = c("new", "new2")
  )
)

home_sites_pca + 
  geom_point(data=filter(home_sites_pca$data, parent.pop == "WL2_Garden"), size=8, shape = 18, show_guide = FALSE) +
  annotate("text", x = -0.4, y= -1.7, label = "WL2 Garden \n 2023", colour = "purple", size=7) +
  annotate("text", x = 1.1, y= 4.9, label = "WL2 Garden \n 2024", colour = "purple", size=7) 
ggsave("../output/Climate/GRWSSN_WL2ONLYHist_PC1-PC2.png", width = 7.4, height = 6, units = "in")
```