---
title: "PreTransplant_Size"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pre-Transplant Size for both gardens 
There shouldn't be any large differences since the plants were grown in the same environment pre-transplant. 
Is there a relationship between pre-transplant size and survival to the end of the first year? Does that differ across gardens?

## Relevant Libraries and Functions

```{r}
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
library(lmerTest) #for mixed effect models
conflicted::conflicts_prefer(lmerTest::lmer)
library(broom.mixed) #tidy method for lmerTest
library(emmeans) #for post-hoc pairwise comparisons 
library(naniar) #replaces values with NA

sem <- function(x, na.rm=FALSE) {           #for caclulating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low") #for proper arrangement in figures 
```

## Load the pop and location data

```{r}
gowersdist_UCD <- read_csv("../output/Climate/Pops_GowersEnvtalDist_UCD.csv") %>% 
  rename(Recent_Gowers_Dist_UCD = Recent_Gowers_Dist, Historic_Gowers_Dist_UCD = Historic_Gowers_Dist)

gowersdist_WL2 <- read_csv("../output/Climate/Pops_GowersEnvtalDist_WL2.csv") %>% 
  rename(Recent_Gowers_Dist_WL2 = Recent_Gowers_Dist, Historic_Gowers_Dist_WL2 = Historic_Gowers_Dist)

gowersdist_all <- full_join(gowersdist_UCD, gowersdist_WL2)
head(gowersdist_all)
```

## Load Pre-Transplant Size data from both Gardens
```{r}
ucd_pretrans_size <- read_csv("../input/UCD_Data/CorrectedCSVs/UCD_garden_size_measurements_20221128_corrected.csv",
                              na = c("", "NA", "-", "N/A"))
head(ucd_pretrans_size) #includes germs and non-germs 

ucd_pretrans_size_germonly <- ucd_pretrans_size %>% 
  rename(germ = `germinated?`, height.cm = `Height (cm)`, long.leaf.cm=`Longest leaf (cm)`) %>% 
  filter(germ==1) %>% 
  unite(Genotype, parent.pop:rep, sep="_", remove = FALSE) %>% 
  unite(pop.mf, parent.pop:mf, sep="_", remove = FALSE)
head(ucd_pretrans_size_germonly)
unique(ucd_pretrans_size_germonly$parent.pop) #looks correct
```

```{r}
wl2_pretrans_size1 <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_DNA_Collection_Size_survey_combined20230703_corrected.csv")
head(wl2_pretrans_size1)

wl2_pretrans_size2 <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_Extras_DNA_collection_size_survey_combined20230706_corrected.csv") %>% 
  filter(!is.na(`height (cm)`)) #to get rid of genotypes that were measured on the other data sheet (NAs on this sheet)
head(wl2_pretrans_size2)

wl2_pretrans_size_all <- bind_rows(wl2_pretrans_size1, wl2_pretrans_size2) %>%
  rename(parent.pop=Pop, height.cm = `height (cm)`, long.leaf.cm=`longest leaf (cm)`) %>% 
  unite(Genotype, parent.pop:rep, sep="_", remove = FALSE) %>% 
  unite(pop.mf, parent.pop:mf, sep="_", remove = FALSE)
head(wl2_pretrans_size_all)
```

## Merge the two sites
```{r}
ucd_pretrans_size_prep <- ucd_pretrans_size_germonly %>% select(Genotype:rep, height.cm:long.leaf.cm) %>%
  mutate(Site="UCD")
names(ucd_pretrans_size_prep)

wl2_pretrans_size_prep <- wl2_pretrans_size_all %>% select(Genotype:rep, height.cm:long.leaf.cm) %>%
  mutate(Site="WL2")
names(wl2_pretrans_size_prep)

pretrans_rxnnorms <- bind_rows(ucd_pretrans_size_prep, wl2_pretrans_size_prep) %>% 
  arrange(pop.mf, Site)
head(pretrans_rxnnorms)
```

## Add in Location Info
```{r}
pretrans_rxnnorms_loc <-left_join(pretrans_rxnnorms, gowersdist_all) 
head(pretrans_rxnnorms_loc)
```

## Load survival to end of first year data
```{r}
firstyear_surv <- read_csv("../output/firstyear_mort_both_sites.csv")
head(firstyear_surv)

firstyear_surv_merge_prep <- firstyear_surv %>% select(Genotype:rep, Site:Historic_Gowers_Dist_WL2)
```

## Merge size and survival
```{r}
pretrans_surv <- left_join(pretrans_rxnnorms_loc, firstyear_surv_merge_prep)
pretrans_surv %>% filter(is.na(Survival)) #12 genotypes with no survival information... should investigate

pretrans_surv_complete <- pretrans_surv %>% filter(!is.na(Survival))
```


## Plot Reaction Norms

### Means by Pop
```{r}
pretrans_rxnnorms_summary <- pretrans_rxnnorms_loc %>% 
  group_by(parent.pop, Site, elev_m) %>% 
  summarise(N_height = sum(!is.na(height.cm)), mean_height.cm = mean(height.cm,na.rm=(TRUE)), 
            sem_height.cm=sem(height.cm, na.rm=(TRUE)), N_length = sum(!is.na(long.leaf.cm)),
            mean_long.leaf.cm=mean(long.leaf.cm, na.rm=(TRUE)), 
            sem_long.leaf.cm=sem(long.leaf.cm, na.rm=TRUE)) %>% 
  mutate(Site=str_replace_all(Site, "UCD", "Low Elev"), 
         Site=str_replace_all(Site, "WL2", "High Elev"))
pretrans_rxnnorms_summary

pretrans_rxnnorms_summary$Site <- factor(pretrans_rxnnorms_summary$Site,
                                               levels = c('Low Elev','High Elev'))
```

### Plot Pop Avgs
```{r}
pretrans_rxnnorms_summary %>% 
  filter(N_height != 1) %>% 
  ggplot(aes(x=Site, y=mean_height.cm, group=parent.pop, color=elev_m)) + 
  geom_point(size=1.5) + geom_line(linewidth=1.5) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Height (cm)", color="Elevation (m)", title = "Pre-Transplant") +
 # ylim(0,4) +
  theme(text=element_text(size=28)) 
ggsave("../output/PreTrans_RxNorms_Height_ALL_PopAvgs.png", width = 14, height = 8, units = "in")

pretrans_rxnnorms_summary %>% 
  filter(N_length != 1) %>% 
  ggplot(aes(x=Site, y=mean_long.leaf.cm, group=parent.pop, color=elev_m)) + 
  geom_point(size=1.5) + geom_line(linewidth=1.5) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,
                    ymax=mean_long.leaf.cm+sem_long.leaf.cm),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Leaf Length (cm)", color="Elevation (m)", title = "Pre-Transplant") +
  theme(text=element_text(size=28))
ggsave("../output/PreTrans_RxNorms_LongestLength_ALL_PopAvgs.png", width = 14, height = 8, units = "in")
```
Plants were overall smaller when planted at WL2 than at Davis.

## Connection to survival to end of first year
```{r}
pretrans_surv_pop_avgs <- pretrans_surv_complete %>% 
  group_by(parent.pop, Site, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(Survival)), mean_Surv = mean(Survival,na.rm=(TRUE)), 
            sem_surv=sem(Survival, na.rm=(TRUE)), 
            N_height = sum(!is.na(height.cm)), mean_height.cm = mean(height.cm,na.rm=(TRUE)), 
            sem_height.cm=sem(height.cm, na.rm=(TRUE)), N_length = sum(!is.na(long.leaf.cm)),
            mean_long.leaf.cm=mean(long.leaf.cm, na.rm=(TRUE)), 
            sem_long.leaf.cm=sem(long.leaf.cm, na.rm=TRUE)) %>% 
  mutate(Site=str_replace_all(Site, "UCD", "Low Elev"), 
         Site=str_replace_all(Site, "WL2", "High Elev"))

pretrans_surv_pop_avgs %>% arrange(desc(mean_height.cm))

pretrans_surv_pop_avgs$Site <- factor(pretrans_surv_pop_avgs$Site,
                                               levels = c('Low Elev','High Elev'))
```

```{r}
pretrans_surv_pop_avgs %>% 
  ggplot(aes(x=mean_height.cm, y=mean_Surv, group=parent.pop, color=elev_m)) +
  geom_point(size=8) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Height (cm)" ,y="Survival", color="Elevation (m)",title = "Pre-Transplant Size") +
  theme(text=element_text(size=28)) +
  facet_wrap(vars(Site), scales="free")
ggsave("../output/Surv_PreTrans_Height_BothGardens.png", width = 18, height = 8, units = "in")

pretrans_surv_pop_avgs %>% 
  ggplot(aes(x=mean_long.leaf.cm, y=mean_Surv, group=parent.pop, color=elev_m)) +
  geom_point(size=8) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Leaf Length (cm)" ,y="Survival", color="Elevation (m)", title = "Pre-Transplant Size") +
  theme(text=element_text(size=28)) +
  facet_wrap(vars(Site), scales="free")
ggsave("../output/Surv_PreTrans_Length_BothGardens.png", width = 18, height = 8, units = "in")
```

### Log Reg survival ~ size 
Height
```{r}
basiclogit_height <- glm(Survival ~ height.cm, data = pretrans_surv_complete, family = "binomial")
basiclogit_height
summary(basiclogit_height)

basiclogit_height2 <- glmer(Survival ~ height.cm + (1|parent.pop/mf), data = pretrans_surv_complete, family = binomial(link = "logit"))
summary(basiclogit_height2)

logit_height <- glm(Survival ~ height.cm*Site*elev_m, data = pretrans_surv_complete, family = "binomial")
summary(logit_height)

logit_height2 <- glmer(Survival ~ height.cm*Site*elev_m + (1|parent.pop/mf), data = pretrans_surv_complete, family = binomial(link = "logit")) 
summary(logit_height2)
emtrends(logit_height2, pairwise ~ Site, var = "height.cm")
emmip(logit_height2, Site ~ height.cm, cov.reduce = range)
```

Length
```{r}
basiclogit_length <- glm(Survival ~ long.leaf.cm, data = pretrans_surv_complete, family = "binomial")
summary(basiclogit_length)

basiclogit_length2 <- glmer(Survival ~ long.leaf.cm + (1|parent.pop/mf), data = pretrans_surv_complete, family = binomial(link = "logit"))
summary(basiclogit_length2)

logit_length2 <- glmer(Survival ~ long.leaf.cm*Site + (1|parent.pop/mf), data = pretrans_surv_complete, family = binomial(link = "logit")) 
summary(logit_length2)
emtrends(logit_length2, pairwise ~ Site, var = "long.leaf.cm")
emmip(logit_length2, Site ~ long.leaf.cm, cov.reduce = range)
```