---
title: "Single Time Size Reaction Norms"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Reaction Norms of Size (between UCD and WL2)
Use maternal family averages instead of individual lines for each rep since don't know if rep# is the same across sites.
- Reference the analysis we used for Xavier's thesis 

## Relevant Libraries and Functions

```{r}
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
library(lmerTest) #for mixed effect models
conflicted::conflicts_prefer(lmerTest::lmer)
library(broom.mixed) #tidy method for lmerTest
library(emmeans) #for post-hoc pairwise comparisons 
library(naniar) #replaces values with NA
library(corrplot) #plotting correlations 
library(rstatix) #performing cor_test
sem <- function(x, na.rm=FALSE) {           #for caclulating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low") #for proper arrangement in figures 
```

## Load the pop and location data

```{r}
gowersdist_UCD <- read_csv("../output/Climate/Pops_GowersEnvtalDist_UCD.csv") %>% 
  rename(Recent_Gowers_Dist_UCD = Recent_Gowers_Dist, Historic_Gowers_Dist_UCD = Historic_Gowers_Dist)
head(gowersdist_UCD)
gowersdist_WL2 <- read_csv("../output/Climate/Pops_GowersEnvtalDist_WL2.csv") %>% 
  rename(Recent_Gowers_Dist_WL2 = Recent_Gowers_Dist, Historic_Gowers_Dist_WL2 = Historic_Gowers_Dist)
head(gowersdist_WL2)

gowersdist_all <- full_join(gowersdist_UCD, gowersdist_WL2)
head(gowersdist_all)
```

## Load Size data from both Gardens
Starting with just 2 months in the field 
```{r}
WL2_twomonths_size <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_size_survey_20230913_corrected.csv", 
                               na = c("", "NA", "-", "N/A")) %>% 
  rename(parent.pop=pop) %>% 
  mutate(parent.pop= str_replace(parent.pop, ".*VTR.*", "LVTR1")) %>% 
  unite(Genotype, parent.pop:rep, sep="_", remove = FALSE) %>% 
  unite(pop.mf, parent.pop:mf, sep="_", remove = FALSE)
head(WL2_twomonths_size)
dim(WL2_twomonths_size)
summary(WL2_twomonths_size)
WL2_twomonths_size %>% rowwise() %>% filter(!is.na(mf), mf != "buffer") %>%  filter(is.na(as.numeric(mf)))
WL2_twomonths_size %>% rowwise() %>% filter(!is.na(rep), rep != "buffer") %>%  filter(is.na(as.numeric(rep))) #all buffer rows 
unique(WL2_twomonths_size$parent.pop)

UCD_twomonths_size <- read_csv("../input/UCD_Data/CorrectedCSVs/Size_survey_transplants_20230127_corrected.csv", 
                               na = c("", "NA", "-", "N/A")) %>% 
  rename(height.cm = `height (cm)`, long.leaf.cm = `longest leaf (cm)`, parent.pop=pop) %>% 
  mutate(parent.pop= str_replace(parent.pop, ".*VTR.*", "LVTR1")) %>% 
  filter(rep != 100)  %>% #get rid of individuals that germinated in the field 
  unite(Genotype, parent.pop:rep, sep="_", remove = FALSE) %>% 
  unite(pop.mf, parent.pop:mf, sep="_", remove = FALSE)
head(UCD_twomonths_size)
dim(UCD_twomonths_size)
summary(UCD_twomonths_size)

UCD_twomonths_size %>% filter(!is.na(...10)) #extra notes 
```

## Merge the Data
```{r}
WL2_twomonths_size_prep <- WL2_twomonths_size %>% select(Genotype:long.leaf.cm) %>% mutate(Site="WL2") %>%
  mutate(mf=as.double(mf), rep=as.double(rep))
names(WL2_twomonths_size_prep)

UCD_twomonths_size_prep <- UCD_twomonths_size %>% select(Genotype:long.leaf.cm) %>% mutate(Site="UCD")
names(UCD_twomonths_size_prep)

twomonths_rxnnorms <- bind_rows(UCD_twomonths_size_prep, WL2_twomonths_size_prep) %>% 
  filter(!str_detect(Genotype, ".*buffer.*")) %>% 
  filter(!is.na(height.cm)) %>% 
  arrange(Genotype, Site)
#xtabs(~Genotype+Site, data=twomonths_rxnnorms)
head(twomonths_rxnnorms)

twomonths_rxnnorms_siten <- twomonths_rxnnorms %>% group_by(Genotype) %>% summarise(SiteN=sum(!is.na(Genotype))) %>% filter(SiteN==2)
twomonths_rxnnorms_siten

twomonths_rxnnorms_bothsites <- left_join(twomonths_rxnnorms_siten, twomonths_rxnnorms)
dim(twomonths_rxnnorms_bothsites)
head(twomonths_rxnnorms_bothsites)
unique(twomonths_rxnnorms_bothsites$parent.pop)
xtabs(~parent.pop + Site, data=twomonths_rxnnorms_bothsites)
```

## Add in location info
```{r}
twomonths_rxnnorms_bothsites_loc <- left_join(twomonths_rxnnorms_bothsites, gowersdist_all)
head(twomonths_rxnnorms_bothsites_loc)

twomonths_rxnnorms_loc <-left_join(twomonths_rxnnorms, gowersdist_all) 
head(twomonths_rxnnorms_loc)
```


## Plot Reaction Norms
### All Genotypes Colored by Pop
```{r}
twomonths_rxnnorms_bothsites_loc %>% 
  ggplot(aes(x=Site, y=height.cm, group=Genotype, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  theme_classic() +
   theme(text=element_text(size=25))
ggsave("../output/TwoMonths_RxNorms_Height_All_Genos.png", width = 12, height = 8, units = "in")

twomonths_rxnnorms_bothsites_loc %>% 
  ggplot(aes(x=Site, y=long.leaf.cm, group=Genotype, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  theme_classic()  +
   theme(text=element_text(size=25))
ggsave("../output/TwoMonths_RxNorms_LongestLength_All_Genos.png", width = 12, height = 8, units = "in")
```
### All Genotypes colored by elevation
```{r}
twomonths_rxnnorms_bothsites_loc %>% 
  ggplot(aes(x=Site, y=height.cm, group=Genotype, color=elev_m)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")

twomonths_rxnnorms_bothsites_loc %>% 
  ggplot(aes(x=Site, y=long.leaf.cm, group=Genotype, color=elev_m)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")
```

### Just TM2 and WL2
```{r}
twomonths_rxnnorms_bothsites_loc %>% 
  filter(parent.pop=="TM2"|parent.pop=="WL2") %>% 
  ggplot(aes(x=Site, y=height.cm, group=Genotype, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  theme_classic()  +
   theme(text=element_text(size=25))
ggsave("../output/TwoMonths_RxNorms_Height_TM2_WL2_Genos.png", width = 12, height = 8, units = "in")

twomonths_rxnnorms_bothsites_loc %>% 
  filter(parent.pop=="TM2"|parent.pop=="WL2") %>% 
  ggplot(aes(x=Site, y=long.leaf.cm, group=Genotype, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  theme_classic()  +
   theme(text=element_text(size=25))
ggsave("../output/TwoMonths_RxNorms_LongestLength_TM2_WL2_Genos.png", width = 12, height = 8, units = "in")
```

### Means by maternal family 
```{r}
twomonths_rxnnorms_summary_mfs <- twomonths_rxnnorms_loc %>% 
  group_by(pop.mf, parent.pop, Site, elev_m) %>% 
  summarise(N_height = sum(!is.na(height.cm)), mean_height.cm = mean(height.cm,na.rm=(TRUE)), 
            sem_height.cm=sem(height.cm, na.rm=(TRUE)), N_length = sum(!is.na(long.leaf.cm)),
            mean_long.leaf.cm=mean(long.leaf.cm, na.rm=(TRUE)), 
            sem_long.leaf.cm=sem(long.leaf.cm, na.rm=TRUE))
twomonths_rxnnorms_summary_mfs %>% arrange(N_height)

#restrict to mfs with plants at both sites
twomonths_rxnnorms_summary_mfs_wide <- twomonths_rxnnorms_summary_mfs %>% 
  select(pop.mf, Site, N_height) %>% 
  spread(Site, N_height) %>% 
  mutate(Both.Sites=if_else(!is.na(UCD) & !is.na(WL2), TRUE, FALSE)) %>% 
  filter(Both.Sites != "FALSE")
twomonths_rxnnorms_summary_mfs_wide %>% arrange(Both.Sites)

twomonths_rxnnorms_summary_mfs_bothsites <- left_join(twomonths_rxnnorms_summary_mfs_wide, twomonths_rxnnorms_summary_mfs)
twomonths_rxnnorms_summary_mfs_bothsites

twomonths_rxnnorms_summary_mfs2 <- twomonths_rxnnorms_summary_mfs_bothsites %>% 
  mutate(Site=str_replace_all(Site, "UCD", "Low Elev"), 
         Site=str_replace_all(Site, "WL2", "High Elev")) 
twomonths_rxnnorms_summary_mfs2$Site <- factor(twomonths_rxnnorms_summary_mfs2$Site,
                                               levels = c('Low Elev','High Elev'))
```

### Plot mf avgs
```{r}
twomonths_rxnnorms_summary_mfs %>% 
   ggplot(aes(x=Site, y=mean_height.cm, group=pop.mf, color=pop.mf)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.2) +
  theme_classic() +
   theme(text=element_text(size=25))

twomonths_rxnnorms_summary_mfs %>% 
   ggplot(aes(x=Site, y=mean_long.leaf.cm, group=pop.mf, color=pop.mf)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,ymax=mean_long.leaf.cm+sem_long.leaf.cm),width=.2) +
  theme_classic() +
   theme(text=element_text(size=25))
```

```{r}
#note that there are a lot of maternal families with only 1 plant - removed those from these figures 
twomonths_rxnnorms_summary_mfs2 %>% 
  filter(N_height != 1) %>% 
  ggplot(aes(x=Site, y=mean_height.cm, group=pop.mf, color=elev_m)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Height (cm)", color="Elevation (m)") +
  theme(text=element_text(size=28))
ggsave("../output/TwoMonths_RxNorms_Height_MFAvgs.png", width = 14, height = 8, units = "in")

twomonths_rxnnorms_summary_mfs2 %>% 
  filter(N_length != 1) %>% 
  ggplot(aes(x=Site, y=mean_long.leaf.cm, group=pop.mf, color=elev_m)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,ymax=mean_long.leaf.cm+sem_long.leaf.cm),
                width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Leaf Length (cm)", color="Elevation (m)") +
  theme(text=element_text(size=28))
ggsave("../output/TwoMonths_RxNorms_LongestLength_MFAvgs.png", width = 14, height = 8, units = "in")
```

```{r}
twomonths_rxnnorms_summary_mfs2 %>% 
  filter(N_height != 1) %>% 
  filter(parent.pop=="TM2"|parent.pop=="WL2") %>%  
  ggplot(aes(x=Site, y=mean_height.cm, group=pop.mf, color=elev_m)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Height (cm)", color="Elevation (m)") +
  theme(text=element_text(size=28))
ggsave("../output/TwoMonths_RxNorms_Height_TM2_WL2_MFAvgs.png", width = 14, height = 8, units = "in")

twomonths_rxnnorms_summary_mfs2 %>% 
  filter(N_length != 1) %>% 
  filter(parent.pop=="TM2"|parent.pop=="WL2") %>%  
  ggplot(aes(x=Site, y=mean_long.leaf.cm, group=pop.mf, color=elev_m)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,ymax=mean_long.leaf.cm+sem_long.leaf.cm),
                width=.1) +
  theme_classic() + scale_color_manual(low = "#F5A540", high = "#0043F0")  +
  labs(y="Leaf Length (cm)", color="Elevation (m)") +
  theme(text=element_text(size=28))
ggsave("../output/TwoMonths_RxNorms_LongestLength_TM2_WL2_MFAvgs.png", width = 14, height = 8, units = "in")
```


### Means by Pop
```{r}
twomonths_rxnnorms_loc %>% filter(parent.pop=="TM2") %>% filter(Site=="UCD")

twomonths_rxnnorms_summary <- twomonths_rxnnorms_loc %>% 
  group_by(parent.pop, Site, elev_m) %>% 
  summarise(N_height = sum(!is.na(height.cm)), mean_height.cm = mean(height.cm,na.rm=(TRUE)), 
            sem_height.cm=sem(height.cm, na.rm=(TRUE)), N_length = sum(!is.na(long.leaf.cm)),
            mean_long.leaf.cm=mean(long.leaf.cm, na.rm=(TRUE)), 
            sem_long.leaf.cm=sem(long.leaf.cm, na.rm=TRUE))
twomonths_rxnnorms_summary

twomonths_rxnnorms_summary %>% arrange(N_height)
twomonths_rxnnorms_summary %>% filter(elev_m>2800) %>% filter(Site=="WL2") %>% summary()
twomonths_rxnnorms_summary %>% filter(elev_m>2800) %>% filter(Site=="UCD") %>% summary()

twomonths_rxnnorms_summary2 <- twomonths_rxnnorms_summary %>% 
  mutate(Site=str_replace_all(Site, "UCD", "Low Elev"), 
         Site=str_replace_all(Site, "WL2", "High Elev"))
twomonths_rxnnorms_summary2$Site <- factor(twomonths_rxnnorms_summary2$Site,
                                               levels = c('Low Elev','High Elev'))

```

### Plot Pop Avgs
```{r}
twomonths_rxnnorms_summary %>% 
   ggplot(aes(x=Site, y=mean_height.cm, group=parent.pop, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.2) +
  theme_classic() +
   theme(text=element_text(size=25))
ggsave("../output/TwoMonths_RxNorms_Height_ALL_PopAvgs2.png", width = 12, height = 8, units = "in")

twomonths_rxnnorms_summary %>% 
   ggplot(aes(x=Site, y=mean_long.leaf.cm, group=parent.pop, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,ymax=mean_long.leaf.cm+sem_long.leaf.cm),width=.2) +
  theme_classic() +
   theme(text=element_text(size=25))
ggsave("../output/TwoMonths_RxNorms_LongestLength_ALL_PopAvgs2.png", width = 12, height = 8, units = "in")
```

```{r}
twomonths_rxnnorms_summary2 %>% 
  filter(N_height != 1) %>% 
  ggplot(aes(x=Site, y=mean_height.cm, group=parent.pop, color=elev_m)) + 
  geom_point(size=1.5) + geom_line(linewidth=1.5) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Height (cm)", color="Elevation (m)") +
 # ylim(0,4) +
  theme(text=element_text(size=28))
ggsave("../output/TwoMonths_RxNorms_Height_ALL_PopAvgs.png", width = 14, height = 8, units = "in")

twomonths_rxnnorms_summary2 %>% 
  filter(N_length != 1) %>% 
  ggplot(aes(x=Site, y=mean_long.leaf.cm, group=parent.pop, color=elev_m)) + 
  geom_point(size=1.5) + geom_line(linewidth=1.5) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,
                    ymax=mean_long.leaf.cm+sem_long.leaf.cm),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Leaf Length (cm)", color="Elevation (m)") +
  theme(text=element_text(size=28))
ggsave("../output/TwoMonths_RxNorms_LongestLength_ALL_PopAvgs.png", width = 14, height = 8, units = "in")
```

Combine mf and pop averages
```{r}
twomonths_rxnnorms_summary2 %>% 
  mutate(pop.mf=parent.pop) %>% 
  filter(N_height != 1) %>% 
  ggplot(aes(x=Site, y=mean_height.cm, group=pop.mf, color=elev_m)) + 
  geom_point(size=1.5) + geom_line(linewidth=1.5, alpha=0.7) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Height (cm)", color="Elevation (m)") +
  theme(text=element_text(size=28)) +
 geom_line(data = twomonths_rxnnorms_summary_mfs2, size=0.2) +
  geom_point(data = twomonths_rxnnorms_summary_mfs2, size=0.8) 
ggsave("../output/TwoMonths_RxNorms_Height_ALL_PopMFAvgs.png", width = 14, height = 8, units = "in")

twomonths_rxnnorms_summary2 %>% 
  mutate(pop.mf=parent.pop) %>% 
  filter(N_length != 1) %>% 
  ggplot(aes(x=Site, y=mean_long.leaf.cm, group=pop.mf, color=elev_m)) + 
  geom_point(size=1.5) + geom_line(linewidth=1.5,  alpha=0.7) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,
                    ymax=mean_long.leaf.cm+sem_long.leaf.cm),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Leaf Length (cm)", color="Elevation (m)") +
  theme(text=element_text(size=28)) +
geom_line(data = twomonths_rxnnorms_summary_mfs2, size=0.2) +
  geom_point(data = twomonths_rxnnorms_summary_mfs2, size=0.8) 
ggsave("../output/TwoMonths_RxNorms_LongestLength_ALL_PopMFAvgs.png", width = 14, height = 8, units = "in")
```


```{r}
twomonths_rxnnorms_summary2 %>% 
  filter(parent.pop=="TM2"|parent.pop=="WL2") %>%  
   ggplot(aes(x=Site, y=mean_height.cm, group=parent.pop, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.2) +
  theme_classic()  +
   theme(text=element_text(size=25))
ggsave("../output/TwoMonths_RxNorms_Height_TM2_WL2_PopAvgs.png", width = 12, height = 8, units = "in")

twomonths_rxnnorms_summary2 %>% 
  filter(parent.pop=="TM2"|parent.pop=="WL2") %>%  
   ggplot(aes(x=Site, y=mean_long.leaf.cm, group=parent.pop, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,ymax=mean_long.leaf.cm+sem_long.leaf.cm),width=.2) +
  theme_classic()  +
   theme(text=element_text(size=25))
ggsave("../output/TwoMonths_RxNorms_LongestLength_TM2_WL2_PopAvgs.png", width = 12, height = 8, units = "in")
```


### Means by Elevation
```{r}
twomonths_rxnnorms_summary2 <- twomonths_rxnnorms_bothsites_loc %>% 
  group_by(elevation.group, Site) %>% 
  summarise(N_height = sum(!is.na(height.cm)), mean_height.cm = mean(height.cm,na.rm=(TRUE)), 
            sem_height.cm=sem(height.cm, na.rm=(TRUE)), N_length = sum(!is.na(long.leaf.cm)),
            mean_long.leaf.cm=mean(long.leaf.cm, na.rm=(TRUE)), 
            sem_long.leaf.cm=sem(long.leaf.cm, na.rm=TRUE))
twomonths_rxnnorms_summary2
```


```{r}
twomonths_rxnnorms_summary2$elevation.group <- factor(twomonths_rxnnorms_summary2$elevation.group, levels = c('High','Mid', 'Low'))
twomonths_rxnnorms_summary2 %>% 
   ggplot(aes(x=Site, y=mean_height.cm, group=elevation.group, color=elevation.group)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.2) +
  theme_classic() + scale_colour_manual(values=elev_three_palette)

twomonths_rxnnorms_summary2 %>% 
   ggplot(aes(x=Site, y=mean_long.leaf.cm, group=elevation.group, color=elevation.group)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,ymax=mean_long.leaf.cm+sem_long.leaf.cm),width=.2) +
  theme_classic() + scale_colour_manual(values=elev_three_palette)
```


## Stats
Height
```{r}
summary(twomonths_rxnnorms_loc)
lmeheight <- lmer(height.cm ~ elev_m*Site + (1|parent.pop/mf), data = twomonths_rxnnorms_loc) 
summary(lmeheight)
ranova(lmeheight) 
anova(lmeheight)

emtrends(lmeheight, pairwise ~ Site, var = "elev_m")
emmip(lmeheight, Site ~ elev_m, cov.reduce = range)
```

Length
```{r}
summary(twomonths_rxnnorms_loc)
lmelength <- lmer(long.leaf.cm ~ elev_m*Site + (1|parent.pop/mf), data = twomonths_rxnnorms_loc) 
summary(lmelength)
ranova(lmelength) 
anova(lmelength)

emmip(lmelength, elev_m ~ Site)


emtrends(lmelength, pairwise ~ Site, var = "elev_m")
emmip(lmelength, Site ~ elev_m, cov.reduce = range)

```

