---
title: "Single_Time_Size"
author: "Brandie Quarles"
date: "2023-11-07"
output: 
  html_document: 
    keep_md: yes
---

To Do:

-   Figure out how to get Ranova output with map syntax 

-   ASK JULIN AND JENNY HOW WE FEEL ABOUT A SAMPLE SIZE OF 2

-   Calculate regression b/t height and longest leaf?

-   Analyze Pre-transplant size

    -   How predictive is this of survival in the field?

-   emmeans package for post-hoc pairwise comparisons

-   Analyze growth (remove rep# 100)

    -   Include pre-transplant size

    -   Get growth curve code from Julin

    -   Is the growth rate lower for high elev b/c drier and biennial?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Relevant Libraries and Functions

```{r}
library(tidyverse)
library(lmerTest) #for mixed effect models
library(emmeans) #for post-hoc pairwise comparisons 
library(naniar) #replaces values with NA
library(corrplot) #plotting correlations 
library(rstatix) #performing cor_test
sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")
```

## Load the pop and location data

```{r}
#pop info
pops_common_garden <- read_csv("../input/UCD_Data/Pops_for_2022_UCD.csv") #pops included in common garden 
summary(pops_common_garden)
pops_common_garden_nonotes <- pops_common_garden %>% select(parent.pop:elevation.group, UCD.seed.year)
pops_common_garden_nonotes$elevation.group <- str_to_title(pops_common_garden_nonotes$elevation.group)

#extra location info 
pop_loc <- read_csv("../input/Strep_tort_locs.csv")

unique(pop_loc$`Site code`)
unique(pops_common_garden_nonotes$parent.pop)
#need to change YOSE to YO
pop_loc_yo <- pop_loc %>% mutate(parent.pop = str_replace(`Site code`, "YOSE(\\d+)", "YO\\1")) %>% select(Lat, Long, elev_m=`Elevation (m)`, parent.pop)
unique(pop_loc_yo$parent.pop)

#merge in location info
pop_elev <- left_join(pops_common_garden_nonotes, pop_loc_yo)
head(pop_elev)
```

## Load size data using map

```{r}
files <- dir("../input/UCD_Data/CorrectedCSVs/",
             pattern="Size_survey_transplants_2022.*csv|Size_survey_transplants.*0127.*csv|Size_survey_transplants.*0217.*csv",
             full.names = TRUE)
files
single_time_all <- tibble(path=files, filename=basename(path)) %>%
  mutate(survey_date=str_extract(filename, "2[0-9]*"),
         survey_date=lubridate::ymd(survey_date),
         survey_month=c("dec_size", "jan_size", "feb_size"))
single_time_all

single_time_all <- single_time_all %>% 
  mutate(sheets=map(path, 
                    read_csv,
                    na = c("", "NA", "-", "N/A"))) %>% 
  select(-path)
single_time_all

map(single_time_all$sheets, head, 10) #check to make sure the datasheets look good 
single_time_all$sheets[[3]] %>%  rowwise() %>% filter(!is.na(mf)) %>%  filter(is.na(as.numeric(mf))) #. for a buffer --> safe to convert to double
single_time_all$sheets[[3]] <- single_time_all$sheets[[3]] %>% mutate(mf=as.double(mf)) #fix mf in third sheet

map(single_time_all$sheets, colnames) %>% #check the column names 
  unlist() %>%
  unique() %>%
  sort()
single_time_all$sheets[[1]] <- single_time_all$sheets[[1]] %>% rename(pop=parent.pop, col=column) #fix names in first sheet

#unnest to perform tidyverse functions 
single_time_all2 <- single_time_all %>% unnest(sheets) %>%
  rename(height_cm = `height (cm)`, longest_leaf_cm = `longest leaf (cm)`, parent.pop=pop) %>% 
  filter(parent.pop != "buffer") %>% 
  mutate(parent.pop= str_replace(parent.pop, ".*VTR.*", "LVTR1")) %>% 
  filter(rep != 100)
head(single_time_all2)
unique(single_time_all2$parent.pop)

#merge in pop loc info
single_time_all2_elev <- left_join(single_time_all2, pop_elev) %>% select(survey_date:Notes, `herbivory (Y or N)`, elevation.group,Lat, Long, elev_m)
head(single_time_all2_elev)
names(single_time_all2_elev)
```

## Check Sample sizes at each survey
If any pops only have 1 individual, filter those pops out 

```{r}
sample_sizes <- single_time_all2_elev %>% 
  filter(!is.na(height_cm)) %>% 
  group_by(survey_month, parent.pop) %>% 
  summarise(samplesize=n()) %>% 
  arrange(survey_month, samplesize)
sample_sizes

#Dec WV has only 2 individuals 
#Jan LV1 and WV only have 1 indiv
#Feb LV1, YO4 have 1 individual, YO11 has only 2 individuals 

single_time_all_sample_sizes <- left_join(single_time_all2_elev, sample_sizes)
single_time_all_sample_sizes %>% filter(samplesize == 1)

single_time_all_GTONE <- single_time_all_sample_sizes %>% filter(samplesize > 1) #only pops with greater than one individual alive at a given survey 
```

## Summary plots

```{r}
single_time_all_GTONE$survey_month <- factor(single_time_all_GTONE$survey_month, levels = c("dec_size","jan_size","feb_size"))
#Histograms
single_time_all_GTONE %>% select(height_cm, longest_leaf_cm) %>% 
  names() %>% 
  map(~ggplot(single_time_all_GTONE, aes_string(x = .)) + 
        geom_histogram() + 
        facet_grid(survey_month ~ ., scales = "free"))

#qq plots 
single_time_all_GTONE %>% select(height_cm, longest_leaf_cm) %>% 
  names() %>% 
  map(~ggplot(single_time_all_GTONE, aes_string(sample = .)) + 
       stat_qq() +
        stat_qq_line() +
        facet_grid(survey_month ~ ., scales="free"))

#all months height is right skewed
#feb longest leaf very slightly right skewed


#Transformations:
single_time_all_GTONE_transf <- single_time_all_GTONE %>% 
  mutate(sheight_cm=sqrt(height_cm), logheight_cm=log(height_cm),log10height_cm=log10(height_cm),
         slongest_leaf_cm=sqrt(longest_leaf_cm), loglongest_leaf_cm=log(longest_leaf_cm),
         log10longest_leaf_cm=log10(longest_leaf_cm))

single_time_all_GTONE_transf$survey_month <- factor(single_time_all_GTONE_transf$survey_month, levels = c("dec_size","jan_size","feb_size"))

single_time_all_GTONE_transf %>% select(ends_with("cm")) %>% 
  names() %>% 
  map(~ggplot(single_time_all_GTONE_transf, aes_string(x = .)) + 
        geom_histogram() + 
        facet_grid(survey_month ~ ., scales = "free"))

single_time_all_GTONE_transf %>% select(ends_with("cm")) %>% 
  names() %>% 
  map(~ggplot(single_time_all_GTONE_transf, aes_string(sample = .)) + 
       stat_qq() +
        stat_qq_line() +
        facet_grid(survey_month ~ ., scales="free"))

#square root transformation is the best for height (all months)
#no transformations greatly improve Feb distribution 
```

## Relationship between heigh and longest leaf 

### Plotting Height and Length Together
```{r}
single_time_all_GTONE_transf %>% ggplot(aes(x=height_cm, y=longest_leaf_cm)) +
  geom_point() +
  geom_smooth() +
  facet_grid(survey_month ~ ., scales="free")

#generally positive relationship, but a lot more variability at lower heights 
#maybe a more logarithmic scale than linear 

single_time_all_GTONE_transf %>% ggplot(aes(x=longest_leaf_cm, y=height_cm)) +
  geom_point() +
  geom_smooth() +
  facet_grid(survey_month ~ ., scales="free") #opposite direction
```

### Calculating correlation b/t the two
```{r}
single_time_all_GTONE_transf %>% cor_test(height_cm, longest_leaf_cm, method = "pearson") #all months 
#r=0.52, P <0.0001

single_time_all_GTONE_transf %>% group_by(survey_month) %>% 
  cor_test(height_cm, longest_leaf_cm, method = "pearson")
#Dec r=0.51, P <0.0001
#Jan r=0.47, P <0.0001
#Feb r=0.51, P <0.0001
```

## Calculating means by pop (and elev) 

```{r}
single_time_all_GTONE_transf_summary <- single_time_all_GTONE_transf %>% 
  group_by(survey_month, parent.pop, elev_m) %>% 
  summarise(mean_height_cm = mean(height_cm,na.rm=(TRUE)), 
            sem_height_cm=sem(height_cm, na.rm=(TRUE)),
            mean_longest_leaf_cm=mean(longest_leaf_cm, na.rm=(TRUE)), 
            sem_longest_leaf_cm=sem(longest_leaf_cm, na.rm=TRUE))
single_time_all_GTONE_transf_summary
```

## Figures of averages 

### Avg Height Figure
```{r}
single_time_all_GTONE_transf_summary$survey_month <- factor(single_time_all_GTONE_transf_summary$survey_month, levels = c("dec_size","jan_size","feb_size"))

single_time_all_GTONE_transf_summary %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_height_cm), y=mean_height_cm, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=mean_height_cm-sem_height_cm,ymax=mean_height_cm+sem_height_cm),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Height (cm)", x="Population") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) +
  facet_grid(survey_month ~ .)
ggsave("../output/UCD_Traits/Single_Time_Height.png", width = 14, height = 18, units = "in")
```


### Avg Longest Leaf Figure
```{r}
single_time_all_GTONE_transf_summary %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_longest_leaf_cm), y=mean_longest_leaf_cm, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=mean_longest_leaf_cm-sem_longest_leaf_cm,ymax=mean_longest_leaf_cm+sem_longest_leaf_cm),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Longest Leaf (cm)", x="Population") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) +
  facet_grid(survey_month ~ .)
ggsave("../output/UCD_Traits/Single_Time_Longest_Leaf.png", width = 14, height = 18, units = "in")
```

## Mixed Effects Models
```{r}
#prep data for model
single_time_split_by_month <- single_time_all_GTONE_transf %>% 
  select(survey_month, elev_m, parent.pop, mf, sheight_cm, longest_leaf_cm) %>% 
  group_nest(survey_month)
single_time_split_by_month

model_by_month <- 
  single_time_split_by_month %>% 
  mutate(height_model_1 = map(data, ~ lmer(sheight_cm ~ (1|parent.pop/mf), data = .x)),
         height_model_2 = map(data, ~ lmer(sheight_cm ~ elev_m + (1|parent.pop/mf), data = .x)),
         leaf_model_1 = map(data, ~ lmer(longest_leaf_cm ~ (1|parent.pop/mf), data = .x)), 
         leaf_model_2 = map(data, ~ lmer(longest_leaf_cm ~ elev_m + (1|parent.pop/mf), data = .x)),)
model_by_month

library(broom.mixed)
model_by_month %>% 
  mutate(coef = map(height_model_1, tidy)) %>% 
  select(survey_month, coef) %>% 
  unnest(cols = c(coef))

#get Julin's code for comparing models 

#figure out how to get the below output:
#ranova(lmeheight)
#VarCorr(lmeheight) %>% 
 # as.data.frame() %>% 
  #mutate(var.pct = vcov/sum(vcov)*100)
```
