---
title: "Total Fitness"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Total Fitness at both sites

Total Fitness: p(Establishment)*p(Surv to Rep - y1)*Fruits(y1) + p(Winter Surv)*p(Surv to Rep - y2)*Fruits(y2)

Relating phenology and plasticity to fitness:

-   Prob of fitness \~

    -   plasticity in size (height, leaf length y1)

    -   days to budding

    -   plasticity in days to budding

    -   Budding duration

-   Total fitness \~

    -   Days to budding

    -   Plasticity in days to budding

    -   Budding duration

-   Standard error correction on scatter plots

## Libraries

```{r}
library(tidyverse)
library(magrittr)
#conflicted::conflicts_prefer(dplyr::filter)
library(viridis) #for pretty colors
library(ggrepel) #for non-overlapping labels on plots
library(ggdist) #visualizations of distributions and uncertainty 
library(ggpubr)

#library(sp) #for calculating geographic distance 
library(geosphere) #for calculating geographic distance
library(corrplot) #plotting correlations  
library(rstatix) #performing cor_test

library(lmerTest) #mixed models
conflicted::conflicts_prefer(lmerTest::lmer)
 library(emmeans) #least squared means 
library(broom.mixed)
library(tidymodels)
library(furrr) #Apply Mapping Functions in Parallel using Futures
tidymodels_prefer()
# install.packages("multilevelmod")
library(multilevelmod)

library(brms)
#conflicted::conflicts_prefer(brms::ar)
#conflicted::conflicts_prefer(dplyr::combine)
#conflicted::conflicts_prefer(brms::dstudent_t)
library(tidybayes) #for extracting and visiaulizing brms model output 
library(modelr) #for data grid

sem <- function(x, na.rm=FALSE) {           #for caclulating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 

#load inv.logit function
inv.logit <- function(f,a) {
  a <- (1-2*a)
  (a*(1+exp(f))+(exp(f)-1))/(2*a*(1+exp(f)))
}

cbbPalette2 <- c("#E69F00","#000000", "#56B4E9","#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
timepd_palette <- c("#56B4E9","#D55E00")

options(mc.cores = parallel::detectCores())
```

## Gower's Distance

```{r}
garden_climate <- read_csv("../output/Climate/flint_climate_UCDpops.csv") %>% 
  filter(parent.pop=="WL2_Garden" | parent.pop=="UCD_Garden") %>% 
  select(parent.pop:Long) %>% 
  distinct()
garden_climate
#UCD LAT/LONG = 38.53250, -121.7830
#WL2 Lat/Long = 38.82599, -120.2509

ucd_gowers <- read_csv("../output/Climate/Gowers_UCD.csv") %>% 
  select(parent.pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  mutate(UCD_Lat=38.53250, UCD_Long=-121.7830, UCD_Elev=16) %>% 
  mutate(Geographic_Dist=distHaversine(cbind(UCD_Long, UCD_Lat), cbind(Long, Lat)),
         Elev_Dist=elev_m-UCD_Elev) %>% # Calculate the distance using the haversine formula (dist in meters)
  #mutate(Lat_Dist=UCD_Lat-Lat, Long_Dist=UCD_Long-Long) %>% #Garden-Home - lat and long per Gerst et al 2011 which kept them separate for some directionality
  rename(pop=parent.pop)

wl2_gowers_2023 <- read_csv("../output/Climate/Gowers_WL2.csv") %>% 
  select(parent.pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  mutate(WL2_Lat=38.82599, WL2_Long=-120.2509, WL2_Elev=2020) %>% 
  mutate(Geographic_Dist=distHaversine(cbind(WL2_Long, WL2_Lat), cbind(Long, Lat)),
         Elev_Dist=elev_m-WL2_Elev) %>% # Calculate the distance using the haversine formula
  #mutate(Lat_Dist=WL2_Lat-Lat, Long_Dist=WL2_Long-Long) %>% #Garden-Home - lat and long per Gerst et al 2011 which kept them separate for some directionality
  rename(pop=parent.pop)

wl2_gowers_2024 <- read_csv("../output/Climate/Gowers_WL2_2024.csv") %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  mutate(WL2_Lat=38.82599, WL2_Long=-120.2509, WL2_Elev=2020) %>% 
  mutate(Geographic_Dist=distHaversine(cbind(WL2_Long, WL2_Lat), cbind(Long, Lat)),
         Elev_Dist=elev_m-WL2_Elev) %>% # Calculate the distance using the haversine formula
  rename(pop=parent.pop)
```

### Climate Subtraction Distance

```{r}
ucd_wtr_year_sub_recent <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_Davis_Recent.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Recent=ann_tmean_dist, Wtr_Year_PPTDist_Recent=ann_ppt_dist)
ucd_wtr_year_sub_historic <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_Davis_Historical.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Historic=ann_tmean_dist, Wtr_Year_PPTDist_Historic=ann_ppt_dist)
ucd_grwssn_sub_recent <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_Home_Davis_Recent.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Recent=ann_tmean_dist, GrwSsn_PPTDist_Recent=ann_ppt_dist)
ucd_grwssn_sub_historic <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_Home_Davis_Historical.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Historic=ann_tmean_dist, GrwSsn_PPTDist_Historic=ann_ppt_dist)
ucd_sub_dist <- ucd_wtr_year_sub_recent %>% 
  left_join(ucd_wtr_year_sub_historic) %>% 
  left_join(ucd_grwssn_sub_recent) %>% 
  left_join(ucd_grwssn_sub_historic) %>% 
  rename(pop=parent.pop) %>% 
  left_join(ucd_gowers)

wl2_wtr_year_sub_recent <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_Recent.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Recent=ann_tmean_dist, Wtr_Year_PPTDist_Recent=ann_ppt_dist)
wl2_wtr_year_sub_historic <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_Historical.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Historic=ann_tmean_dist, Wtr_Year_PPTDist_Historic=ann_ppt_dist)
wl2_grwssn_sub_recent <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_Home_WL2_Recent.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Recent=ann_tmean_dist, GrwSsn_PPTDist_Recent=ann_ppt_dist)
wl2_grwssn_sub_historic <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_Home_WL2_Historical.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Historic=ann_tmean_dist, GrwSsn_PPTDist_Historic=ann_ppt_dist)
wl2_sub_dist_2023 <- wl2_wtr_year_sub_recent %>% 
  left_join(wl2_wtr_year_sub_historic) %>% 
  left_join(wl2_grwssn_sub_recent) %>% 
  left_join(wl2_grwssn_sub_historic) %>% 
  rename(pop=parent.pop) %>% 
  left_join(wl2_gowers_2023)

wl2_wtr_year_sub_recent_2024 <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_2024_Recent.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Recent=ann_tmean_dist, Wtr_Year_PPTDist_Recent=ann_ppt_dist)
wl2_wtr_year_sub_historic_2024 <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_2024_Historical.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Historic=ann_tmean_dist, Wtr_Year_PPTDist_Historic=ann_ppt_dist)
wl2_grwssn_sub_recent_2024 <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_Home_WL2_2024_Recent.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Recent=ann_tmean_dist, GrwSsn_PPTDist_Recent=ann_ppt_dist)
wl2_grwssn_sub_historic_2024 <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_Home_WL2_2024_Historical.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Historic=ann_tmean_dist, GrwSsn_PPTDist_Historic=ann_ppt_dist)
wl2_sub_dist_2024 <- wl2_wtr_year_sub_recent_2024 %>% 
  left_join(wl2_wtr_year_sub_historic_2024) %>% 
  left_join(wl2_grwssn_sub_recent_2024) %>% 
  left_join(wl2_grwssn_sub_historic_2024) %>% 
  rename(pop=parent.pop) %>% 
  left_join(wl2_gowers_2024)
```

### Check Geographic distance

```{r}
ucd_gowers %>% 
  ggplot(aes(x=fct_reorder(pop, Geographic_Dist), y=Geographic_Dist, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  labs(y="Geographic Distance (m)", x="Population", fill="Water Year \n Climate Distance") +
  theme_classic() +
  scale_fill_viridis(option="mako", direction = -1) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1)) 
#ggsave("../output/UCD_GeoDist.png", width = 12, height = 8, units = "in")

wl2_gowers_2023 %>% 
  ggplot(aes(x=fct_reorder(pop, Geographic_Dist), y=Geographic_Dist, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  labs(y="Geographic Distance (m)", x="Population", fill="Water Year \n Climate Distance") +
  theme_classic() +
  scale_fill_viridis(option="mako", direction = -1) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1)) 
#ggsave("../output/WL2_GeoDist.png", width = 12, height = 8, units = "in")
```

### Checking elevation distance

```{r}
ucd_gowers %>% 
  ggplot(aes(x=fct_reorder(pop, Elev_Dist), y=Elev_Dist, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  labs(y="Elevation Distance (m)", x="Population", fill="Water Year \n Climate Distance") +
  theme_classic() +
  scale_fill_viridis(option="mako", direction = -1) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1)) 
#ggsave("../output/UCD_ElevDist.png", width = 12, height = 8, units = "in")

wl2_gowers_2023 %>% 
  ggplot(aes(x=fct_reorder(pop, Elev_Dist), y=Elev_Dist, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  labs(y="Elevation Distance (m)", x="Population", fill="Water Year \n Climate Distance") +
  theme_classic() +
  scale_fill_viridis(option="mako", direction = -1) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1)) 
#ggsave("../output/WL2_ElevDist.png", width = 12, height = 8, units = "in")

## Correlation between geo dist, elev distance and climate distance
dist_normalized_ucd <- ucd_gowers %>% select(GrwSsn_GD_Recent:Wtr_Year_GD_Historical, Geographic_Dist, Elev_Dist) %>% scale() #normalize the data so they're all on the same scale
#head(dist_normalized_ucd)
cor.norm_ucd = cor(dist_normalized_ucd) #test correlations among the traits
cor.sig_ucd <- cor.mtest(dist_normalized_ucd, method="pearson") #test significance of corrs
corrplot(cor.norm_ucd, type = "upper",
         tl.srt = 45, p.mat = cor.sig_ucd$p, 
         sig.level = 0.05, insig="blank") 
#geo dist not strongly correlated to any of the climate distances, yay!
##elev dist strongly neg correlated w/ water year climate distances
cor.norm_ucd
cor.sig_ucd$p

dist_normalized_wl2 <- wl2_gowers_2023 %>% select(GrwSsn_GD_Recent:Wtr_Year_GD_Historical, Geographic_Dist, Elev_Dist) %>% scale() #normalize the data so they're all on the same scale
#head(dist_normalized_wl2)
cor.norm_wl2 = cor(dist_normalized_wl2) #test correlations among the traits
cor.sig_wl2 <- cor.mtest(dist_normalized_wl2, method="pearson") #test significance of corrs
corrplot(cor.norm_wl2, type = "upper",
         tl.srt = 45, p.mat = cor.sig_wl2$p, 
         sig.level = 0.05, insig="blank")  
#geo dist not strongly correlated to any of the climate distances, yay!
#elev positively correlated with recent water year climate distance 
cor.norm_wl2
cor.sig_wl2$p
```

### WL2 Avg Gowers

```{r}
wl2_gowers_2023_prep <- wl2_gowers_2023 %>% rename_with(~paste0(., "_2023"), 6:9)
wl2_gowers_2024_prep <- wl2_gowers_2024 %>% rename_with(~paste0(., "_2024"), 6:9)

wl2_gowers_avg <- bind_rows(wl2_gowers_2023, wl2_gowers_2024) %>% 
  group_by(pop, elevation.group, elev_m, Geographic_Dist, Elev_Dist) %>% 
  summarise_at(c("GrwSsn_GD_Recent",  "GrwSsn_GD_Historical", "Wtr_Year_GD_Recent", "Wtr_Year_GD_Historical"), c(mean), na.rm = TRUE)
wl2_gowers_avg
```

#### For Subtraction Distances

```{r}
wl2_sub_dist_2023_prep <- wl2_sub_dist_2023 %>% 
  select(pop:GrwSsn_PPTDist_Historic)
wl2_sub_dist_2024_prep <- wl2_sub_dist_2024 %>% 
  select(pop:GrwSsn_PPTDist_Historic) 

wl2_sub_dist_avg <- bind_rows(wl2_sub_dist_2023_prep, wl2_sub_dist_2024_prep) %>% 
  group_by(pop) %>% 
  summarise_at(c("Wtr_Year_TempDist_Recent",  "Wtr_Year_PPTDist_Recent", 
                 "Wtr_Year_TempDist_Historic", "Wtr_Year_PPTDist_Historic",
                 "GrwSsn_TempDist_Recent", "GrwSsn_PPTDist_Recent",
                 "GrwSsn_TempDist_Historic", "GrwSsn_PPTDist_Historic"), c(mean), na.rm = TRUE) %>% 
  left_join(wl2_gowers_avg)
wl2_sub_dist_avg
```

Correlations

```{r}
wl2_sub_dist_avgs_scaled <- wl2_sub_dist_avg %>% ungroup() %>% 
  select(Wtr_Year_TempDist_Recent:GrwSsn_PPTDist_Historic, Geographic_Dist) %>% scale() #normalize the data so they're all on the same scale

cor.norm = cor(wl2_sub_dist_avgs_scaled) #test correlations among the traits
cor.sig <- cor.mtest(wl2_sub_dist_avgs_scaled, method = "pearson")

corrplot(cor.norm, type="upper",
         tl.srt = 45, p.mat = cor.sig$p, 
         sig.level = 0.05, insig="blank")
```

## Load Davis Fitness Components

```{r}
ucd_establishment <- read_csv("../output/UCD_Traits/UCD_Establishment.csv")
ucd_surv_to_rep <- read_csv("../output/UCD_Traits/UCD_SurvtoRep.csv")
ucd_fruits <- read_csv("../output/UCD_Traits/UCD_Fruits.csv")
```

### UCD - Check summer survival (equivalent to winter)

```{r}
ucd_surv <- read_csv("../input/UCD_Data/CorrectedCSVs/UCD_transplants_pheno_mort_20231016_corrected.csv") %>% 
  rename(death.date=`Death Date`, bud.date=`Date First Bud`, flower.date=`Date First Flower`, 
         fruit.date=`Date First Fruit`, last.flower.date=`Date Last Flower`, last.fruit.date=`Date Last Fruit`) %>% 
  filter(!is.na(pop)) %>% 
  filter(rep != 100) %>% #get rid of individuals that germinated in the field 
  unite(Genotype, pop:rep, sep="_", remove = FALSE) 
#unique(ucd_surv$pop)
```

```{r}
ucd_surv %>% filter(is.na(death.date)) #filter the alive plants on 10/16/23
#9 plants alive in October, all BH 
#5 did not reproduce in year 1 
```

## Load WL2 Fitness Components

```{r}
wl2_establishment <- read_csv("../output/WL2_Traits/WL2_Establishment.csv")
wl2_surv_to_rep_y1 <- read_csv("../output/WL2_Traits/WL2_SurvtoRep_y1.csv")
wl2_fruits_y1 <- read_csv("../output/WL2_Traits/WL2_Fruits_Y1.csv")
winter_surv <- read_csv("../output/WL2_Traits/WL2_WinterSurv.csv")
wl2_surv_to_rep_y2 <- read_csv("../output/WL2_Traits/WL2_Surv_to_Rep_Y2.csv")
wl2_fruits_y2 <- read_csv("../output/WL2_Traits/WL2_Fruits_Y2.csv")
```

## Davis - No Year 2

p(Establishment)*p(Surv to Rep - y1)*Fruits(y1)

```{r}
#per individual:
ucd_total_fitness <- left_join(ucd_establishment, ucd_surv_to_rep) %>% 
  left_join(ucd_fruits) %>% 
  select(block:Wtr_Year_GD_Historical, Geographic_Dist, Elev_Dist, Establishment, SurvtoRep_Y1:fruits) %>% 
  mutate(SurvtoRep_Y1=if_else(is.na(SurvtoRep_Y1), 0, SurvtoRep_Y1),
         fruits=if_else(is.na(fruits), 0, fruits)) %>% 
  mutate(Total_Fitness=Establishment*SurvtoRep_Y1*fruits)

ucd_total_fitness %>% group_by(pop) %>% summarise(n=n()) %>% arrange(n) #only 2 WV plants 
ucd_total_fitness %>% 
  filter(Total_Fitness>0) %>% 
  group_by(pop) %>% 
  summarise(n=n()) %>% 
  arrange(n) #only TM2 and BH have greater than 1 indiv

unique(ucd_total_fitness$block)
unique(ucd_total_fitness$pop)
unique(ucd_total_fitness$mf)
```

```{r}
ucd_total_fitness_sub_dist <- ucd_total_fitness %>% 
  select(block:rep, Establishment:Total_Fitness) %>% 
  left_join(ucd_sub_dist)

unique(ucd_total_fitness_sub_dist$block)
unique(ucd_total_fitness_sub_dist$pop)
unique(ucd_total_fitness_sub_dist$mf)
```

### Bar Plots

```{r}
ucd_total_fitness %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(Total_Fitness, na.rm = TRUE), semFruits=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/UCD_Traits/UCD_Total_Fitness_GrwSsn_GD_Recent.png", width = 12, height = 8, units = "in")

ucd_total_fitness %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(Total_Fitness, na.rm = TRUE), semFruits=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/UCD_Traits/UCD_Total_Fitness_Wtr_Year_GD_Recent.png", width = 12, height = 8, units = "in")

```

### Scatterplots

```{r}
#scatter plots
GSCD <- ucd_total_fitness %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  pivot_longer(cols = starts_with("GrwSsn"), names_to = "TimePd", values_to = "GrwSsn_CD") %>% 
  mutate(TimePd=str_replace(TimePd, "GrwSsn_GD_", "")) %>% 
  ggplot(aes(x=GrwSsn_CD, y=meanEst, color=TimePd, group = pop)) +
  geom_point(size=6, alpha=0.7) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Growth Season CD", color="Time Period") +
  scale_color_manual(values=timepd_palette) +
  theme(text=element_text(size=30))

WYCD <- ucd_total_fitness %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>%
  pivot_longer(cols = starts_with("Wtr_Year"), names_to = "TimePd", values_to = "Wtr_Year_CD") %>% 
  mutate(TimePd=str_replace(TimePd, "Wtr_Year_GD_", "")) %>% 
  ggplot(aes(x=Wtr_Year_CD, y=meanEst, color=TimePd, group = pop)) +
  geom_point(size=6, alpha=0.7) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Water Year CD", color="Time Period") +
  scale_color_manual(values=timepd_palette) +
  theme(text=element_text(size=30))

GD <- ucd_total_fitness %>% 
  group_by(pop, elev_m, Geographic_Dist) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))


ED <- ucd_total_fitness %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

ucd_total_fitness_FIG <- ggarrange(GSCD, WYCD, GD, ED, ncol=2, nrow=2) 
ggsave("../output/UCD_Traits/UCD_Total_Fitness_SCATTERS.png", width = 24, height = 18, units = "in")
```

## WL2

\*Used average Gowers

p(Establishment)*p(Surv to Rep - y1)*Fruits(y1) + p(Winter Surv)*p(Surv to Rep - y2)*Fruits(y2)

```{r}
wl2_fruits_y1_prep <- wl2_fruits_y1 %>% 
  select(block:BedLoc, Genotype:rep, elevation.group:Long, y1_flowers=flowers, y1_fruits=fruits)

winter_surv_prep <- winter_surv %>% select(block:BedLoc, Genotype:Long, WinterSurv) 
wl2_surv_to_rep_y2_prep <- wl2_surv_to_rep_y2 %>% select(pop:Long, SurvtoRep_y2) 
wl2_fruits_y2_prep <- wl2_fruits_y2 %>% 
  select(pop:Genotype, elevation.group:Long, y2_flowers=flowers, y2_fruits=fruits)

wl2_total_fitness <- left_join(wl2_establishment, wl2_surv_to_rep_y1) %>% 
  left_join(wl2_fruits_y1_prep) %>% 
  left_join(winter_surv_prep) %>% 
  left_join(wl2_surv_to_rep_y2_prep) %>% 
  left_join(wl2_fruits_y2_prep) %>% 
  left_join(wl2_gowers_avg) %>% 
  select(block:Wtr_Year_GD_Historical, Geographic_Dist, Elev_Dist, Establishment, SurvtoRep_Y1:y2_fruits) %>% 
  mutate(SurvtoRep_Y1=if_else(is.na(SurvtoRep_Y1), 0, SurvtoRep_Y1),
         y1_fruits=if_else(is.na(y1_fruits), 0, y1_fruits),
         y2_fruits=if_else(is.na(y2_fruits), 0, y2_fruits),
         WinterSurv=if_else(is.na(WinterSurv), 0, WinterSurv),
         SurvtoRep_y2=if_else(is.na(SurvtoRep_y2), 0, SurvtoRep_y2)) %>% 
  mutate(Total_Fitness=(Establishment*SurvtoRep_Y1*y1_fruits) + (WinterSurv*SurvtoRep_y2*y2_fruits))

wl2_total_fitness %>% group_by(pop) %>% summarise(n=n()) %>% arrange(n) 
wl2_total_fitness %>% 
  filter(Total_Fitness>0) %>% 
  group_by(pop) %>% 
  summarise(n=n()) %>% 
  arrange(n) #SQ1 and WR only have 1 indiv 

unique(wl2_total_fitness$block)
unique(wl2_total_fitness$pop)
unique(wl2_total_fitness$mf)
```

```{r}
wl2_total_fitness_sub_dist <- wl2_total_fitness %>% 
  select(block:rep, Establishment:Total_Fitness) %>% 
  left_join(wl2_sub_dist_avg)

unique(wl2_total_fitness_sub_dist$block)
unique(wl2_total_fitness_sub_dist$pop)
unique(wl2_total_fitness_sub_dist$mf)
```

### Bar Plots

```{r}
wl2_total_fitness %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(Total_Fitness, na.rm = TRUE), semFruits=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),
                width=.2, position = position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/WL2_Traits/WL2_Total_Fitness_GrwSsn_GD_Recent.png", width = 12, height = 8, units = "in")

wl2_total_fitness %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(Total_Fitness, na.rm = TRUE), semFruits=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/WL2_Traits/WL2_Total_Fitness_Wtr_Year_GD_Recent.png", width = 12, height = 8, units = "in")
```

### Scatterplots

```{r}
#scatter plots
GSCD <- wl2_total_fitness %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  pivot_longer(cols = starts_with("GrwSsn"), names_to = "TimePd", values_to = "GrwSsn_CD") %>% 
  mutate(TimePd=str_replace(TimePd, "GrwSsn_GD_", "")) %>% 
  ggplot(aes(x=GrwSsn_CD, y=meanEst, color=TimePd, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Growth Season CD", color="Growth Season \n Climate Distance") +
  scale_color_manual(values=timepd_palette) +
  theme(text=element_text(size=30))

WYCD <- wl2_total_fitness %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  pivot_longer(cols = starts_with("Wtr_Year"), names_to = "TimePd", values_to = "Wtr_Year_CD") %>% 
  mutate(TimePd=str_replace(TimePd, "Wtr_Year_GD_", "")) %>% 
  ggplot(aes(x=Wtr_Year_CD, y=meanEst, color=TimePd, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Water Year CD", color="Water Year \n Climate Distance") +
  scale_color_manual(values=timepd_palette) +
  theme(text=element_text(size=30))

GD <- wl2_total_fitness %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED <- wl2_total_fitness %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

WL2_total_fitness_FIG <- ggarrange(GSCD, WYCD, GD, ED, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_Total_Fitness_SCATTERS.png", width = 24, height = 18, units = "in")
```

## Stats

### Check Distributions

```{r}
wl2_total_fitness %>% 
  ggplot(aes(x=Total_Fitness)) +
  geom_histogram()

ucd_total_fitness %>% 
  ggplot(aes(x=Total_Fitness)) +
  geom_histogram()
```

### Transformations and Scaling

```{r}
wl2_total_fitness_scaled <- wl2_total_fitness %>% 
  mutate_at(c("GrwSsn_GD_Recent","Wtr_Year_GD_Recent",                                                           "GrwSsn_GD_Historical","Wtr_Year_GD_Historical",
              "Geographic_Dist"),scale) %>% 
  mutate(logTotalFitness=log(Total_Fitness+1),
         log10TotalFitness=log10(Total_Fitness+1)) #log transformation, add 1 for 0s 

wl2_total_fitness_scaled %>% #didn't help much 
  ggplot(aes(x=logTotalFitness)) +
  geom_histogram()

wl2_total_fitness_scaled %>% #didn't help much 
  ggplot(aes(x=log10TotalFitness)) +
  geom_histogram()

ucd_total_fitness_scaled <- ucd_total_fitness %>% 
  mutate_at(c("GrwSsn_GD_Recent","Wtr_Year_GD_Recent",
              "GrwSsn_GD_Historical","Wtr_Year_GD_Historical",
              "Geographic_Dist"), scale) %>% 
  mutate(logTotalFitness=log(Total_Fitness+1),
         log10TotalFitness=log10(Total_Fitness+1)) #log transformation, add 1 for 0s 

ucd_total_fitness_scaled %>% #didn't help much 
  ggplot(aes(x=logTotalFitness)) +
  geom_histogram()

ucd_total_fitness_scaled %>% #didn't help much 
  ggplot(aes(x=log10TotalFitness)) +
  geom_histogram()
```

```{r}
wl2_total_fitness_scaled_sub <- wl2_total_fitness_sub_dist %>% 
  mutate_at(c("Wtr_Year_TempDist_Recent",  "Wtr_Year_PPTDist_Recent", 
                 "Wtr_Year_TempDist_Historic", "Wtr_Year_PPTDist_Historic",
                 "GrwSsn_TempDist_Recent", "GrwSsn_PPTDist_Recent",
                 "GrwSsn_TempDist_Historic", "GrwSsn_PPTDist_Historic",
              "Geographic_Dist"), scale) %>% 
  mutate(logTotalFitness=log(Total_Fitness+1),
         log10TotalFitness=log10(Total_Fitness+1)) #log transformation, add 1 for 0s 

ucd_total_fitness_scaled_sub <- ucd_total_fitness_sub_dist %>% 
  mutate_at(c("Wtr_Year_TempDist_Recent",  "Wtr_Year_PPTDist_Recent", 
                 "Wtr_Year_TempDist_Historic", "Wtr_Year_PPTDist_Historic",
                 "GrwSsn_TempDist_Recent", "GrwSsn_PPTDist_Recent",
                 "GrwSsn_TempDist_Historic", "GrwSsn_PPTDist_Historic",
              "Geographic_Dist"), scale) %>% 
  mutate(logTotalFitness=log(Total_Fitness+1),
         log10TotalFitness=log10(Total_Fitness+1)) #log transformation, add 1 for 0s 
```

Split total fitness into probability of fitness and total reproductive output since it is so zero inflated.

### Prob of total fitness \> 0

#### Calculations

```{r}
#wl2_total_fitness_scaled %>% filter(is.na(Total_Fitness)) #no NAs
#ucd_total_fitness_scaled  %>% filter(is.na(Total_Fitness)) #no NAs

wl2_prob_fitness <- wl2_total_fitness_scaled %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1))

wl2_estab_means <- wl2_prob_fitness %>% # summary for plotting
  group_by(pop) %>%
  summarize(Survival = mean(ProbFitness), n=n()) %>% 
  arrange(n)
wl2_estab_means
#wl2_prob_fitness %>% group_by(pop, mf, ProbFitness) %>% summarise(n())

wl2_prob_fitness_to_save <- wl2_total_fitness %>% 
   mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  select(block:Elev_Dist, ProbFitness)
#write_csv(wl2_prob_fitness_to_save, "../output/WL2_Traits/WL2_ProbFitness.csv")
```

```{r}
wl2_prob_fitness_sub <- wl2_total_fitness_scaled_sub %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1))
```

```{r}
ucd_prob_fitness <- ucd_total_fitness_scaled %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1))

ucd_estab_means <- ucd_prob_fitness %>% # summary for plotting
  group_by(pop) %>% #only 2 indivs for WV -remove
  summarize(Survival = mean(ProbFitness),  n=n()) %>% 
  arrange(n)
ucd_estab_means
#ucd_prob_fitness %>% group_by(pop, mf, ProbFitness) %>% summarise(n())

ucd_prob_fitness_to_save <- ucd_total_fitness %>% 
  filter(SurvtoRep_Y1==1) %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFruits=if_else(Total_Fitness==0, 0, 1))
#write_csv(ucd_prob_fitness_to_save, "../output/UCD_Traits/ProbFruits_UCD.csv")
```

```{r}
ucd_prob_fitness_sub <- ucd_total_fitness_scaled_sub %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1))
```

#### Plots

```{r}
wl2_total_fitness_sub_dist %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  filter(!is.na(meanEst)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanEst), y=meanEst, fill=GrwSsn_TempDist_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Parent Population", fill="Temp Dist") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))
```

```{r}
#scatter plots - recent
GSCD_prob_recent <- wl2_total_fitness %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Recent Growth Season CD") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- wl2_total_fitness %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Recent Water Year CD") +
  theme(text=element_text(size=30))

GD_prob <- wl2_total_fitness %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- wl2_total_fitness %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

WL2_total_fitness_FIG_prob_recent_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_ProbFitness_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- wl2_total_fitness %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Historical, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Historic Growth Season CD") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- wl2_total_fitness %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Historical, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Historic Water Year CD") +
  theme(text=element_text(size=30))

WL2_total_fitness_FIG_prob_historic_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_ProbFitness_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
GSCD_prob_recent <- ucd_total_fitness %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Recent Growth Season CD") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- ucd_total_fitness %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Recent Water Year CD") +
  theme(text=element_text(size=30))

GD_prob <- ucd_total_fitness %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- ucd_total_fitness %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

UCD_total_fitness_FIG_prob_recent_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/UCD_Traits/UCD_ProbFitness_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- ucd_total_fitness %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Historical, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Historic Growth Season CD") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- ucd_total_fitness %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Historical, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Historic Water Year CD") +
  theme(text=element_text(size=30))

UCD_total_fitness_FIG_prob_historic_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/UCD_Traits/UCD_ProbFitness_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
GSCD_prob_recent <- wl2_total_fitness_sub_dist %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Recent Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- wl2_total_fitness_sub_dist %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Recent Water Year Temp Dist") +
  theme(text=element_text(size=30))

GD_prob <- wl2_total_fitness_sub_dist %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, Wtr_Year_TempDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- wl2_total_fitness_sub_dist %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

WL2_total_fitness_sub_FIG_prob_recent_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_ProbFitness_TmpSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- wl2_total_fitness_sub_dist %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Historic Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- wl2_total_fitness_sub_dist %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Historic Water Year Temp Dist") +
  theme(text=element_text(size=30))

WL2_total_fitness_sub_FIG_prob_historic_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_ProbFitness_TmpSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
GSCD_prob_recent <- ucd_total_fitness_sub_dist %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Recent Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- ucd_total_fitness_sub_dist %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Recent Water Year Temp Dist") +
  theme(text=element_text(size=30))

GD_prob <- ucd_total_fitness_sub_dist %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, Wtr_Year_TempDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- ucd_total_fitness_sub_dist %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

UCD_total_fitness_sub_FIG_prob_recent_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/UCD_Traits/UCD_ProbFitness_TmpSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- ucd_total_fitness_sub_dist %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Historic Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- ucd_total_fitness_sub_dist %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Historic Water Year Temp Dist") +
  theme(text=element_text(size=30))

UCD_total_fitness_sub_FIG_prob_historic_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/UCD_Traits/UCD_ProbFitness_TmpSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
GSCD_prob_recent <- wl2_total_fitness_sub_dist %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=0.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Recent Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- wl2_total_fitness_sub_dist %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Recent Water Year PPT Dist") +
  theme(text=element_text(size=30))

GD_prob <- wl2_total_fitness_sub_dist %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, Wtr_Year_PPTDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- wl2_total_fitness_sub_dist %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

WL2_total_fitness_sub_FIG_prob_recent_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_ProbFitness_PPTSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- wl2_total_fitness_sub_dist %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Historic Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- wl2_total_fitness_sub_dist %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Historic Water Year PPT Dist") +
  theme(text=element_text(size=30))

WL2_total_fitness_sub_FIG_prob_historic_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_ProbFitness_PPTSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
GSCD_prob_recent <- ucd_total_fitness_sub_dist %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Recent Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- ucd_total_fitness_sub_dist %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Recent Water Year PPT Dist") +
  theme(text=element_text(size=30))

GD_prob <- ucd_total_fitness_sub_dist %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, Wtr_Year_PPTDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- ucd_total_fitness_sub_dist %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

UCD_total_fitness_sub_FIG_prob_recent_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/UCD_Traits/UCD_ProbFitness_PPTSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- ucd_total_fitness_sub_dist %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Historic Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- ucd_total_fitness_sub_dist %>% 
  filter(pop!="WV") %>% #ONLY 2 INDIVS 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of Successfully Reproducing", x="Historic Water Year PPT Dist") +
  theme(text=element_text(size=30))

UCD_total_fitness_sub_FIG_prob_historic_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/UCD_Traits/UCD_ProbFitness_PPTSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

#### Figure for paper

```{r}
GD_prob <- wl2_total_fitness_sub_dist %>%  
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Geographic_Dist) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of \n Successfully Reproducing", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

WYCD_prob_recent <- wl2_total_fitness %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of \n Successfully Reproducing", x="Recent Water Year Climate Dist") +
  theme(text=element_text(size=30))

WYtemp_prob_recent <- wl2_total_fitness_sub_dist %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of \n Successfully Reproducing", x="Recent Water Year Temp Dist (C)") +
  theme(text=element_text(size=30))

WYppt_prob_recent <- wl2_total_fitness_sub_dist %>% 
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(ProbFitness, na.rm = TRUE), semEst=sem(ProbFitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Probability of \n Successfully Reproducing", x="Recent Water Year PPT Dist (mm)") +
  theme(text=element_text(size=30))

WL2_prob_fitness_for_paper <- ggarrange(GD_prob, WYCD_prob_recent, 
                                        WYtemp_prob_recent, WYppt_prob_recent, 
                                        labels = c("A)", "B)", "C)", "D)"), 
                                        font.label = list(size=30, face = "plain"), 
                                        ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_ProbFitness_SCATTERS_Summary_Recent.png", width = 24, height = 18, units = "in")
```

#### Attempt at calculating mean and standard error in logit scale

Get Block means
```{r}
##block summary
wl2_prob_fit_block_summary <- wl2_total_fitness %>%  
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, block, Geographic_Dist, 
           Wtr_Year_GD_Recent, Wtr_Year_GD_Historical, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanProbBlock=mean(ProbFitness, na.rm = TRUE), n=n())

#double check this mostly matches getting means across blocks 
wl2_prob_fit_block_summary %>% group_by(pop, Geographic_Dist) %>% 
  summarise(meanProb=mean(meanProbBlock, na.rm = TRUE), 
            semProb=sem(meanProbBlock,  na.rm=TRUE))
wl2_total_fitness %>%  
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>%
  group_by(pop, Geographic_Dist) %>% 
  summarise(meanProb=mean(ProbFitness, na.rm = TRUE), 
            semProb=sem(ProbFitness,  na.rm=TRUE))
```

Use Car::loigt and Jenny's inv.logit function 
```{r}
#check back transformation works 
wl2_prob_fit_block_summary %>% 
  mutate(logitBlockMean=car::logit(meanProbBlock, adjust = 0.025)) %>% 
  mutate(BTlogit=inv.logit(logitBlockMean, a=0.025))

#check back transformation of means 
wl2_prob_fit_block_summary %>% 
  mutate(logitBlockMean=car::logit(meanProbBlock, adjust = 0.025)) %>% 
  ungroup() %>% 
  group_by(pop, Geographic_Dist, 
           Wtr_Year_GD_Recent, Wtr_Year_GD_Historical, 
           GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanProb=mean(meanProbBlock, na.rm=TRUE), 
            semProb=sem(meanProbBlock, na.rm=TRUE),
            meanLogit=mean(logitBlockMean, na.rm=TRUE), 
            semLogit=sem(logitBlockMean, na.rm=TRUE)) %>% 
  mutate(Btmean=inv.logit(meanLogit, a=0.025)) #much smaller values than meanProb...

#try without 0s and 1s
wl2_logit_compare2 <- wl2_prob_fit_block_summary %>% 
  filter(meanProbBlock>0, meanProbBlock<1) %>% 
  mutate(logitBlockMean=car::logit(meanProbBlock, adjust = 0.025)) %>% 
  ungroup() %>% 
  group_by(pop, Geographic_Dist, 
           Wtr_Year_GD_Recent, Wtr_Year_GD_Historical, 
           GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanProb=mean(meanProbBlock, na.rm=TRUE), 
            semProb=sem(meanProbBlock, na.rm=TRUE),
            meanLogit=mean(logitBlockMean, na.rm=TRUE), 
            semLogit=sem(logitBlockMean, na.rm=TRUE)) %>% 
  mutate(Btmean=inv.logit(meanLogit, a=0.025)) #much smaller values than meanProb...
```

Make comparison figures 
```{r}
wl2_logit_compare <- wl2_prob_fit_block_summary %>% 
  mutate(logitBlockMean=car::logit(meanProbBlock, adjust = 0.025)) %>% 
  ungroup() %>% 
  group_by(pop, Geographic_Dist, 
           Wtr_Year_GD_Recent, Wtr_Year_GD_Historical, 
           GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanProb=mean(meanProbBlock, na.rm=TRUE), 
            semProb=sem(meanProbBlock, na.rm=TRUE),
            meanLogit=mean(logitBlockMean, na.rm=TRUE), 
            semLogit=sem(logitBlockMean, na.rm=TRUE)) %>% 
  mutate(Btmean=inv.logit(meanLogit, a=0.025)) %>% 
  mutate(HighError=meanLogit+semLogit, LowError=meanLogit-semLogit) %>% 
  mutate(Btmean=inv.logit(meanLogit, a=0.025),
         BtHigh=inv.logit(HighError, a=0.025),
         BtLow=inv.logit(LowError, a=0.025))

wl2_logit_compare %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=Btmean, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=BtLow,ymax=BtHigh), width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  ggtitle("Back-Transformed from Logit Scale")

wl2_logit_compare %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanProb, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanProb-semProb,ymax=meanProb+semProb), width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  ggtitle("Raw Means")
```

Make comparison figures - With 0s removed 
```{r}
wl2_logit_compare2 <- wl2_prob_fit_block_summary %>% 
  filter(meanProbBlock>0, meanProbBlock<1) %>% 
  mutate(logitBlockMean=car::logit(meanProbBlock, adjust = 0.025)) %>% 
  ungroup() %>% 
  group_by(pop, Geographic_Dist, 
           Wtr_Year_GD_Recent, Wtr_Year_GD_Historical, 
           GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanProb=mean(meanProbBlock, na.rm=TRUE), 
            semProb=sem(meanProbBlock, na.rm=TRUE),
            meanLogit=mean(logitBlockMean, na.rm=TRUE), 
            semLogit=sem(logitBlockMean, na.rm=TRUE)) %>% 
  mutate(Btmean=inv.logit(meanLogit, a=0.025)) %>% 
  mutate(HighError=meanLogit+semLogit, LowError=meanLogit-semLogit) %>% 
  mutate(Btmean=inv.logit(meanLogit, a=0.025),
         BtHigh=inv.logit(HighError, a=0.025),
         BtLow=inv.logit(LowError, a=0.025))

wl2_logit_compare2 %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=Btmean, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=BtLow,ymax=BtHigh), width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  ggtitle("Back-Transformed from Logit Scale - No 0s and 1s")

wl2_logit_compare2 %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanProb, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanProb-semProb,ymax=meanProb+semProb), width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  ggtitle("Raw Means - No 0s and 1s")
```

Try boot::logit and boot::inv.logit 
```{r}
#double check that back transformation works 
wl2_prob_fit_block_summary %>% 
  mutate(logitBlockMean=boot::logit(meanProbBlock + 0.001)) %>% 
  mutate(btProb=boot::inv.logit(logitBlockMean)-0.001)

#means are wrong b/c of 0 conversions 
boot::inv.logit((-0.911395862 +  -6.906754779)/2)-0.001 ##logit =  0.01866459
(2.857143e-01 + -4.336809e-19)/2 #non-logit = 0.1428571
boot::inv.logit((-0.911395862 +  0)/2)-0.001 #convert logit zeros = 0.3870069 - also wrong ...
#likely b/c when I'm taking the mean, the +0.001 is being manipulated??
#((x - lb) / (ub - lb))
```

```{r}
##logit scale
wl2_prob_fit_block_summary %>% 
  mutate(logitBlockMean=boot::logit(meanProbBlock + 0.001)) %>% 
  group_by(pop, Geographic_Dist) %>% 
  summarise(meanLogit=mean(logitBlockMean), 
            semLogit=sem(logitBlockMean)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanLogit, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanLogit+semLogit,ymax=meanLogit-semLogit),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  ggtitle("Logit Scale")

##back transformed from logit 
wl2_prob_fit_block_summary %>% 
  mutate(logitBlockMean=boot::logit(meanProbBlock + 0.001)) %>% 
  group_by(pop, Geographic_Dist) %>% 
  summarise(meanLogit=mean(logitBlockMean), 
            semLogit=sem(logitBlockMean)) %>% 
    mutate(HighError=meanLogit+semLogit, LowError=meanLogit-semLogit) %>% 
  mutate(btProb=boot::inv.logit(meanLogit)-0.001, 
         btHighError=boot::inv.logit(HighError)-0.001,
         btLowError=boot::inv.logit(LowError)-0.001) %>% 
  ggplot(aes(x=Geographic_Dist, y=btProb, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=btLowError,ymax=btHighError),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  ggtitle("Back-Transformed from Logit Scale")

##non-logit scale 
wl2_total_fitness_sub_dist %>%  
  mutate(ProbFitness=if_else(Total_Fitness==0, 0, 1)) %>% 
  group_by(pop, block, Geographic_Dist) %>% 
  summarise(meanProbBlock=mean(ProbFitness, na.rm = TRUE) + 0.001, n=n()) %>% 
  group_by(pop, Geographic_Dist) %>% 
  summarise(meanProb=mean(meanProbBlock, na.rm = TRUE), 
            semProb=sem(meanProbBlock,  na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanProb, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanProb-semProb,ymax=meanProb+semProb),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  ggtitle("Raw Means")
```

#### Basic Model Workflow

```{r}
glmer.model_binomial <- 
  linear_reg() %>% 
  set_engine("glmer", family=binomial)

prob_fitness_wflow <- workflow() %>% 
  add_variables(outcomes = ProbFitness, predictors = c(pop, mf, block))

prob_fitness_fits <- tibble(wflow=list(
  pop = {prob_fitness_wflow %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ (1|pop))},
  
  pop.mf = {prob_fitness_wflow %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ (1|pop/mf))},
  
  pop.block = {prob_fitness_wflow %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ (1|pop) + (1|block))},
  
  pop.mf.block = {prob_fitness_wflow %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ (1|pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)

prob_fitness_fits_wl2 <- prob_fitness_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_prob_fitness))
#mod_test <- glmer(ProbFitness ~ (1|pop/mf), data=wl2_prob_fitness, family=binomial)
#Warning: Model failed to converge with max|grad| = 0.0846977 (tol = 0.002, component 1)Warning: Model is nearly unidentifiable: very large eigenvalue
 #- Rescale variables? But can't scale character variables, maybe something wrong with mf sample sizes 
#summary(mod_test)

prob_fitness_fits_ucd <- prob_fitness_fits %>%
  mutate(fit = map(wflow, fit, data = ucd_prob_fitness))

prob_fitness_fits_wl2 %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#model with pop and block (but excluding mf) is best by AIC and BIC

prob_fitness_fits_ucd %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#model with pop and mf (but excluding block) preferred by AIC, model with just pop preferred by BIC 
```

#### VISUALIZE

```{r, eval=FALSE}
prob_fitness_nd_wl2 <- data_grid(wl2_prob_fitness, pop)

prob_fitness_predict_fun <- function(m, nd = prob_fitness_nd_wl2) {
  predict(m, 
          newdata = nd,
          type = "response",
          re.form = ~(1|pop)
  ) 
}

# 100 seconds
system.time (prob_fitness_fits_wl2 <- prob_fitness_fits_wl2 %>%
               mutate(
                 bootstrap = map(fit, \(x) {
                   extract_fit_engine(x) %>% 
                     bootMer(FUN = prob_fitness_predict_fun, 
                             re.form = ~ (1|pop),
                             nsim = 1000,
                             parallel = "multicore",
                             ncpus = 7
                     ) } #bootMer
                 ),  
                 tidyboot = map(bootstrap, tidy, conf.int = TRUE),
                 tidyboot = map(tidyboot, \(x) cbind(prob_fitness_nd_wl2, x))
               ) # mutate
) # system.time

prob_fitness_fits_wl2 %>%
  select(name, tidyboot) %>%
  unnest(tidyboot) %>%
  ggplot(aes(x=pop)) +
  geom_col(aes(y=Survival), data = wl2_estab_means, fill="skyblue") +
  geom_pointrange(aes(y=statistic, ymax = conf.high, ymin = conf.low, color=name), position = position_dodge(width = .75)) +
  ylab("Prob Fitness >0") +
  scale_color_viridis_d() +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=.5))
#the model with pop and block seems to be underestimating the means whereas the model with just pop seems to match the data better...

#wl2_estab_means %>% 
#  ggplot(aes(x=pop, y=Survival)) +
#  geom_col()

#prob_fitness_fits_wl2 %>%
#  select(name, bootstrap) %>%
#  mutate(tidyboot = map(bootstrap, tidy, conf.int = TRUE)) %>%
#  unnest(tidyboot)
```

```{r, eval=FALSE}
prob_fitness_nd_ucd <- data_grid(ucd_prob_fitness, pop)

prob_fitness_predict_fun <- function(m, nd = prob_fitness_nd_ucd) {
  predict(m, 
          newdata = nd,
          type = "response",
          re.form = ~(1|pop)
  )
}

# 100 seconds
system.time (prob_fitness_fits_ucd <- prob_fitness_fits_ucd %>%  
               mutate(
                 bootstrap = map(fit, \(x) {
                   extract_fit_engine(x) %>% 
                     bootMer(FUN = prob_fitness_predict_fun, 
                             re.form = ~ (1|pop),
                             nsim = 1000,
                             parallel = "multicore",
                             ncpus = 7
                     ) } #bootMer
                 ),  
                 tidyboot = map(bootstrap, tidy, conf.int = TRUE),
                 tidyboot = map(tidyboot, \(x) cbind(prob_fitness_nd_ucd, x))
               ) # mutate
) # system.time

prob_fitness_fits_ucd %>%
  select(name, tidyboot) %>%
  unnest(tidyboot) %>%
  ggplot(aes(x=pop)) +
  geom_col(aes(y=Survival), data = ucd_estab_means, fill="skyblue") +
  geom_pointrange(aes(y=statistic, ymax = conf.high, ymin = conf.low, color=name), position = position_dodge(width = .75)) +
   ylab("Prob Fitness >0") +
  scale_color_viridis_d() +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=.5))
#pop.mf seems to be underestimating the means, model with just pop seems to match the data better...
```

#### Test climate and geographic distance

```{r}
prob_fitness_GD_wflow_wl2 <- workflow() %>%
  add_variables(outcomes = ProbFitness, predictors = c(pop, block, contains("GD"), Geographic_Dist)) 

prob_fitness_GD_fits_wl2 <- tibble(wflow=list(
  pop.block = {prob_fitness_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ (1|pop) + (1|block))},
  
  GS_Recent = {prob_fitness_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ GrwSsn_GD_Recent + Geographic_Dist + (1|pop) + (1|block))},
  
  GS_Historical = {prob_fitness_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ GrwSsn_GD_Historical + Geographic_Dist + (1|pop) + (1|block))},
  
  WY_Recent = {prob_fitness_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ Wtr_Year_GD_Recent + Geographic_Dist + (1|pop) + (1|block))},
  
  WY_Historical = {prob_fitness_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ Wtr_Year_GD_Historical + Geographic_Dist + (1|pop) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = wl2_prob_fitness))

prob_fitness_GD_fits_wl2 %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#water year models preferred by AIC 

prob_fitness_GD_fits_wl2 %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "GD") | term=="Geographic_Dist") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)
# recent and historical water year CD are sig, geo dist is not sig in those models. The GS_RECENT model had convergence issues.

mod_test <- glmer(ProbFitness ~ GrwSsn_GD_Recent + Geographic_Dist + (1|pop) + (1|block), data=wl2_prob_fitness, family=binomial)
#summary(mod_test)
#Warning: Model failed to converge with max|grad| = 0.0747217 (tol = 0.002, component 1)Warning: Model is nearly unidentifiable: very large eigenvalue
# - Rescale variables? - But the variables were already scaled 
```

```{r}
prob_fitness_SUB_wflow_wl2 <- workflow() %>%
  add_variables(outcomes = ProbFitness, predictors = c(pop, block, contains("Dist"))) 

prob_fitness_SUB_fits_wl2 <- tibble(wflow=list(
  pop.block = {prob_fitness_SUB_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ (1|pop) + (1|block))},
  
  GS_Recent = {prob_fitness_SUB_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ GrwSsn_TempDist_Recent + GrwSsn_PPTDist_Recent+ Geographic_Dist + (1|pop) + (1|block))},
  
  GS_Historical = {prob_fitness_SUB_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ GrwSsn_TempDist_Historic + GrwSsn_PPTDist_Historic + Geographic_Dist + (1|pop) + (1|block))},
  
  WY_Recent = {prob_fitness_SUB_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ Wtr_Year_TempDist_Recent + Wtr_Year_PPTDist_Recent + Geographic_Dist + (1|pop) + (1|block))},
  
  WY_Historical = {prob_fitness_SUB_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ Wtr_Year_TempDist_Historic + Wtr_Year_PPTDist_Historic + Geographic_Dist + (1|pop) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = wl2_prob_fitness_sub))

prob_fitness_SUB_fits_wl2 %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)

prob_fitness_SUB_fits_wl2 %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "Dist") | term=="Geographic_Dist") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)

mod_test <- glmer(ProbFitness ~ GrwSsn_TempDist_Recent + GrwSsn_PPTDist_Recent+ Geographic_Dist + (1|pop) + (1|block), data=wl2_prob_fitness_sub, family=binomial)
summary(mod_test)
```

Model comparison

```{r}
m1 <-  glmer(ProbFitness ~ Wtr_Year_TempDist_Recent + GrwSsn_PPTDist_Recent + Geographic_Dist + (1|pop) + (1|block), data=wl2_prob_fitness_sub, family=binomial)
m2 <- glmer(ProbFitness ~ Wtr_Year_TempDist_Recent + Geographic_Dist + (1|pop) + (1|block), data=wl2_prob_fitness_sub, family=binomial)
m3 <- glmer(ProbFitness ~  GrwSsn_PPTDist_Recent + Geographic_Dist + (1|pop) + (1|block), data=wl2_prob_fitness_sub, family=binomial)
anova(m1, m2, m3)
```

Keep block in for UCD

```{r}
prob_fitness_GD_wflow_ucd <- workflow() %>%
  add_variables(outcomes = ProbFitness, predictors = c(pop, mf, block, contains("GD"), Geographic_Dist)) 

prob_fitness_GD_fits_ucd <- tibble(wflow=list(
  pop.mf.block = {prob_fitness_GD_wflow_ucd %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ (1|pop/mf) + (1|block))},
  
  GS_Recent = {prob_fitness_GD_wflow_ucd %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ GrwSsn_GD_Recent + Geographic_Dist + (1|pop/mf) + (1|block))},
  
  GS_Historical = {prob_fitness_GD_wflow_ucd %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ GrwSsn_GD_Historical + Geographic_Dist + (1|pop/mf) + (1|block))},
  
  WY_Recent = {prob_fitness_GD_wflow_ucd %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ Wtr_Year_GD_Recent + Geographic_Dist + (1|pop/mf) + (1|block))},
  
  WY_Historical = {prob_fitness_GD_wflow_ucd %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ Wtr_Year_GD_Historical + Geographic_Dist + (1|pop/mf) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = ucd_prob_fitness))

prob_fitness_GD_fits_ucd %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#pop.mf.block preferred by AIC & BIC, WY models are close after

prob_fitness_GD_fits_ucd %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "GD") | term=="Geographic_Dist") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)

mod_test <- glmer(ProbFitness ~ Wtr_Year_GD_Historical + Geographic_Dist + (1|pop/mf) + (1|block), data=ucd_prob_fitness, family=binomial)
#model failed to converge for growth season and water year recent 
```

```{r}
prob_fitness_SUB_wflow_ucd <- workflow() %>%
  add_variables(outcomes = ProbFitness, predictors = c(pop, mf, block, contains("Dist"))) 

prob_fitness_SUB_fits_ucd <- tibble(wflow=list(
  pop.block = {prob_fitness_SUB_wflow_ucd %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ (1|pop/mf) + (1|block))},
  
  GS_Recent = {prob_fitness_SUB_wflow_ucd %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ GrwSsn_TempDist_Recent + GrwSsn_PPTDist_Recent+ Geographic_Dist + (1|pop/mf) + (1|block))},
  
  GS_Historical = {prob_fitness_SUB_wflow_ucd %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ GrwSsn_TempDist_Historic + GrwSsn_PPTDist_Historic + Geographic_Dist + (1|pop/mf) + (1|block))},
  
  WY_Recent = {prob_fitness_SUB_wflow_ucd %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ Wtr_Year_TempDist_Recent + Wtr_Year_PPTDist_Recent + Geographic_Dist + (1|pop/mf) + (1|block))},
  
  WY_Historical = {prob_fitness_SUB_wflow_ucd %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ Wtr_Year_TempDist_Historic + Wtr_Year_PPTDist_Historic + Geographic_Dist + (1|pop/mf) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = ucd_prob_fitness_sub))

prob_fitness_SUB_fits_ucd %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)

prob_fitness_SUB_fits_ucd %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "Dist") | term=="Geographic_Dist") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)

mod_test <- glmer(ProbFitness ~ Wtr_Year_TempDist_Historic + Wtr_Year_PPTDist_Historic + Geographic_Dist + (1|pop/mf) + (1|block), data=ucd_prob_fitness_sub, family=binomial)
#Water year recent and historic models failed to converge 
```

#### Test plasticity in size 
```{r}
size_plasticity <- read_csv("../output/TwoMonths_Size_Slopes.csv") %>% rename(pop=parent.pop)

wl2_prob_fit_plas <- wl2_prob_fitness %>% 
  select(block, Genotype:Elev_Dist, ProbFitness) %>% 
  left_join(size_plasticity) #%>% 
  #mutate_at(c("height_slope","length_slope"),scale) 
head(wl2_prob_fit_plas)
wl2_prob_fit_plas %>% 
  ggplot(aes(length_slope)) +
  geom_histogram()
wl2_prob_fit_plas %>% 
  ggplot(aes(height_slope)) +
  geom_histogram()

ucd_prob_fit_plas <- ucd_prob_fitness %>% 
  select(block, Genotype:Elev_Dist, ProbFitness) %>% 
  left_join(size_plasticity) #%>% 
# mutate_at(c("height_slope","length_slope"),scale) 
head(ucd_prob_fit_plas)
ucd_prob_fit_plas %>% 
  ggplot(aes(length_slope)) +
  geom_histogram()
ucd_prob_fit_plas %>% 
  ggplot(aes(height_slope)) +
  geom_histogram()
```

```{r}
wl2_glmer1 <- glmer(ProbFitness ~ height_slope + (1|pop) + (1|block), data=wl2_prob_fit_plas, family = binomial)
summary(wl2_glmer1) #plasticity in height sig and positive 

wl2_glmer2 <- glmer(ProbFitness ~ length_slope + (1|pop) + (1|block), data=wl2_prob_fit_plas, family = binomial)
summary(wl2_glmer2) #plasticity in length sig and positive 

wl2_glmer3 <- glmer(ProbFitness ~ height_slope + length_slope + (1|pop) + (1|block), data=wl2_prob_fit_plas, family = binomial)
summary(wl2_glmer3) #when both in the model, only height plasticity = sig 
tidy(wl2_glmer3)
```

```{r}
ucd_glmer1 <- glmer(ProbFitness ~ height_slope + (1|pop/mf) + (1|block), data=ucd_prob_fit_plas, family = binomial)
#boundary (singular) fit: see help('isSingular') - block explains 0 var 
summary(ucd_glmer1) #plasticity in height is sig and positive 

ucd_glmer2 <- glmer(ProbFitness ~ length_slope + (1|pop/mf) + (1|block), data=ucd_prob_fit_plas, family = binomial) 
summary(ucd_glmer2) #plasticity in length not sig

ucd_glmer3 <- glmer(ProbFitness ~ height_slope + length_slope + (1|pop/mf) + (1|block), data=ucd_prob_fit_plas, family = binomial)
summary(ucd_glmer3)
tidy(ucd_glmer3)
```

### Total Fitness for plants with any rep

#### Calculations

```{r}
wl2_total_fitness_scaled %>% filter(Total_Fitness > 0) %>% 
  group_by(pop) %>% summarise(n=n()) %>% arrange(n) #7 pops with greater than 1 indiv 

wl2_rep_output <- wl2_total_fitness_scaled %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% #remove these pops b/c only 1 individual in dataset 
  mutate(Total_Fitness_Scaled=round(Total_Fitness/mean(Total_Fitness)),
         logTotalFitness=log(Total_Fitness))
dim(wl2_rep_output) #96 plants 

wl2_rep_output %>% #still skewed
  ggplot(aes(x=Total_Fitness)) +
  geom_histogram()

wl2_rep_output %>% #good enough!
  ggplot(aes(x=logTotalFitness)) +
  geom_histogram()

wl2_rep_output %>% #helped some
  ggplot(aes(x=log10TotalFitness)) +
  geom_histogram()

wl2_rep_output %>% #helped some
  ggplot(aes(x=Total_Fitness_Scaled)) +
  geom_histogram()

wl2_fitness_means <- wl2_rep_output %>% # summary for plotting
  group_by(pop) %>%
  summarize(meanTotalFitness = mean(Total_Fitness), n=n())

wl2_rep_output_to_save <- wl2_total_fitness %>% 
  filter(Total_Fitness > 0) %>% 
  mutate(logTotalFitness=log(Total_Fitness)) %>% 
  select(block:Elev_Dist, Total_Fitness:logTotalFitness)
#write_csv(wl2_rep_output_to_save, "../output/WL2_Traits/WL2_TotalRepOutput.csv")
```

```{r}
wl2_rep_output_sub <- wl2_total_fitness_scaled_sub %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% #remove these pops b/c only 1 individual in dataset 
  mutate(logTotalFitness=log(Total_Fitness),
         log10TotalFitness=log10(Total_Fitness))

wl2_rep_output_sub %>% #still skewed
  ggplot(aes(x=Total_Fitness)) +
  geom_histogram()

wl2_rep_output_sub %>% #good enough!
  ggplot(aes(x=logTotalFitness)) +
  geom_histogram()
```

```{r}
ucd_total_fitness_scaled %>% filter(Total_Fitness > 0) %>% 
  group_by(pop) %>% summarise(n=n()) %>% arrange(n) #only BH and TM2 have greater than 1 individual --> DOESN'T MAKE SENSE TO MODEL 

ucd_rep_output <- ucd_total_fitness_scaled %>% 
  filter(pop=="BH" | pop=="TM2") %>%  #all other pops only have 1 indiv 
  filter(Total_Fitness > 0)
dim(ucd_rep_output) #26 plants 

ucd_rep_output %>% 
  ggplot(aes(x=Total_Fitness)) +
  geom_histogram()

ucd_rep_output %>%  #looks fine 
  ggplot(aes(x=logTotalFitness)) +
  geom_histogram()

ucd_rep_output %>% 
  ggplot(aes(x=log10TotalFitness)) +
  geom_histogram()

ucd_fitness_means <- ucd_rep_output %>% 
  group_by(pop) %>%
  summarize(meanTotalFitness = mean(Total_Fitness), semTotalFitness=sem(Total_Fitness), n=n())
ucd_fitness_means
```

```{r}
ucd_rep_output_sub <- ucd_total_fitness_scaled_sub %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop=="BH" | pop=="TM2") %>%  #all other pops only have 1 indiv 
  mutate(logTotalFitness=log(Total_Fitness),
         log10TotalFitness=log10(Total_Fitness))

ucd_rep_output_sub %>% #still skewed
  ggplot(aes(x=Total_Fitness)) +
  geom_histogram()

ucd_rep_output_sub %>% #okayish
  ggplot(aes(x=logTotalFitness)) +
  geom_histogram()
```

#### Plots

```{r}
#scatter plots - recent
GSCD_repoutput_recent <- wl2_total_fitness %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Recent Growth Season CD", color="Growth Season \n Climate Distance") +
  scale_color_manual(values=timepd_palette) +
  theme(text=element_text(size=30))

WYCD_repoutput_recent <- wl2_total_fitness %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Recent Water Year CD", color="Water Year \n Climate Distance") +
  scale_color_manual(values=timepd_palette) +
  theme(text=element_text(size=30))

GD_repoutput <- wl2_total_fitness %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_repoutput <- wl2_total_fitness %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

WL2_total_fitness_FIG_repoutput_recent_recent <- ggarrange(GSCD_repoutput_recent, WYCD_repoutput_recent, GD_repoutput, ED_repoutput, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_Total_RepOutput_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_repoutput_historic <- wl2_total_fitness %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Historical, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Historic Growth Season CD", color="Growth Season \n Climate Distance") +
  scale_color_manual(values=timepd_palette) +
  theme(text=element_text(size=30))

WYCD_repoutput_historic <- wl2_total_fitness %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Historical, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Historic Water Year CD", color="Water Year \n Climate Distance") +
  scale_color_manual(values=timepd_palette) +
  theme(text=element_text(size=30))

WL2_total_fitness_FIG_repoutput_historic_historic <- ggarrange(GSCD_repoutput_historic, WYCD_repoutput_historic, GD_repoutput, ED_repoutput, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_Total_RepOutput_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
GSCD_repoutput_recent <- ucd_total_fitness %>% 
  filter(pop=="BH" | pop=="TM2") %>%  #all other pops only have 1 indiv 
  filter(Total_Fitness > 0) %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Recent Growth Season CD", color="Growth Season \n Climate Distance") +
  scale_color_manual(values=timepd_palette) +
  theme(text=element_text(size=30))

WYCD_repoutput_recent <- ucd_total_fitness %>% 
  filter(pop=="BH" | pop=="TM2") %>%  #all other pops only have 1 indiv 
  filter(Total_Fitness > 0) %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Recent Water Year CD", color="Water Year \n Climate Distance") +
  scale_color_manual(values=timepd_palette) +
  theme(text=element_text(size=30))

GD_repoutput <- ucd_total_fitness %>% 
  filter(pop=="BH" | pop=="TM2") %>%  #all other pops only have 1 indiv 
  filter(Total_Fitness > 0) %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_repoutput <- ucd_total_fitness %>% 
  filter(pop=="BH" | pop=="TM2") %>%  #all other pops only have 1 indiv 
  filter(Total_Fitness > 0) %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

UCD_total_fitness_FIG_repoutput_recent_recent <- ggarrange(GSCD_repoutput_recent, WYCD_repoutput_recent, GD_repoutput, ED_repoutput, ncol=2, nrow=2) 
#ggsave("../output/UCD_Traits/UCD_Total_RepOutput_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_repoutput_historic <- ucd_total_fitness %>% 
  filter(pop=="BH" | pop=="TM2") %>%  #all other pops only have 1 indiv 
  filter(Total_Fitness > 0) %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, GrwSsn_GD_Historical) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Historical, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Historic Growth Season CD") +
  scale_color_manual(values=timepd_palette) +
  theme(text=element_text(size=30))

WYCD_repoutput_historic <- ucd_total_fitness %>% 
  filter(pop=="BH" | pop=="TM2") %>%  #all other pops only have 1 indiv 
  filter(Total_Fitness > 0) %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Wtr_Year_GD_Historical) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Historical, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Historic Water Year CD") +
  scale_color_manual(values=timepd_palette) +
  theme(text=element_text(size=30))

UCD_total_fitness_FIG_repoutput_historic_historic <- ggarrange(GSCD_repoutput_historic, WYCD_repoutput_historic, GD_repoutput, ED_repoutput, ncol=2, nrow=2) 
ggsave("../output/UCD_Traits/UCD_Total_RepOutput_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
GSCD_repoutput_recent <- wl2_total_fitness_sub_dist %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Recent Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_repoutput_recent <- wl2_total_fitness_sub_dist %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Recent Water Year Temp Dist") +
  theme(text=element_text(size=30))

GD_repoutput <- wl2_total_fitness_sub_dist %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, Wtr_Year_TempDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_repoutput <- wl2_total_fitness_sub_dist %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

WL2_total_fitness_sub_dist_FIG_repoutput_recent_recent <- ggarrange(GSCD_repoutput_recent, WYCD_repoutput_recent, GD_repoutput, ED_repoutput, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_Total_RepOutput_TmpSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_repoutput_historic <- wl2_total_fitness_sub_dist %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Historic Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_repoutput_historic <- wl2_total_fitness_sub_dist %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Historic Water Year Temp Dist") +
  theme(text=element_text(size=30))

WL2_total_fitness_sub_dist_FIG_repoutput_historic_historic <- ggarrange(GSCD_repoutput_historic, WYCD_repoutput_historic, GD_repoutput, ED_repoutput, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_Total_RepOutput_TmpSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
GSCD_repoutput_recent <- wl2_total_fitness_sub_dist %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Recent Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_repoutput_recent <- wl2_total_fitness_sub_dist %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Recent Water Year PPT Dist") +
  theme(text=element_text(size=30))

GD_repoutput <- wl2_total_fitness_sub_dist %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, Wtr_Year_PPTDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_repoutput <- wl2_total_fitness_sub_dist %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

WL2_total_fitness_sub_dist_FIG_repoutput_recent_recent <- ggarrange(GSCD_repoutput_recent, WYCD_repoutput_recent, GD_repoutput, ED_repoutput, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_Total_RepOutput_PPTSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_repoutput_historic <- wl2_total_fitness_sub_dist %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Historic Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_repoutput_historic <- wl2_total_fitness_sub_dist %>% 
  filter(Total_Fitness > 0) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(Total_Fitness, na.rm = TRUE), semEst=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fruit Number", x="Historic Water Year PPT Dist") +
  theme(text=element_text(size=30))

WL2_total_fitness_sub_dist_FIG_repoutput_historic_historic <- ggarrange(GSCD_repoutput_historic, WYCD_repoutput_historic, GD_repoutput, ED_repoutput, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_Total_RepOutput_PPTSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

#### Checking Poisson for Total Fitness

Poisson not appropriate b/c assumes mean = variance...

```{r}
wl2_rep_output %>%
  group_by(pop) %>%
  summarize(mean=mean(Total_Fitness),
            var=var(Total_Fitness))
```

no, overdispersed

histograms

```{r}
wl2_rep_output %>%
  select(pop, Total_Fitness) %>%
  ggplot(aes(x=Total_Fitness)) +
  geom_histogram(bins = 10) +
  facet_wrap(~pop, scales="free_y")
```

#### Basic Model Workflow

```{r}
rep_output_modelslog <- tribble(
  ~name,          ~f,
  "1_pop",              "logTotalFitness ~ (1|pop)", 
  "2_pop.mf",           "logTotalFitness ~  (1|pop/mf)", 
  "3_pop.block",        "logTotalFitness ~ (1|pop) + (1|block)", 
  "4_pop.mf.block",     "logTotalFitness ~  (1|pop/mf) + (1|block)"
)

#run the models 
rep_output_modelslog <- rep_output_modelslog %>%
  mutate(lmer = map(f, ~ lmer(as.formula(.), 
                            data = wl2_rep_output)), #run the models 
         predict = map(lmer, predict), # predicting from original data...
         glance = map(lmer, glance)) #glance at the model results

rep_output_modelslog %>% select(-f, -lmer) %>% unnest(glance) %>% arrange(BIC) #look at the model fitting info 
rep_output_modelslog %>% select(-f, -lmer) %>% unnest(glance) %>% arrange(AIC) #look at the model fitting info 
#model with pop.block best by AIC and BIC 

mod_test <- lmer(logTotalFitness ~  (1|pop) + (1|block), data=wl2_rep_output)
plot(mod_test, which = 1) 
qqnorm(resid(mod_test))
qqline(resid(mod_test)) 
summary(mod_test)
#mf gives "boundary (singular) fit: see help('isSingular')" so take it out 
```

#### Predicted vs. Observed Rep Output

```{r}
wl2_rep_output %>% 
  cbind(predicted={rep_output_modelslog %>% filter(name=="3_pop.block") %>% pull(predict) %>% unlist()}) %>%
  ggplot(aes(x=logTotalFitness, y = predicted)) +
  geom_point(alpha=.2) +
  geom_abline(color="skyblue2") +
  facet_wrap(~pop, scales="free")

```

#### Test climate and geographic distance

```{r}
rep_output_models_log_CD_GD <- tribble(
  ~name,          ~f,
  "1_pop.block",      "logTotalFitness ~  (1|pop) + (1|block)", 
  "2_GS_Recent",      "logTotalFitness ~  GrwSsn_GD_Recent + Geographic_Dist + (1|pop) + (1|block)", 
  "3_GS_Historical",  "logTotalFitness ~  GrwSsn_GD_Historical + Geographic_Dist + (1|pop) + (1|block)", 
  "4_WY_Recent",      "logTotalFitness ~  Wtr_Year_GD_Recent + Geographic_Dist +(1|pop) + (1|block)",
  "5_WY_Historical",  "logTotalFitness ~  Wtr_Year_GD_Historical + Geographic_Dist + (1|pop) + (1|block)"
)

#run the models 
rep_output_models_log_CD_GD <- rep_output_models_log_CD_GD %>%
  mutate(lmer = map(f, ~ lmer(as.formula(.), 
                            data = wl2_rep_output)), #run the models 
         predict = map(lmer, predict), # predicting from original data...
         glance = map(lmer, glance)) #glance at the model results

rep_output_models_log_CD_GD %>% select(-f, -lmer) %>% unnest(glance) %>% arrange(BIC) #look at the model fitting info 
rep_output_models_log_CD_GD %>% select(-f, -lmer) %>% unnest(glance) %>% arrange(AIC) #look at the model fitting info 
#pop block best by AIC and BIC ...

#pop.mf preferred by AIC & BIC, WY models are close after
rep_output_models_log_CD_GD %>% mutate(tidy=map(lmer, tidy)) %>% unnest(tidy) %>%
  select(name, term:p.value) %>% 
  filter(str_detect(term, "GD") | term=="Geographic_Dist") %>%
  drop_na(p.value)
#  arrange(p.value)

#only GrwSSn historical marginally significant p=0.056... no other distances significant 
```

```{r}
rep_output_SUB_models_log_CD_GD <- tribble(
  ~name,          ~f,
  "1_pop.block",      "logTotalFitness ~  (1|pop) + (1|block)", 
  "2_GS_Recent",      "logTotalFitness ~  GrwSsn_TempDist_Recent + GrwSsn_PPTDist_Recent + Geographic_Dist + (1|pop) + (1|block)", 
  "3a_GS_Historical",  "logTotalFitness ~  GrwSsn_TempDist_Historic + GrwSsn_PPTDist_Historic + Geographic_Dist + (1|pop) + (1|block)", 
  "3b_GS_Historical",  "logTotalFitness ~  GrwSsn_TempDist_Historic*GrwSsn_PPTDist_Historic + Geographic_Dist + (1|pop) + (1|block)", 
  "3c_GS_Historical",  "logTotalFitness ~  GrwSsn_TempDist_Historic + Geographic_Dist + (1|pop) + (1|block)", 
  "3d_GS_Historical",  "logTotalFitness ~  GrwSsn_PPTDist_Historic + Geographic_Dist + (1|pop) + (1|block)", 
  "4_WY_Recent",      "logTotalFitness ~  Wtr_Year_TempDist_Recent + Wtr_Year_PPTDist_Recent + Geographic_Dist +(1|pop) + (1|block)",
  "5_WY_Historical",  "logTotalFitness ~  Wtr_Year_TempDist_Historic + Wtr_Year_PPTDist_Historic + Geographic_Dist + (1|pop) + (1|block)"
)

#run the models 
rep_output_SUB_models_log_CD_GD <- rep_output_SUB_models_log_CD_GD %>%
  mutate(lmer = map(f, ~ lmer(as.formula(.), 
                            data = wl2_rep_output_sub)), #run the models 
         predict = map(lmer, predict), # predicting from original data...
         glance = map(lmer, glance)) #glance at the model results

rep_output_SUB_models_log_CD_GD %>% select(-f, -lmer) %>% unnest(glance) %>% arrange(BIC) #look at the model fitting info 
rep_output_SUB_models_log_CD_GD %>% select(-f, -lmer) %>% unnest(glance) %>% arrange(AIC) #look at the model fitting info 
#pop block best by AIC and BIC ...

#pop.mf preferred by AIC & BIC, WY models are close after
rep_output_SUB_models_log_CD_GD %>% mutate(tidy=map(lmer, tidy)) %>% unnest(tidy) %>%
  select(name, term, estimate:p.value) %>% 
  filter(str_detect(term, "Dist")) %>%
  drop_na(p.value)
#  arrange(p.value)

mod_test <- lmer(logTotalFitness ~  GrwSsn_TempDist_Historic + Geographic_Dist + (1|pop) + (1|block), data=wl2_rep_output_sub)
plot(mod_test, which = 1) 
qqnorm(resid(mod_test))
qqline(resid(mod_test)) 
summary(mod_test)
#boundary (singular) fit: see help('isSingular') for GRWSSN historic models a&b - pop explains 0 variance 
```

Troubleshoot weird GrwSsn-Temp_Dist results (positive coefficient, but figure indicates it should be neg...)

```{r}
wl2_rep_output_sub %>% group_by(pop, GrwSsn_TempDist_Historic) %>% summarise(meanFruits=mean(logTotalFitness)) %>% arrange(GrwSsn_TempDist_Historic)

mod_test <- lmer(logTotalFitness ~  GrwSsn_TempDist_Historic + GrwSsn_PPTDist_Historic + Geographic_Dist + (1|block), data=wl2_rep_output_sub)
summary(mod_test)
ranova(mod_test)
anova(mod_test)

wl2_rep_output_sub %>% 
  cbind(predicted={rep_output_SUB_models_log_CD_GD %>% filter(name=="3a_GS_Historical") %>% pull(predict) %>% unlist()}) %>%
  ggplot(aes(x=logTotalFitness, y = predicted)) +
  geom_point(alpha=.2) +
  geom_abline(color="skyblue2") 
#overall, the predictions seem to match the observed data...

wl2_rep_output_sub %>% 
  cbind(predicted={rep_output_SUB_models_log_CD_GD %>% filter(name=="3a_GS_Historical") %>% pull(predict) %>% unlist()}) %>%
  ggplot(aes(x=logTotalFitness, y = predicted)) +
  geom_point(alpha=.2) +
  geom_abline(color="skyblue2") +
  facet_wrap(~block, scales="free")
#some blocks with fewer data points --> weaker predictions 


wl2_rep_output_sub %>% 
  cbind(predicted={rep_output_SUB_models_log_CD_GD %>% filter(name=="3a_GS_Historical") %>% pull(predict) %>% unlist()}) %>%
  ggplot(aes(x=GrwSsn_TempDist_Historic, y = predicted)) +
  geom_point() 
```
