---
title: "Total Fitness"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Total Fitness at both sites 
To Do:
- add in annual census/fruit number for WL2 2024
- add in gowers_distance for WL2 2024 
- add in geographic distance 

Total Fitness: p(Establishment)*p(Surv to Rep - y1)*Fruits(y1) + p(Winter Surv)*p(Surv to Rep - y2)*Fruits(y2)

## Libraries
```{r}
library(tidyverse)
library(sp) #for calculating geographic distance 
```


## Need the UCD 2023 and WL2 2023 & 2024 annual census data sheets (for fruit number)
```{r}
ucd_ann_cens <- read_csv("../output/UCD_Traits/UCD2023_Annual_Census_Combined.csv") %>%  #note this is only for plants that survived to rep 
  rename(pop=parent.pop) %>% 
  filter(!is.na(pop)) %>% 
  filter(rep != 100) %>% #get rid of individuals that germinated in the field 
  unite(Genotype, pop:rep, sep="_", remove = FALSE) 

wl2_ann_cens <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_annual_census_20231027_corrected.csv") %>% 
  unite(Genotype, pop:rep, sep="_", remove = FALSE) %>% 
  filter(!is.na(pop), !str_detect(Genotype, ".*buff*")) #remove buffers 

#wl2_ann_cens_2024

```


## Need surv post-transplant shock, surv to rep both years, and winter survival
```{r}
ucd_surv <- read_csv("../input/UCD_Data/CorrectedCSVs/UCD_transplants_pheno_mort_20231016_corrected.csv") %>% 
  rename(death.date=`Death Date`, bud.date=`Date First Bud`, flower.date=`Date First Flower`, 
         fruit.date=`Date First Fruit`, last.flower.date=`Date Last Flower`, last.fruit.date=`Date Last Fruit`) %>% 
  filter(!is.na(pop)) %>% 
  filter(rep != 100) %>% #get rid of individuals that germinated in the field 
  unite(Genotype, pop:rep, sep="_", remove = FALSE) 

wl2_surv_1020 <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_mort_pheno_20231020_corrected.csv") #need to add in 10/27 mortality 
wl2_surv_1027 <- wl2_ann_cens %>% 
  filter(pheno=="X") %>% 
  select(death.date_2=date, block:bed, bed.row=`bed-row`, bed.col=`bed-col`, pop:rep) #add in 10/27 death dates
#note: 10/27 pheno dates (like flowering), not updated with this code 
wl2_surv_y1 <- left_join(wl2_surv_1020, wl2_surv_1027) %>% 
  mutate(death.date=if_else(is.na(death.date), death.date_2, death.date)) %>% 
  unite(Genotype, pop:rep, sep="_", remove = FALSE) %>% 
  filter(!is.na(pop), !str_detect(Genotype, ".*buff*")) %>%  #remove buffers 
  select(block:survey.notes)

wl2_20241023 <- read_csv("../input/WL2_Data/WL2_mort_pheno_20241023_corrected.csv") %>% #note this has 2023 and 2024 plants
  select(-block)
wl2_y2_pops <- read_csv("../input/WL2_Data/Final_2023_2024_Pop_Loc_Info.csv") %>%
  select(Pop.Type:unique.ID) %>% 
  filter(Pop.Type=="2023-survivor") %>% 
  select(Pop.Type, loc:bed, row=bedrow, col=bedcol, pop:unique.ID)
wl2_surv_y2 <- left_join(wl2_y2_pops, wl2_20241023)
```

## Gower's Distance
```{r}
ucd_gowers <- read_csv("../output/Climate/Gowers_UCD.csv") %>% 
  select(parent.pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  rename(pop=parent.pop)

wl2_gowers_2023 <- read_csv("../output/Climate/Gowers_WL2.csv") %>% 
  select(parent.pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  rename(pop=parent.pop)

#wl2_gowers_2024 
```


## p(Establishment)
Davis - used 3 weeks to match WL2
```{r}
unique(ucd_surv$pop)
ucd_establishment <- ucd_surv %>% 
  left_join(ucd_gowers) %>% 
  select(block:rep, elevation.group:Wtr_Year_GD_Historical, death.date) %>% 
  mutate(death.date=mdy(death.date)) %>% 
  mutate(Establishment=if_else(is.na(death.date), 1,
                               if_else(death.date=="2022-11-30" | death.date=="2022-12-13" | death.date=="2022-12-21", 0,
                                       1)))

ucd_establishment %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanEst), y=meanEst, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Establishment)", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

ucd_establishment %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanEst), y=meanEst, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Establishment)", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
```

WL2 - 3 weeks based off survival curve
```{r}
wl2_establishment <- wl2_surv_y1 %>% 
  left_join(wl2_gowers_2023) %>% 
  select(block:rep, elevation.group:Wtr_Year_GD_Historical, death.date) %>% 
  mutate(death.date=mdy(death.date)) %>% 
  mutate(Establishment=if_else(is.na(death.date), 1,
                               if_else(death.date=="2023-07-26" | death.date=="2023-08-02" | death.date=="2023-08-09", 0,
                                       1)))
  
wl2_establishment %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanEst), y=meanEst, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Establishment)", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

wl2_establishment %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanEst), y=meanEst, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Establishment)", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
```


## p(Surv to Rep - y1)
Davis
```{r}
ucd_surv %>% filter(!is.na(bud.date), is.na(fruit.date)) #all plants that initiated reproduction, but did not finish, have a death date 
ucd_surv %>% filter(!is.na(fruit.date), !is.na(death.date)) #many plants with a fruit date, have a later death date 
ucd_surv %>% filter(!is.na(fruit.date), is.na(death.date)) #some fruit plants do not have a death date
ucd_surv %>% filter(is.na(fruit.date), !is.na(last.flower.date)) #did not miss any fruiting plants
ucd_surv %>% filter(is.na(fruit.date), !is.na(last.fruit.date)) #did not miss any fruiting plants 

ucd_surv_to_rep <- ucd_surv %>% 
  left_join(ucd_gowers) %>% 
  mutate(death.date=mdy(death.date)) %>% 
  mutate(SurvtoRep_Y1=if_else(is.na(fruit.date), 0, 1)) %>% 
  select(block:rep, elevation.group:Wtr_Year_GD_Historical, fruit.date, death.date, SurvtoRep_Y1) 

ucd_surv_to_rep %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(SurvtoRep_Y1, na.rm = TRUE), semSurv=sem(SurvtoRep_Y1, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Surv-to-Rep Y1)", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

ucd_surv_to_rep %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(SurvtoRep_Y1, na.rm = TRUE), semSurv=sem(SurvtoRep_Y1, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Surv-to-Rep Y1)", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
```

WL2
```{r}
wl2_surv_y1 %>% filter(!is.na(bud.date), is.na(fruit.date)) #some plants initiated reproduction but did not make fruits
wl2_surv_y1 %>% filter(!is.na(bud.date), !is.na(death.date))  #some of above died, some lived
wl2_surv_y1 %>% filter(!is.na(fruit.date), !is.na(death.date)) #no plants that made fruits have a death date 
wl2_surv_y1 %>% filter(!is.na(last.flower.date)) #NAs
wl2_surv_y1 %>% filter(!is.na(last.fruit.date)) #none 
wl2_surv_y1 %>% filter(!is.na(fruit.date)) 

wl2_surv_to_rep_y1 <- wl2_surv_y1 %>% 
  left_join(wl2_gowers_2023) %>% 
  mutate(death.date=mdy(death.date)) %>% 
  mutate(fruit.date=if_else(Genotype=="TM2_1_12", "10/27/23", fruit.date)) %>% #add in fruit date for plant that started fruiting the week of the annual census 
  mutate(SurvtoRep_Y1=if_else(is.na(fruit.date), 0, 1)) %>% 
  select(block:rep, elevation.group:Wtr_Year_GD_Historical, fruit.date, death.date, SurvtoRep_Y1) 
#this is quite uninteresting since only TM2 reproduced in year 1 

wl2_surv_to_rep_y1 %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(SurvtoRep_Y1, na.rm = TRUE), semSurv=sem(SurvtoRep_Y1, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Surv-to-Rep Y1)", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

wl2_surv_to_rep_y1 %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(SurvtoRep_Y1, na.rm = TRUE), semSurv=sem(SurvtoRep_Y1, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Surv-to-Rep Y1)", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
```


## Fruits(y1) 
*What to do about flower #?

Davis
```{r}
ucd_fruits <- ucd_ann_cens %>% select(block:rep, elevation.group, elev_m, Lat:Long, flowers, fruits) %>%  #note this is only for plants that survived to rep 
  left_join(ucd_gowers)

ucd_fruits %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(fruits, na.rm = TRUE), semFruits=sem(fruits, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Fruit Number", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

ucd_fruits %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(fruits, na.rm = TRUE), semFruits=sem(fruits, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Fruit Number", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
```

WL2
```{r}
wl2_fruits_y1 <- wl2_ann_cens %>% select(block:rep, flowers=num.flw, fruits=num.fruit) %>% 
  left_join(wl2_gowers_2023)

wl2_fruits_y1 %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(fruits, na.rm = TRUE), semFruits=sem(fruits, na.rm=TRUE)) %>%
  filter(pop!="WV") %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Fruit Number", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

wl2_fruits_y1 %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(fruits, na.rm = TRUE), semFruits=sem(fruits, na.rm=TRUE)) %>% 
  filter(pop!="WV") %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Fruit Number", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
```


## p(Winter Surv)
WL2
*What gowers distance should I use for this?
```{r}

```


## p(Surv to Rep - y2)
WL2
```{r}

```


## Fruits(y2)
WL2
```{r}
#wl2_fruits_y2
```


## Total fitness

Davis - No Year 2
p(Establishment)*p(Surv to Rep - y1)*Fruits(y1) 
```{r}
names(ucd_establishment)
names(ucd_surv_to_rep)
names(ucd_fruits)

#per individual:
ucd_total_fitness <- left_join(ucd_establishment, ucd_surv_to_rep) %>% 
  left_join(ucd_fruits) %>% 
  select(block:Wtr_Year_GD_Historical, Establishment, SurvtoRep_Y1:fruits) %>% 
  mutate(fruits=if_else(is.na(fruits), 0, fruits)) %>% 
  mutate(Total_Fitness=Establishment*SurvtoRep_Y1*fruits)
head(ucd_total_fitness)
ucd_total_fitness %>% filter(Total_Fitness>0)

ucd_total_fitness %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(Total_Fitness, na.rm = TRUE), semFruits=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

ucd_total_fitness %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(Total_Fitness, na.rm = TRUE), semFruits=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#could also calculate it for each treatment from avg probabilities?
```

WL2
*Use average Gowers?

p(Establishment)*p(Surv to Rep - y1)*Fruits(y1) + p(Winter Surv)*p(Surv to Rep - y2)*Fruits(y2)
```{r}
wl2_establishment
wl2_surv_to_rep_y1
```

