---
title: "Total Fitness"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Total Fitness at both sites 
To Do:
- add in geographic distance 

Total Fitness: p(Establishment)*p(Surv to Rep - y1)*Fruits(y1) + p(Winter Surv)*p(Surv to Rep - y2)*Fruits(y2)

## Libraries
```{r}
library(tidyverse)
library(sp) #for calculating geographic distance 

library(brms)
conflicts_prefer(brms::ar)
library(tidybayes) #for extracting and visiaulizing brms model output 
library(modelr) #for data grid

sem <- function(x, na.rm=FALSE) {           #for caclulating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 
```

## Year 2 Pop Info
```{r}
wl2_y2_pops <- read_csv("../input/WL2_Data/Final_2023_2024_Pop_Loc_Info.csv") %>%
  select(Pop.Type:unique.ID) %>% 
  filter(Pop.Type=="2023-survivor") %>% 
  select(Pop.Type, loc:bed, row=bedrow, col=bedcol, pop:unique.ID)
```


## Need the UCD 2023 and WL2 2023 & 2024 annual census data sheets (for fruit number)
```{r}
ucd_ann_cens <- read_csv("../output/UCD_Traits/UCD2023_Annual_Census_Combined.csv") %>%  #note this is only for plants that survived to rep 
  rename(pop=parent.pop) %>% 
  filter(!is.na(pop)) %>% 
  filter(rep != 100) %>% #get rid of individuals that germinated in the field 
  unite(Genotype, pop:rep, sep="_", remove = FALSE) 

wl2_ann_cens <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_annual_census_20231027_corrected.csv") %>% 
  unite(Genotype, pop:rep, sep="_", remove = FALSE) %>% 
  unite(BedLoc, bed:`bed-col`, sep="_", remove = FALSE) %>% 
  filter(BedLoc!="K_5_C") %>% #get rid of duplicate locations
  filter(BedLoc!="B_32_A") %>% #get rid of duplicate locations
  filter(!is.na(pop), !str_detect(Genotype, ".*buff*")) #remove buffers 

wl2_ann_cens_2024 <- read_csv("../input/WL2_Data/WL2_Annual_Census_20241023_corrected.csv")
wl2_ann_cens_2024_pops <- left_join(wl2_y2_pops, wl2_ann_cens_2024) %>%  
  rename(Genotype=unique.ID)
```


## Need surv post-transplant shock, surv to rep both years, and winter survival
```{r}
ucd_surv <- read_csv("../input/UCD_Data/CorrectedCSVs/UCD_transplants_pheno_mort_20231016_corrected.csv") %>% 
  rename(death.date=`Death Date`, bud.date=`Date First Bud`, flower.date=`Date First Flower`, 
         fruit.date=`Date First Fruit`, last.flower.date=`Date Last Flower`, last.fruit.date=`Date Last Fruit`) %>% 
  filter(!is.na(pop)) %>% 
  filter(rep != 100) %>% #get rid of individuals that germinated in the field 
  unite(Genotype, pop:rep, sep="_", remove = FALSE) 
#unique(ucd_surv$pop)

wl2_surv_1020 <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_mort_pheno_20231020_corrected.csv") #need to add in 10/27 mortality 
wl2_surv_1027 <- wl2_ann_cens %>% 
  filter(pheno=="X") %>% 
  select(death.date_2=date, block:bed, bed.row=`bed-row`, bed.col=`bed-col`, pop:rep) #add in 10/27 death dates
#note: 10/27 pheno dates (like flowering), not updated with this code 
wl2_surv_y1 <- left_join(wl2_surv_1020, wl2_surv_1027) %>% 
  mutate(death.date=if_else(is.na(death.date), death.date_2, death.date)) %>% 
  mutate(pop= str_replace(pop, "Y08", "YO8")) %>% 
  mutate(pop= str_replace(pop, "Y04", "YO4")) %>% 
  unite(BedLoc, bed:bed.col, sep="_", remove = FALSE) %>% 
  filter(BedLoc!="K_5_C") %>% #get rid of duplicate locations
  filter(BedLoc!="B_32_A") %>% #get rid of duplicate locations
  unite(Genotype, pop:rep, sep="_", remove = FALSE) %>% 
  filter(!is.na(pop), !str_detect(Genotype, ".*buff*")) %>%  #remove buffers 
  select(block:survey.notes)
#wl2_surv_y1_to_export <- wl2_surv_y1 %>% select(block:rep, death.date, survey.notes)
#write_csv(wl2_surv_y1_to_export, "../output/WL2_Traits/WL2_Mortality_2023.csv")
#unique(wl2_surv_y1$pop)

wl2_20241023 <- read_csv("../input/WL2_Data/WL2_mort_pheno_20241023_corrected.csv") %>% #note this has 2023 and 2024 plants
  select(-block)
wl2_surv_y2 <- left_join(wl2_y2_pops, wl2_20241023) %>%  
  rename(Genotype=unique.ID)
#unique(wl2_surv_y2$pop) #only 10 pops with winter surv 

post_winter <- read_csv("../input/WL2_Data/WL2_status_check_20240603_corrected.csv",
                         na = c("", "NA", "-", "N/A")) 
post_winter_clean <- post_winter %>% 
  mutate(pop= str_replace(pop, "iH", "IH")) %>% 
  mutate(pop= str_replace(pop, "1H", "IH")) %>% 
  mutate(pop= str_replace(pop, "cc", "CC")) %>% 
  unite(BedLoc, bed:`bed- col`, sep="_", remove = FALSE) %>% 
  filter(BedLoc!="K_5_C") %>% #get rid of duplicate locations
  filter(BedLoc!="B_32_A") %>% #get rid of duplicate locations
  unite(Genotype, pop:rep, sep="_", remove = FALSE) %>% 
  filter(pop!="buffer", !str_detect(mf, "buf")) %>% 
  mutate(mf=as.double(mf), rep=as.double(rep))
unique(post_winter_clean$pop)
```

## Gower's Distance
```{r}
garden_climate <- read_csv("../output/Climate/flint_climate_UCDpops.csv") %>% 
  filter(parent.pop=="WL2_Garden" | parent.pop=="UCD_Garden") %>% 
  select(parent.pop:Long) %>% 
  distinct()
garden_climate
#UCD LAT/LONG = 38.53250, -121.7830
#WL2 Lat/Long = 38.82599, -120.2509

ucd_gowers <- read_csv("../output/Climate/Gowers_UCD.csv") %>% 
  select(parent.pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  #mutate(Geographic_Dist=spDistsN1()) %>%  #Great Circle distance (shortest distance b/t 2 points on the surface of a sphere)
  #mutate(Lat_Dist=38.53250-Lat, Long_Dist=-121.7830-Long) %>% #Garden-Home - lat and long per Gerst et al 2011 which kept them separate for some directionality
  rename(pop=parent.pop)

wl2_gowers_2023 <- read_csv("../output/Climate/Gowers_WL2.csv") %>% 
  select(parent.pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  #mutate(Lat_Dist=38.82599-Lat, Long_Dist=-120.2509-Long) %>% #Garden-Home - lat and long per Gerst et al 2011 which kept them separate for some directionality
  rename(pop=parent.pop)

wl2_gowers_2024 <- read_csv("../output/Climate/Gowers_WL2_2024.csv") %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  #mutate(Lat_Dist=38.82599-Lat, Long_Dist=-120.2509-Long) %>% #Garden-Home - lat and long per Gerst et al 2011 which kept them separate for some directionality
  rename(pop=parent.pop)
```

### WL2 Avg Gowers
```{r}
wl2_gowers_2023_prep <- wl2_gowers_2023 %>% rename_with(~paste0(., "_2023"), 6:9)
wl2_gowers_2024_prep <- wl2_gowers_2024 %>% rename_with(~paste0(., "_2024"), 6:9)

wl2_gowers_avg <- bind_rows(wl2_gowers_2023, wl2_gowers_2024) %>% 
  group_by(pop, elevation.group, elev_m, Lat, Long) %>% 
  summarise_at(c("GrwSsn_GD_Recent",  "GrwSsn_GD_Historical", "Wtr_Year_GD_Recent", "Wtr_Year_GD_Historical"), c(mean), na.rm = TRUE)
wl2_gowers_avg
```



## p(Establishment)
Davis - used 3 weeks to match WL2
```{r}
unique(ucd_surv$pop)
ucd_establishment <- ucd_surv %>% 
  left_join(ucd_gowers) %>% 
  select(block:rep, elevation.group:Wtr_Year_GD_Historical, death.date) %>% 
  mutate(death.date=mdy(death.date)) %>% 
  mutate(Establishment=if_else(is.na(death.date), 1,
                               if_else(death.date=="2022-11-30" | death.date=="2022-12-13" | death.date=="2022-12-21", 0,
                                       1)))

ucd_establishment %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanEst), y=meanEst, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Establishment)", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/UCD_Traits/UCD_Establishment_GrwSsn_GD_Recent.png", width = 12, height = 8, units = "in")

ucd_establishment %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanEst), y=meanEst, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Establishment)", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/UCD_Traits/UCD_Establishment_Wtr_Year_GD_Recent.png", width = 12, height = 8, units = "in")
```


WL2 - 3 weeks based off survival curve
```{r}
wl2_establishment <- wl2_surv_y1 %>% 
  left_join(wl2_gowers_2023) %>% 
  select(block:rep, elevation.group:Wtr_Year_GD_Historical, death.date) %>% 
  mutate(death.date=mdy(death.date)) %>% 
  mutate(Establishment=if_else(is.na(death.date), 1,
                               if_else(death.date=="2023-07-26" | death.date=="2023-08-02" | death.date=="2023-08-09", 0,
                                       1)))
  
wl2_establishment %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanEst), y=meanEst, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Establishment)", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

wl2_establishment %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanEst=mean(Establishment, na.rm = TRUE), semEst=sem(Establishment, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanEst), y=meanEst, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Establishment)", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
```

### Stats
```{r}

est_UCD_bf1 <- brmsformula(Establishment ~ Wtr_Year_GD_Recent + (1|pop))

get_prior(est_UCD_bf1, family = "bernoulli", data = ucd_establishment)
prior1 <- c(set_prior(prior = 'normal(0, 1)', class='Intercept'),
            set_prior(prior = 'normal(0, 1)', class='sd'))
```

```{r}
est_UCD_m1 <- brm(est_UCD_bf1, 
                             family = "bernoulli",
                             data = ucd_establishment,
                             cores=4,
                             iter = 5000, #increased iterations b/c complex model
                             sample_prior = "yes", # needed for hypothesis testing
                             control = list(adapt_delta = 0.9),
                             prior=prior1) #increased adapt_delta to help with divergent transitions
```


```{r}
prior_summary(est_UCD_m1)
summary(est_UCD_m1)
#Rhat <1.05 (good!)
#ESS > 1000 (good!)
```

```{r}
plot(est_UCD_m1,  nvariables = 3, ask=FALSE) #plots look a little better with prior distribution adjustments 

#pairs(est_UCD_m1)

pp_check(est_UCD_m1)  # posterior predictive checks
#The main use of this function is to check if you model predicts your data accurately (using the estimates). If it does, then you can use that model to generate new data and make accurate predictions.
#light blue = 10 random draws or distributions created by the model
#dark blue = posterior distribution
#some diffs between draws 
```

To calcualte the stats we need to extract the posterior samples, and add the Intercept to each pop random effect, and then compute the stats.

```{r}
intercept <- as_draws_df(est_UCD_m1, variable = "b_Intercept") %>% as_tibble() %>% select(starts_with("b"))

r_pops_raw <- as_draws_df(est_UCD_m1, variable = "*r_", regex = TRUE) %>% as_tibble() %>% select(starts_with("r"))

r_pops <- r_pops_raw %>% mutate(across(everything(), ~ .x + intercept$b_Intercept))

```

#### Hypothesis Testing 

```{r}
hypothesis(est_UCD_m1,
           class = NULL,
           hypothesis = "r_pop[BH,Intercept] > 0",
           scope = "standard")
```

test whether WV increases survival probability *more* than other pops
```{r}
hypothesis(est_UCD_m1,
           class = NULL,
           hypothesis = "r_pop[TM2,Intercept] > r_pop[WL2,Intercept]",
           scope = "standard")
#Posterior probability that TM2 establishment is higher than WL2 is 81%
```

#### Tidy Plot 
```{r}
ucd_establishment %>% 
  data_grid(pop, Wtr_Year_GD_Recent) %>%
  add_epred_draws(est_UCD_m1) %>%
  ggplot(aes(y = pop, x = .epred)) +
  stat_slab() +
  theme_classic() +
  labs(x="Posterior Survival Predictions", y="Parent Population", title="UCD p(Establishment)") #+
 # theme(text=element_text(size=25)) 
```

## p(Surv to Rep - y1)
Davis
```{r}
ucd_surv %>% filter(!is.na(bud.date), is.na(fruit.date)) #all plants that initiated reproduction, but did not finish, have a death date 
ucd_surv %>% filter(!is.na(fruit.date), !is.na(death.date)) #many plants with a fruit date, have a later death date 
ucd_surv %>% filter(!is.na(fruit.date), is.na(death.date)) #some fruit plants do not have a death date
ucd_surv %>% filter(is.na(fruit.date), !is.na(last.flower.date)) #did not miss any fruiting plants
ucd_surv %>% filter(is.na(fruit.date), !is.na(last.fruit.date)) #did not miss any fruiting plants 

ucd_surv_to_rep <- ucd_surv %>% 
  left_join(ucd_gowers) %>% 
  mutate(death.date=mdy(death.date)) %>% 
  mutate(SurvtoRep_Y1=if_else(is.na(fruit.date), 0, 1)) %>% 
  select(block:rep, elevation.group:Wtr_Year_GD_Historical, fruit.date, death.date, SurvtoRep_Y1) 

ucd_surv_to_rep %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(SurvtoRep_Y1, na.rm = TRUE), semSurv=sem(SurvtoRep_Y1, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Surv-to-Rep Y1)", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

ucd_surv_to_rep %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(SurvtoRep_Y1, na.rm = TRUE), semSurv=sem(SurvtoRep_Y1, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Surv-to-Rep Y1)", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
```

WL2
```{r}
wl2_surv_y1 %>% filter(!is.na(bud.date), is.na(fruit.date)) #some plants initiated reproduction but did not make fruits
wl2_surv_y1 %>% filter(!is.na(bud.date), !is.na(death.date))  #some of above died, some lived
wl2_surv_y1 %>% filter(!is.na(fruit.date), !is.na(death.date)) #no plants that made fruits have a death date 
wl2_surv_y1 %>% filter(!is.na(last.flower.date)) #NAs
wl2_surv_y1 %>% filter(!is.na(last.fruit.date)) #none 
wl2_surv_y1 %>% filter(!is.na(fruit.date)) 

wl2_surv_to_rep_y1 <- wl2_surv_y1 %>% 
  left_join(wl2_gowers_2023) %>% 
  mutate(death.date=mdy(death.date)) %>% 
  mutate(fruit.date=if_else(Genotype=="TM2_1_12", "10/27/23", fruit.date)) %>% #add in fruit date for plant that started fruiting the week of the annual census 
  mutate(SurvtoRep_Y1=if_else(is.na(fruit.date), 0, 1)) %>% 
  select(block:rep, elevation.group:Wtr_Year_GD_Historical, fruit.date, death.date, SurvtoRep_Y1) 
#this is quite uninteresting since only TM2 reproduced in year 1 

wl2_surv_to_rep_y1 %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(SurvtoRep_Y1, na.rm = TRUE), semSurv=sem(SurvtoRep_Y1, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Surv-to-Rep Y1)", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

wl2_surv_to_rep_y1 %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(SurvtoRep_Y1, na.rm = TRUE), semSurv=sem(SurvtoRep_Y1, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Surv-to-Rep Y1)", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
```


## Fruits(y1) 
*What to do about flower #?

Davis
```{r}
ucd_fruits <- ucd_ann_cens %>% select(block:rep, elevation.group, elev_m, Lat:Long, flowers, fruits) %>%  #note this is only for plants that survived to rep 
  left_join(ucd_gowers)

ucd_fruits %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(fruits, na.rm = TRUE), semFruits=sem(fruits, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Fruit Number", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

ucd_fruits %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(fruits, na.rm = TRUE), semFruits=sem(fruits, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Fruit Number", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
```

WL2
```{r}
wl2_fruits_y1 <- wl2_ann_cens %>% select(block:rep, flowers=num.flw, fruits=num.fruit) %>% 
  left_join(wl2_gowers_2023)

wl2_fruits_y1 %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(fruits, na.rm = TRUE), semFruits=sem(fruits, na.rm=TRUE)) %>%
  filter(pop!="WV") %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Fruit Number", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

wl2_fruits_y1 %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(fruits, na.rm = TRUE), semFruits=sem(fruits, na.rm=TRUE)) %>% 
  filter(pop!="WV") %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Fruit Number", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
```


## p(Winter Surv)
WL2
*Used only water year gowers distance for 2024 b/c that's the only one that includes the winter months 
```{r}
#wl2_surv_y1 %>% filter(is.na(death.date)) - 470 plants alive on 10/27
post_winter_clean %>% filter(is.na(death.date)) #44 plants with no death date, most have a death date of 10/27
#A-37-B, no death date on any data sheet, was dying on 9/20 so may have just been missed in June 2024

winter_surv <- post_winter_clean %>% 
  left_join(wl2_gowers_2024) %>% 
  filter(death.date == "A" | death.date == "B" | death.date == "C" | death.date == "D") %>% 
  mutate(WinterSurv=if_else(death.date=="D", 0, 1)) %>% 
  select(block:rep, elevation.group:Wtr_Year_GD_Historical, death.date, WinterSurv) 
dim(winter_surv) #only 469 rows because A_37_B not included 

#winter_surv %>% filter(death.date!="D") #135 plants survived 

winter_surv %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  filter(pop != "WR") %>% #only 1 plant 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Winter Surv)", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
```


## p(Surv to Rep - y2)
WL2
```{r}
wl2_surv_y2 %>% filter(!is.na(bud.date), is.na(fruit.date)) #some plants initiated reproduction but did not make fruits
#1 plant with a last flower date, but no fruit date, this plant did not make any fruits per the notes 
wl2_surv_y2 %>% filter(!is.na(bud.date), !is.na(death.date))  #some of above died, some lived
#some rep plants have a death date 
wl2_surv_y2 %>% filter(!is.na(fruit.date), !is.na(death.date)) 
wl2_surv_y2 %>% filter(!is.na(fruit.date), is.na(last.FL.date)) #not all fruiting plants have a last flower date
wl2_surv_y2 %>% filter(!is.na(fruit.date), is.na(last.FR.date)) #not all fruiting plants have a last fruit date
wl2_surv_y2 %>% filter(is.na(fruit.date), !is.na(last.FL.date)) #see note above about this one plant
wl2_surv_y2 %>% filter(is.na(fruit.date), !is.na(last.FR.date)) #none 

wl2_surv_to_rep_y2 <- wl2_surv_y2 %>% 
  left_join(wl2_gowers_2023) %>% 
  mutate(SurvtoRep_y2=if_else(is.na(fruit.date), 0, 1)) %>% 
  select(Pop.Type:Genotype, elevation.group:Wtr_Year_GD_Historical, fruit.date, death.date, SurvtoRep_y2) 

wl2_surv_to_rep_y2 %>% group_by(pop) %>% summarise(n=n()) %>% arrange(n)
#LV1, SQ1, and WR only have 1 indiv alive in 2024

wl2_surv_to_rep_y2 %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(SurvtoRep_y2, na.rm = TRUE), semSurv=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  filter(pop != "LV1", pop !="SQ1", pop !="WR") %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Surv-to-Rep Y2)", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

wl2_surv_to_rep_y2 %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(SurvtoRep_y2, na.rm = TRUE), semSurv=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  filter(pop != "LV1", pop !="SQ1", pop !="WR") %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="p(Surv-to-Rep Y2)", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
```


## Fruits(y2)
WL2
```{r}
wl2_fruits_y2 <- wl2_ann_cens_2024_pops %>% select(Pop.Type:Genotype, flowers=num.flw, fruits=num.fruit) %>% 
  left_join(wl2_gowers_2024)

wl2_fruits_y2 %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(fruits, na.rm = TRUE), semFruits=sem(fruits, na.rm=TRUE)) %>%
  filter(pop != "LV1", pop !="SQ1", pop !="WR") %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Fruit Number", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

wl2_fruits_y2 %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(fruits, na.rm = TRUE), semFruits=sem(fruits, na.rm=TRUE)) %>% 
  filter(pop != "LV1", pop !="SQ1", pop !="WR") %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Fruit Number", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

wl2_fruits_y2 %>% filter(pop=="TM2") #only 3 of the 6 made fruits and those 3 had wide variation (2, 19, 125)
```


## Total fitness

Davis - No Year 2
p(Establishment)*p(Surv to Rep - y1)*Fruits(y1) 
```{r}
names(ucd_establishment)
names(ucd_surv_to_rep)
names(ucd_fruits)

#per individual:
ucd_total_fitness <- left_join(ucd_establishment, ucd_surv_to_rep) %>% 
  left_join(ucd_fruits) %>% 
  select(block:Wtr_Year_GD_Historical, Establishment, SurvtoRep_Y1:fruits) %>% 
  mutate(fruits=if_else(is.na(fruits), 0, fruits)) %>% 
  mutate(Total_Fitness=Establishment*SurvtoRep_Y1*fruits)
head(ucd_total_fitness)
ucd_total_fitness %>% filter(Total_Fitness>0)

ucd_total_fitness %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(Total_Fitness, na.rm = TRUE), semFruits=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

ucd_total_fitness %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(Total_Fitness, na.rm = TRUE), semFruits=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#could also calculate it for each treatment from avg probabilities?
```

WL2
*Use average Gowers?

p(Establishment)*p(Surv to Rep - y1)*Fruits(y1) + p(Winter Surv)*p(Surv to Rep - y2)*Fruits(y2)
```{r}
wl2_fruits_y1_prep <- wl2_fruits_y1 %>% 
  select(block:BedLoc, Genotype:rep, elevation.group:Long, y1_flowers=flowers, y1_fruits=fruits)

winter_surv_prep <- winter_surv %>% select(block:BedLoc, Genotype:Long, WinterSurv) %>% 
  mutate(mf=as.character(mf), rep=as.character(rep))
wl2_surv_to_rep_y2_prep <- wl2_surv_to_rep_y2 %>% select(pop:Long, SurvtoRep_y2) %>% 
  mutate(mf=as.character(mf), rep=as.character(rep))
wl2_fruits_y2_prep <- wl2_fruits_y2 %>% 
  select(pop:Genotype, elevation.group:Long, y2_flowers=flowers, y2_fruits=fruits) %>% 
  mutate(mf=as.character(mf), rep=as.character(rep))

wl2_total_fitness <- left_join(wl2_establishment, wl2_surv_to_rep_y1) %>% 
  left_join(wl2_fruits_y1_prep) %>% 
  left_join(winter_surv_prep) %>% 
  left_join(wl2_surv_to_rep_y2_prep) %>% 
  left_join(wl2_fruits_y2_prep) %>% 
  left_join(wl2_gowers_avg) %>% 
  select(block:Wtr_Year_GD_Historical, Establishment, SurvtoRep_Y1:y2_fruits) %>% 
  mutate(y1_fruits=if_else(is.na(y1_fruits), 0, y1_fruits),
         y2_fruits=if_else(is.na(y2_fruits), 0, y2_fruits),
         WinterSurv=if_else(is.na(WinterSurv), 0, WinterSurv),
         SurvtoRep_y2=if_else(is.na(SurvtoRep_y2), 0, SurvtoRep_y2)) %>% 
  mutate(Total_Fitness=(Establishment*SurvtoRep_Y1*y1_fruits) + (WinterSurv*SurvtoRep_y2*y2_fruits))
wl2_total_fitness

wl2_total_fitness %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(Total_Fitness, na.rm = TRUE), semFruits=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits, color=GrwSsn_GD_Recent),
                show.legend = FALSE, width=.2, position = position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  scale_colour_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/WL2_Traits/WL2_Total_Fitness_GrwSsn_GD_Recent.png", width = 12, height = 8, units = "in")

wl2_total_fitness %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanFruits=mean(Total_Fitness, na.rm = TRUE), semFruits=sem(Total_Fitness, na.rm=TRUE)) %>% 
  ggplot(aes(x=fct_reorder(pop, meanFruits), y=meanFruits, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fitness", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_gradient(low = "#0043F0", high = "#F5A540") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/WL2_Traits/WL2_Total_Fitness_Wtr_Year_GD_Recent.png", width = 12, height = 8, units = "in")
```

