---
title: "Single Time Size Reaction Norms"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reaction Norms of Size (between UCD and WL2) - Two Months Post-Transplanting

## Relevant Libraries and Functions

```{r}
library(tidyverse)
library(lmerTest) #for mixed effect models
library(emmeans) #for post-hoc pairwise comparisons 
library(naniar) #replaces values with NA
library(ggnewscale)
library(corrplot) #plotting correlations 
library(geosphere) #for calculating geographic distance
conflicted::conflicts_prefer(lmerTest::lmer)
conflicted::conflicts_prefer(dplyr::filter)

sem <- function(x, na.rm=FALSE) {           #for calculating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low") #for proper arrangement in figures 
```

## Load Size data from both Gardens

```{r}
WL2_twomonths_size <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_size_survey_20230913_corrected.csv", 
                               na = c("", "NA", "-", "N/A")) %>% 
  filter(!is.na(pop)) %>% 
  rename(parent.pop=pop) %>% 
  mutate(parent.pop= str_replace(parent.pop, ".*VTR.*", "LVTR1")) %>% 
  unite(BedLoc, bed:bed.col, sep="_", remove = FALSE) %>% 
  filter(BedLoc!="K_5_C") %>% 
  filter(BedLoc!="B_32_A") %>% 
  unite(Genotype, parent.pop:rep, sep="_", remove = FALSE) %>% 
  filter(!str_detect(Genotype, ".*buff*")) %>% 
  unite(pop.mf, parent.pop:mf, sep="_", remove = FALSE)
head(WL2_twomonths_size)
dim(WL2_twomonths_size) #2 extra rows...
#summary(WL2_twomonths_size)
WL2_twomonths_size %>% rowwise() %>% filter(!is.na(mf), mf != "buffer") %>%  filter(is.na(as.numeric(mf)))
WL2_twomonths_size %>% rowwise() %>% filter(!is.na(rep), rep != "buffer") %>%  filter(is.na(as.numeric(rep))) #all buffer rows 
unique(WL2_twomonths_size$parent.pop) #all pops with * are buffers

UCD_twomonths_size <- read_csv("../input/UCD_Data/CorrectedCSVs/Size_survey_transplants_20230127_corrected.csv", 
                               na = c("", "NA", "-", "N/A")) %>% 
  rename(height.cm = `height (cm)`, long.leaf.cm = `longest leaf (cm)`, parent.pop=pop) %>% 
  mutate(parent.pop= str_replace(parent.pop, ".*VTR.*", "LVTR1")) %>% 
  unite(BedLoc, block:col, sep="_", remove = FALSE) %>% 
  filter(rep != 100)  %>% #get rid of individuals that germinated in the field 
  unite(Genotype, parent.pop:rep, sep="_", remove = FALSE) %>% 
  filter(!str_detect(Genotype, ".*buff*")) %>% 
  unite(pop.mf, parent.pop:mf, sep="_", remove = FALSE)
head(UCD_twomonths_size)
dim(UCD_twomonths_size) #correct number of rows 
#summary(UCD_twomonths_size)
unique(UCD_twomonths_size$parent.pop) #all pops with * are buffers

UCD_twomonths_size %>% filter(!is.na(...10)) #extra notes 
```

### Merge the Data

```{r}
WL2_twomonths_size_prep <- WL2_twomonths_size %>% 
  filter(BedLoc!="C_4_D", BedLoc!="C_5_D") %>% 
  select(block:BedLoc, Genotype:long.leaf.cm) %>% 
  mutate(Site="WL2") %>%
  mutate(mf=as.double(mf), rep=as.double(rep)) 
names(WL2_twomonths_size_prep)

UCD_twomonths_size_prep <- UCD_twomonths_size %>% 
  select(BedLoc:block, Genotype:long.leaf.cm) %>% mutate(Site="UCD")
names(UCD_twomonths_size_prep)

twomonths_rxnnorms <- bind_rows(UCD_twomonths_size_prep, WL2_twomonths_size_prep) %>% 
  arrange(pop.mf, Site)
head(twomonths_rxnnorms)
```

## Load the pop and location data

```{r}
#pop info
pops_common_garden <- read_csv("../input/WL2_Data/Pops_for_2023_WL2.csv") #pops included in common garden 
pops_common_garden_nonotes <- pops_common_garden %>% select(parent.pop, elevation.group) #subset columns

#extra location info 
pop_loc <- read_csv("../input/Strep_tort_locs.csv")

#need to change YOSE to YO
pop_loc_yo <- pop_loc %>% mutate(parent.pop = str_replace(`Site code`, "YOSE(\\d+)", "YO\\1")) %>% select(Lat, Long, elev_m=`Elevation (m)`, parent.pop)

#merge in location info
pop_elev <- left_join(pops_common_garden_nonotes, pop_loc_yo)
```

## Add in location info

```{r}
twomonths_rxnnorms_loc <-left_join(twomonths_rxnnorms, pop_elev) 
head(twomonths_rxnnorms_loc)

#write_csv(twomonths_rxnnorms_loc, "../output/TwoMonths_Size_BothSites.csv")
```

## Maternal Family Sample Sizes 
```{r}
#restrict to mfs with data at both sites 
twomonths_surv_mfs_wide <- twomonths_rxnnorms_loc %>% 
  group_by(pop.mf, parent.pop, Site, elev_m) %>% 
  summarise(N_height = sum(!is.na(height.cm))) %>% 
  select(pop.mf, Site, N_height) %>% 
  spread(Site, N_height) %>% 
  mutate(Both.Sites=if_else(!is.na(UCD) & !is.na(WL2), TRUE, FALSE)) %>% 
  filter(Both.Sites != "FALSE")

twomonths_surv_mfs_bothsites <- left_join(twomonths_surv_mfs_wide, twomonths_rxnnorms_loc)

mf_sample_size <- twomonths_surv_mfs_bothsites %>% 
  count(parent.pop, pop.mf, Site) %>% 
  arrange(n)
mf_sample_size #108 mfs with data at both sites 
summary(mf_sample_size)
mf_sample_size %>% 
  ggplot(aes(x=pop.mf, y=n)) +
  geom_col() + 
  labs(x="Pop.MF", y="Sample Size", title="Size Two Months Post-Transplant") +
  facet_wrap(~Site)

mf_per_pop <- twomonths_surv_mfs_bothsites %>% 
  select(parent.pop, pop.mf) %>% 
  distinct() %>% 
  group_by(parent.pop) %>% 
  summarise(n = n()) %>% 
  arrange(n)

mf_per_pop %>% 
  ggplot(aes(x=fct_reorder(parent.pop, n), y=n)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  theme_classic() +
  scale_y_continuous(limits = c(0,8), breaks = c(2, 4, 6, 8), expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1)) +
  labs(x="Parent Population", y="# Maternal Families", title="Size Two Months Post-Transplant")
```

## Plot Reaction Norms

### Means by maternal family

```{r}
twomonths_rxnnorms_summary_mfs <- twomonths_rxnnorms_loc %>% 
  group_by(pop.mf, parent.pop, Site, elev_m) %>% 
  summarise(N_height = sum(!is.na(height.cm)), mean_height.cm = mean(height.cm,na.rm=(TRUE)), 
            sem_height.cm=sem(height.cm, na.rm=(TRUE)), N_length = sum(!is.na(long.leaf.cm)),
            mean_long.leaf.cm=mean(long.leaf.cm, na.rm=(TRUE)), 
            sem_long.leaf.cm=sem(long.leaf.cm, na.rm=TRUE))
#twomonths_rxnnorms_summary_mfs %>% arrange(N_height)

#restrict to mfs with plants at both sites
twomonths_rxnnorms_summary_mfs_wide <- twomonths_rxnnorms_summary_mfs %>% 
  select(pop.mf, Site, N_height) %>% 
  spread(Site, N_height) %>% 
  mutate(Both.Sites=if_else(!is.na(UCD) & !is.na(WL2), TRUE, FALSE)) %>% 
  filter(Both.Sites != "FALSE")
#twomonths_rxnnorms_summary_mfs_wide %>% arrange(Both.Sites)

twomonths_rxnnorms_summary_mfs_bothsites <- left_join(twomonths_rxnnorms_summary_mfs_wide, twomonths_rxnnorms_summary_mfs)
#twomonths_rxnnorms_summary_mfs_bothsites

twomonths_rxnnorms_summary_mfs2 <- twomonths_rxnnorms_summary_mfs_bothsites %>% 
  mutate(Site=str_replace_all(Site, "UCD", "Low Elev"), 
         Site=str_replace_all(Site, "WL2", "High Elev")) 
twomonths_rxnnorms_summary_mfs2$Site <- factor(twomonths_rxnnorms_summary_mfs2$Site,
                                               levels = c('Low Elev','High Elev'))

twomonths_rxnnorms_summary_mfs2 %>% filter(N_height!=1) %>% filter(is.na(sem_height.cm))
#20 maternal families with 0 plants left alive at one of the sites by these measurements. 
```

### Plot mf avgs

```{r, eval=FALSE}
twomonths_rxnnorms_summary_mfs %>% 
   ggplot(aes(x=Site, y=mean_height.cm, group=pop.mf, color=pop.mf)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.2) +
  theme_classic() +
   theme(text=element_text(size=25))

twomonths_rxnnorms_summary_mfs %>% 
   ggplot(aes(x=Site, y=mean_long.leaf.cm, group=pop.mf, color=pop.mf)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,ymax=mean_long.leaf.cm+sem_long.leaf.cm),width=.2) +
  theme_classic() +
   theme(text=element_text(size=25))
```

```{r}
#note that there are a lot of maternal families with only 1 plant - removed those from these figures 
twomonths_rxnnorms_summary_mfs2 %>% 
  filter(N_height >2) %>% 
  ggplot(aes(x=Site, y=mean_height.cm, group=pop.mf, color=elev_m)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Height (cm)", color="Elevation (m)") +
  theme(text=element_text(size=28))
ggsave("../output/TwoMonths_RxNorms_Height_MFAvgs.png", width = 14, height = 8, units = "in")

twomonths_rxnnorms_summary_mfs2 %>% 
  filter(N_length>2) %>% 
  ggplot(aes(x=Site, y=mean_long.leaf.cm, group=pop.mf, color=elev_m)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,ymax=mean_long.leaf.cm+sem_long.leaf.cm),
                width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Leaf Length (cm)", color="Elevation (m)") +
  theme(text=element_text(size=28))
ggsave("../output/TwoMonths_RxNorms_LongestLength_MFAvgs.png", width = 14, height = 8, units = "in")
```

```{r}
twomonths_rxnnorms_summary_mfs2 %>% 
  filter(N_height >2) %>% 
  filter(parent.pop=="BH"|parent.pop=="WL1"|parent.pop=="CP2") %>%  
  ggplot(aes(x=Site, y=mean_height.cm, group=pop.mf, color=elev_m)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Height (cm)", color="Elevation (m)") +
  theme(text=element_text(size=28))
#ggsave("../output/TwoMonths_RxNorms_Height_BH_WL1_CP2_MFAvgs.png", width = 14, height = 8, units = "in")

twomonths_rxnnorms_summary_mfs2 %>% 
  filter(N_length >2) %>% 
  filter(parent.pop=="BH"|parent.pop=="WL1"|parent.pop=="CP2") %>%  
  ggplot(aes(x=Site, y=mean_long.leaf.cm, group=pop.mf, color=elev_m)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,ymax=mean_long.leaf.cm+sem_long.leaf.cm),
                width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Leaf Length (cm)", color="Elevation (m)") +
  theme(text=element_text(size=28))
#ggsave("../output/TwoMonths_RxNorms_LongestLength_BH_WL1_CP2_MFAvgs.png", width = 14, height = 8, units = "in")
```

### Means by Pop

```{r}
twomonths_rxnnorms_summary_pops <- twomonths_rxnnorms_loc %>% 
  filter(!is.na(parent.pop)) %>% 
  group_by(parent.pop, Site, elev_m) %>% 
  summarise(N_height = sum(!is.na(height.cm)), mean_height.cm = mean(height.cm,na.rm=(TRUE)), 
            sem_height.cm=sem(height.cm, na.rm=(TRUE)), N_length = sum(!is.na(long.leaf.cm)),
            mean_long.leaf.cm=mean(long.leaf.cm, na.rm=(TRUE)), 
            sem_long.leaf.cm=sem(long.leaf.cm, na.rm=TRUE))
#twomonths_rxnnorms_summary_pops

twomonths_rxnnorms_summary_pops %>% arrange(N_height)
twomonths_rxnnorms_summary_pops %>% filter(elev_m>2800) %>% filter(Site=="WL2") %>% summary()
twomonths_rxnnorms_summary_pops %>% filter(elev_m>2800) %>% filter(Site=="UCD") %>% summary()

twomonths_rxnnorms_summary_pops2 <- twomonths_rxnnorms_summary_pops %>% 
  mutate(Site=str_replace_all(Site, "UCD", "Low Elev"), 
         Site=str_replace_all(Site, "WL2", "High Elev"))
twomonths_rxnnorms_summary_pops2$Site <- factor(twomonths_rxnnorms_summary_pops2$Site,
                                               levels = c('Low Elev','High Elev'))
#no data for WV at high elevation (dead by the time of these measurements)
```

### Plot Pop Avgs

```{r}
twomonths_rxnnorms_summary_pops %>% 
  filter(N_height>2) %>% 
   ggplot(aes(x=Site, y=mean_height.cm, group=parent.pop, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.1) +
  theme_classic() +
   theme(text=element_text(size=25))
ggsave("../output/TwoMonths_RxNorms_Height_ALL_PopAvgs2.png", width = 12, height = 8, units = "in")

twomonths_rxnnorms_summary_pops %>% 
  filter(N_length>2) %>% 
   ggplot(aes(x=Site, y=mean_long.leaf.cm, group=parent.pop, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,ymax=mean_long.leaf.cm+sem_long.leaf.cm),width=.1) +
  theme_classic() +
   theme(text=element_text(size=25))
ggsave("../output/TwoMonths_RxNorms_LongestLength_ALL_PopAvgs2.png", width = 12, height = 8, units = "in")
```

```{r}
twomonths_rxnnorms_summary_pops2 %>% 
  filter(N_height >2) %>% 
  ggplot(aes(x=Site, y=mean_height.cm, group=parent.pop, color=elev_m)) + 
  geom_point(size=1.5) + geom_line(linewidth=1.5) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Height (cm)", color="Elevation (m)") +
 # ylim(0,4) +
  theme(text=element_text(size=28))
ggsave("../output/TwoMonths_RxNorms_Height_ALL_PopAvgs.png", width = 14, height = 8, units = "in")

twomonths_rxnnorms_summary_pops2 %>% 
  filter(N_length >2) %>% 
  ggplot(aes(x=Site, y=mean_long.leaf.cm, group=parent.pop, color=elev_m)) + 
  geom_point(size=1.5) + geom_line(linewidth=1.5) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,
                    ymax=mean_long.leaf.cm+sem_long.leaf.cm),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Leaf Length (cm)", color="Elevation (m)") +
  theme(text=element_text(size=28))
ggsave("../output/TwoMonths_RxNorms_LongestLength_ALL_PopAvgs.png", width = 14, height = 8, units = "in")
```

### Combine mf and pop averages

```{r}
twomonths_rxnnorms_summary_pops2 %>% 
  mutate(pop.mf=parent.pop) %>% 
  filter(N_height >2) %>% 
  ggplot(aes(x=Site, y=mean_height.cm, group=pop.mf, color=elev_m)) + 
  geom_point(size=1.5) + geom_line(linewidth=1.5, alpha=0.7) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Height (cm)", color="Elevation (m)") +
  theme(text=element_text(size=28)) +
 geom_line(data = twomonths_rxnnorms_summary_mfs2, size=0.2) +
  geom_point(data = twomonths_rxnnorms_summary_mfs2, size=0.8) 
ggsave("../output/TwoMonths_RxNorms_Height_ALL_PopMFAvgs.png", width = 14, height = 8, units = "in")

twomonths_rxnnorms_summary_pops2 %>% 
  mutate(pop.mf=parent.pop) %>% 
  filter(N_length >2) %>% 
  ggplot(aes(x=Site, y=mean_long.leaf.cm, group=pop.mf, color=elev_m)) + 
  geom_point(size=1.5) + geom_line(linewidth=1.5,  alpha=0.7) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,
                    ymax=mean_long.leaf.cm+sem_long.leaf.cm),width=.1) +
  theme_classic() + scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(y="Leaf Length (cm)", color="Elevation (m)") +
  theme(text=element_text(size=28)) +
geom_line(data = twomonths_rxnnorms_summary_mfs2, size=0.2) +
  geom_point(data = twomonths_rxnnorms_summary_mfs2, size=0.8) 
ggsave("../output/TwoMonths_RxNorms_LongestLength_ALL_PopMFAvgs.png", width = 14, height = 8, units = "in")
```

```{r}
twomonths_rxnnorms_summary_pops2 %>% 
  filter(parent.pop=="TM2"|parent.pop=="WL2") %>%  
   ggplot(aes(x=Site, y=mean_height.cm, group=parent.pop, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.2) +
  theme_classic()  +
   theme(text=element_text(size=25))
ggsave("../output/TwoMonths_RxNorms_Height_TM2_WL2_PopAvgs.png", width = 12, height = 8, units = "in")

twomonths_rxnnorms_summary_pops2 %>% 
  filter(parent.pop=="TM2"|parent.pop=="WL2") %>%  
   ggplot(aes(x=Site, y=mean_long.leaf.cm, group=parent.pop, color=parent.pop)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,ymax=mean_long.leaf.cm+sem_long.leaf.cm),width=.2) +
  theme_classic()  +
   theme(text=element_text(size=25))
ggsave("../output/TwoMonths_RxNorms_LongestLength_TM2_WL2_PopAvgs.png", width = 12, height = 8, units = "in")
```

### Means by Elevation

```{r}
twomonths_rxnnorms_summary_elev <- twomonths_rxnnorms_loc %>% 
  group_by(elevation.group, Site) %>% 
  summarise(N_height = sum(!is.na(height.cm)), mean_height.cm = mean(height.cm,na.rm=(TRUE)), 
            sem_height.cm=sem(height.cm, na.rm=(TRUE)), N_length = sum(!is.na(long.leaf.cm)),
            mean_long.leaf.cm=mean(long.leaf.cm, na.rm=(TRUE)), 
            sem_long.leaf.cm=sem(long.leaf.cm, na.rm=TRUE))
twomonths_rxnnorms_summary_elev
```

```{r}
twomonths_rxnnorms_summary_elev$elevation.group <- factor(twomonths_rxnnorms_summary_elev$elevation.group, levels = c('High','Mid', 'Low'))
twomonths_rxnnorms_summary_elev %>% 
   ggplot(aes(x=Site, y=mean_height.cm, group=elevation.group, color=elevation.group)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.2) +
  theme_classic() + scale_colour_manual(values=elev_three_palette)

twomonths_rxnnorms_summary_elev %>% 
   ggplot(aes(x=Site, y=mean_long.leaf.cm, group=elevation.group, color=elevation.group)) + 
  geom_point(size=0.8) + geom_line(linewidth=0.8) +
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,ymax=mean_long.leaf.cm+sem_long.leaf.cm),width=.2) +
  theme_classic() + scale_colour_manual(values=elev_three_palette)
```

## Stats

### Scaling

Center and scale/transform height and longest leaf in the plasticity models to get slopes that aren’t biased by plant size

#### Check distributions 
```{r}
twomonths_rxnnorms_loc %>% 
  ggplot(aes(height.cm)) +
  geom_histogram() #pretty skewed

twomonths_rxnnorms_loc %>% 
  mutate(logHeight=log(height.cm)) %>% 
  ggplot(aes(logHeight))+
  geom_histogram() #looks better 

twomonths_rxnnorms_loc %>% 
  ggplot(aes(long.leaf.cm)) +
  geom_histogram() #looks fine 

```

Test that log transformation will help 
```{r}
#Case where size doubles between sites:
##Plant 1: 2.3 --> 4.6
4.6-2.3 #2.3
log(4.6)-log(2.3) #0.6931472

##Plant 2: 4.5 --> 9 
9-4.5 #4.5
log(9)-log(4.5) #0.6931472

#Log transformation makes the difference between the two be on the same scale 
```

Test Centering and Scaling for Length
```{r}
#use same case as above chunk 
test <- tibble(length=c(2.3, 4.6, 4.5, 9), Site=c("UCD", "WL2","UCD", "WL2"), pop=c(1,1,2,2)) %>% 
  group_by(pop) %>% 
  mutate(length_scaled = scale(length))
test #the same 
```


```{r}
twomonths_rxnnorms_loc_scale <- twomonths_rxnnorms_loc %>% 
  mutate(logHeight=log(height.cm)) %>% 
  group_by(parent.pop) %>% 
  mutate(length_scaled = scale(long.leaf.cm)) %>% 
  ungroup()
```


### Height

```{r}
lmeheight2 <- lmer(logHeight ~ Site + (1|parent.pop/mf), data = twomonths_rxnnorms_loc_scale) #no block b/c nested within site?
summary(lmeheight2)
#ranova(lmeheight2) 
#anova(lmeheight2)

lmeheight3 <- lmer(logHeight ~ Site + (Site|parent.pop) + (1|pop.mf), data = twomonths_rxnnorms_loc_scale) 
summary(lmeheight3)
ranova(lmeheight3) 
anova(lmeheight3)
coef(lmeheight3)$parent.pop
plot(lmeheight3, which = 1)
qqnorm(resid(lmeheight3))
qqline(resid(lmeheight3)) #fit diagnostics not great 

height_slopes <- coef(lmeheight3)$parent.pop %>% 
  rename(height_slope=SiteWL2) %>% select(-`(Intercept)`) %>% 
  rownames_to_column(var="parent.pop")
```

### Length

```{r}
lmelength2 <- lmer(length_scaled ~ Site + (1|parent.pop/mf), data = twomonths_rxnnorms_loc_scale) #no block b/c nested within site?
summary(lmelength2) #boundary (singular) fit: see help('isSingular') - mf explains little variation 
#ranova(lmelength2) 
#anova(lmelength2)

lmelength3 <- lmer(length_scaled ~ Site + (1|parent.pop), data = twomonths_rxnnorms_loc_scale) #remove mf 
#boundary (singular) fit: see help('isSingular') - pop also explains little variation 
summary(lmelength3) 
ranova(lmelength3) 
anova(lmelength3)
lmtest <- lm(length_scaled ~ Site, data = twomonths_rxnnorms_loc_scale)
summary(lmtest) #taking pop out doesn't change Site effect - keep pop in 

lmelength4 <- lmer(length_scaled ~ Site + (Site|parent.pop), data = twomonths_rxnnorms_loc_scale) 
summary(lmelength4)
ranova(lmelength4) 
anova(lmelength4)
coef(lmelength4)$parent.pop

length_slopes <- coef(lmelength4)$parent.pop %>% 
  rename(length_slope=SiteWL2) %>% select(-`(Intercept)`) %>% 
  rownames_to_column(var="parent.pop")
```

### Export Site slopes 
```{r}
site_slopes <- left_join(height_slopes, length_slopes)

write_csv(site_slopes, "../output/TwoMonths_Size_Slopes.csv")
```

