---
title: "UCD Climate Distance"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Climate Distance at the Davis Garden

Annual data might give a different result, worth presenting it both ways
because that helps highlight the importance of the life history
differences

\*Used "water year" with a start date of Nov 1 b/c that's the first
month UCD was planted and includes all the months we collected data from
WL2. This (as opposed to calendar year) allowed for the inclusion of the
2022 months at the UCD garden

Options:

-   Basic subtraction as in Moran et al 2017 (garden site - home site)

-   Gower's environmental distance metric as in Rutter and Fenster 2007

-   From PCAs "We calculated an overall environmental distance using
    PCA, with EDij corresponding to the Euclidian distance between i and
    j based on PCA1 & PCA2." as in Moran et al 2017

## Relevant Libraries and Functions

```{r}
library(raster)
library(tidyverse)
library(conflicted)
conflicts_prefer(dplyr::select())
conflicts_prefer(dplyr::filter)
library(ggrepel)
library(cowplot)
library(gridExtra)
library(naniar) #replaces values with NA
library(boot)
library(broom)

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
} #legend function for grid_arrange

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")
```

## Davis Climate Data (Nov 2022-Oct 2023)

### From Flint (changed this from CIMIS)

```{r}
davis_climate <- read_csv("../output/Climate/flint_climate_UCDpops.csv") %>% 
  filter(parent.pop=="UCD_Garden") %>% 
  filter(year>=2022, year<2024) %>% 
  filter(if_else(year==2022, month=="nov" | month=="dec",
                 month!="nov" & month !="dec")) %>% 
  mutate(tavg = (tmn + tmx)/2, t_diurnal = (tmx-tmn))
head(davis_climate)
summary(davis_climate)

davis_climate_flint <- davis_climate %>% #get the means for the variables to compare to home sites 
  summarise(cwd_Davis=mean(cwd),ppt_Davis=mean(ppt), pck_Davis=mean(pck), tmn_Davis=mean(tmn), tmx_Davis=mean(tmx))
davis_climate_flint
```

## Davis Climate Trends

```{r}
davis_climate$month <- factor(davis_climate$month, levels = c("nov","dec","jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct"))

davis_climate %>% 
  ggplot(aes(x=month,y=cwd)) +
  geom_point()

davis_climate %>% 
  ggplot(aes(x=month,y=tmx)) +
  geom_point()

davis_climate %>% 
  ggplot(aes(x=month,y=tmn)) +
  geom_point()

davis_climate %>% 
  ggplot(aes(x=month,y=ppt)) +
  geom_point() 
```

### BioClim

Calculating wettest, driest, warmest, and coldest months

Note I called these "quarter" instead of month for merging with the home
climate quarter data.

```{r}
davis_wettest_quarter <- davis_climate %>%  
  slice_max(ppt)

davis_driest_quarter <- davis_climate %>% 
  slice_min(ppt)

davis_warmest_quarter <- davis_climate %>% 
  slice_max(tavg)

davis_coldest_quarter <- davis_climate %>%
  slice_min(tavg)
```

Bio 1, 2, 4, 7, 12, 15

```{r}
bioclim_davis_calc <- davis_climate %>% 
  summarise(ann_tmean=mean(tavg),  #Bio1 - Annual Mean Temperature
            mean_diurnal_range=mean(t_diurnal), #Bio2 - Mean Diurnal Range
            temp_seasonality=sd(tavg), #Bio4 - Temperature Seasonality
            temp_ann_range=(max(tmx))-(min(tmn)), #bio7 - temp annual range
            ann_ppt=sum(ppt), #bio12 - annual precip
            ppt_seasonality=cv(ppt+1)) #bio15 - Precipitation Seasonality (+1 to avoid strange CVs for areas where mean rainfaill is < 1)
bioclim_davis_calc
```

Bio 8(Q), 9(Q), 18(Q), 19(Q)

```{r}
#bio8 = tmean_wettest_quarter
bio8_davis <- davis_wettest_quarter %>% 
  dplyr::select(tmean_wettest_quarter=tavg)

#bio9 = tmean_driest_quarter
bio9_davis <- davis_driest_quarter %>% 
  dplyr::select(tmean_driest_quarter=tavg)

bio8_9_davis <- bind_cols(bio8_davis, bio9_davis)

#bio18 = ppt_warmest_quarter
bio18_davis <- davis_warmest_quarter %>% 
  dplyr::select(ppt_warmest_quarter=ppt)

#bio19 = ppt_coldest_quarter
bio19_davis <- davis_wettest_quarter %>% 
  dplyr::select(ppt_coldest_quarter=ppt)

bio18_19_davis <- bind_cols(bio18_davis, bio19_davis)

all_periods_davis <- bind_cols(bio8_9_davis, bio18_19_davis)
```

Merge all bioclims

```{r}
davis_bioclim_final <- bind_cols(bioclim_davis_calc, all_periods_davis) %>% 
  rename_with(~paste0(., "_Davis"), 1:10)
summary(davis_bioclim_final)
```

## Gower's Climate Distance

(1/P) \* SUM ((absolute value(Ai - Bi)) / range(i)) for each variable

-   P = number of environmental variables = 14

-   Ai = 30 year avg of that variable for the home site

-   Bi = Nov 2022-Oct 2023 avg of that variable for the Davis garden

-   Range(i) = maximum - minimum of that variable in the whole data set
    (across sites)

### Prep Davis data for range calculations 

```{r}
davis_climate_all <- bind_cols(davis_climate_flint, davis_bioclim_final)

davis_range_prep <- davis_climate_all %>% 
  mutate(parent.pop="Davis") %>% 
  rename_with(~str_remove(., "_Davis"), everything())
```

### Bootstrapping 
#### Load data with all 30 years
```{r}
recent_clim_boot <- read_csv("../output/Climate/fullyear_wtr_year_avgs_Recent.csv")

recent_clim_boot_nest <- recent_clim_boot %>% nest(.by=year) #nest to prepare for bootstrapping 
recent_clim_boot_nest

historical_clim_boot <- read_csv("../output/Climate/fullyear_wtr_year_avgs_Historical.csv")

historical_clim_boot_nest <- historical_clim_boot %>% nest(.by=year)
historical_clim_boot_nest
```

#### Create the gower_calc function 
Note: This code combines the code from chunks 14-17 below 
```{r}
#data <- recent_clim_boot_nest
#P=14

gowers_calc <- function(data, indices, P) { #function with all of the code necessary for calculating gowers distance 
  #data = _clim_boot (recent or historical) - needs to be nested by year; P = # climate variables 

  #need to make davis_range_prep before running this function 
  
  data <-data[indices,] # subset per bootstrap indices
  
  data <- data %>% unnest(data) #unnest so the function can access the climate data
  
  data_means <- data %>% 
    group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
    summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx", "ann_tmean", "mean_diurnal_range", 
                   "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_quarter", "tmean_driest_quarter", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_quarter", "ppt_coldest_quarter"),
               c(mean), na.rm = TRUE) #get 30 year averages for each climate variable 
  
  range_merge <- bind_rows(data_means, davis_range_prep)
  
  davis_home_climate_ranges <- range_merge %>% #calculate ranges
    ungroup() %>% 
  summarise(cwd_range=max(cwd)-min(cwd),
            ppt_range=max(ppt)-min(ppt), 
            tmn_range=max(tmn)-min(tmn), 
            tmx_range=max(tmx)-min(tmx), 
            ann_tmean_range=max(ann_tmean)-min(ann_tmean),
            mean_diurnal_range_range=max(mean_diurnal_range)-min(mean_diurnal_range),
            temp_seasonality_range=max(temp_seasonality)-min(temp_seasonality),
            temp_ann_range_range=max(temp_ann_range)-min(temp_ann_range),
            tmean_wettest_quarter_range=max(tmean_wettest_quarter)-min(tmean_wettest_quarter),
            tmean_driest_quarter_range=max(tmean_driest_quarter)-min(tmean_driest_quarter),
            ann_ppt_range=max(ann_ppt)-min(ann_ppt), 
            ppt_seasonality_range=max(ppt_seasonality)-min(ppt_seasonality),
            ppt_warmest_quarter_range=max(ppt_warmest_quarter)-min(ppt_warmest_quarter), 
            ppt_coldest_quarter_range=max(ppt_coldest_quarter)-min(ppt_coldest_quarter))
  
  davis_home_climate <- bind_cols(davis_climate_all, data_means) #add davis climate data to home climate data 
  
  davis_home_climate_with_ranges <- bind_cols(davis_home_climate, davis_home_climate_ranges) #add in ranges 
  
  gowers_calc_each_var <- davis_home_climate_with_ranges %>% #variable by variable calc
  mutate(cwd_gowers=abs(cwd_Davis-cwd) / cwd_range,
         ppt_gowers=abs(ppt_Davis - ppt) / ppt_range,
         tmn_gowers=abs(tmn_Davis - tmn) / tmn_range,
         tmx_gowers=abs(tmx_Davis - tmx) / tmx_range,
         ann_tmean_gowers=abs(ann_tmean_Davis - ann_tmean) / ann_tmean_range,
         mean_diurnal_range_gowers=abs(mean_diurnal_range_Davis - mean_diurnal_range) / mean_diurnal_range_range,
         temp_seasonality_gowers=abs(temp_seasonality_Davis - temp_seasonality) / temp_seasonality_range,
         temp_ann_range_gowers=abs(temp_ann_range_Davis - temp_ann_range) / temp_ann_range_range,
         tmean_wettest_quarter_gowers=abs(tmean_wettest_quarter_Davis - tmean_wettest_quarter) / tmean_wettest_quarter_range,
         tmean_driest_quarter_gowers=abs(tmean_driest_quarter_Davis - tmean_driest_quarter) / tmean_driest_quarter_range,
         ann_ppt_gowers=abs(ann_ppt_Davis - ann_ppt) / ann_ppt_range,
         ppt_seasonality_gowers=abs(ppt_seasonality_Davis - ppt_seasonality) / ppt_seasonality_range,
         ppt_warmest_quarter_gowers=abs(ppt_warmest_quarter_Davis - ppt_warmest_quarter) / ppt_warmest_quarter_range,
         ppt_coldest_quarter_gowers=abs(ppt_coldest_quarter_Davis - ppt_coldest_quarter) / ppt_coldest_quarter_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))

 gowers_calc_per_pop <- gowers_calc_each_var %>% #final gowers calc 
  mutate(Gowers_Dist=(1/P)*(cwd_gowers + ppt_gowers + tmn_gowers + tmx_gowers +
                                ann_tmean_gowers + mean_diurnal_range_gowers +
                                temp_seasonality_gowers +temp_ann_range_gowers +
                                tmean_wettest_quarter_gowers +
                                tmean_driest_quarter_gowers +ann_ppt_gowers +
                                ppt_seasonality_gowers + ppt_warmest_quarter_gowers +
                                ppt_coldest_quarter_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Gowers_Dist)
  
 gowers_calc_per_pop %>% pull(Gowers_Dist) #make the result a vector 
   }

#gowers_calc(recent_clim_boot_nest, P=14) #the function works
```

#### Perform the bootstrap sampling 
Recent
```{r}
gowers.boot_recent <- boot(data=recent_clim_boot_nest, statistic=gowers_calc, R=1000, P=14) #will sample each row (year) with replacement 
gowers.boot_recent$t0 #looks correct 
#str(gowers.boot_recent)

for(i in 1:23) {
  plot(gowers.boot_recent, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_recent, type="norm", index = 1) # for the first pop
boot_recent_results <- tidy(gowers.boot_recent,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Recent")
```

Historical
```{r}
gowers.boot_historical <- boot(data=historical_clim_boot_nest, statistic=gowers_calc, R=1000, P=14) #will sample each row (year) with replacement 
gowers.boot_historical$t0 #looks correct 
#str(gowers.boot_historical)

for(i in 1:23) {
  plot(gowers.boot_historical, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_historical, type="norm", index = 1) # for the first pop
boot_historical_results <- tidy(gowers.boot_historical,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Historical")
```

#### Plot the results 
```{r}
pops <- recent_clim_boot %>% select(parent.pop, elevation.group, elev_m, Lat, Long) %>% unique()

boot_gowers_recent_pops <- bind_cols(pops, boot_recent_results) %>% arrange(Gowers_Dist)
boot_gowers_historical_pops <- bind_cols(pops, boot_historical_results) %>% arrange(Gowers_Dist)
boot_gowers_results_all <- bind_rows(boot_gowers_recent_pops, boot_gowers_historical_pops)

boot_gowers_results_all %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from Davis", fill="Elevation (m)", x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) +
  facet_wrap(~TimePd)
ggsave("../output/Climate/full_year_Gowers_fromDavis_wtr_year.png", width = 24, height = 8, units = "in")
```

Historical
```{r}
gowers.boot_historical <- boot(data=historical_clim_boot_nest, statistic=gowers_calc, R=1000, P=14) #will sample each row (year) with replacement 
gowers.boot_historical$t0 #looks correct 
#str(gowers.boot_historical)

for(i in 1:23) {
  plot(gowers.boot_historical, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_historical, type="norm", index = 1) # for the first pop
boot_historical_results <- tidy(gowers.boot_historical,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Historical")
```

#### Plot the results 
```{r}
pops <- recent_clim_boot %>% select(parent.pop, elevation.group, elev_m, Lat, Long) %>% unique()

boot_gowers_recent_pops <- bind_cols(pops, boot_recent_results) %>% arrange(Gowers_Dist)
boot_gowers_historical_pops <- bind_cols(pops, boot_historical_results) %>% arrange(Gowers_Dist)
boot_gowers_results_all <- bind_rows(boot_gowers_recent_pops, boot_gowers_historical_pops)
write_csv(boot_gowers_results_all, "../output/Climate/full_year_GowersEnvtalDist_UCD_wtr_year.csv")

boot_gowers_results_all %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from Davis", fill="Elevation (m)", x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) +
  facet_wrap(~TimePd)
ggsave("../output/Climate/full_year_Gowers_fromDavis_wtr_year.png", width = 24, height = 8, units = "in")
```

## Flint Climate Distance

### Gowers
```{r}
#need to modify the gowers_calc function to only include Flint variables 
gowers_calc_flint <- function(data, indices, P) { #function with all of the code necessary for calculating gowers distance 
  #data = _clim_boot (recent or historical) - needs to be nested by year; P = # climate variables 
  #need to make davis_range_prep before running this function 
  
  data <-data[indices,] # subset per bootstrap indices
  
  data <- data %>% unnest(data) #unnest so the function can access the climate data
  
  data_means <- data %>% 
    group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
    summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx"),
               c(mean), na.rm = TRUE) #get 30 year averages for each climate variable 
  
  range_merge <- bind_rows(data_means, davis_range_prep)
  
  davis_home_climate_ranges <- range_merge %>% #calculate ranges
    ungroup() %>% 
  summarise(cwd_range=max(cwd)-min(cwd),
            ppt_range=max(ppt)-min(ppt), 
            tmn_range=max(tmn)-min(tmn), 
            tmx_range=max(tmx)-min(tmx))
  
  davis_home_climate <- bind_cols(davis_climate_all, data_means) #add davis climate data to home climate data 
  
  davis_home_climate_with_ranges <- bind_cols(davis_home_climate, davis_home_climate_ranges) #add in ranges 
  
  gowers_calc_each_var <- davis_home_climate_with_ranges %>% #variable by variable calc
  mutate(cwd_gowers=abs(cwd_Davis-cwd) / cwd_range,
         ppt_gowers=abs(ppt_Davis - ppt) / ppt_range,
         tmn_gowers=abs(tmn_Davis - tmn) / tmn_range,
         tmx_gowers=abs(tmx_Davis - tmx) / tmx_range)

 gowers_calc_per_pop <- gowers_calc_each_var %>% #final gowers calc 
  mutate(Gowers_Dist=(1/P)*(cwd_gowers + ppt_gowers + tmn_gowers + tmx_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Gowers_Dist)
 
 gowers_calc_per_pop %>% pull(Gowers_Dist) #make the result a vector 
   }
```

#### Recent

```{r}
recent_clim_boot_flint_nest <- recent_clim_boot %>% select(parent.pop:tmx) %>% nest(.by=year)

gowers.boot_flint_recent <- boot(data=recent_clim_boot_flint_nest, statistic=gowers_calc_flint, R=1000, P=5) #will sample each row (year) with replacement 
gowers.boot_flint_recent$t0 #looks correct 
#str(gowers.boot_flint_recent)

for(i in 1:23) {
  plot(gowers.boot_flint_recent, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_flint_recent, type="norm", index = 1) # for the first pop
boot_flint_recent_results <- tidy(gowers.boot_flint_recent,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Recent")
```

#### Historic

```{r}
historical_clim_boot_flint_nest <- historical_clim_boot %>% select(parent.pop:tmx) %>% nest(.by=year)

gowers.boot_flint_historical <- boot(data=historical_clim_boot_flint_nest, statistic=gowers_calc_flint, R=1000, P=5) #will sample each row (year) with replacement 
gowers.boot_flint_historical$t0 #looks correct 
#str(gowers.boot_flint_historical)

for(i in 1:23) {
  plot(gowers.boot_flint_historical, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_flint_historical, type="norm", index = 1) # for the first pop
boot_flint_historical_results <- tidy(gowers.boot_flint_historical,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Historical")
```

#### Plot the results 

```{r}
boot_flint_gowers_recent_pops <- bind_cols(pops, boot_flint_recent_results) %>% arrange(Gowers_Dist)
boot_flint_gowers_historical_pops <- bind_cols(pops, boot_flint_historical_results) %>% arrange(Gowers_Dist)
boot_flint_gowers_results_all <- bind_rows(boot_flint_gowers_recent_pops, boot_flint_gowers_historical_pops)
write_csv(boot_flint_gowers_results_all, "../output/Climate/full_year_GowersEnvtalDist_UCDFlint_wtr_year.csv")

boot_flint_gowers_results_all %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from Davis", fill="Elevation (m)", x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) +
  facet_wrap(~TimePd)
ggsave("../output/Climate/full_year_Gowers_Flint_fromDavis_wtr_year.png", width = 24, height = 8, units = "in")
```

### Subtraction

```{r}
pops_flint_avgs <- read_csv("../output/Climate/fullyear_FlintAvgs_wtr_year.csv")
pops_flint_recent_avgs <- pops_flint_avgs %>% filter(TimePd=="Recent")
recent_flint_dist_prep <- bind_cols(davis_climate_flint, pops_flint_recent_avgs)
names(recent_flint_dist_prep)

recent_flint_dist <- recent_flint_dist_prep %>% 
  mutate(cwd_dist=cwd_Davis - cwd,
         ppt_dist=ppt_Davis - ppt,
         tmn_dist=tmn_Davis - tmn,
         tmx_dist=tmx_Davis - tmx) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))

pops_flint_historic_avgs <-  pops_flint_avgs %>% filter(TimePd=="Historical")
historic_flint_dist_prep <- bind_cols(davis_climate_flint, pops_flint_historic_avgs)
names(historic_flint_dist_prep)

historic_flint_dist <- historic_flint_dist_prep %>% 
  mutate(cwd_dist=cwd_Davis - cwd,
         ppt_dist=ppt_Davis - ppt,
         tmn_dist=tmn_Davis - tmn,
         tmx_dist=tmx_Davis - tmx) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
```

Figures Recent (subtraction distance)

```{r}
recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, cwd_dist), y=cwd_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanCWD_DistfromDavis_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_dist), y=ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanPPT_DistfromDavis_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmn_dist), y=tmn_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanTMN_DistfromDavis_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmx_dist), y=tmx_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanTMX_DistfromDavis_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")
```

Figures Historical (subtraction distance)

```{r}
historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, cwd_dist), y=cwd_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanCWD_DistfromDavis_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_dist), y=ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanPPT_DistfromDavis_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmn_dist), y=tmn_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanTMN_DistfromDavis_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmx_dist), y=tmx_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanTMX_DistfromDavis_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")
```

## Bioclim Climate Distance

### Gowers
```{r}
#need to modify the gowers_calc function to only include bioclim variables 
gowers_calc_bioclim <- function(data, indices, P) { #function with all of the code necessary for calculating gowers distance 
  #data = _clim_boot (recent or historical) - needs to be nested by year; P = # climate variables 
  #need to make davis_range_prep before running this function 
  
  data <-data[indices,] # subset per bootstrap indices
  
  data <- data %>% unnest(data) #unnest so the function can access the climate data
  
  data_means <- data %>% 
    group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
    summarise_at(c("ann_tmean", "mean_diurnal_range", 
                   "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_quarter", "tmean_driest_quarter", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_quarter", "ppt_coldest_quarter"),
               c(mean), na.rm = TRUE) #get 30 year averages for each climate variable 
  
  range_merge <- bind_rows(data_means, davis_range_prep)
  
  davis_home_climate_ranges <- range_merge %>% #calculate ranges
    ungroup() %>% 
  summarise(ann_tmean_range=max(ann_tmean)-min(ann_tmean),
            mean_diurnal_range_range=max(mean_diurnal_range)-min(mean_diurnal_range),
            temp_seasonality_range=max(temp_seasonality)-min(temp_seasonality),
            temp_ann_range_range=max(temp_ann_range)-min(temp_ann_range),
            tmean_wettest_quarter_range=max(tmean_wettest_quarter)-min(tmean_wettest_quarter),
            tmean_driest_quarter_range=max(tmean_driest_quarter)-min(tmean_driest_quarter),
            ann_ppt_range=max(ann_ppt)-min(ann_ppt), 
            ppt_seasonality_range=max(ppt_seasonality)-min(ppt_seasonality),
            ppt_warmest_quarter_range=max(ppt_warmest_quarter)-min(ppt_warmest_quarter), 
            ppt_coldest_quarter_range=max(ppt_coldest_quarter)-min(ppt_coldest_quarter))
  
  davis_home_climate <- bind_cols(davis_climate_all, data_means) #add davis climate data to home climate data 
  
  davis_home_climate_with_ranges <- bind_cols(davis_home_climate, davis_home_climate_ranges) #add in ranges 
  
  gowers_calc_each_var <- davis_home_climate_with_ranges %>% #variable by variable calc
  mutate(ann_tmean_gowers=abs(ann_tmean_Davis - ann_tmean) / ann_tmean_range,
         mean_diurnal_range_gowers=abs(mean_diurnal_range_Davis - mean_diurnal_range) / mean_diurnal_range_range,
         temp_seasonality_gowers=abs(temp_seasonality_Davis - temp_seasonality) / temp_seasonality_range,
         temp_ann_range_gowers=abs(temp_ann_range_Davis - temp_ann_range) / temp_ann_range_range,
         tmean_wettest_quarter_gowers=abs(tmean_wettest_quarter_Davis - tmean_wettest_quarter) / tmean_wettest_quarter_range,
         tmean_driest_quarter_gowers=abs(tmean_driest_quarter_Davis - tmean_driest_quarter) / tmean_driest_quarter_range,
         ann_ppt_gowers=abs(ann_ppt_Davis - ann_ppt) / ann_ppt_range,
         ppt_seasonality_gowers=abs(ppt_seasonality_Davis - ppt_seasonality) / ppt_seasonality_range,
         ppt_warmest_quarter_gowers=abs(ppt_warmest_quarter_Davis - ppt_warmest_quarter) / ppt_warmest_quarter_range,
         ppt_coldest_quarter_gowers=abs(ppt_coldest_quarter_Davis - ppt_coldest_quarter) / ppt_coldest_quarter_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))

 gowers_calc_per_pop <- gowers_calc_each_var %>% #final gowers calc 
  mutate(Gowers_Dist=(1/P)*(ann_tmean_gowers + mean_diurnal_range_gowers +
                                temp_seasonality_gowers +temp_ann_range_gowers +
                                tmean_wettest_quarter_gowers +
                                tmean_driest_quarter_gowers +ann_ppt_gowers +
                                ppt_seasonality_gowers + ppt_warmest_quarter_gowers +
                                ppt_coldest_quarter_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Gowers_Dist)
 
 gowers_calc_per_pop %>% pull(Gowers_Dist) #make the result a vector 
   }
```

#### Recent

```{r}
recent_clim_boot_bioclim_nest <- recent_clim_boot %>% select(parent.pop:year, ann_tmean:ppt_coldest_quarter) %>% nest(.by=year)

gowers.boot_bioclim_recent <- boot(data=recent_clim_boot_bioclim_nest, statistic=gowers_calc_bioclim, R=1000, P=10) #will sample each row (year) with replacement 
gowers.boot_bioclim_recent$t0 #looks correct 
#str(gowers.boot_bioclim_recent)

for(i in 1:23) {
  plot(gowers.boot_bioclim_recent, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_bioclim_recent, type="norm", index = 1) # for the first pop
boot_bioclim_recent_results <- tidy(gowers.boot_bioclim_recent,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Recent")
```

#### Historic

```{r}
historical_clim_boot_bioclim_nest <- historical_clim_boot %>% select(parent.pop:year, ann_tmean:ppt_coldest_quarter) %>% nest(.by=year)

gowers.boot_bioclim_historical <- boot(data=historical_clim_boot_bioclim_nest, statistic=gowers_calc_bioclim, R=1000, P=10) #will sample each row (year) with replacement 
gowers.boot_bioclim_historical$t0 #looks correct 
#str(gowers.boot_bioclim_historical)

for(i in 1:23) {
  plot(gowers.boot_bioclim_historical, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_bioclim_historical, type="norm", index = 1) # for the first pop
boot_bioclim_historical_results <- tidy(gowers.boot_bioclim_historical,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Historical")
```

#### Plot the results 

```{r}
boot_bioclim_gowers_recent_pops <- bind_cols(pops, boot_bioclim_recent_results) %>% arrange(Gowers_Dist)
boot_bioclim_gowers_historical_pops <- bind_cols(pops, boot_bioclim_historical_results) %>% arrange(Gowers_Dist)
boot_bioclim_gowers_results_all <- bind_rows(boot_bioclim_gowers_recent_pops, boot_bioclim_gowers_historical_pops)
write_csv(boot_bioclim_gowers_results_all, "../output/Climate/full_year_GowersEnvtalDist_UCDbioclim_wtr_year.csv")

boot_bioclim_gowers_results_all %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from Davis", fill="Elevation (m)", x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) +
  facet_wrap(~TimePd)
ggsave("../output/Climate/full_year_Gowers_BioClim_fromDavis_wtr_year.png", width = 24, height = 8, units = "in")
```

### Subtraction

```{r}
pops_bioclim_avgs <-  read_csv("../output/Climate/fullyear_BioClimAvgs_wtr_year.csv") 
#Recent
pops_bioclim_recent_avgs <- pops_bioclim_avgs %>%  filter(TimePd=="Recent")
recent_bioclim_dist_prep <- bind_cols(davis_bioclim_final, pops_bioclim_recent_avgs)
recent_bioclim_dist <- recent_bioclim_dist_prep %>% 
  mutate(ann_tmean_dist=ann_tmean_Davis - ann_tmean,
         mean_diurnal_range_dist=mean_diurnal_range_Davis - mean_diurnal_range,
         temp_seasonality_dist=temp_seasonality_Davis - temp_seasonality,
         temp_ann_range_dist=temp_ann_range_Davis - temp_ann_range,
         tmean_wettest_quarter_dist=tmean_wettest_quarter_Davis - tmean_wettest_quarter,
         tmean_driest_quarter_dist=tmean_driest_quarter_Davis - tmean_driest_quarter,
         ann_ppt_dist=ann_ppt_Davis - ann_ppt,
         ppt_seasonality_dist=ppt_seasonality_Davis - ppt_seasonality, 
         ppt_warmest_quarter_dist=ppt_warmest_quarter_Davis - ppt_warmest_quarter,
         ppt_coldest_quarter_dist=ppt_coldest_quarter_Davis - ppt_coldest_quarter) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
recent_bioclim_dist

#Historical
pops_bioclim_historical_avgs <- pops_bioclim_avgs %>% filter(TimePd=="Historical")
historical_bioclim_dist_prep <- bind_cols(davis_bioclim_final, pops_bioclim_historical_avgs)
historical_bioclim_dist <- historical_bioclim_dist_prep %>% 
  mutate(ann_tmean_dist=ann_tmean_Davis - ann_tmean,
         mean_diurnal_range_dist=mean_diurnal_range_Davis - mean_diurnal_range,
         temp_seasonality_dist=temp_seasonality_Davis - temp_seasonality,
         temp_ann_range_dist=temp_ann_range_Davis - temp_ann_range,
         tmean_wettest_quarter_dist=tmean_wettest_quarter_Davis - tmean_wettest_quarter,
         tmean_driest_quarter_dist=tmean_driest_quarter_Davis - tmean_driest_quarter,
         ann_ppt_dist=ann_ppt_Davis - ann_ppt,
         ppt_seasonality_dist=ppt_seasonality_Davis - ppt_seasonality, 
         ppt_warmest_quarter_dist=ppt_warmest_quarter_Davis - ppt_warmest_quarter,
         ppt_coldest_quarter_dist=ppt_coldest_quarter_Davis - ppt_coldest_quarter) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
historical_bioclim_dist
```

Figures Recent (subtraction distance)

```{r}
recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_dist), y=ann_tmean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Ann_Tmean_DistfromDavis_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_dist), y=mean_diurnal_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Diurnal_Range_DistfromDavis_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_dist), y=temp_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Seasonality_DistfromDavis_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_dist), y=temp_ann_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Ann_Range_DistfromDavis_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_quarter_dist), y=tmean_wettest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Wet_DistfromDavis_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_quarter_dist), y=tmean_driest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Dry_DistfromDavis_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_dist), y=ann_ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Ann_PPT_DistfromDavis_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_dist), y=ppt_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_PPT_Seasonality_DistfromDavis_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_quarter_dist), y=ppt_warmest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_PPT_Warm_DistfromDavis_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_quarter_dist), y=ppt_coldest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_PPT_Cold_DistfromDavis_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")
```

Historical (subtraction distance)

```{r}
historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_dist), y=ann_tmean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Ann_Tmean_DistfromDavis_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_dist), y=mean_diurnal_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Diurnal_Range_DistfromDavis_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_dist), y=temp_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Seasonality_DistfromDavis_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_dist), y=temp_ann_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Ann_Range_DistfromDavis_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_quarter_dist), y=tmean_wettest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Wet_DistfromDavis_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_quarter_dist), y=tmean_driest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Dry_DistfromDavis_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_dist), y=ann_ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Ann_PPT_DistfromDavis_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_dist), y=ppt_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_PPT_Seasonality_DistfromDavis_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_quarter_dist), y=ppt_warmest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_PPT_Warm_DistfromDavis_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_quarter_dist), y=ppt_coldest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_PPT_Cold_DistfromDavis_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")
```
