---
title: "UCD_SurvAnalysis_Pops_GLMER"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Decide on base models to use for climate 

# Survival Analysis

## Libraries

```{r}
# install.packages("multilevelmod")
library(tidyverse) 
library(magrittr)
library(ggrepel)
library(lme4)
library(ggdist)
library(tidymodels)
library(multilevelmod)
library(broom.mixed)
library(modelr)
library(furrr)
library(zoo)
tidymodels_prefer()
options(mc.cores = parallel::detectCores())
```

## Read in the plant data
```{r}
ucd_surv <- read_csv("../input/UCD_Data/CorrectedCSVs/UCD_transplants_pheno_mort_20231016_corrected.csv") %>% 
  filter(rep != 100) #get rid of individuals that germinated in the field
head(ucd_surv)
unique(ucd_surv$pop) #pops look correct 
```

As weeks
```{r}
ucd_surv %>% filter(is.na(`Death Date`),!is.na(Notes)) #no plants that seemed to disappear 

ucd_surv_dates <- ucd_surv %>%  
  rename(parent.pop = pop, death.date=`Death Date`) %>%
  unite("Genotype", parent.pop:rep, remove = FALSE) %>% 
  filter(!is.na(parent.pop)) %>% 
  mutate(planting.date="11/30/22", #could try to make this more specific to when certain blocks were planted 
         last_fup_date=if_else(is.na(death.date), "10/16/23", death.date)) %>%  #need this to calculate survival times
  mutate(planting.date=mdy(planting.date), last_fup_date=mdy(last_fup_date)) %>% #convert to date objects
  mutate(os_weeks=as.duration(planting.date %--% last_fup_date) / dweeks(1), #observed number of weeks
         status=if_else(is.na(death.date), 0, 1)) %>% #0=censured (alive in this case), 1=dead
  filter(os_weeks > 0) %>% #remove any that died at planting since this is survival post- transplanting
  select(block:rep, death.date, planting.date:status)
head(ucd_surv_dates)
unique(ucd_surv_dates$os_weeks) %>% sort()
```

## Location Info
```{r}
ucd_gowers <- read_csv("../output/Climate/Gowers_UCD.csv") %>% 
  select(parent.pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  mutate(Garden_Lat=38.53250, Garden_Long=-121.7830, Garden_Elev=16) %>% 
  mutate(Geographic_Dist=distHaversine(cbind(Garden_Long, Garden_Lat), cbind(Long, Lat)),
         Elev_Dist=elev_m-Garden_Elev) # Calculate the distance using the haversine formula (dist in meters)
```

```{r}
ucd_surv_dates_loc <- left_join(ucd_surv_dates, ucd_gowers)
```


## Reformat to be one row per plant per week

```{r, eval=TRUE}
ucd_ind <- ucd_surv_dates %>% 
  complete(Genotype, last_fup_date) %>%
  group_by(Genotype) %>%
  mutate(across(c(-status, -last_fup_date, -os_weeks, -planting.date), \(x) ifelse(all(is.na(unique(x))), NA,unique(na.omit(x))))) %>%
  mutate(planting.date = unique(na.omit(planting.date)),
         os_weeks=as.duration(planting.date %--% last_fup_date) / dweeks(1),
           dead = ifelse((is.na(death.date) | mdy(death.date) > last_fup_date), 0, 1)) %>% # 0 is alive, 1 is dead
  filter(cumsum(dead) < 2) %>% # you can't die more than once
  ungroup() %>%  
  left_join(ucd_gowers)

ucd_ind %>% select(Genotype, last_fup_date, os_weeks, dead)
```

# Modeling probability of event rather than event time

An alternative approach to hazard analysis is to model the probably of death as a function of time.  This is approach taken here.

## Discrete intervals

Modify the input data to add some factor variables
```{r}
ucd_ind <- ucd_ind %>% mutate(
  os_weeks_f = factor(str_pad(round(os_weeks, 1), 2, pad = "0")), # padding so it will sort correctly
  os_weeks_pop = factor(str_c(parent.pop,"_", os_weeks_f))
)
ucd_ind
```

Remove Establishment Phase 
```{r}
ucd_ind_surv <- ucd_ind %>% filter(os_weeks > 3) %>%
  mutate(os_weeks_f = factor(os_weeks))

ucd_surv_means <- ucd_ind_surv %>%
  group_by(parent.pop, os_weeks) %>%
  summarize(survival = 1-mean(dead))
```

# Survival data set

```{r}
glmer.model <- 
  linear_reg() %>%
  set_engine("glmer", family = binomial)
```

check basic model forms
```{r}
# Note: StdDev due to mf within parent.pop is essentially zero, so I am omitting this.
##mod_test <- glmer(dead ~ (1|parent.pop/mf) + (1|block), data=ucd_ind_surv, family=binomial)
##summary(mod_test)


surv_wflow <- workflow() %>%
  add_variables(outcomes = dead, predictors = c(parent.pop, mf, block, os_weeks, os_weeks_f)) 

surv_fits_1 <- tibble(wflow=list(

  pop.block = {surv_wflow %>% 
      add_model(glmer.model, formula = dead ~ (1|parent.pop) + (1|block))},
  
  pop.weeks_f.block ={surv_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks_f + (1|parent.pop) + (1|block))},

  pop.weeks.block ={surv_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + (1|parent.pop) + (1|block))},

  pop.weeks.quad.block ={surv_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + I(os_weeks^2) + (1|parent.pop) + (1|block))}
      
  # Models commented out fail to converge, so not running them.
  # Any model that has a RE slope temer for weeks fails.
  # Note:only having os_weeks as RE slope does not fix this.
  
  # pop.weeks_f.int ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks_f + (os_weeks_f|parent.pop))},
  # 
  # pop.weeks_f.int.block ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks_f + (os_weeks_f|parent.pop) + (1|block))},
  # 

  # pop.weeks.int ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks + (os_weeks_f|parent.pop))},
  # 
  # pop.weeks.int.block ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks + (os_weeks_f|parent.pop) + (1|block))},

  # Models below fail to converge, so not running them.
  # Note:only having os_weeks as RE slope does not fix this.
  
  # pop.weeks.quad.int ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks + I(os_weeks^2) + (os_weeks_f|parent.pop))},
  # 
  # pop.weeks.quad.quad.int ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks + I(os_weeks^2) + (os_weeks_f + I(os_weeks^2)|parent.pop))},
  # 
  # pop.weeks.quad.int.block ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks + I(os_weeks^2) + (os_weeks_f|parent.pop) + (1|block))},
  # 
  # pop.weeks.quad.quad.int.block ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks + I(os_weeks^2) + (os_weeks_f + I(os_weeks^2)|parent.pop) + (1|block))}
    
),
name=names(wflow)
) %>% 
  select(name,wflow)
surv_fits_1

#mod_test <- glmer(dead ~ os_weeks + I(os_weeks^2) + (1|parent.pop) + (1|block), data=ucd_ind_surv, family=binomial)
#Warning: Model failed to converge with max|grad| = 0.00241906 (tol = 0.002, component 1) - pop.weeks_f.block model
#Warning: Model failed to converge with max|grad| = 0.0174841 (tol = 0.002, component 1)Warning: Model is nearly unidentifiable: very large eigenvalue - pop.weeks.quad.block
```

```{r}
system.time(
  surv_fits_1 <- surv_fits_1 %>%
  mutate(fit = map(wflow, fit, data = ucd_ind_surv))
)
```
get warnings
```{r}
surv_fits_1 <- surv_fits_1 %>%
  mutate(message=map_chr(fit, \(x) {extract_fit_engine(x) %>% .@optinfo %>% magrittr::extract2("message")}))

surv_fits_1 %>% select(name, message) %>%
  arrange(message, name)
```
A little strange since I get a warning when running about failure to converge for the model with os_weeks_f

```{r}
surv_fits_1 %>% pull(fit)
```

```{r}
surv_fits_1 %>% 
  filter(!str_detect(message, "failure")) %>%
  mutate(glance=map(fit, glance)) %>% 
  unnest(glance) %>% 
  arrange(BIC) %>%
  select(-wflow:-sigma)
```

All models with a random slope for time for each pop don't converge.

get bootstrap predictions and plot
```{r}
surv_nd <- data_grid(ucd_ind_surv, parent.pop, os_weeks) %>%
  mutate(os_weeks_f = factor(os_weeks))

surv_predict_fun <- function(m, nd = surv_nd) {
  predict(m, 
          newdata = nd,
          type = "response",
          re.form = ~(1|parent.pop)
  ) %>%
    subtract(1, .) # convert hazard to survival
}

# ~975 seconds for 100.  Need to go somewhere other than my laptop for more
system.time (surv_fits_1 <- surv_fits_1 %>% 
                 filter(!str_detect(message, "failure")) %>%
               mutate(
                 bootstrap = map(fit, \(x) {
                   extract_fit_engine(x) %>% 
                     bootMer(FUN = surv_predict_fun,
                             re.form = ~ (1|parent.pop),
                             nsim = 100,
                             parallel = "multicore",
                             ncpus = 7
                     ) } #bootMer
                 ),  
                 tidyboot = map(bootstrap, tidy, conf.int = TRUE),
                 tidyboot = map(tidyboot, \(x) cbind(surv_nd, x))
               ) # mutate
) # system.time
```
some bootstrap runs failed (33/100) 

```{r}
save.image("../output/JM_GLMER_UCD.Rdata")
```

```{r}
surv_fits_1 %>%
  select(name, tidyboot) %>%
  unnest(tidyboot) %>%
  mutate(survival = statistic) %>%
  ggplot(aes(x=os_weeks, y = survival)) +
  geom_line(aes(color=name)) +
  geom_point(data = ucd_surv_means, pch=1) + 
  facet_wrap(~parent.pop) +
  scale_color_viridis_d() # +
  #theme(axis.text.x = element_text(angle=90, hjust=1, vjust=.5))
  
```

Go with pop.weeks.block - best of the two models that actually converged

### is Gower's distance an important predictor? 

Also: check mf for final model

```{r}
surv_GD_wflow <- workflow() %>%
  add_variables(outcomes = dead, predictors = c(parent.pop, mf, block, os_weeks,  contains("GD"))) 

surv_GD_fits <- tibble(wflow=list(

  pop.weeks.block = {surv_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + (1|parent.pop) + (1|block))},
  
  pop.mf.weeks.block = {surv_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + (1|parent.pop/mf) + (1|block))},
  
  GS_Recent = {surv_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + GrwSsn_GD_Recent + (1|parent.pop) + (1|block))},
  
  GS_Historical = {surv_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + GrwSsn_GD_Historical + (1|parent.pop) + (1|block))},
  
  WY_Recent = {surv_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + Wtr_Year_GD_Recent + (1|parent.pop) + (1|block))},
  
  WY_Historical = {surv_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + Wtr_Year_GD_Historical + (1|parent.pop) + (1|block))}

),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = ucd_ind_surv))

```

```{r}
surv_GD_fits %>% 
  mutate(glance=map(fit, glance)) %>% 
  unnest(glance) %>% 
  arrange(BIC) %>%
  select(-wflow:-sigma)
```

Model with mf better by AIC and BIC
```{r}
mod_test <- glmer(dead ~ os_weeks + (1|parent.pop/mf) + (1|block), data=ucd_ind_surv, family=binomial)
summary(mod_test)
```


```{r}
surv_GD_fits %>% 
  mutate(tidy=map(fit, tidy)) %>% 
  unnest(tidy) %>% 
  filter(str_detect(term, "GD")) %>%
  select(-wflow:-group) %>%
  arrange(p.value)
```