---
title: "Data Checking"
output: html_notebook:
  keep_md: yes
---
# Annual Census - Transplants 
```{r}
#load the libraries
library(tidyverse)
#tutorial at https://bookdown.org/ansellbr/WEHI_tidyR_course_book/week-4-part-3.html
data_files <- list.files('~/Google Drive/Shared drives/IntBioTeam/Common Gardens/UCD2022_2023/CorrectedCSVs',pattern='Annual_Census_Transplants',ignore.case=TRUE)
data_files #make this specific to mortality or size

data_names <- str_remove(data_files, pattern = ".csv")
data_names

for (i in data_names) {
  #set up variable file names
  in_file <- paste("~/Google Drive/Shared drives/IntBioTeam/Common Gardens/UCD2022_2023/CorrectedCSVs/", i, '.csv', sep = "")   #edit this to be correct path 
  out_table <- paste("../output/", i, '_dataprobs.csv', sep="")
  
  #data checking code
  data <- read_csv(in_file)
  
  #remove spaces in column names
  names(data) <- gsub(" ", "", colnames(data))
  
  #check for spaces in numbers
  data2 <- data %>% filter(grepl(" ", diam_mm) | grepl(" ", height_cm) | grepl(" ", total_branch) | grepl(" ", longest_leaf_cm) | grepl(" ", flowers) | grepl(" ", fruits) | grepl(" ", longest_fruit_cm) | grepl(" ", repro_branch)) %>% mutate(problem="spaces")
  
  #check for extreme values
  data3 <- data
  data3 <- data3 %>% filter(diam_mm > 50, height_cm > 140 | longest_leaf_cm > 12, longest_fruit_cm > 15) %>% mutate(problem="extreme value")
  
  #check for letters in number columns 
  data4 <- data %>% rowwise() %>% filter(!is.na(diam_mm)) %>% filter(is.na(as.numeric(diam_mm))) %>% mutate(problem="character")
  data5 <- data %>% rowwise() %>% filter(!is.na(height_cm)) %>% filter(is.na(as.numeric(height_cm))) %>% mutate(problem="character")
  data6 <- data %>% rowwise() %>% filter(!is.na(total_branch)) %>% filter(is.na(as.numeric(total_branch))) %>% mutate(problem="character")
  data7 <- data %>% rowwise() %>% filter(!is.na(longest_leaf_cm)) %>% filter(is.na(as.numeric(longest_leaf_cm))) %>% mutate(problem="character")
  data8 <- data %>% rowwise() %>% filter(!is.na(flowers)) %>% filter(is.na(as.numeric(flowers))) %>% mutate(problem="character")
  data9 <- data %>% rowwise() %>% filter(!is.na(fruits)) %>% filter(is.na(as.numeric(fruits))) %>% mutate(problem="character")
  data10 <- data %>% rowwise() %>% filter(!is.na(longest_fruit_cm)) %>% filter(is.na(as.numeric(longest_fruit_cm))) %>% mutate(problem="character")
  data11 <- data %>% rowwise() %>% filter(!is.na(repro_branch)) %>% filter(is.na(as.numeric(repro_branch))) %>% mutate(problem="character")
  
  #combine into one problems list
  data_probs <- rbind(data2, data3, data4, data5, data6, data7, data8, data9, data10, data11)
  print(nrow(data_probs))
  write_csv(data_probs, path=out_table)
}
```

# Size Data - Transplants (temporal check)
Check for data when there is no data in previous week 
* Reason: found a case where someone wrote the data in the wrong row (in the row of a dead plant) 

```{r}
files <- dir("../input/UCD_Data/CorrectedCSVs/",
             pattern="Size_survey_transplants.*corrected.*csv",
             full.names=TRUE)
files #list of files 

dat <- tibble(path=files, filename=basename(path)) #storing data in tibble
dat

# create a new column, "date"
dat2 <- dat %>%
  mutate(survey_date=str_extract(filename, "2[0-9]*"),
         survey_date=lubridate::ymd(survey_date))
dat2

#read in the files 
dat3 <- dat2 %>% 
  mutate(sheets=map(path, 
                    read_csv, 
                    col_types=cols(.default=col_character()), # forces all columns to be character
    # this is needed because of data sheet problems that cause some numerics to be imported as character  
                    na = c("", "NA", "-", "N/A") # sets NA strings.
                    )) %>%
  select(-path)

dat3

dat3$sheets[[1]] # extract the first imported sheet to check that the above worked

map(dat3$sheets, colnames) %>% #check column names before merging 
  unlist() %>%
  unique() %>%
  sort()
#have some extra columns ("...10" and "...12"), double check these after unnest 
#have some weird date columns, but not important b/c we have survey_date from the file name 
#have "col" and "column", maybe important 
#have both "parent.pop" and "pop", might be important depending how I code the rest

map(dat3$sheets, head,10) #see which ones have different names
dat3$sheets[[1]] <- dat3$sheets[[1]] %>% rename(pop=parent.pop, col = column) #fix the name in the first sheet
dat3$sheets[[1]]

#unnest
dat4 <- dat3 %>% unnest(sheets) %>%
  rename(height_cm = `height (cm)`, longest_leaf_cm = `longest leaf (cm)`, death_date=`death date`)
head(dat4)
names(dat4)
dat4 %>% filter(!is.na(...10)) #an additional note for one plant 
dat4 %>% filter(!is.na(...12)) #all NA

#check for data issues
dat4 %>% rowwise() %>% filter(!is.na(row)) %>% filter(is.na(as.numeric(row))) #none
dat4 %>% rowwise() %>% filter(!is.na(mf)) %>% filter(is.na(as.numeric(mf))) #only weird for a buffer
dat4 %>% rowwise() %>% filter(!is.na(rep)) %>% filter(is.na(as.numeric(rep))) #none
dat4 %>% rowwise() %>% filter(!is.na(height_cm)) %>% filter(is.na(as.numeric(height_cm))) #only weird for a dead plant 
dat4 %>% rowwise() %>% filter(!is.na(longest_leaf_cm)) %>% filter(is.na(as.numeric(longest_leaf_cm))) #none
unique(dat4$pop) #many versions of LVTR1 
dat4 %>% filter(is.na(pop), !is.na(height_cm))

unique(dat4$mf)
dat4 %>% filter(mf==".") #on a buffer
dat4 %>% filter(is.na(mf), !is.na(height_cm)) #on a buffer
dat4 %>% filter(mf < 0) #H1 12 B, mf = -2, fixed it in the Corrected CSV

unique(dat4$rep) #some rep numbers = 100 
dat4 %>% filter(is.na(rep), !is.na(height_cm)) #on a buffer
dat4 %>% filter(as.numeric(rep)==100) 
#L1 5 A - original died, new one germinated, so rep changed to 100
#L2 37 C, should be SQ2 6 1 NOT 100 (from 03-03 to 07-24) - original died and new one germinated later
#L2 33 D, should be WL2 4 10 NOT 100 (from 03-17 to 07-24) - original died and new one germinated later

unique(dat4$block)
dat4 %>% filter(is.na(block), !is.na(height_cm))
unique(dat4$row)
dat4 %>% filter(is.na(row), !is.na(height_cm))
unique(dat4$col) #some cols named "A." and "&"  SHOULD FIX!!*******
dat4 %>% filter(col=="A.")
dat4 %>% filter(col=="&")
dat4 %>% filter(is.na(col), !is.na(height_cm))

#lag lead function
dat5 <- dat4 %>% filter(pop!="buffer") %>% mutate(across(c(row, mf, rep,longest_leaf_cm, height_cm), as.numeric)) %>% mutate(pop=str_replace(pop, ".*VTR.*", "LVTR1")) %>%  group_by(block, row, col) %>% mutate(previous_census_height=lag(height_cm, order_by = survey_date), previous_census_length=lag(longest_leaf_cm, order_by = survey_date)) %>% select(survey_date, block, row, col, previous_census_height, height_cm, previous_census_length, longest_leaf_cm, death_date, Notes)
dat5

#Check for data when there is no data in previous week 
dat6 <- dat5 %>% filter(survey_date!="2022-12-13") %>% filter(is.na(previous_census_height), !is.na(height_cm))
#write.csv(dat6, "../output/UCD_MissingDataPrevWeek_BlockRowCol.csv", row.names = FALSE)
dat6
#J1 12 C on 12/13, the block, row, column, didn't show on the scan so it was NA, fixed it on the CorrectedCSV
#F2 31 D was dead, Amazon had turned a "-" into a "1" for height, fixed it  
#H2 28 C on 0217 had a column name of ": C" - amazon error - fixed it on the corrected csv

#16 other cases to check:
#J2 28 A was marked dead on 0127, but was measured on 0210 w/ no note and prob too big to be a new germ 
#L1 05 A 0210 - original died 0127, new one germinated, so rep changed to 100
#H1 19 B 0303
#H2 28 A 0303
#L2 37 C 0303 - original died and new one germinated later
#D1 05 D 0317
#L2 33 D 0317 - original died and new one germinated later
#J2 31 D 0324
#L1 03 C 0324
#L1 03 D 0324
#J1 06 A 0403
#D1 09 A 0410 - note says new germ, species ID unconfirmed 
#J1 06 B 0410
#L1 03 C 0410
#J2 23 D 0417
#L1 03 C 0424

dat7 <- dat4 %>% filter(pop!="buffer") %>% mutate(across(c(row, mf, rep,longest_leaf_cm, height_cm), as.numeric)) %>% mutate(pop=str_replace(pop, ".*VTR.*", "LVTR1")) %>%  group_by(pop, mf, rep) %>% mutate(previous_census_height=lag(height_cm, order_by = survey_date), previous_census_length=lag(longest_leaf_cm, order_by = survey_date)) %>% select(survey_date, block, row, col, pop, mf, rep, previous_census_height, height_cm, previous_census_length, longest_leaf_cm, death_date, Notes)
dat7

#Check for data when there is no data in previous week 
dat8 <- dat7 %>% filter(survey_date!="2022-12-13") %>% filter(is.na(previous_census_height), !is.na(height_cm)) #same number rows as the above 
#write.csv(dat8, "../output/UCD_MissingDataPrevWeek_PopMFREp.csv", row.names = FALSE)
dat8
```


# Size Data - Transplants (basic check)
```{r}
#load the libraries
library(tidyverse)
#tutorial at https://bookdown.org/ansellbr/WEHI_tidyR_course_book/week-4-part-3.html
data_files <- list.files('~/Google Drive/Shared drives/IntBioTeam/Common Gardens/UCD2022_2023/CorrectedCSVs',pattern='Size_survey_transplants.*csv',ignore.case=FALSE) 

data_names <- str_remove(data_files, pattern = ".csv")
data_names
data_names2 <- data_names[-c(1,2)]
data_names2 #the first two censuses don't have a herbivory column 

for (i in data_names2) {
  #set up variable file names
  in_file <- paste("~/Google Drive/Shared drives/IntBioTeam/Common Gardens/UCD2022_2023/CorrectedCSVs/", i, '.csv', sep = "")   #edit this to be correct path 
  out_table <- paste("../output/", i, '_dataprobs.csv', sep="")
 
  #data checking code
  data <- read_csv(in_file)
  
  #remove spaces in column names
  names(data) <- gsub(" ", "", colnames(data)) 
  names(data)
  
  #data <- data %>% rename(height.cm= `height(cm)`, long.leaf.cm=`longestleaf(cm)`)
  data <- data %>% rename(height.cm= `height(cm)`, long.leaf.cm=`longestleaf(cm)`, herbiv.y.n=`herbivory(YorN)`)
  
  #check for spaces in numbers
  data2 <- data %>% filter(grepl(" ", height.cm) | grepl(" ", long.leaf.cm)) %>% mutate(problem="spaces")
  
  #check for extreme values
  data3 <- data
  data3$height.cm <- as.double(data3$height.cm)
  data3$long.leaf.cm <- as.double(data3$long.leaf.cm)
  data3 <- data3 %>% filter(height.cm > 140 | long.leaf.cm > 15) %>% mutate(problem="extreme value")
  
  #check for spread out data 
  #data4 <- data
  #data4 <- data4 %>% mutate(Notes = na_if(Notes, "N/A")) %>% mutate(Notes = na_if(Notes, "-"))
  #data4$Notes <- str_to_upper(data4$Notes)
  #data4 <- data4 %>% filter(!is.character(Notes)) %>% mutate(problem="notes col")
  data4 <- data
  data4 <- data4 %>% mutate(herbiv.y.n = na_if(herbiv.y.n, "N/A")) %>% mutate(herbiv.y.n = na_if(herbiv.y.n, "-"))
  data4$herbiv.y.n <- str_to_upper(data4$herbiv.y.n)
  data4 <- data4 %>% filter(herbiv.y.n != "Y" & herbiv.y.n != "N") %>% mutate(problem="herbiv col")
  
  #check for letters in number columns 
  data5 <- data %>% rowwise() %>% filter(!is.na(height.cm) & height.cm !="-") %>% filter(is.na(as.numeric(height.cm))) %>% mutate(problem="character")
  data6 <- data %>% rowwise() %>% filter(!is.na(long.leaf.cm) & long.leaf.cm!="-") %>% filter(is.na(as.numeric(long.leaf.cm))) %>% mutate(problem="character")
  
  #combine into one problems list
  data_probs <- rbind(data2, data3, data4, data5, data6)
  print(nrow(data_probs))
  write_csv(data_probs, path=out_table)
}
```

# Mortality/Phenology - Transplants
```{r}
#load the libraries
library(tidyverse)
#tutorial at https://bookdown.org/ansellbr/WEHI_tidyR_course_book/week-4-part-3.html
data_files <- list.files('~/Google Drive/Shared drives/IntBioTeam/Common Gardens/UCD2022_2023/CorrectedCSVs',pattern='transplants.*mort.*csv',ignore.case=TRUE)
data_files #make this specific to mortality or size

data_names <- str_remove(data_files, pattern = ".csv")
data_names

for (i in data_names) {
  #set up variable file names
  in_file <- paste("~/Google Drive/Shared drives/IntBioTeam/Common Gardens/UCD2022_2023/CorrectedCSVs/", i, '.csv', sep = "") 
  out_table <- paste("../output/", i, '_dataprobs.csv', sep="")
  
  #data checking code
  data <- read_csv(in_file)
  
  #remove spaces in column names
  names(data) <- gsub(" ", "", colnames(data))
  names(data)
  
  #check for spaces in numbers
  data2 <- data %>% filter(grepl(" ", DateFirstBud) | grepl(" ", DateFirstFlower) | grepl(" ", DateFirstFruit) | grepl(" ", DateLastFlower) | grepl(" ", DateLastFruit) | grepl(" ", DeathDate)) %>% mutate(problem="spaces")
  
  #combine into one problems list
  data_probs <- data2
  print(nrow(data_probs))
  write_csv(data_probs, path=out_table)
}
```

# Germination Beds
Note: In germination beds sheets, B2 34 B - had “w” as population name. It is actually WV 3 7 (on “Pops_for_2022_UCD > New Bed Map_racks” google sheet). BQC changed it in all the CorrectedCSV files. 

# Annual Census - Germination Beds 
```{r}
#load the libraries
library(tidyverse)
#tutorial at https://bookdown.org/ansellbr/WEHI_tidyR_course_book/week-4-part-3.html
data_files <- list.files('~/Google Drive/Shared drives/IntBioTeam/Common Gardens/UCD2022_2023/CorrectedCSVs',pattern='Annual_Census_Germination',ignore.case=TRUE)
data_files #make this specific to mortality or size

data_names <- str_remove(data_files, pattern = ".csv")
data_names

for (i in data_names) {
  #set up variable file names
  in_file <- paste("~/Google Drive/Shared drives/IntBioTeam/Common Gardens/UCD2022_2023/CorrectedCSVs/", i, '.csv', sep = "")  
  out_table <- paste("../output/", i, '_dataprobs.csv', sep="")
  
  #data checking code
  data <- read_csv(in_file)
  
  #remove spaces in column names
  names(data) <- gsub(" ", "_", colnames(data))
  names(data)
  
  #check for spaces in numbers
  data2 <- data %>% filter(grepl(" ", diam_mm) | grepl(" ", height_cm) | grepl(" ", total_branch) | grepl(" ", longest_leaf_cm) | grepl(" ", flowers) | grepl(" ", fruits) | grepl(" ", longest_fruit_cm) | grepl(" ", repro__branch)) %>% mutate(problem="spaces")
  
  #check for extreme values
  data3 <- data
  data3$diam_mm <- as.double(data3$diam_mm)
  data3$height_cm <- as.double(data3$height_cm)
  data3$longest_leaf_cm <- as.double(data3$longest_leaf_cm)
  data3$longest_fruit_cm <- as.double(data3$longest_fruit_cm)
  data3 <- data3 %>% filter(diam_mm > 50, height_cm > 140 | longest_leaf_cm > 12, longest_fruit_cm > 15) %>% mutate(problem="extreme value")
  
  #check for letters in number columns 
  data4 <- data %>% rowwise() %>% filter(!is.na(diam_mm) & diam_mm!="n/a") %>% filter(is.na(as.numeric(diam_mm))) %>% mutate(problem="character")
  data5 <- data %>% rowwise() %>% filter(!is.na(height_cm) & height_cm !="n/a") %>% filter(is.na(as.numeric(height_cm))) %>% mutate(problem="character")
  data6 <- data %>% rowwise() %>% filter(!is.na(total_branch) & total_branch != "n/a") %>% filter(is.na(as.numeric(total_branch))) %>% mutate(problem="character")
  data7 <- data %>% rowwise() %>% filter(!is.na(longest_leaf_cm) & longest_leaf_cm != "n/a") %>% filter(is.na(as.numeric(longest_leaf_cm))) %>% mutate(problem="character")
  data8 <- data %>% rowwise() %>% filter(!is.na(flowers) & flowers != "n/a") %>% filter(is.na(as.numeric(flowers))) %>% mutate(problem="character")
  data9 <- data %>% rowwise() %>% filter(!is.na(fruits) & fruits != "n/a") %>% filter(is.na(as.numeric(fruits))) %>% mutate(problem="character")
  data10 <- data %>% rowwise() %>% filter(!is.na(longest_fruit_cm) & longest_fruit_cm != "n/a") %>% filter(is.na(as.numeric(longest_fruit_cm))) %>% mutate(problem="character")
  data11 <- data %>% rowwise() %>% filter(!is.na(repro__branch) & repro__branch != "n/a") %>% filter(is.na(as.numeric(repro__branch))) %>% mutate(problem="character")
  
  #combine into one problems list
  data_probs <- rbind(data2, data3, data4, data5, data6, data7, data8, data9, data10, data11)
  print(nrow(data_probs))
  write_csv(data_probs, path=out_table)
}
```


# Size Data - Germination Beds 
```{r}
#load the libraries
library(tidyverse)
#tutorial at https://bookdown.org/ansellbr/WEHI_tidyR_course_book/week-4-part-3.html
data_files <- list.files('~/Google Drive/Shared drives/IntBioTeam/Common Gardens/UCD2022_2023/CorrectedCSVs',pattern='Size_survey_germination.*csv',ignore.case=FALSE) 
data_files

data_names <- str_remove(data_files, pattern = ".csv")
data_names

for (i in data_names) {
  #set up variable file names
  in_file <- paste("~/Google Drive/Shared drives/IntBioTeam/Common Gardens/UCD2022_2023/CorrectedCSVs/", i, '.csv', sep = "")   #edit this to be correct path 
  out_table <- paste("../output/", i, '_dataprobs.csv', sep="")
  
  #data checking code
  data <- read_csv(in_file)
  
  #remove spaces in column names
  names(data) <- gsub(" ", "", colnames(data)) 
  names(data)
  
  data <- data %>% rename(height.cm= `height(cm)`, long.leaf.cm=`longestleaf(cm)`, herbiv.y.n=`herbivory(YorN)`)
  
  #check for spaces in numbers
  data2 <- data %>% filter(grepl(" ", height.cm) | grepl(" ", long.leaf.cm)) %>% mutate(problem="spaces")
  
  #check for extreme values
  data3 <- data
  data3$height.cm <- as.double(data3$height.cm)
  data3$long.leaf.cm <- as.double(data3$long.leaf.cm)
  data3 <- data3 %>% filter(height.cm > 140 | long.leaf.cm > 15) %>% mutate(problem="extreme value")
  
  #check for spread out data 
  data4 <- data
  data4 <- data4 %>% mutate(herbiv.y.n = na_if(herbiv.y.n, "N/A")) %>% mutate(herbiv.y.n = na_if(herbiv.y.n, "-"))
  data4$herbiv.y.n <- str_to_upper(data4$herbiv.y.n)
  data4 <- data4 %>% filter(herbiv.y.n != "Y" & herbiv.y.n != "N") %>% mutate(problem="herbiv col")
  
   #check for letters in number columns 
  data5 <- data %>% rowwise() %>% filter(!is.na(height.cm) & height.cm !="-") %>% filter(is.na(as.numeric(height.cm))) %>% mutate(problem="character")
  data6 <- data %>% rowwise() %>% filter(!is.na(long.leaf.cm) & long.leaf.cm!="-") %>% filter(is.na(as.numeric(long.leaf.cm))) %>% mutate(problem="character")
  
  #combine into one problems list
  data_probs <- rbind(data2, data3, data4, data5, data6)
  print(nrow(data_probs))
  write_csv(data_probs, path=out_table)
}
```

# Mortality/Phenology - Germination Beds
```{r}
#load the libraries
library(tidyverse)
#tutorial at https://bookdown.org/ansellbr/WEHI_tidyR_course_book/week-4-part-3.html
data_files <- list.files('~/Google Drive/Shared drives/IntBioTeam/Common Gardens/UCD2022_2023/CorrectedCSVs',pattern='Germination_survey.*csv',ignore.case=TRUE)
data_files #make this specific to mortality or size

data_files_phen <- list.files('~/Google Drive/Shared drives/IntBioTeam/Common Gardens/UCD2022_2023/CorrectedCSVs',pattern='phenology.*germination.*csv',ignore.case=TRUE)
data_files_phen #make this specific to mortality or size

data_names <- str_remove(data_files, pattern = ".csv")
data_names
data_names2 <- data_names[-1]
data_names2

data_names_phen <- str_remove(data_files_phen, pattern = ".csv")
data_names_phen

for (i in data_names_phen) {
  #set up variable file names
  in_file <- paste("~/Google Drive/Shared drives/IntBioTeam/Common Gardens/UCD2022_2023/CorrectedCSVs/", i, '.csv', sep = "") 
  out_table <- paste("../output/", i, '_dataprobs.csv', sep="")
  in_file
  #data checking code
  data <- read_csv(in_file)
  
  #remove spaces in column names
  names(data) <- gsub(" ", "", colnames(data))
  names(data)
  
  #check for spaces in numbers
  #data2 <- data %>% filter(grepl(" ", deathdate)) %>% mutate(problem="spaces")
  #data2 <- data %>% filter(grepl(" ", deathdate1) | grepl(" ", deathdate2) | grepl(" ", deathdate3)) %>% mutate(problem="spaces")
  data2 <- data %>% filter(grepl(" ", DateFirstBud) | grepl(" ", DateFirstFlower) | grepl(" ", DateFirstFruit) | grepl(" ", DateLastFlower) | grepl(" ", DateLastFruit)) %>% mutate(problem="spaces")
  
  #combine into one problems list
  data_probs <- data2
  print(nrow(data_probs))
  write_csv(data_probs, path=out_table)
}
```
