---
title: "UCD_SurvAnalysis_Pops_GLMER_Add_Climate"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Add climate to best base model

# Survival Analysis

## Libraries

```{r}
# install.packages("multilevelmod")
library(tidyverse) 
library(magrittr)
library(ggrepel)
library(lme4)
library(ggdist)
library(tidymodels)
library(multilevelmod)
library(broom.mixed)
library(modelr)
library(furrr)
library(zoo)
tidymodels_prefer()
options(mc.cores = parallel::detectCores())
```

## Read in the plant data
```{r}
ucd_surv <- read_csv("../input/UCD_Data/CorrectedCSVs/UCD_transplants_pheno_mort_20231016_corrected.csv") %>% 
  filter(rep != 100) #get rid of individuals that germinated in the field
head(ucd_surv)
unique(ucd_surv$pop) #pops look correct 
```

As weeks
```{r}
ucd_surv %>% filter(is.na(`Death Date`),!is.na(Notes)) #no plants that seemed to disappear 

ucd_surv_dates <- ucd_surv %>%  
  rename(parent.pop = pop, death.date=`Death Date`) %>%
  unite("Genotype", parent.pop:rep, remove = FALSE) %>% 
  filter(!is.na(parent.pop)) %>% 
  mutate(planting.date="11/30/22", #could try to make this more specific to when certain blocks were planted 
         last_fup_date=if_else(is.na(death.date), "10/16/23", death.date)) %>%  #need this to calculate survival times
  mutate(planting.date=mdy(planting.date), last_fup_date=mdy(last_fup_date)) %>% #convert to date objects
  mutate(os_weeks=as.duration(planting.date %--% last_fup_date) / dweeks(1), #observed number of weeks
         status=if_else(is.na(death.date), 0, 1)) %>% #0=censured (alive in this case), 1=dead
  filter(os_weeks > 0) %>% #remove any that died at planting since this is survival post- transplanting
  select(block:rep, death.date, planting.date:status)
head(ucd_surv_dates)
unique(ucd_surv_dates$os_weeks) %>% sort()
```

## Read in soil temp
Transplants only in beds D, F, H, J, L - Remove Bed A from data set 
No temperatures were taken from 2/23/23 to 3/17/23 and after 8/4
```{r}
temp <- read_csv("../input/UCD_Data/iButton_Data/UCD_iButton_Data_Corrected.csv") %>%
  filter(Bed != "A") %>% 
  mutate(Date_Time = mdy_hm(Date_Time))
head(temp)
skimr::skim(temp)
```

```{r}
temp %>% ggplot(aes(x=Date_Time, y = SoilTemp, color= Bed)) + 
  geom_line()

#Take a closer look
temp %>% mutate(month = month(Date_Time)) %>%
  filter(Date_Time > ymd("2023-07-17"), Date_Time <= ymd("2023-07-24")) %>%
  ggplot(aes(x=Date_Time, y = SoilTemp, color= Bed, lty = Bed)) + 
  geom_line() +
  scale_color_brewer(type="qual", palette = "Accent")

temp %>% mutate(month = month(Date_Time)) %>%
  filter(month==7) %>%
  ggplot(aes(x=Date_Time, y = SoilTemp)) + 
  geom_line() +
  scale_color_brewer(type="qual", palette = "Accent") +
  facet_wrap(~Bed)
```

Average across beds 

```{r}
temp_summary <- temp %>%
  group_by(Date_Time) %>% 
  summarise(AvgSoilTemp=mean(SoilTemp, na.rm=TRUE)) %>% #avg across beds 
  mutate(Date=as.Date(Date_Time)) %>%
  filter(Date != min(Date), Date != max(Date)) %>% # trim ragged ends
  group_by(Date) %>% #summarize hourly data by date
  summarize(
    min_temp_d = min(AvgSoilTemp),
    max_temp_d = max(AvgSoilTemp),
    mean_temp_d = mean(AvgSoilTemp)
  ) %>%
  mutate(
    across(ends_with("temp_d"), \(x) rollmean(x, k = 7, align = "right", fill = NA), .names="{.col}1_7"),
    across(ends_with("temp_d"), \(x) rollmean(x, k = 13, align = "right", fill = NA), .names="{.col}1_13"), # 13 so I can get the first survey date
    across(ends_with("1_7"), \(x) dplyr::lag(x, 6), .names = "{.col}_13") 
  ) %>%
  rename_with(\(x) str_replace(x, "1_7_13", "6_13"))

temp_summary
```

## Read in soil moisture
Remove Beds E, N since no transplants in those beds 

H - Started logging 4/6 at 2pm- port 1 error/unplugged until 4/10 - 6 probes
L - Started logging 4/6 at 2pm - 5 probes 
End date = 10/16
```{r}
moisture <- read_csv("../input/UCD_Data/Soil_Moisture/UCD_Soil_VWC_Corrected.csv") %>%
  filter(Bed != "E", Bed!="N") %>% 
  mutate(Date_Time = mdy_hm(Date_Time)) 
head(moisture)
skimr::skim(moisture)
```


```{r}
moisture %>% pivot_longer(Port_1:Port_6) %>%
  ggplot(aes(x=Date_Time, y = value, color= name)) + 
  geom_line() +
  facet_wrap(~Bed)
```

Will average across the ports and beds 

```{r}
moisture_summary <- moisture %>%
  mutate(Port_1=if_else(Port_1<0, 0, Port_1),
         Port_2=if_else(Port_2<0, 0, Port_2),
         Port_3=if_else(Port_3<0, 0, Port_3),
         Port_4=if_else(Port_4<0, 0, Port_4),
         Port_5=if_else(Port_5<0, 0, Port_5),
         Port_6=if_else(Port_5<0, 0, Port_6)) %>%  #convert negatives to zeros (negatives result from extremely dry soil --> air pockets)
  group_by(Date_Time) %>%
  summarise(s_moisture=mean(c_across(Port_1:Port_6), na.rm=TRUE)) %>% 
  mutate(Date=as.Date(Date_Time)) %>%
  group_by(Date) %>%
  summarize(
    mean_moisture_d = mean(s_moisture)
  ) %>%
  mutate(
    s_moisture_1_7 = rollmean(mean_moisture_d, k = 7, align = "right", fill = "extend"),
    s_moisture_1_14 = rollmean(mean_moisture_d, k = 13, align = "right", fill = "extend"),
    s_moisture_7_14 = dplyr::lag(s_moisture_1_7, 7)
  )

moisture_summary
```

## Reformat to be one row per plant per week

```{r, eval=TRUE}
ucd_ind <- ucd_surv_dates %>% 
  complete(Genotype, last_fup_date) %>%
  group_by(Genotype) %>%
  mutate(across(c(-status, -last_fup_date, -os_weeks, -planting.date), \(x) ifelse(all(is.na(unique(x))), NA,unique(na.omit(x))))) %>%
  mutate(planting.date = unique(na.omit(planting.date)),
         os_weeks=as.duration(planting.date %--% last_fup_date) / dweeks(1),
           dead = ifelse((is.na(death.date) | mdy(death.date) > last_fup_date), 0, 1)) %>% # 0 is alive, 1 is dead
  filter(cumsum(dead) < 2) %>% # you can't die more than once
  ungroup() 

ucd_ind %>% select(Genotype, last_fup_date, os_weeks, dead)
```

Split data into establishment and survival data sets. (We will only use survival set for this analysis)

```{r}
ucd_ind_surv <- ucd_ind %>% filter(os_weeks > 3) %>%
  mutate(os_weeks_f = factor(os_weeks))

ucd_surv_means <- ucd_ind_surv %>%
  group_by(parent.pop, os_weeks) %>%
  summarize(survivorship = 1-mean(dead))
```

## add climate data

```{r}
ucd_ind_surv_climate <- ucd_ind_surv %>% 
  left_join(temp_summary, by = c("last_fup_date" = "Date")) %>%
  left_join(moisture_summary, by = c("last_fup_date" = "Date"))
```

## Temp analysis

check basic model forms

```{r}
glmer.model <- 
  linear_reg() %>%
  set_engine("glmer", family = binomial)

temp_wflow <- workflow() %>%
  add_variables(outcomes = dead, predictors = c(parent.pop, mf, block, os_weeks, contains("temp_d"))) 

temp_fits <- tibble(wflow=list(

  # The best base model from earlier analyses:
  pop.mf.weeks.block ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + (1|parent.pop/mf) + (1|block))},
  
  min.temp.1.7 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ min_temp_d1_7 + os_weeks + (1|parent.pop/mf) + (1|block))},
  
  mean.temp.1.7 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ mean_temp_d1_7 + os_weeks + (1|parent.pop/mf) + (1|block))},
  
  max.temp.1.7 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ max_temp_d1_7 + os_weeks + (1|parent.pop/mf) + (1|block))},
  
  min.temp.1.13 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ min_temp_d1_13 + os_weeks + (1|parent.pop/mf) + (1|block))},
  
  mean.temp.1.13 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ mean_temp_d1_13 + os_weeks + (1|parent.pop/mf) + (1|block))},
  
  max.temp.1.13 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ max_temp_d1_13 + os_weeks + (1|parent.pop/mf) + (1|block))},
  
  min.temp.6.13 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ min_temp_d6_13 + os_weeks + (1|parent.pop/mf) + (1|block))},
  
  mean.temp.6.13 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ mean_temp_d6_13 + os_weeks + (1|parent.pop/mf) + (1|block))},
  
  max.temp.6.13 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ max_temp_d6_13 + os_weeks + (1|parent.pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)
temp_fits
```

```{r}
system.time(
  temp_fits <- temp_fits %>%
  mutate(fit = map(wflow, fit, data = drop_na(ucd_ind_surv_climate, contains("temp_d")), .progress = TRUE))
)
```

```{r}
temp_fits %>% pull(fit)
```

```{r}
temp_fits %>% 
  mutate(glance=map(fit, glance)) %>% 
  unnest(glance) %>% 
  arrange(BIC) %>%
  select(-wflow:-fit, -sigma)
```

get p-values

```{r}
temp_fits %>%
  mutate(tidy=map(fit, tidy)) %>%
  select(name, tidy) %>%
  unnest(tidy) %>%
  filter(str_detect(term, "temp")) %>%
  arrange(p.value)
```

## Moisture analysis

check basic model forms

```{r}
glmer.model <- 
  linear_reg() %>%
  set_engine("glmer", family = binomial)

moisture_wflow <- workflow() %>%
  add_variables(outcomes = dead, predictors = c(parent.pop, mf, block, os_weeks, contains("s_moisture") )) 

moisture_fits <- tibble(wflow=list(

  # The best base model from earlier analyses:
  pop.weeks.block ={moisture_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + (1|parent.pop/mf) + (1|block))},
  
    s_moisture1.7 ={moisture_wflow %>% 
      add_model(glmer.model, formula = dead ~ s_moisture_1_7 + os_weeks + (1|parent.pop/mf) + (1|block))},
  
      s_moisture1.14 ={moisture_wflow %>% 
      add_model(glmer.model, formula = dead ~ s_moisture_1_14 + os_weeks + (1|parent.pop/mf) + (1|block))},
      
          s_moisture7.14 ={moisture_wflow %>% 
      add_model(glmer.model, formula = dead ~ s_moisture_7_14 + os_weeks + (1|parent.pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)
moisture_fits
```

```{r}
system.time(
  moisture_fits <- moisture_fits %>%
  mutate(fit = map(wflow, fit, data = drop_na(ucd_ind_surv_climate, contains("s_moisture")), .progress = TRUE))
)
```
Singularity warnings for 3/4 models b/c of mf 
```{r}
moisture_fits %>% pull(fit)
```

```{r}
moisture_fits %>% 
  mutate(glance=map(fit, glance)) %>% 
  unnest(glance) %>% 
  arrange(BIC) %>%
  select(-wflow:-fit, -sigma)
```
Only 905 observations since soil moisture was only measured starting in April  

get p-values

```{r}
moisture_fits %>%
  mutate(tidy=map(fit, tidy)) %>%
  select(name, tidy) %>%
  unnest(tidy) %>%
  filter(str_detect(term, "moisture"))
```