---
title: "UCD_Growth_Climate"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# UCD Growth + Climate

Add climate to best base model

## Libraries

```{r}
library(tidyverse)
library(brms)
library(zoo) #rollmean
```

## Need to make a csv file with all survey data for each plant

### Get list of size files

```{r}
pretrans_ucd <-  dir("../input/UCD_Data/CorrectedCSVs/",
             pattern=".*20221128.*csv",
             full.names=TRUE)
pretrans_ucd

posttrans_ucd <- dir("../input/UCD_Data/CorrectedCSVs/",
             pattern="Size_survey_transplants.*corrected.*csv",
             full.names=TRUE)
posttrans_ucd
```

### Add data to a tibble

```{r}
pretrans_dat_ucd <- tibble(path=pretrans_ucd, filename=basename(path))
pretrans_dat_ucd
posttrans_dat_ucd <- tibble(path=posttrans_ucd, filename=basename(path))
posttrans_dat_ucd

all_ucd_size <- bind_rows(pretrans_dat_ucd, posttrans_dat_ucd) %>% 
  mutate(survey_date=str_extract(filename, "2[0-9]*"),
         survey_date=lubridate::ymd(survey_date))
all_ucd_size
```

### Read in the files

```{r}
all_ucd_size2 <- all_ucd_size %>%
  mutate(sheets=map(path, 
                    read_csv, 
                    #col_types=cols(.default=col_character()), # forces all columns to be character
                                                              # this is needed because of data sheet problems that cause some numerics to be imported as character  
                    na = c("", "NA", "-", "N/A") # sets NA strings. 
                    )) %>%
  select(-path)

all_ucd_size2
```

### Check for data issues

```{r}
all_ucd_size2$sheets[[1]] <- all_ucd_size2$sheets[[1]] %>% rename(pop=parent.pop, `height (cm)`=`Height (cm)`, `longest leaf (cm)`=`Longest leaf (cm)`) 
all_ucd_size2$sheets[[2]] <- all_ucd_size2$sheets[[2]] %>% rename(pop=parent.pop) 
all_ucd_size2$sheets[[5]] <- all_ucd_size2$sheets[[5]] %>% mutate(mf=as.double(mf)) #convert mf to numeric
all_ucd_size2$sheets[[1]] <- all_ucd_size2$sheets[[1]] %>% filter(`germinated?`== 1) #keep germs only from pretrans size 
#ALL ISSUES FIXED!
```

### Combine the data

```{r}
all_ucd_size_all <- all_ucd_size2 %>% unnest(sheets) %>%
  rename(germ = `germinated?`, height.cm = `height (cm)`, 
         long.leaf.cm = `longest leaf (cm)`, parent.pop=pop) %>% 
  filter(parent.pop != "buffer") %>% 
  mutate(parent.pop= str_replace(parent.pop, ".*VTR.*", "LVTR1")) %>% 
  filter(rep != 100) %>% #get rid of individuals that germinated in the field 
  unite(Genotype, parent.pop:rep, sep="_", remove = FALSE) %>% 
  unite(pop.mf, parent.pop:mf, sep="_", remove = FALSE) %>% 
  select(survey_date, block, Genotype:Notes)
head(all_ucd_size_all)
unique(all_ucd_size_all$parent.pop) #all 23 pops!

#write_csv(all_ucd_size_all, "../output/UCD_Traits/UCD_Size_Combined.csv")
```

## Pop Location Info

```{r}
ucd_gowers <- read_csv("../output/Climate/Gowers_UCD.csv") %>% 
  select(parent.pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) 
head(ucd_gowers)
```

Merge

```{r}
all_ucd_size_loc <- left_join(all_ucd_size_all, ucd_gowers) %>% 
  mutate(day=as.numeric(survey_date-min(survey_date)))
head(all_ucd_size_loc)
unique(all_ucd_size_loc$day)
```

## Data set for the model

```{r}
all_ucd_size_for_models <- all_ucd_size_loc %>% 
  filter(parent.pop != "WV", parent.pop!="LV1") %>% #consider also removing CP3 since no data after 1/27 
  filter(Genotype !="SQ3_4_3") %>% 
  filter(!is.na(height.cm))
```

## Read in soil temp

Transplants only in beds D, F, H, J, L - Remove Bed A from data set No
temperatures were taken from 2/23/23 to 3/17/23 and after 8/4

```{r}
temp_UCD <- read_csv("../input/UCD_Data/iButton_Data/UCD_iButton_Data_Corrected.csv") %>%
  filter(Bed != "A") %>% 
  mutate(Date_Time = mdy_hm(Date_Time))
head(temp_UCD)
skimr::skim(temp_UCD)
```

Average across beds

```{r}
temp_UCD_summary <- temp_UCD %>%
  group_by(Date_Time) %>% 
  summarise(AvgSoilTemp=mean(SoilTemp, na.rm=TRUE)) %>% #avg across beds 
  mutate(Date=as.Date(Date_Time)) %>%
  filter(Date != min(Date), Date != max(Date)) %>% # trim ragged ends
  group_by(Date) %>% #summarize hourly data by date
  summarize(
    min_temp_d = min(AvgSoilTemp),
    max_temp_d = max(AvgSoilTemp),
    mean_temp_d = mean(AvgSoilTemp)
  ) %>%
  mutate(
    across(ends_with("temp_d"), \(x) rollmean(x, k = 7, align = "right", fill = NA), .names="{.col}1_7"),
    across(ends_with("temp_d"), \(x) rollmean(x, k = 13, align = "right", fill = NA), .names="{.col}1_13"), # 13 so I can get the first survey date
    across(ends_with("1_7"), \(x) dplyr::lag(x, 6), .names = "{.col}_13") 
  ) %>%
  rename_with(\(x) str_replace(x, "1_7_13", "6_13"))

temp_UCD_summary
```

## add climate data

```{r}
all_ucd_size_for_models_climate <- all_ucd_size_for_models %>% 
  left_join(temp_UCD_summary, by = c("survey_date" = "Date")) #%>%
  #filter(day>0) #remove day pre-field size b/c no climate data #without day 0, min is different...

all_ucd_size_for_models_climate %>% 
  ggplot(aes(group=Genotype, x=survey_date, y=height.cm, col=elev_m)) + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_line() + facet_wrap(~parent.pop, scales = "free")
```

## Models

```{r}
#best base model:
climate_f1 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  Hmax + k + delta ~ (1|parent.pop), 
                  nl=TRUE)

climate_f2 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  Hmax + k + delta ~ min_temp_d1_7 + (1|parent.pop), 
                  nl=TRUE)

climate_f3 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  Hmax + k + delta ~ mean_temp_d1_7 + (1|parent.pop), 
                  nl=TRUE)

climate_f4 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  Hmax + k + delta ~ max_temp_d1_7 + (1|parent.pop), 
                  nl=TRUE)

climate_f5 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  Hmax + k + delta ~ min_temp_d1_13 + (1|parent.pop), 
                  nl=TRUE)

climate_f6 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  Hmax + k + delta ~ max_temp_d1_13 + (1|parent.pop), 
                  nl=TRUE)

climate_f7 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  Hmax + k + delta ~ (1|parent.pop), 
                  nl=TRUE)

climate_f8 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  Hmax + k + delta ~ min_temp_d6_13 + (1|parent.pop), 
                  nl=TRUE)

climate_f9 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  Hmax + k + delta ~ mean_temp_d6_13 + (1|parent.pop), 
                  nl=TRUE)

climate_f10 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  Hmax + k + delta ~ max_temp_d6_13 + (1|parent.pop), 
                  nl=TRUE)
```

## Priors

```{r}
prior1 <- c(set_prior("normal(5,3)", nlpar="Hmin"),
            set_prior("normal(60,15)",  nlpar="Hmax"),
            set_prior("gamma(2,2)", nlpar = "k", lb = 0), #gamma constrains to be > 0 
            set_prior("gamma(20,3)", nlpar = "delta", lb = 0)
)
            #first number = avg, second number = the space around that average to search 
```

## Model Fits

### Base model (fit1)

```{r}
climate_fit1 <- brm(formula=climate_f1, data=all_ucd_size_for_models_climate, prior=prior1, iter = 3800, cores=4)
#Warning: There were 2500 divergent transitions after warmup. See https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup to find out why this is a problem and how to eliminate them.Warning: Examine the pairs() plot to diagnose sampling problems

summary(climate_fit1)
plot(climate_fit1, nvariables = 3, ask=FALSE)

pred.df <- expand_grid(day=min(all_ucd_size_for_models_climate$day):max(all_ucd_size_for_models_climate$day),
                       parent.pop=unique(all_ucd_size_for_models_climate$parent.pop))
climate_fit1.predictions <- cbind(pred.df, prediction=predict(climate_fit1, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_ucd_size_for_models_climate, by=c("day", "parent.pop"))

climate_fit1.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)
```

### Min D1-7 (fit2)

```{r}
all_ucd_size_for_models_climate_fit2 <- all_ucd_size_for_models_climate %>% drop_na(min_temp_d1_7)

climate_fit2 <- brm(formula=climate_f2, data=all_ucd_size_for_models_climate_fit2, prior=prior1, iter = 3800, cores=4)
#See warnings below 

summary(climate_fit2) #pretty bad diagnostics
plot(climate_fit2, nvariables = 3, ask=FALSE) #pretty bad 

#pred.df <- expand_grid(day=min(all_ucd_size_for_models_climate_fit2$day):max(all_ucd_size_for_models_climate_fit2$day),
                  #     parent.pop=unique(all_ucd_size_for_models_climate_fit2$parent.pop),
                   #    min_temp_d1_7=unique(all_ucd_size_for_models_climate_fit2$min_temp_d1_7))
#climate_fit2.predictions <- cbind(pred.df, prediction=predict(climate_fit2, newdata = pred.df)[,"Estimate"]) %>%
#  full_join(all_ucd_size_for_models_climate_fit2, by=c("day", "parent.pop"))
#Error: vector memory limit of 16.0 Gb reached, see mem.maxVSize()

#climate_fit2.predictions %>% ggplot(aes(x=day)) +
#  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
#  geom_line(aes(y=prediction, color=parent.pop)) +
#  facet_wrap(~parent.pop)
```

Warning: There were 6576 divergent transitions after warmup. See https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup to find out why this is a problem and how to eliminate them.
Warning: There were 1 chains where the estimated Bayesian Fraction of Missing Information was low. See https://mc-stan.org/misc/warnings.html#bfmi-low
Warning: Examine the pairs() plot to diagnose sampling problems
Warning: The largest R-hat is 3.69, indicating chains have not mixed. Running the chains for more iterations may help. See https://mc-stan.org/misc/warnings.html#r-hat
Warning: Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable. Running the chains for more iterations may help. See https://mc-stan.org/misc/warnings.html#bulk-ess
Warning: Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable. Running the chains for more iterations may help. See https://mc-stan.org/misc/warnings.html#tail-ess

### Mean D1-7 (fit3)

```{r}
all_ucd_size_for_models_climate_fit3 <- all_ucd_size_for_models_climate %>% drop_na(mean_temp_d1_7)

climate_fit3 <- brm(formula=climate_f3, data=all_ucd_size_for_models_climate_fit3, prior=prior1, iter = 3800, cores=4)
#See warnings below 

summary(climate_fit3)
plot(climate_fit3, nvariables = 3, ask=FALSE)
#pretty bad diagnostics 
```

Warning: There were 2854 divergent transitions after warmup. See https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup
to find out why this is a problem and how to eliminate them.
Warning: Examine the pairs() plot to diagnose sampling problems
Warning: The largest R-hat is 1.51, indicating chains have not mixed. Running the chains for more iterations may help. See https://mc-stan.org/misc/warnings.html#r-hat
Warning: Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable. Running the chains for more iterations may help. See https://mc-stan.org/misc/warnings.html#bulk-ess
Warning: Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable.
Running the chains for more iterations may help. See https://mc-stan.org/misc/warnings.html#tail-ess

### Max D1-7 (fit4)

```{r}
all_ucd_size_for_models_climate_fit4 <- all_ucd_size_for_models_climate %>% drop_na(max_temp_d1_7)

climate_fit4 <- brm(formula=climate_f4, data=all_ucd_size_for_models_climate_fit4, prior=prior1, iter = 3800, cores=4)
#Warning: There were 2498 divergent transitions after warmup. See https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup to find out why this is a problem and how to eliminate them.Warning: Examine the pairs() plot to diagnose sampling problems

summary(climate_fit4)
plot(climate_fit4, nvariables = 3, ask=FALSE)
```

### Min D1-13 (fit5)

```{r}
all_ucd_size_for_models_climate_fit5 <- all_ucd_size_for_models_climate %>% drop_na(min_temp_d1_13)

climate_fit5 <- brm(formula=climate_f5, data=all_ucd_size_for_models_climate_fit5, prior=prior1, iter = 3800, cores=4)
#Warning: There were 2498 divergent transitions after warmup. See https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup to find out why this is a problem and how to eliminate them.Warning: Examine the pairs() plot to diagnose sampling problems

summary(climate_fit5)
plot(climate_fit5, nvariables = 3, ask=FALSE)
```

### Mean D1-13 (fit6)

```{r}
all_ucd_size_for_models_climate_fit6 <- all_ucd_size_for_models_climate %>% drop_na(meam_temp_d1_13)

climate_fit6 <- brm(formula=climate_f6, data=all_ucd_size_for_models_climate_fit6, prior=prior1, iter = 3800, cores=4)
#Warning: There were 2498 divergent transitions after warmup. See https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup to find out why this is a problem and how to eliminate them.Warning: Examine the pairs() plot to diagnose sampling problems

summary(climate_fit6)
plot(climate_fit6, nvariables = 3, ask=FALSE)
```

### Max D1-13 (fit7)

```{r}
all_ucd_size_for_models_climate_fit7 <- all_ucd_size_for_models_climate %>% drop_na(max_temp_d1_13)

climate_fit7 <- brm(formula=climate_f7, data=all_ucd_size_for_models_climate_fit7, prior=prior1, iter = 3800, cores=4)
#Warning: There were 2498 divergent transitions after warmup. See https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup to find out why this is a problem and how to eliminate them.Warning: Examine the pairs() plot to diagnose sampling problems

summary(climate_fit7)
plot(climate_fit7, nvariables = 3, ask=FALSE)
```

### Min D6-13 (fit8)

```{r}
all_ucd_size_for_models_climate_fit8 <- all_ucd_size_for_models_climate %>% drop_na(min_temp_d6_13)

climate_fit8 <- brm(formula=climate_f8, data=all_ucd_size_for_models_climate_fit8, prior=prior1, iter = 3800, cores=4)
#Warning: There were 2498 divergent transitions after warmup. See https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup to find out why this is a problem and how to eliminate them.Warning: Examine the pairs() plot to diagnose sampling problems

summary(climate_fit8)
plot(climate_fit8, nvariables = 3, ask=FALSE)
```

### Mean D6-13 (fit9)

```{r}
all_ucd_size_for_models_climate_fit9 <- all_ucd_size_for_models_climate %>% drop_na(mean_temp_d6_13)

climate_fit9 <- brm(formula=climate_f9, data=all_ucd_size_for_models_climate_fit9, prior=prior1, iter = 3800, cores=4)
#Warning: There were 2498 divergent transitions after warmup. See https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup to find out why this is a problem and how to eliminate them.Warning: Examine the pairs() plot to diagnose sampling problems

summary(climate_fit9)
plot(climate_fit9, nvariables = 3, ask=FALSE)
```

### Max D6-13 (fit10)

```{r}
all_ucd_size_for_models_climate_fit10 <- all_ucd_size_for_models_climate %>% drop_na(max_temp_d6_13)

climate_fit10 <- brm(formula=climate_f10, data=all_ucd_size_for_models_climate_fit10, prior=prior1, iter = 3800, cores=4)
#Warning: There were 2498 divergent transitions after warmup. See https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup to find out why this is a problem and how to eliminate them.Warning: Examine the pairs() plot to diagnose sampling problems

summary(climate_fit10)
plot(climate_fit10, nvariables = 3, ask=FALSE)
```

## Model Comparison

```{r}
climate_fit1 <- add_criterion(climate_fit1, "loo")
climate_fit2 <- add_criterion(climate_fit2, "loo")
climate_fit3 <- add_criterion(climate_fit3, "loo")
climate_fit4 <- add_criterion(climate_fit4, "loo")
climate_fit5 <- add_criterion(climate_fit5, "loo") 
climate_fit6 <- add_criterion(climate_fit6, "loo") 
climate_fit7 <- add_criterion(climate_fit7, "loo") 
climate_fit8 <- add_criterion(climate_fit8, "loo") 
climate_fit9 <- add_criterion(climate_fit9, "loo") 
climate_fit10 <- add_criterion(climate_fit10, "loo") 
loo_compare(climate_fit1, climate_fit2, climate_fit3, climate_fit4, climate_fit5, climate_fit6, climate_fit7,climate_fit8, climate_fit9, climate_fit10)
```
