---
title: "UCD_PreTransplantSize-MortPred"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---

# Analysis of Pre-transplant size

To Do:
-   Since such low survival to bud and fruit might be good to have an earlier time point too, like survival through winter?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Relevant Libraries and Functions

```{r}
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
library(lmerTest) #for mixed effect models
conflicted::conflicts_prefer(lmerTest::lmer)
library(broom.mixed) #tidy method for lmerTest
library(emmeans) #for post-hoc pairwise comparisons 
library(naniar) #replaces values with NA
library(corrplot) #plotting correlations 
library(rstatix) #performing cor_test
library(lubridate) #formatting dates
sem <- function(x, na.rm=FALSE) {           #for caclulating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low") #for proper arrangement in figures 
```

## Load the pop and location data

```{r}
#pop info
pops_common_garden <- read_csv("../input/UCD_Data/Pops_for_2022_UCD.csv") #pops included in common garden 
summary(pops_common_garden)
pops_common_garden_nonotes <- pops_common_garden %>% select(parent.pop:elevation.group, UCD.seed.year)
pops_common_garden_nonotes$elevation.group <- str_to_title(pops_common_garden_nonotes$elevation.group)

#extra location info 
pop_loc <- read_csv("../input/Strep_tort_locs.csv")

unique(pop_loc$`Site code`)
unique(pops_common_garden_nonotes$parent.pop)
#need to change YOSE to YO
pop_loc_yo <- pop_loc %>% mutate(parent.pop = str_replace(`Site code`, "YOSE(\\d+)", "YO\\1")) %>% select(Lat, Long, elev_m=`Elevation (m)`, parent.pop)
unique(pop_loc_yo$parent.pop)

#merge in location info
pop_elev <- left_join(pops_common_garden_nonotes, pop_loc_yo)
head(pop_elev)
```

## Load Size Data
```{r}
ucd_pretransplant_size <- read_csv("../input/UCD_Data/CorrectedCSVs/UCD_garden_size_measurements_20221128_corrected.csv")
head(ucd_pretransplant_size) #includes germinated and non germ plants 

ucd_pretransplant_size_germonly <- ucd_pretransplant_size %>% rename(germ = `germinated?`, height_cm = `Height (cm)`, longest_leaf_cm=`Longest leaf (cm)`) %>% filter(germ==1)
head(ucd_pretransplant_size_germonly)
tail(ucd_pretransplant_size_germonly)

#merge with pop loc info 
ucd_pretransplant_size_elev <- left_join(ucd_pretransplant_size_germonly, pop_elev)
head(ucd_pretransplant_size_elev)
tail(ucd_pretransplant_size_elev)

##check sample sizes 
ucd_pretransplant_size_elev %>% 
  filter(!is.na(height_cm)) %>% 
  group_by(parent.pop) %>% 
  summarise(samplesize=n()) %>% 
  arrange(samplesize)
#only 2 WV individuals, 3 LV1 individuals  
```

## Summary plots
```{r}
#Histograms
ucd_pretransplant_size_elev %>% select(height_cm, longest_leaf_cm) %>% 
  names() %>% 
  map(~ggplot(ucd_pretransplant_size_elev, aes_string(x = .)) + 
        geom_histogram())
#prints height first then longest leaf 

#qq plots 
ucd_pretransplant_size_elev %>% select(height_cm, longest_leaf_cm) %>% 
  names() %>% 
  map(~ggplot(ucd_pretransplant_size_elev, aes_string(sample = .)) + 
       stat_qq() +
        stat_qq_line())

#height is right skewed 
#longest leaf is fairly normal 


#Transformations:
ucd_pretransplant_size_elev_transf <- ucd_pretransplant_size_elev %>% 
  mutate(sheight_cm=sqrt(height_cm), logheight_cm=log(height_cm),log10height_cm=log10(height_cm))

ucd_pretransplant_size_elev_transf %>% select(ends_with("cm"))%>% 
  names() %>% 
  map(~ggplot(ucd_pretransplant_size_elev_transf, aes_string(x = .)) + 
        geom_histogram())

ucd_pretransplant_size_elev_transf %>% select(ends_with("cm")) %>% 
  names() %>% 
  map(~ggplot(ucd_pretransplant_size_elev_transf, aes_string(sample = .)) + 
       stat_qq() +
        stat_qq_line())
#square root transformation looks better, but not perfect 
```

## Box Plots
```{r}
ucd_pretransplant_size_elev_transf %>% ggplot(aes(x=parent.pop, y=height_cm)) + geom_boxplot()

ucd_pretransplant_size_elev_transf %>% ggplot(aes(x=parent.pop, y=longest_leaf_cm)) + geom_boxplot()
```

## Calculating means by pop (and elev)

```{r}
ucd_pretransplant_size_elev_transf_summary <- ucd_pretransplant_size_elev_transf %>% 
  group_by(parent.pop, elev_m) %>% 
  summarise(N_height = sum(!is.na(height_cm)), mean_height_cm = mean(height_cm,na.rm=(TRUE)), 
            sem_height_cm=sem(height_cm, na.rm=(TRUE)), N_length = sum(!is.na(longest_leaf_cm)),
            mean_longest_leaf_cm=mean(longest_leaf_cm, na.rm=(TRUE)), 
            sem_longest_leaf_cm=sem(longest_leaf_cm, na.rm=TRUE)) %>% 
  arrange(elev_m)
ucd_pretransplant_size_elev_transf_summary
#write_csv(ucd_pretransplant_size_elev_transf_summary, file ="../output/UCD_Traits/pretransplant_size_summary.csv")
```


## Figures of averages

### Avg Height Figure

```{r}
ucd_pretransplant_size_elev_transf_summary %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_height_cm), y=mean_height_cm, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=mean_height_cm-sem_height_cm,ymax=mean_height_cm+sem_height_cm),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Height (cm)", x="Population", title ="Pre-Transplant") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) 
#ggsave("../output/UCD_Traits/PreTransplant_Height.png", width = 14, height = 9, units = "in")
```

### Avg Longest Leaf Figure

```{r}
ucd_pretransplant_size_elev_transf_summary %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_longest_leaf_cm), y=mean_longest_leaf_cm, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=mean_longest_leaf_cm-sem_longest_leaf_cm,ymax=mean_longest_leaf_cm+sem_longest_leaf_cm),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Longest Leaf (cm)", x="Population", title ="Pre-Transplant") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) 
#ggsave("../output/UCD_Traits/PreTransplant_Longest_Leaf.png", width = 14, height = 9, units = "in")
```

## Mixed Effects Models
```{r}
names(ucd_pretransplant_size_elev_transf)

lmeheight = lmerTest::lmer(sheight_cm ~ (1|parent.pop/mf), data=ucd_pretransplant_size_elev_transf)
summary(lmeheight)
ranova(lmeheight) #both pop and mf significant 
VarCorr(lmeheight) %>% #mf = 9.3% of var and pop = 50.4% of var 
  as.data.frame() %>% 
  mutate(var.pct = vcov/sum(vcov)*100)

lmeheight2 = lmerTest::lmer(sheight_cm ~ elev_m + (1|parent.pop/mf), data=ucd_pretransplant_size_elev_transf)
summary(lmeheight2) #elevation is significant 
ranova(lmeheight2) #both pop and mf still significant 

lmelength = lmerTest::lmer(longest_leaf_cm ~ (1|parent.pop/mf), data=ucd_pretransplant_size_elev_transf)
summary(lmelength)
ranova(lmelength) #both pop and mf significant 
VarCorr(lmelength) %>% #mf = 7.2% of var and pop = 29.1% of var 
  as.data.frame() %>% 
  mutate(var.pct = vcov/sum(vcov)*100)

lmelength2 = lmerTest::lmer(longest_leaf_cm ~ elev_m + (1|parent.pop/mf), data=ucd_pretransplant_size_elev_transf)
summary(lmelength2) #elevaition is significant 
ranova(lmelength2) #both pop and mf still significant 
```


## Load mort/pheno data 
```{r}
ucdmort.pheno <- read_csv("../input/UCD_Data/CorrectedCSVs/UCD_transplants_pheno_mort_20231016_corrected.csv")
head(ucdmort.pheno)

ucdmort.pheno2 <- ucdmort.pheno %>% 
  rename(parent.pop=pop, date.first.bud=`Date First Bud`, date.first.fl=`Date First Flower`, date.first.fruit=`Date First Fruit`, date.last.fl=`Date Last Flower`, date.last.fruit=`Date Last Fruit`, death.date=`Death Date`) %>% 
  filter(parent.pop!="buffer") #remove buffers
head(ucdmort.pheno2,60)
unique(ucdmort.pheno2$parent.pop)
unique(ucdmort.pheno2$date.first.bud)

ucdmort.pheno2 %>% filter(is.na(date.first.bud), !is.na(date.last.fruit)) #no cases where there is later rep info and not a bud date 

ucdmort.pheno2 %>% filter(is.na(date.first.fruit), !is.na(date.last.fruit)) #no cases with last fruit, but not first fruit

ucdmort.pheno2 %>% filter(is.na(date.first.bud), is.na(death.date)) #5 cases where there is no death date and no bud date (so a plant did not reproduce or die)

ucdmort.pheno2 %>% filter(is.na(date.first.fruit), is.na(death.date)) #5 cases where there is no death date and no fruit date (same as above)

ucdmort.pheno2 %>% filter(!is.na(date.first.bud), is.na(date.first.fruit), !is.na(death.date)) #18 cases where a plant started to bud and didn't survive to mature fruits 

ucdmort.pheno2 %>% filter(!is.na(date.first.fruit), !is.na(death.date)) #28 cases with fruit date and death date 

#convert dates to 0s (mortality) or 1s (survival)
#Surv to Bud = survival to bud.date or until other plants were budding (for non-rep, potentially perennial plants)
ucd.survtobud <- ucdmort.pheno2 %>% mutate(SurvtoBud=ifelse(!is.na(date.first.bud), 1,
                                                            ifelse(!is.na(death.date), 0, 1))) %>% 
  select(block:rep, date.first.bud, death.date, SurvtoBud)
head(ucd.survtobud, 60)
summary(ucd.survtobud) #avg = 7.3% surv 

ucd.survtobud %>% filter(is.na(date.first.bud), SurvtoBud==1) #correct!
ucd.survtobud %>% filter(!is.na(death.date), SurvtoBud==1) #46 cases with survival to bud, but later death date 

#Surv to Fruit = survival to fruit date or until other plants were fruiting 
ucd.survtofruit <- ucdmort.pheno2 %>% mutate(SurvtoFruit=ifelse(!is.na(date.first.fruit), 1,
                                                            ifelse(!is.na(death.date), 0, 1))) %>% 
  select(block:rep, date.first.fruit, death.date, SurvtoFruit)
head(ucd.survtofruit, 60)
summary(ucd.survtofruit) #avg = 4.9% survival 

ucd.survtofruit %>% filter(is.na(date.first.fruit), SurvtoFruit==1) #5 potential perennials (all BH)
```

## Merge survival with size
```{r}
head(ucd.survtobud)
head(ucd.survtofruit)
head(ucd_pretransplant_size_elev_transf)

names(ucd.survtobud)
names(ucd.survtofruit)
names(ucd_pretransplant_size_elev_transf)

ucd.all.surv <- full_join(ucd.survtobud, ucd.survtofruit) %>% select(block:rep, death.date, SurvtoBud, SurvtoFruit)
ucd.all.surv

ucd_pretransplant_size_surv <- left_join(ucd_pretransplant_size_elev_transf, ucd.all.surv) %>%
  select(parent.pop:germ, UCD.seed.year, Lat, Long, elev_m, block, row, col, height_cm, longest_leaf_cm, sheight_cm, death.date, SurvtoBud, SurvtoFruit) %>% filter(!is.na(SurvtoBud)) %>% filter(!is.na(height_cm))
dim(ucd_pretransplant_size_surv) #3 plants with pre-transplant size, but no death info, 7 plants with no size measurements
head(ucd_pretransplant_size_surv)
```

## Mean survival by pop
```{r}
ucd_surv_mean <- ucd_pretransplant_size_surv %>% group_by(parent.pop, elev_m) %>% 
  summarise(meanSurvtoBud=mean(SurvtoBud, na.rm = TRUE), sem.SurvtoBud=sem(SurvtoBud, na.rm=TRUE),
            meanSurvtoFruit=mean(SurvtoFruit, na.rm = TRUE), sem.SurvtoFruit=sem(SurvtoFruit, na.rm=TRUE),
            mean_height_cm = mean(height_cm,na.rm=(TRUE)), sem_height_cm=sem(height_cm, na.rm=(TRUE)), 
            mean_longest_leaf_cm=mean(longest_leaf_cm, na.rm=(TRUE)), sem_longest_leaf_cm=sem(longest_leaf_cm, na.rm=TRUE))
ucd_surv_mean

ucd_surv_mean %>% 
  ggplot(aes(x=fct_reorder(parent.pop, meanSurvtoBud), y=meanSurvtoBud, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurvtoBud-sem.SurvtoBud,ymax=meanSurvtoBud+sem.SurvtoBud),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Surv to Bud", x="Population") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) 
ggsave("../output/UCD_Traits/SurvtoBud.png", width = 14, height = 9, units = "in")

ucd_surv_mean %>% 
  ggplot(aes(x=fct_reorder(parent.pop, meanSurvtoFruit), y=meanSurvtoFruit, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurvtoFruit-sem.SurvtoFruit,ymax=meanSurvtoFruit+sem.SurvtoFruit),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Surv to Bud", x="Population") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) 
ggsave("../output/UCD_Traits/SurvtoFruit.png", width = 14, height = 9, units = "in")
```

## Plot size and survival
```{r}
ucd_pretransplant_size_surv %>% ggplot(aes(x=height_cm, y=SurvtoBud)) + geom_point()
ucd_pretransplant_size_surv %>% ggplot(aes(x=height_cm, y=SurvtoFruit)) + geom_point()

ucd_pretransplant_size_surv %>% ggplot(aes(x=longest_leaf_cm, y=SurvtoBud)) + geom_point()
ucd_pretransplant_size_surv %>% ggplot(aes(x=longest_leaf_cm, y=SurvtoFruit)) + geom_point()

#plot w/ population averages 
ucd_surv_mean %>% ggplot(aes(x=mean_height_cm, y=meanSurvtoBud)) + geom_point()
ucd_surv_mean %>% ggplot(aes(x=mean_height_cm, y=meanSurvtoFruit)) + geom_point()

ucd_surv_mean %>% ggplot(aes(x=mean_longest_leaf_cm, y=meanSurvtoBud)) + geom_point()
ucd_surv_mean %>% ggplot(aes(x=mean_longest_leaf_cm, y=meanSurvtoFruit)) + geom_point()
```

## Logistic Regresion 
```{r}
xtabs(~SurvtoBud + height_cm, data = ucd_pretransplant_size_surv) #contingency table 
mylogit_height = glm(SurvtoBud ~ height_cm, data = ucd_pretransplant_size_surv, family = binomial())
summary(mylogit_height) #P<0.0001, int= -4.2632, height= 0.6953
#signficant positive relationship between height and survival 

xtabs(~SurvtoBud + longest_leaf_cm, data = ucd_pretransplant_size_surv)
mylogit_longleaf = glm(SurvtoBud ~ longest_leaf_cm, data = ucd_pretransplant_size_surv, family = binomial())
summary(mylogit_longleaf) #P<0.0001, int=-4.4529, longleaf= 0.8777
#significant positive relationship between longest leaf and survival 
```
The bigger you are pre-transplant, the more likely you are to survive 
