---
title: "Data Exploration"
author: "Brandie Quarles"
date: "2023-08-03"
output: html_document:
  keep_md: yes
---

## Note: There are 3 individuals with rep \# 100 in transplant data sheets that should be analyzed with germination bed data instead!!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Relevant Libraries and Functions

```{r}
library(tidyverse)
library(naniar) #replaces values with NA
sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("high", "mid", "low")
```

# Figuring out what the different data are

## Annual Censuses of Germination and Transplant Beds

Purpose: get more thorough data at peak fruiting

```{r}
#Note: BQC edited the csvs prior to importing to R b/c of issues with the naming of columns and the first row of the transplants sheets not being the heading row (veg, repo info and not block, mf etc...) 
anncensusgerm <- read_csv("../input/UCD_Data/CorrectedCSVs/Annual_census_germination_20230626_corrected.csv")
head(anncensusgerm)
summary(anncensusgerm)

anncensustrans_0522 <- read_csv("../input/UCD_Data/CorrectedCSVs/Annual_Census_Transplants_20230522_corrected.csv")
head(anncensustrans_0522)
summary(anncensustrans_0522)

anncensustrans_0626 <- read_csv("../input/UCD_Data/CorrectedCSVs/Annual_Census_transplants_20230626_corrected.csv")
head(anncensustrans_0626)
summary(anncensustrans_0626)
```

## Germination Surveys Purpose: monitor germination of planted seeds in germ beds (plants that didn't germinate in the lab)

```{r}
germsurvey_0127 = read_csv("../input/UCD_Data/CorrectedCSVs/Germination_survey_20230127_corrected.csv")
summary(germsurvey_0127)
head(germsurvey_0127)

germsurvey_0210 = read_csv("../input/UCD_Data/CorrectedCSVs/Germination_survey_20230210_corrected.csv")
summary(germsurvey_0210)
head(germsurvey_0210)

germsurvey_0508 = read_csv("../input/UCD_Data/CorrectedCSVs/Germination_survey_20230508_corrected.csv")
summary(germsurvey_0508)
head(germsurvey_0508)
```

## Mortality Survey? See the later mort_pheno surveys

```{r}
mortsurvey <- read_csv("../input/UCD_Data/CorrectedCSVs/Mortality_survey_20230217_corrected.csv")
summary(mortsurvey) #859 rows so proably only transplant beds?
head(mortsurvey)

```

## Phenology Survey of Germination Bed? See the later mort-pheno surveys

**If a plant didn't fruit, did flowers fail, or did it flower too late?**

```{r}
pheno_germs <- read_csv("../input/UCD_Data/CorrectedCSVs/Phenology_survey_germination_20230515_corrected.csv")
summary(pheno_germs)
head(pheno_germs)
```

## Weekly Size Surveys of Germination and Transplant Beds - weekly except Yos trip or flooded

```{r}
size_germs_0217
size_germs_0303
size_germs_0317
size_germs_0324
size_germs_0403
size_germs_0410
size_germs_0417
size_germs_0424
size_germs_0501
size_germs_0508
size_germs_0515
size_germs_0605
size_germs_0703
size_germs_0710
size_germs_0717
size_germs_0724


size_trans_1213
size_trans_0127
size_trans_0210
size_trans_0217
size_trans_0303
size_trans_0317
size_trans_0324
size_trans_0403
size_trans_0410
size_trans_0417
size_trans_0424
size_trans_0501
size_trans_0508
size_trans_0605
size_trans_0612
size_trans_0703
size_trans_0710
size_trans_0717
size_trans_0724

ucd_garden_size_1128 <- read_csv("../input/UCD_Data/CorrectedCSVs/UCD_garden_size_measurements_20221128_corrected.csv")
summary(ucd_garden_size_1128) #2264 rows... maybe germs and trans combined?
head (ucd_garden_size_1128)
```
