---
title: "UCD Climate Distance"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  markdown: 
    wrap: 72
---
To Do: 
- Copy this code and substitute the full year climate data for the growth season inputs used here.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Climate Distance at the Davis Garden
What happens if I drop the bioclim?
CWD might change the results too
But maybe similar because the life history differences are allowing them to pick the same time
Annual data might give a different result, worth presenting it both ways because that helps highlight the importance of the life history differences 
To get confidence intervals could do some bootstrapping - samples some of the 30 years and calculates it and so on - Julin could help with the code 


Options:

-   Basic subtraction as in Moran et al 2017 (garden site - home site)

-   Gower's environmental distance metric as in Rutter and Fenster 2007

-   From PCAs "We calculated an overall environmental distance using
    PCA, with EDij corresponding to the Euclidian distance between i and
    j based on PCA1 & PCA2." as in Moran et al 2017

## Relevant Libraries and Functions

```{r}
library(raster)
library(tidyverse)
library(conflicted)
conflicts_prefer(dplyr::select())
conflicts_prefer(dplyr::filter)
library(ggrepel)
library(cowplot)
library(gridExtra)
library(naniar) #replaces values with NA

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
} #legend function for grid_arrange

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")
```

## Home Climates

### Flint

```{r}
pops_flint_avgs <- read_csv("../output/Climate/growthseason_FlintAvgs.csv")
head(pops_flint_avgs)
unique(pops_flint_avgs$parent.pop) #only home pops
unique(pops_flint_avgs$TimePd) # recent and historical timeperiod 

pops_flint_recent_avgs <- pops_flint_avgs %>% filter(TimePd=="Recent")
head(pops_flint_recent_avgs)

pops_flint_historic_avgs <-  pops_flint_avgs %>% filter(TimePd=="Historical")
head(pops_flint_historic_avgs)
names(pops_flint_historic_avgs)
```

### BioClim

```{r}
pops_bioclim_avgs <-  read_csv("../output/Climate/growthseason_BioClimAvgs.csv") 
head(pops_bioclim_avgs)
unique(pops_bioclim_avgs$parent.pop) #only home pops
unique(pops_bioclim_avgs$TimePd) # recent and historical timeperiod 

pops_bioclim_recent_avgs <- pops_bioclim_avgs %>%  filter(TimePd=="Recent")
head(pops_bioclim_recent_avgs)

pops_bioclim_historical_avgs <- pops_bioclim_avgs %>% filter(TimePd=="Historical")
head(pops_bioclim_historical_avgs)
names(pops_bioclim_historical_avgs)
```

## Davis Climate Data (Nov 2022-Oct 2023)

### From Flint (changed this from CIMIS)

```{r}
davis_climate <- read_csv("../output/Climate/flint_climate_UCDpops.csv") %>% 
  filter(parent.pop=="UCD_Garden") %>% 
  filter(year>=2022, year<2024) %>% 
  filter(if_else(year==2022, month=="nov" | month=="dec",
                 month!="nov" & month !="dec")) %>% 
  mutate(tavg = (tmn + tmx)/2, t_diurnal = (tmx-tmn))
head(davis_climate)
summary(davis_climate)

davis_climate_flint <- davis_climate %>% #get the means for the variables to compare to home sites 
  summarise(cwd_Davis=mean(cwd),ppt_Davis=mean(ppt), pck_Davis=mean(pck), tmn_Davis=mean(tmn), tmx_Davis=mean(tmx))
davis_climate_flint
```
## Davis Climate Trends
```{r}
davis_climate$month <- factor(davis_climate$month, levels = c("nov","dec","jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct"))

davis_climate %>% 
  ggplot(aes(x=month,y=cwd)) +
  geom_point()

davis_climate %>% 
  ggplot(aes(x=month,y=tmx)) +
  geom_point()

davis_climate %>% 
  ggplot(aes(x=month,y=tmn)) +
  geom_point()

davis_climate %>% 
  ggplot(aes(x=month,y=ppt)) +
  geom_point()
```

### BioClim

Calculating wettest, driest, warmest, and coldest months

```{r}
davis_wettest_month <- davis_climate %>%  
  slice_max(ppt)

davis_driest_month <- davis_climate %>% 
  slice_min(ppt)

davis_warmest_month <- davis_climate %>% 
  slice_max(tavg)

davis_coldest_month <- davis_climate %>%
  slice_min(tavg)
```

Bio 1, 2, 4, 7, 12, 15

```{r}
bioclim_davis_calc <- davis_climate %>% 
  summarise(ann_tmean=mean(tavg),  #Bio1 - Annual Mean Temperature
            mean_diurnal_range=mean(t_diurnal), #Bio2 - Mean Diurnal Range
            temp_seasonality=sd(tavg), #Bio4 - Temperature Seasonality
            temp_ann_range=(max(tmx))-(min(tmn)), #bio7 - temp annual range
            ann_ppt=sum(ppt), #bio12 - annual precip
            ppt_seasonality=cv(ppt+1)) #bio15 - Precipitation Seasonality (+1 to avoid strange CVs for areas where mean rainfaill is < 1)
bioclim_davis_calc
```

Bio 8(Q), 9(Q), 18(Q), 19(Q)

```{r}
#bio8 = tmean_wettest_month
bio8_davis <- davis_wettest_month %>% 
  dplyr::select(tmean_wettest_month=tavg)

#bio9 = tmean_driest_month
bio9_davis <- davis_driest_month %>% 
  dplyr::select(tmean_driest_month=tavg)

bio8_9_davis <- bind_cols(bio8_davis, bio9_davis)

#bio18 = ppt_warmest_month
bio18_davis <- davis_warmest_month %>% 
  dplyr::select(ppt_warmest_month=ppt)

#bio19 = ppt_coldest_month
bio19_davis <- davis_wettest_month %>% 
  dplyr::select(ppt_coldest_month=ppt)

bio18_19_davis <- bind_cols(bio18_davis, bio19_davis)

all_periods_davis <- bind_cols(bio8_9_davis, bio18_19_davis)
```

Merge all bioclims

```{r}
davis_bioclim_final <- bind_cols(bioclim_davis_calc, all_periods_davis) %>% 
  rename_with(~paste0(., "_Davis"), 1:10)
summary(davis_bioclim_final)
```

## Gower's Climate Distance

(1/P) \* SUM ((absolute value(Ai - Bi)) / range(i)) for each variable

-   P = number of environmental variables = 13 (without CWD)

-   Ai = 30 year avg of that variable for the home site

-   Bi = Nov 2022-Oct 2023 avg of that variable for the Davis garden

-   Range(i) = maximum - minimum of that variable in the whole data set
    (across sites)

### Combine the flint and bioclim variables 

```{r}
davis_climate_all <- bind_cols(davis_climate_flint, davis_bioclim_final)
dim(davis_climate_all)

home_climates_all_recent <- left_join(pops_flint_recent_avgs, pops_bioclim_recent_avgs)
home_climates_all_historic <- left_join(pops_flint_historic_avgs, pops_bioclim_historical_avgs)

davis_home_climate_recent <- bind_cols(davis_climate_all, home_climates_all_recent)
summary(davis_home_climate_recent)

davis_home_climate_historic <- bind_cols(davis_climate_all, home_climates_all_historic)
summary(davis_home_climate_historic)

names(davis_home_climate_historic)
```

### Figure out the range for each variable 

```{r}
davis_range_prep <- davis_climate_all %>% 
  mutate(parent.pop="Davis") %>% 
  rename_with(~str_remove(., "_Davis"), everything())
```

#### Recent
```{r}
range_merge_recent <- bind_rows(home_climates_all_recent, davis_range_prep)
names(range_merge_recent)
summary(range_merge_recent)

davis_home_climate_ranges_recent <- range_merge_recent %>% 
  summarise(cwd_range=max(cwd)-min(cwd),
            ppt_range=max(ppt)-min(ppt), 
            tmn_range=max(tmn)-min(tmn), 
            tmx_range=max(tmx)-min(tmx), 
            ann_tmean_range=max(ann_tmean)-min(ann_tmean),
            mean_diurnal_range_range=max(mean_diurnal_range)-min(mean_diurnal_range),
            temp_seasonality_range=max(temp_seasonality)-min(temp_seasonality),
            temp_ann_range_range=max(temp_ann_range)-min(temp_ann_range),
            tmean_wettest_month_range=max(tmean_wettest_month)-min(tmean_wettest_month),
            tmean_driest_month_range=max(tmean_driest_month)-min(tmean_driest_month),
            ann_ppt_range=max(ann_ppt)-min(ann_ppt), 
            ppt_seasonality_range=max(ppt_seasonality)-min(ppt_seasonality),
            ppt_warmest_month_range=max(ppt_warmest_month)-min(ppt_warmest_month), 
            ppt_coldest_month_range=max(ppt_coldest_month)-min(ppt_coldest_month))
davis_home_climate_ranges_recent

davis_home_climate_with_ranges_recent <- bind_cols(davis_home_climate_recent, davis_home_climate_ranges_recent)
names(davis_home_climate_with_ranges_recent)
```

#### Historic
```{r}
range_merge_historic <- bind_rows(home_climates_all_historic, davis_range_prep)
names(range_merge_historic)
summary(range_merge_historic)

davis_home_climate_ranges_historic <- range_merge_historic %>% 
  summarise(cwd_range=max(cwd)-min(cwd),
            ppt_range=max(ppt)-min(ppt), 
            tmn_range=max(tmn)-min(tmn), 
            tmx_range=max(tmx)-min(tmx), 
            ann_tmean_range=max(ann_tmean)-min(ann_tmean),
            mean_diurnal_range_range=max(mean_diurnal_range)-min(mean_diurnal_range),
            temp_seasonality_range=max(temp_seasonality)-min(temp_seasonality),
            temp_ann_range_range=max(temp_ann_range)-min(temp_ann_range),
            tmean_wettest_month_range=max(tmean_wettest_month)-min(tmean_wettest_month),
            tmean_driest_month_range=max(tmean_driest_month)-min(tmean_driest_month),
            ann_ppt_range=max(ann_ppt)-min(ann_ppt), 
            ppt_seasonality_range=max(ppt_seasonality)-min(ppt_seasonality),
            ppt_warmest_month_range=max(ppt_warmest_month)-min(ppt_warmest_month), 
            ppt_coldest_month_range=max(ppt_coldest_month)-min(ppt_coldest_month))
davis_home_climate_ranges_historic

davis_home_climate_with_ranges_historic <- bind_cols(davis_home_climate_historic, davis_home_climate_ranges_historic)
names(davis_home_climate_with_ranges_historic)
```

### Recent Gowers Calc
```{r}
gowers_calc_each_var_recent <- davis_home_climate_with_ranges_recent %>% 
  mutate(cwd_gowers=abs(cwd_Davis-cwd) / cwd_range,
         ppt_gowers=abs(ppt_Davis - ppt) / ppt_range,
         tmn_gowers=abs(tmn_Davis - tmn) / tmn_range,
         tmx_gowers=abs(tmx_Davis - tmx) / tmx_range,
         ann_tmean_gowers=abs(ann_tmean_Davis - ann_tmean) / ann_tmean_range,
         mean_diurnal_range_gowers=abs(mean_diurnal_range_Davis - mean_diurnal_range) / mean_diurnal_range_range,
         temp_seasonality_gowers=abs(temp_seasonality_Davis - temp_seasonality) / temp_seasonality_range,
         temp_ann_range_gowers=abs(temp_ann_range_Davis - temp_ann_range) / temp_ann_range_range,
         tmean_wettest_month_gowers=abs(tmean_wettest_month_Davis - tmean_wettest_month) / tmean_wettest_month_range,
         tmean_driest_month_gowers=abs(tmean_driest_month_Davis - tmean_driest_month) / tmean_driest_month_range,
         ann_ppt_gowers=abs(ann_ppt_Davis - ann_ppt) / ann_ppt_range,
         ppt_seasonality_gowers=abs(ppt_seasonality_Davis - ppt_seasonality) / ppt_seasonality_range,
         ppt_warmest_month_gowers=abs(ppt_warmest_month_Davis - ppt_warmest_month) / ppt_warmest_month_range,
         ppt_coldest_month_gowers=abs(ppt_coldest_month_Davis - ppt_coldest_month) / ppt_coldest_month_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))

gowers_calc_per_pop_recent <- gowers_calc_each_var_recent %>% 
  mutate(Recent_Gowers_Dist=(1/14)*(cwd_gowers + ppt_gowers + tmn_gowers + tmx_gowers +
                                ann_tmean_gowers + mean_diurnal_range_gowers +
                                temp_seasonality_gowers +temp_ann_range_gowers +
                                tmean_wettest_month_gowers +
                                tmean_driest_month_gowers +ann_ppt_gowers +
                                ppt_seasonality_gowers + ppt_warmest_month_gowers +
                                ppt_coldest_month_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Recent_Gowers_Dist)

gowers_calc_per_pop_recent
```

### Historic Gowers Calc
```{r}
gowers_calc_each_var_historic <- davis_home_climate_with_ranges_historic %>% 
  mutate(cwd_gowers=abs(cwd_Davis-cwd) / cwd_range,
         ppt_gowers=abs(ppt_Davis - ppt) / ppt_range,
         tmn_gowers=abs(tmn_Davis - tmn) / tmn_range,
         tmx_gowers=abs(tmx_Davis - tmx) / tmx_range,
         ann_tmean_gowers=abs(ann_tmean_Davis - ann_tmean) / ann_tmean_range,
         mean_diurnal_range_gowers=abs(mean_diurnal_range_Davis - mean_diurnal_range) / mean_diurnal_range_range,
         temp_seasonality_gowers=abs(temp_seasonality_Davis - temp_seasonality) / temp_seasonality_range,
         temp_ann_range_gowers=abs(temp_ann_range_Davis - temp_ann_range) / temp_ann_range_range,
         tmean_wettest_month_gowers=abs(tmean_wettest_month_Davis - tmean_wettest_month) / tmean_wettest_month_range,
         tmean_driest_month_gowers=abs(tmean_driest_month_Davis - tmean_driest_month) / tmean_driest_month_range,
         ann_ppt_gowers=abs(ann_ppt_Davis - ann_ppt) / ann_ppt_range,
         ppt_seasonality_gowers=abs(ppt_seasonality_Davis - ppt_seasonality) / ppt_seasonality_range,
         ppt_warmest_month_gowers=abs(ppt_warmest_month_Davis - ppt_warmest_month) / ppt_warmest_month_range,
         ppt_coldest_month_gowers=abs(ppt_coldest_month_Davis - ppt_coldest_month) / ppt_coldest_month_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))

  
gowers_calc_per_pop_historic <- gowers_calc_each_var_historic %>% 
  mutate(Historic_Gowers_Dist=(1/14)*(cwd_gowers + ppt_gowers + tmn_gowers + tmx_gowers +
                                ann_tmean_gowers + mean_diurnal_range_gowers +
                                temp_seasonality_gowers +temp_ann_range_gowers +
                                tmean_wettest_month_gowers +
                                tmean_driest_month_gowers +ann_ppt_gowers +
                                ppt_seasonality_gowers + ppt_warmest_month_gowers +
                                ppt_coldest_month_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Historic_Gowers_Dist)

gowers_calc_per_pop_historic
```

### Merge recent and historic
```{r}
gowers_all_time <- full_join(gowers_calc_per_pop_recent, gowers_calc_per_pop_historic)
gowers_all_time
names(gowers_all_time)
write_csv(gowers_all_time, "../output/Climate/growthseason_GowersEnvtalDist_UCD.csv")
```

Figures
```{r}
gowers_rec_fig <- gowers_all_time %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Recent_Gowers_Dist), y=Recent_Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from Davis", fill="Elevation (m)", x="Population", title="Recent Climate") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_Gowers_Recent_fromDavis.png", width = 12, height = 6, units = "in")

gowers_hist_fig <- gowers_all_time %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Historic_Gowers_Dist), y=Historic_Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population", title="Historic Climate", y="Gowers Envtal Distance \n from Davis") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/growthseason_Gowers_Historic_fromDavis.png", width = 12, height = 6, units = "in")

#should combine these into one figure and save that instead
legend <- get_legend(gowers_rec_fig)
gowers_hist_fig <- gowers_hist_fig + theme(legend.position="none")
gowers_rec_fig <- gowers_rec_fig + theme(legend.position="none")
grid.arrange(gowers_hist_fig, gowers_rec_fig, legend, ncol=3, widths=c(3.12, 3.12, 1.09))
 #2000 x 850
```


## Flint Climate Distance

### Gowers

#### Recent
```{r}
gowers_calc_each_var_recent_flint <- davis_home_climate_with_ranges_recent %>% 
  mutate(cwd_gowers=abs(cwd_Davis-cwd) / cwd_range,
         ppt_gowers=abs(ppt_Davis - ppt) / ppt_range,
         tmn_gowers=abs(tmn_Davis - tmn) / tmn_range,
         tmx_gowers=abs(tmx_Davis - tmx) / tmx_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))
  
gowers_calc_per_pop_recent_flint <- gowers_calc_each_var_recent_flint %>% 
  mutate(Recent_Gowers_Dist=(1/4)*(cwd_gowers + ppt_gowers + tmn_gowers + tmx_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Recent_Gowers_Dist)

gowers_calc_per_pop_recent_flint
```

#### Historic
```{r}
gowers_calc_each_var_historic_flint <- davis_home_climate_with_ranges_historic %>% 
  mutate(cwd_gowers=abs(cwd_Davis-cwd) / cwd_range,
         ppt_gowers=abs(ppt_Davis - ppt) / ppt_range,
         tmn_gowers=abs(tmn_Davis - tmn) / tmn_range,
         tmx_gowers=abs(tmx_Davis - tmx) / tmx_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))

  
gowers_calc_per_pop_historic_flint <- gowers_calc_each_var_historic_flint %>% 
  mutate(Historic_Gowers_Dist=(1/4)*(cwd_gowers + ppt_gowers + tmn_gowers + tmx_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Historic_Gowers_Dist)

gowers_calc_per_pop_historic_flint
```

#### Merge recent and historic
```{r}
gowers_all_time_flint <- full_join(gowers_calc_per_pop_recent_flint, gowers_calc_per_pop_historic_flint)
gowers_all_time_flint
names(gowers_all_time_flint)
write_csv(gowers_all_time_flint, "../output/Climate/growthseason_GowersEnvtalDist_UCDFlint.csv")
```

Figures
```{r}
gowers_rec_fig_flint <- gowers_all_time_flint %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Recent_Gowers_Dist), y=Recent_Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from Davis", fill="Elevation (m)", x="Population", title="Recent Climate") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/growthseason_Gowers_RecentFlint_fromDavis.png", width = 12, height = 6, units = "in")

gowers_hist_fig_flint <- gowers_all_time_flint %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Historic_Gowers_Dist), y=Historic_Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population", title="Historic Climate", y="Gowers Envtal Distance \n from Davis") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/growthseason_Gowers_HistoricFlint_fromDavis.png", width = 12, height = 6, units = "in")

#should combine these into one figure and save that instead
legend <- get_legend(gowers_rec_fig_flint)
gowers_hist_fig_flint <- gowers_hist_fig_flint + theme(legend.position="none")
gowers_rec_fig_flint <- gowers_rec_fig_flint + theme(legend.position="none")
grid.arrange(gowers_hist_fig_flint, gowers_rec_fig_flint, legend, ncol=3, widths=c(3.12, 3.12, 1.09))
 #2000 x 850
```

### Subtraction

```{r}
recent_flint_dist_prep <- bind_cols(davis_climate_flint, pops_flint_recent_avgs)
names(recent_flint_dist_prep)

recent_flint_dist <- recent_flint_dist_prep %>% 
  mutate(cwd_dist=cwd_Davis - cwd,
         ppt_dist=ppt_Davis - ppt,
         tmn_dist=tmn_Davis - tmn,
         tmx_dist=tmx_Davis - tmx) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))

historic_flint_dist_prep <- bind_cols(davis_climate_flint, pops_flint_historic_avgs)
names(historic_flint_dist_prep)

historic_flint_dist <- historic_flint_dist_prep %>% 
  mutate(cwd_dist=cwd_Davis - cwd,
         ppt_dist=ppt_Davis - ppt,
         tmn_dist=tmn_Davis - tmn,
         tmx_dist=tmx_Davis - tmx) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
```

Figures Recent (subtraction distance)

```{r}
recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, cwd_dist), y=cwd_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_MeanCWD_DistfromDavis_RecentClim.png", width = 12, height = 6, units = "in")

recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_dist), y=ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_MeanPPT_DistfromDavis_RecentClim.png", width = 12, height = 6, units = "in")

recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmn_dist), y=tmn_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_MeanTMN_DistfromDavis_RecentClim.png", width = 12, height = 6, units = "in")

recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmx_dist), y=tmx_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_MeanTMX_DistfromDavis_RecentClim.png", width = 12, height = 6, units = "in")
```

Figures Historical (subtraction distance)

```{r}
historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, cwd_dist), y=cwd_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_MeanCWD_DistfromDavis_HistoricalClim.png", width = 12, height = 6, units = "in")

historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_dist), y=ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_MeanPPT_DistfromDavis_HistoricalClim.png", width = 12, height = 6, units = "in")

historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmn_dist), y=tmn_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_MeanTMN_DistfromDavis_HistoricalClim.png", width = 12, height = 6, units = "in")

historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmx_dist), y=tmx_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_MeanTMX_DistfromDavis_HistoricalClim.png", width = 12, height = 6, units = "in")
```

## Bioclim Climate Distance

### Gowers

#### Recent
```{r}
gowers_calc_each_var_recent_bioclim <- davis_home_climate_with_ranges_recent %>% 
  mutate(ann_tmean_gowers=abs(ann_tmean_Davis - ann_tmean) / ann_tmean_range,
         mean_diurnal_range_gowers=abs(mean_diurnal_range_Davis - mean_diurnal_range) / mean_diurnal_range_range,
         temp_seasonality_gowers=abs(temp_seasonality_Davis - temp_seasonality) / temp_seasonality_range,
         temp_ann_range_gowers=abs(temp_ann_range_Davis - temp_ann_range) / temp_ann_range_range,
         tmean_wettest_month_gowers=abs(tmean_wettest_month_Davis - tmean_wettest_month) / tmean_wettest_month_range,
         tmean_driest_month_gowers=abs(tmean_driest_month_Davis - tmean_driest_month) / tmean_driest_month_range,
         ann_ppt_gowers=abs(ann_ppt_Davis - ann_ppt) / ann_ppt_range,
         ppt_seasonality_gowers=abs(ppt_seasonality_Davis - ppt_seasonality) / ppt_seasonality_range,
         ppt_warmest_month_gowers=abs(ppt_warmest_month_Davis - ppt_warmest_month) / ppt_warmest_month_range,
         ppt_coldest_month_gowers=abs(ppt_coldest_month_Davis - ppt_coldest_month) / ppt_coldest_month_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))
  
gowers_calc_per_pop_recent_bioclim <- gowers_calc_each_var_recent_bioclim %>% 
  mutate(Recent_Gowers_Dist=(1/10)*(ann_tmean_gowers + mean_diurnal_range_gowers +
                                temp_seasonality_gowers +temp_ann_range_gowers +
                                tmean_wettest_month_gowers +
                                tmean_driest_month_gowers +ann_ppt_gowers +
                                ppt_seasonality_gowers + ppt_warmest_month_gowers +
                                ppt_coldest_month_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Recent_Gowers_Dist)

gowers_calc_per_pop_recent_bioclim
```

#### Historic
```{r}
gowers_calc_each_var_historic_bioclim <- davis_home_climate_with_ranges_historic %>% 
  mutate(ann_tmean_gowers=abs(ann_tmean_Davis - ann_tmean) / ann_tmean_range,
         mean_diurnal_range_gowers=abs(mean_diurnal_range_Davis - mean_diurnal_range) / mean_diurnal_range_range,
         temp_seasonality_gowers=abs(temp_seasonality_Davis - temp_seasonality) / temp_seasonality_range,
         temp_ann_range_gowers=abs(temp_ann_range_Davis - temp_ann_range) / temp_ann_range_range,
         tmean_wettest_month_gowers=abs(tmean_wettest_month_Davis - tmean_wettest_month) / tmean_wettest_month_range,
         tmean_driest_month_gowers=abs(tmean_driest_month_Davis - tmean_driest_month) / tmean_driest_month_range,
         ann_ppt_gowers=abs(ann_ppt_Davis - ann_ppt) / ann_ppt_range,
         ppt_seasonality_gowers=abs(ppt_seasonality_Davis - ppt_seasonality) / ppt_seasonality_range,
         ppt_warmest_month_gowers=abs(ppt_warmest_month_Davis - ppt_warmest_month) / ppt_warmest_month_range,
         ppt_coldest_month_gowers=abs(ppt_coldest_month_Davis - ppt_coldest_month) / ppt_coldest_month_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))

  
gowers_calc_per_pop_historic_bioclim <- gowers_calc_each_var_historic_bioclim %>% 
  mutate(Historic_Gowers_Dist=(1/10)*(ann_tmean_gowers + mean_diurnal_range_gowers +
                                temp_seasonality_gowers +temp_ann_range_gowers +
                                tmean_wettest_month_gowers +
                                tmean_driest_month_gowers +ann_ppt_gowers +
                                ppt_seasonality_gowers + ppt_warmest_month_gowers +
                                ppt_coldest_month_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Historic_Gowers_Dist)

gowers_calc_per_pop_historic_bioclim
```

#### Merge recent and historic
```{r}
gowers_all_time_bioclim <- full_join(gowers_calc_per_pop_recent_bioclim, gowers_calc_per_pop_historic_bioclim)
gowers_all_time_bioclim
names(gowers_all_time_bioclim)
write_csv(gowers_all_time_bioclim, "../output/Climate/growthseason_GowersEnvtalDist_UCDbioclim.csv")
```

Figures
```{r}
gowers_rec_fig_bioclim <- gowers_all_time_bioclim %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Recent_Gowers_Dist), y=Recent_Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from Davis", fill="Elevation (m)", x="Population", title="Recent Climate") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/growthseason_Gowers_Recentbioclim_fromDavis.png", width = 12, height = 6, units = "in")

gowers_hist_fig_bioclim <- gowers_all_time_bioclim %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Historic_Gowers_Dist), y=Historic_Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population", title="Historic Climate", y="Gowers Envtal Distance \n from Davis") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/growthseason_Gowers_Historicbioclim_fromDavis.png", width = 12, height = 6, units = "in")

#should combine these into one figure and save that instead
legend <- get_legend(gowers_rec_fig_bioclim)
gowers_hist_fig_bioclim <- gowers_hist_fig_bioclim + theme(legend.position="none")
gowers_rec_fig_bioclim <- gowers_rec_fig_bioclim + theme(legend.position="none")
grid.arrange(gowers_hist_fig_bioclim, gowers_rec_fig_bioclim, legend, ncol=3, widths=c(3.12, 3.12, 1.09))
 #2000 x 850
```

### Subtraction

```{r}
#Recent
names(pops_bioclim_recent_avgs)
recent_bioclim_dist_prep <- bind_cols(davis_bioclim_final, pops_bioclim_recent_avgs)
recent_bioclim_dist <- recent_bioclim_dist_prep %>% 
  mutate(ann_tmean_dist=ann_tmean_Davis - ann_tmean,
         mean_diurnal_range_dist=mean_diurnal_range_Davis - mean_diurnal_range,
         temp_seasonality_dist=temp_seasonality_Davis - temp_seasonality,
         temp_ann_range_dist=temp_ann_range_Davis - temp_ann_range,
         tmean_wettest_month_dist=tmean_wettest_month_Davis - tmean_wettest_month,
         tmean_driest_month_dist=tmean_driest_month_Davis - tmean_driest_month,
         ann_ppt_dist=ann_ppt_Davis - ann_ppt,
         ppt_seasonality_dist=ppt_seasonality_Davis - ppt_seasonality, 
         ppt_warmest_month_dist=ppt_warmest_month_Davis - ppt_warmest_month,
         ppt_coldest_month_dist=ppt_coldest_month_Davis - ppt_coldest_month) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
recent_bioclim_dist

#Historical
names(pops_bioclim_historical_avgs)
historical_bioclim_dist_prep <- bind_cols(davis_bioclim_final, pops_bioclim_historical_avgs)
historical_bioclim_dist <- historical_bioclim_dist_prep %>% 
  mutate(ann_tmean_dist=ann_tmean_Davis - ann_tmean,
         mean_diurnal_range_dist=mean_diurnal_range_Davis - mean_diurnal_range,
         temp_seasonality_dist=temp_seasonality_Davis - temp_seasonality,
         temp_ann_range_dist=temp_ann_range_Davis - temp_ann_range,
         tmean_wettest_month_dist=tmean_wettest_month_Davis - tmean_wettest_month,
         tmean_driest_month_dist=tmean_driest_month_Davis - tmean_driest_month,
         ann_ppt_dist=ann_ppt_Davis - ann_ppt,
         ppt_seasonality_dist=ppt_seasonality_Davis - ppt_seasonality, 
         ppt_warmest_month_dist=ppt_warmest_month_Davis - ppt_warmest_month,
         ppt_coldest_month_dist=ppt_coldest_month_Davis - ppt_coldest_month) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
historical_bioclim_dist
```

Figures Recent (subtraction distance)

```{r}
recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_dist), y=ann_tmean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_Ann_Tmean_DistfromDavis_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_dist), y=mean_diurnal_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_Diurnal_Range_DistfromDavis_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_dist), y=temp_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_Temp_Seasonality_DistfromDavis_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_dist), y=temp_ann_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_Temp_Ann_Range_DistfromDavis_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_month_dist), y=tmean_wettest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_Temp_Wet_DistfromDavis_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_month_dist), y=tmean_driest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_Temp_Dry_DistfromDavis_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_dist), y=ann_ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_Ann_PPT_DistfromDavis_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_dist), y=ppt_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_PPT_Seasonality_DistfromDavis_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_month_dist), y=ppt_warmest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_PPT_Warm_DistfromDavis_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_month_dist), y=ppt_coldest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_PPT_Cold_DistfromDavis_RecentClim.png", width = 12, height = 6, units = "in")
```

Historical (subtraction distance)

```{r}
historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_dist), y=ann_tmean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_Ann_Tmean_DistfromDavis_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_dist), y=mean_diurnal_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_Diurnal_Range_DistfromDavis_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_dist), y=temp_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_Temp_Seasonality_DistfromDavis_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_dist), y=temp_ann_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_Temp_Ann_Range_DistfromDavis_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_month_dist), y=tmean_wettest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_Temp_Wet_DistfromDavis_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_month_dist), y=tmean_driest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_Temp_Dry_DistfromDavis_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_dist), y=ann_ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_Ann_PPT_DistfromDavis_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_dist), y=ppt_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_PPT_Seasonality_DistfromDavis_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_month_dist), y=ppt_warmest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_PPT_Warm_DistfromDavis_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_month_dist), y=ppt_coldest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/grwssn_PPT_Cold_DistfromDavis_HistoricalClim.png", width = 12, height = 6, units = "in")
```
