---
title: "UCD Climate Distance"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Climate Distance at the Davis Garden
Options:
-   Basic subtraction as in Moran et al 2017 (garden site - home site)
-   Gower’s environmental distance metric as in Rutter and Fenster 2007
-   From PCAs "We calculated an overall environmental distance using PCA, with EDij correspond- ing to the Euclidian distance between i and j based on PCA1 & PCA2." as in Moran et al 2017

## Relevant Libraries and Functions

```{r}
library(tidyverse)
library(ggrepel)
library(cowplot)
library(gridExtra)
library(corrplot) #plotting correlations 
library(rstatix) #performing cor_test
library(naniar) #replaces values with NA
library(QBMS) #for function calc_biovars to calculate bioclim variables
sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
} #legend function for grid_arrange

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")

#For scree plots 
#library("devtools") #The package devtools is required for the installation as factoextra is hosted on github.
#1install_github("kassambara/factoextra")
library("factoextra")
```

## Home Climates (Flint Data)
```{r}

```


## Davis Climate Data (Nov 2022-Oct 2023)
Paltineanu et al 2012 calcualted climate water deficit as ppt (as a sum) - ref evaopotranspiration (refet). But I don't think this is correct b/c I got negative numbers for that calcualtion and minimum CWD is supposed to be 0...
```{r}
davis_climate <- read_csv("../input/CIMS_Davis_monthly.csv")
head(davis_climate)

davis_climate_short <- davis_climate %>% 
  select(month_year=`Month Year`, refet=`Total ETo (mm)`, ppt=`Total Precip (mm)`,
         avg_sol_rad=`Avg Sol Rad (W/sq.m)`, avg_vapor_pres=`Avg Vap Pres (kPa)`,
         tmx=`Avg Max Air Temp (C)`, tmn=`Avg Min Air Temp (C)`, tavg=`Avg Air Temp (C)`,
         avg_max_rel_hum=`Avg Max Rel Hum (%)`, avg_min_rel_hum=`Avg Min Rel Hum (%)`,
         avg_rel_hum=`Avg Rel Hum (%)`, avg_dew_pt=`Avg Dew Point (C)`,
         avg_wind_spd=`Avg Wind Speed (m/s)`, avg_soil_temp=`Avg Soil Temp (C)`) %>% 
  separate(month_year, c("month", "year"), " ", remove = FALSE) %>% 
  mutate(month=match(month, month.abb), cwd=ppt-refet)
davis_climate_short
summary(davis_climate_short)
```

## BioClim Variables (https://www.worldclim.org/data/bioclim.html#google_vignette)
-   annual mean temperature (BIO1)
-   mean diurnal range (BIO2)
-   temperature seasonality (BIO4) (standard deviation *100)
-   temperature annual range (BIO7) (Max Temperature of Warmest Month - Min Temperature of Coldest Month)
-   mean temp of wettest quarter (BIO8)
-   mean temp of driest quarter (BIO9)
-   annual precipitation (BIO12) - sum of ppt for the entire year (not the avg)
-   precipitation seasonality (BIO15)  (Coefficient of Variation)
-   precip of warmest quarter (BIO18)
-   precip of coldest quarter (BIO19)

Use library QBMS, calc_biovars function: https://search.r-project.org/CRAN/refmans/QBMS/html/calc_biovars.html 
Data.frame has 4 mandatory columns (year, ppt, tmin, and tmax), and 12 rows (months) for each year sorted from Jan to Dec.

For home climate 
```{r}
pops_bioclim_recent_avgs <- read_csv("../output/Climate/Pops_BioClimAvgs_Recent.csv") %>% 
  select(parent.pop:ppt_coldest_quarter_avg)
head(pops_bioclim_recent_avgs)

pops_bioclim_historical_avgs <- read_csv("../output/Climate/Pops_BioClimAvgs_Historical.csv")  %>% 
  select(parent.pop:ppt_coldest_quarter_avg)
head(pops_bioclim_historical_avgs)
```

For Davis Climate
Had to exclude 2022
```{r}
davis_bioclim <- davis_climate_short %>% 
  filter(year=="2023") %>% 
  rename(tmin=tmn, tmax=tmx) %>% 
  calc_biovars()
davis_bioclim

davis_bioclim_final <- davis_bioclim %>% 
  mutate(across(c(bio1, bio2, bio4, bio7, bio8, bio9, bio12, bio15, bio18, bio19), as.numeric)) %>% 
  select(ann_tmean_Davis=bio1, mean_diurnal_range_Davis=bio2,
         temp_seasonality_Davis=bio4, temp_ann_range_Davis=bio7,
         tmean_wettest_quarter_Davis=bio8, tmean_driest_quarter_Davis=bio9,
         ann_ppt_Davis=bio12, ppt_seasonality_Davis=bio15, 
         ppt_warmest_quarter_Davis=bio18, ppt_coldest_quarter_Davis=bio19) 
davis_bioclim_final
```


## Davis Climate Trends
```{r}
davis_climate_short$month_year <- factor(davis_climate_short$month_year, levels = c("Nov 2022","Dec 2022","Jan 2023", "Feb 2023", "Mar 2023", "Apr 2023", "May 2023", "Jun 2023", "Jul 2023", "Aug 2023", "Sep 2023", "Oct 2023"))
davis_climate_short %>% ggplot(aes(x=month_year, y=tavg)) + geom_point() + theme_classic()
davis_climate_short %>% ggplot(aes(x=month_year, y=ppt)) + geom_point() + theme_classic()
davis_climate_short %>% ggplot(aes(x=month_year, y=tmx)) + geom_point() + theme_classic()
davis_climate_short %>% ggplot(aes(x=month_year, y=avg_rel_hum)) + geom_point() + theme_classic()
davis_climate_short %>% ggplot(aes(x=month_year, y=avg_wind_spd)) + geom_point() + theme_classic()
davis_climate_short %>% ggplot(aes(x=month_year, y=avg_sol_rad)) + geom_point() + theme_classic()
```


## Bioclim Climate Distance

### Subtraction
```{r}
#Recent
names(pops_bioclim_recent_avgs)
recent_bioclim_dist <- bind_cols(davis_bioclim_final, pops_bioclim_recent_avgs)
recent_bioclim_dist <- recent_bioclim_dist %>% 
  mutate(ann_tmean_dist=ann_tmean_Davis - ann_tmean_avg,
         mean_diurnal_range_dist=mean_diurnal_range_Davis - mean_diurnal_range_avg,
         temp_seasonality_dist=temp_seasonality_Davis - temp_seasonality_avg,
         temp_ann_range_dist=temp_ann_range_Davis - temp_ann_range_avg,
         tmean_wettest_quarter_dist=tmean_wettest_quarter_Davis - tmean_wettest_quarter_avg,
         tmean_driest_quarter_dist=tmean_driest_quarter_Davis - tmean_driest_quarter_avg,
         ann_ppt_dist=ann_ppt_Davis - ann_ppt_avg,
         ppt_seasonality_dist=ppt_seasonality_Davis - ppt_seasonality_avg, 
         ppt_warmest_quarter_dist=ppt_warmest_quarter_Davis - ppt_warmest_quarter_avg,
         ppt_coldest_quarter_dist=ppt_coldest_quarter_Davis - ppt_coldest_quarter_avg) %>% 
  select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
recent_bioclim_dist

#Historical
names(pops_bioclim_historical_avgs)
historical_bioclim_dist <- bind_cols(davis_bioclim_final, pops_bioclim_historical_avgs)
historical_bioclim_dist <- historical_bioclim_dist %>% 
  mutate(ann_tmean_dist=ann_tmean_Davis - ann_tmean_avg,
         mean_diurnal_range_dist=mean_diurnal_range_Davis - mean_diurnal_range_avg,
         temp_seasonality_dist=temp_seasonality_Davis - temp_seasonality_avg,
         temp_ann_range_dist=temp_ann_range_Davis - temp_ann_range_avg,
         tmean_wettest_quarter_dist=tmean_wettest_quarter_Davis - tmean_wettest_quarter_avg,
         tmean_driest_quarter_dist=tmean_driest_quarter_Davis - tmean_driest_quarter_avg,
         ann_ppt_dist=ann_ppt_Davis - ann_ppt_avg,
         ppt_seasonality_dist=ppt_seasonality_Davis - ppt_seasonality_avg, 
         ppt_warmest_quarter_dist=ppt_warmest_quarter_Davis - ppt_warmest_quarter_avg,
         ppt_coldest_quarter_dist=ppt_coldest_quarter_Davis - ppt_coldest_quarter_avg) %>% 
  select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
historical_bioclim_dist
```

Figures
Recent (subtraction distance)
```{r}
recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_dist), y=ann_tmean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_dist), y=mean_diurnal_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_dist), y=temp_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_dist), y=temp_ann_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_quarter_dist), y=tmean_wettest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_quarter_dist), y=tmean_driest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_dist), y=ann_ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_dist), y=ppt_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_quarter_dist), y=ppt_warmest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_quarter_dist), y=ppt_coldest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))
```

Historical (subtraction distance)
```{r}
historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_dist), y=ann_tmean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_dist), y=mean_diurnal_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_dist), y=temp_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_dist), y=temp_ann_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_quarter_dist), y=tmean_wettest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_quarter_dist), y=tmean_driest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_dist), y=ann_ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_dist), y=ppt_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_quarter_dist), y=ppt_warmest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_quarter_dist), y=ppt_coldest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=14), axis.text.x = element_text(angle = 45,  hjust = 1))
```

