---
title: "UCD Climate Distance"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Climate Distance at the Davis Garden

Options:

-   Basic subtraction as in Moran et al 2017 (garden site - home site)

-   Gower's environmental distance metric as in Rutter and Fenster 2007

-   From PCAs "We calculated an overall environmental distance using
    PCA, with EDij corresponding to the Euclidian distance between i and
    j based on PCA1 & PCA2." as in Moran et al 2017

Davis CWD

Evapotranspiration can be computed as reference, potential, or actual
evapotranspiration.   Reference evapotranspiration is that from a grass
surface that is well-watered.  Potential evapotranspiration is that from
a surface that has unlimited water (such as a lake). 

-   CWD = PET - AET

    -   PET = modeled on an hourly basis from solar radiation that is
        modeled using topgraphic shading and cloudiness

        -   Priestley-Taylor Equation

        -   Still figuring out how to calculate this. The people we
            reached out to say Eto(ref) is similar to PET, but not quite
            the same...

    -   AET = PET x monthly Kv for each vegtype when soil is above
        wilting point

        -   Kv values used by flint for different veg types is in the
            control file - can't access the control file because they're
            currently revising the data release

## Relevant Libraries and Functions

```{r}
library(raster)
library(tidyverse)
library(conflicted)
conflicts_prefer(dplyr::select())
conflicts_prefer(dplyr::filter)
library(ggrepel)
library(cowplot)
library(gridExtra)
library(corrplot) #plotting correlations 
library(rstatix) #performing cor_test
library(naniar) #replaces values with NA
sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
} #legend function for grid_arrange

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")

#For scree plots 
#library("devtools") #The package devtools is required for the installation as factoextra is hosted on github.
#1install_github("kassambara/factoextra")
library("factoextra")
```

## Home Climates

### Flint

```{r}
pops_flint_recent_avgs <- read_csv("../output/Climate/Pops_FlintAvgs_Recent.csv") %>% 
  select(parent.pop:tmx_mean)
head(pops_flint_recent_avgs)

pops_flint_historic_avgs <- read_csv("../output/Climate/Pops_FlintAvgs_Historic.csv") %>% 
  select(parent.pop:tmx_mean)
head(pops_flint_historic_avgs)
names(pops_flint_historic_avgs)
```

### BioClim

```{r}
pops_bioclim_recent_avgs <- read_csv("../output/Climate/Pops_BioClimAvgs_Recent.csv") %>% 
  select(parent.pop:ppt_coldest_month_avg)
head(pops_bioclim_recent_avgs)

pops_bioclim_historical_avgs <- read_csv("../output/Climate/Pops_BioClimAvgs_Historical.csv")  %>% 
  select(parent.pop:ppt_coldest_month_avg)
head(pops_bioclim_historical_avgs)

names(pops_bioclim_historical_avgs)
```

## Davis Climate Data (Nov 2022-Oct 2023)

### From CIMIS

```{r}
davis_climate <- read_csv("../input/CIMS_Davis_monthly.csv")
head(davis_climate)

davis_climate_short <- davis_climate %>% 
  select(month_year=`Month Year`, refet=`Total ETo (mm)`, ppt=`Total Precip (mm)`,
         avg_sol_rad=`Avg Sol Rad (W/sq.m)`, avg_vapor_pres=`Avg Vap Pres (kPa)`,
         tmx=`Avg Max Air Temp (C)`, tmn=`Avg Min Air Temp (C)`, tavg=`Avg Air Temp (C)`,
         avg_max_rel_hum=`Avg Max Rel Hum (%)`, avg_min_rel_hum=`Avg Min Rel Hum (%)`,
         avg_rel_hum=`Avg Rel Hum (%)`, avg_dew_pt=`Avg Dew Point (C)`,
         avg_wind_spd=`Avg Wind Speed (m/s)`, avg_soil_temp=`Avg Soil Temp (C)`) %>% 
  separate(month_year, c("month", "year"), " ", remove = FALSE) %>% 
  mutate(month=match(month, month.abb),  t_diurnal = (tmx-tmn))
summary(davis_climate_short)

names(davis_climate_short)

davis_climate_short_flint <- davis_climate_short %>% #get the means for the variables to compare to home sites 
  summarise(ppt_mean_Davis=mean(ppt), tmn_mean_Davis=mean(tmn), tmx_mean_Davis=mean(tmx))
```

### BioClim

-   annual mean temperature (BIO1)

-   mean diurnal range (BIO2) - (Mean of monthly (max temp - min temp))

-   temperature seasonality (BIO4) (standard deviation \*100)

-   temperature annual range (BIO7) (Max Temperature of Warmest Month -
    Min Temperature of Coldest Month)

-   annual precipitation (BIO12) - sum of ppt for the entire year (not
    the avg)

-   precipitation seasonality (BIO15) (Coefficient of Variation)

-   mean temp of wettest quarter (BIO8) - CHANGED TO mean temp of
    wettest month

-   mean temp of driest quarter (BIO9) - CHANGED TO mean temp of driest
    month

-   precip of warmest quarter (BIO18) - CHANGED TO precip of warmest
    month

-   precip of coldest quarter (BIO19) - CHANGED TO precip of coldest
    month

Calculating wettest, driest, warmest, and coldest months

```{r}
davis_wettest_month <- davis_climate_short %>%  
  slice_max(ppt)

davis_driest_month <- davis_climate_short %>%  #there's more than one driest month
  slice_min(ppt)

davis_warmest_month <- davis_climate_short %>% 
  slice_max(tavg)

davis_coldest_month <- davis_climate_short %>%
  slice_min(tavg)
```

Bio 1, 2, 4, 7, 12, 15

```{r}
bioclim_davis_calc <- davis_climate_short %>% 
  summarise(ann_tmean=mean(tavg),  #Bio1 - Annual Mean Temperature
            mean_diurnal_range=mean(t_diurnal), #Bio2 - Mean Diurnal Range
            temp_seasonality=sd(tavg), #Bio4 - Temperature Seasonality
            temp_ann_range=(max(tmx))-(min(tmn)), #bio7 - temp annual range
            ann_ppt=sum(ppt), #bio12 - annual precip
            ppt_seasonality=cv(ppt+1)) #bio15 - Precipitation Seasonality (+1 to avoid strange CVs for areas where mean rainfaill is < 1)
bioclim_davis_calc
```

Bio 8(Q), 9(Q), 18(Q), 19(Q)

```{r}
#bio8 = tmean_wettest_month
bio8_davis <- davis_wettest_month %>% 
  dplyr::select(tmean_wettest_month=tavg)

#bio9 = tmean_driest_month
bio9_davis <- davis_driest_month %>% 
  summarise(tmean_driest_month=mean(tavg)) #taking the average b/c multiple driest months 

bio8_9_davis <- bind_cols(bio8_davis, bio9_davis)

#bio18 = ppt_warmest_month
bio18_davis <- davis_warmest_month %>% 
  dplyr::select(ppt_warmest_month=ppt)

#bio19 = ppt_coldest_month
bio19_davis <- davis_wettest_month %>% 
  dplyr::select(ppt_coldest_month=ppt)

bio18_19_davis <- bind_cols(bio18_davis, bio19_davis)

all_periods_davis <- bind_cols(bio8_9_davis, bio18_19_davis)
```

Merge all bioclims

```{r}
davis_bioclim_final <- bind_cols(bioclim_davis_calc, all_periods_davis) %>% 
  select_all(.funs = funs(paste0(., "_Davis")))
summary(davis_bioclim_final)
```

## Davis Climate Trends

```{r}
davis_climate_short$month_year <- factor(davis_climate_short$month_year, levels = c("Nov 2022","Dec 2022","Jan 2023", "Feb 2023", "Mar 2023", "Apr 2023", "May 2023", "Jun 2023", "Jul 2023", "Aug 2023", "Sep 2023", "Oct 2023"))
davis_climate_short %>% ggplot(aes(x=month_year, y=tavg)) + geom_point() + theme_classic()
davis_climate_short %>% ggplot(aes(x=month_year, y=ppt)) + geom_point() + theme_classic()
davis_climate_short %>% ggplot(aes(x=month_year, y=tmx)) + geom_point() + theme_classic()
davis_climate_short %>% ggplot(aes(x=month_year, y=avg_rel_hum)) + geom_point() + theme_classic()
davis_climate_short %>% ggplot(aes(x=month_year, y=avg_wind_spd)) + geom_point() + theme_classic()
davis_climate_short %>% ggplot(aes(x=month_year, y=avg_sol_rad)) + geom_point() + theme_classic()
```

## Gower's Climate Distance

(1/P) \* SUM ((absolute value(Ai - Bi)) / range(i)) for each variable

-   P = number of environmental variables = 13 (without CWD)

-   Ai = 30 year avg of that variable for the home site

-   Bi = Nov 2022-Oct 2023 avg of that variable for the Davis garden

-   Range(i) = maximum - minimum of that variable in the whole data set
    (across sites)

### Combine the flint and bioclim variables 

```{r}
davis_climate_all <- bind_cols(davis_climate_short_flint, davis_bioclim_final)
dim(davis_climate_all)

home_climates_all_recent <- left_join(pops_flint_recent_avgs, pops_bioclim_recent_avgs)
home_climates_all_historic <- left_join(pops_flint_historic_avgs, pops_bioclim_historical_avgs)

davis_home_climate_recent <- bind_cols(davis_climate_all, home_climates_all_recent)
summary(davis_home_climate_recent)

davis_home_climate_historic <- bind_cols(davis_climate_all, home_climates_all_historic)
summary(davis_home_climate_historic)

variables <- c("ppt_mean", "tmn_mean", "tmx_mean", "ann_tmean_avg",
                 "mean_diurnal_range_avg","temp_seasonality_avg","temp_ann_range_avg",
                 "tmean_wettest_month_avg", "tmean_driest_month_avg","ann_ppt_avg", 
                 "ppt_seasonality_avg", "ppt_warmest_month_avg", "ppt_coldest_month_avg")
```

### Figure out the range for each variable 

```{r}
davis_range_prep <- davis_climate_all %>% 
  mutate(parent.pop="Davis")

colnames(davis_range_prep) <- sub("_Davis", "_avg", colnames(davis_range_prep)) #remove Davis from column names
davis_range_prep <- davis_range_prep %>% rename(ppt_mean=ppt_mean_avg, tmn_mean=tmn_mean_avg, tmx_mean=tmx_mean_avg)

##Recent
range_merge_recent <- bind_rows(home_climates_all_recent, davis_range_prep)
names(range_merge_recent)
summary(range_merge_recent)

davis_home_climate_ranges_recent <- range_merge_recent %>% 
  summarise(ppt_mean_range=max(ppt_mean)-min(ppt_mean), 
            tmn_mean_range=max(tmn_mean)-min(tmn_mean), 
            tmx_mean_range=max(tmx_mean)-min(tmx_mean), 
            ann_tmean_avg_range=max(ann_tmean_avg)-min(ann_tmean_avg),
            mean_diurnal_range_avg_range=max(mean_diurnal_range_avg)-min(mean_diurnal_range_avg),
            temp_seasonality_avg_range=max(temp_seasonality_avg)-min(temp_seasonality_avg),
            temp_ann_range_avg_range=max(temp_ann_range_avg)-min(temp_ann_range_avg),
            tmean_wettest_month_avg_range=max(tmean_wettest_month_avg)-min(tmean_wettest_month_avg),
            tmean_driest_month_avg_range=max(tmean_driest_month_avg)-min(tmean_driest_month_avg),
            ann_ppt_avg_range=max(ann_ppt_avg)-min(ann_ppt_avg), 
            ppt_seasonality_avg_range=max(ppt_seasonality_avg)-min(ppt_seasonality_avg),
            ppt_warmest_month_avg_range=max(ppt_warmest_month_avg)-min(ppt_warmest_month_avg), 
            ppt_coldest_month_avg_range=max(ppt_coldest_month_avg)-min(ppt_coldest_month_avg))
davis_home_climate_ranges_recent

davis_home_climate_with_ranges_recent <- bind_cols(davis_home_climate_recent, davis_home_climate_ranges_recent)
names(davis_home_climate_with_ranges_recent)

##Historic
range_merge_historic <- bind_rows(home_climates_all_historic, davis_range_prep)
names(range_merge_historic)
summary(range_merge_historic)

davis_home_climate_ranges_historic <- range_merge_historic %>% 
  summarise(ppt_mean_range=max(ppt_mean)-min(ppt_mean), 
            tmn_mean_range=max(tmn_mean)-min(tmn_mean), 
            tmx_mean_range=max(tmx_mean)-min(tmx_mean), 
            ann_tmean_avg_range=max(ann_tmean_avg)-min(ann_tmean_avg),
            mean_diurnal_range_avg_range=max(mean_diurnal_range_avg)-min(mean_diurnal_range_avg),
            temp_seasonality_avg_range=max(temp_seasonality_avg)-min(temp_seasonality_avg),
            temp_ann_range_avg_range=max(temp_ann_range_avg)-min(temp_ann_range_avg),
            tmean_wettest_month_avg_range=max(tmean_wettest_month_avg)-min(tmean_wettest_month_avg),
            tmean_driest_month_avg_range=max(tmean_driest_month_avg)-min(tmean_driest_month_avg),
            ann_ppt_avg_range=max(ann_ppt_avg)-min(ann_ppt_avg), 
            ppt_seasonality_avg_range=max(ppt_seasonality_avg)-min(ppt_seasonality_avg),
            ppt_warmest_month_avg_range=max(ppt_warmest_month_avg)-min(ppt_warmest_month_avg), 
            ppt_coldest_month_avg_range=max(ppt_coldest_month_avg)-min(ppt_coldest_month_avg))
davis_home_climate_ranges_historic

davis_home_climate_with_ranges_historic <- bind_cols(davis_home_climate_historic, davis_home_climate_ranges_historic)
names(davis_home_climate_with_ranges_historic)
```

### Recent Gowers Calc
```{r}
gowers_calc_each_var_recent <- davis_home_climate_with_ranges_recent %>% 
  mutate(ppt_mean_gowers=abs(ppt_mean_Davis - ppt_mean) / ppt_mean_range,
         tmn_mean_gowers=abs(tmn_mean_Davis - tmn_mean) / tmn_mean_range,
         tmx_mean_gowers=abs(tmx_mean_Davis - tmx_mean) / tmx_mean_range,
         ann_tmean_gowers=abs(ann_tmean_Davis - ann_tmean_avg) / ann_tmean_avg_range,
         mean_diurnal_range_gowers=abs(mean_diurnal_range_Davis - mean_diurnal_range_avg) / mean_diurnal_range_avg_range,
         temp_seasonality_gowers=abs(temp_seasonality_Davis - temp_seasonality_avg) / temp_seasonality_avg_range,
         temp_ann_range_gowers=abs(temp_ann_range_Davis - temp_ann_range_avg) / temp_ann_range_avg_range,
         tmean_wettest_month_gowers=abs(tmean_wettest_month_Davis - tmean_wettest_month_avg) / tmean_wettest_month_avg_range,
         tmean_driest_month_gowers=abs(tmean_driest_month_Davis - tmean_driest_month_avg) / tmean_driest_month_avg_range,
         ann_ppt_gowers=abs(ann_ppt_Davis - ann_ppt_avg) / ann_ppt_avg_range,
         ppt_seasonality_gowers=abs(ppt_seasonality_Davis - ppt_seasonality_avg) / ppt_seasonality_avg_range,
         ppt_warmest_month_gowers=abs(ppt_warmest_month_Davis - ppt_warmest_month_avg) / ppt_warmest_month_avg_range,
         ppt_coldest_month_gowers=abs(ppt_coldest_month_Davis - ppt_coldest_month_avg) / ppt_coldest_month_avg_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))

davis_home_climate_with_ranges_recent
gowers_calc_each_var_recent
  
gowers_calc_per_pop_recent <- gowers_calc_each_var_recent %>% 
  mutate(Recent_Gowers_Dist=(1/13)*(ppt_mean_gowers + tmn_mean_gowers + tmx_mean_gowers +
                                ann_tmean_gowers + mean_diurnal_range_gowers +
                                temp_seasonality_gowers +temp_ann_range_gowers +
                                tmean_wettest_month_gowers +
                                tmean_driest_month_gowers +ann_ppt_gowers +
                                ppt_seasonality_gowers + ppt_warmest_month_gowers +
                                ppt_coldest_month_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Recent_Gowers_Dist)

gowers_calc_per_pop_recent
```

### Historic Gowers Calc
```{r}
gowers_calc_each_var_historic <- davis_home_climate_with_ranges_historic %>% 
  mutate(ppt_mean_gowers=abs(ppt_mean_Davis - ppt_mean) / ppt_mean_range,
         tmn_mean_gowers=abs(tmn_mean_Davis - tmn_mean) / tmn_mean_range,
         tmx_mean_gowers=abs(tmx_mean_Davis - tmx_mean) / tmx_mean_range,
         ann_tmean_gowers=abs(ann_tmean_Davis - ann_tmean_avg) / ann_tmean_avg_range,
         mean_diurnal_range_gowers=abs(mean_diurnal_range_Davis - mean_diurnal_range_avg) / mean_diurnal_range_avg_range,
         temp_seasonality_gowers=abs(temp_seasonality_Davis - temp_seasonality_avg) / temp_seasonality_avg_range,
         temp_ann_range_gowers=abs(temp_ann_range_Davis - temp_ann_range_avg) / temp_ann_range_avg_range,
         tmean_wettest_month_gowers=abs(tmean_wettest_month_Davis - tmean_wettest_month_avg) / tmean_wettest_month_avg_range,
         tmean_driest_month_gowers=abs(tmean_driest_month_Davis - tmean_driest_month_avg) / tmean_driest_month_avg_range,
         ann_ppt_gowers=abs(ann_ppt_Davis - ann_ppt_avg) / ann_ppt_avg_range,
         ppt_seasonality_gowers=abs(ppt_seasonality_Davis - ppt_seasonality_avg) / ppt_seasonality_avg_range,
         ppt_warmest_month_gowers=abs(ppt_warmest_month_Davis - ppt_warmest_month_avg) / ppt_warmest_month_avg_range,
         ppt_coldest_month_gowers=abs(ppt_coldest_month_Davis - ppt_coldest_month_avg) / ppt_coldest_month_avg_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))

  
gowers_calc_per_pop_historic <- gowers_calc_each_var_historic %>% 
  mutate(Historic_Gowers_Dist=(1/13)*(ppt_mean_gowers + tmn_mean_gowers + tmx_mean_gowers +
                                ann_tmean_gowers + mean_diurnal_range_gowers +
                                temp_seasonality_gowers +temp_ann_range_gowers +
                                tmean_wettest_month_gowers +
                                tmean_driest_month_gowers +ann_ppt_gowers +
                                ppt_seasonality_gowers + ppt_warmest_month_gowers +
                                ppt_coldest_month_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Historic_Gowers_Dist)

gowers_calc_per_pop_historic
```

### Merge recent and historic
```{r}
gowers_all_time <- full_join(gowers_calc_per_pop_recent, gowers_calc_per_pop_historic)
gowers_all_time
names(gowers_all_time)
write_csv(gowers_all_time, "../output/Climate/Pops_GowersEnvtalDist.csv")
```

Figures
```{r}
gowers_rec_fig <- gowers_all_time %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Recent_Gowers_Dist), y=Recent_Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
   labs(fill="Elevation (m)",x="Population", title="Recent Climate", y="Gowers Envtal Distance \n from Davis") +
  theme_classic() +
  scale_y_continuous(limits=c(0, 0.6)) +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Gowers_Recent.png", width = 12, height = 6, units = "in")

gowers_hist_fig <- gowers_all_time %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Historic_Gowers_Dist), y=Historic_Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population", title="Historic Climate", y="Gowers Envtal Distance \n from Davis") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Gowers_Historic.png", width = 12, height = 6, units = "in")

#should combine these into one figure and save that instead
legend <- get_legend(gowers_rec_fig)
gowers_hist_fig <- gowers_hist_fig + theme(legend.position="none")
gowers_rec_fig <- gowers_rec_fig + theme(legend.position="none")
grid.arrange(gowers_hist_fig, gowers_rec_fig, legend, ncol=3, widths=c(3.12, 3.12, 1.09))
 #2000 x 850
```


## Flint Climate Distance

### Subtraction

```{r}
recent_flint_dist_prep <- bind_cols(davis_climate_short_flint, pops_flint_recent_avgs)
names(recent_flint_dist_prep)

recent_flint_dist <- recent_flint_dist_prep %>% 
  mutate(ppt_mean_dist=ppt_mean_Davis - ppt_mean,
         tmn_mean_dist=tmn_mean_Davis - tmn_mean,
         tmx_mean_dist=tmx_mean_Davis - tmx_mean) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))

historic_flint_dist_prep <- bind_cols(davis_climate_short_flint, pops_flint_historic_avgs)
names(historic_flint_dist_prep)

historic_flint_dist <- historic_flint_dist_prep %>% 
  mutate(ppt_mean_dist=ppt_mean_Davis - ppt_mean,
         tmn_mean_dist=tmn_mean_Davis - tmn_mean,
         tmx_mean_dist=tmx_mean_Davis - tmx_mean) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
```

Figures Recent (subtraction distance)

```{r}
recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_mean_dist), y=ppt_mean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/MeanPPT_Dist_RecentClim.png", width = 12, height = 6, units = "in")

recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmn_mean_dist), y=tmn_mean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/MeanTMN_Dist_RecentClim.png", width = 12, height = 6, units = "in")

recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmx_mean_dist), y=tmx_mean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/MeanTMX_Dist_RecentClim.png", width = 12, height = 6, units = "in")
```

Figures Historical (subtraction distance)

```{r}
historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_mean_dist), y=ppt_mean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/MeanPPT_Dist_HistoricalClim.png", width = 12, height = 6, units = "in")

historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmn_mean_dist), y=tmn_mean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/MeanTMN_Dist_HistoricalClim.png", width = 12, height = 6, units = "in")

historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmx_mean_dist), y=tmx_mean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/MeanTMX_Dist_HistoricalClim.png", width = 12, height = 6, units = "in")
```

## Bioclim Climate Distance

### Subtraction

```{r}
#Recent
names(pops_bioclim_recent_avgs)
recent_bioclim_dist_prep <- bind_cols(davis_bioclim_final, pops_bioclim_recent_avgs)
recent_bioclim_dist <- recent_bioclim_dist_prep %>% 
  mutate(ann_tmean_dist=ann_tmean_Davis - ann_tmean_avg,
         mean_diurnal_range_dist=mean_diurnal_range_Davis - mean_diurnal_range_avg,
         temp_seasonality_dist=temp_seasonality_Davis - temp_seasonality_avg,
         temp_ann_range_dist=temp_ann_range_Davis - temp_ann_range_avg,
         tmean_wettest_month_dist=tmean_wettest_month_Davis - tmean_wettest_month_avg,
         tmean_driest_month_dist=tmean_driest_month_Davis - tmean_driest_month_avg,
         ann_ppt_dist=ann_ppt_Davis - ann_ppt_avg,
         ppt_seasonality_dist=ppt_seasonality_Davis - ppt_seasonality_avg, 
         ppt_warmest_month_dist=ppt_warmest_month_Davis - ppt_warmest_month_avg,
         ppt_coldest_month_dist=ppt_coldest_month_Davis - ppt_coldest_month_avg) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
recent_bioclim_dist

#Historical
names(pops_bioclim_historical_avgs)
historical_bioclim_dist_prep <- bind_cols(davis_bioclim_final, pops_bioclim_historical_avgs)
historical_bioclim_dist <- historical_bioclim_dist_prep %>% 
  mutate(ann_tmean_dist=ann_tmean_Davis - ann_tmean_avg,
         mean_diurnal_range_dist=mean_diurnal_range_Davis - mean_diurnal_range_avg,
         temp_seasonality_dist=temp_seasonality_Davis - temp_seasonality_avg,
         temp_ann_range_dist=temp_ann_range_Davis - temp_ann_range_avg,
         tmean_wettest_month_dist=tmean_wettest_month_Davis - tmean_wettest_month_avg,
         tmean_driest_month_dist=tmean_driest_month_Davis - tmean_driest_month_avg,
         ann_ppt_dist=ann_ppt_Davis - ann_ppt_avg,
         ppt_seasonality_dist=ppt_seasonality_Davis - ppt_seasonality_avg, 
         ppt_warmest_month_dist=ppt_warmest_month_Davis - ppt_warmest_month_avg,
         ppt_coldest_month_dist=ppt_coldest_month_Davis - ppt_coldest_month_avg) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
historical_bioclim_dist
```

Figures Recent (subtraction distance)

```{r}
recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_dist), y=ann_tmean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Ann_Tmean_Dist_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_dist), y=mean_diurnal_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Diurnal_Range_Dist_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_dist), y=temp_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Seasonality_Dist_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_dist), y=temp_ann_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Ann_Range_Dist_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_month_dist), y=tmean_wettest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Wet_Dist_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_month_dist), y=tmean_driest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Dry_Dist_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_dist), y=ann_ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Ann_PPT_Dist_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_dist), y=ppt_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/PPT_Seasonality_Dist_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_month_dist), y=ppt_warmest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/PPT_Warm_Dist_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_month_dist), y=ppt_coldest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/PPT_Cold_Dist_RecentClim.png", width = 12, height = 6, units = "in")
```

Historical (subtraction distance)

```{r}
historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_dist), y=ann_tmean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Ann_Tmean_Dist_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_dist), y=mean_diurnal_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Diurnal_Range_Dist_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_dist), y=temp_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Seasonality_Dist_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_dist), y=temp_ann_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Ann_Range_Dist_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_month_dist), y=tmean_wettest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Wet_Dist_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_month_dist), y=tmean_driest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Dry_Dist_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_dist), y=ann_ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Ann_PPT_Dist_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_dist), y=ppt_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/PPT_Seasonality_Dist_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_month_dist), y=ppt_warmest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/PPT_Warm_Dist_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_month_dist), y=ppt_coldest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/PPT_Cold_Dist_HistoricalClim.png", width = 12, height = 6, units = "in")
```
