---
title: "UCD_Growth_Climate"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# UCD Growth + Climate

Add climate to best base model

-   Weekly climate as predictor of K

-   Summary of climate as predictor of Hmax and maybe delta

## Libraries

```{r}
library(tidyverse)
library(brms)
library(zoo) #rollmean
```

## Load the growth data

```{r}
all_ucd_size_all <- read_csv("../output/UCD_Traits/UCD-2023_Size_Combined.csv")
```


## Pop Location Info

```{r}
ucd_gowers <- read_csv("../output/Climate/Gowers_UCD.csv") %>% 
  select(parent.pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) 
head(ucd_gowers)
```

Merge

```{r}
all_ucd_size_loc <- left_join(all_ucd_size_all, ucd_gowers) %>% 
  mutate(day=as.numeric(survey_date-min(survey_date)))
head(all_ucd_size_loc)
unique(all_ucd_size_loc$day)
```

## Data set for the model

```{r}
all_ucd_size_for_models <- all_ucd_size_loc %>% 
  filter(parent.pop != "WV", parent.pop!="LV1") %>% #consider also removing CP3 since no data after 1/27 
  filter(Genotype !="SQ3_4_3") %>% 
  filter(!is.na(height.cm))
```

## Read in soil temp

Transplants only in beds D, F, H, J, L - Remove Bed A from data set No
temperatures were taken from 2/23/23 to 3/17/23 and after 8/4

```{r}
temp_UCD <- read_csv("../input/UCD_Data/iButton_Data/UCD_iButton_Data_Corrected.csv") %>%
  filter(Bed != "A") %>% 
  mutate(Date_Time = mdy_hm(Date_Time))
head(temp_UCD)
skimr::skim(temp_UCD)
```

Average across beds

```{r}
temp_UCD_summary <- temp_UCD %>%
  group_by(Date_Time) %>% 
  summarise(AvgSoilTemp=mean(SoilTemp, na.rm=TRUE)) %>% #avg across beds 
  mutate(Date=as.Date(Date_Time)) %>%
  filter(Date != min(Date), Date != max(Date)) %>% # trim ragged ends
  group_by(Date) %>% #summarize hourly data by date
  summarize(
    min_temp_d = min(AvgSoilTemp),
    max_temp_d = max(AvgSoilTemp),
    mean_temp_d = mean(AvgSoilTemp)
  ) %>%
  mutate(
    across(ends_with("temp_d"), \(x) rollmean(x, k = 7, align = "right", fill = NA), .names="{.col}1_7"),
    across(ends_with("temp_d"), \(x) rollmean(x, k = 13, align = "right", fill = NA), .names="{.col}1_13") # 13 so I can get the first survey date
  ) 

temp_UCD_summary
```

## add climate data

```{r}
all_ucd_size_for_models_climate <- all_ucd_size_for_models %>% 
  left_join(temp_UCD_summary, by = c("survey_date" = "Date")) %>%
  filter(day>0) #remove day pre-field size b/c no climate data 

all_ucd_size_for_models_climate %>% 
  ggplot(aes(group=Genotype, x=survey_date, y=height.cm, col=elev_m)) + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_line() + facet_wrap(~parent.pop, scales = "free")
#without day 0, min is different...
```
## Priors

```{r}
prior1 <- c(set_prior("normal(5,3)", nlpar="Hmin"),
            set_prior("normal(60,15)",  nlpar="Hmax"),
            set_prior("gamma(2,2)", nlpar = "k", lb = 0), #gamma constrains to be > 0 
            set_prior("gamma(20,3)", nlpar = "delta", lb = 0)
)
            #first number = avg, second number = the space around that average to search 
```

## Base Models
```{r}
#best base model from non-climate data:
base_f1 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  Hmax + k + delta ~ (1|parent.pop) + (1|Genotype), 
                  nl=TRUE)

#alternate base model
base_f2 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin + Hmax + k + delta ~ (1|parent.pop) + (1|Genotype), 
                  nl=TRUE)
```

## Base Model Fits

### Base model 1 - Hmax, k, delta vary

```{r}
base_fit1 <- brm(formula=base_f1, data=all_ucd_size_for_models_climate, prior=prior1, iter = 3800, cores=4)

saveRDS(temp_fit1, file = "brm_object_UCD_base.rds")
summary(base_fit1)
plot(base_fit1, nvariables = 3, ask=FALSE)

pred.df <- all_ucd_size_for_models_climate %>% select(day, parent.pop, Genotype)
base_fit1.predictions <- cbind(pred.df, prediction=predict(base_fit1, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_ucd_size_for_models_climate, by=c("day", "parent.pop", "Genotype"))

base_fit1.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(group=Genotype,y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop, scales = "free")
ggsave("../output/UCD_GrowthClim_Predictions_Base1.png", width = 14, height = 8, units = "in")
```

### Base model 2 - Hmin varies too 

```{r}
base_fit2 <- brm(formula=climate_f1, data=all_ucd_size_for_models_climate, prior=prior1, iter = 3800, cores=4)
saveRDS(temp_fit1, file = "brm_object_UCD_base.rds")
summary(base_fit2)
plot(base_fit2, nvariables = 3, ask=FALSE)

pred.df <- all_ucd_size_for_models_climate %>% select(day, parent.pop, Genotype)
base_fit2.predictions <- cbind(pred.df, prediction=predict(base_fit2, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_ucd_size_for_models_climate, by=c("day", "parent.pop", "Genotype"))

base_fit2.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(group=Genotype, y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop, scales = "free")
ggsave("../output/UCD_GrowthClim_Predictions_Base2.png", width = 14, height = 8, units = "in")
```

### Model comparison 
```{r}
base_fit1 <- add_criterion(base_fit1, "loo")
base_fit2 <- add_criterion(base_fit2, "loo")
loo_compare(base_fit1, base_fit2)
```

```{r}
save.image("../output/BQC_UCD_Growth_Climate.Rdata")
```

## Temp Models

```{r, eval=FALSE}
climate_f2 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  k ~ min_temp_d1_7 + (1|parent.pop),
                  Hmax + delta ~ (1|parent.pop), 
                  nl=TRUE)

climate_f3 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  k ~  mean_temp_d1_7 + (1|parent.pop),
                  Hmax + delta ~ (1|parent.pop), 
                  nl=TRUE)

climate_f4 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  k ~ max_temp_d1_7 + (1|parent.pop), 
                  Hmax + delta ~ (1|parent.pop), 
                  nl=TRUE)

climate_f5 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  k ~ min_temp_d1_13 + (1|parent.pop), 
                  Hmax + delta ~ (1|parent.pop), 
                  nl=TRUE)

climate_f6 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  k ~ mean_temp_d1_13 + (1|parent.pop), 
                  Hmax + delta ~ (1|parent.pop), 
                  nl=TRUE)

climate_f7 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  k ~  max_temp_d1_13 + (1|parent.pop), 
                  Hmax + delta ~ (1|parent.pop), 
                  nl=TRUE)
```

## Temp Model Fits

### Min D1-7 (fit2)

```{r, eval=FALSE}
all_ucd_size_for_models_climate_fit2 <- all_ucd_size_for_models_climate %>% drop_na(min_temp_d1_7)

climate_fit2 <- brm(formula=climate_f2, data=all_ucd_size_for_models_climate_fit2, prior=prior1, iter = 3800, cores=4)

summary(climate_fit2) 
plot(climate_fit2, nvariables = 3, ask=FALSE)  
```

### Mean D1-7 (fit3)

```{r, eval=FALSE}
all_ucd_size_for_models_climate_fit3 <- all_ucd_size_for_models_climate %>% drop_na(mean_temp_d1_7)

climate_fit3 <- brm(formula=climate_f3, data=all_ucd_size_for_models_climate_fit3, prior=prior1, iter = 3800, cores=4)

summary(climate_fit3)
plot(climate_fit3, nvariables = 3, ask=FALSE)
```

### Max D1-7 (fit4)

```{r, eval=FALSE}
all_ucd_size_for_models_climate_fit4 <- all_ucd_size_for_models_climate %>% drop_na(max_temp_d1_7)

climate_fit4 <- brm(formula=climate_f4, data=all_ucd_size_for_models_climate_fit4, prior=prior1, iter = 3800, cores=4)

summary(climate_fit4)
plot(climate_fit4, nvariables = 3, ask=FALSE)
```

### Min D1-13 (fit5)

```{r, eval=FALSE}
all_ucd_size_for_models_climate_fit5 <- all_ucd_size_for_models_climate %>% drop_na(min_temp_d1_13)

climate_fit5 <- brm(formula=climate_f5, data=all_ucd_size_for_models_climate_fit5, prior=prior1, iter = 3800, cores=4)

summary(climate_fit5)
plot(climate_fit5, nvariables = 3, ask=FALSE)
```

### Mean D1-13 (fit6)

```{r, eval=FALSE}
all_ucd_size_for_models_climate_fit6 <- all_ucd_size_for_models_climate %>% drop_na(meam_temp_d1_13)

climate_fit6 <- brm(formula=climate_f6, data=all_ucd_size_for_models_climate_fit6, prior=prior1, iter = 3800, cores=4)

summary(climate_fit6)
plot(climate_fit6, nvariables = 3, ask=FALSE)
```

### Max D1-13 (fit7)

```{r, eval=FALSE}
all_ucd_size_for_models_climate_fit7 <- all_ucd_size_for_models_climate %>% drop_na(max_temp_d1_13)

climate_fit7 <- brm(formula=climate_f7, data=all_ucd_size_for_models_climate_fit7, prior=prior1, iter = 3800, cores=4)

summary(climate_fit7)
plot(climate_fit7, nvariables = 3, ask=FALSE)
```

## Model Comparison

```{r, eval=FALSE}
base_fit1 <- add_criterion(base_fit1, "loo")
base_fit2 <- add_criterion(base_fit2, "loo")
climate_fit2 <- add_criterion(climate_fit2, "loo")
climate_fit3 <- add_criterion(climate_fit3, "loo")
climate_fit4 <- add_criterion(climate_fit4, "loo")
climate_fit5 <- add_criterion(climate_fit5, "loo") 
climate_fit6 <- add_criterion(climate_fit6, "loo") 
climate_fit7 <- add_criterion(climate_fit7, "loo") 
loo_compare(base_fit1, base_fit2, climate_fit2, climate_fit3, climate_fit4, climate_fit5, climate_fit6, climate_fit7)
```
