---
title: "UCD_SurvAnalysis"
output: 
  html_document: 
    keep_md: true
date: "2024-10-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Survival Analysis 

See tutorial: https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html

## Libraries
```{r}
# install.packages(c("ggsurvfit", "gtsummary"))
library(tidyverse) #includes lubridate, which we need 
library(ggsurvfit)
library(gtsummary)
library(survival)
```

## Read in the data
Need to come back to this to check the data for issues!!
```{r}
ucd_surv <- read_csv("../input/UCD_Data/CorrectedCSVs/UCD_transplants_pheno_mort_20231016_corrected.csv",
                     na = c("", "NA", "-", "N/A")) %>% #note this only goes to 10/20 need to come back and change this to include dates after this
  select(block:rep, death.date=`Death Date`, Notes) %>% 
  rename(parent.pop=pop) %>% 
  filter(rep != 100) %>% #get rid of individuals that germinated in the field 
  filter(!is.na(parent.pop)) %>% 
   unite(BedLoc, block:col, sep="_", remove = FALSE) %>% 
  unite(Genotype, parent.pop:rep, sep="_", remove = FALSE) %>% 
  filter(!str_detect(Genotype, ".*buff*")) %>% 
  unite(pop.mf, parent.pop:mf, sep="_", remove = FALSE)
head(ucd_surv)
unique(ucd_surv$parent.pop) #there's an SQ with no number - need to check on this 

ucd_surv %>% rowwise() %>%  #checking if mf and rep can be converted to numeric 
  filter(!is.na(rep)) %>%  
  filter(is.na(as.numeric(rep)))
```

As weeks
```{r}
ucd_surv_dates <- ucd_surv %>% 
  mutate(mf=as.double(mf), rep=as.double(rep)) %>% 
  mutate(planting.date="11/30/22", #could try to make this more specific to when certain blocks were planted 
         last_fup_date=if_else(is.na(death.date), "10/16/23", death.date)) %>%  #need this to calculate survival times
  mutate(planting.date=mdy(planting.date), last_fup_date=mdy(last_fup_date)) %>% #convert to date objects
  mutate(os_weeks=as.duration(planting.date %--% last_fup_date) / dweeks(1), #observed number of weeks
         status=if_else(is.na(death.date), 0, 1)) %>% #0=censured (alive in this case), 1=dead
  filter(os_weeks > 0) %>% #there is one case of a plant that was dead at planting, so just removed it since this is survival post- transplanting
  select(-Notes)
head(ucd_surv_dates)
unique(ucd_surv_dates$os_weeks)
```

## Location Info
Note: the below files include climate distance which is not finalized yet (as of 10/3)
```{r}
gowersdist_UCD <- read_csv("../output/Climate/Pops_GowersEnvtalDist_UCD.csv") %>% 
  rename(Recent_Gowers_Dist_UCD = Recent_Gowers_Dist, Historic_Gowers_Dist_UCD = Historic_Gowers_Dist)
head(gowersdist_UCD)
gowersdist_WL2 <- read_csv("../output/Climate/Pops_GowersEnvtalDist_WL2.csv") %>% 
  rename(Recent_Gowers_Dist_WL2 = Recent_Gowers_Dist, Historic_Gowers_Dist_WL2 = Historic_Gowers_Dist)
head(gowersdist_WL2)

gowersdist_all <- full_join(gowersdist_UCD, gowersdist_WL2)
head(gowersdist_all)
```

```{r}
ucd_surv_dates_loc <- left_join(ucd_surv_dates, gowersdist_all)
```

## Create Survival Objects and Curves
```{r}
Surv(ucd_surv_dates_loc$os_weeks, ucd_surv_dates_loc$status)[1:10] #show the first 10 observations of the survival object
#There will be one entry for each subject that is the survival time, which is followed by a + if the subject was censored.
```

"The survfit() function creates survival curves using the Kaplan-Meier method based on a formula. Let’s generate the overall survival curve for the entire cohort, assign it to object s1, and look at the structure using str():"
```{r}
s1 <- survfit(Surv(os_weeks, status) ~ 1, data = ucd_surv_dates_loc)
str(s1)
#time: the timepoints at which the curve has a step, i.e. at least one event occurred
#surv: the estimate of survival at the corresponding time
```

## Survival Plot
```{r}
survfit2(Surv(os_weeks, status) ~ 1, data = ucd_surv_dates_loc) %>% 
  ggsurvfit() +
  labs(
    x = "Weeks",
    y = "Overall survival probability"
  ) + 
  add_confidence_interval() +
  add_risktable() #at risk = plants still alive, event=plants dead
```

## Estimating x-week survival
```{r}
summary(survfit(Surv(os_weeks, status) ~ 1, data = ucd_surv_dates_loc), times = 8) #survival for 2 months
#2 month survival probability is 84%
```

## Estimating median survival time
```{r}
survfit(Surv(os_weeks, status) ~ 1, data = ucd_surv_dates_loc) 
#median survival time is 13.3 weeks 
```

## Compare Survival between Populations 
Plot survival differences
```{r}
survfit2(Surv(os_weeks, status) ~ elevation.group, data = ucd_surv_dates_loc) %>% 
  ggsurvfit() +
  labs(
    x = "Weeks",
    y = "Overall survival probability") + 
  add_confidence_interval() 

survfit(Surv(os_weeks, status) ~ parent.pop, data = ucd_surv_dates_loc) #median survival lengths for each pop
```

"We can conduct between-group significance tests using a log-rank test. The log-rank test equally weights observations over the entire follow-up time and is the most common way to compare survival times between groups."
```{r}
survdiff(Surv(os_weeks, status) ~ parent.pop, data = ucd_surv_dates_loc)
#significant difference in overall survival according to population
```

## Cox regression model
"We may want to quantify an effect size for a single variable, or include more than one variable into a regression model to account for the effects of multiple variables.
The Cox regression model is a semi-parametric model that can be used to fit univariable and multivariable regression models that have survival outcomes.:
```{r}
coxph(Surv(os_weeks, status) ~ parent.pop, data = ucd_surv_dates_loc) %>% 
  tbl_regression(exp = TRUE) 

# hazard ratio (HR) represents the ratio of hazards between two groups at any particular point in time. 
#A HR < 1 indicates reduced hazard of death whereas a HR > 1 indicates an increased hazard of death.
```
