---
title: "Annual Census Analysis"
author: "Brandie Quarles"
date: "2023-09-14"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  markdown: 
    wrap: 72
---

To Do:

-   Add climate to the models!

Original Design - 23 pops, 7 mfs, 13 reps

-   Note: this analysis is for transplant beds only

## Questions:

-   Is there more variation b/t populations than b/t mfs?
-   Are certain traits related to fitness?
-   Did low elevation populations perform better than high elevation
    populations?
-   Connection to climate of home site?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Relevant Libraries and Functions

```{r}
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
library(lmerTest) #for mixed effect models
conflicted::conflicts_prefer(lmerTest::lmer)
library(broom.mixed) #tidy method for lmerTest and glmer
library(corrplot) #plotting correlations 
library(rstatix) #performing cor_test
library(naniar) #replaces values with NA
sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")

#For scree plots 
#library("devtools") #The package devtools is required for the installation as factoextra is hosted on github.
#1install_github("kassambara/factoextra")
library("factoextra")
```

## Load the pop and location data

```{r}
#pop info
pops_common_garden <- read_csv("../input/UCD_Data/Pops_for_2022_UCD.csv") #pops included in common garden 
summary(pops_common_garden)
pops_common_garden_nonotes <- pops_common_garden %>% select(parent.pop:elevation.group, UCD.seed.year)
pops_common_garden_nonotes$elevation.group <- str_to_title(pops_common_garden_nonotes$elevation.group)

#extra location info 
pop_loc <- read_csv("../input/Strep_tort_locs.csv")

unique(pop_loc$`Site code`)
unique(pops_common_garden_nonotes$parent.pop)
#need to change YOSE to YO
pop_loc_yo <- pop_loc %>% mutate(parent.pop = str_replace(`Site code`, "YOSE(\\d+)", "YO\\1")) %>% select(Lat, Long, elev_m=`Elevation (m)`, parent.pop)
unique(pop_loc_yo$parent.pop)

#merge in location info
pop_elev <- left_join(pops_common_garden_nonotes, pop_loc_yo)
head(pop_elev)
```

## Load the annual census data

```{r}
#Note: BQC edited the csvs prior to importing to R b/c the first row of the transplants sheets was not the heading row (veg, repo info and not block, mf etc...) 
anncensustrans.0522 <- read_csv("../input/UCD_Data/CorrectedCSVs/Annual_Census_Transplants_All_May_2023_corrected.csv")
names(anncensustrans.0522) <- gsub(" ", "", colnames(anncensustrans.0522)) 
anncensustrans.0522 %>% rowwise() %>% filter(!is.na(mf)) %>% filter(is.na(as.numeric(mf))) #mf = X when pop = buffer
anncensustrans.0522 %>% rowwise() %>% filter(!is.na(rep)) %>% filter(is.na(as.numeric(rep)))
anncensustrans.0522 %>% filter(pheno != "X" & is.na(diam_mm)) #no living plants without a diameter
anncensustrans.0522_alive <- anncensustrans.0522 %>% filter(!is.na(diam_mm)) %>% mutate_at(c("mf", "rep"), as.double) #used diameter instead of pheno to filter b/c some dying plants were measured 
summary(anncensustrans.0522_alive)

anncensustrans.0626 <- read_csv("../input/UCD_Data/CorrectedCSVs/Annual_Census_transplants_20230626_corrected.csv")
names(anncensustrans.0626) <- gsub(" ", "", colnames(anncensustrans.0626))
anncensustrans.0626 %>% filter(pheno == "X" & !is.na(diam_mm)) #no dead plants that have size info
anncensustrans.0626_alive <- anncensustrans.0626 %>% filter(pheno != "X")
summary(anncensustrans.0626_alive) #56 plants alive 


#quick summary info
levels(as.factor(anncensustrans.0522_alive$pop)) #out of 23 pops, only 10 alive still (BH, CP2, DPR, IH, SC, SQ3, TM2, WL1, WL2, YO7)
levels(as.factor(anncensustrans.0626_alive$pop)) #only 9 pops alive (no DPR)

#B = budding
#F = floweirng
#P = post flowering (has fruits)
#S = senescing (applies to TM2 only)
#V = vegetative 
xtabs(~pheno+pop, data = anncensustrans.0522_alive) 
xtabs(~pheno+pop, data = anncensustrans.0626_alive) #vegetative decreases for BH & CP2 --> may have reproductive data for some new individuals 
```

## Prep for Merge

```{r}
anncensustrans_0522_alive_tomerge <- anncensustrans.0522_alive %>% select(block:col, parent.pop=pop, mf:rep, pheno.0522=pheno, diam.0522=diam_mm, height.0522=height_cm, total_branch.0522=total_branch, herb_dam.0522=herb_dam, wilt_status.0522=wilt_status, longest_leaf.0522=longest_leaf_cm, flowers.0522=flowers, fruits.0522=fruits, longest_fruit.0522=longest_fruit_cm, repro_branch.0522=repro_branch)
names(anncensustrans_0522_alive_tomerge)

anncensustrans_0626_alive_tomerge <- anncensustrans.0626_alive %>% select(block:col, parent.pop=pop, mf:rep, pheno.0626=pheno, diam.0626=diam_mm, height.0626=height_cm, total_branch.0626=total_branch, herb_dam.0626=herb_dam, wilt_status.0626=wilt_status, longest_leaf.0626=longest_leaf_cm, flowers.0626=flowers, fruits.0626=fruits, longest_fruit.0626=longest_fruit_cm, repro_branch.0626=repro_branch)
names(anncensustrans_0626_alive_tomerge)
```

## Merge dataframes

-   Combine 0522 and 0626
    -   if an individual is sampled twice use the maximum value - add in
        pop elevation info

```{r}
anncensustrans_all_alive <- full_join(anncensustrans_0522_alive_tomerge, anncensustrans_0626_alive_tomerge)
dim(anncensustrans_all_alive) #64 rows , 28 cols

#convert to long 
anncensustrans_all_alive_long <- reshape(anncensustrans_all_alive, direction = "long", sep = ".",
                                      idvar=c("block", "row", "col", "parent.pop", "mf", "rep"),
                                      varying =c("pheno.0522", "diam.0522", "height.0522", "total_branch.0522", "herb_dam.0522", "wilt_status.0522", "longest_leaf.0522", "flowers.0522", "fruits.0522", "longest_fruit.0522", "repro_branch.0522", "pheno.0626", "diam.0626", "height.0626", "total_branch.0626", "herb_dam.0626", "wilt_status.0626", "longest_leaf.0626", "flowers.0626", "fruits.0626", "longest_fruit.0626", "repro_branch.0626"))
dim(anncensustrans_all_alive_long)

anncensustrans_all_alive_max <- anncensustrans_all_alive_long %>% 
  group_by(block, row, col, parent.pop, mf, rep) %>% 
  summarise_if(is.double, max, na.rm=TRUE) %>% 
  select(-time)
head(anncensustrans_all_alive_max) #some cases where max of different variables comes from different dates 
summary(anncensustrans_all_alive_max) #getting -Inf for longest_leaf and repro_branch b/c N/A at both times
#anncensustrans_all_alive_max %>% filter(longest_leaf==-Inf)
#anncensustrans_all_alive_long %>% filter(block=="D1", row==4)
#head(anncensustrans_all_alive_max)

#add in pop elevation info
anncensustrans_all_alive_max_elev <- left_join(anncensustrans_all_alive_max, pop_elev)
names(anncensustrans_all_alive_max_elev)

anncensustrans_all_alive_max_elev %>% group_by(parent.pop, mf) %>% 
  filter(!is.na(diam)) %>% 
  summarise(n=n()) #low replication at maternal family level 

anncensustrans_all_alive_max_elev_noinf <- anncensustrans_all_alive_max_elev %>% 
  mutate_if(is.double, list(~na_if(., -Inf))) %>%  #replace -INf with NAs
  filter(parent.pop != "WL1") %>%                  #take out pops with only 1 individual alive 
  filter(parent.pop != "SQ3") %>% 
  mutate(longest_fruit = na_if(longest_fruit, 0)) %>%  #change 0 branches to NA
  filter(rep != 100) #get rid of individuals that germinated in the field 
summary(anncensustrans_all_alive_max_elev_noinf)
```

## Summary plots

-   diam
-   height
-   total_branch
-   longest_leaf
-   flowers
-   fruits
-   longest_fruit
-   repro_branch

```{r}
anncensustrans_all_alive_max_elev_noinf <- as_tibble(anncensustrans_all_alive_max_elev_noinf) #convert to tibble so that the map function can use it 

#Histograms
anncensustrans_all_alive_max_elev_noinf %>% 
  select(diam, height, total_branch, longest_leaf, flowers, fruits, longest_fruit, repro_branch) %>% 
  names() %>% 
  map(~ggplot(anncensustrans_all_alive_max_elev_noinf, aes_string(x = .)) + 
        geom_histogram())

#qq plots 
anncensustrans_all_alive_max_elev_noinf %>% 
  select(diam, height, total_branch, longest_leaf, flowers, fruits, longest_fruit, repro_branch) %>% 
  names() %>% 
  map(~ggplot(anncensustrans_all_alive_max_elev_noinf, aes_string(sample = .)) + 
       stat_qq() +
        stat_qq_line())


# diam close to normal 
# height close to normal 
# total_branch heavily right skewed 
# longest_leaf close to normal 
# flowers heavily right skewed (0 inflated)
# fruits heavily right skewed (0 inflated)
# longest_fruit roughly normal, but low replication --> no stats
# repro_branch right skewed 

#Transformations:
#anncensustrans_all_alive_max_elev_noinf %>% mutate(stbr = sqrt(total_branch)) %>% ggplot(aes(x=stbr)) + geom_histogram() #log, 1og10, and sqrt transformation didn't work 

#anncensustrans_all_alive_max_elev_noinf %>% mutate(lrbr = log10(repro_branch)) %>% ggplot(aes(x=lrbr)) + geom_histogram() #log, 1og10, and sqrt transformation didn't work 

#Box Plots 
anncensustrans_all_alive_max_elev_noinf %>% ggplot(aes(x=parent.pop, y=diam)) + geom_boxplot()
anncensustrans_all_alive_max_elev_noinf %>% ggplot(aes(x=parent.pop, y=height)) + geom_boxplot()
anncensustrans_all_alive_max_elev_noinf %>% ggplot(aes(x=parent.pop, y=total_branch)) + geom_boxplot()
anncensustrans_all_alive_max_elev_noinf %>% ggplot(aes(x=parent.pop, y=longest_leaf)) + geom_boxplot()
anncensustrans_all_alive_max_elev_noinf %>% ggplot(aes(x=parent.pop, y=flowers)) + geom_boxplot()
anncensustrans_all_alive_max_elev_noinf %>% ggplot(aes(x=parent.pop, y=fruits)) + geom_boxplot()
anncensustrans_all_alive_max_elev_noinf %>% ggplot(aes(x=parent.pop, y=longest_fruit)) + geom_boxplot()
anncensustrans_all_alive_max_elev_noinf %>% ggplot(aes(x=parent.pop, y=repro_branch)) + geom_boxplot()
#variation across populations for diam, height, and longest leaf (maybe total_branch & repro_branch, not # flowers, prob not # fruits (all low except TM2), not longest_fruit)
```

## Relationships between traits

-   Plot branching pattern with other size metrics for idea of shape

```{r}
anncensustrans_all_alive_max_elev_noinf %>% ggplot(aes(x=total_branch, y=repro_branch)) +
  geom_point()  #repro_branches is so 0 skewed it's hard to tell if there's a relationship 

anncensustrans_all_alive_max_elev_noinf %>% ggplot(aes(x=total_branch, y=diam)) +
  geom_point() +
  geom_smooth()

anncensustrans_all_alive_max_elev_noinf %>% ggplot(aes(x=total_branch, y=height)) +
  geom_point() +
  geom_smooth()

anncensustrans_all_alive_max_elev_noinf %>% ggplot(aes(x=total_branch, y=longest_leaf)) +
  geom_point() +
  geom_smooth()
```

No clear relationships between branching and other size traits.

What about diameter and height?

```{r}
anncensustrans_all_alive_max_elev_noinf %>% ggplot(aes(x=diam, y=height)) +
  geom_point() +
  geom_smooth()
```

There seems to be a positive relationship between height and diameter.

What about diameter and longest leaf?

```{r}
anncensustrans_all_alive_max_elev_noinf %>% ggplot(aes(x=diam, y=longest_leaf)) +
  geom_point() +
  geom_smooth()
```

There seems to be a positive relationship between height and length of
the longest leaf.

What about height and longest leaf?

```{r}
anncensustrans_all_alive_max_elev_noinf %>% ggplot(aes(x=height, y=longest_leaf)) +
  geom_point() +
  geom_smooth()
```

Positive relationship between height and longest leaf seems to plateau.

## Relationships between Climate and Traits (fix this after figuring out climate data)

```{r, eval=FALSE}
names(anncensustrans_all_alive_max_elev)
ggplot(anncensustrans_all_alive_max_elev, aes(x = tmx, y = diam))+
  geom_point()

ggplot(anncensustrans_all_alive_max_elev, aes(x = ppt, y = diam))+
  geom_point()


ggplot(anncensustrans_all_alive_max_elev, aes(x = pck, y = diam))+
  geom_point()

ggplot(anncensustrans_all_alive_max_elev, aes(x = tmn, y = height))+
  geom_point()
ggplot(anncensustrans_all_alive_max_elev, aes(x = tmn, y = longest_leaf))+
  geom_point()
```

## Calculating means by pop (and elev)

```{r}
anncensustrans_all_alive_summary <- anncensustrans_all_alive_max_elev_noinf %>% 
  group_by(parent.pop, elevation.group, elev_m) %>% 
  summarise(meanDiam=mean(diam, na.rm = TRUE), sem_diam=sem(diam, na.rm=TRUE), 
            meanHeight = mean(height,na.rm=(TRUE)), sem_height=sem(height, na.rm=(TRUE)), 
            meanLongestLeaf=mean(longest_leaf, na.rm=(TRUE)), sem_leaf=sem(longest_leaf, na.rm=TRUE), 
            meanTotBr=mean(total_branch, na.rm = TRUE), sem_totbr=sem(total_branch, na.rm=TRUE), 
            meanRepBr=mean(repro_branch, na.rm=TRUE), semRepBr=sem(repro_branch, na.rm=TRUE), 
            meanFruits=mean(fruits, na.rm = TRUE), semFruits=sem(fruits, na.rm=TRUE), 
            meanLongFruit=mean(longest_fruit, na.rm = TRUE), semLongFruit=sem(longest_fruit, na.rm = TRUE), 
            meanFl=mean(flowers, na.rm=TRUE), semFl=sem(flowers, na.rm=TRUE))
anncensustrans_all_alive_summary 
#write_csv(anncensustrans_all_alive_summary, file ="../output/UCD_Traits/anncensus_traitavgs.csv")
anncensustrans_all_alive_max_elev_noinf %>% filter(parent.pop=="DPR") %>% select(flowers) #one DPR plant had 1 flower and the other had 116 (massive variation)

anncensustrans_all_alive_max_elev_logregs <- anncensustrans_all_alive_max_elev_noinf %>% 
  mutate(bTotBr=ifelse(total_branch==1, 0, 1), 
         bRepBr=ifelse(repro_branch==0, 0, 1), 
         bFruits=ifelse(fruits==0, 0, 1), 
         bFlowers=ifelse(flowers==0, 0, 1)) 

anncensustrans_all_alive_summary_logregs <- anncensustrans_all_alive_max_elev_logregs %>% 
  group_by(parent.pop, elevation.group, elev_m) %>% 
  summarise(probTotBr = mean(bTotBr, na.rm = TRUE), sem_totbr=sem(bTotBr, na.rm = TRUE),
            probRepBr=mean(bRepBr, na.rm = TRUE), semRepBr=sem(bRepBr,na.rm = TRUE), 
            probFruits=mean(bFruits, na.rm = TRUE), semFruits=sem(bFruits, na.rm = TRUE), 
            probFlowers=mean(bFlowers, na.rm = TRUE), semFl=sem(bFlowers, na.rm = TRUE))
anncensustrans_all_alive_summary_logregs 
#write_csv(anncensustrans_all_alive_summary_logregs, file ="../output/UCD_Traits/anncensus_traitprobs.csv")
```

## Mixed Effects Models (and figures of trait means)

### Diam

```{r}
anncensustrans_all_alive_summary %>% ggplot(aes(x=fct_reorder(parent.pop, meanDiam), y=meanDiam, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanDiam-sem_diam,ymax=meanDiam+sem_diam),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  #scale_fill_manual(values=elev_three_palette) + 
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Diam (mm)", x="Population") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/UCD_Traits/AvgDiam.png", width = 8, height = 6, units = "in")

lmediam2 = lmerTest::lmer(diam ~ (1|parent.pop/mf), data=anncensustrans_all_alive_max_elev_noinf)
summary(lmediam2)
ranova(lmediam2) #no maternal family effect b/c maybe not enough individuals 
#could try assigning unique names and "un-nesting" them 
VarCorr(lmediam2) %>% 
  as.data.frame() %>% 
  mutate(var.pct = vcov/sum(vcov)*100)

#EMM_diam=emmeans(lmediam2,pairwise~parent.pop)
#EMM_diam

#COME BACK TO THE BELOW WHEN CLIMATE DATA FIGURED OUT 
#lmediam3 = lmer(diam ~ ppt + (1|parent.pop/mf), #data=anncensustrans_all_alive_max_elev_noinf)
#summary(lmediam3)
#ranova(lmediam3) 
##singular warning b/c we only have one value for each pop #(multiple maternal families have one value)
#
#lmediam4 = lmer(diam ~ ppt + (1|parent.pop), #data=anncensustrans_all_alive_max_elev_noinf)
#summary(lmediam4)
#ranova(lmediam4) 
```

### Total_branch

-   Tried code for poisson regression or logistic regression with mixed
    effects

    -   Had to use glmer for this

        -   Used family = poisson for traits with poisson distribution

            -   Couldn't use ranova so used a loglikihood test to test
                for the significance of the random effects

        -   Use family = binomial(logit) for logit link

            -   Tried family = binomial for Prob_Total_Branch (1 or
                greater than 1). Couldn't include random effects so
                parent.pop is in the model as a fixed effect.

```{r}
anncensustrans_all_alive_summary %>% ggplot(aes(x=fct_reorder(parent.pop, meanTotBr), y=meanTotBr, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanTotBr-sem_totbr,ymax=meanTotBr+sem_totbr),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  #scale_fill_manual(values=elev_three_palette) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", x="Population", y="Avg # Total Branches") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) #large standard error bars 
#ggsave("../output/UCD_Traits/AvgTotBsBr.png", width = 8, height = 6, units = "in")

#Poisson regression 
lmetotal_branch <- glmer(total_branch ~ (1|parent.pop/mf), data = anncensustrans_all_alive_max_elev_noinf, family = poisson(link = "log")) 
summary(lmetotal_branch)
VarCorr(lmetotal_branch) %>% 
  as.data.frame() %>% 
  mutate(var.pct = vcov/sum(vcov)*100)
glance(lmetotal_branch)
#ranova(lmetotal_branch) doesn't work for lme4
#to test for significance of random effect: https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#testing-significance-of-random-effects
#the most common way to do this is to use a likelihood ratio test, i.e. fit the full and reduced models (the reduced model is the model with the focal variance(s) set to zero). 
m0 <- glmer(total_branch ~ (1|parent.pop), data = anncensustrans_all_alive_max_elev_noinf, family = poisson(link = "log")) 
m00 <- lm(total_branch~1, anncensustrans_all_alive_max_elev_noinf)
anova(lmetotal_branch,m0, m00) #model with both random effects has a higher likelihood (better fit)


anncensustrans_all_alive_summary_logregs %>% ggplot(aes(x=fct_reorder(parent.pop, probTotBr), y=probTotBr, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=probTotBr-sem_totbr,ymax=probTotBr+sem_totbr),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0.01)) +
  #scale_fill_manual(values=elev_three_palette) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", x="Population", y="Prob. of Total Branches > 1") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) 
#ggsave("../output/UCD_Traits/ProbTotBsBr.png", width = 8, height = 6, units = "in")

totbr_logreg=glm(bTotBr ~ parent.pop, data=anncensustrans_all_alive_max_elev_logregs, family = binomial())
summary(totbr_logreg) #this is not the correct model b/c it treats parent pop as a fixed effect and doesn't take into account maternal families 
```

### Repro_branch

```{r}
anncensustrans_all_alive_summary %>% ggplot(aes(x=fct_reorder(parent.pop, meanRepBr), y=meanRepBr, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanRepBr-semRepBr,ymax=meanRepBr+semRepBr),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0.01, 0)) +
  #scale_fill_manual(values=elev_three_palette) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", x="Population", y="Avg # Reproductive Branches") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) #mostly the same except 1 SQ3 plant with a lot
#ggsave("../output/UCD_Traits/AvgRepBsBr.png", width = 8, height = 6, units = "in")

#poisson regression 
lmerepro_branch <- glmer(repro_branch ~ (1|parent.pop/mf), data = anncensustrans_all_alive_max_elev_noinf, family = poisson(link = "log")) 
summary(lmerepro_branch)
VarCorr(lmerepro_branch) %>% 
  as.data.frame() %>% 
  mutate(var.pct = vcov/sum(vcov)*100)
glance(lmerepro_branch)
m0 <- glmer(repro_branch ~ (1|parent.pop), data = anncensustrans_all_alive_max_elev_noinf, family = poisson(link = "log")) 
m00 <- lm(repro_branch~1, anncensustrans_all_alive_max_elev_noinf)
anova(lmerepro_branch,m0, m00) #model with both random effects has a higher likelihood (better fit)


anncensustrans_all_alive_summary_logregs %>% ggplot(aes(x=fct_reorder(parent.pop, probRepBr), y=probRepBr, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=probRepBr-semRepBr,ymax=probRepBr+semRepBr),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0.01, 0)) +
  #scale_fill_manual(values=elev_three_palette) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", x="Population", y="Prob. Repro. Branches > 0") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/UCD_Traits/ProbRepBsBr.png", width = 8, height = 6, units = "in")
```

### Fruits

```{r}
anncensustrans_all_alive_summary %>% ggplot(aes(x=fct_reorder(parent.pop, meanFruits), y=meanFruits, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0.01, 0)) +
  #scale_fill_manual(values=elev_three_palette) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", x="Population", y="Avg # Fruits") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) #mostly the same (low), except TM2 had way more fruits than the other populations 
#ggsave("../output/UCD_Traits/AvgFruits.png", width = 8, height = 6, units = "in")

#poisson regression 
lmefruits <- glmer(fruits ~ (1|parent.pop/mf), data = anncensustrans_all_alive_max_elev_noinf, family = poisson(link = "log")) 
summary(lmefruits)
VarCorr(lmefruits) %>% 
  as.data.frame() %>% 
  mutate(var.pct = vcov/sum(vcov)*100)
glance(lmefruits)
m0 <- glmer(fruits ~ (1|parent.pop), data = anncensustrans_all_alive_max_elev_noinf, family = poisson(link = "log")) 
m00 <- lm(fruits~1, anncensustrans_all_alive_max_elev_noinf)
anova(lmefruits,m0, m00) #model without random effects has a higher likelihood

anncensustrans_all_alive_summary_logregs %>% ggplot(aes(x=fct_reorder(parent.pop, probFruits), y=probFruits, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=probFruits-semFruits,ymax=probFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0.01, 0)) +
  #scale_fill_manual(values=elev_three_palette) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", x="Population", y="Prob. of Fruits") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) 
#ggsave("../output/UCD_Traits/ProbFruits.png", width = 8, height = 6, units = "in")
```

### Longest_fruit

```{r}
#only 2 populations with multiple fruits to calculate a mean from 
anncensustrans_all_alive_summary %>% ggplot(aes(x=parent.pop, y=meanLongFruit, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanLongFruit-semLongFruit,ymax=meanLongFruit+semLongFruit),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0.01, 0)) +
  #scale_fill_manual(values=elev_three_palette) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", x="Population", y="Avg Longest Fruit (cm)") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) #mostly the same (low), except TM2 had way more fruits than the other populations 
#ggsave("../output/UCD_Traits/AvgLongestFruits.png", width = 8, height = 6, units = "in")

#stats not necessary 
#lmelongest_fruit = lmer(longest_fruit ~ (1|parent.pop/mf), data=anncensustrans_all_alive_max_elev_noinf)
#summary(lmelongest_fruit)
#ranova(lmelongest_fruit)
#VarCorr(lmelongest_fruit) %>% 
#  as.data.frame() %>% 
#  mutate(var.pct = vcov/sum(vcov)*100)
```

### Flowers

```{r}
anncensustrans_all_alive_summary %>% ggplot(aes(x=fct_reorder(parent.pop, meanFl), y=meanFl, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFl-semFl,ymax=meanFl+semFl),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0.01, 0)) +
  #scale_fill_manual(values=elev_three_palette) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", x="Population", y="Avg # Flowers") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) #mostly the same (low), except TM2 had way more fruits than the other populations 
#ggsave("../output/UCD_Traits/AvgFlowers.png", width = 8, height = 6, units = "in")

#poisson regression 
lmeflowers <- glmer(flowers ~ (1|parent.pop/mf), data = anncensustrans_all_alive_max_elev_noinf, family = poisson(link = "log")) 
summary(lmeflowers)
VarCorr(lmeflowers) %>% 
  as.data.frame() %>% 
  mutate(var.pct = vcov/sum(vcov)*100)
glance(lmeflowers)
m0 <- glmer(flowers ~ (1|parent.pop), data = anncensustrans_all_alive_max_elev_noinf, family = poisson(link = "log")) 
m00 <- lm(flowers~1, anncensustrans_all_alive_max_elev_noinf)
anova(lmeflowers,m0, m00) #model without random effects has a higher likelihood

anncensustrans_all_alive_summary_logregs %>% ggplot(aes(x=fct_reorder(parent.pop, probFlowers), y=probFlowers, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=probFlowers-semFl,ymax=probFlowers+semFl),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0.01, 0)) +
  #scale_fill_manual(values=elev_three_palette) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", x="Population", y="Prob. of Flowers") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) 
#ggsave("../output/UCD_Traits/ProbFlowers.png", width = 8, height = 6, units = "in")
```

### Height

```{r}
anncensustrans_all_alive_summary %>% ggplot(aes(x=fct_reorder(parent.pop, meanHeight), y=meanHeight, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanHeight-sem_height,ymax=meanHeight+sem_height),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  #scale_fill_manual(values=elev_three_palette) + 
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation", x="Population") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/UCD_Traits/AvgHeight_AnnCensus.png", width = 8, height = 6, units = "in")

lmeheight = lmer(height ~ (1|parent.pop/mf), data=anncensustrans_all_alive_max_elev_noinf)
summary(lmeheight)
ranova(lmeheight)
```

### Longest_leaf

```{r}
anncensustrans_all_alive_summary %>% ggplot(aes(x=parent.pop, y=meanLongestLeaf, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanLongestLeaf-sem_leaf,ymax=meanLongestLeaf+sem_leaf),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
   #scale_fill_manual(values=elev_three_palette) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/UCD_Traits/AvgLongLeaf_AnnCensus.png", width = 8, height = 6, units = "in")

lmelongest_leaf = lmer(longest_leaf ~ (1|parent.pop/mf), data=anncensustrans_all_alive_max_elev_noinf)
summary(lmelongest_leaf)
ranova(lmelongest_leaf)
```
