---
title: "UCD_Growth"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---

# Analyze growth

-   Get growth curve code from Julin

-   Is the growth rate lower for high elev b/c drier and biennial?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries 
```{r}
library(tidyverse)
library(brms)
```

## Need to make a csv file with all survey data for each plant 

Get list of size files 
```{r}
pretrans_ucd <-  dir("../input/UCD_Data/CorrectedCSVs/",
             pattern=".*20221128.*csv",
             full.names=TRUE)
pretrans_ucd

posttrans_ucd <- dir("../input/UCD_Data/CorrectedCSVs/",
             pattern="Size_survey_transplants.*corrected.*csv",
             full.names=TRUE)
posttrans_ucd
```

Add data to a tibble
```{r}
pretrans_dat_ucd <- tibble(path=pretrans_ucd, filename=basename(path))
pretrans_dat_ucd
posttrans_dat_ucd <- tibble(path=posttrans_ucd, filename=basename(path))
posttrans_dat_ucd

all_ucd_size <- bind_rows(pretrans_dat_ucd, posttrans_dat_ucd) %>% 
  mutate(survey_date=str_extract(filename, "2[0-9]*"),
         survey_date=lubridate::ymd(survey_date))
all_ucd_size
```

Read in the files
```{r}
all_ucd_size2 <- all_ucd_size %>%
  mutate(sheets=map(path, 
                    read_csv, 
                    #col_types=cols(.default=col_character()), # forces all columns to be character
                                                              # this is needed because of data sheet problems that cause some numerics to be imported as character  
                    na = c("", "NA", "-", "N/A") # sets NA strings. 
                    )) %>%
  select(-path)

all_ucd_size2
```


Check for data issues
```{r, eval=FALSE}
map(all_ucd_size2$sheets, colnames) %>% #get unique column names 
  unlist() %>%
  unique() %>%
  sort()
#size columns sometimes uppercase for initial size census 
#pop is "parent.pop" for 1128 and 1213 

map(all_ucd_size2$sheets, head,10) #look at first 10 lines of each sheet 
#1213 read in height and longest leaf as characters - check why 
#0210 read in height and longest leaf as characters - check why 

all_ucd_size2$sheets[[1]] <- all_ucd_size2$sheets[[1]] %>% rename(pop=parent.pop, `height (cm)`=`Height (cm)`, `longest leaf (cm)`=`Longest leaf (cm)`) 
all_ucd_size2$sheets[[2]] <- all_ucd_size2$sheets[[2]] %>% rename(pop=parent.pop) 

all_ucd_size2$sheets[[2]] %>% rowwise() %>% filter(!is.na(`height (cm)`)) %>%  filter(is.na(as.numeric(`height (cm)`))) #character issue was caused by "N/A" so I added that to the previous code chunk so R could recognize it as NA
all_ucd_size2$sheets[[4]] %>% rowwise() %>% filter(!is.na(`height (cm)`)) %>%  filter(is.na(as.numeric(`height (cm)`))) #character issue due to a ) in height column - fixed this in the original csv file 
all_ucd_size2$sheets[[5]] %>% rowwise() %>% filter(!is.na(mf)) %>%  filter(is.na(as.numeric(mf))) #"." for mf on a buffer row 
all_ucd_size2$sheets[[5]] <- all_ucd_size2$sheets[[5]] %>% mutate(mf=as.double(mf)) #convert mf to numeric
all_ucd_size2$sheets[[1]] <- all_ucd_size2$sheets[[1]] %>% filter(`germinated?`== 1) #keep germs only from pretrans size 
#ALL ISSUES FIXED!
```

Combine the data
```{r}
all_ucd_size_all <- all_ucd_size2 %>% unnest(sheets) %>%
  rename(germ = `germinated?`, height.cm = `height (cm)`, 
         long.leaf.cm = `longest leaf (cm)`, parent.pop=pop) %>% 
  filter(parent.pop != "buffer") %>% 
  mutate(parent.pop= str_replace(parent.pop, ".*VTR.*", "LVTR1")) %>% 
  filter(rep != 100) %>% #get rid of individuals that germinated in the field 
  unite(Genotype, parent.pop:rep, sep="_", remove = FALSE) %>% 
  unite(pop.mf, parent.pop:mf, sep="_", remove = FALSE) %>% 
  select(survey_date, block, Genotype:Notes)
head(all_ucd_size_all)
unique(all_ucd_size_all$parent.pop) #all 23 pops!

#write_csv(all_ucd_size_all, "../output/UCD_Traits/UCD_Size_Combined.csv")
```

## Pop Location Info
```{r}
gowersdist_UCD <- read_csv("../output/Climate/Gowers_UCD.csv")

gowersdist_UCD_wider <- gowersdist_UCD %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, GrwSsn_FLINT_GD, GrwSsn_BIOCLIM_GD, Wtr_Year_GD, Wtr_Year_FLINT_GD, Wtr_Year_BIOCLIM_GD))
head(gowersdist_UCD_wider)
```

Merge
```{r}
all_ucd_size_loc <- left_join(all_ucd_size_all, gowersdist_UCD_wider) %>% 
  mutate(day=as.numeric(survey_date-min(survey_date)))
head(all_ucd_size_loc)
```

## Check Sample Sizes
```{r}
all_ucd_size_loc %>% 
  filter(!is.na(height.cm)) %>% 
  group_by(parent.pop, day) %>% 
  summarise(Sample_Size=n()) %>% 
  filter(day==0) %>% 
  arrange(Sample_Size)

#WV only has 2 individuals so going to remove it 
#LV1 only has 3 individuals, still keep it?

all_ucd_size_for_models <- all_ucd_size_loc %>% 
  filter(parent.pop != "WV") %>% 
  filter(!is.na(height.cm))

#Sam said that you want 10x the number of parameters (4) you are estimating. So should I remove everything with a saple size of less than ~35?
```

## Quick figure

Height
```{r}
all_ucd_size_all$parent.pop = factor(all_ucd_size_all$parent.pop, levels=c('CC','TM2','SC','IH','BH','WV','FR','DPR','WR','WL1','SQ1','SQ2','WL2','YO4','CP2','CP3','LV3','SQ3','YO7','YO8','LV1','LVTR1','YO11'))
  
all_ucd_size_loc %>% 
  #filter(parent.pop=="LV3") %>% 
  ggplot(aes(group=Genotype, x=survey_date, y=height.cm, col=elev_m)) + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_line() + facet_wrap(~parent.pop, scales = "free")
#ggsave("../output/UCD_Traits/Growth_AllPops.png", width = 14, height = 8, units = "in")
```
Longest Leaf
```{r}
all_ucd_size_loc %>% 
  #filter(parent.pop=="SQ3") %>% 
  ggplot(aes(group=Genotype, x=survey_date, y=long.leaf.cm, col=elev_m)) + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_line() + facet_wrap(~parent.pop, scales = "free")

```

### Check weird cases
```{r}
all_ucd_size_all %>% filter(parent.pop=="WV") %>% arrange(Genotype) #no issues 
all_ucd_size_all %>% filter(parent.pop=="LV3") %>% filter(Genotype=="LV3_6_3") #grew and then immediately started to shrink 

all_ucd_size_all %>% filter(parent.pop=="CP3") %>% arrange(Genotype) #Weird drop then increase in height for CP3_12_1
all_ucd_size_all %>% filter(parent.pop=="SQ3") %>% filter(Genotype=="SQ3_4_3") #SOMETHING WEIRD HAPPENED WITH THIS PLANT 

```

## Model growth with the Weibull model 

### Base Model
```{r}
f1 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta), #Weibull formula, divide k by 100 to get it on a more reasonable scale
    Hmax + Hmin + k + delta ~ 1, # how each parameter should be fit with respect to the predictors 
    #we want a single parameter value (Intercept) for each of the four parameters
   nl=TRUE) #tells brms that this is a non-linear formula
```

### Set Priors
```{r}
prior1 <- c(set_prior("normal(5,3)", nlpar="Hmin"),
            set_prior("normal(60,15)",  nlpar="Hmax"),
            set_prior("gamma(2,2)", nlpar = "k"), #gamma constrains to be > 0 
            set_prior("gamma(20,3)", nlpar = "delta")
)

#Hmax = alpha -  Sets the asymptote (max growth)
#Hmin = beta - Changes the y intercept (the lower limit)
#k - Often described as the growth rate --\> related to the slope of the line
#delta - Sets the inflection point - when you shift b/t concave to convex
```

### Fit the model
```{r}
fit1 <- brm(formula=f1, data=all_ucd_size_for_models, prior=prior1, cores = 4)
```

Summary
```{r}
summary(fit1)
#Rhat and ESS are fine 
```

Plots
```{r}
plot(fit1, nvariables = 3, ask=FALSE)
#posterior distributions look fine 
```

### Predictions
```{r}
fit1.prediction <- cbind(all_ucd_size_for_models, pred1=predict(fit1)[,"Estimate"])

fit1.prediction %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm), alpha=.1) +
  geom_line(aes(y=pred1), color="blue") +
  facet_wrap(. ~parent.pop)

#Predictions are clearly off, esp for TM2, good for BH 
```

## Let Hmax vary 
```{r}
random_f2 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin + k + delta ~ 1,
                  Hmax ~ (1|parent.pop), #different hmax for each pop + random effect of pop
                  nl=TRUE)
```

Default Priors
```{r}
get_prior(random_f2, data = all_ucd_size_for_models)
#there's now a prior for sd of Hmax due to pop
#don't need a specific prior for each pop now 

prior2 <- c(set_prior("normal(5,3)", nlpar="Hmin"),
            set_prior("normal(60,15)",  nlpar="Hmax"),
            set_prior("gamma(2,2)", nlpar = "k"), #gamma constrains to be > 0 
            set_prior("gamma(20,3)", nlpar = "delta")
) #same as prior 1 for now 
```

### Fit the model
```{r}
random_fit2 <- brm(formula=random_f2, data=all_ucd_size_for_models, prior=prior2, cores = 4, iter = 3000) #increase iterations b/c more complex (helps with Rhat and ESS)
```

Summary
```{r}
summary(random_fit2)
#rhat and ESS look good
```

```{r}
plot(random_fit2, nvariables = 3, ask=FALSE)
```

### Predictions
The function expand_grid creates a data frame with all combinations of the items it is given (days and pops in this case).
```{r}
pred.df <- expand_grid(day=min(all_ucd_size_for_models$day):max(all_ucd_size_for_models$day),
                       parent.pop=unique(all_ucd_size_for_models$parent.pop))
pred.df
```

```{r}
random_fit2.predictions <- cbind(pred.df, prediction=predict(random_fit2, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_ucd_size_for_models, by=c("day", "parent.pop"))

random_fit2.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  #scale_color_brewer(type="qual", palette = "Set2", guide="none") +
  facet_wrap(~parent.pop)

random_fit2.predictions %>% filter(parent.pop=="BH") %>% 
  ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)

#not really capturing the curve well for any parent.pop
```

## Vary slope (k) and asymptote (Hmax) by pop
```{r}
random_f3 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin + delta ~ 1,
                  Hmax ~ (1|parent.pop), #different hmax for each pop 
                  #could also do Hmax + k  ~pop
                  k ~ (1|parent.pop), #different k for each pop
                  nl=TRUE)
```

### Priors
```{r}
get_prior(random_f3, data=all_ucd_size_for_models)

prior3 <- c(set_prior("normal(5,3)", nlpar="Hmin"),
            set_prior("normal(60,15)",  nlpar="Hmax"),
            set_prior("normal(0,1)", nlpar = "k"), #gamma constrains to be > 0 
            set_prior("gamma(20,3)", nlpar = "delta")
)
```

### Fit the model
```{r}
random_fit3 <- brm(formula=random_f3, data=all_ucd_size_for_models, prior=prior3, cores = 4, iter = 4000)
#Warning: There were 1299 divergent transitions after warmup. See
#https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup
#to find out why this is a problem and how to eliminate them.Warning: There were 181 transitions after warmup that exceeded #the maximum treedepth. Increase max_treedepth above 10. See
#https://mc-stan.org/misc/warnings.html#maximum-treedepth-exceededWarning: Examine the pairs() plot to diagnose sampling #problems
```

Summary
```{r}
summary(random_fit3)
#Rhat is high, but ESS is low for sd(k_Intercept) and k_Intercept
#Warning: There were 2350 divergent transitions after warmup. Increasing adapt_delta above 0.8 may help.
```

Plots
```{r}
plot(random_fit3, N = 3, ask=FALSE)
#some weirdness with the chains 
```

### Pairs 
```{r}
pairs(random_fit3, variable = "TM2", regex = TRUE) 
#look at correlation b/t hmax and k for each pop 
```

Predictions 
```{r}
random_fit3.predictions <- cbind(pred.df, prediction=predict(random_fit3, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_ucd_size_for_models, by=c("day", "parent.pop"))

random_fit3.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)
#looks better even though the fit diagnostics aren't perfect
```

## Let delta and hmax vary 
```{r}
random_f4 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin + k ~ 1,
                  Hmax ~ (1|parent.pop), #different hmax for each pop #could also do Hmax + k  ~pop
                  delta ~ (1|parent.pop), #different hmax for each pop
                  nl=TRUE)
```

```{r}
prior4 <- c(set_prior("normal(5,3)", nlpar="Hmin"),
            set_prior("normal(60,15)",  nlpar="Hmax"),
            set_prior("gamma(2,2)", nlpar = "k"), #gamma constrains to be > 0 
            set_prior("gamma(20,3)", nlpar = "delta")
)
```

```{r}
random_fit4 <- brm(formula=random_f4, data=all_ucd_size_for_models, prior=prior4, cores = 4, iter = 4000)
```

```{r}
summary(random_fit4)
```

```{r}
plot(random_fit4, N = 3, ask=FALSE)
```

Predictions 
```{r}
random_fit4.predictions <- cbind(pred.df, prediction=predict(random_fit4, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_ucd_size_for_models, by=c("day", "parent.pop"))

random_fit4.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)
```


## Let delta, k, and hmax vary
```{r}
random_f5 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin ~ 1,
                  Hmax ~ (1|parent.pop), #different hmax for each pop #could also do Hmax + k  ~pop
                  k ~ (1|parent.pop), #different hmax for each pop
                  delta ~ (1|parent.pop),
                  nl=TRUE)
```

```{r}
prior5 <- c(set_prior("normal(5,3)", nlpar="Hmin"),
            set_prior("normal(60,15)",  nlpar="Hmax"),
            set_prior("gamma(2,2)", nlpar = "k"), #gamma constrains to be > 0 
            set_prior("gamma(20,3)", nlpar = "delta")
)
            #first number = avg, second number = the space around that average to search 
```

```{r}
random_fit5 <- brm(formula=random_f5, data=all_ucd_size_for_models, prior=prior5, cores = 4, iter = 4000)
```

```{r}
summary(random_fit5)
```

```{r}
plot(random_fit5, N = 3, ask=FALSE)

pairs(random_fit5, variable = "DPR", regex = TRUE) #Want the correlations to be more cloud like (no obvious correlations among parameters)
```

Predictions 
```{r}
random_fit5.predictions <- cbind(pred.df, prediction=predict(random_fit5, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_ucd_size_for_models, by=c("day", "parent.pop"))

random_fit5.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)
```


## Model Comparison 
```{r}
fit1 <- add_criterion(fit1, "loo")
random_fit2 <- add_criterion(random_fit2, "loo")
random_fit3 <- add_criterion(random_fit3, "loo")
random_fit4 <- add_criterion(random_fit4, "loo")
random_fit5 <- add_criterion(random_fit5, "loo")
loo_compare(fit1, random_fit2, random_fit3, random_fit4, random_fit5)
```