---
title: "Annual Census Analysis"
author: "Brandie Quarles"
date: "2023-09-14"
output: html_document:
  keep_md: yes
---

##Note: this is for transplant beds only 

Original Design
- 23 pops, 7 mfs, 13 reps 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Relevant Libraries and Functions
```{r}
library(tidyverse)
library(corrplot) #plotting correlations 
library(rstatix) #performing cor_test
library(lmerTest)
library(naniar) #replaces values with NA
sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")

#For scree plots 
#library("devtools") #The package devtools is required for the installation as factoextra is hosted on github.
#1install_github("kassambara/factoextra")
library("factoextra")
```

#Load the data 
```{r}
#Note: BQC edited the csvs prior to importing to R b/c of issues with the naming of columns and the first row of the transplants sheets not being the heading row (veg, repo info and not block, mf etc...) 
pop_elev <- read_csv("../input/UCD Data/Pops_for_2022_UCD.csv") #elevation info for each population 
summary(pop_elev)
pop_elev_nonotes <- pop_elev %>% select(-notes, -JGI.DNA)
pop_elev_nonotes
pop_elev_nonotes$elevation.group <- str_to_title(pop_elev_nonotes$elevation.group)

anncensustrans.0522 <- read_csv("../input/UCD Data/Annual_Census_Transplants_All_May_2023_corrected.csv")
head(anncensustrans.0522)
names(anncensustrans.0522) <- gsub(" ", "", colnames(anncensustrans.0522))
summary(anncensustrans.0522)
anncensustrans.0522 %>% filter(diam_mm=="X") #747 rows, X means the plant is dead 
anncensustrans.0522 %>% filter(is.na(diam_mm)) #53 rows 
anncensustrans.0522_numcorrected <- anncensustrans.0522 %>% mutate_at(c("diam_mm", "height_cm", "total_branch", "longest_leaf_cm", "flowers", "fruits", "longest_fruit_cm", "repro_branch"), as.double)
summary(anncensustrans.0522_numcorrected) #correct number of NAs
anncensustrans.0522_numcorrected %>% filter(pheno !="X" & is.na(diam_mm)) #all alive plants have a diam
anncensustrans.0522_alive <- anncensustrans.0522_numcorrected %>% filter(!is.na(diam_mm))
summary(anncensustrans.0522_alive) #64 plants alive - includes early measured plants (marked as dead)

anncensustrans.0626 <- read_csv("../input/UCD Data/Annual_Census_transplants_20230626_corrected.csv")
head(anncensustrans.0626)
names(anncensustrans.0626) <- gsub(" ", "", colnames(anncensustrans.0626))
anncensustrans.0626 %>% filter(diam_mm=="X") 
summary(anncensustrans.0626)
anncensustrans.0626$mf <- as.character(anncensustrans.0626$mf) #convert mf to character 
anncensustrans.0626$rep <- as.character(anncensustrans.0626$rep) #convert rep to character 
anncensustrans.0626 %>% filter(pheno == "X", !is.na(diam_mm)) #no dead plants that have size info
anncensustrans.0626_alive <- anncensustrans.0626 %>% filter(pheno != "X")
summary(anncensustrans.0626_alive) #56 plants alive 


#prep for merge 
levels(as.factor(anncensustrans.0522_alive$pop)) #out of 23 pops, only 10 alive still (BH, CP2, DPR, IH, SC, SQ3, TM2, WL1, WL2, YO7)
levels(as.factor(anncensustrans.0626_alive$pop)) #only 9 pops alive (no DPR)

#B = budding
#F = floweirng
#P = post flowering (has fruits)
#S = senescing (applies to TM2 only)
#V = vegetative 
xtabs(~pheno+pop, data = anncensustrans.0522_alive) 
xtabs(~pheno+pop, data = anncensustrans.0626_alive) #vegetative decreases for BH & CP2 --> may have reproductive data for some new individuals 


names(anncensustrans.0522_alive)
names(anncensustrans.0626_alive)
names(pop_elev_nonotes)

anncensustrans_0522_alive_tomerge <- anncensustrans.0522_alive %>% select(block:col, parent.pop=pop, mf:rep, pheno.0522=pheno, diam.0522=diam_mm, height.0522=height_cm, total_branch.0522=total_branch, herb_dam.0522=herb_dam, wilt_status.0522=wilt_status, longest_leaf.0522=longest_leaf_cm, flowers.0522=flowers, fruits.0522=fruits, longest_fruit.0522=longest_fruit_cm, repro_branch.0522=repro_branch)
names(anncensustrans_0522_alive_tomerge)

anncensustrans_0626_alive_tomerge <- anncensustrans.0626_alive %>% select(block:col, parent.pop=pop, mf:rep, pheno.0626=pheno, diam.0626=diam_mm, height.0626=height_cm, total_branch.0626=total_branch, herb_dam.0626=herb_dam, wilt_status.0626=wilt_status, longest_leaf.0626=longest_leaf_cm, flowers.0626=flowers, fruits.0626=fruits, longest_fruit.0626=longest_fruit_cm, repro_branch.0626=repro_branch)
names(anncensustrans_0626_alive_tomerge)
```

#Climate Data (1895-2022)
- aet (mm) = actual evapotranspiration (water transpired by plants, if water not limited)
- cwd (mm) = climate water deficit (estimate of drought)
- pck (mm) = snowpack (prior month pck + snow - lost snow)
- pet (mm) = potential evapotranspiration (total amount of water that can evaporate from the ground or be trasnpired by plants)
- ppt (mm) = precipitation
- rch (mm) = recharge (water that penetrates below the root zone)
- run (mm) = runoff (water that becomes streamflow)
- str (mm) = soil water storage (avg amount of water stored in the soil annually)
all above are totals/sums per month 
- tmn (deg C) = min air temp (for a given month)
- tmx (deg C) = max air temp (for a given month)
127 years

*Notes:
- temperature highly correlated within pops w/in years - could just take avg temp for the 3 months when we think germination is occurring (Oct-Jan & April-June)
--try having separate columns for both and see how they load on the PCA or just pick the months based on elevation - base times on AJB paper 


To do:
- do some lit review to see what variables other people have focused on or found to be significant 
-Sam said CWD or snow-related variables - if all correlated could pick the one that makes teh most sense
-Could also consider PCR (principal components regression)
```{r}
climate_data <- read_csv("../input/Dimensions_All_1895-2022.csv")
head(climate_data)
names(climate_data)
unique(climate_data$pop)
climate_data$year = as.character(climate_data$year)
head(climate_data)

climate_data_yo <- climate_data %>% mutate(pop = str_replace(pop, "YOSE(\\d+)", "YO\\1")) #changing YOSE to YO 
#`(\\d+)`: This part contains a regular expression within parentheses: 
#`\\d`: Matches a digit (equivalent to `[0-9]`). 
#The double backslash `\\` is used to escape the `\` #character in R strings, so `\\d` is interpreted as `\d` in the regex, denoting a digit.
#`+`: This qualifier means "one or more of the preceding element," so `\\d+` will match one or more digits.
#`()`: Parentheses are used to define a capturing group, which means that the part of the regex within the parentheses `(\\d+)` is captured to be used in back-references or for extracting matched parts.
#\\1 references the first capturing group
unique(climate_data_yo$pop)

#names(pop_elev_nonotes)
pop_elev_climate <- left_join(pop_elev_nonotes, climate_data_yo, by=c("parent.pop"="pop")) %>% select(parent.pop, elevation.group, year:tmx)
unique(pop_elev_climate$parent.pop)
#unique(pop_elev_nonotes$parent.pop)
head(pop_elev_climate, 30)
names(pop_elev_climate)
```

#Climate traits across pops
```{r}
pop_elev_climate_yearlytot <- pop_elev_climate %>% group_by(parent.pop, year, elevation.group) %>% summarise_at(c("pck", "ppt"), sum, na.rm = TRUE)
names(pop_elev_climate_yearlytot)
pop_elev_climate_yearlytot %>% filter(year>1992) %>% ggplot(aes(x=year, y=pck, group=parent.pop, color=elevation.group)) + geom_point() + geom_line() + scale_colour_manual(values=elev_three_palette)
pop_elev_climate_yearlytot %>% filter(year>1992) %>% ggplot(aes(x=year, y=ppt, group=parent.pop, color=elevation.group)) + geom_point() + geom_line() + scale_colour_manual(values=elev_three_palette)

pop_elev_climate_avgs <- pop_elev_climate %>% filter(year>1992) %>% group_by(parent.pop, elevation.group) %>% summarise_at(c("aet", "cwd", "pck", "pet", "ppt", "rch", "run", "str", "tmn", "tmx"), mean, na.rm = TRUE)
pop_elev_climate_avgs #30 year averages of all climate variables 
pop_elev_climate_avgs$elevation.group <- factor(pop_elev_climate_avgs$elevation.group, levels=elev_order)

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, aet), y=aet, fill=elevation.group)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, cwd), y=cwd, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, pck), y=pck, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, pet), y=pet, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, ppt), y=ppt, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, rch), y=rch, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, run), y=run, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, str), y=str, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1)) #this could be a better metric of available water (compared to ppt)
#TM2 may be higher b/c of clay in the area that was being mapped (off the outcrops)

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmn), y=tmn, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

pop_elev_climate_avgs %>% ggplot(aes(x=fct_reorder(parent.pop, tmx), y=tmx, fill=elevation.group)) + 
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))
```

#Climate trait correlations
```{r}
names(pop_elev_climate)
pop_elev_climate %>% cor_test(aet, cwd, pck, pet, ppt, rch, run, str, tmn, tmx, method="pearson")

cor.mat <- pop_elev_climate %>% select(aet, cwd, pck, pet, ppt, rch, run, str, tmn, tmx) %>% cor_mat()
cor.mat
	
mydata.cor = pop_elev_climate %>% select(aet, cwd, pck, pet, ppt, rch, run, str, tmn, tmx) %>% cor()
mydata.cor
corrplot(mydata.cor)

#PCA
#normalize the data
climate_normalized <- pop_elev_climate %>% select(aet, cwd, pck, pet, ppt, rch, run, str, tmn, tmx) %>% scale() #normalize the data so they're all on the same scale
head(climate_normalized)
cor.norm = cor(climate_normalized) #test correlations among the traits
corrplot(cor.norm)

climate.pca <- princomp(cor.norm) 
summary(climate.pca) #PC1 explains 80% of the variation, PC2 explains 7.9%, together they explain 88%
climate.pca$loadings[, 1:2] #PC1 has high positive loading values for tmn, tmx, pet, cwd, aet
fviz_eig(climate.pca) #scree plot 
fviz_pca_var(climate.pca, col.var = "black") #biplot

#The goal of the third visualization is to determine how much each variable is represented in a given component. Such a quality of representation is called the Cos2 and corresponds to the square cosine, and it is computed using the fviz_cos2 function. (https://www.datacamp.com/tutorial/pca-analysis-r)
#A low value means that the variable is not perfectly represented by that component. 
#A high value, on the other hand, means a good representation of the variable on that component.
fviz_cos2(climate.pca, choice = "var", axes = 1:2) #axes says which PCs to use 
#biplot combined wiht cos2
fviz_pca_var(climate.pca, col.var = "cos2", 
            gradient.cols = c("black", "orange", "green"),
            repel = TRUE)

#Should use this code to check the significance of the PCA: https://github.com/StatQuest/pca_demo/blob/master/pca_demo.R
#Remember this paper: Björklund, M. 2019. Be careful with your principal components. Evolution 73: 2151–2158.

```


#Merge dataframes
- combine 0522 and 0626
- if an individual is sampled twice use the maximum value 
- add in pop elevation info 
```{r}
anncensustrans_all_alive <- full_join(anncensustrans_0522_alive_tomerge, anncensustrans_0626_alive_tomerge)
dim(anncensustrans_all_alive) #64 rows , 28 cols
head(anncensustrans_all_alive)
names(anncensustrans_all_alive)

#convert to long 
anncensustrans_all_alive_long <- reshape(anncensustrans_all_alive, direction = "long", sep = ".",
                                      idvar=c("block", "row", "col", "parent.pop", "mf", "rep"),
                                      varying =c("pheno.0522", "diam.0522", "height.0522", "total_branch.0522", "herb_dam.0522", "wilt_status.0522", "longest_leaf.0522", "flowers.0522", "fruits.0522", "longest_fruit.0522", "repro_branch.0522", "pheno.0626", "diam.0626", "height.0626", "total_branch.0626", "herb_dam.0626", "wilt_status.0626", "longest_leaf.0626", "flowers.0626", "fruits.0626", "longest_fruit.0626", "repro_branch.0626"))
head(anncensustrans_all_alive_long)
dim(anncensustrans_all_alive_long)
names(anncensustrans_all_alive_long)

anncensustrans_all_alive_max <- anncensustrans_all_alive_long %>% group_by(block, row, col, parent.pop, mf, rep) %>% summarise_if(is.double, max, na.rm=TRUE) %>% select(-time)
head(anncensustrans_all_alive_max) #some cases where max of different variables comes from different dates 
summary(anncensustrans_all_alive_max) #getting -Inf for longest_leaf and repro_branch b/c N/A at both times
#anncensustrans_all_alive_max %>% filter(longest_leaf==-Inf)
#anncensustrans_all_alive_long %>% filter(block=="D1", row==4)
#head(anncensustrans_all_alive_max)

#add in pop elevation info
names(anncensustrans_all_alive_max)
names(pop_elev_nonotes)
anncensustrans_all_alive_max_elev <- left_join(anncensustrans_all_alive_max, pop_elev_climate_avgs)
names(anncensustrans_all_alive_max_elev)
```

#Summary plots 
- diam
- height
- total_branch
- longest_leaf
- flowers
- fruits
- longest_fruit
- repro_branch 
```{r}
anncensustrans_all_alive_max_elev %>% ggplot(aes(x=diam)) + geom_histogram()
anncensustrans_all_alive_max_elev %>% ggplot(aes(x=height)) + geom_histogram()
anncensustrans_all_alive_max_elev %>% ggplot(aes(x=total_branch)) + geom_histogram()
anncensustrans_all_alive_max_elev %>% ggplot(aes(x=longest_leaf)) + geom_histogram()
anncensustrans_all_alive_max_elev %>% ggplot(aes(x=flowers)) + geom_histogram()
anncensustrans_all_alive_max_elev %>% ggplot(aes(x=fruits)) + geom_histogram()
anncensustrans_all_alive_max_elev %>% ggplot(aes(x=longest_fruit)) + geom_histogram()
anncensustrans_all_alive_max_elev %>% ggplot(aes(x=repro_branch)) + geom_histogram()


anncensustrans_all_alive_max_elev %>% ggplot(aes(x=parent.pop, y=diam)) + geom_boxplot()
anncensustrans_all_alive_max_elev %>% ggplot(aes(x=parent.pop, y=height)) + geom_boxplot()
anncensustrans_all_alive_max_elev %>% ggplot(aes(x=parent.pop, y=total_branch)) + geom_boxplot()
anncensustrans_all_alive_max_elev %>% ggplot(aes(x=parent.pop, y=longest_leaf)) + geom_boxplot()
anncensustrans_all_alive_max_elev %>% ggplot(aes(x=parent.pop, y=flowers)) + geom_boxplot()
anncensustrans_all_alive_max_elev %>% ggplot(aes(x=parent.pop, y=fruits)) + geom_boxplot()
anncensustrans_all_alive_max_elev %>% ggplot(aes(x=parent.pop, y=longest_fruit)) + geom_boxplot()
anncensustrans_all_alive_max_elev %>% ggplot(aes(x=parent.pop, y=repro_branch)) + geom_boxplot()
#variation across populations for diam, height, and longest leaf (maybe total_branch & repro_branch, not # flowers, prob not # fruits (all low escept TM2), not longest_fruit)
```

#Relationships between Climate and Traits
```{r}
names(anncensustrans_all_alive_max_elev)
ggplot(anncensustrans_all_alive_max_elev, aes(x = tmx, y = diam))+
  geom_point()

ggplot(anncensustrans_all_alive_max_elev, aes(x = ppt, y = diam))+
  geom_point()


ggplot(anncensustrans_all_alive_max_elev, aes(x = pck, y = diam))+
  geom_point()

ggplot(anncensustrans_all_alive_max_elev, aes(x = tmn, y = height))+
  geom_point()
ggplot(anncensustrans_all_alive_max_elev, aes(x = tmn, y = longest_leaf))+
  geom_point()
```


#Calculating and plotting means by pop (and elev)
First Priority
- diam
- height
- longest_leaf

Second Priority
- total_branch
- repro_branch 
- fruits
```{r}
head(anncensustrans_all_alive_max_elev)
summary(anncensustrans_all_alive_max_elev)
names(anncensustrans_all_alive_max_elev)

anncensustrans_all_alive_max_elev_noinf <- anncensustrans_all_alive_max_elev %>% mutate_if(is.double, list(~na_if(., -Inf))) #replace -INf with NAs
summary(anncensustrans_all_alive_max_elev_noinf)
head(anncensustrans_all_alive_max_elev_noinf)

anncensustrans_all_alive_summary <- anncensustrans_all_alive_max_elev_noinf %>% group_by(parent.pop, elevation.group, aet, cwd, pck, pet, ppt, rch, run, str, tmn, tmx) %>% summarise(meanDiam=mean(diam, na.rm = TRUE), sem_diam=sem(diam, na.rm=TRUE), meanHeight = mean(height,na.rm=(TRUE)), sem_height=sem(height, na.rm=(TRUE)), meanLongestLeaf=mean(longest_leaf, na.rm=(TRUE)), sem_leaf=sem(longest_leaf, na.rm=TRUE), meanTotBr=mean(total_branch, na.rm = TRUE), sem_totbr=sem(total_branch, na.rm=TRUE), meanRepBr=mean(repro_branch, na.rm=TRUE), semRepBr=sem(repro_branch, na.rm=TRUE), meanFruits=mean(fruits, na.rm = TRUE), semFruits=sem(fruits, na.rm=TRUE)) 
anncensustrans_all_alive_summary 
```


Mixed Effects Models 
#Figure out PCRs as part of mixed effects models?
```{r}
#Diam
anncensustrans_all_alive_summary %>% ggplot(aes(x=fct_reorder(parent.pop, meanDiam), y=meanDiam, fill=elevation.group)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanDiam-sem_diam,ymax=meanDiam+sem_diam),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))


lmediam2 = lmer(diam ~ (1|parent.pop/mf), data=anncensustrans_all_alive_max_elev_noinf)
summary(lmediam2)
ranova(lmediam2) #no maternal family effect b/c maybe not enough individuals 
#could try assinging unique names and "un-nesting" them 

lmediam3 = lmer(diam ~ ppt + (1|parent.pop/mf), data=anncensustrans_all_alive_max_elev_noinf)
summary(lmediam3)
ranova(lmediam3) 
#singular warning b/c we only have one value for each pop (multiple maternal families have one value)

lmediam4 = lmer(diam ~ ppt + (1|parent.pop), data=anncensustrans_all_alive_max_elev_noinf)
summary(lmediam4)
ranova(lmediam4) 
```


```{r}
#Height
anncensustrans_all_alive_summary %>% ggplot(aes(x=fct_reorder(parent.pop, meanHeight), y=meanHeight, fill=elevation.group)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanHeight-sem_height,ymax=meanHeight+sem_height),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

lmeheight = lmer(height ~ ppt + (1|parent.pop/mf), data=anncensustrans_all_alive_max_elev_noinf)
summary(lmeheight)
ranova(lmeheight)
```


```{r}
#Longest_leaf
anncensustrans_all_alive_summary %>% ggplot(aes(x=parent.pop, y=meanLongestLeaf, fill=elevation.group)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanLongestLeaf-sem_leaf,ymax=meanLongestLeaf+sem_leaf),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1))

lmelongest_leaf = lmer(longest_leaf ~ ppt + (1|parent.pop/mf), data=anncensustrans_all_alive_max_elev_noinf)
summary(lmelongest_leaf)
ranova(lmelongest_leaf)
```


```{r}
#Total_branch
anncensustrans_all_alive_summary %>% ggplot(aes(x=fct_reorder(parent.pop, meanTotBr), y=meanTotBr, fill=elevation.group)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanTotBr-sem_totbr,ymax=meanTotBr+sem_totbr),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1)) #large standard error bars 
```


```{r}
#Repro_branch
anncensustrans_all_alive_summary %>% ggplot(aes(x=fct_reorder(parent.pop, meanRepBr), y=meanRepBr, fill=elevation.group)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanRepBr-semRepBr,ymax=meanRepBr+semRepBr),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1)) #mostly the same except 1 SQ3 plant with a lot
```


```{r}
#Fruits
anncensustrans_all_alive_summary %>% ggplot(aes(x=fct_reorder(parent.pop, meanFruits), y=meanFruits, fill=elevation.group)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanFruits-semFruits,ymax=meanFruits+semFruits),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values=elev_three_palette) + 
  labs(fill="Elevation") +
   theme_classic() +
  xlab("Population") +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1)) #mostly the same (low), except TM2 had way more fruits than the other populations 
```


