---
title: "Single_Time_Size"
author: "Brandie Quarles"
date: "2023-11-07"
output: 
  html_document: 
    keep_md: yes
---

# Comparison of Size Traits at WL2 Garden



To Do:
-   Fix summary plots code to account for cases where N for length is </= 2
-   Fix mixed effects model code to account for cases where N for length is </= 2
-   Add climate to the models!


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Relevant Libraries and Functions

```{r}
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
library(lmerTest) #for mixed effect models
conflicted::conflicts_prefer(lmerTest::lmer)
library(broom.mixed) #tidy method for lmerTest
library(emmeans) #for post-hoc pairwise comparisons 
library(naniar) #replaces values with NA
library(corrplot) #plotting correlations 
library(rstatix) #performing cor_test
sem <- function(x, na.rm=FALSE) {           #for caclulating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low") #for proper arrangement in figures 
```

## Load the pop and location data

```{r}
#pop info
pops_common_garden <- read_csv("../input/WL2_Data/Pops_for_2023_WL2.csv") #pops included in common garden 
summary(pops_common_garden)
pops_common_garden_nonotes <- pops_common_garden %>% select(parent.pop:`seed year`)
pops_common_garden_nonotes$elevation.group <- str_to_title(pops_common_garden_nonotes$elevation.group)

#extra location info 
pop_loc <- read_csv("../input/Strep_tort_locs.csv")

unique(pop_loc$`Site code`)
unique(pops_common_garden_nonotes$parent.pop)
#need to change YOSE to YO
pop_loc_yo <- pop_loc %>% mutate(parent.pop = str_replace(`Site code`, "YOSE(\\d+)", "YO\\1")) %>% select(Lat, Long, elev_m=`Elevation (m)`, parent.pop)
unique(pop_loc_yo$parent.pop)

#merge in location info
pop_elev <- left_join(pops_common_garden_nonotes, pop_loc_yo)
head(pop_elev)
```

## Load size data using map

Want equivalent dates to Davis 
Dec (2 weeks in field) --> 0802
Jan (2 months in field) --> 0913
Feb (3 months in field) --> 1020 (this is only 1 week away from the annual census at WL2)
```{r}
files <- dir("../input/WL2_Data/CorrectedCSVs/",
             pattern="WL2_size_survey.*0802.*csv|WL2_size_survey.*0913.*csv|WL2_size_survey.*1020.*csv",
             full.names = TRUE)   #get list of desired files 
files

single_time_all <- tibble(path=files, filename=basename(path)) %>%  #make a tibble with the filenames and paths 
  mutate(survey_date=str_extract(filename, "20[0-9]*"),   #extract survey date from filenames
        survey_date=ymd(survey_date),                   #format the date column 
         survey_month=c("aug_size", "sept_size", "oct_size"))        #make a column for month
single_time_all

single_time_all <- single_time_all %>%                              #read in the data to "sheets" column
  mutate(sheets=map(path, 
                    read_csv,
                    na = c("", "NA", "-", "N/A"))) %>%              #set NA strings 
  select(-path)
single_time_all

map(single_time_all$sheets, head, 10)          #check to make sure the datasheets look good 
single_time_all$sheets[[3]] %>%  rowwise() %>% filter(!is.na(mf), mf != "buffer") %>%  filter(is.na(as.numeric(mf))) #all buffers --> safe to convert to double
single_time_all$sheets[[1]] %>%  rowwise() %>% filter(!is.na(rep), rep != "buffer") %>% filter(is.na(as.numeric(rep))) #all buffers --> safe to convert to double
single_time_all$sheets[[1]] <- single_time_all$sheets[[1]] %>% mutate(mf=as.double(mf),
                                                                      rep=as.double(rep)) 
single_time_all$sheets[[2]] <- single_time_all$sheets[[2]] %>% mutate(mf=as.double(mf),
                                                                      rep=as.double(rep)) 
single_time_all$sheets[[3]] <- single_time_all$sheets[[3]] %>% mutate(mf=as.double(mf),
                                                                      rep=as.double(rep)) 

map(single_time_all$sheets, colnames) %>% #check the column names 
  unlist() %>%
  unique() %>%
  sort()

#unnest to perform tidyverse functions 
single_time_all2 <- single_time_all %>% unnest(sheets) %>%
  rename(parent.pop=pop) %>% 
  filter(parent.pop != "buffer") %>% 
  filter(!is.na(mf)) %>% #make sure buffers are removed
  mutate(parent.pop= str_replace(parent.pop, ".*VTR.*", "LVTR1"))
head(single_time_all2)
unique(single_time_all2$parent.pop)

#merge in pop loc info
single_time_all2_elev <- left_join(single_time_all2, pop_elev) %>% select(survey_date:survey_month, block:survey.notes, elevation.group,Lat, Long, elev_m)
head(single_time_all2_elev)
names(single_time_all2_elev)
```

## Check Sample sizes at each survey

No pops with only 1 individual 

Create a data set that removes pops with 2 or less individuals for comparison.

```{r}
sample_sizes <- single_time_all2_elev %>% 
  group_by(survey_month, parent.pop) %>% 
  summarise(N_height = sum(!is.na(height.cm)), N_length = sum(!is.na(long.leaf.cm))) %>% 
  arrange(N_height)
sample_sizes

#Oct WR has only 2 individuals 
#WV has 0 individuals at all times

single_time_all_sample_sizes <- left_join(single_time_all2_elev, sample_sizes) %>% 
  filter(N_height != 0) #remove pops with no individuals alive for surveys
single_time_all_sample_sizes %>% filter(N_height == 0)
single_time_all_sample_sizes %>% filter(N_length == 0) #some pops with height data but no length data 

single_time_all_GTTWO <- single_time_all_sample_sizes %>% filter(N_height > 2) #only pops with greater than two individuals alive at a given survey 
```

## Summary plots

When sample size is greater than one:

```{r}
single_time_all_sample_sizes$survey_month <- factor(single_time_all_sample_sizes$survey_month, levels = c("aug_size","sept_size","oct_size")) #ensures the months are printed in chronological order 

#Histograms
single_time_all_sample_sizes %>% select(height.cm, long.leaf.cm) %>% 
  names() %>% 
  map(~ggplot(single_time_all_sample_sizes, aes_string(x = .)) + 
        geom_histogram() + 
        facet_grid(survey_month ~ ., scales = "free"))
#prints height first then longest leaf 

#qq plots 
single_time_all_sample_sizes %>% select(height.cm, long.leaf.cm) %>% 
  names() %>% 
  map(~ggplot(single_time_all_sample_sizes, aes_string(sample = .)) + 
       stat_qq() +
        stat_qq_line() +
        facet_grid(survey_month ~ ., scales="free"))

#all months height is right skewed
#oct longest leaf slightly right skewed


#Transformations:
single_time_all_sample_sizes_transf <- single_time_all_sample_sizes %>% 
  mutate(sheight.cm=sqrt(height.cm), logheight.cm=log(height.cm),log10height.cm=log10(height.cm),
         slong.leaf.cm=sqrt(long.leaf.cm), loglong.leaf.cm=log(long.leaf.cm),
         log10long.leaf.cm=log10(long.leaf.cm))

single_time_all_sample_sizes_transf$survey_month <- factor(single_time_all_sample_sizes_transf$survey_month, levels = c("aug_size","sept_size","oct_size"))

single_time_all_sample_sizes_transf %>% select(ends_with("cm")) %>% 
  names() %>% 
  map(~ggplot(single_time_all_sample_sizes_transf, aes_string(x = .)) + 
        geom_histogram() + 
        facet_grid(survey_month ~ ., scales = "free"))
#prints in the following order: height, longest leaf, sheight, logheight, log10height, slong.leaf.cm, loglong.leaf.cm, log10long.leaf.cm

single_time_all_sample_sizes_transf %>% select(ends_with("cm")) %>% 
  names() %>% 
  map(~ggplot(single_time_all_sample_sizes_transf, aes_string(sample = .)) + 
       stat_qq() +
        stat_qq_line() +
        facet_grid(survey_month ~ ., scales="free"))

#log transformation is the best for height (all months)
#no transformations greatly improve Oct distribution 
```

When sample size is greater than two:

```{r}
single_time_all_GTTWO$survey_month <- factor(single_time_all_GTTWO$survey_month, levels = c("aug_size","sept_size","oct_size")) #ensures the months are printed in chronological order

#Histograms
single_time_all_GTTWO %>% select(height.cm, long.leaf.cm) %>% 
  names() %>% 
  map(~ggplot(single_time_all_GTTWO, aes_string(x = .)) + 
        geom_histogram() + 
        facet_grid(survey_month ~ ., scales = "free"))
#prints height first then longest leaf 

#qq plots 
single_time_all_GTTWO %>% select(height.cm, long.leaf.cm) %>% 
  names() %>% 
  map(~ggplot(single_time_all_GTTWO, aes_string(sample = .)) + 
       stat_qq() +
        stat_qq_line() +
        facet_grid(survey_month ~ ., scales="free"))

#all months height is right skewed, similar to GTONE results 
#oct longest leaf very slightly right skewed, similar to GTONE results 


#Transformations:
single_time_all_GTTWO_transf <- single_time_all_GTTWO %>% 
  mutate(sheight.cm=sqrt(height.cm), logheight.cm=log(height.cm),log10height.cm=log10(height.cm),
         slong.leaf.cm=sqrt(long.leaf.cm), loglong.leaf.cm=log(long.leaf.cm),
         log10long.leaf.cm=log10(long.leaf.cm))

single_time_all_GTTWO_transf$survey_month <- factor(single_time_all_GTTWO_transf$survey_month, levels = c("aug_size","sept_size","oct_size"))

single_time_all_GTTWO_transf %>% select(ends_with("cm")) %>% 
  names() %>% 
  map(~ggplot(single_time_all_GTTWO_transf, aes_string(x = .)) + 
        geom_histogram() + 
        facet_grid(survey_month ~ ., scales = "free"))
#prints in the following order: height, longest leaf, sheight, logheight, log10height, slong.leaf.cm, loglong.leaf.cm, log10long.leaf.cm

single_time_all_GTTWO_transf %>% select(ends_with("cm")) %>% 
  names() %>% 
  map(~ggplot(single_time_all_GTTWO_transf, aes_string(sample = .)) + 
       stat_qq() +
        stat_qq_line() +
        facet_grid(survey_month ~ ., scales="free"))

#log transformation is the best for height (all months), similar to GTONE results 
#no transformations greatly improve Oct distribution, similar to GTONE results 
```

## Relationship between height and longest leaf

### Plotting Height and Length Together

```{r}
single_time_all_sample_sizes_transf %>% ggplot(aes(x=height.cm, y=long.leaf.cm)) +
  geom_point() +
  geom_smooth() +
  facet_grid(survey_month ~ ., scales="free")

#generally positive relationship at lower heights, but a lot more variability at lower heights
#tall plants have smaller leaves (likely b/c tall plants = reproductive)
#relationship not linear

single_time_all_sample_sizes_transf %>% ggplot(aes(x=long.leaf.cm, y=height.cm)) +
  geom_point() +
  geom_smooth() +
  facet_grid(survey_month ~ ., scales="free") #flipped axes
#not much of a relationship  
```

### Calculating correlation b/t the two

```{r}
single_time_all_sample_sizes_transf %>% cor_test(height.cm, long.leaf.cm, method = "pearson") #all months 
#Note, I used the pearson method event though height is not normally distributed...something to consider if we actually want to use this analysis
#r=0.24, P <0.0001

single_time_all_sample_sizes_transf %>% group_by(survey_month) %>% 
  cor_test(height.cm, long.leaf.cm, method = "pearson") #separate test for each month
#Aug r=0.37, P <0.0001
#Sept r=0.19, P <0.0001
#Oct r=0.36, P <0.0001

summary(lm(sheight.cm ~ long.leaf.cm, data=single_time_all_sample_sizes_transf))
```

## Calculating means by pop (and elev)

```{r}
single_time_all_sample_sizes_transf_summary <- single_time_all_sample_sizes_transf %>% 
  group_by(survey_month, parent.pop, elev_m) %>% 
  summarise(N_height = sum(!is.na(height.cm)), mean_height.cm = mean(height.cm,na.rm=(TRUE)), 
            sem_height.cm=sem(height.cm, na.rm=(TRUE)), N_length = sum(!is.na(long.leaf.cm)),
            mean_long.leaf.cm=mean(long.leaf.cm, na.rm=(TRUE)), 
            sem_long.leaf.cm=sem(long.leaf.cm, na.rm=TRUE)) %>% 
  arrange(survey_month, elev_m)
single_time_all_sample_sizes_transf_summary

single_time_all_sample_sizes_transf_summary %>% filter(N_height < 3)
single_time_all_sample_sizes_transf_summary %>% filter(N_length < 3) #4 plants that have at least 2 individuals for height, have limited sample size for length

#write_csv(single_time_all_sample_sizes_transf_summary, file ="../output/WL2_Traits/single_time_size_summary.csv")

#greater than 0 N for leaf length
single_time_all_sample_sizes_transf_summary_GTONE <- single_time_all_sample_sizes_transf_summary %>% 
  filter(N_length > 1)

#february overall avg
single_time_all_sample_sizes_transf %>% filter(survey_month=="oct_size") %>% summarise(AvgHeight=mean(height.cm, na.rm=(TRUE)), semHeight=sem(height.cm, na.rm=(TRUE)))

single_time_all_sample_sizes_transf %>% filter(survey_month=="oct_size") %>% filter(parent.pop!="TM2") %>% summarise(MAX=max(height.cm, na.rm=(TRUE)))

single_time_all_sample_sizes_transf %>% filter(survey_month=="oct_size") %>% filter(parent.pop!="TM2") %>% arrange(desc(height.cm))
```

## Figures of averages

### Avg Height Figure

```{r}
single_time_all_sample_sizes_transf_summary_GTONE$survey_month <- factor(single_time_all_sample_sizes_transf_summary_GTONE$survey_month, levels = c("aug_size","sept_size","oct_size"))

single_time_all_sample_sizes_transf_summary_GTONE %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_height.cm), y=mean_height.cm, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=mean_height.cm-sem_height.cm,ymax=mean_height.cm+sem_height.cm),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Height (cm)", x="Population") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) +
  facet_grid(survey_month ~ .)
ggsave("../output/WL2_Traits/Single_Time_Height.png", width = 14, height = 18, units = "in")
```

### Avg Longest Leaf Figure

```{r}
single_time_all_sample_sizes_transf_summary_GTONE %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_long.leaf.cm), y=mean_long.leaf.cm, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=mean_long.leaf.cm-sem_long.leaf.cm,ymax=mean_long.leaf.cm+sem_long.leaf.cm),width=.2, position = 
                  position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)", y="Avg Longest Leaf (cm)", x="Population") +
   theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) +
  facet_grid(survey_month ~ .)
ggsave("../output/WL2_Traits/Single_Time_Longest_Leaf.png", width = 14, height = 18, units = "in")
```

## Mixed Effects Models

### With a sample size of greater than 1 per pop
Prep data for model
```{r}
single_time_split_by_month_GTONE <- single_time_all_sample_sizes_transf %>% 
  select(survey_month, elev_m, parent.pop, mf, logheight.cm, long.leaf.cm) %>% 
  group_nest(survey_month)
single_time_split_by_month_GTONE #the data column contains the parameters in the model 

model_by_month_GTONE <- 
  single_time_split_by_month_GTONE %>% 
  mutate(height_model_1 = map(data, ~ lmer(logheight.cm ~ (1|parent.pop/mf), data = .x)),
         height_model_2 = map(data, ~ lmer(logheight.cm ~ elev_m + (1|parent.pop/mf), data = .x)),
         leaf_model_1 = map(data, ~ lmer(long.leaf.cm ~ (1|parent.pop/mf), data = .x)), 
         leaf_model_2 = map(data, ~ lmer(long.leaf.cm ~ elev_m + (1|parent.pop/mf), data = .x)),)
model_by_month_GTONE
```

Summary output for one model across survey_months 
```{r}
model_by_month_GTONE %>% 
  mutate(coef = map(height_model_1, tidy)) %>% 
  select(survey_month, coef) %>% 
  unnest(cols = c(coef))
```

Ranova output for each model (across months)
```{r}
#Model without random effects printed first, then nested predictor and then the predictor that nest was in
#(1 | mf:parent.pop) 
#(1 | parent.pop)
model_by_month_GTONE %>% #height with just pop and mf
  mutate(ranova = map(height_model_1, ranova)) %>% 
  select(survey_month, ranova) %>% 
  unnest(cols = c(ranova))
```

```{r}
model_by_month_GTONE %>% #height with elev, pop, and mf
  mutate(ranova = map(height_model_2, ranova)) %>% 
  select(survey_month, ranova) %>% 
  unnest(cols = c(ranova))
```

```{r}
model_by_month_GTONE %>% #length with just pop and mf
  mutate(ranova = map(leaf_model_1, ranova)) %>% 
  select(survey_month, ranova) %>% 
  unnest(cols = c(ranova))
```

```{r}
model_by_month_GTONE %>% #length with elev, pop, and mf
  mutate(ranova = map(leaf_model_2, ranova)) %>% 
  select(survey_month, ranova) %>% 
  unnest(cols = c(ranova))
```

Summary for all models per survey month
```{r}
tidy_models_GTONE <- model_by_month_GTONE %>% #get tidy summary of each model 
  mutate(coef_1 = map(height_model_1, tidy),
            coef_2 = map(height_model_2, tidy),
            coef_3 = map(leaf_model_1, tidy),
            coef_4 = map(leaf_model_2, tidy)) %>% 
  select(survey_month, coef_1:coef_4)
tidy_models_GTONE

tidy_models_GTONE_aug <- tidy_models_GTONE %>% filter(survey_month == "aug_size")
tidy_models_GTONE_sept <- tidy_models_GTONE %>% filter(survey_month == "sept_size")
tidy_models_GTONE_oct <- tidy_models_GTONE %>% filter(survey_month == "oct_size")

#convert to long 
models_long_GTONEaug <- reshape(tidy_models_GTONE_aug, direction = "long", sep = "_",
                           idvar = "survey_month", 
                            varying = c("coef_1", "coef_2", "coef_3", "coef_4")) %>% 
                unnest(coef) %>% 
  rename("model"="time") %>% 
  mutate(model = if_else(model==1, "height_model_1", 
                         if_else(model==2, "height_model_2",
                                 if_else(model==3, "leaf_model_1",
                                         "leaf_model_2"))))
models_long_GTONEaug
write_csv(models_long_GTONEaug, "../output/WL2_Traits/08022023_Size_Models_GTONE.csv")

models_long_GTONEsept <- reshape(tidy_models_GTONE_sept, direction = "long", sep = "_",
                           idvar = "survey_month", 
                            varying = c("coef_1", "coef_2", "coef_3", "coef_4")) %>% 
                unnest(coef) %>% 
  rename("model"="time") %>% 
  mutate(model = if_else(model==1, "height_model_1", 
                         if_else(model==2, "height_model_2",
                                 if_else(model==3, "leaf_model_1",
                                         "leaf_model_2"))))
models_long_GTONEsept
write_csv(models_long_GTONEsept, "../output/WL2_Traits/09132023_Size_Models_GTONE.csv")

models_long_GTONEoct <-  reshape(tidy_models_GTONE_oct, direction = "long", sep = "_",
                           idvar = "survey_month", 
                            varying = c("coef_1", "coef_2", "coef_3", "coef_4")) %>% 
                unnest(coef) %>% 
  rename("model"="time") %>% 
  mutate(model = if_else(model==1, "height_model_1", 
                         if_else(model==2, "height_model_2",
                                 if_else(model==3, "leaf_model_1",
                                         "leaf_model_2"))))
models_long_GTONEoct
write_csv(models_long_GTONEoct, "../output/WL2_Traits/10202023_Size_Models_GTONE.csv")
```
Elevation is negatively associated with height and length 


### With a sample size of greater than 2 per pop
Prep data for model
```{r}
single_time_split_by_month_GTTWO <- single_time_all_GTTWO_transf %>% 
  select(survey_month, elev_m, parent.pop, mf, logheight.cm, long.leaf.cm) %>% 
  group_nest(survey_month)
single_time_split_by_month_GTTWO #the data column contains the parameters in the model 

model_by_month_GTTWO <- 
  single_time_split_by_month_GTTWO %>% 
  mutate(height_model_1 = map(data, ~ lmer(logheight.cm ~ (1|parent.pop/mf), data = .x)),
         height_model_2 = map(data, ~ lmer(logheight.cm ~ elev_m + (1|parent.pop/mf), data = .x)),
         leaf_model_1 = map(data, ~ lmer(long.leaf.cm ~ (1|parent.pop/mf), data = .x)), 
         leaf_model_2 = map(data, ~ lmer(long.leaf.cm ~ elev_m + (1|parent.pop/mf), data = .x)),)
model_by_month_GTTWO
```

Summary output for one model across survey_months 
```{r}
model_by_month_GTTWO %>% 
  mutate(coef = map(height_model_1, tidy)) %>% 
  select(survey_month, coef) %>% 
  unnest(cols = c(coef))
```

Ranova output for each model (across months)
```{r}
#Model without random effects printed first, then nested predictor and then the predictor that nest was in
#(1 | mf:parent.pop) 
#(1 | parent.pop)
model_by_month_GTTWO %>% #height with just pop and mf
  mutate(ranova = map(height_model_1, ranova)) %>% 
  select(survey_month, ranova) %>% 
  unnest(cols = c(ranova))
```

```{r}
model_by_month_GTTWO %>% #height with elev, pop and mf
  mutate(ranova = map(height_model_2, ranova)) %>% 
  select(survey_month, ranova) %>% 
  unnest(cols = c(ranova))
```

```{r}
model_by_month_GTTWO %>% #length with just pop and mf
  mutate(ranova = map(leaf_model_1, ranova)) %>% 
  select(survey_month, ranova) %>% 
  unnest(cols = c(ranova))
```

```{r}
model_by_month_GTTWO %>% #length with elev pop and mf
  mutate(ranova = map(leaf_model_2, ranova)) %>% 
  select(survey_month, ranova) %>% 
  unnest(cols = c(ranova))
```


Summary for all models per survey month
```{r}
tidy_models_GTTWO <- model_by_month_GTTWO %>% #get tidy summary of each model 
  mutate(coef_1 = map(height_model_1, tidy),
            coef_2 = map(height_model_2, tidy),
            coef_3 = map(leaf_model_1, tidy),
            coef_4 = map(leaf_model_2, tidy)) %>% 
  select(survey_month, coef_1:coef_4)
tidy_models_GTTWO

tidy_models_GTTWO_aug <- tidy_models_GTTWO %>% filter(survey_month == "aug_size")
tidy_models_GTTWO_sept <- tidy_models_GTTWO %>% filter(survey_month == "sept_size")
tidy_models_GTTWO_oct <- tidy_models_GTTWO %>% filter(survey_month == "oct_size")

#convert to long 
models_long_GTTWOaug <- reshape(tidy_models_GTTWO_aug, direction = "long", sep = "_",
                           idvar = "survey_month", 
                            varying = c("coef_1", "coef_2", "coef_3", "coef_4")) %>% 
                unnest(coef) %>% 
  rename("model"="time") %>% 
  mutate(model = if_else(model==1, "height_model_1", 
                         if_else(model==2, "height_model_2",
                                 if_else(model==3, "leaf_model_1",
                                         "leaf_model_2"))))
models_long_GTTWOaug
write_csv(models_long_GTTWOaug, "../output/WL2_Traits/08022023_Size_Models_GTTWO.csv")

models_long_GTTWOsept <- reshape(tidy_models_GTTWO_sept, direction = "long", sep = "_",
                           idvar = "survey_month", 
                            varying = c("coef_1", "coef_2", "coef_3", "coef_4")) %>% 
                unnest(coef) %>% 
  rename("model"="time") %>% 
  mutate(model = if_else(model==1, "height_model_1", 
                         if_else(model==2, "height_model_2",
                                 if_else(model==3, "leaf_model_1",
                                         "leaf_model_2"))))
models_long_GTTWOsept
write_csv(models_long_GTTWOsept, "../output/WL2_Traits/09132023_Size_Models_GTTWO.csv")

models_long_GTTWOoct <-  reshape(tidy_models_GTTWO_oct, direction = "long", sep = "_",
                           idvar = "survey_month", 
                            varying = c("coef_1", "coef_2", "coef_3", "coef_4")) %>% 
                unnest(coef) %>% 
  rename("model"="time") %>% 
  mutate(model = if_else(model==1, "height_model_1", 
                         if_else(model==2, "height_model_2",
                                 if_else(model==3, "leaf_model_1",
                                         "leaf_model_2"))))
models_long_GTTWOoct
write_csv(models_long_GTTWOoct, "../output/WL2_Traits/10202023_Size_Models_GTTWO.csv")
```
Results similar to GTONE case for ranova and elev_m as a fixed effect. 


