---
title: "WL2_Survival-Analysis"
author: "Julin Maloof"
date: '`r Sys.Date()`'
output: 
  html_document: 
    keep_md: yes
---

YOU PROBABLY DON'T WANT THIS, IT DOESN'T WORK WELL.  SEE `WL2_SurvAnalysis_BRMS_JM.Rmd` INSTEAD



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Survival Analysis 

See tutorial: https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html

## Libraries
```{r}
# install.packages(c("ggsurvfit", "gtsummary"))
library(tidyverse) #includes lubridate, which we need 
library(ggrepel)
library(ggsurvfit)
library(gtsummary)
library(rstanarm)
library(bayesplot)
options(mc.cores = parallel::detectCores())
```

## Read in the data
```{r}
wl2_surv <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_mort_pheno_20231020_corrected.csv",
                     na = c("", "NA", "-", "N/A")) %>% #note this only goes to 10/20 need to come back and change this to include dates after this
  select(block:rep, death.date, survey.notes) %>% 
  rename(parent.pop=pop) %>% 
  mutate(parent.pop= str_replace(parent.pop, "Y08", "YO8")) %>% 
  mutate(parent.pop= str_replace(parent.pop, "Y04", "YO4")) %>% 
  filter(!is.na(parent.pop)) %>% 
  unite(BedLoc, bed:bed.col, sep="_", remove = FALSE) %>% 
  unite(Genotype, parent.pop:rep, sep="_", remove = FALSE) %>% 
  filter(!str_detect(Genotype, ".*buff*")) %>% 
  unite(pop.mf, parent.pop:mf, sep="_", remove = FALSE)
head(wl2_surv)
unique(wl2_surv$parent.pop)

wl2_surv %>% filter(Genotype=="CC_1_2") #there are 2 CC_1_2 plants (in different field locs), CC-1-2 was planted in 13-A and not 5C according to planting notes 
wl2_surv %>% filter(Genotype=="IH_4_5") #there are 2 IH_4_5 plants (in different field locs), IH_4_5 was planted in 22B and not 32A according to planting notes 
wl2_surv %>% rowwise() %>%  #checking if mf and rep can be converted to numeric 
  filter(!is.na(rep)) %>%  
  filter(is.na(as.numeric(rep)))
```

As weeks
```{r}
wl2_surv %>% filter(is.na(death.date),!is.na(survey.notes)) #no plants that seemed to disappear 
wl2_surv %>% filter(Genotype=="YO7_4_2")

wl2_surv_dates <- wl2_surv %>% 
  filter(BedLoc!="K_5_C") %>% 
  filter(BedLoc!="B_32_A") %>% 
  mutate(mf=as.double(mf), rep=as.double(rep)) %>% 
  mutate(planting.date="7/19/23", #could try to make this more specific to when certain blocks were planted 
         last_fup_date=if_else(is.na(death.date), "10/20/23", death.date)) %>%  #need this to calculate survival times
  mutate(planting.date=mdy(planting.date), last_fup_date=mdy(last_fup_date)) %>% #convert to date objects
  mutate(os_weeks=as.duration(planting.date %--% last_fup_date) / dweeks(1), #observed number of weeks
         status=if_else(is.na(death.date), 0, 1)) %>% #0=censured (alive in this case), 1=dead
  filter(os_weeks > 0) %>% #there is one case of a plant that was dead at planting, so just removed it since this is survival post- transplanting
  select(-survey.notes)
head(wl2_surv_dates)
unique(wl2_surv_dates$os_weeks)
```


```{r}
wl2_surv_dates %>% filter(parent.pop=="WL2") #there's a different number of WL2 plants here compared to annual census (see FirstYear Survv)
```

## Location Info
```{r}
gowersdist_UCD <- read_csv("../output/Climate/Pops_GowersEnvtalDist_UCD.csv") %>% 
  rename(Recent_Gowers_Dist_UCD = Recent_Gowers_Dist, Historic_Gowers_Dist_UCD = Historic_Gowers_Dist)
head(gowersdist_UCD)
gowersdist_WL2 <- read_csv("../output/Climate/Pops_GowersEnvtalDist_WL2.csv") %>% 
  rename(Recent_Gowers_Dist_WL2 = Recent_Gowers_Dist, Historic_Gowers_Dist_WL2 = Historic_Gowers_Dist)
head(gowersdist_WL2)

gowersdist_all <- full_join(gowersdist_UCD, gowersdist_WL2)
head(gowersdist_all)
```

```{r}
wl2_surv_dates_loc <- left_join(wl2_surv_dates, gowersdist_all) %>%
  mutate(elevation.group = factor(elevation.group, levels= c("Low", "Mid", "High")))
wl2_surv_dates_loc
```
## plot observed data

Need to summarize separately for pops and for elevation.  It is a mistake to average pops and then elevation, at least with respect to comparing to population naive models

### population level summary:
```{r}
# set up week 0 for 100% survival
week0.pop <- wl2_surv_dates_loc %>% 
  select(parent.pop, elevation.group) %>% 
  unique() %>%
  bind_cols(os_weeks=0,
            prop.surviving=1 )

wl2_obs.pop <- 
  wl2_surv_dates_loc %>% select(parent.pop, death.date, os_weeks, status, elevation.group) %>%
  group_by(parent.pop) %>%
  mutate(total=n()) %>%
  group_by(parent.pop, os_weeks) %>%
  summarize(n.newly.dead=sum(status),
            total=unique(total),
            elevation.group=unique(elevation.group)) %>%
  ungroup() %>%
  complete(parent.pop, os_weeks, fill = list(n.newly.dead=0)) %>% # we have missing data for weeks where nobody died and we need to fill that in
  group_by(parent.pop) %>%
  mutate(total=unique(na.omit(total)), elevation.group=unique(na.omit(elevation.group)), # filling in the blanks from the complete() step above
         total.dead=cumsum(n.newly.dead),
         surviving = total - total.dead,
         prop.surviving = surviving/total) %>%
  ungroup() %>%
  bind_rows(week0.pop) %>%
  select(parent.pop, elevation.group, os_weeks, prop.surviving) %>%
  arrange(parent.pop, os_weeks)

wl2_obs.pop %>%
  ggplot(aes(x=os_weeks, y = prop.surviving, color = elevation.group)) +
  geom_smooth() +
  geom_point()

wl2_obs.pop %>%
  ggplot(aes(x=os_weeks, y = prop.surviving, color = parent.pop)) +
  geom_line() +
  geom_point()
```

### elevation group summary
```{r}
week0.elev <- wl2_surv_dates_loc %>% 
  select(elevation.group) %>% 
  unique() %>%
  bind_cols(os_weeks=0,
            prop.surviving=1 )

wl2_obs.elev <-
  wl2_surv_dates_loc %>% select(death.date, os_weeks, status, elevation.group) %>%
  group_by(elevation.group) %>%
  mutate(total=n()) %>%
  group_by(elevation.group, os_weeks) %>%
  summarize(n.newly.dead=sum(status),
            total=unique(total)
            ) %>%
  ungroup() %>%
  complete(elevation.group, os_weeks, fill = list(n.newly.dead=0)) %>% # we have missing data for weeks where nobody died and we need to fill that in
  group_by(elevation.group) %>%
  mutate(total=unique(na.omit(total)), # filling in the blanks from the complete() step above
         total.dead=cumsum(n.newly.dead),
         surviving = total - total.dead,
         prop.surviving = surviving/total) %>%
  ungroup() %>%
  bind_rows(week0.elev) %>%
  select(elevation.group, os_weeks, prop.surviving) %>%
  arrange(elevation.group, os_weeks)


wl2_obs_elev.summary <- wl2_obs.elev %>%
  group_by(os_weeks, elevation.group) %>%
  summarize(prop.surviving = mean(prop.surviving))


wl2_obs_elev.summary %>%
  ggplot(aes(x=os_weeks, y=prop.surviving, color=elevation.group )) +
  geom_point() +
  geom_line()

```

## Simple fit

```{r}
f1 <- Surv(os_weeks, status) ~ 1
m1 <- stan_surv(f1, data= wl2_surv_dates_loc, chains=4)
m1_loo <- loo(m1)
```
```{r}
summary(m1)
```


```{r}
m1
```

```{r}
f2 <- Surv(os_weeks, status) ~ elevation.group
m2 <- stan_surv(f2, data= wl2_surv_dates_loc, chains=4)
m2_loo <- loo(m2)
```

```{r}
summary(m2)
```

```{r}
print(m2)
```





```{r}
# different # of knots for the m-splines
mdf10 <- update(m2, basehaz_ops = list(df = 10))
mdf10_loo <- loo(mdf10)

# quadratic instead of cubic splines
mdeg4 <- update(m2, basehaz_ops = list(degree = 4))
mdeg4_loo <- loo(mdeg4)

# quadratic splines with and 10 df
mdf10deg4 <- update(m3, basehaz_ops = list(degree = 4))
mdf10deg4_loo <- loo(mdf10deg4)

# specify knot location
mknots5 <- update(m2, basehaz_ops = list (knots = c(1,2,3,5,7)))
mknots5_loo <- loo(mknots5)
```


```{r}
loo_compare(m2_loo, mdf10_loo, mdeg4_loo, mdf10deg4_loo, mknots5_loo)
```

create a function to plot fits along with observations
```{r}
plot_fit <- function(fit, obs, newdata, title=NULL) {
  preds <- posterior_survfit(fit, newdata = newdata, times = 0, extrapolate = TRUE)
  
  preds <- preds %>%
    as_tibble() %>%
    mutate(elevation.group = newdata$elevation.group[id])
  
  pl <- preds %>%
    ggplot(aes(x=time, y = median, ymin=ci_lb, ymax = ci_ub, group = elevation.group)) +
    geom_ribbon(fill="grey70", alpha=.5) +
    geom_line(aes(color=elevation.group), lwd=1.5) +
    scale_color_viridis_d() +
    
    geom_point(aes(x = os_weeks, y=prop.surviving, color = elevation.group), data=wl2_obs_elev.summary, inherit.aes = FALSE) +
    ylab("proportion surviving") +
    xlab("weeks after transplant") 
  
  if(!is.null(title)) pl <- pl + ggtitle(title)
  
  pl
}
```

Make some plots
```{r}
nd <- data.frame(elevation.group = levels(wl2_surv_dates_loc$elevation.group))

plot_fit(m2, wl2_obs_elev.summary, nd, "m2 default") 
plot_fit(mdeg4, wl2_obs_elev.summary, nd, "quadratic splines")
plot_fit(mdf10, wl2_obs_elev.summary, nd, "10 degrees of freedom")
plot_fit(mdf10deg4, wl2_obs_elev.summary, nd, "10 degrees of freedom and quadratic splines")
plot_fit(mknots5, wl2_obs_elev.summary, nd, "5 manual knots")


```
Let's try some variations on having tve coefficients

```{r}
mtve_default <- stan_surv(Surv(os_weeks, status) ~ tve(elevation.group), data = wl2_surv_dates_loc, chains=4) # 900 seconds
mtve_default_loo <- loo(mtve_default)

mtve_base_10df <- update(mtve_default, basehaz_ops = list(df = 10) ) # 900 seconds
mtve_base_10df_loo <- loo(mtve_base_10df)

mtve_tve_5knots <- update(mtve_default, formula = Surv(os_weeks, status) ~ tve(elevation.group, knots=c(1,2,3,5,7)) )
mtve_tve_5knots_loo <- loo(mtve_tve_5knots)

mtve_base_10df_tve_5knots <- update(mtve_tve_5knots, basehaz_ops = list(df = 10) ) # 1 hour
mtve_base_10df_tve_5knots_loo <- loo(mtve_base_10df_tve_5knots)

```


```{r}
loo_compare(mget(ls(pattern="loo")))
```
```{r}
plot_fit(m2, wl2_obs_elev.summary, nd, "m2 default") 
plot_fit(mdf10, wl2_obs_elev.summary, nd, "10 degrees of freedom")
plot_fit(mtve_default, wl2_obs_elev.summary, nd, "tve default")

plot_fit(mtve_tve_5knots, wl2_obs_elev.summary, nd, "tve 5 tve knots")

plot_fit(mtve_base_10df, wl2_obs_elev.summary, nd, "tve 10 base df")

plot_fit(mtve_base_10df_tve_5knots, wl2_obs_elev.summary, nd, "tve 10 base df 5 tve knots")


```


```{r}
mtve_tve_7knots <- update(mtve_default, formula = Surv(os_weeks, status) ~ tve(elevation.group, knots=c(1, 2, 3, 4, 6, 7,10)) )
mtve_tve_7knots_loo <- loo(mtve_tve_7knots)
```

exponentional model

```{r}
m_exp <- stan_surv(Surv(os_weeks, status) ~ elevation.group, basehaz = "exp" , data = wl2_surv_dates_loc, chains=4) 
m_exp_loo <- loo(m_exp)

m_expb <- stan_surv(Surv(os_weeks, status) ~ elevation.group, basehaz = "exp" , data = {wl2_surv_dates_loc %>% mutate(elevation.group=relevel(elevation.group, ref="High"))}
  , chains=4) 
m_expb_loo <- loo(m_expb)

m_exp_tve <- stan_surv(Surv(os_weeks, status) ~ tve(elevation.group), basehaz = "exp" , data = wl2_surv_dates_loc, chains=4) 
m_exp_tve_loo <- loo(m_exp)
```

```{r}
plot_fit(m2, wl2_obs_elev.summary, nd, "m2 default") 
plot_fit(m_exp, wl2_obs_elev.summary, nd, "exp model")
plot_fit(m_expb, wl2_obs_elev.summary, nd, "exp model b")

plot_fit(m_exp_tve, wl2_obs_elev.summary, nd, "exp tve model")
```

Seems like priors are too stringent?

```{r}
m_exp
```


```{r}
prior_summary(m_exp)
```
```{r}
m_exp %>%
  as.matrix() %>%
  mcmc_areas()
```

save the models so far

```{r}
save(list=ls(pattern="^m"), file="../output/WL2_Stan_Surv_Models_JM.Rdata")
```


Fit a random effects model
```{r}
f3 <- Surv(os_weeks, status) ~ elevation.group + (1|parent.pop)
m3_ran <- update(m3, formula = f3)
m3_ran_loo <- loo(m3_ran)
loo_compare(m3_loo, m3_ran_loo)
```

```{r}
m3_ran
summary(m3_ran)
```


```{r}
nd2 <- wl2_surv_dates_loc %>%
  select(parent.pop, elevation.group) %>%
  unique()

preds <- posterior_survfit(m3_ran, newdata = nd2, times = 0, extrapolate = TRUE)


preds <- preds %>%
  as_tibble() %>%
  mutate(elevation.group = nd2$elevation.group[id],
         parent.pop = nd2$parent.pop[id]) 

pred.labels <- preds %>%
  filter(time==max(time))

preds %>%
  ggplot(aes(x=time, y = median, ymin=ci_lb, ymax = ci_ub, group = parent.pop)) +
  geom_ribbon(fill="grey70", alpha=.5) +
  geom_line(aes(color=elevation.group)) +
  geom_label_repel(aes(label = parent.pop), data = pred.labels, max.overlaps = 20, xlim=c(14, 17), direction = "both", size=2.5, label.padding = .1, force_pull=10) +
  #facet_wrap(~ elevation.group)
  scale_color_viridis_d() +
  coord_cartesian(xlim = c(0, 16)) +
  ylab("proportion surviving") +
  xlab("weeks after transplant")
```

```{r}
preds %>%
  group_by(elevation.group, time) %>%
  summarize(median = mean(median),
            ci_lb = mean(ci_lb), # this is probably not correct
            ci_ub = mean(ci_ub)) %>%
  ggplot(aes(x=time, y = median, ymin=ci_lb, ymax = ci_ub, group = elevation.group)) +
  geom_ribbon(fill="grey70", alpha=.5) +
  geom_line(aes(color=elevation.group), lwd=2) +
  
  geom_point(aes(x = os_weeks, y=prop.surviving, color = elevation.group), data=wl2_obs_elev.summary, inherit.aes = FALSE) +
  geom_errorbar(aes(x = os_weeks, ymin=prop.surviving-sem, ymax= prop.surviving+sem,  color = elevation.group), data=wl2_obs_elev.summary, inherit.aes = FALSE, width=.25, lwd=.75) +
  
  #facet_wrap(~ elevation.group)
  scale_color_viridis_d() 
```

But can I have tve for group? (yes)

deegree=0 fits piecewise constant function for elevation group.   Otherwise b-splines.



```{r}
f10  <- Surv(os_weeks, status) ~ tve(elevation.group)
m10 <- stan_surv(f10, data= wl2_surv_dates_loc, chains=4)
m10_loo <- loo(m10)
```
```{r}
m10
```

```{r}
summary(m10)
```

```{r}
preds <- posterior_survfit(m10, newdata = nd, times = 0, extrapolate = TRUE)

preds <- preds %>%
  as_tibble() %>%
  mutate(elevation.group = nd$elevation.group[id])

preds %>%
  ggplot(aes(x=time, y = median, ymin=ci_lb, ymax = ci_ub, group = elevation.group)) +
  geom_ribbon(fill="grey70", alpha=.5) +
  geom_line(aes(color=elevation.group), lwd=1.5) +
  
  geom_point(aes(x = os_weeks, y=prop.surviving, color = elevation.group), data=wl2_obs_elev.summary, inherit.aes = FALSE) +
  geom_errorbar(aes(x = os_weeks, ymin=prop.surviving-sem, ymax= prop.surviving+sem,  color = elevation.group), data=wl2_obs_elev.summary, inherit.aes = FALSE, width=.25, lwd=.75) +
  
  scale_color_viridis_d()
```
```{r}
loo_compare(m2_loo, m3_loo, m10_loo)
```




"The survfit() function creates survival curves using the Kaplan-Meier method based on a formula. Let’s generate the overall survival curve for the entire cohort, assign it to object s1, and look at the structure using str():"
```{r}
s1 <- survfit(Surv(os_weeks, status) ~ 1, data = wl2_surv_dates_loc)
str(s1)
#time: the timepoints at which the curve has a step, i.e. at least one event occurred
#surv: the estimate of survival at the corresponding time
```

## Survival Plot
```{r}
survfit2(Surv(os_weeks, status) ~ 1, data = wl2_surv_dates_loc) %>% 
  ggsurvfit() +
  labs(
    x = "Weeks",
    y = "Overall survival probability"
  ) + 
  add_confidence_interval() +
  add_risktable() #at risk = plants still alive, event=plants dead
```

## Estimating x-week survival
```{r}
summary(survfit(Surv(os_weeks, status) ~ 1, data = wl2_surv_dates_loc), times = 8) #survival for 2 months
#2 month survival probability is 41%
```

## Estimating median survival time
```{r}
survfit(Surv(os_weeks, status) ~ 1, data = wl2_surv_dates_loc) 
#median survival time is 3 weeks
```

## Compare Survival between Populations 
Plot survival differences
```{r}
survfit2(Surv(os_weeks, status) ~ elevation.group, data = wl2_surv_dates_loc) %>% 
  ggsurvfit() +
  labs(
    x = "Weeks",
    y = "Overall survival probability") + 
  add_confidence_interval() 

survfit(Surv(os_weeks, status) ~ parent.pop, data = wl2_surv_dates_loc) #median survival lengths for each pop
```

"We can conduct between-group significance tests using a log-rank test. The log-rank test equally weights observations over the entire follow-up time and is the most common way to compare survival times between groups."
```{r}
survdiff(Surv(os_weeks, status) ~ parent.pop, data = wl2_surv_dates_loc)
#significant difference in overall survival according to population
```

## Cox regression model
"We may want to quantify an effect size for a single variable, or include more than one variable into a regression model to account for the effects of multiple variables.
The Cox regression model is a semi-parametric model that can be used to fit univariable and multivariable regression models that have survival outcomes.:
```{r}
coxph(Surv(os_weeks, status) ~ parent.pop, data = wl2_surv_dates_loc) %>% 
  tbl_regression(exp = TRUE) 

# hazard ratio (HR) represents the ratio of hazards between two groups at any particular point in time. 
#A HR < 1 indicates reduced hazard of death whereas a HR > 1 indicates an increased hazard of death.
```
