---
title: "WL2_Spatial_Variation"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Evaluating the phenotypic spatial variation (across beds) at WL2

Purpose: See if we could take advantage of the natural variation at WL2 for the 2026 planting

-   Some packages in R for autocorrelation

To Do:

-   Add size 

-   Spatial plots for all variables 

-   For TDR data, compare top of beds avg across the season (See Rachel's github)

-   Correlate TDR data to fitness/size data somehow 

## Load Libraries

```{r}
library(tidyverse)
library(viridis)

sem <- function(x, na.rm=FALSE) {           #for caclulating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 
```

## Load Fitness Data

```{r}
wl2_est <- read_csv("../output/WL2_Traits/WL2_Establishment.csv") %>% select(block:rep, Establishment) %>% 
  mutate(bed.col=if_else(bed.col=="A -", "A", bed.col))

y1_surv <- read_csv("../output/WL2_Traits/WL2_Y1Surv.csv") %>% 
  left_join(wl2_est) %>% 
  select(block, bed:bed.col, Genotype:rep, Y1Surv=Survival) 

wintersurv <- read_csv("../output/WL2_Traits/WL2_WinterSurv_2324.csv") %>% 
  rename(bed.row=`bed- row`, bed.col=`bed- col`) %>% 
  select(block, bed:bed.col, Genotype:rep, WinterSurv)

repsurvy2 <- read_csv("../output/WL2_Traits/WL2_Surv_to_Rep_Y2_2324.csv") %>% 
  select(block, bed:col, Genotype, pop:rep, SurvtoRep_y2)

fruitsy2 <- read_csv("../output/WL2_Traits/WL2_Fruits_Y2_2324.csv") %>% 
  select(block, bed:col, Genotype, pop:rep, fruits)

probfruit <- read_csv("../output/WL2_Traits/WL2_ProbFruits_Y2_2324.csv") %>% 
  select(block, bed:col, Genotype, pop:rep, ProbFruits)

totfruit <- read_csv("../output/WL2_Traits/WL2_TotalRepOutput.csv") %>% 
  select(block:rep, Total_Fitness)
```

## Load Size (Height 2 months post transplant, stem diameter at end of season 2, canopy area?)
```{r}

```

## Box Plots
```{r}
wl2_est %>% 
  ggplot(aes(x=bed, y=Establishment)) +
  geom_boxplot()

y1_surv %>% 
  ggplot(aes(x=bed, y=Y1Surv)) +
  geom_boxplot()

wintersurv %>% 
  ggplot(aes(x=bed, y=WinterSurv)) +
  geom_boxplot()

repsurvy2 %>% 
  ggplot(aes(x=bed, y=SurvtoRep_y2)) +
  geom_boxplot()

fruitsy2 %>% 
  ggplot(aes(x=bed, y=fruits)) +
  geom_boxplot()

probfruit %>% 
  ggplot(aes(x=bed, y=ProbFruits)) +
  geom_boxplot()

totfruit %>% 
  ggplot(aes(x=bed, y=Total_Fitness)) +
  geom_boxplot()
```

## Summaries
```{r}
wl2_est_summary <- wl2_est %>% 
  group_by(bed) %>% 
  summarise(meanSurv=mean(Establishment, na.rm=TRUE), semSurv=sem(Establishment, na.rm=TRUE))

y1surv_summary <- y1_surv %>% 
  group_by(bed) %>% 
  summarise(meanSurv=mean(Y1Surv, na.rm=TRUE), semSurv=sem(Y1Surv, na.rm=TRUE))

wintersurv_summary <- wintersurv %>% 
  group_by(bed) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm=TRUE), semSurv=sem(WinterSurv, na.rm=TRUE))

repsurvy2_summary <- repsurvy2 %>% 
  group_by(bed) %>% 
  summarise(meanSurv=mean(SurvtoRep_y2, na.rm=TRUE), semSurv=sem(SurvtoRep_y2, na.rm=TRUE))

fruitsy2_summary <- fruitsy2 %>% 
  group_by(bed) %>% 
  summarise(meanSurv=mean(fruits, na.rm=TRUE), semSurv=sem(fruits, na.rm=TRUE))

probfruit_summary <- probfruit %>% 
  group_by(bed) %>% 
  summarise(meanSurv=mean(ProbFruits, na.rm=TRUE), semSurv=sem(ProbFruits, na.rm=TRUE))

totfruit_summary <- totfruit %>% 
  group_by(bed) %>% 
  summarise(meanSurv=mean(Total_Fitness, na.rm=TRUE), semSurv=sem(Total_Fitness, na.rm=TRUE))
```

## Bar Plots
```{r}
wl2_est_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanSurv), y=meanSurv)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Establishment")

y1surv_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanSurv), y=meanSurv)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Y1 Surv")

wintersurv_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanSurv), y=meanSurv)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Winter Surv")

repsurvy2_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanSurv), y=meanSurv)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Surv to Bud Y2")

fruitsy2_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanSurv), y=meanSurv)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Fruits Y2")

probfruit_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanSurv), y=meanSurv)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Prob. Fruit")

totfruit_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanSurv), y=meanSurv)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Total Fruit")
```

## Spatial Plots Prep
x = beds + cols
y = row 
```{r}
bed_col_info <- wl2_est %>% 
  select(bed, bed.col) %>% 
  distinct() %>% 
  mutate(x_pos = if_else(bed.col=="A", 1,
                         if_else(bed.col=="B", 2, 
                                 if_else(bed.col=="C", 3, 4)))) %>% 
  mutate(x_pos=if_else(bed=="B", x_pos+6,
                       if_else(bed=="C", x_pos+12,
                               if_else(bed=="D", x_pos+18,
                                       if_else(bed=="E", x_pos+24,
                                               if_else(bed=="F", x_pos+30,
                                                       if_else(bed=="G", x_pos+36,
                                                               if_else(bed=="H", x_pos+42, 
                                                                       if_else(bed=="I", x_pos+48,
                                                                               if_else(bed=="J", x_pos+54, 
                                                                                       if_else(bed=="K", x_pos+60, x_pos))))))))))) %>% 
  arrange(bed, bed.col)

bed_col_info
```

## Spatial Plots
```{r}
wl2_est %>% 
  left_join(bed_col_info) %>% 
  mutate(Establishment=as.character(Establishment)) %>% 
  ggplot(aes(x=x_pos, y=bed.row, colour=Establishment)) +
  geom_point(size=3) +
  scale_y_reverse() +
  scale_color_manual(values = c("gray", "darkgreen")) + 
  theme_classic()
```
