---
title: "WL2_climate_distance"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Climate Distance at the WL2 Garden

## Relevant Libraries and Functions

```{r}
library(raster)
library(tidyverse)
library(conflicted)
conflicts_prefer(dplyr::select())
conflicts_prefer(dplyr::filter)
library(cowplot)
library(QBMS) #for function calc_biovars to calculate bioclim variables
library(boot)
library(broom)
library(ggpubr) #for ggarange 

sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
} #legend function for grid_arrange

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")
```

## WL2 Climate Data (For plants measured July 2023-Oct 2023)

### From Flint (changed this from Magney MetStation)

```{r}
WL2_climate <- read_csv("../output/Climate/flint_climate_UCDpops.csv") %>% 
  filter(parent.pop=="WL2_Garden") %>% 
  filter(year>=2022, year<2024) %>% 
  filter(if_else(year==2022, month=="nov" | month=="dec", #2022-2023 Water Year 
                 month!="nov" & month !="dec")) %>% 
  mutate(tavg = (tmn + tmx)/2, t_diurnal = (tmx-tmn))
head(WL2_climate)
summary(WL2_climate)

WL2_climate_flint <- WL2_climate %>% #get the means for the variables to compare to home sites 
  summarise(cwd_WL2=mean(cwd),ppt_WL2=mean(ppt), pck_WL2=mean(pck), tmn_WL2=mean(tmn), tmx_WL2=mean(tmx))
WL2_climate_flint
```

## WL2 Climate Trends (2023)
```{r}
WL2_climate$month <- factor(WL2_climate$month, levels = c("nov","dec","jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct"))

WL2_climate %>% 
  ggplot(aes(x=month,y=cwd)) +
  geom_point()

WL2_climate %>% 
  ggplot(aes(x=month,y=pck)) +
  geom_point() 

WL2_climate %>% 
  ggplot(aes(x=month,y=tmx)) +
  geom_point()

WL2_climate %>% 
  ggplot(aes(x=month,y=tmn)) +
  geom_point()

WL2_climate %>% 
  ggplot(aes(x=month,y=ppt)) +
  geom_point()
```

### BioClim

#### Prep
```{r}
wl2_bioclim_allyear_prep <- WL2_climate %>% 
  mutate(wtr_year="2023") %>% 
  rename(tmin=tmn, tmax=tmx, year_cal=year, year=wtr_year) %>% #rename columns to match what calc_biovars expects 
  arrange(parent.pop, year, month)
```

### Calculation
```{r}
wl2_bioclim_all_year <- wl2_bioclim_allyear_prep %>% calc_biovars() 
```

#### SUBSET
```{r}
wl2_bioclim_all_year_final <- wl2_bioclim_all_year %>% 
  select(ann_tmean_WL2=bio1, mean_diurnal_range_WL2=bio2, 
         temp_seasonality_WL2=bio4, temp_ann_range_WL2=bio7, tmean_wettest_quarter_WL2=bio8,
         tmean_driest_quarter_WL2=bio9, ann_ppt_WL2=bio12, ppt_seasonality_WL2=bio15,
         ppt_warmest_quarter_WL2=bio18, ppt_coldest_quarter_WL2=bio19) %>% 
  mutate_if(is.character, as.numeric)
wl2_bioclim_all_year_final
```

## WL2 Climate Data for 2024

### From Flint

```{r}
WL2_climate_2024 <- read_csv("../output/Climate/flint_climate_UCDpops.csv") %>% 
  filter(parent.pop=="WL2_Garden") %>% 
  filter(year>=2023, year<=2024) %>% 
  filter(if_else(year==2023, month=="nov" | month=="dec", #2023-2024 Water Year 
                 month!="nov" & month !="dec")) %>% 
  mutate(tavg = (tmn + tmx)/2, t_diurnal = (tmx-tmn))
head(WL2_climate_2024)
summary(WL2_climate_2024)

WL2_climate_2024_flint <- WL2_climate_2024 %>% #get the means for the variables to compare to home sites 
  summarise(cwd_WL2=mean(cwd),ppt_WL2=mean(ppt), pck_WL2=mean(pck), tmn_WL2=mean(tmn), tmx_WL2=mean(tmx))
WL2_climate_2024_flint
```

## WL2 Climate Trends (2024)
```{r}
WL2_climate_2024$month <- factor(WL2_climate_2024$month, levels = c("nov","dec","jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct"))

WL2_climate_2024 %>% 
  ggplot(aes(x=month,y=cwd)) +
  geom_point()

WL2_climate_2024 %>% 
  ggplot(aes(x=month,y=pck)) +
  geom_point() 

WL2_climate_2024 %>% 
  ggplot(aes(x=month,y=tmx)) +
  geom_point()

WL2_climate_2024 %>% 
  ggplot(aes(x=month,y=tmn)) +
  geom_point()

WL2_climate_2024 %>% 
  ggplot(aes(x=month,y=ppt)) +
  geom_point()
```

### BioClim

#### Prep
```{r}
wl2_bioclim_allyear_prep_2024 <- WL2_climate_2024 %>% 
  mutate(wtr_year="2024") %>% 
  rename(tmin=tmn, tmax=tmx, year_cal=year, year=wtr_year) %>% #rename columns to match what calc_biovars expects 
  arrange(parent.pop, year, month)
```

### Calculation
```{r}
wl2_bioclim_all_year_2024 <- wl2_bioclim_allyear_prep_2024 %>% calc_biovars() 
```

#### SUBSET
```{r}
wl2_bioclim_all_year_final_2024 <- wl2_bioclim_all_year_2024 %>% 
  select(ann_tmean_WL2=bio1, mean_diurnal_range_WL2=bio2, 
         temp_seasonality_WL2=bio4, temp_ann_range_WL2=bio7, tmean_wettest_quarter_WL2=bio8,
         tmean_driest_quarter_WL2=bio9, ann_ppt_WL2=bio12, ppt_seasonality_WL2=bio15,
         ppt_warmest_quarter_WL2=bio18, ppt_coldest_quarter_WL2=bio19) %>% 
  mutate_if(is.character, as.numeric)
wl2_bioclim_all_year_final_2024
```


## Gower's Climate Distance

(1/P) \* SUM ((absolute value(Ai - Bi)) / range(i)) for each variable

-   P = number of environmental variables = 13 (without CWD)

-   Ai = 30 year avg of that variable for the home site

-   Bi = July 2023-Dec 2023 avg of that variable for the WL2 garden

-   Range(i) = maximum - minimum of that variable in the whole data set
    (across sites)

### Prep WL2 data for range calculations

```{r}
WL2_climate_all <- bind_cols(WL2_climate_flint, wl2_bioclim_all_year_final)

WL2_range_prep <- WL2_climate_all %>% 
  mutate(parent.pop="WL2_Garden") %>% 
  rename_with(~str_remove(., "_WL2"), everything())
```

### Bootstrapping 
#### Load data with all 30 years
```{r}
recent_clim_boot <- read_csv("../output/Climate/fullyear_wtr_year_avgs_Recent.csv")

recent_clim_boot_nest <- recent_clim_boot %>% nest(.by=year) #nest to prepare for bootstrapping 
recent_clim_boot_nest

historical_clim_boot <- read_csv("../output/Climate/fullyear_wtr_year_avgs_Historical.csv")

historical_clim_boot_nest <- historical_clim_boot %>% nest(.by=year)
historical_clim_boot_nest
```

#### Create the gower_calc function 
```{r}
#data <- recent_clim_boot_nest
#P=14

gowers_calc <- function(data, indices, P) { #function with all of the code necessary for calculating gowers distance 
  #data = _clim_boot (recent or historical) - needs to be nested by year; P = # climate variables 

  #need to make WL2_range_prep before running this function 
  
  data <-data[indices,] # subset per bootstrap indices
  
  data <- data %>% unnest(data) #unnest so the function can access the climate data
  
  data_means <- data %>% 
    group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
    summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx", "ann_tmean", "mean_diurnal_range", 
                   "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_quarter", "tmean_driest_quarter", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_quarter", "ppt_coldest_quarter"),
               c(mean), na.rm = TRUE) #get 30 year averages for each climate variable 
  
  range_merge <- bind_rows(data_means, WL2_range_prep)
  
  WL2_home_climate_ranges <- range_merge %>% #calculate ranges
    ungroup() %>% 
  summarise(cwd_range=max(cwd)-min(cwd),
            pck_range=max(pck)-min(pck),
            ppt_range=max(ppt)-min(ppt), 
            tmn_range=max(tmn)-min(tmn), 
            tmx_range=max(tmx)-min(tmx), 
            ann_tmean_range=max(ann_tmean)-min(ann_tmean),
            mean_diurnal_range_range=max(mean_diurnal_range)-min(mean_diurnal_range),
            temp_seasonality_range=max(temp_seasonality)-min(temp_seasonality),
            temp_ann_range_range=max(temp_ann_range)-min(temp_ann_range),
            tmean_wettest_quarter_range=max(tmean_wettest_quarter)-min(tmean_wettest_quarter),
            tmean_driest_quarter_range=max(tmean_driest_quarter)-min(tmean_driest_quarter),
            ann_ppt_range=max(ann_ppt)-min(ann_ppt), 
            ppt_seasonality_range=max(ppt_seasonality)-min(ppt_seasonality),
            ppt_warmest_quarter_range=max(ppt_warmest_quarter)-min(ppt_warmest_quarter), 
            ppt_coldest_quarter_range=max(ppt_coldest_quarter)-min(ppt_coldest_quarter))
  
  WL2_home_climate <- bind_cols(WL2_climate_all, data_means) #add WL2 climate data to home climate data 
  
  WL2_home_climate_with_ranges <- bind_cols(WL2_home_climate, WL2_home_climate_ranges) #add in ranges 
  
  gowers_calc_each_var <- WL2_home_climate_with_ranges %>% #variable by variable calc
  mutate(cwd_gowers=abs(cwd_WL2-cwd) / cwd_range,
         pck_gowers=abs(pck_WL2-pck) / pck_range,
         ppt_gowers=abs(ppt_WL2 - ppt) / ppt_range,
         tmn_gowers=abs(tmn_WL2 - tmn) / tmn_range,
         tmx_gowers=abs(tmx_WL2 - tmx) / tmx_range,
         ann_tmean_gowers=abs(ann_tmean_WL2 - ann_tmean) / ann_tmean_range,
         mean_diurnal_range_gowers=abs(mean_diurnal_range_WL2 - mean_diurnal_range) / mean_diurnal_range_range,
         temp_seasonality_gowers=abs(temp_seasonality_WL2 - temp_seasonality) / temp_seasonality_range,
         temp_ann_range_gowers=abs(temp_ann_range_WL2 - temp_ann_range) / temp_ann_range_range,
         tmean_wettest_quarter_gowers=abs(tmean_wettest_quarter_WL2 - tmean_wettest_quarter) / tmean_wettest_quarter_range,
         tmean_driest_quarter_gowers=abs(tmean_driest_quarter_WL2 - tmean_driest_quarter) / tmean_driest_quarter_range,
         ann_ppt_gowers=abs(ann_ppt_WL2 - ann_ppt) / ann_ppt_range,
         ppt_seasonality_gowers=abs(ppt_seasonality_WL2 - ppt_seasonality) / ppt_seasonality_range,
         ppt_warmest_quarter_gowers=abs(ppt_warmest_quarter_WL2 - ppt_warmest_quarter) / ppt_warmest_quarter_range,
         ppt_coldest_quarter_gowers=abs(ppt_coldest_quarter_WL2 - ppt_coldest_quarter) / ppt_coldest_quarter_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))

 gowers_calc_per_pop <- gowers_calc_each_var %>% #final gowers calc 
  mutate(Gowers_Dist=(1/P)*(cwd_gowers + pck_gowers + ppt_gowers + tmn_gowers + tmx_gowers +
                                ann_tmean_gowers + mean_diurnal_range_gowers +
                                temp_seasonality_gowers +temp_ann_range_gowers +
                                tmean_wettest_quarter_gowers +
                                tmean_driest_quarter_gowers +ann_ppt_gowers +
                                ppt_seasonality_gowers + ppt_warmest_quarter_gowers +
                                ppt_coldest_quarter_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Gowers_Dist)
  
 gowers_calc_per_pop %>% pull(Gowers_Dist) #make the result a vector 
   }

#gowers_calc(recent_clim_boot_nest, P=15) #the function works
```

#### Perform the bootstrap sampling 
Recent
```{r}
gowers.boot_recent <- boot(data=recent_clim_boot_nest, statistic=gowers_calc, R=1000, P=15) #will sample each row (year) with replacement 
gowers.boot_recent$t0 #looks correct 
#str(gowers.boot_recent)

for(i in 1:23) {
  plot(gowers.boot_recent, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_recent, type="norm", index = 1) # for the first pop
boot_recent_results <- tidy(gowers.boot_recent,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Recent")
```

Historical
```{r}
gowers.boot_historical <- boot(data=historical_clim_boot_nest, statistic=gowers_calc, R=1000, P=15) #will sample each row (year) with replacement 
gowers.boot_historical$t0 #looks correct 
#str(gowers.boot_historical)

for(i in 1:23) {
  plot(gowers.boot_historical, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_historical, type="norm", index = 1) # for the first pop
boot_historical_results <- tidy(gowers.boot_historical,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Historical")
```

#### Plot the results 
```{r}
pops <- recent_clim_boot %>% select(parent.pop, elevation.group, elev_m, Lat, Long) %>% unique()

boot_gowers_recent_pops <- bind_cols(pops, boot_recent_results) %>% arrange(Gowers_Dist)
boot_gowers_historical_pops <- bind_cols(pops, boot_historical_results) %>% arrange(Gowers_Dist)
boot_gowers_results_all <- bind_rows(boot_gowers_recent_pops, boot_gowers_historical_pops)
#write_csv(boot_gowers_results_all, "../output/Climate/full_year_GowersEnvtalDist_WL2_wtr_year.csv")

recent_fig <- boot_gowers_results_all %>% 
  filter(TimePd=="Recent") %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population", title = "Recent Climate") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

historical_fig <- boot_gowers_results_all %>% 
  filter(TimePd=="Historical") %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population", title="Historic Climate") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) 
  

plot_grid(historical_fig, recent_fig)
#ggsave("../output/Climate/full_year_Gowers_fromWL2_wtr_year.png", width = 24, height = 8, units = "in")
```

## Gower's Distance For 2024
```{r}
WL2_climate_all_2024 <- bind_cols(WL2_climate_2024_flint, wl2_bioclim_all_year_final_2024)

WL2_range_prep_2024 <- WL2_climate_all_2024 %>% 
  mutate(parent.pop="WL2_Garden") %>% 
  rename_with(~str_remove(., "_WL2"), everything())
```

### Bootstrapping with home climates with avgs up to 2023

#### Create the gower_calc function 
```{r}
#data <- recent_clim_boot_nest
#P=14

gowers_calc_2324 <- function(data, indices, P) { #function with all of the code necessary for calculating gowers distance 
  #data = _clim_boot (recent or historical) - needs to be nested by year; P = # climate variables 

  #need to make WL2_range_prep_2024 before running this function 
  
  data <-data[indices,] # subset per bootstrap indices
  
  data <- data %>% unnest(data) #unnest so the function can access the climate data
  
  data_means <- data %>% 
    group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
    summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx", "ann_tmean", "mean_diurnal_range", 
                   "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_quarter", "tmean_driest_quarter", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_quarter", "ppt_coldest_quarter"),
               c(mean), na.rm = TRUE) #get 30 year averages for each climate variable 
  
  range_merge <- bind_rows(data_means, WL2_range_prep_2024)
  
  WL2_home_climate_ranges <- range_merge %>% #calculate ranges
    ungroup() %>% 
  summarise(cwd_range=max(cwd)-min(cwd),
            pck_range=max(pck)-min(pck),
            ppt_range=max(ppt)-min(ppt), 
            tmn_range=max(tmn)-min(tmn), 
            tmx_range=max(tmx)-min(tmx), 
            ann_tmean_range=max(ann_tmean)-min(ann_tmean),
            mean_diurnal_range_range=max(mean_diurnal_range)-min(mean_diurnal_range),
            temp_seasonality_range=max(temp_seasonality)-min(temp_seasonality),
            temp_ann_range_range=max(temp_ann_range)-min(temp_ann_range),
            tmean_wettest_quarter_range=max(tmean_wettest_quarter)-min(tmean_wettest_quarter),
            tmean_driest_quarter_range=max(tmean_driest_quarter)-min(tmean_driest_quarter),
            ann_ppt_range=max(ann_ppt)-min(ann_ppt), 
            ppt_seasonality_range=max(ppt_seasonality)-min(ppt_seasonality),
            ppt_warmest_quarter_range=max(ppt_warmest_quarter)-min(ppt_warmest_quarter), 
            ppt_coldest_quarter_range=max(ppt_coldest_quarter)-min(ppt_coldest_quarter))
  
  WL2_home_climate <- bind_cols(WL2_climate_all_2024, data_means) #add WL2 climate data to home climate data 
  
  WL2_home_climate_with_ranges <- bind_cols(WL2_home_climate, WL2_home_climate_ranges) #add in ranges 
  
  gowers_calc_each_var <- WL2_home_climate_with_ranges %>% #variable by variable calc
  mutate(cwd_gowers=abs(cwd_WL2-cwd) / cwd_range,
         pck_gowers=abs(pck_WL2-pck) / pck_range,
         ppt_gowers=abs(ppt_WL2 - ppt) / ppt_range,
         tmn_gowers=abs(tmn_WL2 - tmn) / tmn_range,
         tmx_gowers=abs(tmx_WL2 - tmx) / tmx_range,
         ann_tmean_gowers=abs(ann_tmean_WL2 - ann_tmean) / ann_tmean_range,
         mean_diurnal_range_gowers=abs(mean_diurnal_range_WL2 - mean_diurnal_range) / mean_diurnal_range_range,
         temp_seasonality_gowers=abs(temp_seasonality_WL2 - temp_seasonality) / temp_seasonality_range,
         temp_ann_range_gowers=abs(temp_ann_range_WL2 - temp_ann_range) / temp_ann_range_range,
         tmean_wettest_quarter_gowers=abs(tmean_wettest_quarter_WL2 - tmean_wettest_quarter) / tmean_wettest_quarter_range,
         tmean_driest_quarter_gowers=abs(tmean_driest_quarter_WL2 - tmean_driest_quarter) / tmean_driest_quarter_range,
         ann_ppt_gowers=abs(ann_ppt_WL2 - ann_ppt) / ann_ppt_range,
         ppt_seasonality_gowers=abs(ppt_seasonality_WL2 - ppt_seasonality) / ppt_seasonality_range,
         ppt_warmest_quarter_gowers=abs(ppt_warmest_quarter_WL2 - ppt_warmest_quarter) / ppt_warmest_quarter_range,
         ppt_coldest_quarter_gowers=abs(ppt_coldest_quarter_WL2 - ppt_coldest_quarter) / ppt_coldest_quarter_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))

 gowers_calc_per_pop <- gowers_calc_each_var %>% #final gowers calc 
  mutate(Gowers_Dist=(1/P)*(cwd_gowers + pck_gowers + ppt_gowers + tmn_gowers + tmx_gowers +
                                ann_tmean_gowers + mean_diurnal_range_gowers +
                                temp_seasonality_gowers +temp_ann_range_gowers +
                                tmean_wettest_quarter_gowers +
                                tmean_driest_quarter_gowers +ann_ppt_gowers +
                                ppt_seasonality_gowers + ppt_warmest_quarter_gowers +
                                ppt_coldest_quarter_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Gowers_Dist)
  
 gowers_calc_per_pop %>% pull(Gowers_Dist) #make the result a vector 
   }

#gowers_calc(recent_clim_boot_nest, P=15) #the function works
```

#### Perform the bootstrap sampling 
Recent
```{r}
gowers.boot_recent_2324 <- boot(data=recent_clim_boot_nest, statistic=gowers_calc_2324, R=1000, P=15) #will sample each row (year) with replacement 
gowers.boot_recent_2324$t0 #looks correct 
#str(gowers.boot_recent)

for(i in 1:23) {
  plot(gowers.boot_recent_2324, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_recent, type="norm", index = 1) # for the first pop
boot_recent_results_2324 <- tidy(gowers.boot_recent_2324,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Recent")
```

Historical
```{r}
gowers.boot_historical_2324 <- boot(data=historical_clim_boot_nest, statistic=gowers_calc_2324, R=1000, P=15) #will sample each row (year) with replacement 
gowers.boot_historical_2324$t0 #looks correct 
#str(gowers.boot_historical)

for(i in 1:23) {
  plot(gowers.boot_historical_2324, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_historical, type="norm", index = 1) # for the first pop
boot_historical_results_2324 <- tidy(gowers.boot_historical_2324,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Historical")
```

#### Plot the results 
```{r}
pops <- recent_clim_boot %>% select(parent.pop, elevation.group, elev_m, Lat, Long) %>% unique()

boot_gowers_recent_pops <- bind_cols(pops, boot_recent_results_2324) %>% arrange(Gowers_Dist)
boot_gowers_historical_pops <- bind_cols(pops, boot_historical_results_2324) %>% arrange(Gowers_Dist)
boot_gowers_results_2324_all <- bind_rows(boot_gowers_recent_pops, boot_gowers_historical_pops)
write_csv(boot_gowers_results_2324_all, "../output/Climate/full_year_GowersEnvtalDist_WL2_wtr_year_2324.csv")

recent_fig <- boot_gowers_results_2324_all %>% 
  filter(TimePd=="Recent") %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population", title = "Recent Climate") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

historical_fig <- boot_gowers_results_2324_all %>% 
  filter(TimePd=="Historical") %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population", title="Historic Climate") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) 
  

plot_grid(historical_fig, recent_fig)
ggsave("../output/Climate/full_year_Gowers_fromWL2_wtr_year_2324.png", width = 24, height = 8, units = "in")
```

### Bootstrapping with home climates with avgs up to 2024

#### Load data with all 30 years
```{r}
recent_clim_boot_2024 <- read_csv("../output/Climate/fullyear_wtr_year_avgs_Recent_2024.csv")

recent_clim_boot_2024_nest <- recent_clim_boot_2024 %>% nest(.by=year) #nest to prepare for bootstrapping 
recent_clim_boot_2024_nest

historical_clim_boot_2024 <- read_csv("../output/Climate/fullyear_wtr_year_avgs_Historical_2024.csv")

historical_clim_boot_2024_nest <- historical_clim_boot_2024 %>% nest(.by=year)
historical_clim_boot_2024_nest
```

#### Create the gower_calc function 
```{r}
#data <- recent_clim_boot_2024_nest
#P=14

gowers_calc_2024 <- function(data, indices, P) { #function with all of the code necessary for calculating gowers distance 
  #data = _clim_boot_2024 (recent or historical) - needs to be nested by year; P = # climate variables 

  #need to make WL2_range_prep_2024 before running this function 
  
  data <-data[indices,] # subset per bootstrap indices
  
  data <- data %>% unnest(data) #unnest so the function can access the climate data
  
  data_means <- data %>% 
    group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
    summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx", "ann_tmean", "mean_diurnal_range", 
                   "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_quarter", "tmean_driest_quarter", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_quarter", "ppt_coldest_quarter"),
               c(mean), na.rm = TRUE) #get 30 year averages for each climate variable 
  
  range_merge <- bind_rows(data_means, WL2_range_prep_2024)
  
  WL2_home_climate_ranges <- range_merge %>% #calculate ranges
    ungroup() %>% 
  summarise(cwd_range=max(cwd)-min(cwd),
            pck_range=max(pck)-min(pck),
            ppt_range=max(ppt)-min(ppt), 
            tmn_range=max(tmn)-min(tmn), 
            tmx_range=max(tmx)-min(tmx), 
            ann_tmean_range=max(ann_tmean)-min(ann_tmean),
            mean_diurnal_range_range=max(mean_diurnal_range)-min(mean_diurnal_range),
            temp_seasonality_range=max(temp_seasonality)-min(temp_seasonality),
            temp_ann_range_range=max(temp_ann_range)-min(temp_ann_range),
            tmean_wettest_quarter_range=max(tmean_wettest_quarter)-min(tmean_wettest_quarter),
            tmean_driest_quarter_range=max(tmean_driest_quarter)-min(tmean_driest_quarter),
            ann_ppt_range=max(ann_ppt)-min(ann_ppt), 
            ppt_seasonality_range=max(ppt_seasonality)-min(ppt_seasonality),
            ppt_warmest_quarter_range=max(ppt_warmest_quarter)-min(ppt_warmest_quarter), 
            ppt_coldest_quarter_range=max(ppt_coldest_quarter)-min(ppt_coldest_quarter))
  
  WL2_home_climate <- bind_cols(WL2_climate_all_2024, data_means) #add WL2 climate data to home climate data 
  
  WL2_home_climate_with_ranges <- bind_cols(WL2_home_climate, WL2_home_climate_ranges) #add in ranges 
  
  gowers_calc_each_var <- WL2_home_climate_with_ranges %>% #variable by variable calc
  mutate(cwd_gowers=abs(cwd_WL2-cwd) / cwd_range,
         pck_gowers=abs(pck_WL2-pck) / pck_range,
         ppt_gowers=abs(ppt_WL2 - ppt) / ppt_range,
         tmn_gowers=abs(tmn_WL2 - tmn) / tmn_range,
         tmx_gowers=abs(tmx_WL2 - tmx) / tmx_range,
         ann_tmean_gowers=abs(ann_tmean_WL2 - ann_tmean) / ann_tmean_range,
         mean_diurnal_range_gowers=abs(mean_diurnal_range_WL2 - mean_diurnal_range) / mean_diurnal_range_range,
         temp_seasonality_gowers=abs(temp_seasonality_WL2 - temp_seasonality) / temp_seasonality_range,
         temp_ann_range_gowers=abs(temp_ann_range_WL2 - temp_ann_range) / temp_ann_range_range,
         tmean_wettest_quarter_gowers=abs(tmean_wettest_quarter_WL2 - tmean_wettest_quarter) / tmean_wettest_quarter_range,
         tmean_driest_quarter_gowers=abs(tmean_driest_quarter_WL2 - tmean_driest_quarter) / tmean_driest_quarter_range,
         ann_ppt_gowers=abs(ann_ppt_WL2 - ann_ppt) / ann_ppt_range,
         ppt_seasonality_gowers=abs(ppt_seasonality_WL2 - ppt_seasonality) / ppt_seasonality_range,
         ppt_warmest_quarter_gowers=abs(ppt_warmest_quarter_WL2 - ppt_warmest_quarter) / ppt_warmest_quarter_range,
         ppt_coldest_quarter_gowers=abs(ppt_coldest_quarter_WL2 - ppt_coldest_quarter) / ppt_coldest_quarter_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))

 gowers_calc_per_pop <- gowers_calc_each_var %>% #final gowers calc 
  mutate(Gowers_Dist=(1/P)*(cwd_gowers + pck_gowers + ppt_gowers + tmn_gowers + tmx_gowers +
                                ann_tmean_gowers + mean_diurnal_range_gowers +
                                temp_seasonality_gowers +temp_ann_range_gowers +
                                tmean_wettest_quarter_gowers +
                                tmean_driest_quarter_gowers +ann_ppt_gowers +
                                ppt_seasonality_gowers + ppt_warmest_quarter_gowers +
                                ppt_coldest_quarter_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Gowers_Dist)
  
 gowers_calc_per_pop %>% pull(Gowers_Dist) #make the result a vector 
   }

#gowers_calc_2024(recent_clim_boot_2024_nest, P=15) #the function works
```

#### Perform the bootstrap sampling 
Recent
```{r}
gowers.boot_recent_2024 <- boot(data=recent_clim_boot_2024_nest, statistic=gowers_calc_2024, R=1000, P=15) #will sample each row (year) with replacement 
gowers.boot_recent_2024$t0 #looks correct 
#str(gowers.boot_recent)

for(i in 1:23) {
  plot(gowers.boot_recent_2024, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_recent, type="norm", index = 1) # for the first pop
boot_recent_results_2024 <- tidy(gowers.boot_recent_2024,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Recent")
```

Historical
```{r}
gowers.boot_historical_2024 <- boot(data=historical_clim_boot_2024_nest, statistic=gowers_calc_2024, R=1000, P=15) #will sample each row (year) with replacement 
gowers.boot_historical_2024$t0 #looks correct 
#str(gowers.boot_historical)

for(i in 1:23) {
  plot(gowers.boot_historical_2024, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_historical, type="norm", index = 1) # for the first pop
boot_historical_results_2024 <- tidy(gowers.boot_historical_2024,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Historical")
```

#### Plot the results 
```{r}
pops <- recent_clim_boot_2024 %>% select(parent.pop, elevation.group, elev_m, Lat, Long) %>% unique()

boot_gowers_recent_pops_2024 <- bind_cols(pops, boot_recent_results_2024) %>% arrange(Gowers_Dist)
boot_gowers_historical_pops_2024 <- bind_cols(pops, boot_historical_results_2024) %>% arrange(Gowers_Dist)
boot_gowers_results_all_2024 <- bind_rows(boot_gowers_recent_pops_2024, boot_gowers_historical_pops_2024)
#write_csv(boot_gowers_results_all_2024, "../output/Climate/full_year_GowersEnvtalDist_WL2_wtr_year_2024.csv")

recent_fig <- boot_gowers_results_all_2024 %>% 
  filter(TimePd=="Recent") %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population", title = "Recent Climate") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

historical_fig <- boot_gowers_results_all_2024 %>% 
  filter(TimePd=="Historical") %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population", title="Historic Climate") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) 
  

plot_grid(historical_fig, recent_fig)
#ggsave("../output/Climate/full_year_Gowers_fromWL2_wtr_year_2024.png", width = 24, height = 8, units = "in")
```

## Flint Climate Distance

### Gowers
```{r}
#need to modify the gowers_calc function to only include Flint variables 
gowers_calc_flint <- function(data, indices, P) { #function with all of the code necessary for calculating gowers distance 
  #data = _clim_boot (recent or historical) - needs to be nested by year; P = # climate variables 
  #need to make WL2_range_prep before running this function 
  
  data <-data[indices,] # subset per bootstrap indices
  
  data <- data %>% unnest(data) #unnest so the function can access the climate data
  
  data_means <- data %>% 
    group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
    summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx"),
               c(mean), na.rm = TRUE) #get 30 year averages for each climate variable 
  
  range_merge <- bind_rows(data_means, WL2_range_prep)
  
  WL2_home_climate_ranges <- range_merge %>% #calculate ranges
    ungroup() %>% 
  summarise(cwd_range=max(cwd)-min(cwd),
            pck_range=max(pck)-min(pck),
            ppt_range=max(ppt)-min(ppt), 
            tmn_range=max(tmn)-min(tmn), 
            tmx_range=max(tmx)-min(tmx))
  
  WL2_home_climate <- bind_cols(WL2_climate_all, data_means) #add WL2 climate data to home climate data 
  
  WL2_home_climate_with_ranges <- bind_cols(WL2_home_climate, WL2_home_climate_ranges) #add in ranges 
  
  gowers_calc_each_var <- WL2_home_climate_with_ranges %>% #variable by variable calc
  mutate(cwd_gowers=abs(cwd_WL2-cwd) / cwd_range,
         pck_gowers=abs(pck_WL2-pck) / pck_range,
         ppt_gowers=abs(ppt_WL2 - ppt) / ppt_range,
         tmn_gowers=abs(tmn_WL2 - tmn) / tmn_range,
         tmx_gowers=abs(tmx_WL2 - tmx) / tmx_range)

 gowers_calc_per_pop <- gowers_calc_each_var %>% #final gowers calc 
  mutate(Gowers_Dist=(1/P)*(cwd_gowers + pck_gowers + ppt_gowers + tmn_gowers + tmx_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Gowers_Dist)
 
 gowers_calc_per_pop %>% pull(Gowers_Dist) #make the result a vector 
   }
```

#### Recent

```{r}
recent_clim_boot_flint_nest <- recent_clim_boot %>% select(parent.pop:tmx) %>% nest(.by=year)

gowers.boot_flint_recent <- boot(data=recent_clim_boot_flint_nest, statistic=gowers_calc_flint, R=1000, P=5) #will sample each row (year) with replacement 
gowers.boot_flint_recent$t0 #looks correct 
#str(gowers.boot_flint_recent)

for(i in 1:23) {
  plot(gowers.boot_flint_recent, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_flint_recent, type="norm", index = 1) # for the first pop
boot_flint_recent_results <- tidy(gowers.boot_flint_recent,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Recent")
```

#### Historic

```{r}
historical_clim_boot_flint_nest <- historical_clim_boot %>% select(parent.pop:tmx) %>% nest(.by=year)

gowers.boot_flint_historical <- boot(data=historical_clim_boot_flint_nest, statistic=gowers_calc_flint, R=1000, P=5) #will sample each row (year) with replacement 
gowers.boot_flint_historical$t0 #looks correct 
#str(gowers.boot_flint_historical)

for(i in 1:23) {
  plot(gowers.boot_flint_historical, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_flint_historical, type="norm", index = 1) # for the first pop
boot_flint_historical_results <- tidy(gowers.boot_flint_historical,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Historical")
```

#### Plot the results 

```{r}
boot_flint_gowers_recent_pops <- bind_cols(pops, boot_flint_recent_results) %>% arrange(Gowers_Dist)
boot_flint_gowers_historical_pops <- bind_cols(pops, boot_flint_historical_results) %>% arrange(Gowers_Dist)
boot_flint_gowers_results_all <- bind_rows(boot_flint_gowers_recent_pops, boot_flint_gowers_historical_pops)
#write_csv(boot_flint_gowers_results_all, "../output/Climate/full_year_GowersEnvtalDist_WL2Flint_wtr_year.csv")

recent_fig <- boot_flint_gowers_results_all %>% 
  filter(TimePd=="Recent") %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population", title = "Recent Climate") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

historical_fig <- boot_flint_gowers_results_all %>% 
  filter(TimePd=="Historical") %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population", title="Historic Climate") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) 
  

plot_grid(historical_fig, recent_fig)
#ggsave("../output/Climate/full_year_Gowers_Flint_fromWL2_wtr_year.png", width = 24, height = 8, units = "in")
```

### Subtraction

```{r}
pops_flint_avgs <- read_csv("../output/Climate/fullyear_FlintAvgs_wtr_year.csv")

pops_flint_recent_avgs <- pops_flint_avgs %>% filter(TimePd=="Recent")
recent_flint_dist_prep <- bind_cols(WL2_climate_flint, pops_flint_recent_avgs)
names(recent_flint_dist_prep)

recent_flint_dist <- recent_flint_dist_prep %>% 
  mutate(ppt_dist=ppt - ppt_WL2,
         cwd_dist=cwd - cwd_WL2,
         pck_dist=pck - pck_WL2,
         tmn_dist=tmn - tmn_WL2,
         tmx_dist=tmx - tmx_WL2) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))

pops_flint_historic_avgs <-  pops_flint_avgs %>% filter(TimePd=="Historical")
historic_flint_dist_prep <- bind_cols(WL2_climate_flint, pops_flint_historic_avgs)
names(historic_flint_dist_prep)

historic_flint_dist <- historic_flint_dist_prep %>% 
  mutate(ppt_dist=ppt - ppt_WL2,
         cwd_dist=cwd - cwd_WL2,
         pck_dist=pck - pck_WL2,
         tmn_dist=tmn - tmn_WL2,
         tmx_dist=tmx - tmx_WL2) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
```

Figures Recent (subtraction distance)

```{r}
cwd_dist_fig_recent <- recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, cwd_dist), y=cwd_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanCWD_DistfromHome_WL2_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

pck_dist_fig_recent <- recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, pck_dist), y=pck_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanPCK_DistfromHome_WL2_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

ppt_dist_fig_recent <- recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_dist), y=ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanPPT_DistfromHome_WL2_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

tmn_dist_fig_recent <- recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmn_dist), y=tmn_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanTMN_DistfromHome_WL2_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

tmx_dist_fig_recent <- recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmx_dist), y=tmx_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanTMX_DistfromHome_WL2_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")
```

Figures Historical (subtraction distance)

```{r}
cwd_dist_fig_historical <- historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, cwd_dist), y=cwd_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanCWD_DistfromHome_WL2_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

pck_dist_fig_historical <- historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, pck_dist), y=pck_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanPCK_DistfromHome_WL2_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

ppt_dist_fig_historical <- historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_dist), y=ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanPPT_DistfromHome_WL2_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

tmn_dist_fig_historical <- historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmn_dist), y=tmn_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanTMN_DistfromHome_WL2_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

tmx_dist_fig_historical <- historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmx_dist), y=tmx_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_MeanTMX_DistfromHome_WL2_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")
```

#### For 2024 (from 2023 home avgs)

```{r}
recent_flint_dist_prep_2324 <- bind_cols(WL2_climate_2024_flint, pops_flint_recent_avgs)
names(recent_flint_dist_prep_2324)

recent_flint_dist_2324 <- recent_flint_dist_prep_2324 %>% 
  mutate(ppt_dist=ppt - ppt_WL2,
         cwd_dist=cwd - cwd_WL2,
         pck_dist=pck - pck_WL2,
         tmn_dist=tmn - tmn_WL2,
         tmx_dist=tmx - tmx_WL2) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))

historic_flint_dist_prep_2324 <- bind_cols(WL2_climate_2024_flint, pops_flint_historic_avgs)
names(historic_flint_dist_prep_2324)

historic_flint_dist_2324 <- historic_flint_dist_prep_2324 %>% 
  mutate(ppt_dist=ppt - ppt_WL2,
         cwd_dist=cwd - cwd_WL2,
         pck_dist=pck - pck_WL2,
         tmn_dist=tmn - tmn_WL2,
         tmx_dist=tmx - tmx_WL2) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
```

#### For 2024
```{r}
pops_flint_avgs_2024 <- read_csv("../output/Climate/fullyear_FlintAvgs_wtr_year_2024.csv")

pops_flint_avgs_recent_2024 <- pops_flint_avgs_2024 %>% filter(TimePd=="Recent")
recent_flint_dist_prep_2024 <- bind_cols(WL2_climate_2024_flint, pops_flint_avgs_recent_2024)
names(recent_flint_dist_prep_2024)

recent_flint_dist_2024 <- recent_flint_dist_prep_2024 %>% 
  mutate(ppt_dist=ppt - ppt_WL2,
         cwd_dist=cwd - cwd_WL2,
         pck_dist=pck - pck_WL2,
         tmn_dist=tmn - tmn_WL2,
         tmx_dist=tmx - tmx_WL2) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))

pops_flint_avgs_historic_2024 <- pops_flint_avgs_2024 %>% filter(TimePd=="Historical")
historic_flint_dist_prep_2024 <- bind_cols(WL2_climate_2024_flint, pops_flint_avgs_historic_2024)
names(historic_flint_dist_prep_2024)

historic_flint_dist_2024 <- historic_flint_dist_prep_2024 %>% 
  mutate(ppt_dist=ppt - ppt_WL2,
         cwd_dist=cwd - cwd_WL2,
         pck_dist=pck - pck_WL2,
         tmn_dist=tmn - tmn_WL2,
         tmx_dist=tmx - tmx_WL2) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
```


## Bioclim Climate Distance

### Gowers
```{r}
#need to modify the gowers_calc function to only include bioclim variables 
gowers_calc_bioclim <- function(data, indices, P) { #function with all of the code necessary for calculating gowers distance 
  #data = _clim_boot (recent or historical) - needs to be nested by year; P = # climate variables 
  #need to make WL2_range_prep before running this function 
  
  data <-data[indices,] # subset per bootstrap indices
  
  data <- data %>% unnest(data) #unnest so the function can access the climate data
  
  data_means <- data %>% 
    group_by(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
    summarise_at(c("ann_tmean", "mean_diurnal_range", 
                   "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_quarter", "tmean_driest_quarter", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_quarter", "ppt_coldest_quarter"),
               c(mean), na.rm = TRUE) #get 30 year averages for each climate variable 
  
  range_merge <- bind_rows(data_means, WL2_range_prep)
  
  WL2_home_climate_ranges <- range_merge %>% #calculate ranges
    ungroup() %>% 
  summarise(ann_tmean_range=max(ann_tmean)-min(ann_tmean),
            mean_diurnal_range_range=max(mean_diurnal_range)-min(mean_diurnal_range),
            temp_seasonality_range=max(temp_seasonality)-min(temp_seasonality),
            temp_ann_range_range=max(temp_ann_range)-min(temp_ann_range),
            tmean_wettest_quarter_range=max(tmean_wettest_quarter)-min(tmean_wettest_quarter),
            tmean_driest_quarter_range=max(tmean_driest_quarter)-min(tmean_driest_quarter),
            ann_ppt_range=max(ann_ppt)-min(ann_ppt), 
            ppt_seasonality_range=max(ppt_seasonality)-min(ppt_seasonality),
            ppt_warmest_quarter_range=max(ppt_warmest_quarter)-min(ppt_warmest_quarter), 
            ppt_coldest_quarter_range=max(ppt_coldest_quarter)-min(ppt_coldest_quarter))
  
  WL2_home_climate <- bind_cols(WL2_climate_all, data_means) #add WL2 climate data to home climate data 
  
  WL2_home_climate_with_ranges <- bind_cols(WL2_home_climate, WL2_home_climate_ranges) #add in ranges 
  
  gowers_calc_each_var <- WL2_home_climate_with_ranges %>% #variable by variable calc
  mutate(ann_tmean_gowers=abs(ann_tmean_WL2 - ann_tmean) / ann_tmean_range,
         mean_diurnal_range_gowers=abs(mean_diurnal_range_WL2 - mean_diurnal_range) / mean_diurnal_range_range,
         temp_seasonality_gowers=abs(temp_seasonality_WL2 - temp_seasonality) / temp_seasonality_range,
         temp_ann_range_gowers=abs(temp_ann_range_WL2 - temp_ann_range) / temp_ann_range_range,
         tmean_wettest_quarter_gowers=abs(tmean_wettest_quarter_WL2 - tmean_wettest_quarter) / tmean_wettest_quarter_range,
         tmean_driest_quarter_gowers=abs(tmean_driest_quarter_WL2 - tmean_driest_quarter) / tmean_driest_quarter_range,
         ann_ppt_gowers=abs(ann_ppt_WL2 - ann_ppt) / ann_ppt_range,
         ppt_seasonality_gowers=abs(ppt_seasonality_WL2 - ppt_seasonality) / ppt_seasonality_range,
         ppt_warmest_quarter_gowers=abs(ppt_warmest_quarter_WL2 - ppt_warmest_quarter) / ppt_warmest_quarter_range,
         ppt_coldest_quarter_gowers=abs(ppt_coldest_quarter_WL2 - ppt_coldest_quarter) / ppt_coldest_quarter_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))

 gowers_calc_per_pop <- gowers_calc_each_var %>% #final gowers calc 
  mutate(Gowers_Dist=(1/P)*(ann_tmean_gowers + mean_diurnal_range_gowers +
                                temp_seasonality_gowers +temp_ann_range_gowers +
                                tmean_wettest_quarter_gowers +
                                tmean_driest_quarter_gowers +ann_ppt_gowers +
                                ppt_seasonality_gowers + ppt_warmest_quarter_gowers +
                                ppt_coldest_quarter_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Gowers_Dist)
 
 gowers_calc_per_pop %>% pull(Gowers_Dist) #make the result a vector 
   }
```

#### Recent

```{r}
recent_clim_boot_bioclim_nest <- recent_clim_boot %>% select(parent.pop:year, ann_tmean:ppt_coldest_quarter) %>% nest(.by=year)

gowers.boot_bioclim_recent <- boot(data=recent_clim_boot_bioclim_nest, statistic=gowers_calc_bioclim, R=1000, P=10) #will sample each row (year) with replacement 
gowers.boot_bioclim_recent$t0 #looks correct 
#str(gowers.boot_bioclim_recent)

for(i in 1:23) {
  plot(gowers.boot_bioclim_recent, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_bioclim_recent, type="norm", index = 1) # for the first pop
boot_bioclim_recent_results <- tidy(gowers.boot_bioclim_recent,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Recent")
```

#### Historic

```{r}
historical_clim_boot_bioclim_nest <- historical_clim_boot %>% select(parent.pop:year, ann_tmean:ppt_coldest_quarter) %>% nest(.by=year)

gowers.boot_bioclim_historical <- boot(data=historical_clim_boot_bioclim_nest, statistic=gowers_calc_bioclim, R=1000, P=10) #will sample each row (year) with replacement 
gowers.boot_bioclim_historical$t0 #looks correct 
#str(gowers.boot_bioclim_historical)

for(i in 1:23) {
  plot(gowers.boot_bioclim_historical, index=i) #distributions look normal for the most part 
}

#boot.ci(gowers.boot_bioclim_historical, type="norm", index = 1) # for the first pop
boot_bioclim_historical_results <- tidy(gowers.boot_bioclim_historical,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(TimePd="Historical")
```

#### Plot the results 

```{r}
boot_bioclim_gowers_recent_pops <- bind_cols(pops, boot_bioclim_recent_results) %>% arrange(Gowers_Dist)
boot_bioclim_gowers_historical_pops <- bind_cols(pops, boot_bioclim_historical_results) %>% arrange(Gowers_Dist)
boot_bioclim_gowers_results_all <- bind_rows(boot_bioclim_gowers_recent_pops, boot_bioclim_gowers_historical_pops)
#write_csv(boot_bioclim_gowers_results_all, "../output/Climate/full_year_GowersEnvtalDist_WL2bioclim_wtr_year.csv")

recent_fig <- boot_bioclim_gowers_results_all %>% 
  filter(TimePd=="Recent") %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population", title = "Recent Climate") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

historical_fig <- boot_bioclim_gowers_results_all %>% 
  filter(TimePd=="Historical") %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population", title="Historic Climate") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) 
  

plot_grid(historical_fig, recent_fig)
#ggsave("../output/Climate/full_year_Gowers_BioClim_fromWL2_wtr_year.png", width = 24, height = 8, units = "in")
```

### Subtraction

```{r}
pops_bioclim_avgs <-  read_csv("../output/Climate/fullyear_BioClimAvgs_wtr_year.csv") 

#Recent
pops_bioclim_recent_avgs <- pops_bioclim_avgs %>%  filter(TimePd=="Recent")
recent_bioclim_dist_prep <- bind_cols(wl2_bioclim_all_year_final, pops_bioclim_recent_avgs)
recent_bioclim_dist <- recent_bioclim_dist_prep %>% 
  mutate(ann_tmean_dist=ann_tmean - ann_tmean_WL2,
         mean_diurnal_range_dist=mean_diurnal_range - mean_diurnal_range_WL2,
         temp_seasonality_dist=temp_seasonality - temp_seasonality_WL2,
         temp_ann_range_dist=temp_ann_range - temp_ann_range_WL2,
         tmean_wettest_quarter_dist=tmean_wettest_quarter - tmean_wettest_quarter_WL2,
         tmean_driest_quarter_dist=tmean_driest_quarter - tmean_driest_quarter_WL2,
         ann_ppt_dist=ann_ppt - ann_ppt_WL2,
         ppt_seasonality_dist=ppt_seasonality - ppt_seasonality_WL2, 
         ppt_warmest_quarter_dist=ppt_warmest_quarter - ppt_warmest_quarter_WL2,
         ppt_coldest_quarter_dist=ppt_coldest_quarter - ppt_coldest_quarter_WL2) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
recent_bioclim_dist

#Historical
pops_bioclim_historical_avgs <- pops_bioclim_avgs %>% filter(TimePd=="Historical")
historical_bioclim_dist_prep <- bind_cols(wl2_bioclim_all_year_final, pops_bioclim_historical_avgs)
historical_bioclim_dist <- historical_bioclim_dist_prep %>% 
  mutate(ann_tmean_dist=ann_tmean - ann_tmean_WL2,
         mean_diurnal_range_dist=mean_diurnal_range - mean_diurnal_range_WL2,
         temp_seasonality_dist=temp_seasonality - temp_seasonality_WL2,
         temp_ann_range_dist=temp_ann_range - temp_ann_range_WL2,
         tmean_wettest_quarter_dist=tmean_wettest_quarter - tmean_wettest_quarter_WL2,
         tmean_driest_quarter_dist=tmean_driest_quarter - tmean_driest_quarter_WL2,
         ann_ppt_dist=ann_ppt - ann_ppt_WL2,
         ppt_seasonality_dist=ppt_seasonality - ppt_seasonality_WL2, 
         ppt_warmest_quarter_dist=ppt_warmest_quarter - ppt_warmest_quarter_WL2,
         ppt_coldest_quarter_dist=ppt_coldest_quarter - ppt_coldest_quarter_WL2) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
historical_bioclim_dist
```

Figures Recent (subtraction distance)

```{r}
ann_tmean_dist_fig_recent <- recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_dist), y=ann_tmean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Ann_Tmean_DistfromHome_WL2_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

diurnal_range_dist_fig_recent <- recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_dist), y=mean_diurnal_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Diurnal_Range_DistfromHome_WL2_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

tmp_seasonality_dist_fig_recent <- recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_dist), y=temp_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Seasonality_DistfromHome_WL2_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

tmp_ann_range_dist_fig_recent <- recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_dist), y=temp_ann_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Ann_Range_DistfromHome_WL2_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

tmean_wet_dist_fig_recent <- recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_quarter_dist), y=tmean_wettest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Wet_DistfromHome_WL2_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

tmean_dry_dist_fig_recent <- recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_quarter_dist), y=tmean_driest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Dry_DistfromHome_WL2_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

ann_ppt_dist_fig_recent <- recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_dist), y=ann_ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Ann_PPT_DistfromHome_WL2_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

ppt_seasonality_dist_fig_recent <- recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_dist), y=ppt_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_PPT_Seasonality_DistfromHome_WL2_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

ppt_warm_dist_fig_recent <- recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_quarter_dist), y=ppt_warmest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_PPT_Warm_DistfromHome_WL2_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")

ppt_cold_dist_fig_recent <- recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_quarter_dist), y=ppt_coldest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_PPT_Cold_DistfromHome_WL2_RecentClim_wtr_year.png", width = 12, height = 6, units = "in")
```

Historical (subtraction distance)

```{r}
ann_tmean_dist_fig_historical <- historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_dist), y=ann_tmean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Ann_Tmean_DistfromHome_WL2_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

diurnal_range_dist_fig_historical <- historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_dist), y=mean_diurnal_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Diurnal_Range_DistfromHome_WL2_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

tmp_seasonality_dist_fig_historical <- historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_dist), y=temp_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Seasonality_DistfromHome_WL2_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

tmp_ann_range_dist_fig_historical <- historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_dist), y=temp_ann_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Ann_Range_DistfromHome_WL2_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

tmean_wet_dist_fig_historical <- historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_quarter_dist), y=tmean_wettest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Wet_DistfromHome_WL2_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

tmean_dry_dist_fig_historical <- historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_quarter_dist), y=tmean_driest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Temp_Dry_DistfromHome_WL2_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

ann_ppt_dist_fig_historical <- historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_dist), y=ann_ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_Ann_PPT_DistfromHome_WL2_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

ppt_seasonality_dist_fig_historical <- historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_dist), y=ppt_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_PPT_Seasonality_DistfromHome_WL2_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

ppt_warm_dist_fig_historical <- historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_quarter_dist), y=ppt_warmest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_PPT_Warm_DistfromHome_WL2_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")

ppt_cold_dist_fig_historical <- historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_quarter_dist), y=ppt_coldest_quarter_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/Climate/all-year_PPT_Cold_DistfromHome_WL2_HistoricalClim_wtr_year.png", width = 12, height = 6, units = "in")
```

#### For 2024 (from 2023 home avgs)

```{r}
#Recent
recent_bioclim_dist_prep_2324 <- bind_cols(wl2_bioclim_all_year_final_2024, pops_bioclim_recent_avgs) 
recent_bioclim_dist_2324 <- recent_bioclim_dist_prep_2324 %>% 
  mutate(ann_tmean_dist=ann_tmean - ann_tmean_WL2,
         mean_diurnal_range_dist=mean_diurnal_range - mean_diurnal_range_WL2,
         temp_seasonality_dist=temp_seasonality - temp_seasonality_WL2,
         temp_ann_range_dist=temp_ann_range - temp_ann_range_WL2,
         tmean_wettest_quarter_dist=tmean_wettest_quarter - tmean_wettest_quarter_WL2,
         tmean_driest_quarter_dist=tmean_driest_quarter - tmean_driest_quarter_WL2,
         ann_ppt_dist=ann_ppt - ann_ppt_WL2,
         ppt_seasonality_dist=ppt_seasonality - ppt_seasonality_WL2, 
         ppt_warmest_quarter_dist=ppt_warmest_quarter - ppt_warmest_quarter_WL2,
         ppt_coldest_quarter_dist=ppt_coldest_quarter - ppt_coldest_quarter_WL2) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
#recent_bioclim_dist_2324

#Historical
historical_bioclim_dist_prep_2324 <- bind_cols(wl2_bioclim_all_year_final_2024, pops_bioclim_historical_avgs)  
historical_bioclim_dist_2324 <- historical_bioclim_dist_prep_2324 %>% 
  mutate(ann_tmean_dist=ann_tmean - ann_tmean_WL2,
         mean_diurnal_range_dist=mean_diurnal_range - mean_diurnal_range_WL2,
         temp_seasonality_dist=temp_seasonality - temp_seasonality_WL2,
         temp_ann_range_dist=temp_ann_range - temp_ann_range_WL2,
         tmean_wettest_quarter_dist=tmean_wettest_quarter - tmean_wettest_quarter_WL2,
         tmean_driest_quarter_dist=tmean_driest_quarter - tmean_driest_quarter_WL2,
         ann_ppt_dist=ann_ppt - ann_ppt_WL2,
         ppt_seasonality_dist=ppt_seasonality - ppt_seasonality_WL2, 
         ppt_warmest_quarter_dist=ppt_warmest_quarter - ppt_warmest_quarter_WL2,
         ppt_coldest_quarter_dist=ppt_coldest_quarter - ppt_coldest_quarter_WL2) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
#historical_bioclim_dist_2324
```


#### For 2024
```{r}
pops_bioclim_avgs_2024 <-  read_csv("../output/Climate/fullyear_BioClimAvgs_wtr_year_2024.csv") 

#Recent
pops_bioclim_recent_avgs_2024 <- pops_bioclim_avgs_2024 %>%  filter(TimePd=="Recent")
recent_bioclim_dist_prep_2024 <- bind_cols(wl2_bioclim_all_year_final_2024, pops_bioclim_recent_avgs_2024)
recent_bioclim_dist_2024 <- recent_bioclim_dist_prep_2024 %>% 
  mutate(ann_tmean_dist=ann_tmean - ann_tmean_WL2,
         mean_diurnal_range_dist=mean_diurnal_range - mean_diurnal_range_WL2,
         temp_seasonality_dist=temp_seasonality - temp_seasonality_WL2,
         temp_ann_range_dist=temp_ann_range - temp_ann_range_WL2,
         tmean_wettest_quarter_dist=tmean_wettest_quarter - tmean_wettest_quarter_WL2,
         tmean_driest_quarter_dist=tmean_driest_quarter - tmean_driest_quarter_WL2,
         ann_ppt_dist=ann_ppt - ann_ppt_WL2,
         ppt_seasonality_dist=ppt_seasonality - ppt_seasonality_WL2, 
         ppt_warmest_quarter_dist=ppt_warmest_quarter - ppt_warmest_quarter_WL2,
         ppt_coldest_quarter_dist=ppt_coldest_quarter - ppt_coldest_quarter_WL2) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
#recent_bioclim_dist_2024

#Historical
pops_bioclim_historical_avgs_2024 <- pops_bioclim_avgs_2024 %>% filter(TimePd=="Historical")
historical_bioclim_dist_prep_2024 <- bind_cols(wl2_bioclim_all_year_final_2024, pops_bioclim_historical_avgs_2024)
historical_bioclim_dist_2024 <- historical_bioclim_dist_prep_2024 %>% 
  mutate(ann_tmean_dist=ann_tmean - ann_tmean_WL2,
         mean_diurnal_range_dist=mean_diurnal_range - mean_diurnal_range_WL2,
         temp_seasonality_dist=temp_seasonality - temp_seasonality_WL2,
         temp_ann_range_dist=temp_ann_range - temp_ann_range_WL2,
         tmean_wettest_quarter_dist=tmean_wettest_quarter - tmean_wettest_quarter_WL2,
         tmean_driest_quarter_dist=tmean_driest_quarter - tmean_driest_quarter_WL2,
         ann_ppt_dist=ann_ppt - ann_ppt_WL2,
         ppt_seasonality_dist=ppt_seasonality - ppt_seasonality_WL2, 
         ppt_warmest_quarter_dist=ppt_warmest_quarter - ppt_warmest_quarter_WL2,
         ppt_coldest_quarter_dist=ppt_coldest_quarter - ppt_coldest_quarter_WL2) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
#historical_bioclim_dist_2024
```


## Combine all Subtraction Figs
```{r}
ggarrange(cwd_dist_fig_recent, pck_dist_fig_recent, ppt_dist_fig_recent, tmn_dist_fig_recent, tmx_dist_fig_recent,
          ann_tmean_dist_fig_recent, diurnal_range_dist_fig_recent, tmp_seasonality_dist_fig_recent,
          tmp_ann_range_dist_fig_recent, tmean_wet_dist_fig_recent, tmean_dry_dist_fig_recent,
          ann_ppt_dist_fig_recent, ppt_seasonality_dist_fig_recent, ppt_warm_dist_fig_recent,
          ppt_cold_dist_fig_recent, ncol=3,nrow=5)
#ggsave("../output/Climate/all_year_Sub_dist_from_Home_WL2_Recent.png", width = 24, height = 24, units = "in")

ggarrange(cwd_dist_fig_historical, pck_dist_fig_historical, ppt_dist_fig_historical, 
          tmn_dist_fig_historical, tmx_dist_fig_historical,
          ann_tmean_dist_fig_historical, diurnal_range_dist_fig_historical, tmp_seasonality_dist_fig_historical,
          tmp_ann_range_dist_fig_historical, tmean_wet_dist_fig_historical, tmean_dry_dist_fig_historical,
          ann_ppt_dist_fig_historical, ppt_seasonality_dist_fig_historical, ppt_warm_dist_fig_historical,
          ppt_cold_dist_fig_historical, ncol=3,nrow=5)
#ggsave("../output/Climate/all_year_Sub_dist_from_Home_WL2_Historical.png", width = 24, height = 24, units = "in")
```

## Combine all Subtraction data
```{r}
recent_sub_dist_from_wl2 <- full_join(recent_flint_dist, recent_bioclim_dist)
#write_csv(recent_sub_dist_from_wl2, "../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_Recent.csv")

historic_sub_dist_from_wl2 <- full_join(historic_flint_dist, historical_bioclim_dist)
#write_csv(historic_sub_dist_from_wl2, "../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_Historical.csv")
```

#### For 2024 (from 2023 home avgs)
```{r}
recent_sub_dist_2324_from_wl2 <- full_join(recent_flint_dist_2324, recent_bioclim_dist_2324)
write_csv(recent_sub_dist_2324_from_wl2, "../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_2324_Recent.csv")

historic_sub_dist_2324_from_wl2 <- full_join(historic_flint_dist_2324, historical_bioclim_dist_2324)
write_csv(historic_sub_dist_2324_from_wl2, "../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_2324_Historical.csv")
```

### For 2024
```{r}
recent_sub_dist_2024_from_wl2 <- full_join(recent_flint_dist_2024, recent_bioclim_dist_2024)
#write_csv(recent_sub_dist_2024_from_wl2, "../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_2024_Recent.csv")

historic_sub_dist_2024_from_wl2 <- full_join(historic_flint_dist_2024, historical_bioclim_dist_2024)
#write_csv(historic_sub_dist_2024_from_wl2, "../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_2024_Historical.csv")
```

