---
title: "WL2_OverWinter_Survival"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Over-winter survival at WL2 garden 

To Do:
- Check the data - when looking at Oct 2024 survival, only found 131 "2023-survivors" whereas in this code there are 135. Are any missclassified as TM2 fruiting?
- Upate stats 

## Libraries
```{r}
library(tidyverse)
library(magrittr)
#conflicted::conflicts_prefer(dplyr::filter)
library(viridis) #for pretty colors
library(ggrepel) #for non-overlapping labels on plots
library(ggdist) #visualizations of distributions and uncertainty 
library(ggpubr)

#library(sp) #for calculating geographic distance 
library(geosphere) #for calculating geographic distance
library(corrplot) #plotting correlations  
library(rstatix) #performing cor_test

library(lmerTest) #mixed models
conflicted::conflicts_prefer(lmerTest::lmer)
library(broom.mixed)
library(tidymodels)
library(furrr) #Apply Mapping Functions in Parallel using Futures
tidymodels_prefer()
# install.packages("multilevelmod")
library(multilevelmod)

library(brms)
#conflicted::conflicts_prefer(brms::ar)
#conflicted::conflicts_prefer(dplyr::combine)
#conflicted::conflicts_prefer(brms::dstudent_t)
library(tidybayes) #for extracting and visiaulizing brms model output 
library(modelr) #for data grid

sem <- function(x, na.rm=FALSE) {           #for caclulating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 

cbbPalette2 <- c("#E69F00","#000000", "#56B4E9","#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
timepd_palette <- c("#56B4E9","#D55E00")

options(mc.cores = parallel::detectCores())
```

## Load the post-winter data
```{r}
post_winter <- read_csv("../input/WL2_Data/WL2_status_check_20240603_corrected.csv",
                         na = c("", "NA", "-", "N/A")) 
post_winter_clean <- post_winter %>% 
  mutate(pop= str_replace(pop, "iH", "IH")) %>% 
  mutate(pop= str_replace(pop, "1H", "IH")) %>% 
  mutate(pop= str_replace(pop, "cc", "CC")) %>% 
  unite(BedLoc, bed:`bed- col`, sep="_", remove = FALSE) %>% 
  filter(BedLoc!="K_5_C") %>% #get rid of duplicate locations
  filter(BedLoc!="B_32_A") %>% #get rid of duplicate locations
  unite(Genotype, pop:rep, sep="_", remove = FALSE) %>% 
  filter(pop!="buffer", !str_detect(mf, "buf")) %>% 
  mutate(mf=as.double(mf), rep=as.double(rep))
unique(post_winter_clean$pop)
```

## Gower's Distance
```{r}
garden_climate <- read_csv("../output/Climate/flint_climate_UCDpops.csv") %>% 
  filter(parent.pop=="WL2_Garden" | parent.pop=="UCD_Garden") %>% 
  select(parent.pop:Long) %>% 
  distinct()
garden_climate
#UCD LAT/LONG = 38.53250, -121.7830
#WL2 Lat/Long = 38.82599, -120.2509

wl2_gowers_2024 <- read_csv("../output/Climate/Gowers_WL2_2024.csv") %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  mutate(WL2_Lat=38.82599, WL2_Long=-120.2509, WL2_Elev=2020) %>% 
  mutate(Geographic_Dist=distHaversine(cbind(WL2_Long, WL2_Lat), cbind(Long, Lat)),
         Elev_Dist=WL2_Elev-elev_m) %>% # Calculate the distance using the haversine formula
  rename(pop=parent.pop)
```

## Calculate Winter Surv

*Used only water year gowers distance for 2024 b/c that's the only one that includes the winter months 

Categories 
A - alive, “happy” leaves
B = “barely alive”, still has leaves but major damage
C = no leaves, stem not brittle
D = brittle, no leaves, definitely dead 

```{r}
#wl2_surv_y1 %>% filter(is.na(death.date)) - 470 plants alive on 10/27
post_winter_clean %>% filter(is.na(death.date)) #44 plants with no death date, most have a death date of 10/27
#A-37-B, no death date on any data sheet, was dying on 9/20 so may have just been missed in June 2024

winter_surv <- post_winter_clean %>% 
  left_join(wl2_gowers_2024) %>% 
  filter(death.date == "A" | death.date == "B" | death.date == "C" | death.date == "D") %>% 
  mutate(WinterSurv=if_else(death.date=="D", 0, 1)) %>% 
  select(block:rep, elevation.group:Wtr_Year_GD_Historical, Geographic_Dist, Elev_Dist, death.date, WinterSurv) 
dim(winter_surv) #only 469 rows because A_37_B not included 

#winter_surv %>% filter(death.date!="D") #135 plants survived 

write_csv(winter_surv, "../output/WL2_Traits/WL2_WinterSurv.csv")
```

### Descriptive tables 
```{r}
#overall
xtabs(~WinterSurv, data=winter_surv) #335 dead post-winter out of 470 alive pre-winter = 71% overwinter mortality 

xtabs(~WinterSurv+pop, data=winter_surv)
```

### Bar Plots 
```{r}
winter_surv %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  filter(pop != "WR") %>% #only 1 plant 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/WL2_Traits/WL2_WinterSurv_Wtr_Year_GD_Recent.png", width = 12, height = 8, units = "in")
```

### Scatterplots
```{r}
#scatter plots
GSCD <- winter_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Recent, y=meanSurv, color=GrwSsn_GD_Recent, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  #geom_text_repel(aes(x = GrwSsn_GD_Recent, y = meanSurv,
  #          label = `pop`),
  #      min.segment.length = 0.8,
  #      max.overlaps = 100,
  #      #label.padding = 1,
  #      #point.padding = 0.8,
  #      size = 4) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Growth Season CD", color="Growth Season \n Climate Distance") +
  scale_color_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=25))

WYCD <- winter_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanSurv, color=Wtr_Year_GD_Recent, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02,linewidth = 2) +
  #geom_text_repel(aes(x = Wtr_Year_GD_Recent, y = meanSurv,
  #          label = `pop`),
  #      min.segment.length = 0.8,
  #      max.overlaps = 100,
  #      #label.padding = 1,
  #      #point.padding = 0.8,
  #      size = 4) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Water Year CD", color="Water Year \n Climate Distance") +
  scale_color_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=25))

GD <- winter_surv %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent, Geographic_Dist) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanSurv, color=Geographic_Dist, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  #geom_text_repel(aes(x = Geographic_Dist, y = meanSurv,
  #          label = `pop`),
  #      min.segment.length = 0.8,
  #      max.overlaps = 100,
  #      #label.padding = 1,
  #      #point.padding = 0.8,
  #      size = 4) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Geographic Distance (m)", color="Geographic Distance") +
  scale_color_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

wl2_winter_surv_FIG <- ggarrange(GSCD, WYCD, GD, ncol=2, nrow=2) 
#ggsave("../output/WL2_Traits/WL2_WinterSurv_SCATTERS.png", width = 24, height = 18, units = "in")
```


## Stats - need to update!
```{r, eval=FALSE}
lmesurv1 <- glmer(Survival ~ Recent_Gowers_Dist_WL2 + (1|parent.pop/mf) + (1|block), 
                  data = alive_oct, 
                  family = binomial(link = "logit")) 
summary(lmesurv1)
anova(lmesurv1)
glance(lmesurv1)

lmesurv2 <- glmer(Survival ~ Historic_Gowers_Dist_WL2 + (1|parent.pop/mf) + (1|block), 
                  data = alive_oct, 
                  family = binomial(link = "logit")) 
summary(lmesurv2)
anova(lmesurv2)
glance(lmesurv2)

alive_oct_scale <- alive_oct %>% 
  mutate(elev_m_s = scale_this(elev_m))
lmesurv3 <- glmer(Survival ~ elev_m_s + (1|parent.pop/mf) + (1|block), 
                  data = alive_oct_scale, 
                  family = binomial(link = "logit")) #get a failed to converge and rescale variables warning for this model 
summary(lmesurv3)
anova(lmesurv3)
glance(lmesurv3)

#to test for significance of random effect: 
#the most common way to do this is to use a likelihood ratio test, i.e. fit the full and reduced models (the reduced model is the model with the focal variance(s) set to zero). 
m0 <- glmer(Survival ~ elev_m_s + (1|parent.pop), data = alive_oct_scale, family = binomial("logit")) 
m00 <- glm(Survival~ elev_m_s, alive_oct_scale, family = binomial("logit"))
anova(lmesurv3,m0, m00) #model with both random effects has a higher likelihood (better fit)
```

