---
title: "WL2_Spatial_Variation"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Evaluating the phenotypic spatial variation (across beds) at WL2

Purpose: See if we could take advantage of the natural variation at WL2 for the 2026 planting

-   Note: There are some packages in R for autocorrelation

To Do:

-   Consider looking at stem diameter and canopy area in year 2

-   For TDR data, compare top of beds avg across the season (See Rachel's github)

-   Correlate TDR data to fitness/size data somehow 

## Load Libraries

```{r}
library(tidyverse)
library(viridis)

sem <- function(x, na.rm=FALSE) {           #for calculating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 
```

## Load Fitness Data

```{r}
wl2_est <- read_csv("../output/WL2_Traits/WL2_Establishment.csv") %>% select(block:rep, Establishment) %>% 
  mutate(bed.col=if_else(bed.col=="A -", "A", bed.col))

y1_surv <- read_csv("../output/WL2_Traits/WL2_Y1Surv.csv") %>% 
  left_join(wl2_est) %>% 
  select(block, bed:bed.col, Genotype:rep, Y1Surv=Survival) 

wintersurv <- read_csv("../output/WL2_Traits/WL2_WinterSurv_2324.csv") %>% 
  rename(bed.row=`bed- row`, bed.col=`bed- col`) %>% 
  select(block, bed:bed.col, Genotype:rep, WinterSurv)

repsurvy2 <- read_csv("../output/WL2_Traits/WL2_Surv_to_Rep_Y2_2324.csv") %>% 
  select(block, bed:col, Genotype, pop:rep, SurvtoRep_y2)

fruitsy2 <- read_csv("../output/WL2_Traits/WL2_Fruits_Y2_2324.csv") %>% 
  select(block, bed:col, Genotype, pop:rep, fruits)

probfruit <- read_csv("../output/WL2_Traits/WL2_ProbFruits_Y2_2324.csv") %>% 
  select(block, bed:col, Genotype, pop:rep, ProbFruits)

totfruit <- read_csv("../output/WL2_Traits/WL2_TotalRepOutput.csv") %>% 
  select(block:rep, Total_Fitness)
```

## Load Size (Height and herbivory 2 months post transplant; stem diameter end of first year)
```{r}
WL2_twomonths_size <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_size_survey_20230913_corrected.csv", 
                               na = c("", "NA", "-", "N/A")) %>% 
  filter(!is.na(pop)) %>% 
  rename(parent.pop=pop) %>% 
  mutate(parent.pop= str_replace(parent.pop, ".*VTR.*", "LVTR1")) %>% 
  unite(BedLoc, bed:bed.col, sep="_", remove = FALSE) %>% 
  filter(BedLoc!="K_5_C", BedLoc!="B_32_A", BedLoc!="C_4_D", BedLoc!="C_5_D") %>% 
  mutate(mf=as.double(mf), rep=as.double(rep)) %>% 
  unite(Genotype, parent.pop:rep, sep="_", remove = FALSE) %>% 
  filter(!str_detect(Genotype, ".*buff*")) %>% 
  select(block:height.cm, herbiv.y.n)

wl2_anncensus <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_annual_census_20231027_corrected.csv") %>% 
  select(block, bed, bed.row=`bed-row`, bed.col=`bed-col`, pop:diam.mm) %>% 
  filter(!is.na(pop)) %>% 
  rename(parent.pop=pop) %>% 
  mutate(parent.pop= str_replace(parent.pop, ".*VTR.*", "LVTR1")) %>% 
  unite(BedLoc, bed:bed.col, sep="_", remove = FALSE) %>% 
  filter(BedLoc!="K_5_C", BedLoc!="B_32_A", BedLoc!="C_4_D", BedLoc!="C_5_D") %>% 
  mutate(mf=as.double(mf), rep=as.double(rep)) %>% 
  unite(Genotype, parent.pop:rep, sep="_", remove = FALSE) %>% 
  filter(!str_detect(Genotype, ".*buff*"))
```

## Box Plots
```{r}
wl2_est %>% 
  ggplot(aes(x=bed, y=Establishment)) +
  geom_boxplot()

y1_surv %>% 
  ggplot(aes(x=bed, y=Y1Surv)) +
  geom_boxplot()

wintersurv %>% 
  ggplot(aes(x=bed, y=WinterSurv)) +
  geom_boxplot()

repsurvy2 %>% 
  ggplot(aes(x=bed, y=SurvtoRep_y2)) +
  geom_boxplot()

fruitsy2 %>% 
  ggplot(aes(x=bed, y=fruits)) +
  geom_boxplot()

probfruit %>% 
  ggplot(aes(x=bed, y=ProbFruits)) +
  geom_boxplot()

totfruit %>% 
  ggplot(aes(x=bed, y=Total_Fitness)) +
  geom_boxplot()

WL2_twomonths_size %>% 
  ggplot(aes(x=bed, y=height.cm)) +
  geom_boxplot()

WL2_twomonths_size %>% 
  mutate(herbiv.bin = if_else(herbiv.y.n=="Y", 1, 0)) %>% 
  ggplot(aes(x=bed, y=herbiv.bin)) +
  geom_boxplot()

wl2_anncensus %>% 
  ggplot(aes(x=bed, y=diam.mm)) +
  geom_boxplot()
```

## Summaries
```{r}
wl2_est_summary <- wl2_est %>% 
  group_by(bed) %>% 
  summarise(meanSurv=mean(Establishment, na.rm=TRUE), semSurv=sem(Establishment, na.rm=TRUE))

y1surv_summary <- y1_surv %>% 
  group_by(bed) %>% 
  summarise(meanSurv=mean(Y1Surv, na.rm=TRUE), semSurv=sem(Y1Surv, na.rm=TRUE))

wintersurv_summary <- wintersurv %>% 
  group_by(bed) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm=TRUE), semSurv=sem(WinterSurv, na.rm=TRUE))

repsurvy2_summary <- repsurvy2 %>% 
  group_by(bed) %>% 
  summarise(meanSurv=mean(SurvtoRep_y2, na.rm=TRUE), semSurv=sem(SurvtoRep_y2, na.rm=TRUE))

fruitsy2_summary <- fruitsy2 %>% 
  group_by(bed) %>% 
  summarise(meanSurv=mean(fruits, na.rm=TRUE), semSurv=sem(fruits, na.rm=TRUE))

probfruit_summary <- probfruit %>% 
  group_by(bed) %>% 
  summarise(meanSurv=mean(ProbFruits, na.rm=TRUE), semSurv=sem(ProbFruits, na.rm=TRUE))

totfruit_summary <- totfruit %>% 
  group_by(bed) %>% 
  summarise(meanSurv=mean(Total_Fitness, na.rm=TRUE), semSurv=sem(Total_Fitness, na.rm=TRUE))

height_summary <- WL2_twomonths_size %>% 
  group_by(bed) %>% 
  summarise(meanHeight=mean(height.cm, na.rm=TRUE), semHeight=sem(height.cm, na.rm=TRUE))

herbiv_summary <- WL2_twomonths_size %>% 
  mutate(herbiv.bin = if_else(herbiv.y.n=="Y", 1, 0)) %>% 
  group_by(bed) %>% 
  summarise(meanHerb=mean(herbiv.bin, na.rm=TRUE), semHerb=sem(herbiv.bin, na.rm = TRUE))

diam_summary <- wl2_anncensus %>% 
  group_by(bed) %>% 
  summarise(meanDiam=mean(diam.mm, na.rm=TRUE), semDiam=sem(diam.mm, na.rm=TRUE))
```

## Bar Plots
```{r}
wl2_est_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanSurv), y=meanSurv)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Establishment")

y1surv_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanSurv), y=meanSurv)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Y1 Surv")

wintersurv_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanSurv), y=meanSurv)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Winter Surv")

repsurvy2_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanSurv), y=meanSurv)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Surv to Bud Y2")

fruitsy2_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanSurv), y=meanSurv)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Fruits Y2")

probfruit_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanSurv), y=meanSurv)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Prob. Fruit")

totfruit_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanSurv), y=meanSurv)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Total Fruit")

height_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanHeight), y=meanHeight)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanHeight-semHeight,ymax=meanHeight+semHeight),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Height (cm)", title = "Two Months Post-Transplant")

herbiv_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanHerb), y=meanHerb)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanHerb-semHerb,ymax=meanHerb+semHerb),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Prob of Herbivory", title = "Two Months Post-Transplant")

diam_summary %>% 
  ggplot(aes(x=fct_reorder(bed, meanDiam), y=meanDiam)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanDiam-semDiam,ymax=meanDiam+semDiam),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() +
  labs(x="Bed", y="Stem Diameter (mm)")
```

## Spatial Plots Prep
x = beds + cols
y = row 
```{r}
bed_col_info <- wl2_est %>% 
  select(bed, bed.col) %>% 
  distinct() %>% 
  mutate(x_pos = if_else(bed.col=="A", 1,
                         if_else(bed.col=="B", 2, 
                                 if_else(bed.col=="C", 3, 4)))) %>% 
  mutate(x_pos=if_else(bed=="B", x_pos+6,
                       if_else(bed=="C", x_pos+12,
                               if_else(bed=="D", x_pos+18,
                                       if_else(bed=="E", x_pos+24,
                                               if_else(bed=="F", x_pos+30,
                                                       if_else(bed=="G", x_pos+36,
                                                               if_else(bed=="H", x_pos+42, 
                                                                       if_else(bed=="I", x_pos+48,
                                                                               if_else(bed=="J", x_pos+54, 
                                                                                       if_else(bed=="K", x_pos+60, x_pos))))))))))) %>% 
  arrange(bed, bed.col)

bed_col_info

bed_rect <- data.frame( #data frame for bed rectangles 
  xmin = c(1, 7, 13, 19, 25, 31, 37, 43, 49, 55, 61),
  xmax = c(4, 10, 16, 22, 28, 34, 40, 46, 52, 58, 64),
  ymin = 1,
  ymax = c(59, 59, 58, 59, 51, 46, 43, 34, 22, 18, 13)
)
```

## Spatial Plots
```{r}
wl2_est %>% 
  left_join(bed_col_info) %>% 
  mutate(Establishment=as.character(Establishment)) %>% 
  ggplot(aes(x=x_pos, y=bed.row, colour=Establishment)) +
  geom_point(size=3) +
  scale_y_reverse() +
  scale_color_manual(values = c("gray", "darkgreen")) + 
  theme_classic()

y1_surv %>% 
  left_join(bed_col_info) %>% 
  mutate(Y1Surv=as.character(Y1Surv)) %>% 
  ggplot(aes(x=x_pos, y=bed.row, colour=Y1Surv)) +
  geom_point(size=3) +
  scale_y_reverse() +
  scale_color_manual(values = c("gray", "darkgreen")) + 
  theme_classic()

wintersurv %>% 
  left_join(bed_col_info) %>% 
  mutate(WinterSurv=as.character(WinterSurv)) %>% 
  ggplot(aes(x=x_pos, y=bed.row, colour=WinterSurv)) +
  geom_point(size=3) +
  scale_y_reverse() +
  scale_color_manual(values = c("gray", "darkgreen")) + 
  theme_classic()

repsurvy2 %>% 
  rename(bed.row=row, bed.col=col) %>% 
  left_join(bed_col_info) %>% 
  mutate(SurvtoRep_y2=as.character(SurvtoRep_y2)) %>% 
  ggplot(aes(x=x_pos, y=bed.row, colour=SurvtoRep_y2)) +
  geom_point(size=3) +
  scale_y_reverse() +
  scale_color_manual(values = c("gray", "darkgreen")) + 
  geom_rect(data=bed_rect, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), colour = "black", alpha=0, inherit.aes = FALSE) +
  theme_classic()

probfruit %>% 
  rename(bed.row=row, bed.col=col) %>% 
  left_join(bed_col_info) %>% 
  mutate(ProbFruits=as.character(ProbFruits)) %>% 
  ggplot(aes(x=x_pos, y=bed.row, colour=ProbFruits)) +
  geom_point(size=3) +
  scale_y_reverse() +
  scale_color_manual(values = c("gray", "darkgreen")) + 
  geom_rect(data=bed_rect, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), colour = "black", alpha=0, inherit.aes = FALSE) +
  theme_classic()

 WL2_twomonths_size %>% 
  mutate(herbiv.bin = if_else(herbiv.y.n=="Y", 1, 0)) %>% 
  left_join(bed_col_info) %>% 
  mutate(herbiv.bin=as.character(herbiv.bin)) %>% 
   filter(!is.na(herbiv.bin)) %>% 
  ggplot(aes(x=x_pos, y=bed.row, colour=herbiv.bin)) +
  geom_point(size=3) +
  scale_y_reverse() +
  scale_color_manual(values = c("gray", "darkgreen")) + 
  theme_classic()
```

### Continuous Traits (fruit number and size)
```{r}
fruitsy2 %>% 
  rename(bed.row=row, bed.col=col) %>% 
  left_join(bed_col_info) %>% 
  ggplot(aes(x=x_pos, y=bed.row, colour=fruits)) +
  geom_point(size=3) +
  scale_y_reverse() +
  scale_color_viridis() +
  geom_rect(data=bed_rect, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), colour = "black", alpha=0, inherit.aes = FALSE) +
  theme_classic()

totfruit %>% 
  left_join(bed_col_info) %>% 
  ggplot(aes(x=x_pos, y=bed.row, colour=Total_Fitness)) +
  geom_point(size=3) +
  scale_y_reverse() +
  scale_color_viridis() +
  geom_rect(data=bed_rect, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), colour = "black", alpha=0, inherit.aes = FALSE) +
  theme_classic()

WL2_twomonths_size %>% 
  left_join(bed_col_info) %>% 
  filter(!is.na(height.cm)) %>% 
  ggplot(aes(x=x_pos, y=bed.row, colour=height.cm)) +
  geom_point(size=3) +
  scale_y_reverse() +
  scale_color_viridis() +
  theme_classic()

wl2_anncensus %>% 
  left_join(bed_col_info) %>% 
  filter(!is.na(diam.mm)) %>% 
  ggplot(aes(x=x_pos, y=bed.row, colour=diam.mm)) +
  geom_point(size=3) +
  scale_y_reverse() +
  scale_color_viridis() +
  theme_classic()
```

