---
title: "WL2_Survival-Analysis--GLMER"
author: "Julin Maloof"
date: '`r Sys.Date()`'
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script focuses on using populations as random effects.  See other scripts for analysis by elevation

# Survival Analysis 

See chapters 9 - 13 in tutorial: https://bookdown.org/content/4253/extending-the-discrete-time-hazard-model.html

## Libraries
```{r}
# install.packages("multilevelmod")
library(tidyverse) #includes lubridate, which we need 
library(magrittr)
library(ggrepel)
library(lme4)
library(ggdist)
library(tidymodels)
library(multilevelmod)
library(broom.mixed)
library(modelr)
library(furrr)
tidymodels_prefer()
options(mc.cores = parallel::detectCores())
```

## Read in the data
```{r}
wl2_surv <- read_csv("../output/WL2_Traits/WL2_Mortality_2023.csv")
```

As weeks
```{r}
wl2_surv %>% filter(is.na(death.date),!is.na(survey.notes)) #no plants that seemed to disappear 
wl2_surv %>% filter(Genotype=="YO7_4_2")

wl2_surv_dates <- wl2_surv %>%  
  rename(parent.pop = pop) %>%
  mutate(parent.pop= str_replace(parent.pop, "Y08", "YO8")) %>% 
  mutate(parent.pop= str_replace(parent.pop, "Y04", "YO4")) %>% 
  filter(!is.na(parent.pop)) %>% 
  mutate(planting.date="7/19/23", #could try to make this more specific to when certain blocks were planted 
         last_fup_date=if_else(is.na(death.date), "10/27/23", death.date)) %>%  #need this to calculate survival times
  mutate(planting.date=mdy(planting.date), last_fup_date=mdy(last_fup_date)) %>% #convert to date objects
  mutate(os_weeks=as.duration(planting.date %--% last_fup_date) / dweeks(1), #observed number of weeks
         status=if_else(is.na(death.date), 0, 1)) %>% #0=censured (alive in this case), 1=dead
  filter(os_weeks > 0) %>% #there is one case of a plant that was dead at planting, so just removed it since this is survival post- transplanting
  select(-survey.notes)
head(wl2_surv_dates)
unique(wl2_surv_dates$os_weeks) %>% sort()
```


## Location Info

```{r}
gowersdist_WL2 <- read_csv("../output/Climate/Gowers_WL2.csv") 
gowersdist_WL2 %>% arrange(parent.pop)
```

```{r, fig.asp=1}
gowersdist_WL2 %>% 
  select(ends_with("GD")) %>%
  GGally::ggpairs(progress = FALSE)
```

```{r}
gowersdist_WL2_wide <- gowersdist_WL2 %>% select(-matches("BIOCLIM|FLINT")) %>%
  pivot_wider(values_from=c(GrwSsn_GD, Wtr_Year_GD), names_from = TimePd)
```


```{r}
wl2_surv_dates_loc <- left_join(wl2_surv_dates, gowersdist_WL2_wide)  %>%
  mutate(elevation.group = factor(elevation.group, levels= c("Low", "Mid", "High")))
wl2_surv_dates_loc 
```


## plot observed data

Need to summarize separately for pops and for elevation.  It is a mistake to average pops and then elevation, at least with respect to comparing to population naive models

### population-level summary
```{r}
wl2_obs.pop <- 
  wl2_surv_dates_loc %>% select(parent.pop, death.date, os_weeks, status, elevation.group, ends_with("GD")) %>%
  group_by(parent.pop) %>%
  mutate(total=n()) %>%
  group_by(parent.pop, os_weeks) %>%
  summarize(n.newly.dead=sum(status),
            across(c(total, elevation.group, ends_with("GD")), unique)) %>%
  ungroup() %>%
  complete(parent.pop, os_weeks, fill = list(n.newly.dead=0)) %>% # we have missing data for weeks where nobody died and we need to fill that in
  group_by(parent.pop) %>%
  mutate(total=unique(na.omit(total)), elevation.group=unique(na.omit(elevation.group)), # filling in the blanks from the complete() step above
         total.dead=cumsum(n.newly.dead),
         surviving = total - total.dead,
         prop.surviving = surviving/total,
         across(ends_with("GD"), \(x) unique(na.omit(x)))) %>%
  ungroup() %>%
#  bind_rows(week0.pop) %>%
#  select(parent.pop, elevation.group, os_weeks, prop.surviving) %>%
  arrange(parent.pop, os_weeks)
```


## Reformat to be one row per plant per week

```{r, eval=TRUE}
wl2_ind <- wl2_surv_dates %>% 
  complete(Genotype, last_fup_date) %>%
  group_by(Genotype) %>%
  mutate(across(c(-status, -last_fup_date, -os_weeks, -planting.date), \(x) ifelse(all(is.na(unique(x))), NA,unique(na.omit(x))))) %>%
  mutate(planting.date = unique(na.omit(planting.date)),
         os_weeks=as.duration(planting.date %--% last_fup_date) / dweeks(1),
           dead = ifelse((is.na(death.date) | mdy(death.date) > last_fup_date), 0, 1)) %>% # 0 is alive, 1 is dead
  filter(cumsum(dead) < 2) %>% # you can't die more than once
  ungroup() %>%  
  left_join(gowersdist_WL2_wide)

wl2_ind %>% select(Genotype, last_fup_date, os_weeks, dead)
```

# Modeling probability of event rather than event time

An alternative approach to hazard analysis is to model the probably of death as a function of time.  This is approach taken here.

## Discrete intervals

Modify the input data to add some factor variables
```{r}
wl2_ind <- wl2_ind %>% mutate(
  os_weeks_f = factor(str_pad(round(os_weeks, 1), 2, pad = "0")), # padding so it will sort correctly
  os_weeks_pop = factor(str_c(parent.pop,"_", os_weeks_f))
)
wl2_ind
```

Split data into establishment and survival data sets
```{r}
wl2_ind_estab <- wl2_ind %>% filter(os_weeks <= 3) %>%
  # compute status at week 3 for each plant
  group_by(Genotype) %>%
  summarize(dead=max(dead),
            across(c(block, bed, parent.pop, mf, contains("GD")), unique))

wl2_ind_surv <- wl2_ind %>% filter(os_weeks > 3) %>%
  mutate(os_weeks_f = factor(os_weeks))

wl2_estab_means <- wl2_ind_estab %>% # summary for plotting
  group_by(parent.pop) %>%
  summarize(survivorship = 1-mean(dead))

wl2_surv_means <- wl2_ind_surv %>%
  group_by(parent.pop, os_weeks) %>%
  summarize(survivorship = 1-mean(dead))
```

# Establishment data set

```{r}
glmer.model <- 
  linear_reg() %>%
  set_engine("glmer", family = binomial)
```

```{r}
estab_wflow <- workflow() %>%
  add_variables(outcomes = dead, predictors = c(parent.pop, mf, block)) 

estab_fits <- tibble(wflow=list(
  pop = {estab_wflow %>% 
      add_model(glmer.model, formula = dead ~ (1|parent.pop))},
  
  pop.mf = {estab_wflow %>% 
      add_model(glmer.model, formula = dead ~ (1|parent.pop/mf))},
  
  pop.block = {estab_wflow %>% 
      add_model(glmer.model, formula = dead ~ (1|parent.pop) + (1|block))},
  
  pop.mf.block = {estab_wflow %>% 
      add_model(glmer.model, formula = dead ~ (1|parent.pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)
estab_fits
```

```{r}
estab_fits <- estab_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_ind_estab))
```

Trying out multicore mapping
```{r, eval=FALSE}
estab_fits_test <- estab_fits %>%
  select(-fit) %>%
  mutate(fit = future_map(wflow, fit, data = wl2_ind_estab,
                          .options = furrr_options(
                            packages = c("tidymodels", "multilevelmod")
                          )))
```


```{r}
estab_fits %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
```

model with pop and block (but excluding mf) is best by AIC and BIC

get bootstrap predictions and plot
```{r}
estab_nd <- data_grid(wl2_ind_estab, parent.pop)

estab_predict_fun <- function(m, nd = estab_nd) {
  predict(m, 
          newdata = nd,
          type = "response",
          re.form = ~(1|parent.pop)
  ) %>%
    subtract(1, .) # convert hazard to survivorship
}

# 100 seconds
system.time (estab_fits <- estab_fits %>% 
               mutate(
                 bootstrap = map(fit, \(x) {
                   extract_fit_engine(x) %>% 
                     bootMer(FUN = estab_predict_fun,
                             re.form = ~ (1|parent.pop),
                             nsim = 1000,
                             parallel = "multicore",
                             ncpus = 7
                     ) } #bootMer
                 ),  
                 tidyboot = map(bootstrap, tidy, conf.int = TRUE),
                 tidyboot = map(tidyboot, \(x) cbind(estab_nd, x))
               ) # mutate
) # system.time
```


```{r}
estab_fits %>%
  select(name, tidyboot) %>%
  unnest(tidyboot) %>%
  ggplot(aes(x=parent.pop)) +
  geom_col(aes(y=survivorship), data = wl2_estab_means, fill="skyblue") +
  geom_pointrange(aes(y=statistic, ymax = conf.high, ymin = conf.low, color=name), position = position_dodge(width = .75)) +
  scale_color_viridis_d() +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=.5))
  

# NOTE: need bootmer for standard errors on predictions
```

```{r}
estab_fits %>%
  select(name, bootstrap) %>%
  mutate(tidyboot = map(bootstrap, tidy, conf.int = TRUE)) %>%
  unnest(tidyboot)
```

Keep block but not mf

### is Gower's distance an important predictor?


```{r}
estab_GD_wflow <- workflow() %>%
  add_variables(outcomes = dead, predictors = c(parent.pop, mf, block, contains("GD"))) 

estab_GD_fits <- tibble(wflow=list(
  pop.block = {estab_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ (1|parent.pop) + (1|block))},
  
  GS_Recent = {estab_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ GrwSsn_GD_Recent + (1|parent.pop) + (1|block))},
  
  GS_Historical = {estab_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ GrwSsn_GD_Historical + (1|parent.pop) + (1|block))},
  
  WY_Recent = {estab_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ Wtr_Year_GD_Recent + (1|parent.pop) + (1|block))},
  
  WY_Historical = {estab_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ Wtr_Year_GD_Historical + (1|parent.pop) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = wl2_ind_estab))

```

```{r}
estab_GD_fits %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
```
For any given GD, including a random effect for (GD | parent.pop) fits worse

```{r}
estab_GD_fits %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "GD")) %>%
  drop_na(p.value) %>%
  select(-wflow:-group) %>%
  arrange(p.value)

```

# Survival data set

check basic model forms
```{r}
# Note: StdDev due to mf withint parent.pop is essentially zero, so I am omitting this.

surv_wflow <- workflow() %>%
  add_variables(outcomes = dead, predictors = c(parent.pop, mf, block, os_weeks, os_weeks_f)) 

surv_fits_1 <- tibble(wflow=list(

  pop.block = {surv_wflow %>% 
      add_model(glmer.model, formula = dead ~ (1|parent.pop) + (1|block))},
  
  pop.weeks_f.block ={surv_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks_f + (1|parent.pop) + (1|block))},

  pop.weeks.block ={surv_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + (1|parent.pop) + (1|block))},

  pop.weeks.quad.block ={surv_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + I(os_weeks^2) + (1|parent.pop) + (1|block))}
      
  # Models commented out fail to converge, so not running them.
  # Any model that has a RE slope temer for weeks fails.
  # Note:only having os_weeks as RE slope does not fix this.
  
  # pop.weeks_f.int ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks_f + (os_weeks_f|parent.pop))},
  # 
  # pop.weeks_f.int.block ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks_f + (os_weeks_f|parent.pop) + (1|block))},
  # 

  # pop.weeks.int ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks + (os_weeks_f|parent.pop))},
  # 
  # pop.weeks.int.block ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks + (os_weeks_f|parent.pop) + (1|block))},

  # Models below fail to converge, so not running them.
  # Note:only having os_weeks as RE slope does not fix this.
  
  # pop.weeks.quad.int ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks + I(os_weeks^2) + (os_weeks_f|parent.pop))},
  # 
  # pop.weeks.quad.quad.int ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks + I(os_weeks^2) + (os_weeks_f + I(os_weeks^2)|parent.pop))},
  # 
  # pop.weeks.quad.int.block ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks + I(os_weeks^2) + (os_weeks_f|parent.pop) + (1|block))},
  # 
  # pop.weeks.quad.quad.int.block ={surv_wflow %>% 
  #     add_model(glmer.model, formula = dead ~ os_weeks + I(os_weeks^2) + (os_weeks_f + I(os_weeks^2)|parent.pop) + (1|block))}
    
),
name=names(wflow)
) %>% 
  select(name,wflow)
surv_fits_1
```

```{r}
system.time(
  surv_fits_1 <- surv_fits_1 %>%
  mutate(fit = map(wflow, fit, data = wl2_ind_surv))
)
```
get warnings
```{r}
surv_fits_1 <- surv_fits_1 %>%
  mutate(message=map_chr(fit, \(x) {extract_fit_engine(x) %>% .@optinfo %>% magrittr::extract2("message")}))

surv_fits_1 %>% select(name, message) %>%
  arrange(message, name)
```
A little strange since I get a warning when running about failure to converge for the model with os_weeks_f

```{r}
surv_fits_1 %>% pull(fit)
```

```{r}
surv_fits_1 %>% 
  filter(!str_detect(message, "failure")) %>%
  mutate(glance=map(fit, glance)) %>% 
  unnest(glance) %>% 
  arrange(BIC) %>%
  select(-wflow:-sigma)
```

All models with a random slope for time for each pop don't converge.

get bootstrap predictions and plot
```{r}
surv_nd <- data_grid(wl2_ind_surv, parent.pop, os_weeks) %>%
  mutate(os_weeks_f = factor(os_weeks))

surv_predict_fun <- function(m, nd = surv_nd) {
  predict(m, 
          newdata = nd,
          type = "response",
          re.form = ~(1|parent.pop)
  ) %>%
    subtract(1, .) # convert hazard to survivorship
}

# ~975 seconds for 100.  Need to go somewhere other than my laptop for more
system.time (surv_fits_1 <- surv_fits_1 %>% 
                 filter(!str_detect(message, "failure")) %>%
               mutate(
                 bootstrap = map(fit, \(x) {
                   extract_fit_engine(x) %>% 
                     bootMer(FUN = surv_predict_fun,
                             re.form = ~ (1|parent.pop),
                             nsim = 100,
                             parallel = "multicore",
                             ncpus = 7
                     ) } #bootMer
                 ),  
                 tidyboot = map(bootstrap, tidy, conf.int = TRUE),
                 tidyboot = map(tidyboot, \(x) cbind(surv_nd, x))
               ) # mutate
) # system.time
```

```{r}
save.image("../output/JM_GLMER.Rdata")
```

```{r}
surv_fits_1 %>%
  select(name, tidyboot) %>%
  unnest(tidyboot) %>%
  mutate(survivorship = statistic) %>%
  ggplot(aes(x=os_weeks, y = survivorship)) +
  geom_line(aes(color=name)) +
  geom_point(data = wl2_surv_means, pch=1) + 
  facet_wrap(~parent.pop) +
  scale_color_viridis_d() # +
  #theme(axis.text.x = element_text(angle=90, hjust=1, vjust=.5))
  
```

I AM HERE.  Code below should be ~ Correct

Choose final model

Check mf

Add GDs

### is Gower's distance an important predictor? 

Also: check mf for final model

```{r}
surv_GD_wflow <- workflow() %>%
  add_variables(outcomes = dead, predictors = c(parent.pop, mf, block, os_weeks,  contains("GD"))) 

surv_GD_fits <- tibble(wflow=list(

  pop.weeks.block = {surv_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + (1|parent.pop) + (1|block))},
  
  pop.mf.weeks.block = {surv_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + (1|parent.pop/mf) + (1|block))},
  
  GS_Recent = {surv_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + GrwSsn_GD_Recent + (1|parent.pop) + (1|block))},
  
  GS_Historical = {surv_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + GrwSsn_GD_Historical + (1|parent.pop) + (1|block))},
  
  WY_Recent = {surv_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + Wtr_Year_GD_Recent + (1|parent.pop) + (1|block))},
  
  WY_Historical = {surv_GD_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + Wtr_Year_GD_Historical + (1|parent.pop) + (1|block))}

),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = wl2_ind_surv))

```

```{r}
surv_GD_fits %>% 
  mutate(glance=map(fit, glance)) %>% 
  unnest(glance) %>% 
  arrange(BIC) %>%
  select(-wflow:-sigma)
```

```{r}
surv_GD_fits %>% 
  mutate(tidy=map(fit, tidy)) %>% 
  unnest(tidy) %>% 
  filter(str_detect(term, "GD")) %>%
  select(-wflow:-group) %>%
  arrange(p.value)
```