---
title: "WL2_OverWinter_Survival"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Over-winter survival at WL2 garden 

Note: Made this climate distance of 2024 garden climate - recent and historical home climates (up to 2023)

## Libraries
```{r}
library(tidyverse)
library(magrittr)
#conflicted::conflicts_prefer(dplyr::filter)
library(viridis) #for pretty colors
library(ggrepel) #for non-overlapping labels on plots
library(ggdist) #visualizations of distributions and uncertainty 
library(ggpubr)

#library(sp) #for calculating geographic distance 
library(geosphere) #for calculating geographic distance
library(corrplot) #plotting correlations  
library(rstatix) #performing cor_test

library(lmerTest) #mixed models
conflicted::conflicts_prefer(lmerTest::lmer)
library(broom.mixed)
library(tidymodels)
library(furrr) #Apply Mapping Functions in Parallel using Futures
tidymodels_prefer()
# install.packages("multilevelmod")
library(multilevelmod)

library(brms)
#conflicted::conflicts_prefer(brms::ar)
#conflicted::conflicts_prefer(dplyr::combine)
#conflicted::conflicts_prefer(brms::dstudent_t)
library(tidybayes) #for extracting and visiaulizing brms model output 
library(modelr) #for data grid

sem <- function(x, na.rm=FALSE) {           #for caclulating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 

cbbPalette2 <- c("#E69F00","#000000", "#56B4E9","#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
timepd_palette <- c("#56B4E9","#D55E00")

options(mc.cores = parallel::detectCores())
```

## Load the post-winter data
```{r}
post_winter <- read_csv("../input/WL2_Data/WL2_status_check_20240603_corrected.csv",
                         na = c("", "NA", "-", "N/A")) 
post_winter_clean <- post_winter %>% 
  mutate(pop= str_replace(pop, "iH", "IH")) %>% 
  mutate(pop= str_replace(pop, "1H", "IH")) %>% 
  mutate(pop= str_replace(pop, "cc", "CC")) %>% 
  unite(BedLoc, bed:`bed- col`, sep="_", remove = FALSE) %>% 
  filter(BedLoc!="K_5_C") %>% #get rid of duplicate locations
  filter(BedLoc!="B_32_A") %>% #get rid of duplicate locations
  unite(Genotype, pop:rep, sep="_", remove = FALSE) %>% 
  filter(pop!="buffer", !str_detect(mf, "buf")) %>% 
  mutate(mf=as.double(mf), rep=as.double(rep))
unique(post_winter_clean$pop)
```

## Gower's Distance
```{r}
garden_climate <- read_csv("../output/Climate/flint_climate_UCDpops.csv") %>% 
  filter(parent.pop=="WL2_Garden" | parent.pop=="UCD_Garden") %>% 
  select(parent.pop:Long) %>% 
  distinct()
garden_climate
#UCD LAT/LONG = 38.53250, -121.7830
#WL2 Lat/Long = 38.82599, -120.2509

wl2_gowers_2324 <- read_csv("../output/Climate/Gowers_WL2_2324.csv") %>% 
  select(parent.pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  mutate(WL2_Lat=38.82599, WL2_Long=-120.2509, WL2_Elev=2020) %>% 
  mutate(Geographic_Dist=distHaversine(cbind(WL2_Long, WL2_Lat), cbind(Long, Lat)),
         Elev_Dist=elev_m-WL2_Elev) %>% # Calculate the distance using the haversine formula
  #mutate(Lat_Dist=WL2_Lat-Lat, Long_Dist=WL2_Long-Long) %>% #Garden-Home - lat and long per Gerst et al 2011 which kept them separate for some directionality
  rename(pop=parent.pop)

wl2_gowers_2024 <- read_csv("../output/Climate/Gowers_WL2_2024.csv") %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  mutate(WL2_Lat=38.82599, WL2_Long=-120.2509, WL2_Elev=2020) %>% 
  mutate(Geographic_Dist=distHaversine(cbind(WL2_Long, WL2_Lat), cbind(Long, Lat)),
         Elev_Dist=elev_m-WL2_Elev) %>% # Calculate the distance using the haversine formula
  rename(pop=parent.pop)
```

## Climate Subtraction Distance

```{r}
wl2_wtr_year_sub_recent_2324 <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_2324_Recent.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Recent=ann_tmean_dist, 
         Wtr_Year_PPTDist_Recent=ann_ppt_dist,
         Wtr_Year_CWDDist_Recent=cwd_dist)
wl2_wtr_year_sub_historic_2324 <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_2324_Historical.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Historic=ann_tmean_dist, 
         Wtr_Year_PPTDist_Historic=ann_ppt_dist,
         Wtr_Year_CWDDist_Historic=cwd_dist)

wl2_sub_dist_2324 <- wl2_wtr_year_sub_recent_2324 %>% 
  left_join(wl2_wtr_year_sub_historic_2324) %>% 
  rename(pop=parent.pop) %>% 
  left_join(wl2_gowers_2324) %>% 
  mutate(Wtr_Year_TempPpt_Recent=scale(Wtr_Year_TempDist_Recent,center=FALSE) +
           scale(Wtr_Year_PPTDist_Recent,center=FALSE),
         Wtr_Year_TempPpt_Historic=scale(Wtr_Year_TempDist_Historic,center=FALSE) +
           scale(Wtr_Year_PPTDist_Historic,center=FALSE))
```

```{r}
wl2_wtr_year_sub_recent_2024 <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_2024_Recent.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Recent=ann_tmean_dist, Wtr_Year_PPTDist_Recent=ann_ppt_dist)
wl2_wtr_year_sub_historic_2024 <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_2024_Historical.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Historic=ann_tmean_dist, Wtr_Year_PPTDist_Historic=ann_ppt_dist)

wl2_sub_dist_2024 <- wl2_wtr_year_sub_recent_2024 %>% 
  left_join(wl2_wtr_year_sub_historic_2024) %>% 
  rename(pop=parent.pop) %>% 
  left_join(wl2_gowers_2024) %>% 
  mutate(Wtr_Year_TempPpt_Recent=scale(Wtr_Year_TempDist_Recent,center=FALSE) +
           scale(Wtr_Year_PPTDist_Recent,center=FALSE),
         Wtr_Year_TempPpt_Historic=scale(Wtr_Year_TempDist_Historic,center=FALSE) +
           scale(Wtr_Year_PPTDist_Historic,center=FALSE))
```

## Calculate Winter Surv

*Used only water year gowers distance b/c that's the only one that includes the winter months 

Categories 
A - alive, “happy” leaves
B = “barely alive”, still has leaves but major damage
C = no leaves, stem not brittle
D = brittle, no leaves, definitely dead 

```{r}
#wl2_surv_y1 %>% filter(is.na(death.date)) - 470 plants alive on 10/27
#post_winter_clean %>% filter(is.na(death.date)) #44 plants with no death date, most have a death date of 10/27
#A-37-B, no death date on any data sheet, was dying on 9/20 so may have just been missed in June 2024

winter_surv <- post_winter_clean %>% 
  left_join(wl2_gowers_2024) %>% 
  filter(death.date == "A" | death.date == "B" | death.date == "C" | death.date == "D") %>% 
  mutate(WinterSurv=if_else(death.date=="D", 0, 1)) %>% 
  select(block:rep, elevation.group:Wtr_Year_GD_Historical, Geographic_Dist, Elev_Dist, death.date, WinterSurv) 
dim(winter_surv) #only 469 rows because A_37_B not included 

#winter_surv %>% filter(death.date!="D") #135 plants survived 

#write_csv(winter_surv, "../output/WL2_Traits/WL2_WinterSurv.csv")
```

```{r}
winter_surv_sub_dist <- winter_surv %>% 
  select(block:rep, WinterSurv) %>% 
  left_join(wl2_sub_dist_2024)
```

### Wtih 2023 Gowers
```{r}
winter_surv_2324 <- post_winter_clean %>% 
  left_join(wl2_gowers_2324) %>% 
  filter(death.date == "A" | death.date == "B" | death.date == "C" | death.date == "D") %>% 
  mutate(WinterSurv=if_else(death.date=="D", 0, 1)) %>% 
  select(block:rep, elevation.group:Wtr_Year_GD_Historical, Geographic_Dist, Elev_Dist, death.date, WinterSurv) 
#write_csv(winter_surv, "../output/WL2_Traits/WL2_WinterSurv_2324.csv")
```

```{r}
winter_surv_sub_dist_2324 <- winter_surv_2324 %>% 
  select(block:rep, WinterSurv) %>% 
  left_join(wl2_sub_dist_2324)
```

### Descriptive tables 
```{r}
#overall
xtabs(~WinterSurv, data=winter_surv) #334 dead post-winter out of 469 alive pre-winter = 71% overwinter mortality 

xtabs(~WinterSurv+pop, data=winter_surv)
```

### Bar Plots 
```{r}
winter_surv %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  filter(pop != "WR") %>% #only 1 plant 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/WL2_Traits/WL2_WinterSurv_Wtr_Year_GD_Recent.png", width = 12, height = 8, units = "in")
```

## Figure for Paper
```{r}
WYCD_recent <- winter_surv %>%
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Survival", x="Recent Water Year CD") +
  theme(text=element_text(size=25))
#ggsave("../output/WL2_Traits/WL2_WinterSurv_ClimDist.png", width = 8, height = 5, units = "in")
```

```{r}
winter_surv_2324 %>%
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanSurv, group = pop, colour = elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02,linewidth = 2) +
  geom_text_repel(aes(x = Wtr_Year_GD_Recent, y = meanSurv,
          label = `pop`),
      min.segment.length = 1,
      max.overlaps = 100,
      #label.padding = 1,
      point.padding = 1,
      size = 4) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Survival", x="Recent Water Year CD (2023)") +
  theme(text=element_text(size=25))
```

```{r}
WYCD_recent <- winter_surv_2324 %>%
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanSurv, group = pop, colour = elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02,linewidth = 2) +
  annotate("text", x = 0.227, y= 0.24, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Survival", x="Recent Water Year Climate Dist", colour="Elevation (m)") +
  theme(text=element_text(size=25))
ggsave("../output/WL2_Traits/WL2_WinterSurv_2324ClimDist.png", width = 10, height = 5, units = "in")
```

### Check CWD Dist
```{r}
winter_surv_sub_dist_2324 %>% 
  group_by(pop, elev_m, Wtr_Year_CWDDist_Recent, Wtr_Year_CWDDist_Historic) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_CWDDist_Recent, y=meanSurv, group = pop, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.3,linewidth = 2) +
  #annotate("text", x = 12.1639444	, y= 0.65, label = "WL2", 
   #       colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Survival", x="Recent Water Year CWD Dist", 
       color="Elevation (m)") +
  theme(text=element_text(size=30))

winter_surv_sub_dist_2324 %>% 
  group_by(pop, elev_m, Wtr_Year_CWDDist_Recent, Wtr_Year_CWDDist_Historic) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_CWDDist_Historic, y=meanSurv, group = pop, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.3,linewidth = 2) +
  #annotate("text", x = 8.9845278		, y= 0.65, label = "WL2", 
   #       colour = "purple", fontface="bold", size = 22 / .pt) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Survival", x="Historic Water Year CWD Dist", 
       color="Elevation (m)") +
  theme(text=element_text(size=30))
```

### Check for adaptational lag of WL2 pop
T-test compare WL2 to the pop that performed the best?
```{r}
winter_surv_sub_dist_2324 %>% 
  group_by(pop, elev_m, Wtr_Year_CWDDist_Recent, Wtr_Year_CWDDist_Historic) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  arrange(meanSurv)

wl2_winter_surv_adapt <- winter_surv_2324 %>% filter(pop=="WL2" | pop=="IH")

adaptlagm1 <- glmer(WinterSurv ~ pop + (1|block), data=wl2_winter_surv_adapt, family = binomial)
summary(adaptlagm1) #WL2 sig < than IH
```

## Scatterplots
```{r}
#scatter plots
WYCD_recent <- winter_surv %>%
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Recent Water Year CD") +
  theme(text=element_text(size=25))

GD <- winter_surv %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Recent, Geographic_Dist) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Geographic Distance (m)") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

ED <- winter_surv %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Elevation Distance (m)") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

wl2_winter_surv_FIG_recent <- ggarrange("",WYCD_recent, GD,ED, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_WinterSurv_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots
WYCD_historic <- winter_surv %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_GD_Historical) %>% 
  summarise(meanSurv=mean(WinterSurv, na.rm = TRUE), semSurv=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Historical, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Historic Water Year CD") +
  theme(text=element_text(size=25))

wl2_winter_surv_FIG_historic <- ggarrange("", WYCD_historic, GD,ED, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_WinterSurv_SCATTERS_Historical.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
WYCD_prob_recent <- winter_surv_sub_dist %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(WinterSurv, na.rm = TRUE), semEst=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Recent Water Year Temp Dist") +
  theme(text=element_text(size=30))

GD_prob <- winter_surv_sub_dist %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(WinterSurv, na.rm = TRUE), semEst=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- winter_surv_sub_dist %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(WinterSurv, na.rm = TRUE), semEst=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

wl2_winter_surv_sub_FIG_recent <- ggarrange("", WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_WinterSurv_TmpSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
WYCD_prob_historic <- winter_surv_sub_dist %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(WinterSurv, na.rm = TRUE), semEst=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Historic Water Year Temp Dist") +
  theme(text=element_text(size=30))

wl2_winter_surv_sub_FIG_historic <- ggarrange("", WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_WinterSurv_TmpSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
WYCD_prob_recent <- winter_surv_sub_dist %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(WinterSurv, na.rm = TRUE), semEst=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Recent Water Year PPT Dist") +
  theme(text=element_text(size=30))

GD_prob <- winter_surv_sub_dist %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(WinterSurv, na.rm = TRUE), semEst=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- winter_surv_sub_dist %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(WinterSurv, na.rm = TRUE), semEst=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

wl2_winter_surv_sub_FIG_recent <- ggarrange("", WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_WinterSurv_PPTSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
WYCD_prob_historic <- winter_surv_sub_dist %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(WinterSurv, na.rm = TRUE), semEst=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Historic Water Year PPT Dist") +
  theme(text=element_text(size=30))

wl2_winter_surv_sub_FIG_historic <- ggarrange("", WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_WinterSurv_PPTSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
WYCD_prob_recent <- winter_surv_sub_dist %>%
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_TempPpt_Recent, Wtr_Year_TempPpt_Historic) %>% 
  summarise(meanEst=mean(WinterSurv, na.rm = TRUE), semEst=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempPpt_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Recent Water Year Temp + PPT Dist") +
  theme(text=element_text(size=30))

GD_prob <- winter_surv_sub_dist %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_TempPpt_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(WinterSurv, na.rm = TRUE), semEst=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- winter_surv_sub_dist %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(WinterSurv, na.rm = TRUE), semEst=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

wl2_winter_surv_sub_FIG_recent <- ggarrange("", WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_WinterSurv_TempPptSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
WYCD_prob_historic <- winter_surv_sub_dist %>% 
  filter(pop!="WR") %>% 
  group_by(pop, elev_m, Wtr_Year_TempPpt_Recent, Wtr_Year_TempPpt_Historic) %>% 
  summarise(meanEst=mean(WinterSurv, na.rm = TRUE), semEst=sem(WinterSurv, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempPpt_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Winter Surv", x="Historic Water Year Temp + PPT Dist") +
  theme(text=element_text(size=30))

wl2_winter_surv_sub_FIG_historic <- ggarrange("", WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_WinterSurv_TempPptSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

## Stats
### Scaling 
```{r}
#winter_surv %>% group_by(pop) %>% 
 # summarise(n=n()) %>% 
  #arrange(n)#WR only has 1 individual, remove 

wl2_winter_surv_scaled <- winter_surv %>% 
  filter(pop!="WR") %>% 
  mutate_at(c("Wtr_Year_GD_Recent","Wtr_Year_GD_Historical", 
              "Geographic_Dist"), scale) 

wl2_winter_surv_scaled_2324 <- winter_surv_2324 %>% 
  filter(pop!="WR") %>% 
  mutate_at(c("Wtr_Year_GD_Recent","Wtr_Year_GD_Historical", 
              "Geographic_Dist"), scale) 

#check predictors
unique(wl2_winter_surv_scaled$pop)
unique(wl2_winter_surv_scaled$mf)
unique(wl2_winter_surv_scaled$block)
```

```{r}
winter_surv_sub_dist_scaled <- winter_surv_sub_dist %>% 
  filter(pop!="WR") %>% 
  mutate_at(c("Wtr_Year_TempDist_Recent",  "Wtr_Year_PPTDist_Recent", 
                 "Wtr_Year_TempDist_Historic", "Wtr_Year_PPTDist_Historic",
              "Geographic_Dist"), scale) 

winter_surv_sub_dist_scaled_2324 <- winter_surv_sub_dist_2324 %>% 
  filter(pop!="WR") %>% 
  mutate_at(c("Wtr_Year_TempDist_Recent",  "Wtr_Year_PPTDist_Recent", 
                 "Wtr_Year_TempDist_Historic", "Wtr_Year_PPTDist_Historic",
              "Geographic_Dist"), scale) 
```

```{r}
winter_surv_tp_sub_dist_scaled <- winter_surv_sub_dist %>% 
  filter(pop!="WR") %>% 
  mutate_at(c("Wtr_Year_TempPpt_Recent", "Wtr_Year_TempPpt_Historic",
              "Geographic_Dist"), scale) 

winter_surv_tp_sub_dist_scaled_2324 <- winter_surv_sub_dist_2324 %>% 
  filter(pop!="WR") %>% 
  mutate_at(c("Wtr_Year_TempPpt_Recent", "Wtr_Year_TempPpt_Historic",
              "Geographic_Dist"), scale) 
```


### Basic Model Workflow 
```{r}
glmer.model_binomial <- 
  linear_reg() %>% 
  set_engine("glmer", family=binomial)

surv_wflow <- workflow() %>% 
  add_variables(outcomes = WinterSurv, predictors = c(pop, mf, block))

surv_fits <- tibble(wflow=list(
  pop = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ (1|pop))},
  
  pop.mf = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ (1|pop/mf))},
  
  pop.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ (1|pop) + (1|block))},
  
  pop.mf.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ (1|pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)

surv_fits_wl2 <- surv_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_winter_surv_scaled))
#mod_test <- glmer(WinterSurv ~ (1|pop/mf) + (1|block), data=wl2_winter_surv_scaled, family=binomial)
#summary(mod_test)
surv_fits_wl2 %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#model with just pop best by AIC and BIC, but nothing wrong with the full model, so use that 

surv_fits_wl2_2324 <- surv_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_winter_surv_scaled_2324))
#mod_test <- glmer(WinterSurv ~ (1|pop/mf) + (1|block), data=wl2_winter_surv_scaled_2324, family=binomial)
#summary(mod_test)
surv_fits_wl2_2324 %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#model with just pop best by AIC and BIC, but nothing wrong with the full model, so use that 
```

#### Test climate and geographic distance 
```{r}
surv_GD_wflow_wl2 <- workflow() %>%
  add_variables(outcomes = WinterSurv, predictors = c(pop, mf, block, contains("GD"), Geographic_Dist)) 

surv_GD_fits_wl2 <- tibble(wflow=list(
  pop.mf.block = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ (1|pop/mf) + (1|block))},
  
  WY_Recent = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ Wtr_Year_GD_Recent + Geographic_Dist + (1|pop/mf) + (1|block))},
  
  WY_Historical = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ Wtr_Year_GD_Historical + Geographic_Dist + (1|pop/mf) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = wl2_winter_surv_scaled_2324))

surv_GD_fits_wl2 %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#WY recent model the best 

surv_GD_fits_wl2 %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "GD") | term=="Geographic_Dist") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)

#mod_test <- glmer(WinterSurv ~ Wtr_Year_GD_Recent + Geographic_Dist + (1|pop/mf) + (1|block), data=wl2_winter_surv_scaled, family=binomial)
#summary(mod_test)
```

```{r}
surv_GD_wflow_wl2_sub <- workflow() %>%
  add_variables(outcomes = WinterSurv, predictors = c(pop, mf, block, contains("Dist"))) 

surv_GD_fits_wl2_sub <- tibble(wflow=list(
  pop.mf.block = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ (1|pop/mf) + (1|block))},
  
  WY_Recent = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ Wtr_Year_TempDist_Recent + Wtr_Year_PPTDist_Recent + Geographic_Dist + (1|pop/mf) + (1|block))},
  
  WY_Historical = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ Wtr_Year_TempDist_Historic + Wtr_Year_PPTDist_Historic + Geographic_Dist + (1|pop/mf) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = winter_surv_sub_dist_scaled_2324))

surv_GD_fits_wl2_sub %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)


surv_GD_fits_wl2_sub %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "Dist")) %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)
```

```{r}
surv_GD_wflow_wl2_sub_tp <- workflow() %>%
  add_variables(outcomes = WinterSurv, predictors = c(pop, mf, block, contains("TempPpt"), Geographic_Dist)) 

surv_GD_fits_wl2_sub_tp <- tibble(wflow=list(
  pop.mf.block = {surv_GD_wflow_wl2_sub_tp %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ (1|pop/mf) + (1|block))},
  
  WY_Recent = {surv_GD_wflow_wl2_sub_tp %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ Wtr_Year_TempPpt_Recent + Geographic_Dist + (1|pop/mf) + (1|block))},
  
  WY_Historical = {surv_GD_wflow_wl2_sub_tp %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ Wtr_Year_TempPpt_Historic + Geographic_Dist + (1|pop/mf) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = winter_surv_tp_sub_dist_scaled_2324))

surv_GD_fits_wl2_sub_tp %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)


surv_GD_fits_wl2_sub_tp %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "TempPpt") | term=="Geographic_Dist") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)
```


## Fitness \~ Size

### Load the size data & Combine with Survival 

#### Stem Diameter and Basal Branches from Annual Census 

```{r}
ann_cens_size <- read_csv("../output/AnnCens_Size_BothSites_Y1.csv")
names(ann_cens_size)

ann_cens_size_wl2 <- ann_cens_size %>% filter(Site=="WL2") %>% select(BedLoc:diam.mm, total.branch, Site) %>% rename(pop=parent.pop)
head(ann_cens_size_wl2)
winter_surv_annsize <- left_join(winter_surv, ann_cens_size_wl2)
summary(winter_surv_annsize)
```

#### Height and Leaf Length from 10/20 Census
```{r}
prewint_size <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_size_survey_20231020_corrected.csv") %>% 
  select(block, `bed- row`=bed.row, `bed- col`=bed.col, pop:long.leaf.cm) %>% 
  filter(!str_detect(mf, "buff")) %>% 
  mutate(mf=as.double(mf), rep=as.double(rep))
head(prewint_size)

winter_surv_Octsize <- left_join(winter_surv, prewint_size)
summary(winter_surv_Octsize)
```

#### Merge all 
```{r}
winter_surv_size <- left_join(winter_surv_Octsize, winter_surv_annsize)
```

### Figures of Survival \~ Size 
```{r}
winter_surv_size %>% 
  ggplot(aes(x=height.cm, y=WinterSurv)) +
  geom_point()

winter_surv_size %>% 
  ggplot(aes(x=long.leaf.cm, y=WinterSurv)) +
  geom_point()

winter_surv_size %>% 
  ggplot(aes(x=diam.mm, y=WinterSurv)) +
  geom_point()

winter_surv_size %>% #not enough variation in branch #?
  ggplot(aes(x=total.branch, y=WinterSurv)) +
  geom_point()
```

```{r}
winter_surv_size_pop_avgs <- winter_surv_size %>% 
  group_by(pop, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(WinterSurv)), 
            mean_Surv = mean(WinterSurv,na.rm=(TRUE)), sem_surv=sem(WinterSurv, na.rm=(TRUE)), 
            N_height = sum(!is.na(height.cm)), 
            mean_height.cm = mean(height.cm,na.rm=(TRUE)), sem_height.cm=sem(height.cm, na.rm=(TRUE)), 
            N_length = sum(!is.na(long.leaf.cm)),
            mean_long.leaf.cm=mean(long.leaf.cm, na.rm=(TRUE)), sem_long.leaf.cm=sem(long.leaf.cm, na.rm=TRUE),
            N_diam = sum(!is.na(diam.mm)), 
            mean_diam.mm = mean(diam.mm,na.rm=(TRUE)), sem_diam.mm=sem(diam.mm, na.rm=(TRUE)), 
            N_total.branch = sum(!is.na(total.branch)),
            mean_total.branch=mean(total.branch, na.rm=(TRUE)), sem_total.branch=sem(total.branch, na.rm=TRUE))

winter_surv_size_pop_avgs %>% arrange(N_height)
```
```{r}
winter_surv_size %>% 
  drop_na(WinterSurv, height.cm) %>% 
  group_by(pop, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(WinterSurv)), 
            mean_Surv = mean(WinterSurv,na.rm=(TRUE)), sem_surv=sem(WinterSurv, na.rm=(TRUE)), 
            N_height = sum(!is.na(height.cm)), 
            mean_height.cm = mean(height.cm,na.rm=(TRUE)), sem_height.cm=sem(height.cm, na.rm=(TRUE))) %>% 
  filter(N_height>2) %>% 
  ggplot(aes(x=mean_height.cm, y=mean_Surv, group=pop, color=elev_m)) +
  geom_point(size=4) +
  theme_classic() + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Height (cm)", y="Winter Survival", color="Elevation (m)", title="High Elevation Garden") +
  theme(text=element_text(size=25))
  
winter_surv_size %>% 
  drop_na(WinterSurv, long.leaf.cm) %>% 
  group_by(pop, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(WinterSurv)), 
            mean_Surv = mean(WinterSurv,na.rm=(TRUE)), sem_surv=sem(WinterSurv, na.rm=(TRUE)), 
            N_length = sum(!is.na(long.leaf.cm)),
            mean_long.leaf.cm=mean(long.leaf.cm, na.rm=(TRUE)), sem_long.leaf.cm=sem(long.leaf.cm, na.rm=TRUE)) %>% 
  filter(N_length>2) %>% 
  ggplot(aes(x=mean_long.leaf.cm, y=mean_Surv, group=pop, color=elev_m)) +
  geom_point(size=4) +
  scale_y_continuous(expand = c(0.01, 0.01)) +
  theme_classic() + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Leaf Length (cm)" ,y="Winter Survival", color="Elevation (m)", title="High Elevation Garden") +
  theme(text=element_text(size=25))

winter_surv_size %>% 
  drop_na(WinterSurv, diam.mm) %>% 
  group_by(pop, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(WinterSurv)), 
            mean_Surv = mean(WinterSurv,na.rm=(TRUE)), sem_surv=sem(WinterSurv, na.rm=(TRUE)), 
            N_diam = sum(!is.na(diam.mm)), 
            mean_diam.mm = mean(diam.mm,na.rm=(TRUE)), sem_diam.mm=sem(diam.mm, na.rm=(TRUE))) %>% 
  filter(N_diam>2) %>% 
  ggplot(aes(x=mean_diam.mm, y=mean_Surv, group=pop, color=elev_m)) +
  geom_point(size=4) +
  scale_y_continuous(expand = c(0.01, 0.01)) +
  theme_classic() + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Stem Diameter (mm)" ,y="Winter Survival", color="Elevation (m)", title="High Elevation Garden") +
  theme(text=element_text(size=25))

winter_surv_size %>% 
  drop_na(WinterSurv, total.branch) %>% 
  group_by(pop, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(WinterSurv)), 
            mean_Surv = mean(WinterSurv,na.rm=(TRUE)), sem_surv=sem(WinterSurv, na.rm=(TRUE)), 
            N_total.branch = sum(!is.na(total.branch)),
            mean_total.branch=mean(total.branch, na.rm=(TRUE)), sem_total.branch=sem(total.branch, na.rm=TRUE)) %>% 
  filter(N_total.branch>2) %>% 
  ggplot(aes(x=mean_total.branch, y=mean_Surv, group=pop, color=elev_m)) +
  geom_point(size=4) +
  scale_y_continuous(expand = c(0.01, 0.01)) +
  theme_classic() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Basal Branch N" ,y="Winter Survival", color="Elevation (m)", title="High Elevation Garden") +
  theme(text=element_text(size=25))
```

### Stats

Log Reg survival \~ size

#### Check for correlations between traits 
```{r}
size_normalized_wl2 <- winter_surv_size %>% 
  select(height.cm, long.leaf.cm, diam.mm, total.branch) %>% 
  drop_na(height.cm, long.leaf.cm, diam.mm, total.branch) %>% scale() #normalize the data so they're all on the same scale
#head(size_normalized_wl2)
cor.norm_wl2 = cor(size_normalized_wl2) #test correlations among the traits
cor.sig_wl2 <- cor.mtest(size_normalized_wl2, method="pearson") #test significance of corrs
corrplot(cor.norm_wl2, type = "upper",
         tl.srt = 45, p.mat = cor.sig_wl2$p, 
         sig.level = 0.05, insig="blank")  
#longest leaf and diam are 70% correlated, no other strong corrs
cor.norm_wl2
cor.sig_wl2$p
```

```{r}
#summary(winter_surv_size) different amounts of NAs depending on trait 
winter_surv_height_formod <- winter_surv_size %>% 
  drop_na(WinterSurv, height.cm) %>% 
  filter(pop!="WR")

winter_surv_length_formod <- winter_surv_size %>% 
  drop_na(WinterSurv, long.leaf.cm) %>% 
  filter(pop!="WR", pop!="LV3", pop!="FR", pop!="LVTR1")

winter_surv_diam_formod <- winter_surv_size %>% 
  drop_na(WinterSurv, diam.mm) %>% 
  filter(pop!="WR")

winter_surv_branch_formod <- winter_surv_size %>% 
  drop_na(WinterSurv, total.branch) %>% 
  filter(pop!="WR")
```

#### Height
```{r}
wl2_basiclogit_height <- glm(WinterSurv ~ height.cm, data = winter_surv_height_formod, family = "binomial")
summary(wl2_basiclogit_height)

wl2_logit_height2 <- glmer(WinterSurv ~ height.cm + (1|pop/mf) + (1|block), data = winter_surv_height_formod, family = binomial(link = "logit"))
summary(wl2_logit_height2) #height not significant in full model 
```

#### Length
```{r}
wl2_basiclogit_length <- glm(WinterSurv ~ long.leaf.cm, data = winter_surv_length_formod, family = "binomial")
summary(wl2_basiclogit_length)

wl2_logit_length2 <- glmer(WinterSurv ~ long.leaf.cm + (1|pop/mf) + (1|block), data = winter_surv_length_formod, family = binomial(link = "logit"))
summary(wl2_logit_length2)
#positive effect of leaf length in both models 
```

#### Stem Diameter 
```{r}
wl2_basiclogit_diam <- glm(WinterSurv ~ diam.mm, data = winter_surv_diam_formod, family = "binomial")
summary(wl2_basiclogit_diam)

wl2_logit_diam2 <- glmer(WinterSurv ~ diam.mm + (1|pop/mf) + (1|block), data = winter_surv_diam_formod, family = binomial(link = "logit"))
summary(wl2_logit_diam2)
#positive effect of stem diam in both models
```

#### Basal Branches
```{r}
wl2_basiclogit_branch <- glm(WinterSurv ~ total.branch, data = winter_surv_branch_formod, family = "binomial")
summary(wl2_basiclogit_branch)

wl2_logit_branch2 <- glmer(WinterSurv ~ total.branch + (1|pop/mf) + (1|block), data = winter_surv_branch_formod, family = binomial(link = "logit"))
summary(wl2_logit_branch2)
#no effect of branch # 
```
