---
title: "WL2_Growth_Climate"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# WL2 Growth + Climate

Add climate to best base model

-   Weekly climate as predictor of K

-   Summary of climate as predictor of Hmax and maybe delta

## Libraries

```{r}
library(tidyverse)
library(brms)
library(zoo) #rollmean
```

## Load the growth data

```{r}
all_wl2_size_all_nobuffers <- read_csv("../output/WL2_Traits/WL2-2023_Size_Combined.csv")
```

## Pop Location Info

```{r}
wl2_gowers_2023 <- read_csv("../output/Climate/Gowers_WL2.csv") %>% 
  select(parent.pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) 
head(wl2_gowers_2023)
```

Merge

```{r}
all_wl2_size_loc <- left_join(all_wl2_size_all_nobuffers, wl2_gowers_2023) %>% 
  mutate(day=if_else(survey_date=="2023-07-03" | survey_date=="2023-07-06", 0, #day 0 = pre-transplant
                 as.numeric(survey_date-min(survey_date)-7))) #days in the field starts on 7/11 (day all plants taken to the field) rather than the exact planting date 
head(all_wl2_size_loc)
```

## Data set for the model (remove WV and shrinkage months)

```{r}
all_wl2_size_for_models <- all_wl2_size_loc %>% 
  filter(parent.pop!="WV") %>% 
  group_by(Genotype) %>% 
  mutate(height_next = lead(height.cm, order_by = survey_date),
         height_diff=height_next-height.cm,
         shrink_day=if_else(height_diff>-2, NA, day),
         min_shrink_day=min(shrink_day, na.rm = TRUE)) %>% #deals with problem of if something shrinks and then starts to grow again
  filter(day<min_shrink_day | min_shrink_day==Inf) %>% 
  ungroup() %>% 
  drop_na(height.cm)
```

## Read in soil temp

```{r}
temp_WL2 <- read_csv("../input/WL2_Data/WL2_2022_2023_iButton_Data_Corrected.csv") %>%
  select(-`...3`) %>%
  mutate(Date_Time = mdy_hm(Date_Time)) %>%
  filter(Date_Time > ymd("2023-07-06"))
head(temp_WL2)
skimr::skim(temp_WL2)
```

Average across beds.

```{r}
temp_WL2_summary <- temp_WL2 %>%
  group_by(Date_Time) %>% 
  summarise(AvgSoilTemp=mean(SoilTemp, na.rm=TRUE)) %>% #avg across beds 
  mutate(Date=as.Date(Date_Time)) %>%
  filter(Date != min(Date), Date != max(Date)) %>% # trim ragged ends
  group_by(Date) %>% #summmarise hourly data by date
  summarize(
    min_temp_d = min(AvgSoilTemp),
    max_temp_d = max(AvgSoilTemp),
    mean_temp_d = mean(AvgSoilTemp)
  ) %>%
  mutate(
    across(ends_with("temp_d"), \(x) rollmean(x, k = 7, align = "right", fill = NA), .names="{.col}1_7"),
    across(ends_with("temp_d"), \(x) rollmean(x, k = 13, align = "right", fill = NA), .names="{.col}1_13") # 13 so I can get the first survey date
  ) 

temp_WL2_summary
```

## Read in soil moisture

```{r}
moisture_WL2 <- read_csv("../input/WL2_Data/WL2_2023_Bed_C_Soil_Moisture_Corrected.csv") %>%
  mutate(Date_Time = mdy_hm(Date_Time)) 
head(moisture_WL2)
skimr::skim(moisture_WL2)
```

Will average across the ports

```{r}
moisture_WL2_summary <- moisture_WL2 %>%
  mutate(Port_1=if_else(Port_1<0, 0, Port_1),
         Port_2=if_else(Port_2<0, 0, Port_2),
         Port_3=if_else(Port_3<0, 0, Port_3),
         Port_4=if_else(Port_4<0, 0, Port_4),
         Port_5=if_else(Port_5<0, 0, Port_5)) %>%  #convert negatives to zeros (negatives result from extremely dry soil --> air pockets)
  rowwise() %>%
  mutate(s_moisture = mean(c_across(-Date_Time)) ) %>%
  select(Date_Time, s_moisture) %>%
  mutate(Date=as.Date(Date_Time)) %>%
  group_by(Date) %>%
  summarize(
    mean_moisture_d = mean(s_moisture)
  ) %>%
  mutate(
    s_moisture_1_7 = rollmean(mean_moisture_d, k = 7, align = "right", fill = "extend"),
    s_moisture_1_13 = rollmean(mean_moisture_d, k = 13, align = "right", fill = "extend")
  )

moisture_WL2_summary
```

## add climate data

```{r}
all_wl2_size_for_models_climate <- all_wl2_size_for_models %>% 
  left_join(temp_WL2_summary, by = c("survey_date" = "Date")) %>%
  left_join(moisture_WL2_summary, by = c("survey_date" = "Date")) %>% 
  filter(day>0) #remove day pre-field size b/c no climate data 

all_wl2_size_for_models_climate %>% 
  ggplot(aes(group=Genotype, x=survey_date, y=height.cm, col=elev_m)) + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_line() + facet_wrap(~parent.pop, scales = "free")
#without day 0, min is different...
```

## Priors

```{r}
prior1_wl2 <- c(set_prior("normal(5,3)", nlpar="Hmin"),
            set_prior("normal(60,15)",  nlpar="Hmax"),
            set_prior("gamma(2,2)", nlpar = "k", lb = 0), #gamma constrains to be > 0 
            set_prior("gamma(20,3)", nlpar = "delta", lb = 0)
)
            #first number = avg, second number = the space around that average to search 
```

## Base Models
```{r}
#best base model:
base_f1 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin + k + delta ~ 1,
                  Hmax ~ (1|parent.pop),
                  nl=TRUE)

base_f2 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  k + delta ~ 1,
                  Hmin + Hmax ~ (1|parent.pop),
                  nl=TRUE)
```


## Temp Models

```{r}
temp_f2 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  delta ~ 1,
                  k ~ min_temp_d1_7,
                  Hmin + Hmax ~ (1|parent.pop),
                  nl=TRUE)

temp_f3 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  delta ~ 1,
                  k ~ mean_temp_d1_7,
                  Hmin + Hmax ~ (1|parent.pop),
                  nl=TRUE)

temp_f4 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  delta ~ 1,
                  k ~ max_temp_d1_7,
                  Hmin + Hmax ~ (1|parent.pop),
                  nl=TRUE)

temp_f5 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  delta ~ 1,
                  k ~ min_temp_d1_13, 
                  Hmin + Hmax ~ (1|parent.pop),
                  nl=TRUE)

temp_f6 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  delta ~ 1,
                  k ~ mean_temp_d1_13,
                  Hmin + Hmax ~ (1|parent.pop),
                  nl=TRUE)

temp_f7 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  delta ~ 1,
                  k ~ max_temp_d1_13,
                  Hmin + Hmax ~ (1|parent.pop),
                  nl=TRUE)
```

## Base Model Fits

### Base model 1 - Hmax vary by pop 

```{r}
base_fit1 <- brm(formula=base_f1, data=all_wl2_size_for_models_climate, prior=prior1_wl2, iter = 3800, cores=4)
saveRDS(base_fit1, file = "brm_object_wl2_base1.rds")
summary(base_fit1)
plot(base_fit1, nvariables = 3, ask=FALSE)

pred.df <- expand_grid(day=min(all_wl2_size_for_models_climate$day):max(all_wl2_size_for_models_climate$day),
                       parent.pop=unique(all_wl2_size_for_models_climate$parent.pop))
base_fit1.predictions <- cbind(pred.df, prediction=predict(base_fit1, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_wl2_size_for_models_climate, by=c("day", "parent.pop"))

base_fit1.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop, scales = "free")
ggsave("../output/WL2_GrowthClim_Predictions_Hmax.png", width = 14, height = 8, units = "in")
```

### Base model 2 - Hmin + Hmax vary by pop

```{r}
base_fit2 <- brm(formula=base_f2, data=all_wl2_size_for_models_climate, prior=prior1_wl2, iter = 3800, cores=4)
saveRDS(base_fit2, file = "brm_object_wl2_base2.rds")
summary(base_fit2)
plot(base_fit2, nvariables = 3, ask=FALSE)

pred.df <- expand_grid(day=min(all_wl2_size_for_models_climate$day):max(all_wl2_size_for_models_climate$day),
                       parent.pop=unique(all_wl2_size_for_models_climate$parent.pop))
base_fit2.predictions <- cbind(pred.df, prediction=predict(base_fit2, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_wl2_size_for_models_climate, by=c("day", "parent.pop"))

base_fit2.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop, scales = "free")
ggsave("../output/WL2_GrowthClim_Predictions_HmaxHmin.png", width = 14, height = 8, units = "in")
```

### Model comparison 
```{r}
base_fit1 <- add_criterion(base_fit1, "loo")
base_fit2 <- add_criterion(base_fit2, "loo")
loo_compare(base_fit1, base_fit2)
#model with Hmin is better 
```

## Temp Model Fits

### Min D1-7 (fit2)

```{r}
all_wl2_size_for_models_temp_fit2 <- all_wl2_size_for_models_climate %>% drop_na(min_temp_d1_7)

temp_fit2 <- brm(formula=temp_f2, data=all_wl2_size_for_models_temp_fit2, prior=prior1_wl2, iter = 3800, cores=4)
#Warning: There were 257 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10. See https://mc-stan.org/misc/warnings.html#maximum-treedepth-exceededWarning: Examine the pairs() plot to diagnose sampling problems
summary(temp_fit2) #some ESS kind of low 
plot(temp_fit2, nvariables = 3, ask=FALSE) #some Hmax chain weirdness 
```

### Mean D1-7 (fit3)

```{r}
all_wl2_size_for_models_temp_fit3 <- all_wl2_size_for_models_climate %>% drop_na(mean_temp_d1_7)

temp_fit3 <- brm(formula=temp_f3, data=all_wl2_size_for_models_temp_fit3, prior=prior1_wl2, iter = 3800, cores=4)

summary(temp_fit3)
plot(temp_fit3, nvariables = 3, ask=FALSE)

temp_fit3.predictions <- cbind(pred.df, prediction=predict(temp_fit3, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_wl2_size_for_models, by=c("day", "parent.pop"))

temp_fit3.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)
```

### Max D1-7 (fit4)

```{r}
all_wl2_size_for_models_temp_fit4 <- all_wl2_size_for_models_climate %>% drop_na(max_temp_d1_7)

temp_fit4 <- brm(formula=temp_f4, data=all_wl2_size_for_models_temp_fit4, prior=prior1_wl2, iter = 3800, cores=4)

summary(temp_fit4)
plot(temp_fit4, nvariables = 3, ask=FALSE)

temp_fit4.predictions <- cbind(pred.df, prediction=predict(temp_fit4, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_wl2_size_for_models, by=c("day", "parent.pop"))

temp_fit4.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)
```

### Min D1-13 (fit5)

```{r}
all_wl2_size_for_models_temp_fit5 <- all_wl2_size_for_models_climate %>% drop_na(min_temp_d1_13)

temp_fit5 <- brm(formula=temp_f5, data=all_wl2_size_for_models_temp_fit5, prior=prior1_wl2, iter = 3800, cores=4)

summary(temp_fit5)
plot(temp_fit5, nvariables = 3, ask=FALSE)

temp_fit5.predictions <- cbind(pred.df, prediction=predict(temp_fit5, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_wl2_size_for_models, by=c("day", "parent.pop"))

temp_fit5.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)
```

### Mean D1-13 (fit6)

```{r}
all_wl2_size_for_models_temp_fit6 <- all_wl2_size_for_models_climate %>% drop_na(meam_temp_d1_13)

temp_fit6 <- brm(formula=temp_f6, data=all_wl2_size_for_models_temp_fit6, prior=prior1_wl2, iter = 3800, cores=4)

summary(temp_fit6)
plot(temp_fit6, nvariables = 3, ask=FALSE)

temp_fit6.predictions <- cbind(pred.df, prediction=predict(temp_fit6, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_wl2_size_for_models, by=c("day", "parent.pop"))

temp_fit6.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)
```

### Max D1-13 (fit7)

```{r}
all_wl2_size_for_models_temp_fit7 <- all_wl2_size_for_models_climate %>% drop_na(max_temp_d1_13)

temp_fit7 <- brm(formula=temp_f7, data=all_wl2_size_for_models_temp_fit7, prior=prior1_wl2, iter = 3800, cores=4)

summary(temp_fit7)
plot(temp_fit7, nvariables = 3, ask=FALSE)

temp_fit7.predictions <- cbind(pred.df, prediction=predict(temp_fit7, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_wl2_size_for_models, by=c("day", "parent.pop"))

temp_fit7.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)
```

## Temp Model Comparison

```{r}
base_fit1 <- add_criterion(base_fit1, "loo")
base_fit2 <- add_criterion(base_fit2, "loo")
temp_fit2 <- add_criterion(temp_fit2, "loo")
temp_fit3 <- add_criterion(temp_fit3, "loo")
temp_fit4 <- add_criterion(temp_fit4, "loo")
temp_fit5 <- add_criterion(temp_fit5, "loo") 
temp_fit6 <- add_criterion(temp_fit6, "loo") 
temp_fit7 <- add_criterion(temp_fit7, "loo") 
loo_compare(base_fit1, base_fit2, temp_fit2, temp_fit3, temp_fit4, temp_fit5, temp_fit6, temp_fit7)
```

## Moisture Models

```{r}
moisture_f2 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  delta ~ 1,
                  k ~ s_moisture_1_7, 
                  Hmin + Hmax ~ (1|parent.pop),
                  nl=TRUE)

moisture_f3 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  delta ~ 1,
                  k ~ s_moisture_1_13,
                  Hmin + Hmax ~ (1|parent.pop),
                  nl=TRUE)
```

## Moisture Model Fits

### Min D1-7 (fit2)

```{r}
all_wl2_size_for_models_moisture_fit2 <- all_wl2_size_for_models_climate %>% drop_na(s_moisture_1_7)

moisture_fit2 <- brm(formula=moisture_f2, data=all_wl2_size_for_models_moisture_fit2, prior=prior1_wl2, iter = 3800, cores=4)

summary(moisture_fit2)
plot(moisture_fit2, nvariables = 3, ask=FALSE)

moisture_fit2.predictions <- cbind(pred.df, prediction=predict(moisture_fit2, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_wl2_size_for_models, by=c("day", "parent.pop"))

moisture_fit2.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)
```

### Mean D1-13 (fit3)

```{r}
all_wl2_size_for_models_moisture_fit3 <- all_wl2_size_for_models_climate %>% drop_na(s_moisture_1_13)

moisture_fit3 <- brm(formula=moisture_f3, data=all_wl2_size_for_models_moisture_fit3, prior=prior1_wl2, iter = 3800, cores=4)

summary(moisture_fit3)
plot(moisture_fit3, nvariables = 3, ask=FALSE)

moisture_fit3.predictions <- cbind(pred.df, prediction=predict(moisture_fit3, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_wl2_size_for_models, by=c("day", "parent.pop"))

moisture_fit3.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)
```

## Moisture Model Comparison

```{r}
base_fit1 <- add_criterion(base_fit1, "loo")
base_fit2 <- add_criterion(base_fit2, "loo")
moisture_fit2 <- add_criterion(moisture_fit2, "loo")
moisture_fit3 <- add_criterion(moisture_fit3, "loo")
loo_compare(base_fit1, base_fit2, moisture_fit2, moisture_fit3)
```
