---
title: "WL2_Surv_to_Rep_Y2"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Survival to Bolting Year 2 at WL2

To Do:

-   Look at relationship between size and survival

    -   size post winter

    -   stem diameter y2

    -   total basal branches y2

    -   canopy area y2 - 2 months in?

<!-- -->

-   Standard error correction on scatter plots

## Libraries

```{r}
library(tidyverse)
library(magrittr)
#conflicted::conflicts_prefer(dplyr::filter)
library(viridis) #for pretty colors
library(ggrepel) #for non-overlapping labels on plots
library(ggdist) #visualizations of distributions and uncertainty 
library(ggpubr)

#library(sp) #for calculating geographic distance 
library(geosphere) #for calculating geographic distance
library(corrplot) #plotting correlations  
library(rstatix) #performing cor_test

library(lmerTest) #mixed models
conflicted::conflicts_prefer(lmerTest::lmer)
library(broom.mixed)
library(tidymodels)
library(furrr) #Apply Mapping Functions in Parallel using Futures
tidymodels_prefer()
# install.packages("multilevelmod")
library(multilevelmod)

library(brms)
#conflicted::conflicts_prefer(brms::ar)
#conflicted::conflicts_prefer(dplyr::combine)
#conflicted::conflicts_prefer(brms::dstudent_t)
library(tidybayes) #for extracting and visiaulizing brms model output 
library(modelr) #for data grid

sem <- function(x, na.rm=FALSE) {           #for caclulating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 

cbbPalette2 <- c("#E69F00","#000000", "#56B4E9","#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
timepd_palette <- c("#56B4E9","#D55E00")

options(mc.cores = parallel::detectCores())
```

## Year 2 Pop Info

```{r}
wl2_y2_pops <- read_csv("../input/WL2_Data/Final_2023_2024_Pop_Loc_Info.csv") %>%
  select(Pop.Type:unique.ID) %>% 
  filter(Pop.Type=="2023-survivor") %>% 
  select(Pop.Type, loc:bed, row=bedrow, col=bedcol, pop:unique.ID)

wl2_blocks <- read_csv("../input/WL2_Data/CorrectedCSVs/WL2_mort_pheno_20231020_corrected.csv") %>% 
  unite(BedLoc, bed:bed.col, sep="_", remove = FALSE) %>% 
  filter(BedLoc!="K_5_C") %>% #get rid of duplicate locations
  select(block, pop, mf, rep) %>% #add in block info 
  mutate(mf=as.double(mf), rep=as.double(rep)) #convert to num

#wl2_blocks %>% rowwise() %>%  #checking if mf and rep can be converted to numeric (all buffers)
#  filter(!is.na(mf)) %>%  
#  filter(is.na(as.numeric(mf)))

wl2_y2_pops_blocks <- left_join(wl2_y2_pops, wl2_blocks)
```

## Year 2 Surv Data

```{r}
wl2_20241023 <- read_csv("../input/WL2_Data/WL2_mort_pheno_20241023_corrected.csv") %>% #note this has 2023 and 2024 plants
  select(-block)
wl2_surv_y2 <- left_join(wl2_y2_pops_blocks, wl2_20241023) %>%  
  rename(Genotype=unique.ID)
#unique(wl2_surv_y2$pop) #only 10 pops with winter surv 
```

## Gower's Distance

```{r}
garden_climate <- read_csv("../output/Climate/flint_climate_UCDpops.csv") %>% 
  filter(parent.pop=="WL2_Garden" | parent.pop=="UCD_Garden") %>% 
  select(parent.pop:Long) %>% 
  distinct()
garden_climate
#UCD LAT/LONG = 38.53250, -121.7830
#WL2 Lat/Long = 38.82599, -120.2509

wl2_gowers_2024 <- read_csv("../output/Climate/Gowers_WL2_2024.csv") %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) %>% 
  mutate(WL2_Lat=38.82599, WL2_Long=-120.2509, WL2_Elev=2020) %>% 
  mutate(Geographic_Dist=distHaversine(cbind(WL2_Long, WL2_Lat), cbind(Long, Lat)),
         Elev_Dist=elev_m-WL2_Elev) %>% # Calculate the distance using the haversine formula
  rename(pop=parent.pop)
```

## Climate Subtraction Distance

```{r}
wl2_wtr_year_sub_recent_2024 <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_2024_Recent.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Recent=ann_tmean_dist, Wtr_Year_PPTDist_Recent=ann_ppt_dist)
wl2_wtr_year_sub_historic_2024 <- read_csv("../output/Climate/full_year_Subtraction_Dist_from_Home_WL2_2024_Historical.csv") %>% 
  select(parent.pop, Wtr_Year_TempDist_Historic=ann_tmean_dist, Wtr_Year_PPTDist_Historic=ann_ppt_dist)
wl2_grwssn_sub_recent_2024 <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_Home_WL2_2024_Recent.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Recent=ann_tmean_dist, GrwSsn_PPTDist_Recent=ann_ppt_dist)
wl2_grwssn_sub_historic_2024 <- read_csv("../output/Climate/grwssn_Subtraction_Dist_from_Home_WL2_2024_Historical.csv") %>% 
  select(parent.pop, GrwSsn_TempDist_Historic=ann_tmean_dist, GrwSsn_PPTDist_Historic=ann_ppt_dist)
wl2_sub_dist_2024 <- wl2_wtr_year_sub_recent_2024 %>% 
  left_join(wl2_wtr_year_sub_historic_2024) %>% 
  left_join(wl2_grwssn_sub_recent_2024) %>% 
  left_join(wl2_grwssn_sub_historic_2024) %>% 
  rename(pop=parent.pop) %>% 
  left_join(wl2_gowers_2024)
```

## Calculate Survival to Bolting

```{r}
wl2_surv_y2 %>% filter(!is.na(bud.date), is.na(fruit.date)) #some plants initiated reproduction but did not make fruits
#1 plant with a last flower date, but no fruit date, this plant did not make any fruits per the notes 
wl2_surv_y2 %>% filter(!is.na(bud.date), !is.na(death.date))  #bud date and death date for a lot of plants
wl2_surv_y2 %>% filter(is.na(bud.date), !is.na(flower.date)) #3 plants without a bud date, but with a flower date
wl2_surv_y2 %>% filter(is.na(bud.date), !is.na(fruit.date)) #4 plants without a bud date, but with a fruit date  
wl2_surv_y2 %>% filter(is.na(bud.date), !is.na(last.FL.date)) #same plants as above
wl2_surv_y2 %>% filter(is.na(bud.date), !is.na(last.FR.date)) #same plants as above 

wl2_surv_to_rep_y2 <- wl2_surv_y2 %>% 
  left_join(wl2_gowers_2024) %>% 
  mutate(bud.date=if_else(Genotype=="SC_6_15" | Genotype=="SC_5_6" | 
                            Genotype=="CC_9_6" | Genotype=="WL2_8_4",
                          "Missed", bud.date)) %>% #add in bud date for plants with a later rep date 
  mutate(SurvtoRep_y2=if_else(is.na(bud.date), 0, 1)) %>% 
  select(Pop.Type:block, elevation.group:Wtr_Year_GD_Historical, Geographic_Dist, Elev_Dist, bud.date, death.date, SurvtoRep_y2) 

wl2_surv_to_rep_y2 %>% group_by(pop) %>% summarise(n=n()) %>% arrange(n)
#LV1, SQ1, and WR only have 1 indiv alive in 2024

#write_csv(wl2_surv_to_rep_y2, "../output/WL2_Traits/WL2_Surv_to_Rep_Y2.csv")
```

```{r}
wl2_surv_to_rep_y2_sub_dist <- wl2_surv_to_rep_y2 %>% 
  select(loc:block, SurvtoRep_y2) %>% 
  left_join(wl2_sub_dist_2024)

unique(wl2_surv_to_rep_y2_sub_dist$block)
unique(wl2_surv_to_rep_y2_sub_dist$pop)
unique(wl2_surv_to_rep_y2_sub_dist$mf)
```

### Bar Plots

```{r}
wl2_surv_to_rep_y2 %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(SurvtoRep_y2, na.rm = TRUE), semSurv=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  filter(pop != "LV1", pop !="SQ1", pop !="WR") %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=GrwSsn_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Parent Population", fill="Growth Season Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/WL2_Traits/WL2_SurvtoRepY2_GrwSsn_GD_Recent.png", width = 12, height = 8, units = "in")

wl2_surv_to_rep_y2 %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(SurvtoRep_y2, na.rm = TRUE), semSurv=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  filter(pop != "LV1", pop !="SQ1", pop !="WR") %>% 
  ggplot(aes(x=fct_reorder(pop, meanSurv), y=meanSurv, fill=Wtr_Year_GD_Recent)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.2, position = 
                  position_dodge(0.75)) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Parent Population", fill="Water Year Gowers") +
  scale_fill_viridis(option="mako", direction = -1) +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
#ggsave("../output/WL2_Traits/WL2_SurvtoRepY2_Wtr_Year_GD_Recent.png", width = 12, height = 8, units = "in")
```

### Scatterplots

```{r}
#scatter plots
GSCD_recent <- wl2_surv_to_rep_y2 %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(SurvtoRep_y2, na.rm = TRUE), semSurv=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Recent, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Recent Growth Season CD") +
  theme(text=element_text(size=25))

WYCD_recent <- wl2_surv_to_rep_y2 %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(SurvtoRep_y2, na.rm = TRUE), semSurv=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Recent Water Year CD") +
  theme(text=element_text(size=25))

GD <- wl2_surv_to_rep_y2 %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent, Geographic_Dist) %>% 
  summarise(meanSurv=mean(SurvtoRep_y2, na.rm = TRUE), semSurv=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Geographic Distance (m)") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

ED <- wl2_surv_to_rep_y2 %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanSurv=mean(SurvtoRep_y2, na.rm = TRUE), semSurv=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Elevation Distance (m)") +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))

wl2_surv_to_rep_y2_FIG <- ggarrange(GSCD_recent, WYCD_recent, GD, ED, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_SurvtoRep_y2_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots
GSCD_historic <- wl2_surv_to_rep_y2 %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, GrwSsn_GD_Historical, Wtr_Year_GD_Historical) %>% 
  summarise(meanSurv=mean(SurvtoRep_y2, na.rm = TRUE), semSurv=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_GD_Historical, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Historic Growth Season CD") +
  theme(text=element_text(size=25))

WYCD_historic <- wl2_surv_to_rep_y2 %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, GrwSsn_GD_Historical, Wtr_Year_GD_Historical) %>% 
  summarise(meanSurv=mean(SurvtoRep_y2, na.rm = TRUE), semSurv=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Historical, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Historic Water Year CD") +
  theme(text=element_text(size=25))

wl2_surv_to_rep_y2_FIG <- ggarrange(GSCD_historic, WYCD_historic, GD, ED, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_SurvtoRep_y2_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

#### Directional Distance

```{r}
#scatter plots - recent
GSCD_prob_recent <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Recent Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Recent Water Year Temp Dist") +
  theme(text=element_text(size=30))

GD_prob <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, Wtr_Year_TempDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

WL2_surv_to_rep_y2_sub_FIG_prob_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_SurvtoRep_y2_TmpSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, GrwSsn_TempDist_Recent, GrwSsn_TempDist_Historic) %>% 
  summarise(meanEst=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Historic Growth Season Temp Dist") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Historic Water Year Temp Dist") +
  theme(text=element_text(size=30))

WL2_surv_to_rep_y2_sub_FIG_prob_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_SurvtoRep_y2_TmpSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - recent
GSCD_prob_recent <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=0.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Recent Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_prob_recent <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Recent Water Year PPT Dist") +
  theme(text=element_text(size=30))

GD_prob <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, Wtr_Year_PPTDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

ED_prob <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, Elev_Dist) %>% 
  summarise(meanEst=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Elev_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Elevation Distance (m)") +
  theme(text=element_text(size=30))

WL2_surv_to_rep_y2_sub_FIG_prob_recent <- ggarrange(GSCD_prob_recent, WYCD_prob_recent, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_SurvtoRep_y2_PPTSubDist_SCATTERS_Recent.png", width = 24, height = 18, units = "in")
```

```{r}
#scatter plots - historic
GSCD_prob_historic <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, GrwSsn_PPTDist_Historic) %>% 
  summarise(meanEst=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=GrwSsn_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Historic Growth Season PPT Dist") +
  theme(text=element_text(size=30))

WYCD_prob_historic <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Historic, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Surv-to-Rep Y2", x="Historic Water Year PPT Dist") +
  theme(text=element_text(size=30))

WL2_surv_to_rep_y2_sub_FIG_prob_historic <- ggarrange(GSCD_prob_historic, WYCD_prob_historic, GD_prob, ED_prob, ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_SurvtoRep_y2_PPTSubDist_SCATTERS_Historic.png", width = 24, height = 18, units = "in")
```

#### Figure for paper

```{r}
GD_prob <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, GrwSsn_PPTDist_Recent, Wtr_Year_PPTDist_Recent, Geographic_Dist) %>% 
  summarise(meanEst=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Geographic_Dist, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.2, linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Survival to Budding Y2", x="Geographic Distance (m)") +
  theme(text=element_text(size=30), axis.text.x = element_text(angle = 45,  hjust = 1))

WYCD_recent <- wl2_surv_to_rep_y2 %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, GrwSsn_GD_Recent, Wtr_Year_GD_Recent) %>% 
  summarise(meanSurv=mean(SurvtoRep_y2, na.rm = TRUE), semSurv=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_GD_Recent, y=meanSurv, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanSurv-semSurv,ymax=meanSurv+semSurv),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Survival to Budding Y2", x="Recent Water Year CD") +
  theme(text=element_text(size=25))

WYtemp_prob_recent <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, Wtr_Year_TempDist_Recent, Wtr_Year_TempDist_Historic) %>% 
  summarise(meanEst=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_TempDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Survival to Budding Y2", x="Recent Water Year Temp Dist (°C)") +
  theme(text=element_text(size=30))

WYppt_prob_recent <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  group_by(pop, elev_m, Wtr_Year_PPTDist_Recent, Wtr_Year_PPTDist_Historic) %>% 
  summarise(meanEst=mean(SurvtoRep_y2, na.rm = TRUE), semEst=sem(SurvtoRep_y2, na.rm=TRUE)) %>% 
  ggplot(aes(x=Wtr_Year_PPTDist_Recent, y=meanEst, group = pop)) +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanEst-semEst,ymax=meanEst+semEst),width=.3,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Survival to Budding Y2", x="Recent Water Year PPT Dist (mm)") +
  theme(text=element_text(size=30))

WL2_surv_to_rep_y2_for_paper <- ggarrange(GD_prob, WYCD_recent, 
                                        WYtemp_prob_recent, WYppt_prob_recent, 
                                        labels = c("A)", "B)", "C)", "D)"), 
                                        font.label = list(size=30, face = "plain"), 
                                        ncol=2, nrow=2) 
ggsave("../output/WL2_Traits/WL2_SurvtoRep_y2_SCATTERS_Summary_Recent.png", width = 24, height = 18, units = "in")
```

## Stats

### Scaling

```{r}
wl2_surv_to_rep_y2 %>% 
  group_by(pop) %>% 
  summarise(n=n()) %>% 
  arrange(n) #LV1, SQ1, WR only 1 indiv, remove 

wl2_surv_to_rep_y2_scaled <- wl2_surv_to_rep_y2 %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  mutate_at(c("GrwSsn_GD_Recent","Wtr_Year_GD_Recent",                                                           "GrwSsn_GD_Historical","Wtr_Year_GD_Historical","Geographic_Dist"),
                                                            scale) 
```

```{r}
wl2_surv_to_rep_y2_sub_dist %>% 
  group_by(pop) %>% 
  summarise(n=n()) %>% 
  arrange(n) #LV1, SQ1, WR only 1 indiv, remove 

wl2_surv_to_rep_y2_scaled_sub <- wl2_surv_to_rep_y2_sub_dist %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% 
  mutate_at(c("Wtr_Year_TempDist_Recent",  "Wtr_Year_PPTDist_Recent", 
                 "Wtr_Year_TempDist_Historic", "Wtr_Year_PPTDist_Historic",
                 "GrwSsn_TempDist_Recent", "GrwSsn_PPTDist_Recent",
                 "GrwSsn_TempDist_Historic", "GrwSsn_PPTDist_Historic",
              "Geographic_Dist"), scale)

unique(wl2_surv_to_rep_y2_scaled_sub$pop)
unique(wl2_surv_to_rep_y2_scaled_sub$mf)
unique(wl2_surv_to_rep_y2_scaled_sub$block)
```

### Basic Model Workflow

```{r}
glmer.model_binomial <- 
  linear_reg() %>% 
  set_engine("glmer", family=binomial)

surv_wflow <- workflow() %>% 
  add_variables(outcomes = SurvtoRep_y2, predictors = c(pop, mf, block))

surv_fits <- tibble(wflow=list(
  pop = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ (1|pop))},
  
  pop.mf = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ (1|pop/mf))},
  
  pop.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ (1|pop) + (1|block))},
  
  pop.mf.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ (1|pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)

surv_fits_wl2 <- surv_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_surv_to_rep_y2_scaled))
#mod_test <- glmer(SurvtoRep_y2 ~ (1|pop/mf) + (1|block), data=wl2_surv_to_rep_y2_scaled, family=binomial)
#summary(mod_test)
#Warning: Model failed to converge with max|grad| = 0.0354704 (tol = 0.002, component 1)

surv_fits_wl2 %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#model with pop and block (but excluding mf) is best by AIC and BIC, convergence issues with mf included 
```

#### Test climate and geographic distance

```{r}
surv_GD_wflow_wl2 <- workflow() %>%
  add_variables(outcomes = SurvtoRep_y2, predictors = c(pop, mf, block, contains("GD"), Geographic_Dist)) 

surv_GD_fits_wl2 <- tibble(wflow=list(
  pop.block = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ (1|pop) + (1|block))},
  
  GS_Recent = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ GrwSsn_GD_Recent + Geographic_Dist + (1|pop) + (1|block))},
  
  GS_Historical = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ GrwSsn_GD_Historical + Geographic_Dist + (1|pop) + (1|block))},
  
  WY_Recent = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ Wtr_Year_GD_Recent + Geographic_Dist + (1|pop) + (1|block))},
  
  WY_Historical = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ Wtr_Year_GD_Historical + Geographic_Dist + (1|pop) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = wl2_surv_to_rep_y2_scaled))

surv_GD_fits_wl2 %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#water year models preferred by AIC 

surv_GD_fits_wl2 %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "GD") | term=="Geographic_Dist") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)
# recent water year = marginally sig, historical water year = sig, but historical water year had a singular boundary warning 

mod_test <- glmer(SurvtoRep_y2 ~ Wtr_Year_GD_Recent + Geographic_Dist + (1|pop) + (1|block), data=wl2_surv_to_rep_y2_scaled, family=binomial)
summary(mod_test)
#boundary (singular) fit: see help('isSingular') for water year recent and historical - when you add water year climate distance, pop explains little variation 
```

```{r}
surv_GD_wflow_wl2_sub <- workflow() %>%
  add_variables(outcomes = SurvtoRep_y2, predictors = c(pop, mf, block, contains("Dist"))) 

surv_GD_fits_wl2_sub <- tibble(wflow=list(
  pop.block = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ (1|pop) + (1|block))},
  
  GS_Recent = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ GrwSsn_TempDist_Recent + GrwSsn_PPTDist_Recent + Geographic_Dist + (1|pop) + (1|block))},
  
  GS_Historical = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ GrwSsn_TempDist_Historic + GrwSsn_PPTDist_Historic + Geographic_Dist + (1|pop) + (1|block))},
  
  WY_Recent = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ Wtr_Year_TempDist_Recent + Wtr_Year_PPTDist_Recent + Geographic_Dist + (1|pop) + (1|block))},
  
  WY_Historical = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ Wtr_Year_TempDist_Historic + Wtr_Year_PPTDist_Historic + Geographic_Dist + (1|pop) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = wl2_surv_to_rep_y2_scaled_sub))

surv_GD_fits_wl2_sub %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% arrange(AIC) %>% select(-wflow:-sigma)
#water year models preferred by AIC 

surv_GD_fits_wl2_sub %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "Dist")) %>%
  drop_na(p.value) %>%
  select(-wflow:-group)# %>%
#  arrange(p.value)
```

## Fitness \~ Size

### Load the size data & Combine with Survival - need to update 

#### Size (post winter)

```{r}
postwinter_size <- read_csv("../output/WL2_Traits/WL2_PostWinter_Size.csv")
names(postwinter_size)

wl2_surv_to_rep_y2_mossize <- left_join(wl2_surv_to_rep_y2, postwinter_size)
```

#### Stem Diameter, Basal Branches, and Canopy Area from Annual Census  

```{r}
wl2_ann_cens_2024 <- read_csv("../input/WL2_Data/WL2_Annual_Census_20241023_CanopyAdded_corrected.csv") #added canopy area for the rows were we skipped b/c we also measured it in the size survey 
wl2_ann_cens_2024_pops <- left_join(wl2_y2_pops_blocks, wl2_ann_cens_2024) %>%  
  rename(Genotype=unique.ID) %>% 
  select(Pop.Type:block, diam.mm, total.branch, overhd.diam, overhd.perp)

wl2_surv_to_rep_y2_annsize <- left_join(wl2_surv_to_rep_y2, wl2_ann_cens_2024_pops)
```

#### Merge all 
```{r}
wl2_surv_to_rep_y2_size <- left_join(wl2_surv_to_rep_y2_mossize, wl2_surv_to_rep_y2_annsize)
wl2_surv_to_rep_y2_size %>% filter(!is.na(diam.mm))
```

### Figures of Survival \~ Size 

```{r}
wl2_surv_to_rep_y2_size %>% 
  ggplot(aes(x=height.cm, y=SurvtoRep_y2)) +
  geom_point()

wl2_surv_to_rep_y2_size %>% 
  ggplot(aes(x=long.leaf.cm, y=SurvtoRep_y2)) +
  geom_point()

wl2_surv_to_rep_y2_size %>% #prob not enough var in surv
  ggplot(aes(x=diam.mm, y=SurvtoRep_y2)) +
  geom_point()

wl2_surv_to_rep_y2_size %>% #not enough variation in surv
  ggplot(aes(x=total.branch, y=SurvtoRep_y2)) +
  geom_point()

wl2_surv_to_rep_y2_size %>% #prob not enough var in surv
  ggplot(aes(x=overhd.diam, y=SurvtoRep_y2)) +
  geom_point()

wl2_surv_to_rep_y2_size %>% #prob not enough var in surv
  ggplot(aes(x=overhd.perp, y=SurvtoRep_y2)) +
  geom_point()

wl2_surv_to_rep_y2_size %>% filter(SurvtoRep_y2==0, !is.na(overhd.diam)) #only 6 dead indivs with size info (died the day of the census, likely )
```

```{r}
wl2_size_surv_pop_avgs <- wl2_surv_to_rep_y2_size %>% 
  group_by(pop, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(SurvtoRep_y2)), 
            mean_Surv = mean(SurvtoRep_y2,na.rm=(TRUE)), sem_surv=sem(SurvtoRep_y2, na.rm=(TRUE)), 
            N_height = sum(!is.na(height.cm)), 
            mean_height.cm = mean(height.cm,na.rm=(TRUE)), sem_height.cm=sem(height.cm, na.rm=(TRUE)), 
            N_length = sum(!is.na(long.leaf.cm)),
            mean_long.leaf.cm=mean(long.leaf.cm, na.rm=(TRUE)), sem_long.leaf.cm=sem(long.leaf.cm, na.rm=TRUE),
            N_diam = sum(!is.na(diam.mm)), 
            mean_diam.mm = mean(diam.mm,na.rm=(TRUE)), sem_diam.mm=sem(diam.mm, na.rm=(TRUE)), 
            N_total.branch = sum(!is.na(total.branch)),
            mean_total.branch=mean(total.branch, na.rm=(TRUE)), sem_total.branch=sem(total.branch, na.rm=TRUE),
            N_overhd.diam = sum(!is.na(overhd.diam)), 
            mean_overhd.diam = mean(overhd.diam,na.rm=(TRUE)), sem_overhd.diam=sem(overhd.diam, na.rm=(TRUE)), 
            N_overhd.perp = sum(!is.na(overhd.perp)),
            mean_overhd.perp=mean(overhd.perp, na.rm=(TRUE)), sem_overhd.perp=sem(overhd.perp, na.rm=TRUE))

wl2_size_surv_pop_avgs %>% arrange(N_height)
```

```{r}
wl2_surv_to_rep_y2_size %>% 
  drop_na(SurvtoRep_y2, height.cm) %>% 
  group_by(pop, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(SurvtoRep_y2)), 
            mean_Surv = mean(SurvtoRep_y2,na.rm=(TRUE)), sem_surv=sem(SurvtoRep_y2, na.rm=(TRUE)), 
            N_height = sum(!is.na(height.cm)), 
            mean_height.cm = mean(height.cm,na.rm=(TRUE)), sem_height.cm=sem(height.cm, na.rm=(TRUE))) %>% 
  filter(N_height>2) %>% 
  ggplot(aes(x=mean_height.cm, y=mean_Surv, group=pop, color=elev_m)) +
  geom_point(size=4) +
  theme_classic() + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Height (cm)", y="Survival to Budding Y2", color="Elevation (m)") +
  theme(text=element_text(size=25))
  
wl2_surv_to_rep_y2_size %>% 
  drop_na(SurvtoRep_y2, long.leaf.cm) %>% 
  group_by(pop, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(SurvtoRep_y2)), 
            mean_Surv = mean(SurvtoRep_y2,na.rm=(TRUE)), sem_surv=sem(SurvtoRep_y2, na.rm=(TRUE)), 
            N_length = sum(!is.na(long.leaf.cm)),
            mean_long.leaf.cm=mean(long.leaf.cm, na.rm=(TRUE)), sem_long.leaf.cm=sem(long.leaf.cm, na.rm=TRUE)) %>% 
  filter(N_length>2) %>% 
  ggplot(aes(x=mean_long.leaf.cm, y=mean_Surv, group=pop, color=elev_m)) +
  geom_point(size=4) +
  scale_y_continuous(expand = c(0.01, 0.01)) +
  theme_classic() + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Leaf Length (cm)" ,y="Survival to Budding Y2", color="Elevation (m)") +
  theme(text=element_text(size=25))

wl2_surv_to_rep_y2_size %>% 
  drop_na(SurvtoRep_y2, diam.mm) %>% 
  group_by(pop, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(SurvtoRep_y2)), 
            mean_Surv = mean(SurvtoRep_y2,na.rm=(TRUE)), sem_surv=sem(SurvtoRep_y2, na.rm=(TRUE)), 
            N_diam = sum(!is.na(diam.mm)), 
            mean_diam.mm = mean(diam.mm,na.rm=(TRUE)), sem_diam.mm=sem(diam.mm, na.rm=(TRUE))) %>% 
  filter(N_diam>2) %>% 
  ggplot(aes(x=mean_diam.mm, y=mean_Surv, group=pop, color=elev_m)) +
  geom_point(size=4) +
  scale_y_continuous(expand = c(0.01, 0.01)) +
  theme_classic() + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Stem Diameter (mm)" ,y="Survival to Budding Y2", color="Elevation (m)") +
  theme(text=element_text(size=25))

wl2_surv_to_rep_y2_size %>% 
  drop_na(SurvtoRep_y2, total.branch) %>% 
  group_by(pop, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(SurvtoRep_y2)), 
            mean_Surv = mean(SurvtoRep_y2,na.rm=(TRUE)), sem_surv=sem(SurvtoRep_y2, na.rm=(TRUE)), 
            N_total.branch = sum(!is.na(total.branch)),
            mean_total.branch=mean(total.branch, na.rm=(TRUE)), sem_total.branch=sem(total.branch, na.rm=TRUE)) %>% 
  filter(N_total.branch>2) %>% 
  ggplot(aes(x=mean_total.branch, y=mean_Surv, group=pop, color=elev_m)) +
  geom_point(size=4) +
  scale_y_continuous(expand = c(0.01, 0.01)) +
  theme_classic() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Basal Branch N" ,y="Survival to Budding Y2", color="Elevation (m)") +
  theme(text=element_text(size=25))

wl2_surv_to_rep_y2_size %>% 
  drop_na(SurvtoRep_y2, overhd.diam) %>% 
  group_by(pop, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(SurvtoRep_y2)), 
            mean_Surv = mean(SurvtoRep_y2,na.rm=(TRUE)), sem_surv=sem(SurvtoRep_y2, na.rm=(TRUE)), 
            N_overhd.diam = sum(!is.na(overhd.diam)), 
            mean_overhd.diam = mean(overhd.diam,na.rm=(TRUE)), sem_overhd.diam=sem(overhd.diam, na.rm=(TRUE))) %>% 
  filter(N_overhd.diam>2) %>% 
  ggplot(aes(x=mean_overhd.diam, y=mean_Surv, group=pop, color=elev_m)) +
  geom_point(size=4) +
  scale_y_continuous(expand = c(0.01, 0.01)) +
  theme_classic() + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Over-head Diameter (cm)" ,y="Survival to Budding Y2", color="Elevation (m)") +
  theme(text=element_text(size=25))

wl2_surv_to_rep_y2_size %>% 
  drop_na(SurvtoRep_y2, overhd.perp) %>% 
  group_by(pop, elev_m) %>% 
  summarise(N_Surv = sum(!is.na(SurvtoRep_y2)), 
            mean_Surv = mean(SurvtoRep_y2,na.rm=(TRUE)), sem_surv=sem(SurvtoRep_y2, na.rm=(TRUE)), 
            N_overhd.perp = sum(!is.na(overhd.perp)),
            mean_overhd.perp=mean(overhd.perp, na.rm=(TRUE)), sem_overhd.perp=sem(overhd.perp, na.rm=TRUE)) %>% 
  filter(N_overhd.perp>2) %>% 
  ggplot(aes(x=mean_overhd.perp, y=mean_Surv, group=pop, color=elev_m)) +
  geom_point(size=4) +
  scale_y_continuous(expand = c(0.01, 0.01)) +
  theme_classic() +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0")  +
  labs(x="Over-head Perp-Diameter (cm)" ,y="Survival to Budding Y2", color="Elevation (m)") +
  theme(text=element_text(size=25))
```

### Stats

Log Reg survival \~ size

#### Check for correlations between traits 

```{r}
size_normalized_wl2 <- wl2_surv_to_rep_y2_size %>% 
  select(height.cm, long.leaf.cm, diam.mm, total.branch, overhd.diam, overhd.perp) %>% 
  drop_na(height.cm, long.leaf.cm, diam.mm, total.branch, overhd.diam, overhd.perp) %>% scale() #normalize the data so they're all on the same scale
#head(size_normalized_wl2)
cor.norm_wl2 = cor(size_normalized_wl2) #test correlations among the traits
cor.sig_wl2 <- cor.mtest(size_normalized_wl2, method="pearson") #test significance of corrs
corrplot(cor.norm_wl2, type = "upper",
         tl.srt = 45, p.mat = cor.sig_wl2$p, 
         sig.level = 0.05, insig="blank")  
cor.norm_wl2
cor.sig_wl2$p
#overhead.diam and perp 85% correlated
#stem diam is 78% and 75% corr with overhead.diam and perp
#leaf length is 72%, 72%, 55% corr with diam, overhd.diam, perp
#height is 69%, 78%, 74%, 68% corr with length, diam, overhd.diam, perp
```

```{r}
wl2_surv_to_rep_y2_height_formod <- wl2_surv_to_rep_y2_size %>% 
  drop_na(SurvtoRep_y2, height.cm) %>% 
  filter(pop!="LV1", pop!="SQ1", pop!="WR")

wl2_surv_to_rep_y2_length_formod <- wl2_surv_to_rep_y2_size %>% 
  drop_na(SurvtoRep_y2, long.leaf.cm) %>% 
  filter(pop!="LV1", pop!="SQ1", pop!="WR")

wl2_surv_to_rep_y2_diam_formod <- wl2_surv_to_rep_y2_size %>% 
  drop_na(SurvtoRep_y2, diam.mm) %>% 
  filter(pop!="LV1", pop!="SQ1", pop!="WR")

wl2_surv_to_rep_y2_overhd.diam_formod <- wl2_surv_to_rep_y2_size %>% 
  drop_na(SurvtoRep_y2, overhd.diam) %>% 
  filter(pop!="LV1", pop!="SQ1", pop!="WR")
```


#### Height

```{r}
wl2_basiclogit_height <- glm(SurvtoRep_y2 ~ height.cm, data = wl2_surv_to_rep_y2_height_formod, family = "binomial")
summary(wl2_basiclogit_height)

wl2_logit_height2 <- glmer(SurvtoRep_y2 ~ height.cm + (1|pop) + (1|block), data = wl2_surv_to_rep_y2_height_formod, family = binomial(link = "logit")) #, nAGQ=0) 
summary(wl2_logit_height2)
#boundary (singular) fit: see help('isSingular') block explains 0 var
#sig relat, estimate = 0.1760
```

#### Length

```{r}
wl2_logit_length2 <- glmer(SurvtoRep_y2 ~ long.leaf.cm + (1|pop) + (1|block), data = wl2_surv_to_rep_y2_length_formod, family = binomial(link = "logit")) 
summary(wl2_logit_length2)
#boundary (singular) fit: see help('isSingular') block explains little var
#sig relat, estimate = 0.86951
```

#### Stem Diameter
```{r}
wl2_logit_diam2 <- glmer(SurvtoRep_y2 ~ diam.mm + (1|pop) + (1|block), data = wl2_surv_to_rep_y2_diam_formod, family = binomial(link = "logit")) 
summary(wl2_logit_diam2)
#boundary (singular) fit: see help('isSingular') block explains 0 var
#marg relat estiamte =  1.3730 
```


#### Overhead diam
```{r}
wl2_logit_overhd.diam2 <- glmer(SurvtoRep_y2 ~ overhd.diam + (1|pop) + (1|block), data = wl2_surv_to_rep_y2_overhd.diam_formod, family = binomial(link = "logit")) 
summary(wl2_logit_overhd.diam2)
#boundary (singular) fit: see help('isSingular') pop explains little var
#no sig relat
```

