---
title: "WL2_Growth"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---

# Analyze Growth

To Do

-   Add individual as a random effect to account for repeated measures


Notes:

-   Used sample size cut off of 3 or greater post-establishment, removed WV

-   Truncated analysis to only grow months 

-   Best base model = fit 2 (hmax varies)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)
library(brms)
```

## Need to make a csv file with all survey data for each plant

Get list of size files

```{r}
pretrans_wl2 <-  dir("../input/WL2_Data/CorrectedCSVs/",
             pattern=".*DNA.*csv",
             full.names=TRUE)
pretrans_wl2

posttrans_wl2 <- dir("../input/WL2_Data/CorrectedCSVs/",
             pattern="WL2_size_survey.*corrected.*csv",
             full.names=TRUE)
posttrans_wl2
```

Add data to a tibble

```{r}
pretrans_dat_wl2 <- tibble(path=pretrans_wl2, filename=basename(path))
pretrans_dat_wl2
posttrans_dat_wl2 <- tibble(path=posttrans_wl2, filename=basename(path))
posttrans_dat_wl2

all_wl2_size <- bind_rows(pretrans_dat_wl2, posttrans_dat_wl2) %>% 
  mutate(survey_date=str_extract(filename, "20[0-9]*"),
         survey_date=lubridate::ymd(survey_date))
all_wl2_size
```

Read in the files

```{r}
all_wl2_size2 <- all_wl2_size %>%
  mutate(sheets=map(path, 
                    read_csv, 
                    #col_types=cols(.default=col_character()), # forces all columns to be character
                                                              # this is needed because of data sheet problems that cause some numerics to be imported as character  
                    na = c("", "NA", "-", "N/A") # sets NA strings. 
                    )) %>%
  select(-path)

all_wl2_size2
```

Check for merging issues

```{r}
map(all_wl2_size2$sheets, colnames) %>% #get unique column names 
  unlist() %>%
  unique() %>%
  sort()
#height sometimes with space and sometimes with "."
#herbivory sometimes has space between y and n
#longest leaf has many different versions 
#pop is "Pop" 
#Notes are "Notes", survey-notes, and survey.notes

map(all_wl2_size2$sheets, head,10) #look at first 10 lines of each sheet 

all_wl2_size2$sheets[[3]] %>% rowwise() %>% filter(!is.na(mf)) %>%  filter(is.na(as.numeric(mf))) #all buffers
all_wl2_size2$sheets[[3]] %>% rowwise() %>% filter(!is.na(rep)) %>%  filter(is.na(as.numeric(rep))) #all buffers

all_wl2_size2$sheets[[1]] <- all_wl2_size2$sheets[[1]] %>% 
  rename(pop=Pop, height.cm=`height (cm)`, long.leaf.cm=`longest leaf (cm)`, survey.notes=Notes) %>% 
  mutate(mf=as.character(mf), rep=as.character(rep)) #convert to character since the majority of the sheets have these columns as characters 

all_wl2_size2$sheets[[2]] <- all_wl2_size2$sheets[[2]] %>% 
  rename(pop=Pop, height.cm=`height (cm)`, long.leaf.cm=`longest leaf (cm)`, survey.notes=Notes) %>% 
  filter(!is.na(height.cm))  %>% #to get rid of genotypes that were measured on the other data sheet (NAs on this sheet)
  mutate(mf=as.character(mf), rep=as.character(rep)) #convert to character since the majority of the sheets have these columns as characters 

all_wl2_size2$sheets[[3]] <- all_wl2_size2$sheets[[3]] %>% 
  rename(long.leaf.cm=`long.leaf. cm`, herbiv.y.n=`herbiv.y. n`)

all_wl2_size2$sheets[[5]] <- all_wl2_size2$sheets[[5]] %>% 
  rename(long.leaf.cm=`long.leaf.c m`)

all_wl2_size2$sheets[[7]] <- all_wl2_size2$sheets[[7]] %>% 
  rename(long.leaf.cm=`long.leaf.c m`)

all_wl2_size2$sheets[[11]] <- all_wl2_size2$sheets[[11]] %>% 
  rename(survey.notes=`survey-notes`) %>% 
  filter(block=="A"|block=="B"|block=="C"|block=="D"|block=="E"|block=="F")
unique(all_wl2_size2$sheets[[11]]$block) #only blocks A-F censused on this day 

all_wl2_size2$sheets[[12]] <- all_wl2_size2$sheets[[12]] %>% 
  filter(block=="G"|block=="H"|block=="I"|block=="J"|block=="K"|block=="L"|block=="M") #only blocks G-M censused on this day 
unique(all_wl2_size2$sheets[[12]]$block)
```

Combine the data

```{r}
all_wl2_size_all <- all_wl2_size2 %>% unnest(sheets) %>%
  rename(parent.pop=pop) %>% 
  filter(parent.pop != "buffer") %>% 
  mutate(parent.pop= str_replace(parent.pop, ".*VTR.*", "LVTR1")) %>% 
  unite(Genotype, parent.pop:rep, sep="_", remove = FALSE) %>% 
  unite(pop.mf, parent.pop:mf, sep="_", remove = FALSE) %>% 
  select(survey_date, block, Genotype:survey.notes)
head(all_wl2_size_all)
unique(all_wl2_size_all$parent.pop) #all 23 pops! + buffers though 

all_wl2_size_all %>% filter(!str_detect(mf, "buff")) %>% rowwise() %>% 
  filter(!is.na(mf)) %>%  filter(is.na(as.numeric(mf))) #all buffers --> safe to convert to numeric 
all_wl2_size_all %>% filter(!str_detect(rep, "buff")) %>% rowwise() %>% 
  filter(!is.na(rep)) %>%  filter(is.na(as.numeric(rep))) #all buffers --> safe to convert to numeric 

all_wl2_size_all_nobuffers <- all_wl2_size_all %>% 
  filter(!str_detect(mf, "buff")) %>% 
  mutate(mf=as.double(mf), rep=as.double(rep)) %>% 
  select(!DNA)
unique(all_wl2_size_all_nobuffers$parent.pop) #buffers removed!
unique(all_wl2_size_all_nobuffers$survey_date) 
tail(all_wl2_size_all_nobuffers)

unique(all_wl2_size_all_nobuffers$block)

#write_csv(all_wl2_size_all_nobuffers, "../output/WL2_Traits/WL2-2023_Size_Combined.csv")
```

## Pop Location Info

```{r}
wl2_gowers_2023 <- read_csv("../output/Climate/Gowers_WL2.csv") %>% 
  select(parent.pop:GrwSsn_GD, Wtr_Year_GD) %>% 
  pivot_wider(names_from = TimePd, values_from = c(GrwSsn_GD, Wtr_Year_GD)) 
head(wl2_gowers_2023)
```

Merge

```{r}
all_wl2_size_loc <- left_join(all_wl2_size_all_nobuffers, wl2_gowers_2023) %>% 
  mutate(day=if_else(survey_date=="2023-07-03" | survey_date=="2023-07-06", 0, #day 0 = pre-transplant
                 as.numeric(survey_date-min(survey_date)-7))) #days in the field starts on 7/11 (day all plants taken to the field) rather than the exact planting date 
head(all_wl2_size_loc)

all_wl2_size_loc %>% select(survey_date, day) %>% unique()
```

## Check Sample Sizes

```{r}
#all_wl2_size_loc %>% arrange(Genotype, day) %>% select(Genotype, day, survey_date)

all_wl2_size_loc %>% 
  filter(!is.na(height.cm)) %>% 
  group_by(day, survey_date, parent.pop) %>% 
  summarise(Sample_Size=n()) %>% 
  filter(day<38) %>% #look at first 4 weeks (1st 3 = establishment) 
  arrange(Sample_Size)

 
#WV only has 3 individuals, none survived transplant
all_wl2_size_loc %>% 
  filter(!is.na(height.cm)) %>% 
  group_by(day, survey_date, parent.pop) %>% 
  summarise(Sample_Size=n()) %>% 
  filter(day<38) %>% #look at first 4 weeks (1st 3 = establishment) 
  arrange(Sample_Size) %>% 
  filter(parent.pop=="WV")
```

## Quick figure

Height

```{r}
all_wl2_size_forfig <- all_wl2_size_loc
all_wl2_size_forfig$parent.pop = factor(all_wl2_size_forfig$parent.pop, levels=c('CC','TM2','SC','IH','BH','WV','FR','DPR','WR','WL1','SQ1','SQ2','WL2','YO4','CP2','CP3','LV3','SQ3','YO7','YO8','LV1','LVTR1','YO11'))
  
all_wl2_size_forfig %>% 
  ggplot(aes(group=Genotype, x=survey_date, y=height.cm, col=elev_m)) + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_line() + facet_wrap(~parent.pop, scales = "free")
#ggsave("../output/WL2_Traits/Growth_AllPops.png", width = 14, height = 8, units = "in")
```

Leaf Length

```{r}
all_wl2_size_loc %>% 
  ggplot(aes(group=Genotype, x=survey_date, y=long.leaf.cm, col=elev_m)) + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_line() + facet_wrap(~parent.pop, scales = "free")
```

### Check weird cases

```{r, eval=FALSE}
#Check strange lines
all_wl2_size_loc %>% filter(parent.pop=="WL2") %>% filter(height.cm>7) #data entry error, fixed on csv
all_wl2_size_loc %>% filter(parent.pop=="YO4") %>% filter(Genotype=="YO4_2_1") #data entry error, fixed on csv
all_wl2_size_loc %>% filter(parent.pop=="CP2") %>% filter(Genotype=="CP2_2_11") #data entry error, fixed on csv
all_wl2_size_loc %>% filter(parent.pop=="CP2") %>% filter(Genotype=="CP2_4_14") #data entry error, fixed on csv
all_wl2_size_loc %>% filter(parent.pop=="CP3") %>% filter(Genotype=="CP3_4_8") #data entry error, fixed on csv
all_wl2_size_loc %>% filter(parent.pop=="CP3") %>% filter(Genotype=="CP3_4_2") #data entry error, fixed on csv
all_wl2_size_loc %>% filter(parent.pop=="LV3") %>% filter(height.cm>6) #data entry error, fixed on csv
all_wl2_size_loc %>% filter(parent.pop=="LV3") %>% filter(Genotype=="LV3_7_7") #not an error, the plant had a long leaf that week 
all_wl2_size_loc %>% filter(parent.pop=="SQ3") %>% arrange(Genotype, survey_date) #data entry error, fixed on csv
all_wl2_size_loc %>% filter(parent.pop=="SQ3") %>% filter(Genotype=="SQ3_4_7") #not an error, the plant had a long leaf that week 
all_wl2_size_loc %>% filter(parent.pop=="YO8") %>% filter(height.cm>7) #data entry error, fixed on csv
all_wl2_size_loc %>% filter(parent.pop=="YO8") %>% filter(Genotype=="YO8_3_8") #no error, the plant grew then died 
all_wl2_size_loc %>% filter(parent.pop=="YO8") %>% filter(Genotype=="YO8_2_8") #no error, the plant grew then shrinked
all_wl2_size_loc %>% filter(parent.pop=="LV1") %>% filter(height.cm>10) #data entry error, fixed on csv
all_wl2_size_loc %>% filter(parent.pop=="LV1") %>% filter(Genotype=="LV1_6_1") #no error, the plant grew then shrinked
all_wl2_size_loc %>% filter(parent.pop=="LV1") %>% filter(Genotype=="LV1_2_1") #no error, herbivory caused no leaves one week in Sept

all_wl2_size_loc %>% filter(parent.pop=="FR") %>% filter(Genotype=="FR_7_3") #no error, the plant grew and then started dying
all_wl2_size_loc %>% filter(parent.pop=="BH") %>% filter(long.leaf.cm>7.5) #no error, the plant grew and then shrinked 
```

## Data set for the model (remove WV and shrinkage months)
```{r}
#all_wl2_size_loc %>% filter(Genotype=="YO8_2_8") %>% #shrinkage test 
# arrange(survey_date) %>% 
# select(survey_date:height.cm, day) %>% 
# mutate(height_next=lead(height.cm)) %>% 
# mutate(height_diff=height_next-height.cm) %>% 
#  mutate(shrink_day=if_else(height_diff>-2, NA, day)) %>% 
#  mutate(min_shrink_day=min(shrink_day, na.rm = TRUE)) %>% 
#  filter(day<min_shrink_day)

all_wl2_size_for_models <- all_wl2_size_loc %>% 
  filter(parent.pop!="WV") %>% 
  group_by(Genotype) %>% 
  mutate(height_next = lead(height.cm, order_by = survey_date),
         height_diff=height_next-height.cm,
         shrink_day=if_else(height_diff>-2, NA, day),
         min_shrink_day=min(shrink_day, na.rm = TRUE)) %>% #deals with problem of if something shrinks and then starts to grow again
  filter(day<min_shrink_day | min_shrink_day==Inf) %>% 
  ungroup() %>% 
  drop_na(height.cm)

all_wl2_size_for_models %>% 
  ggplot(aes(group=Genotype, x=survey_date, y=height.cm, col=elev_m)) + 
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_line() + facet_wrap(~parent.pop, scales = "free")

all_wl2_size_for_models %>% 
  filter(!is.na(height.cm)) %>% 
  group_by(day, survey_date, parent.pop) %>% 
  summarise(Sample_Size=n()) %>% 
  filter(day<38) %>% #look at first 4 weeks (1st 3 = establishment) 
  arrange(Sample_Size)
#no new pops to remove 

write_csv(all_wl2_size_for_models, "../output/WL2_Traits/WL2_growth_for_models.csv")
```


## Model growth with the Weibull model

```{r}
weibull <- function (t, alpha, beta, k, delta) {
  result <- alpha - (alpha - beta) * exp(-(k * t)^delta)
  return(result)
}

growth <- tibble(time = seq(0,100,.1)) %>%
  mutate(model1 = weibull(t = time,
                          alpha = 40, #Hmax = alpha -  Sets the asymptote (max growth)
                          beta = 0.5, #Hmin = beta - Changes the y intercept (the lower limit)
                          k = 0.01, #k - Often described as the growth rate --\> related to the slope of the line
                          delta = 0.5)) #delta - Sets the inflection point - when you shift b/t concave to convex

growth %>%  ggplot(aes(x=time, y=model1)) +
  geom_line()
```

### Base Model

```{r}
f1_WL2 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta), #Weibull formula, divide k by 100 to get it on a more reasonable scale
    Hmax + Hmin + k + delta ~ 1, # how each parameter should be fit with respect to the predictors 
    #we want a single parameter value (Intercept) for each of the four parameters
   nl=TRUE) #tells brms that this is a non-linear formula
```

### Set Priors

```{r}
prior1_WL2 <- c(set_prior("normal(4,3)", nlpar="Hmin"), #(mean, sd)
            set_prior("normal(25,10)",  nlpar="Hmax"),
            set_prior("gamma(1,2)", nlpar = "k", lb = 0), #gamma constrains to be > 0 
            set_prior("gamma(4,2)", nlpar = "delta", lb = 0)
)

#Hmax = alpha -  Sets the asymptote (max growth)
#Hmin = beta - Changes the y intercept (the lower limit)
#k - Often described as the growth rate --\> related to the slope of the line
#delta - Sets the inflection point - when you shift b/t concave to convex
```

Code for plotting different normal distributions

```{r}
#define population mean and standard deviation
population_mean <- 25
population_sd <- 10

#define upper and lower bound
lower_bound <- population_mean - population_sd
upper_bound <- population_mean + population_sd

#Create a sequence of 1000 x values based on population mean and standard deviation
x <- seq(-4, 4, length = 1000) * population_sd + population_mean

#create a vector of values that shows the height of the probability distribution
#for each value in x
y <- dnorm(x, population_mean, population_sd)

#plot normal distribution with customized x-axis labels
plot(x,y, type = "l", lwd = 2, axes = FALSE, xlab = "", ylab = "")
sd_axis_bounds = 5
axis_bounds <- seq(-sd_axis_bounds * population_sd + population_mean,
                    sd_axis_bounds * population_sd + population_mean,
                    by = population_sd)
axis(side = 1, at = axis_bounds, pos = 0)
```

Gamma dist:

```{r}
#define population mean and standard deviation
population_mean <- 4
population_sd <- 2

#define upper and lower bound
lower_bound <- population_mean - population_sd
upper_bound <- population_mean + population_sd

#Create a sequence of 1000 x values based on population mean and standard deviation
x <- seq(-4, 4, length = 1000) * population_sd + population_mean

#create a vector of values that shows the height of the probability distribution
#for each value in x
y <- dgamma(x, population_mean, population_sd)

#plot normal distribution with customized x-axis labels
plot(x,y, type = "l", lwd = 2, axes = FALSE, xlab = "", ylab = "")
sd_axis_bounds = 5
axis_bounds <- seq(-sd_axis_bounds * population_sd + population_mean,
                    sd_axis_bounds * population_sd + population_mean,
                    by = population_sd)
axis(side = 1, at = axis_bounds, pos = 0)
```

### Fit the model

```{r}
fit1_WL2 <- brm(formula=f1_WL2, data=all_wl2_size_for_models, prior=prior1_WL2, cores = 4, iter=3000)

#Warning: There were 17 divergent transitions after warmup. See https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup to find out why this is a problem and how to eliminate them.Warning: Examine the pairs() plot to diagnose sampling problems
```

Summary

```{r}
summary(fit1_WL2)
#Rhat and ESS are good

pairs(fit1_WL2, regex = TRUE) 
#some don't look great, these should be more cloud like than show patterns
#k and Hmax negatively correlated
#delta and k positively correlated 
```

Plots

```{r}
plot(fit1_WL2, nvariables = 3, ask=FALSE)
#some chains don't look good (hmax, k)
```

### Predictions

```{r}
fit1_WL2.prediction <- cbind(all_wl2_size_for_models, pred1=predict(fit1_WL2)[,"Estimate"])

fit1_WL2.prediction %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm), alpha=.1) +
  geom_line(aes(y=pred1), color="blue") +
  facet_wrap(. ~parent.pop)

#Predictions are clearly off, esp for TM2, good for BH 
```

## Let Hmax vary

```{r}
random_f2_WL2 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin + k + delta ~ 1,
                  Hmax ~ (1|parent.pop), #different hmax for each pop + random effect of pop
                  nl=TRUE)
```

Default Priors

```{r}
get_prior(random_f2_WL2, data = all_wl2_size_for_models)
#there's now a prior for sd of Hmax due to pop
#don't need a specific prior for each pop now 

prior2_WL2 <- c(set_prior("normal(4,3)", nlpar="Hmin"),
            set_prior("normal(25,10)",  nlpar="Hmax"),
            set_prior("gamma(1,2)", nlpar = "k", lb = 0), #gamma constrains to be > 0 
            set_prior("gamma(4,2)", nlpar = "delta", lb = 0)
) #same as prior 1 for now 
```

### Fit the model

```{r}
random_fit2_WL2 <- brm(formula=random_f2_WL2, data=all_wl2_size_for_models, prior=prior2_WL2, cores = 4, 
                       iter = 3000, #increase iterations b/c more complex (helps with Rhat and ESS)
                       control = list(max_treedepth = 12)) #default = 10 
```

Summary

```{r}
summary(random_fit2_WL2)
#rhat and ESS mostly good 
#hmax intercept ESS < 1000
```

```{r}
plot(random_fit2_WL2, nvariables = 3, ask=FALSE)
#Hmax chains look a little weird 
```

### Predictions

The function expand_grid creates a data frame with all combinations of the items it is given (days and pops in this case).

```{r}
pred.df <- expand_grid(day=min(all_wl2_size_for_models$day):max(all_wl2_size_for_models$day),
                       parent.pop=unique(all_wl2_size_for_models$parent.pop))
#pred.df
```

```{r}
random_fit2_WL2.predictions <- cbind(pred.df, prediction=predict(random_fit2_WL2, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_wl2_size_for_models, by=c("day", "parent.pop"))

random_fit2_WL2.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  #scale_color_brewer(type="qual", palette = "Set2", guide="none") +
  facet_wrap(~parent.pop)

random_fit2_WL2.predictions %>% filter(parent.pop=="BH") %>% 
  ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)

#predictions look pretty good 
```

## Let Hmax vary - Add Indiv for Repeated Measures 

```{r}
random_f2b_WL2 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin + k + delta ~ 1,
                  Hmax ~ (1|parent.pop) + (1|Genotype), #different hmax for each pop + random effect of pop + random effect of genotype for repeated measures 
                  nl=TRUE)
```

Default Priors

```{r}
get_prior(random_f2b_WL2, data = all_wl2_size_for_models)
#there's now a prior for sd of Hmax due to pop
#don't need a specific prior for each pop now 

prior2_WL2 <- c(set_prior("normal(4,3)", nlpar="Hmin"),
            set_prior("normal(25,10)",  nlpar="Hmax"),
            set_prior("gamma(1,2)", nlpar = "k", lb = 0), #gamma constrains to be > 0 
            set_prior("gamma(4,2)", nlpar = "delta", lb = 0)
) #same as prior 1 for now 
```

### Fit the model

```{r}
random_fit2b_WL2 <- brm(formula=random_f2b_WL2, data=all_wl2_size_for_models, prior=prior2_WL2, cores = 4, 
                       iter = 3800, #increase iterations b/c more complex (helps with Rhat and ESS)
                       control = list(max_treedepth = 12)) #default = 10 
#Warning: Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable. Running the chains for more iterations may help. See https://mc-stan.org/misc/warnings.html#bulk-ess
```

Summary

```{r}
summary(random_fit2b_WL2)
#rhat good 
#ESS too low
```

```{r}
plot(random_fit2b_WL2, nvariables = 3, ask=FALSE)
#Hmax chains look a little weird 
```

### Predictions

The function expand_grid creates a data frame with all combinations of the items it is given (days and pops in this case).

```{r}
pred.df <- expand_grid(day=min(all_wl2_size_for_models$day):max(all_wl2_size_for_models$day),
                       parent.pop=unique(all_wl2_size_for_models$parent.pop),
                       Genotype=unique(all_wl2_size_for_models$Genotype),)
#pred.df
```

```{r}
random_fit2b_WL2.predictions <- cbind(pred.df, prediction=predict(random_fit2b_WL2, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_wl2_size_for_models, by=c("day", "parent.pop", "Genotype"))
#Error: vector memory limit of 16.0 Gb reached, see mem.maxVSize()

random_fit2b_WL2.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  #scale_color_brewer(type="qual", palette = "Set2", guide="none") +
  facet_wrap(~parent.pop)

random_fit2b_WL2.predictions %>% filter(parent.pop=="BH") %>% 
  ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)

#predictions look pretty good 
```

## Vary slope (k) and asymptote (Hmax) by pop

```{r}
random_f3_WL2 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin + delta ~ 1,
                  Hmax ~ (1|parent.pop), #different hmax for each pop 
                  #could also do Hmax + k  ~pop
                  k ~ (1|parent.pop), #different k for each pop
                  nl=TRUE)
```

### Priors

```{r}
get_prior(random_f3_WL2, data=all_wl2_size_for_models)

prior3_WL2 <- c(set_prior("normal(4,3)", nlpar="Hmin"),
            set_prior("normal(25,10)",  nlpar="Hmax"),
            set_prior("normal(1,2)", nlpar = "k", lb = 0), #gamma constrains to be > 0 
            set_prior("gamma(4,2)", nlpar = "delta", lb = 0)
)
```

### Fit the model

```{r}
random_fit3_WL2 <- brm(formula=random_f3_WL2, data=all_wl2_size_for_models, prior=prior3_WL2, cores = 4, iter = 3000)
#Warning: There were 5024 divergent transitions after warmup. See https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup to find out why this is a problem and how to eliminate them.
#Warning: There were 3 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10. See https://mc-stan.org/misc/warnings.html#maximum-treedepth-exceeded
#Warning: There were 1 chains where the estimated Bayesian Fraction of Missing Information was low. See https://mc-stan.org/misc/warnings.html#bfmi-lowWarning: Examine the pairs() plot to diagnose sampling problems
#Warning: The largest R-hat is 2.52, indicating chains have not mixed. Running the chains for more iterations may help. See https://mc-stan.org/misc/warnings.html#r-hat
#Warning: Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable. Running the chains for more iterations may help. Seehttps://mc-stan.org/misc/warnings.html#bulk-ess
#Warning: Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable. Running the chains for more iterations may help. See https://mc-stan.org/misc/warnings.html#tail-ess
```

Summary

```{r}
summary(random_fit3_WL2)
#pretty terrible
```

Plots

```{r}
plot(random_fit3_WL2, N = 3, ask=FALSE)
#pretty terrible
```

### Pairs

```{r}
pairs(random_fit3_WL2, variable = "TM2", regex = TRUE) 
#look at correlation b/t hmax and k for each pop 
#looks bad
```

Predictions

```{r}
random_fit3_WL2.predictions <- cbind(pred.df, prediction=predict(random_fit3_WL2, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_wl2_size_for_models, by=c("day", "parent.pop"))

random_fit3_WL2.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)
#predictions look fine, but pretty similar to just Hmax varying 
```

## Let delta and hmax vary

```{r}
random_f4_WL2 <- brmsformula(height.cm ~ Hmax - (Hmax - Hmin) * exp(-(k/100 * day)^delta),
                  Hmin + k ~ 1,
                  Hmax ~ (1|parent.pop), #different hmax for each pop #could also do Hmax + k  ~pop
                  delta ~ (1|parent.pop), #different hmax for each pop
                  nl=TRUE)
```

```{r}
prior4_WL2 <- c(set_prior("normal(4,3)", nlpar="Hmin"),
            set_prior("normal(25,10)",  nlpar="Hmax"),
            set_prior("gamma(1,2)", nlpar = "k", lb = 0), #gamma constrains to be > 0 
            set_prior("gamma(4,2)", nlpar = "delta", lb = 0)
)
```

```{r}
random_fit4_WL2 <- brm(formula=random_f4_WL2, data=all_wl2_size_for_models, prior=prior4_WL2, cores = 4, iter = 3000)

#Warning: There were 6000 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10. See https://mc-stan.org/misc/warnings.html#maximum-treedepth-exceeded
#Warning: There were 1 chains where the estimated Bayesian Fraction of Missing Information was low. See https://mc-stan.org/misc/warnings.html#bfmi-lowWarning: Examine the pairs() plot to diagnose sampling problems
#Warning: The largest R-hat is 3.16, indicating chains have not mixed. Running the chains for more iterations may help. See https://mc-stan.org/misc/warnings.html#r-hat
#Warning: Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable. Running the chains for more iterations may help. See https://mc-stan.org/misc/warnings.html#bulk-ess
#Warning: Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable. Running the chains for more iterations may help. See https://mc-stan.org/misc/warnings.html#tail-ess
```

```{r}
summary(random_fit4_WL2)
#pretty terrible
```

```{r}
plot(random_fit4_WL2, N = 3, ask=FALSE)
#pretty terrible
```

Predictions

```{r}
random_fit4_WL2.predictions <- cbind(pred.df, prediction=predict(random_fit4_WL2, newdata = pred.df)[,"Estimate"]) %>%
  full_join(all_wl2_size_for_models, by=c("day", "parent.pop"))

random_fit4_WL2.predictions %>% ggplot(aes(x=day)) +
  geom_line(aes(group=Genotype, y=height.cm, color=parent.pop), alpha=.2) +
  geom_line(aes(y=prediction, color=parent.pop)) +
  facet_wrap(~parent.pop)
#predictions look fine but not much different from Hmax model 
```

## Model Comparison

```{r}
fit1_WL2 <- add_criterion(fit1_WL2, "loo")
random_fit2_WL2 <- add_criterion(random_fit2_WL2, "loo")
random_fit3_WL2 <- add_criterion(random_fit3_WL2, "loo") #Warning: Found 43 observations with a pareto_k > 0.7 in model 'random_fit3_WL2'. We recommend to set 'moment_match = TRUE' in order to perform moment matching for problematic observations.
random_fit4_WL2 <- add_criterion(random_fit4_WL2, "loo") #Warning: Found 275 observations with a pareto_k > 0.7 in model 'random_fit4_WL2'. We recommend to set 'moment_match = TRUE' in order to perform moment matching for problematic observations. 
loo_compare(fit1_WL2, random_fit2_WL2, random_fit3_WL2, random_fit4_WL2)
#fit 2 the best as expected 
```

