---
title: "WL2_climate_distance"
author: "Brandie Quarles"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---
To Do: 
- Copy this code and substitute the full year climate data for the growth season inputs used here.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Climate Distance at the WL2 Garden

Note, there is no air temperature data for WL2 in July 2023!! 
Add data for November and December once we get the data after snowmelt? 
- Reference growth season criteria 

## Relevant Libraries and Functions

```{r}
library(raster)
library(tidyverse)
library(conflicted)
conflicts_prefer(dplyr::select())
conflicts_prefer(dplyr::filter)
library(ggrepel)
library(cowplot)
library(gridExtra)
library(naniar) #replaces values with NA
sem <- function(x, na.rm=FALSE) {
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} #standard error function 

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
} #legend function for grid_arrange

elev_three_palette <- c("#0043F0", "#C9727F", "#F5A540") #colors from Gremer et al 2019
elev_order <- c("High", "Mid", "Low")
```

## Home Climates

### Flint

```{r}
pops_flint_recent_avgs <- read_csv("../output/Climate/Pops_FlintAvgs_Recent.csv") %>% 
  select(parent.pop:tmx_mean)
head(pops_flint_recent_avgs)

pops_flint_historic_avgs <- read_csv("../output/Climate/Pops_FlintAvgs_Historic.csv") %>% 
  select(parent.pop:tmx_mean)
head(pops_flint_historic_avgs)
names(pops_flint_historic_avgs)
```

### BioClim

```{r}
pops_bioclim_recent_avgs <- read_csv("../output/Climate/Pops_BioClimAvgs_Recent.csv") %>% 
  select(parent.pop:ppt_coldest_month_avg)
head(pops_bioclim_recent_avgs)

pops_bioclim_historical_avgs <- read_csv("../output/Climate/Pops_BioClimAvgs_Historical.csv")  %>% 
  select(parent.pop:ppt_coldest_month_avg)
head(pops_bioclim_historical_avgs)

names(pops_bioclim_historical_avgs)
```

## WL2 Climate Data (July 2023-Oct 2023)

### From Magney MetStation (half hour data)

```{r}
WL2_climate_early <- read_csv("../input/WL2_2023_Early_CR1000XSeries_HalfHourTable.csv")
head(WL2_climate_early) #note, cannot use any of the pyranometer data from this dataset 

WL2_climate_late <- read_csv("../input/WL2_2023_CR1000XSeries_HalfHourTable.csv")
head(WL2_climate_late, 12)

WL2_climate <- bind_rows(WL2_climate_early, WL2_climate_late) %>% 
  separate(TIMESTAMP, c("date", "time"), sep = " ")
head(WL2_climate)
tail(WL2_climate)
```

Convert to monthly
```{r}
WL2_climate_monthly_prep <- WL2_climate %>% 
  mutate(date=as.Date(date, format="%m/%d/%y"), month=month(date), year=year(date)) %>% 
  unite("month_year", month:year, remove = FALSE) %>% 
  filter(month_year!="6_2023") %>%  #remove june because planting started July 11
  mutate(across(where(is.numeric), \(x) ifelse(is.nan(x), NA, x))) %>% 
  dplyr::select(date, month, year, month_year,time, PTemp_C_Max:BulkEC_50cm_Std)
head(WL2_climate_monthly_prep)

WL2_climate_monthly_prep %>% filter(month_year=="7_2023")

WL2_climate_monthly_temp_ppt <- WL2_climate_monthly_prep %>% 
  group_by(month_year) %>% 
  filter(month_year!="7_2023") %>% #remove July for now since no temp data, could add it back in for ppt 
  summarise(tmn = min(AirTC_Min, na.rm = TRUE), 
            tmx=max(AirTC_Max, na.rm = TRUE), 
            tavg=mean(AirTC_Avg, na.rm = TRUE), 
            ppt=sum(Rain_mm_Tot, na.rm = TRUE)) %>% 
  mutate(t_diurnal = (tmx-tmn))
WL2_climate_monthly_temp_ppt
```


```{r}
WL2_climate_short_flint <- WL2_climate_monthly_temp_ppt %>% #get the means for the variables to compare to home sites 
  summarise(ppt_mean_WL2=mean(ppt), tmn_mean_WL2=mean(tmn), tmx_mean_WL2=mean(tmx))
WL2_climate_short_flint
```



### BioClim

-   annual mean temperature (BIO1)

-   mean diurnal range (BIO2) - (Mean of monthly (max temp - min temp))

-   temperature seasonality (BIO4) (standard deviation \*100)

-   temperature annual range (BIO7) (Max Temperature of Warmest Month -
    Min Temperature of Coldest Month)

-   annual precipitation (BIO12) - sum of ppt for the entire year (not
    the avg)

-   precipitation seasonality (BIO15) (Coefficient of Variation)

-   mean temp of wettest quarter (BIO8) - CHANGED TO mean temp of
    wettest month

-   mean temp of driest quarter (BIO9) - CHANGED TO mean temp of driest
    month

-   precip of warmest quarter (BIO18) - CHANGED TO precip of warmest
    month

-   precip of coldest quarter (BIO19) - CHANGED TO precip of coldest
    month

Calculating wettest, driest, warmest, and coldest months

```{r}
WL2_wettest_month <- WL2_climate_monthly_temp_ppt %>%  
  slice_max(ppt)

WL2_driest_month <- WL2_climate_monthly_temp_ppt %>% 
  slice_min(ppt)

WL2_warmest_month <- WL2_climate_monthly_temp_ppt %>% 
  slice_max(tavg)

WL2_coldest_month <- WL2_climate_monthly_temp_ppt %>%
  slice_min(tavg)
```

Bio 1, 2, 4, 7, 12, 15

```{r}
bioclim_WL2_calc <- WL2_climate_monthly_temp_ppt %>% 
  summarise(ann_tmean=mean(tavg),  #Bio1 - Annual Mean Temperature
            mean_diurnal_range=mean(t_diurnal), #Bio2 - Mean Diurnal Range
            temp_seasonality=sd(tavg), #Bio4 - Temperature Seasonality
            temp_ann_range=(max(tmx))-(min(tmn)), #bio7 - temp annual range
            ann_ppt=sum(ppt), #bio12 - annual precip
            ppt_seasonality=cv(ppt+1)) #bio15 - Precipitation Seasonality (+1 to avoid strange CVs for areas where mean rainfaill is < 1)
bioclim_WL2_calc
```

Bio 8(Q), 9(Q), 18(Q), 19(Q)

```{r}
#bio8 = tmean_wettest_month
bio8_WL2 <- WL2_wettest_month %>% 
  dplyr::select(tmean_wettest_month=tavg)

#bio9 = tmean_driest_month
bio9_WL2 <- WL2_driest_month %>% 
  dplyr::select(tmean_driest_month=tavg)

bio8_9_WL2 <- bind_cols(bio8_WL2, bio9_WL2)

#bio18 = ppt_warmest_month
bio18_WL2 <- WL2_warmest_month %>% 
  dplyr::select(ppt_warmest_month=ppt)

#bio19 = ppt_coldest_month
bio19_WL2 <- WL2_wettest_month %>% 
  dplyr::select(ppt_coldest_month=ppt)

bio18_19_WL2 <- bind_cols(bio18_WL2, bio19_WL2)

all_periods_WL2 <- bind_cols(bio8_9_WL2, bio18_19_WL2)
```

Merge all bioclims

```{r}
WL2_bioclim_final <- bind_cols(bioclim_WL2_calc, all_periods_WL2) %>% 
  select_all(.funs = funs(paste0(., "_WL2")))
summary(WL2_bioclim_final)
```

## WL2 Climate Trends

```{r}
WL2_climate_monthly_temp_ppt$month_year <- factor(WL2_climate_monthly_temp_ppt$month_year, levels = c("8_2023", "9_2023", "10_2023"))
WL2_climate_monthly_temp_ppt %>% ggplot(aes(x=month_year, y=tavg)) + geom_point() + theme_classic()
WL2_climate_monthly_temp_ppt %>% ggplot(aes(x=month_year, y=ppt)) + geom_point() + theme_classic()
WL2_climate_monthly_temp_ppt %>% ggplot(aes(x=month_year, y=tmx)) + geom_point() + theme_classic()
```

## Gower's Climate Distance

(1/P) \* SUM ((absolute value(Ai - Bi)) / range(i)) for each variable

-   P = number of environmental variables = 13 (without CWD)

-   Ai = 30 year avg of that variable for the home site

-   Bi = Nov 2022-Oct 2023 avg of that variable for the WL2 garden

-   Range(i) = maximum - minimum of that variable in the whole data set
    (across sites)

### Combine the flint and bioclim variables 

```{r}
WL2_climate_all <- bind_cols(WL2_climate_short_flint, WL2_bioclim_final)
dim(WL2_climate_all)

home_climates_all_recent <- left_join(pops_flint_recent_avgs, pops_bioclim_recent_avgs)
home_climates_all_historic <- left_join(pops_flint_historic_avgs, pops_bioclim_historical_avgs)

WL2_home_climate_recent <- bind_cols(WL2_climate_all, home_climates_all_recent)
summary(WL2_home_climate_recent)

WL2_home_climate_historic <- bind_cols(WL2_climate_all, home_climates_all_historic)
summary(WL2_home_climate_historic)

variables <- c("ppt_mean", "tmn_mean", "tmx_mean", "ann_tmean_avg",
                 "mean_diurnal_range_avg","temp_seasonality_avg","temp_ann_range_avg",
                 "tmean_wettest_month_avg", "tmean_driest_month_avg","ann_ppt_avg", 
                 "ppt_seasonality_avg", "ppt_warmest_month_avg", "ppt_coldest_month_avg")
```

### Figure out the range for each variable 

```{r}
WL2_range_prep <- WL2_climate_all %>% 
  mutate(parent.pop="WL2")

colnames(WL2_range_prep) <- sub("_WL2", "_avg", colnames(WL2_range_prep)) #remove WL2 from column names
WL2_range_prep <- WL2_range_prep %>% rename(ppt_mean=ppt_mean_avg, tmn_mean=tmn_mean_avg, tmx_mean=tmx_mean_avg)

##Recent
range_merge_recent <- bind_rows(home_climates_all_recent, WL2_range_prep)
names(range_merge_recent)
summary(range_merge_recent)

WL2_home_climate_ranges_recent <- range_merge_recent %>% 
  summarise(ppt_mean_range=max(ppt_mean)-min(ppt_mean), 
            tmn_mean_range=max(tmn_mean)-min(tmn_mean), 
            tmx_mean_range=max(tmx_mean)-min(tmx_mean), 
            ann_tmean_avg_range=max(ann_tmean_avg)-min(ann_tmean_avg),
            mean_diurnal_range_avg_range=max(mean_diurnal_range_avg)-min(mean_diurnal_range_avg),
            temp_seasonality_avg_range=max(temp_seasonality_avg)-min(temp_seasonality_avg),
            temp_ann_range_avg_range=max(temp_ann_range_avg)-min(temp_ann_range_avg),
            tmean_wettest_month_avg_range=max(tmean_wettest_month_avg)-min(tmean_wettest_month_avg),
            tmean_driest_month_avg_range=max(tmean_driest_month_avg)-min(tmean_driest_month_avg),
            ann_ppt_avg_range=max(ann_ppt_avg)-min(ann_ppt_avg), 
            ppt_seasonality_avg_range=max(ppt_seasonality_avg)-min(ppt_seasonality_avg),
            ppt_warmest_month_avg_range=max(ppt_warmest_month_avg)-min(ppt_warmest_month_avg), 
            ppt_coldest_month_avg_range=max(ppt_coldest_month_avg)-min(ppt_coldest_month_avg))
WL2_home_climate_ranges_recent

WL2_home_climate_with_ranges_recent <- bind_cols(WL2_home_climate_recent, WL2_home_climate_ranges_recent)
names(WL2_home_climate_with_ranges_recent)

##Historic
range_merge_historic <- bind_rows(home_climates_all_historic, WL2_range_prep)
names(range_merge_historic)
summary(range_merge_historic)

WL2_home_climate_ranges_historic <- range_merge_historic %>% 
  summarise(ppt_mean_range=max(ppt_mean)-min(ppt_mean), 
            tmn_mean_range=max(tmn_mean)-min(tmn_mean), 
            tmx_mean_range=max(tmx_mean)-min(tmx_mean), 
            ann_tmean_avg_range=max(ann_tmean_avg)-min(ann_tmean_avg),
            mean_diurnal_range_avg_range=max(mean_diurnal_range_avg)-min(mean_diurnal_range_avg),
            temp_seasonality_avg_range=max(temp_seasonality_avg)-min(temp_seasonality_avg),
            temp_ann_range_avg_range=max(temp_ann_range_avg)-min(temp_ann_range_avg),
            tmean_wettest_month_avg_range=max(tmean_wettest_month_avg)-min(tmean_wettest_month_avg),
            tmean_driest_month_avg_range=max(tmean_driest_month_avg)-min(tmean_driest_month_avg),
            ann_ppt_avg_range=max(ann_ppt_avg)-min(ann_ppt_avg), 
            ppt_seasonality_avg_range=max(ppt_seasonality_avg)-min(ppt_seasonality_avg),
            ppt_warmest_month_avg_range=max(ppt_warmest_month_avg)-min(ppt_warmest_month_avg), 
            ppt_coldest_month_avg_range=max(ppt_coldest_month_avg)-min(ppt_coldest_month_avg))
WL2_home_climate_ranges_historic

WL2_home_climate_with_ranges_historic <- bind_cols(WL2_home_climate_historic, WL2_home_climate_ranges_historic)
names(WL2_home_climate_with_ranges_historic)
```

### Recent Gowers Calc
```{r}
gowers_calc_each_var_recent <- WL2_home_climate_with_ranges_recent %>% 
  mutate(ppt_mean_gowers=abs(ppt_mean_WL2 - ppt_mean) / ppt_mean_range,
         tmn_mean_gowers=abs(tmn_mean_WL2 - tmn_mean) / tmn_mean_range,
         tmx_mean_gowers=abs(tmx_mean_WL2 - tmx_mean) / tmx_mean_range,
         ann_tmean_gowers=abs(ann_tmean_WL2 - ann_tmean_avg) / ann_tmean_avg_range,
         mean_diurnal_range_gowers=abs(mean_diurnal_range_WL2 - mean_diurnal_range_avg) / mean_diurnal_range_avg_range,
         temp_seasonality_gowers=abs(temp_seasonality_WL2 - temp_seasonality_avg) / temp_seasonality_avg_range,
         temp_ann_range_gowers=abs(temp_ann_range_WL2 - temp_ann_range_avg) / temp_ann_range_avg_range,
         tmean_wettest_month_gowers=abs(tmean_wettest_month_WL2 - tmean_wettest_month_avg) / tmean_wettest_month_avg_range,
         tmean_driest_month_gowers=abs(tmean_driest_month_WL2 - tmean_driest_month_avg) / tmean_driest_month_avg_range,
         ann_ppt_gowers=abs(ann_ppt_WL2 - ann_ppt_avg) / ann_ppt_avg_range,
         ppt_seasonality_gowers=abs(ppt_seasonality_WL2 - ppt_seasonality_avg) / ppt_seasonality_avg_range,
         ppt_warmest_month_gowers=abs(ppt_warmest_month_WL2 - ppt_warmest_month_avg) / ppt_warmest_month_avg_range,
         ppt_coldest_month_gowers=abs(ppt_coldest_month_WL2 - ppt_coldest_month_avg) / ppt_coldest_month_avg_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))

WL2_home_climate_with_ranges_recent
gowers_calc_each_var_recent
  
gowers_calc_per_pop_recent <- gowers_calc_each_var_recent %>% 
  mutate(Recent_Gowers_Dist=(1/13)*(ppt_mean_gowers + tmn_mean_gowers + tmx_mean_gowers +
                                ann_tmean_gowers + mean_diurnal_range_gowers +
                                temp_seasonality_gowers +temp_ann_range_gowers +
                                tmean_wettest_month_gowers +
                                tmean_driest_month_gowers +ann_ppt_gowers +
                                ppt_seasonality_gowers + ppt_warmest_month_gowers +
                                ppt_coldest_month_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Recent_Gowers_Dist)

gowers_calc_per_pop_recent
```

### Historic Gowers Calc
```{r}
gowers_calc_each_var_historic <- WL2_home_climate_with_ranges_historic %>% 
  mutate(ppt_mean_gowers=abs(ppt_mean_WL2 - ppt_mean) / ppt_mean_range,
         tmn_mean_gowers=abs(tmn_mean_WL2 - tmn_mean) / tmn_mean_range,
         tmx_mean_gowers=abs(tmx_mean_WL2 - tmx_mean) / tmx_mean_range,
         ann_tmean_gowers=abs(ann_tmean_WL2 - ann_tmean_avg) / ann_tmean_avg_range,
         mean_diurnal_range_gowers=abs(mean_diurnal_range_WL2 - mean_diurnal_range_avg) / mean_diurnal_range_avg_range,
         temp_seasonality_gowers=abs(temp_seasonality_WL2 - temp_seasonality_avg) / temp_seasonality_avg_range,
         temp_ann_range_gowers=abs(temp_ann_range_WL2 - temp_ann_range_avg) / temp_ann_range_avg_range,
         tmean_wettest_month_gowers=abs(tmean_wettest_month_WL2 - tmean_wettest_month_avg) / tmean_wettest_month_avg_range,
         tmean_driest_month_gowers=abs(tmean_driest_month_WL2 - tmean_driest_month_avg) / tmean_driest_month_avg_range,
         ann_ppt_gowers=abs(ann_ppt_WL2 - ann_ppt_avg) / ann_ppt_avg_range,
         ppt_seasonality_gowers=abs(ppt_seasonality_WL2 - ppt_seasonality_avg) / ppt_seasonality_avg_range,
         ppt_warmest_month_gowers=abs(ppt_warmest_month_WL2 - ppt_warmest_month_avg) / ppt_warmest_month_avg_range,
         ppt_coldest_month_gowers=abs(ppt_coldest_month_WL2 - ppt_coldest_month_avg) / ppt_coldest_month_avg_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))

  
gowers_calc_per_pop_historic <- gowers_calc_each_var_historic %>% 
  mutate(Historic_Gowers_Dist=(1/13)*(ppt_mean_gowers + tmn_mean_gowers + tmx_mean_gowers +
                                ann_tmean_gowers + mean_diurnal_range_gowers +
                                temp_seasonality_gowers +temp_ann_range_gowers +
                                tmean_wettest_month_gowers +
                                tmean_driest_month_gowers +ann_ppt_gowers +
                                ppt_seasonality_gowers + ppt_warmest_month_gowers +
                                ppt_coldest_month_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Historic_Gowers_Dist)

gowers_calc_per_pop_historic
```

### Merge recent and historic
```{r}
gowers_all_time <- full_join(gowers_calc_per_pop_recent, gowers_calc_per_pop_historic)
gowers_all_time
names(gowers_all_time)
write_csv(gowers_all_time, "../output/Climate/Pops_GowersEnvtalDist_WL2.csv")
```

Figures
```{r}
gowers_rec_fig <- gowers_all_time %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Recent_Gowers_Dist), y=Recent_Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
   labs(fill="Elevation (m)",x="Population", title="Recent Climate", y="Gowers Envtal Distance \n from WL2") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Gowers_Recent_fromWL2.png", width = 12, height = 6, units = "in")

gowers_hist_fig <- gowers_all_time %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Historic_Gowers_Dist), y=Historic_Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population", title="Historic Climate", y="Gowers Envtal Distance \n from WL2") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Gowers_Historic_fromWL2.png", width = 12, height = 6, units = "in")

#should combine these into one figure and save that instead
legend <- get_legend(gowers_rec_fig)
gowers_hist_fig <- gowers_hist_fig + theme(legend.position="none")
gowers_rec_fig <- gowers_rec_fig + theme(legend.position="none")
grid.arrange(gowers_hist_fig, gowers_rec_fig, legend, ncol=3, widths=c(3.12, 3.12, 1.09))
 #2000 x 850
```


## Flint Climate Distance

### Subtraction

```{r}
recent_flint_dist_prep <- bind_cols(WL2_climate_short_flint, pops_flint_recent_avgs)
names(recent_flint_dist_prep)

recent_flint_dist <- recent_flint_dist_prep %>% 
  mutate(ppt_mean_dist=ppt_mean_WL2 - ppt_mean,
         tmn_mean_dist=tmn_mean_WL2 - tmn_mean,
         tmx_mean_dist=tmx_mean_WL2 - tmx_mean) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))

historic_flint_dist_prep <- bind_cols(WL2_climate_short_flint, pops_flint_historic_avgs)
names(historic_flint_dist_prep)

historic_flint_dist <- historic_flint_dist_prep %>% 
  mutate(ppt_mean_dist=ppt_mean_WL2 - ppt_mean,
         tmn_mean_dist=tmn_mean_WL2 - tmn_mean,
         tmx_mean_dist=tmx_mean_WL2 - tmx_mean) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
```

Figures Recent (subtraction distance)

```{r}
recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_mean_dist), y=ppt_mean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/MeanPPT_DistfromWL2_RecentClim.png", width = 12, height = 6, units = "in")

recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmn_mean_dist), y=tmn_mean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/MeanTMN_DistfromWL2_RecentClim.png", width = 12, height = 6, units = "in")

recent_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmx_mean_dist), y=tmx_mean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/MeanTMX_DistfromWL2_RecentClim.png", width = 12, height = 6, units = "in")
```

Figures Historical (subtraction distance)

```{r}
historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_mean_dist), y=ppt_mean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/MeanPPT_DistfromWL2_HistoricalClim.png", width = 12, height = 6, units = "in")

historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmn_mean_dist), y=tmn_mean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/MeanTMN_DistfromWL2_HistoricalClim.png", width = 12, height = 6, units = "in")

historic_flint_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmx_mean_dist), y=tmx_mean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/MeanTMX_DistfromWL2_HistoricalClim.png", width = 12, height = 6, units = "in")
```

## Bioclim Climate Distance

### Subtraction

```{r}
#Recent
names(pops_bioclim_recent_avgs)
recent_bioclim_dist_prep <- bind_cols(WL2_bioclim_final, pops_bioclim_recent_avgs)
recent_bioclim_dist <- recent_bioclim_dist_prep %>% 
  mutate(ann_tmean_dist=ann_tmean_WL2 - ann_tmean_avg,
         mean_diurnal_range_dist=mean_diurnal_range_WL2 - mean_diurnal_range_avg,
         temp_seasonality_dist=temp_seasonality_WL2 - temp_seasonality_avg,
         temp_ann_range_dist=temp_ann_range_WL2 - temp_ann_range_avg,
         tmean_wettest_month_dist=tmean_wettest_month_WL2 - tmean_wettest_month_avg,
         tmean_driest_month_dist=tmean_driest_month_WL2 - tmean_driest_month_avg,
         ann_ppt_dist=ann_ppt_WL2 - ann_ppt_avg,
         ppt_seasonality_dist=ppt_seasonality_WL2 - ppt_seasonality_avg, 
         ppt_warmest_month_dist=ppt_warmest_month_WL2 - ppt_warmest_month_avg,
         ppt_coldest_month_dist=ppt_coldest_month_WL2 - ppt_coldest_month_avg) %>% 
 dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
recent_bioclim_dist

#Historical
names(pops_bioclim_historical_avgs)
historical_bioclim_dist_prep <- bind_cols(WL2_bioclim_final, pops_bioclim_historical_avgs)
historical_bioclim_dist <- historical_bioclim_dist_prep %>% 
  mutate(ann_tmean_dist=ann_tmean_WL2 - ann_tmean_avg,
         mean_diurnal_range_dist=mean_diurnal_range_WL2 - mean_diurnal_range_avg,
         temp_seasonality_dist=temp_seasonality_WL2 - temp_seasonality_avg,
         temp_ann_range_dist=temp_ann_range_WL2 - temp_ann_range_avg,
         tmean_wettest_month_dist=tmean_wettest_month_WL2 - tmean_wettest_month_avg,
         tmean_driest_month_dist=tmean_driest_month_WL2 - tmean_driest_month_avg,
         ann_ppt_dist=ann_ppt_WL2 - ann_ppt_avg,
         ppt_seasonality_dist=ppt_seasonality_WL2 - ppt_seasonality_avg, 
         ppt_warmest_month_dist=ppt_warmest_month_WL2 - ppt_warmest_month_avg,
         ppt_coldest_month_dist=ppt_coldest_month_WL2 - ppt_coldest_month_avg) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_dist"))
historical_bioclim_dist
```

Figures Recent (subtraction distance)

```{r}
recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_dist), y=ann_tmean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Ann_Tmean_DistfromWL2_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_dist), y=mean_diurnal_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Diurnal_Range_DistfromWL2_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_dist), y=temp_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Seasonality_DistfromWL2_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_dist), y=temp_ann_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Ann_Range_DistfromWL2_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_month_dist), y=tmean_wettest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Wet_DistfromWL2_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_month_dist), y=tmean_driest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Dry_DistfromWL2_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_dist), y=ann_ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Ann_PPT_DistfromWL2_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_dist), y=ppt_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/PPT_Seasonality_DistfromWL2_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_month_dist), y=ppt_warmest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/PPT_Warm_DistfromWL2_RecentClim.png", width = 12, height = 6, units = "in")

recent_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_month_dist), y=ppt_coldest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/PPT_Cold_DistfromWL2_RecentClim.png", width = 12, height = 6, units = "in")
```

Historical (subtraction distance)

```{r}
historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_tmean_dist), y=ann_tmean_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Ann_Tmean_DistfromWL2_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, mean_diurnal_range_dist), y=mean_diurnal_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Diurnal_Range_DistfromWL2_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_seasonality_dist), y=temp_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Seasonality_DistfromWL2_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, temp_ann_range_dist), y=temp_ann_range_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Ann_Range_DistfromWL2_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_wettest_month_dist), y=tmean_wettest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Wet_DistfromWL2_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, tmean_driest_month_dist), y=tmean_driest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Temp_Dry_DistfromWL2_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ann_ppt_dist), y=ann_ppt_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/Ann_PPT_DistfromWL2_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_seasonality_dist), y=ppt_seasonality_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/PPT_Seasonality_DistfromWL2_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_warmest_month_dist), y=ppt_warmest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/PPT_Warm_DistfromWL2_HistoricalClim.png", width = 12, height = 6, units = "in")

historical_bioclim_dist %>% 
  ggplot(aes(x=fct_reorder(parent.pop, ppt_coldest_month_dist), y=ppt_coldest_month_dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(fill="Elevation (m)",x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1))
ggsave("../output/Climate/PPT_Cold_DistfromWL2_HistoricalClim.png", width = 12, height = 6, units = "in")
```