---
title: "WL2_Survival-Analysis--GLMER"
author: "Julin Maloof"
date: '`r Sys.Date()`'
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Having settled on suitable models for survival, this script explores adding climate data from the growing season to see if they add any explanatory power.

# Survival Analysis 

See chapters 9 - 13 in tutorial: https://bookdown.org/content/4253/extending-the-discrete-time-hazard-model.html

## Libraries
```{r}
# install.packages("multilevelmod")
library(tidyverse) 
library(magrittr)
library(ggrepel)
library(lme4)
library(ggdist)
library(tidymodels)
library(multilevelmod)
library(broom.mixed)
library(modelr)
library(furrr)
library(zoo)
tidymodels_prefer()
options(mc.cores = parallel::detectCores())
```

## Read in the plant data
```{r}
wl2_surv <- read_csv("../output/WL2_Traits/WL2_Mortality_2023.csv")
```

As weeks
```{r}
wl2_surv %>% filter(is.na(death.date),!is.na(survey.notes)) #no plants that seemed to disappear 
wl2_surv %>% filter(Genotype=="YO7_4_2")

wl2_surv_dates <- wl2_surv %>%  
  rename(parent.pop = pop) %>%
  mutate(parent.pop= str_replace(parent.pop, "Y08", "YO8")) %>% 
  mutate(parent.pop= str_replace(parent.pop, "Y04", "YO4")) %>% 
  filter(!is.na(parent.pop)) %>% 
  mutate(planting.date="7/19/23", #could try to make this more specific to when certain blocks were planted 
         last_fup_date=if_else(is.na(death.date), "10/27/23", death.date)) %>%  #need this to calculate survival times
  mutate(planting.date=mdy(planting.date), last_fup_date=mdy(last_fup_date)) %>% #convert to date objects
  mutate(os_weeks=as.duration(planting.date %--% last_fup_date) / dweeks(1), #observed number of weeks
         status=if_else(is.na(death.date), 0, 1)) %>% #0=censured (alive in this case), 1=dead
  filter(os_weeks > 0) %>% #there is one case of a plant that was dead at planting, so just removed it since this is survival post- transplanting
  select(-survey.notes)
head(wl2_surv_dates)
unique(wl2_surv_dates$os_weeks) %>% sort()
```


## Read in soil temp

```{r}
temp <- read_csv("../input/WL2_Data/WL2_2022_2023_iButton_Data_Corrected.csv") %>%
  select(-`...3`) %>%
  mutate(Date_Time = mdy_hm(Date_Time)) %>%
  filter(Date_Time > ymd("2023-07-06"))
head(temp)
skimr::skim(temp)
```
```{r}
temp %>% ggplot(aes(x=Date_Time, y = SoilTemp, color= Bed)) + 
  geom_line()

#Take a closer look
temp %>% mutate(month = month(Date_Time)) %>%
  filter(Date_Time > ymd("2023-07-17"), Date_Time <= ymd("2023-07-24")) %>%
  ggplot(aes(x=Date_Time, y = SoilTemp, color= Bed, lty = Bed)) + 
  geom_line() +
  scale_color_brewer(type="qual", palette = "Accent")

temp %>% mutate(month = month(Date_Time)) %>%
  filter(month==7) %>%
  ggplot(aes(x=Date_Time, y = SoilTemp)) + 
  geom_line() +
  scale_color_brewer(type="qual", palette = "Accent") +
  facet_wrap(~Bed)


```
Not sure how to deal with the differences here since we don't have something in every bed or block. For now I am going to take A_2

```{r}
temp_summary <- temp %>%
  filter(Bed=="A_2") %>%
  mutate(Date=as.Date(Date_Time)) %>%
  filter(Date != min(Date), Date != max(Date)) %>% # trim ragged ends
  group_by(Date) %>%
  summarize(
    min_temp_d = min(SoilTemp),
    max_temp_d = max(SoilTemp),
    mean_temp_d = mean(SoilTemp)
  ) %>%
  mutate(
    roll_mean_min_temp_7 = rollmean(min_temp_d, k = 7, align = "right", fill = NA),
    roll_mean_max_temp_7 = rollmean(max_temp_d, k = 7, align = "right", fill = NA),
    roll_mean_mean_temp_7 = rollmean(mean_temp_d, k = 7, align = "right", fill = NA),
    roll_max_max_temp_7 = rollmax(max_temp_d, k = 7, align = "right", fill = NA),
  )

temp_summary
```

## Read in soil moisture

```{r}
moisture <- read_csv("../input/WL2_Data/WL2_2023_Bed_C_Soil_Moisture_Corrected.csv") %>%
  mutate(Date_Time = mdy_hm(Date_Time)) 
head(moisture)
skimr::skim(moisture)
```
```{r}
moisture %>% pivot_longer(-Date_Time) %>%
  ggplot(aes(x=Date_Time, y = value, color= name)) + 
  geom_line()
```
Will average across the ports

```{r}
moisture_summary <- moisture %>%
  rowwise() %>%
  mutate(s_moisture = mean(c_across(-Date_Time)) ) %>%
  select(Date_Time, s_moisture) %>%
  mutate(Date=as.Date(Date_Time)) %>%
  group_by(Date) %>%
  summarize(
    mean_moisture_d = mean(s_moisture)
  ) %>%
  mutate(
    roll_mean_soil_moisture_7 = rollmean(mean_moisture_d, k = 7, align = "right", fill = "extend"),
  )

moisture_summary
```

## Reformat to be one row per plant per week

```{r, eval=TRUE}
wl2_ind <- wl2_surv_dates %>% 
  complete(Genotype, last_fup_date) %>%
  group_by(Genotype) %>%
  mutate(across(c(-status, -last_fup_date, -os_weeks, -planting.date), \(x) ifelse(all(is.na(unique(x))), NA,unique(na.omit(x))))) %>%
  mutate(planting.date = unique(na.omit(planting.date)),
         os_weeks=as.duration(planting.date %--% last_fup_date) / dweeks(1),
           dead = ifelse((is.na(death.date) | mdy(death.date) > last_fup_date), 0, 1)) %>% # 0 is alive, 1 is dead
  filter(cumsum(dead) < 2) %>% # you can't die more than once
  ungroup() 

wl2_ind %>% select(Genotype, last_fup_date, os_weeks, dead)
```

Split data into establishment and survival data sets.  (We will only use survival set for this analysis)
```{r}
wl2_ind_estab <- wl2_ind %>% filter(os_weeks <= 3) %>%
  # compute status at week 3 for each plant
  group_by(Genotype) %>%
  summarize(dead=max(dead),
            across(c(block, bed, parent.pop, mf, contains("GD")), unique))

wl2_ind_surv <- wl2_ind %>% filter(os_weeks > 3) %>%
  mutate(os_weeks_f = factor(os_weeks))

wl2_estab_means <- wl2_ind_estab %>% # summary for plotting
  group_by(parent.pop) %>%
  summarize(survivorship = 1-mean(dead))

wl2_surv_means <- wl2_ind_surv %>%
  group_by(parent.pop, os_weeks) %>%
  summarize(survivorship = 1-mean(dead))
```

## add climate data

```{r}
wl2_ind_surv_climate <- wl2_ind_surv %>% 
  left_join(temp_summary, by = c("last_fup_date" = "Date")) %>%
  left_join(moisture_summary, by = c("last_fup_date" = "Date"))
```

## Temp analysis
check basic model forms
```{r}
# Note: StdDev due to mf within parent.pop is essentially zero, so I am omitting this.
glmer.model <- 
  linear_reg() %>%
  set_engine("glmer", family = binomial)

temp_wflow <- workflow() %>%
  add_variables(outcomes = dead, predictors = c(parent.pop, block, os_weeks, roll_mean_min_temp_7, roll_mean_max_temp_7, roll_mean_mean_temp_7, roll_max_max_temp_7 )) 

temp_fits <- tibble(wflow=list(

  # The best base model from earlier analyses:
  pop.weeks.block ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + (1|parent.pop) + (1|block))},
  
    mean.min.temp ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ roll_mean_min_temp_7 + os_weeks + (1|parent.pop) + (1|block))},
  
    mean.mean.temp ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ roll_mean_mean_temp_7 + os_weeks + (1|parent.pop) + (1|block))},
  
  mean.mean.temp.re_slope ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ roll_mean_mean_temp_7 + os_weeks + (roll_mean_mean_temp_7|parent.pop) + (1|block))},
  
    mean.max.temp ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ roll_mean_max_temp_7 + os_weeks + (1|parent.pop) + (1|block))},
  
      max.max.temp ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ roll_max_max_temp_7 + os_weeks + (1|parent.pop) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)
temp_fits
```

```{r}
system.time(
  temp_fits <- temp_fits %>%
  mutate(fit = map(wflow, fit, data = drop_na(wl2_ind_surv_climate, roll_mean_min_temp_7, roll_mean_max_temp_7, roll_mean_mean_temp_7, roll_max_max_temp_7), .progress = TRUE))
)
```

```{r}
temp_fits %>% pull(fit)
```

```{r}
temp_fits %>% 
  mutate(glance=map(fit, glance)) %>% 
  unnest(glance) %>% 
  arrange(BIC) %>%
  select(-wflow:-fit, -sigma)
```

get p-values

```{r}
temp_fits %>%
  mutate(tidy=map(fit, tidy)) %>%
  select(name, tidy) %>%
  unnest(tidy) %>%
  filter(str_detect(term, "^roll"))
```


## Moisture analysis
check basic model forms
```{r}
# Note: StdDev due to mf within parent.pop is essentially zero, so I am omitting this.
glmer.model <- 
  linear_reg() %>%
  set_engine("glmer", family = binomial)

moisture_wflow <- workflow() %>%
  add_variables(outcomes = dead, predictors = c(parent.pop, block, os_weeks, roll_mean_soil_moisture_7 )) 

moisture_fits <- tibble(wflow=list(

  # The best base model from earlier analyses:
  pop.weeks.block ={moisture_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + (1|parent.pop) + (1|block))},
  
    mean.moisture ={moisture_wflow %>% 
      add_model(glmer.model, formula = dead ~ roll_mean_soil_moisture_7 + os_weeks + (1|parent.pop) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)
moisture_fits
```

```{r}
system.time(
  moisture_fits <- moisture_fits %>%
  mutate(fit = map(wflow, fit, data = drop_na(wl2_surv_ind_climate, roll_mean_soil_moisture_7), .progress = TRUE))
)
```

```{r}
moisture_fits %>% pull(fit)
```

```{r}
moisture_fits %>% 
  mutate(glance=map(fit, glance)) %>% 
  unnest(glance) %>% 
  arrange(BIC) %>%
  select(-wflow:-fit, -sigma)
```

get p-values

```{r}
moisture_fits %>%
  mutate(tidy=map(fit, tidy)) %>%
  select(name, tidy) %>%
  unnest(tidy) %>%
  filter(str_detect(term, "moisture"))
```


```{r}
save.image("../output/JM_GLMER_Climate.Rdata")
```
